### 2016-06-08 paper.tex
first commit

% mnras_template.tex
%
% LaTeX template for creating an MNRAS paper
%
% v3.0 released 14 May 2015
% (version numbers match those of mnras.cls)
%
% Copyright (C) Royal Astronomical Society 2015
% Authors:
% Keith T. Smith (Royal Astronomical Society)

% Change log
%
% v3.0 May 2015
%    Renamed to match the new package name
%    Version number matches mnras.cls
%    A few minor tweaks to wording
% v1.0 September 2013
%    Beta testing only - never publicly released
%    First version: a simple (ish) template for creating an MNRAS paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[a4paper,fleqn,usenatbib]{mnras}

% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
%\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
\usepackage{txfonts}

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}

%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
%\usepackage{wasysym}
\usepackage{savesym}
\savesymbol{iint}
\savesymbol{iiint}
\savesymbol{iiiint}
\savesymbol{idotsint}
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Sparse Redundant Calibration]{SpARC: Sparse Redundant Calibration}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[K. T. Smith et al.]{
Keith T. Smith,$^{1}$\thanks{E-mail: mn@ras.org.uk (KTS)}
A. N. Other,$^{2}$
Third Author$^{2,3}$
and Fourth Author$^{3}$
\\
% List of institutions
$^{1}$Royal Astronomical Society, Burlington House, Piccadilly, London W1J 0BQ, UK\\
$^{2}$Department, Institution, Street Address, City Postal Code, Country\\
$^{3}$Another Department, Different Institution, Street Address, City Postal Code, Country
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2015}

% Don't change these lines
\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle

% Abstract of the paper
\begin{abstract}
This is a simple template for authors to write new MNRAS papers.
The abstract should briefly describe the aims, methods, and main results of the paper.
It should be a single paragraph not more than 250 words (200 words for Letters).
No references should appear in the abstract.
\end{abstract}

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
keyword1 -- keyword2 -- keyword3
\end{keywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%

\section{Introduction}

%All papers should start with an Introduction section, which sets the work
%in context, cites relevant earlier studies in the field by \citet{Others2013},
%and describes the problem the authors aim to solve \citep[e.g.][]{Author2012}.

\section{Historical Overview}

Write a little about how redundant calibration developed, a basic lit review.
Maybe write a little bit about LOGCAL here as well.

\subsection{EW regular-layout}
Describe the EW-layout and its redundant baselines

\subsection{LINCAL}
Briefly review LINCAL here and show that it is equivalent to the GN algorithm.

\section{Complex Optimization}
Describe complex optimization, Jacobian, Hessian etc... Benefits drawbacks compared to LINCAL... etc..

\section{Conjugate Gradient method}
brief description and lit review of the method.
computational complexity of the algorithm.
sparsity and condition number.

\section{Results: EW regular-layout}

\subsection{Jacobian derivation}
\subsection{Hessian derivation}
\subsection{Jacobian-residual product}

\subsection{Computational complexity using CG}
\subsubsection{Sparsity}
\subsubsection{Condition number}

\begin{enumerate}
 \item Plot eigenvalue distribution for different SNR and antenna numbers and different lambda values
 \item Plot computational complexity bound plus error reduction for different setups
 \item Plot number of iterations required by CG as a function of number of antennas
 \item Plot redundant baseline visibilities to show that they are approximately equal.
\end{enumerate}

NB:: Repeat the second last plot for the positive-semi definite case (GN), showing that even if the Hessian is singular we get fast convergence.

\section{Results: Regualar square grid}
If there is time

\section{Results: Hexanol grid}
NB:: Need to test if it works for this redundant layout.

% \section{Methods, Observations, Simulations etc.}
% 
% Normally the next section describes the techniques the authors used.
% It is frequently split into subsections, such as Section~\ref{sec:maths} below.
% 
% \subsection{Maths}
% \label{sec:maths} % used for referring to this section from elsewhere
% 
% Simple mathematics can be inserted into the flow of the text e.g. $2\times3=6$
% or $v=220$\,km\,s$^{-1}$, but more complicated expressions should be entered
% as a numbered equation:
% 
% \begin{equation}
%     x=\frac{-b\pm\sqrt{b^2-4ac}}{2a}.
% 	\label{eq:quadratic}
% \end{equation}
% 
% Refer back to them as e.g. equation~(\ref{eq:quadratic}).
% 
% \subsection{Figures and tables}
% 
% Figures and tables should be placed at logical positions in the text. Don't
% worry about the exact layout, which will be handled by the publishers.
% 
% Figures are referred to as e.g. Fig.~\ref{fig:example_figure}, and tables as
% e.g. Table~\ref{tab:example_table}.
% 
% % Example figure
% \begin{figure}
% 	% To include a figure from a file named example.*
% 	% Allowable file formats are eps or ps if compiling using latex
% 	% or pdf, png, jpg if compiling using pdflatex
% 	\includegraphics[width=\columnwidth]{example}
%     \caption{This is an example figure. Captions appear below each figure.
% 	Give enough detail for the reader to understand what they're looking at,
% 	but leave detailed discussion to the main body of the text.}
%     \label{fig:example_figure}
% \end{figure}
% 
% % Example table
% \begin{table}
% 	\centering
% 	\caption{This is an example table. Captions appear above each table.
% 	Remember to define the quantities, symbols and units used.}
% 	\label{tab:example_table}
% 	\begin{tabular}{lccr} % four columns, alignment for each
% 		\hline
% 		A & B & C & D\\
% 		\hline
% 		1 & 2 & 3 & 4\\
% 		2 & 4 & 6 & 8\\
% 		3 & 5 & 7 & 9\\
% 		\hline
% 	\end{tabular}
% \end{table}
% 
% 
% \section{Conclusions}
% 
% The last numbered section should briefly summarise what has been done, and describe
% the final conclusions which the authors draw from their work.

\section*{Acknowledgements}

The Acknowledgements section is not numbered. Here you can thank helpful
colleagues, acknowledge funding agencies, telescopes and facilities used etc.
Try to keep it short.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%

% The best way to enter references is to use BibTeX:

%\bibliographystyle{mnras}
%\bibliography{example} % if your bibtex file is called example.bib

% Alternatively you could enter them by hand, like this:
% This method is tedious and prone to error if you have lots of references
\begin{thebibliography}{99}
\bibitem[\protect\citeauthoryear{Author}{2012}]{Author2012}
Author A.~N., 2013, Journal of Improbable Astronomy, 1, 1
\bibitem[\protect\citeauthoryear{Others}{2013}]{Others2013}
Others S., 2012, Journal of Interesting Stuff, 17, 198
\end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%

\appendix

\section{Some extra material}

If you want to present additional material which would interrupt the flow of the main paper,
it can be placed in an Appendix which appears after the list of references.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}
\end{document}

% End of mnras_template.tex

### 2016-06-08 expression_calc/H_gen.py

@@ -1,0 +1,1030 @@
+import numpy as np
+import pylab as plt
+import pickle
+from scipy import optimize
+import matplotlib as mpl
+from copy import deepcopy
+import networkx as nx
+from networkx.utils import reverse_cuthill_mckee_ordering, cuthill_mckee_ordering
+
+
+class factor():
+
+    def __init__(self,
+                type_factor = "g",
+                index = 1, 
+                exponent = 1,
+                conjugate_value = False,
+                mag = False):
+                self.type_factor = type_factor
+                self.index = index
+                self.exponent = exponent
+                self.conjugate_value = conjugate_value
+                self.mag = mag
+
+    def to_string(self):
+        string_out = self.type_factor + "_" + str(self.index)
+        if self.conjugate_value:
+           string_out = string_out + "^" + "*"
+           if self.exponent > 1:
+              string_out = "("+string_out+")"+"^"+str(self.exponent)
+        else:
+           if self.mag:
+              string_out = "|"+string_out+"|^2"
+           else:  
+              if self.exponent > 1:
+                 string_out = string_out+"^"+str(self.exponent)
+        return string_out   
+           
+    def conjugate(self):
+        if self.conjugate_value:
+           self.conjugate_value = False
+        else:
+           self.conjugate_value = True
+
+    def equal(self,factor_in):
+        if self.type_factor == factor_in.type_factor:
+           if self.index == factor_in.index:
+              if self.exponent == factor_in.exponent:
+                 if self.conjugate_value == factor_in.conjugate_value:
+                    return true
+        return false
+
+class term():
+      def __init__(self):
+          self.g_array = np.array([],dtype=object)
+          self.y_array = np.array([],dtype=object)
+          self.yc_array = np.array([],dtype=object) 
+          self.gc_array = np.array([],dtype=object)
+          self.ga_array = np.array([],dtype=object)
+          self.ya_array = np.array([],dtype=object)
+          self.zero = False
+
+      def substitute(self,g_v,y_v):
+          if self.zero:
+             return 0.0
+          number = 1
+          for factor in self.g_array:
+              indx = factor.index
+              number = number * g_v[indx-1]
+          for factor in self.y_array:
+              indx = factor.index
+              number = number * y_v[indx-1]
+          for factor in self.gc_array:
+              indx = factor.index
[…]

### 2016-06-08 expression_calc/convergence.py

@@ -1,0 +1,172 @@
+import numpy as np
+import pylab as plt
+import scipy as sp
+
+import networkx as nx
+from networkx.utils import reverse_cuthill_mckee_ordering, cuthill_mckee_ordering
+
+ITR = 0
+compute_error = False
+error_list = []
+x_c = 0
+M = np.array([])
+
+def report(xk):
+    #print "IN"
+    global ITR
+    global compute_error
+    global error_list
+    global M
+
+    ITR = ITR + 1
+    if compute_error:
+       #print "len xk = ",len(xk)
+       e = x_c - xk
+       e_temp = np.dot(M,e)
+       #print "e_temp = ",len(e_temp)
+       e_v = np.sqrt(np.sum(e_temp*e))
+       #print "e_v = ",len(e_v)
+       e2 = x_c - np.ones(x_c.shape)
+       e_temp2 = np.dot(M,e2)   
+       e_v2 = np.sqrt(np.sum(e_temp2*e2))       
+
+       error_list.append(e_v.real/e_v2.real) 
+       
+    
+
+def create_matrix():
+    A = np.array([[3,2],[2,6]])
+    b = np.array([2,-8])
+    return A,b
+
+def tridiag(a, b, c, N, k1=-1, k2=0, k3=1):
+    t_1 = np.ones((N,))
+    return np.diag(a*t_1[1:], k1) + np.diag(b*t_1, k2) + np.diag(c*t_1[1:], k3)
+
+def quaddiag_p1(a,b,c,d,e,N,k1=-2,k2=-1,k3=0,k4=1,k5=2):
+    t_1 = np.ones((N,))
+    D = np.diag(a*t_1[2:], k1) + np.diag(b*t_1[1:], k2) + np.diag(c*t_1, k3) + np.diag(d*t_1[1:], k4) + np.diag(e*t_1[2:], k5)
+    D[0,0] = D[0,0]-1
+    D[N-1,N-1] = D[N-1,N-1]-1
+    return D
+    
+if  __name__=="__main__":
+    #global ITR
+    #global compute_error
+    #global error_list
+    #global M
+    '''
+    #a = 1./9; b = 5./18; c = 1./9; m = 50; N = m**2
+    a = -1; b = 2; c = -1; m = 50; N = m**2
+    T1 = tridiag(a, b, c, m)
+
+    a = 0; b = 1; c = 0
+    I = tridiag(a, b, c, m)
+
[…]

### 2016-06-08 expression_calc/kappa.py

@@ -1,0 +1,205 @@
+import numpy as np
+import pylab as plt
+import scipy as sp
+import H_gen
+
+class redundant_calibration():
+      def __init__(self):
+          self.H_symbols = np.array([])
+          pass
+
+      def compute_kappa(self,H_num):
+          #ones = np.ones((H.shape[0],),dtype=float)
+          #I = np.diag(ones)
+          d = np.diag(H_num)
+          d = d**(-1)
+          D = np.diag(d)
+          w = np.linalg.eigvals(np.dot(D,H_num))
+          lam_max = np.amax(w.real[w.real>1e-6])
+          lam_min = np.min(w.real[w.real>1e-6])
+          print "lam_max = ",lam_max
+          print "lam_min = ",lam_min
+          print "w = ",w
+          return lam_max/lam_min,lam_min,lam_max
+ 
+      def computeH(self,z,N):
+          g = z[:N]
+          y = z[N:] 
+          print "len(g) = ",len(g)
+          print "len(y) = ",len(y)      
+          r = H_gen.redundant(N)
+          r.create_regular()
+          r.create_J1()
+          r.create_J2()
+          r.conjugate_J1_J2()
+          r.create_J()
+          r.hermitian_transpose_J()
+          r.compute_H()
+          H_int = r.to_int_H()
+          #plt.imshow(H_int,interpolation="nearest")
+          #plt.show()
+          H_num = r.substitute_H(g,y)
+          #plt.imshow(np.real(H_num),interpolation="nearest")
+          #plt.show()
+          return H_num,H_int
+
+
+      def create_vis_mat(self,point_sources,u_m,v_m,g,sig):
+          D = np.zeros(u_m.shape)
+          G = np.diag(g)
+          print "D.shape = ",D.shape
+          print "G.shape = ",G.shape
+          #Step 1: Create Model Visibility Matrix
+          for k in xrange(len(point_sources)): #for each point source
+              
+              l_0 = point_sources[k,1]
+              m_0 = point_sources[k,2]
+              print "point_sources[k,0] = ",point_sources[k,0]
[…]

### 2016-06-08 expression_calc/ratio_calc.py

@@ -1,0 +1,46 @@
+import numpy as np
+import pylab as plt
+
+def vector_random_complex_number(sigma,mu,N):
+    x = (sigma * np.random.randn(N) + mu) + (sigma * np.random.randn(N) + mu)*1j
+    return x 
+
+def specific_sum(x):
+    num = np.cumsum(x.conj())
+    den = np.cumsum(x*x.conj())
+    rat = num/den
+    return rat
+
+
+
+if __name__ == "__main__":
+   
+   '''
+   x = vector_random_complex_number(10,5,10)
+   one = np.ones(x.shape,dtype=int)
+   one = np.cumsum(one)
+
+   rat = specific_sum(x)
+   
+   plt.plot(one,np.absolute(x))
+   plt.show()
+
+   plt.plot(one,1/np.absolute(x))
+   plt.plot(one,np.absolute(rat))
+   plt.show()   
+   
+   
+   '''
+   x = np.random.uniform(0.9,1.1,1000)
+   y = np.random.uniform(1,10,1000)
+   one = np.ones(x.shape,dtype=int)
+   one = np.cumsum(one)
+   num = np.cumsum(x*y**2 + x**2*y)
+   den = np.cumsum(x**2*y**2)
+   rat = 1 + num/den
+
+   plt.plot(one,rat)
+   plt.show()	    
+
+
+ 

### 2016-06-08 expression_calc/red_test.py

@@ -1,0 +1,865 @@
+import scipy as sp
+import numpy as np
+import pylab as plt
+import pickle
+from scipy import optimize
+import matplotlib as mpl
+from copy import deepcopy
+import networkx as nx
+from networkx.utils import reverse_cuthill_mckee_ordering, cuthill_mckee_ordering
+import networkx as nx
+from networkx.utils import reverse_cuthill_mckee_ordering, cuthill_mckee_ordering
+
+class test_redundant_calibration():
+
+      def __init__(self):
+          self.itr = 0
+          self.itr_predict = 0
+          self.mylist = []
+          self.mylist2 = []
+          self.compute_error = False
+          self.error_list = []
+          self.M = np.array([])
+          pass
+      
+      def create_vis_mat(self,point_sources,u_m,v_m,g,sig):
+          D = np.zeros(u_m.shape)
+          G = np.diag(g)
+          print "D.shape = ",D.shape
+          print "G.shape = ",G.shape
+          #Step 1: Create Model Visibility Matrix
+          for k in xrange(len(point_sources)): #for each point source
+              
+              l_0 = point_sources[k,1]
+              m_0 = point_sources[k,2]
+              print "point_sources[k,0] = ",point_sources[k,0]
+              D = D + point_sources[k,0]*np.exp(-2*np.pi*1j*(u_m*l_0+v_m*m_0))
+    
+          for t in xrange(D.shape[2]): #for each time-step
+        
+              #Step 2: Corrupting the Visibilities 
+              D[:,:,t] = np.dot(G,D[:,:,t])
+              D[:,:,t] = np.dot(D[:,:,t],G.conj()) 
+        
+              #Step 3: Adding Noise
+              D[:,:,t] = D[:,:,t] + sig*np.random.randn(u_m.shape[0],u_m.shape[1]) + sig*np.random.randn(u_m.shape[0],u_m.shape[1])*1j
+    
+          return D
+
+      def create_uv_regular_line(self,spacing_in_meters=20,antennas=5,lam=0.21428571428571427,dec=60,H_min=-6,H_max=6,time_steps=600,plot=True):
+          b = np.ones((antennas,),dtype=float)*spacing_in_meters
+          b = np.cumsum(b)
+          if plot:
[…]

### 2016-06-08 expression_calc/redcal.py

@@ -1,0 +1,62 @@
+import numpy as np
+import pylab as plt
+
+import networkx as nx
+from networkx.utils import reverse_cuthill_mckee_ordering, cuthill_mckee_ordering
+
+def testH(g,y):
+
+    gc = g.conjugate() 
+    J_1 = np.array([[y[0]*gc[1],0,0,0,0,g[0]*gc[1],0,0,0],[0,y[0]*gc[2],0,0,0,g[1]*gc[2],0,0,0],[0,0,y[0]*gc[3],0,0,g[2]*gc[3],0,0,0],[0,0,0,y[0]*gc[4],0,g[3]*gc[4],0,0,0],[y[1]*gc[2],0,0,0,0,0,g[0]*gc[2],0,0],[0,y[1]*gc[3],0,0,0,0,g[1]*gc[3],0,0],[0,0,y[1]*gc[4],0,0,0,g[2]*gc[4],0,0],[y[2]*gc[3],0,0,0,0,0,0,g[0]*gc[3],0],[0,y[2]*gc[4],0,0,0,0,0,g[1]*gc[4],0],[y[3]*gc[4],0,0,0,0,0,0,0,g[0]*gc[4]]])
+    print "J_1 = ",J_1
+
+    J_2 = np.array([[0,g[0]*y[0],0,0,0,0,0,0,0],[0,0,g[1]*y[0],0,0,0,0,0,0],[0,0,0,g[2]*y[0],0,0,0,0,0],[0,0,0,0,g[3]*y[0],0,0,0,0],[0,0,g[0]*y[1],0,0,0,0,0,0],[0,0,0,g[1]*y[1],0,0,0,0,0],[0,0,0,0,g[2]*y[1],0,0,0,0],[0,0,0,g[0]*y[2],0,0,0,0,0],[0,0,0,0,g[1]*y[2],0,0,0,0],[0,0,0,0,g[0]*y[3],0,0,0,0]])
+    print "J_2 = ",J_2
+
+    J = np.zeros((20,18),dtype=complex)
+  
+    J[0:10,0:9] = J_1
+    J[0:10:,9:] = J_2
+    J[10:,0:9] = J_2.conjugate()
+    J[10:,9:] = J_1.conjugate()
+
+    print "J = ",J
+    plt.imshow(np.absolute(J),interpolation="nearest")
+    plt.colorbar()
+    plt.show()
+    J_H = J.transpose()
+    J_H = J_H.conjugate()
+
[…]

### 2016-06-08 expression_calc/simulator.py

@@ -1,0 +1,335 @@
+import numpy as np
+import pylab as plt
+
+'''
+Simulator that can create uv-tracks and scalar visibilities associated with a regular east-west array
+'''
+
+class sim():
+
+      '''
+      Constructor
+      '''
+      def __init__(self):
+          pass
+	
+      '''
+      Generates the uv-coverage of a regularly spaced east-west (EW) array
+      RETURNS:
+      u_m - a N x N matrix containing the baseline u coordinates of a regurlarly spaced EW array
+      v_m - a N x N matrix containing the baseline v coordinates of a regurlarly spaced EW array
+      w_m - a N x N matrix containing the baseline w coordinates of a regurlarly spaced EW array
+      
+      INPUTS:
+      spacing_in_meters - regular spacing between the array elements in m
+      N - the amount of elements in the array
+      lam - observational wavelength
+      dec - declination of observation
+      H_min - starting hour angle (measured in hours)
+      H_max - maximum hour angle (measured in hours)
+      time_steps - amount of timesteps
+      plot - plots the antenna layout and uv-coverage if True
+      '''
+      def create_uv_regular_line(self,spacing_in_meters=20,N=5,lam=0.21428571428571427,dec=60,H_min=-6,H_max=6,time_steps=600,plot=True):
+          b = np.ones((N,),dtype=float)*spacing_in_meters
+          b = np.cumsum(b)
+          
+          if plot:
+             plt.plot(b,np.zeros((N,),dtype=float),'ro')
+             plt.xlabel("East-West [m]", fontsize=18)
+             plt.ylabel("North-South [m]", fontsize=18)
+             plt.title("ENU-coordinates of interferometer.", fontsize=18)
+             plt.show()
+          
+          b = b/lam
+          
+          bas = np.zeros(((N**2-N)/2,),dtype=float)
+          bas_2 = np.zeros(((N**2-N)/2,),dtype=float)
+          counter = 0
+          
+          for p in xrange(0,len(b)-1):
+              for q in xrange(p+1,len(b)):
+                  #bas[counter] = b[q] - b[p] #Taking Oleg's definition of the baseline vector
+                  bas[counter] = b[p] - b[q] 
+                  bas_2[counter] = q - p
+                  counter = counter + 1          
+ 
[…]

### 2016-06-08 expression_calc/solver.py

@@ -1,0 +1,756 @@
+import numpy as np
+import pylab as plt
+import H_gen
+import simulator
+from scipy import sparse
+from scipy import optimize
+import os
+import pickle
+
+
+'''
+Solver class that can use different methods to do redundant calibration. The aim being to compare these methods with one another
+'''
+
+class solver():
+      
+      '''
+      Constructor 
+      '''
+      def __init__(self):
+          self.J_sym = np.array([])
+          self.J = np.array([])
+          self.JH = np.array([])
+          self.J_int = np.array([])
+          self.H_sym = np.array([])
+          self.H = np.array([])
+          self.H_int = np.array([])
+          #self.r_vec = np.array([])
+          self.itr = 0
+          
+          self.itr_vec_mean = np.array([])
+          self.itr_vec_std = np.array([])
+ 
+          self.itr_list = []
+
+          self.directory = "EXPERIMENT_NAME"
+          self.txtfile = "TXTFILE.txt"
+          self.verbose = True
+
+      def write_to_current_textfile(self,str_to_write):
+          if self.verbose:
+             if not os.path.isdir("./"+self.directory):
+                os.system("mkdir "+self.directory)
+             if not os.path.isfile("./"+self.directory+"/"+self.txtfile):
+                f = open("./"+self.directory+"/"+self.txtfile,"w")
+             else:
+                f = open("./"+self.directory+"/"+self.txtfile,"a")
+          
+             f.write(str_to_write+"\n")
+             f.close()
+            
+      '''
+      Creates the redundant Jacobian from variables. Stores the result in self.J_sym
+      
+      RETURNS:
+      N/A
+ 
+      INPUTS:
+      N - Number of elements in array.     
+      '''
+      def construct_sym_J(self,N):
+          r = H_gen.redundant(N)
+          r.create_regular()
+          r.create_J1()
+          r.create_J2()
+          r.conjugate_J1_J2()
+          r.create_J()
+          self.J_sym = r.J
+          
+      '''
+      Substitutes the vector z into the symbolic redundant Jacobian. Stores the result in self.J
+      
+      RETURNS:
+      N/A
+ 
+      INPUTS:
+      z - Input vector.
+      N - Number of elements in array.
[…]

### 2016-06-08 expression_calc/spasity_ratio.py

@@ -1,0 +1,87 @@
+import numpy as np
+import pylab as plt
+
+def spar_ratio(N):
+    num = 5*N**2-7*N+2
+    den = 1.0*8*N**2-8*N+2
+    
+    gamma_N = (1.0)*num*den**(-1)  
+    return gamma_N
+
+def P_com(N):
+    return 4*N-2
+
+def k_com(N):
+    P = P_com(N)
+    k = (np.log(1-spar_ratio(N))+2*np.log(P))/np.log(P)
+    return k
+
+def c_com(N):
+    P = P_com(N)
+    c = np.log(0.25*P**2-1)/np.log(P)
+    return c
+
+if __name__ == "__main__":
+   
+   N_v = np.ones((200,),dtype=int)
+   N_v = np.cumsum(N_v)
+   N_v = N_v[N_v > 4]
+   
+   ones = np.ones(N_v.shape,dtype=int)
+   gamma = ones*5./8
+   gamma_v = spar_ratio(N_v)
+
+   k_v = k_com(N_v)
+   c_v = c_com(N_v)
+   P_v = P_com(N_v)
+   
+   #print "gamma_v = ",gamma_v
+   plt.plot(N_v,gamma_v)
+   plt.plot(N_v,gamma,"r--")
+
+   plt.xlabel("$N$")
+   plt.ylabel("$\gamma_N$")
+   plt.xlim([5,200])
+   plt.show()   
+   
+   plt.plot(P_v,k_v,label="$k_N$")
+   plt.plot(P_v,c_v,label="$c_N$")
+   plt.plot(P_v,2*ones,"r--",label="Limit")
+   plt.ylim([1.5,2.1])
+   plt.xlim([4*5-2,4*200-2])
+   plt.xlabel("$P$")
+   plt.legend()
+   plt.ylabel("Computational cost order")
+   plt.show()
+
+   plt.plot(N_v,k_v,label="$k_N$")
+   plt.plot(N_v,c_v,label="$c_N$")
+   plt.plot(N_v,2*ones,"r--",label="Limit")
+   plt.ylim([1.5,2.1])
+   plt.xlim([5,200])
+   plt.xlabel("$N$")
+   plt.legend()
+   plt.ylabel("Computational cost order")
+   plt.show()
+
+   """
+   plt.plot(P_v,c_v)
+   plt.plot(P_v,2*ones,"r--")
[…]

### 2016-07-04 old_latex/lincal_GN.tex
LINCAL and GN

\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amsfonts,amssymb,graphicx,subfigure}
\usepackage{amsmath}
\usepackage{rotating}
\usepackage{lscape}
%opening
\title{LINCAL and Gauss-Newton}
\author{T.L. Grobler}

\begin{document}

\maketitle

\begin{equation}
c_{ij} = y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}
\end{equation}

\begin{eqnarray}
\frac{c_{ij}}{\partial \eta_i} &=& y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\ 
\frac{c_{ij}}{\partial \varphi_i} &=&  -i y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\
\frac{c_{ij}}{\partial \eta_j} &=& y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\ 
\frac{c_{ij}}{\partial \varphi_j} &=&  i y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\
\frac{y_{i-j}}{\partial \varphi_j} &=&  e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}
\end{eqnarray}

The Wirtinger derivative is used in the last equation.

\begin{eqnarray}
c_{ij} &\approx& c_{ij}^0 + \Delta \eta_i(y_{i-j}^0 e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}) + \Delta \eta_j(y_{i-j}^0 e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0})\\
&& -i\Delta\varphi_i (y_{i-j}^0e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}) +i\Delta\varphi_j (y_{i-j}^0e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0})\\
&& y_{i-j}^1(y_{i-j}^0e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0})\\
&\approx&  c_{ij}^0 + e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_i+ \Delta \eta_j - i\Delta\varphi_i + i\Delta\varphi_i))
\end{eqnarray}

\begin{equation}
\label{eq:residual}
r_{ij} = \delta_{ij} = c_{ij}-c_{ij}^0 = e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_i+ \Delta \eta_j - i\Delta\varphi_i + i\Delta\varphi_i)) 
\end{equation}

Eq.~\ref{eq:residual} allows us to construct the following linear system:

\begin{equation}
\boldsymbol{J}[\boldsymbol{\Delta \eta},\boldsymbol{\Delta \varphi},\boldsymbol{\Delta y}]^T = \boldsymbol{r}, 
\end{equation}

where $\boldsymbol{J}$ is equal to 
\begin{equation}
\boldsymbol{J} = \bigg [\overbrace{\frac{c_{pq}}{\partial \eta_i}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial \varphi_i}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial y_{t}}}^{t=1\cdots r_s} \bigg ] \bigg \} [pq] = 1\cdots N_{\textrm{b}} (p<q) 
\end{equation}
or
\begin{equation}
\boldsymbol{J} = \bigg [\frac{c_{pq}}{\partial \boldsymbol{\eta}},~\frac{c_{pq}}{\partial \boldsymbol{\varphi}},~\frac{c_{pq}}{\partial \boldsymbol{y}} \bigg ] \bigg \} [pq] = 1\cdots N_{\textrm{b}} (p<q). 
\end{equation}

The last column is again a Wirtinger derivative.

Which means we can obtain the least-squares estimate as follows:

\begin{equation}
[\boldsymbol{\Delta \eta},\boldsymbol{\Delta \varphi},\boldsymbol{\Delta y}]^T = [\boldsymbol{J}^H\boldsymbol{J}]^{-1}\boldsymbol{J}^H\boldsymbol{r}.
\end{equation}

What is interesting about LINCAL is that it seems to be a combination of standard least-squares (splitting the problem into amplitude and phase) and 
complex least-squares. The gains are solved with a standard approach (splitting into amp and phase) and the redundant visibilities use a Wirtinger derivative. What is strange though is that in normal complex optimization, if we optimize towards a complex variable x we also need to optimize towards
its complex conjugate. My best guess is that LINCAL works in any case since the conjugate redundant baselines are not used to calibrate ($p<q$). The only difference between the framework I was using is that
we regard the whole problem as complex, we construct our Jacobian by calculating the derivative towards $g$ and its conjugate as well as $y$ and its conjugate - Smirnov \& Tasse (2016).

Now that the relation between LINCAL and GN has been shown we can start focusing on the speedup that the conjugate gradient method provides as Liu (2010) originally proposed.
Also worth checking: whether using an LM step instead of a GN step improves your convergence properties.
\end{document}

### 2016-07-04 paper.tex
LINCAL and GN

@@
 \subsection{LINCAL}
+\begin{equation}
+c_{ij} = y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}
+\end{equation}
+
+\begin{eqnarray}
+\frac{c_{ij}}{\partial \eta_i} &=& y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\ 
+\frac{c_{ij}}{\partial \varphi_i} &=&  -i y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\
+\frac{c_{ij}}{\partial \eta_j} &=& y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\ 
+\frac{c_{ij}}{\partial \varphi_j} &=&  i y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\
+\frac{y_{i-j}}{\partial \varphi_j} &=&  e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}
+\end{eqnarray}
+
+The Wirtinger derivative is used in the last equation.
+
+\begin{eqnarray}
+c_{ij} &\approx& c_{ij}^0 + \Delta \eta_i(y_{i-j}^0 e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}) + \Delta \eta_j(y_{i-j}^0 e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0})\\
+&& -i\Delta\varphi_i (y_{i-j}^0e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}) +i\Delta\varphi_j (y_{i-j}^0e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0})\\
+&& y_{i-j}^1(y_{i-j}^0e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0})\\
+&\approx&  c_{ij}^0 + e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_i+ \Delta \eta_j - i\Delta\varphi_i + i\Delta\varphi_i))
+\end{eqnarray}
+
+\begin{equation}
+\label{eq:residual}
+r_{ij} = \delta_{ij} = c_{ij}-c_{ij}^0 = e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_i+ \Delta \eta_j - i\Delta\varphi_i + i\Delta\varphi_i)) 
+\end{equation}
+
+Eq.~\ref{eq:residual} allows us to construct the following linear system:
+
+\begin{equation}
+\boldsymbol{J}[\boldsymbol{\Delta \eta},\boldsymbol{\Delta \varphi},\boldsymbol{\Delta y}]^T = \boldsymbol{r}, 
+\end{equation}
+
+where $\boldsymbol{J}$ is equal to 
+\begin{equation}
+\boldsymbol{J} = \bigg [\overbrace{\frac{c_{pq}}{\partial \eta_i}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial \varphi_i}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial y_{t}}}^{t=1\cdots r_s} \bigg ] \bigg \} [pq] = 1\cdots N_{\textrm{b}} (p<q)
+\end{equation}
+or
+\begin{equation}
+\boldsymbol{J} = \bigg [\frac{c_{pq}}{\partial \boldsymbol{\eta}},~\frac{c_{pq}}{\partial \boldsymbol{\varphi}},~\frac{c_{pq}}{\partial \boldsymbol{y}} \bigg ] \bigg \} [pq] = 1\cdots N_{\textrm{b}} (p<q).
+\end{equation}
+
+The last column is again a Wirtinger derivative.
+
+Which means we can obtain the least-squares estimate as follows:
+
+\begin{equation}
+[\boldsymbol{\Delta \eta},\boldsymbol{\Delta \varphi},\boldsymbol{\Delta y}]^T = [\boldsymbol{J}^H\boldsymbol{J}]^{-1}\boldsymbol{J}^H\boldsymbol{r}.
+\end{equation}
+
+What is interesting about LINCAL is that it seems to be a combination of standard least-squares (splitting the problem into amplitude and phase) and 
+complex least-squares.
+The gains are solved with a standard approach (splitting into amp and phase) and the redundant visibilities use a Wirtinger derivative.
+What is strange though is that in normal complex optimization, if we optimize towards a complex variable x we also need to optimize towards
+its complex conjugate.
+My best guess is that LINCAL works in any case since the conjugate redundant baselines are not used to calibrate ($p<q$).
+The only difference between the framework I was using is that
+we regard the whole problem as complex, we construct our Jacobian by calculating the derivative towards $g$ and its conjugate as well as $y$ and its conjugate - Smirnov \& Tasse (2016).
+
+Now that the relation between LINCAL and GN has been shown we can start focusing on the speedup that the conjugate gradient method provides as Liu (2010) originally proposed.
+Also worth checking: whether using an LM step instead of a GN step improves your convergence properties.
 Briefly review LINCAL here and show that it is equivalent to the GN algorithm.

### 2016-07-04 paper.tex
Hex layout of antennas

@@
 \subsection{Hessian derivation}
+\begin{equation}
+\boldsymbol{H} = 
+\begin{bmatrix}
+\boldsymbol{A} & \boldsymbol{B}\\
+\boldsymbol{B}^* & \boldsymbol{A}
+\end{bmatrix}
+\end{equation}
+
+With
+
+\begin{equation}
+\boldsymbol{A} = 
+\begin{bmatrix}
+\boldsymbol{C} & \boldsymbol{D}\\
+\boldsymbol{D^H} & \boldsymbol{E}
+\end{bmatrix}
+\end{equation}
+
+\begin{equation}
+\boldsymbol{B} = 
+\begin{bmatrix}
+\boldsymbol{F} & \boldsymbol{G}\\
+\boldsymbol{G}^T & \boldsymbol{0}
+\end{bmatrix}
+\end{equation}
+
+\begin{equation}
+[\boldsymbol{C}]_{ij} = 
+\begin{cases}
+ \sum_{k \neq i} \left | g_k \right |^2 \left | y_{|i-k|} \right |^2 & \textrm{if} ~ i=j\\
+ 0 & \textrm{otherwise}
+\end{cases}
+\end{equation}
+
+\begin{equation}
+[\boldsymbol{D}]_{ij} = 
+\begin{cases}
+ g_i y_j^*  \left | g_{i+j} \right |^2  & \textrm{if} ~ i<j\\
+ 0 & \textrm{otherwise}
+\end{cases}
+\end{equation}
+
+\begin{equation}
+[\boldsymbol{E}]_{ij} = 
+\begin{cases}
+ \sum_{pq \in \mathcal{I}_i} \left | g_p \right |^2 \left | g_q \right |^2  & \textrm{if} ~ i=j\\
+ 0 & \textrm{otherwise}
+\end{cases}
+\end{equation}
+
+\begin{equation}
+[\boldsymbol{F}]_{ij} = 
+\begin{cases}
+ g_i g_j  \left | y_{|i-j|} \right |^2  & \textrm{if} ~ i \neq j\\
+ 0 & \textrm{otherwise}
+\end{cases}
+\end{equation}
+
+\begin{equation}
+[\boldsymbol{G}]_{ij} = 
+\begin{cases}
+ g_i y_j  \left | g_{i-j} \right |^2  & \textrm{if} ~ i > j\\
+ 0 & \textrm{otherwise}
+\end{cases}
+\end{equation}
+
+Moreover,
+
+\begin{equation}
+\mathcal{I}_i = \left\{pq|(q - p) = i \right\}
+\end{equation}
+
+The dimensions of the matrices:
+
+\begin{enumerate}
+\item $\boldsymbol{H}$: is a $(4N-2)\times(4N-2)$ or a $P \times P$ matrix. $N$ denotes the number of antennas and $P$ denotes the number of parameters. This matrix
+contains $6N^2 - 2N - 2$ non-zero entries. It contains $16N^2 - 16N + 4$ entries. We can construct this matrix with $4N^2 -  4N$ or $\frac{1}{4} P^2 -1$ elementary operations.
+\item $\boldsymbol{A}$: is a $(2N-1)\times(2N-1)$ matrix. This matrix has $N^2 + N - 1$ non-zero entries. We can construct this matrix with $2\frac{1}{2} (N^2 -  N)$ elementary operations.  
+\item $\boldsymbol{B}$: is a $(2N-1)\times(2N-1)$ matrix. This matrix has $2N^2 - 2N$ non-zero entries. We can construct this matrix with $1\frac{1}{2} (N^2 -  N)$ elementary operations.
+\item $\boldsymbol{C}$: is a $N\times N$ matrix. This matrix has $N$ non-zero entries. We can construct this matrix with $N^2-N$ elementary operations.
+\item $\boldsymbol{D}$: is a $N \times (N-1)$ matrix. This matrix has $\frac{1}{2} (N^2 -  N)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations. 
+\item $\boldsymbol{E}$: is a $(N-1) \times (N-1)$ matrix. This matrix has $(N-1)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.  
+\item $\boldsymbol{F}$: is a $N \times N$ matrix. This matrix has $N^2 - N$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.
+\item $\boldsymbol{G}$: is a $N \times (N-1)$ matrix. This matrix has $\frac{1}{2} (N^2 -  N)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.
+\item $\boldsymbol{0}$: is a $(N-1) \times (N-1)$ all zero matrix. 
+\end{enumerate}
+
+We can construct $\boldsymbol{C}-\boldsymbol{G}$ in $O(N^2)$
+
+The sparsity ratio of $\boldsymbol{H}$ is equal to
+\begin{equation}
+\gamma_N = \frac{10N^2-14N+6}{16N^2-16N+4} 
+\end{equation}
+
+The asymptotic sparsity ratio of $\boldsymbol{H}$ is equal to
+
+\begin{equation}
+\gamma = \lim_{N\rightarrow \infty}\frac{5N^2-7N+3}{8N^2-8N+2} = \frac{5}{8}
+\end{equation}
+
+Asymptotically the computational cost of the matrix-vector product between $\boldsymbol{H}$ and a vector of appropriate size is $O(P^2)$. However, the computational cost converges 
+very slowly to its asymptotic value, and in general is equal to $O(P^{k_N})$, where $k_N < 2$.
+
+\begin{eqnarray}
+P^{k_N}_N &=& (1 - \gamma_N)P^2_N\\
+k_N &=& \log_{P_N}(1 - \gamma_N) + 2\\
+k &=& \lim_{N\rightarrow \infty} \log_{P_N}(1 - \gamma_N) + 2 = 2
+\end{eqnarray}
+
+Moreover,
+
+\begin{eqnarray}
+P^{c_N}_N &=& \frac{1}{4} P_N^2 -1\\
+c_N &=& \log_{P_N}\left (\frac{1}{4} P_N^2-1 \right )\\
+c &=& \lim_{N\rightarrow \infty} \log_{P_N}\left (\frac{1}{4} P_N^2-1\right ) = 2
+\end{eqnarray}
+
+NB $P_c$ IS WRONG HERE DID NOT TAKE INTO ACCOUNT THE CONJUGANTION OF $\boldsymbol{B}$ (need multiply its computational complexity by two).
+
 \subsection{Jacobian-residual product}

### 2016-07-04 expression_calc/hex_grid.py

@@ -1,0 +1,12 @@
+import numpy as np
+import pylab as plt
+
+class Hgrid():
+
+    def __init__(self):
+        pass
+
+if __name__ == "__main__":
+   h = Hgrid()
+   print "Hallo"
+

### 2016-07-04 expression_calc/hex_grid.py
initial hex grid

@@ -6,7 +6,45 @@ class Hgrid():
     def __init__(self):
         pass
 
+    def hex_grid(self,hex_dim,l):
+        side = hex_dim + 1
+        ant_main_row = side + (side+1)
+        ant_x = np.zeros((6*hex_dim+1,),dtype=float)
+        ant_y = np.zeros((6*hex_dim+1,),dtype=float)
+        x = 0.0
+        y = 0.0
+
+        counter = 0
+        
+        for k in xrange(side):
+            x_row = x
+            y_row = y
+            for i in xrange(ant_main_row):
+                if k == 0:
+                   ant_x[counter] = x_row 
+                   ant_y[counter] = y_row
+                   x_row = x_row + l
+                   counter = counter + 1 
+                else:
+                   ant_x[counter] = x_row
+                   ant_y[counter] = y_row
+                   counter = counter + 1
+                   
+                   ant_x[counter] = x_row
+                   ant_y[counter] = -1*y_row
+                   x_row = x_row + l
+                   counter = counter + 1   
+            x = x + l/2.0
+            y = y + (np.sqrt(3)/2.0)*l                 
+            ant_main_row = ant_main_row - 1
+    
+        return ant_x,ant_y 
+
+
 if __name__ == "__main__":
    h = Hgrid()
+   ant_x,ant_y = h.hex_grid(1,20)
+   plt.plot(ant_x,ant_y,"ro")
+   plt.show()
    print "Hallo"
 

### 2016-07-04 expression_calc/hex_grid.py
hex_grid.py

@@ -8,7 +8,7 @@ class Hgrid():
 
     def hex_grid(self,hex_dim,l):
         side = hex_dim + 1
-        ant_main_row = side + (side+1)
+        ant_main_row = side + hex_dim
         ant_x = np.zeros((6*hex_dim+1,),dtype=float)
         ant_y = np.zeros((6*hex_dim+1,),dtype=float)
         x = 0.0

### 2016-07-04 expression_calc/hex_grid.py
fixed hex layout code

@@ -9,8 +9,17 @@ class Hgrid():
     def hex_grid(self,hex_dim,l):
         side = hex_dim + 1
         ant_main_row = side + hex_dim
-        ant_x = np.zeros((6*hex_dim+1,),dtype=float)
-        ant_y = np.zeros((6*hex_dim+1,),dtype=float)
+        
+        elements = 1
+
+        #summing the antennas in hexactoganal rings 
+        for k in xrange(hex_dim):
+            elements = elements + (k+1)*6
+                
+        ant_x = np.zeros((elements,),dtype=float)
+        ant_y = np.zeros((elements,),dtype=float)
+        print "len(ant_x) = ",len(ant_x)
+        print "len(ant_y) = ",len(ant_y)
         x = 0.0
         y = 0.0
 
@@ -43,7 +52,7 @@ class Hgrid():
 
 if __name__ == "__main__":
    h = Hgrid()
-   ant_x,ant_y = h.hex_grid(1,20)
+   ant_x,ant_y = h.hex_grid(4,20)
    plt.plot(ant_x,ant_y,"ro")
    plt.show()
    print "Hallo"

### 2016-07-04 expression_calc/hex_grid.py
end of day commit

@@ -47,7 +47,12 @@ class Hgrid():
             y = y + (np.sqrt(3)/2.0)*l                 
             ant_main_row = ant_main_row - 1
     
-        return ant_x,ant_y 
+        return ant_x,ant_y
+
+    def calculate_phi(self,ant_x,ant_y):
+        for k in xrange(len(ant_x)):
+            for j in xrnage(k+1,len(ant_x)) 
+
 
 
 if __name__ == "__main__":

### 2016-07-04 expression_calc/hex_grid.py
end of day commit

@@ -52,7 +52,7 @@ class Hgrid():
     def calculate_phi(self,ant_x,ant_y):
         for k in xrange(len(ant_x)):
             for j in xrnage(k+1,len(ant_x)) 
-
+ 
 
 
 if __name__ == "__main__":

### 2016-07-05 expression_calc/hex_grid.py
phi implmented

@@ -49,10 +49,29 @@ class Hgrid():
     
         return ant_x,ant_y
 
-    def calculate_phi(self,ant_x,ant_y):
+    def determine_phi_value(self,red_vec_x,red_vec_y,ant_x_p,ant_x_q,ant_y_p,ant_y_q):
+        red_x = ant_x_q - ant_x_p
+        red_y = ant_y_q - ant_y_p
+
+        for l in xrange(len(red_vec_x)):
+            if (np.allclose(red_x,red_vec_x[l]) and np.allclose(red_y,red_vec_y[l])):
+               return red_vec_x,red_vec_y,l+1
+
+        red_vec_x = np.append(red_vec_x,np.array[red_x])
+        red_vec_y = np.append(red_vec_y,np.array[red_y])
+        return red_vec_x,red_vec_y,len(red_vec_x) 
+
+    def calculate_phi(self,ant_x,ant_y,plot=True):
+        phi = np.zeros((len(ant_x),len(ant_y)))
+        
         for k in xrange(len(ant_x)):
-            for j in xrnage(k+1,len(ant_x)) 
- 
+            for j in xrnage(k+1,len(ant_x))
+                red_vec_x,red_vec_y,phi[k,j]  = self.determine_phi_value(red_vec_x,red_vec_y,ant_x[k],ant_x[j],ant_y[k],ant_y[j])           
+                phi[j,k] = phi[k,j]
+
+        if plot:
+           plt.imshow(phi)
+           plt.show()
 
 
 if __name__ == "__main__":
@@ -60,5 +79,6 @@ if __name__ == "__main__":
    ant_x,ant_y = h.hex_grid(4,20)
    plt.plot(ant_x,ant_y,"ro")
    plt.show()
-   print "Hallo"
+   h.calculate_phi(ant_x,ant_y)
+   #print "Hallo"
 

### 2016-07-05 expression_calc/hex_grid.py
phi calc fix

@@ -57,15 +57,16 @@ class Hgrid():
             if (np.allclose(red_x,red_vec_x[l]) and np.allclose(red_y,red_vec_y[l])):
                return red_vec_x,red_vec_y,l+1
 
-        red_vec_x = np.append(red_vec_x,np.array[red_x])
-        red_vec_y = np.append(red_vec_y,np.array[red_y])
+        red_vec_x = np.append(red_vec_x,np.array([red_x]))
+        red_vec_y = np.append(red_vec_y,np.array([red_y]))
         return red_vec_x,red_vec_y,len(red_vec_x) 
 
     def calculate_phi(self,ant_x,ant_y,plot=True):
         phi = np.zeros((len(ant_x),len(ant_y)))
-        
+        red_vec_x = np.array([])
+        red_vec_y = np.array([])
         for k in xrange(len(ant_x)):
-            for j in xrnage(k+1,len(ant_x))
+            for j in xrange(k+1,len(ant_x)):
                 red_vec_x,red_vec_y,phi[k,j]  = self.determine_phi_value(red_vec_x,red_vec_y,ant_x[k],ant_x[j],ant_y[k],ant_y[j])           
                 phi[j,k] = phi[k,j]
 

### 2016-07-05 expression_calc/hex_grid.py
phi

@@ -71,13 +71,22 @@ class Hgrid():
                 phi[j,k] = phi[k,j]
 
         if plot:
-           plt.imshow(phi)
+           plt.imshow(phi,interpolation="nearest")
+           x = np.arange(len(ant_x))
+           plt.xticks(x, x+1)
+           y = np.arange(len(ant_x))
+           plt.yticks(y, y+1)
+           plt.colorbar() 
+           #plt.yticks(y, y+1)
+           
            plt.show()
+
+        print "phi = ",phi
 
 
 if __name__ == "__main__":
    h = Hgrid()
-   ant_x,ant_y = h.hex_grid(4,20)
+   ant_x,ant_y = h.hex_grid(3,20)
    plt.plot(ant_x,ant_y,"ro")
    plt.show()
    h.calculate_phi(ant_x,ant_y)

### 2016-07-05 expression_calc/H_gen.py
tried phi

@@ -15,15 +15,28 @@ class factor():
                 index = 1, 
                 exponent = 1,
                 conjugate_value = False,
-                mag = False):
+                mag = False,
+                ant_p = 0,
+                ant_q = 0,
+                print_f = True):
                 self.type_factor = type_factor
                 self.index = index
                 self.exponent = exponent
                 self.conjugate_value = conjugate_value
                 self.mag = mag
+                self.ant_p = ant_p
+                self.ant_q = ant_q
+                self.print_f = print_f 
+             
 
     def to_string(self):
-        string_out = self.type_factor + "_" + str(self.index)
+        if self.print_f:
+           if self.type_factor == "y":
+              string_out = self.type_factor + "_{" + str(self.ant_p)+str(self.ant_q)+"}"
+           else:
+              string_out = self.type_factor + "_" + str(self.index)
+        else:   
+           string_out = self.type_factor + "_" + str(self.index)
         if self.conjugate_value:
            string_out = string_out + "^" + "*"
            if self.exponent > 1:
@@ -563,7 +576,7 @@ class redundant():
                   J_temp[r,c].conjugate()
           self.JH = J_temp.transpose()   
           
-      def create_regular(self):
+      def create_regular(self,print_pq=False):
           spacing_vector = np.arange(self.N)[1:]  
           redundant_baselines = spacing_vector[::-1]          
           
@@ -573,9 +586,9 @@ class redundant():
               ant1 = 1
               for j in xrange(redundant_baselines[k]):
                   ant2 = ant1 + spacing_vector[k]
-                  g_fact = factor("g",ant1,1,False)
-                  y_fact = factor("y",spacing_vector[k],1,False)
-                  gc_fact = factor("g",ant2,1,True)
+                  g_fact = factor("g",ant1,1,False,False,0,0,False)
+                  y_fact = factor("y",spacing_vector[k],1,False,False,ant1,ant1,print_pq)
+                  gc_fact = factor("g",ant2,1,True,False,0,0,False)
                   t_temp = term()
                   t_temp.append_factor(g_fact)
                   t_temp.append_factor(y_fact)
@@ -583,15 +596,15 @@ class redundant():
                   self.regular_array = np.append(self.regular_array,t_temp)
                   ant1 = ant1 + 1
 
-      def create_normal(self):
+      def create_normal(self,print_pq=False):
[…]

### 2016-07-05 expression_calc/hex_grid.py
newest hex_grid

@@ -86,7 +86,7 @@ class Hgrid():
 
 if __name__ == "__main__":
    h = Hgrid()
-   ant_x,ant_y = h.hex_grid(3,20)
+   ant_x,ant_y = h.hex_grid(1,20)
    plt.plot(ant_x,ant_y,"ro")
    plt.show()
    h.calculate_phi(ant_x,ant_y)

### 2016-07-05 expression_calc/H_gen.py
working hex Hessian

@@ -381,12 +381,15 @@ class redundant():
           self.J = np.array([],dtype=object)
           self.JH = np.array([],dtype=object)
           self.H = np.array([],dtype=object)
+          self.phi = np.array([])
          
       def create_J1(self,type_v="RED"):
           rows = ((self.N*self.N) - self.N)/2
 
           if type_v == "RED":
              columns = self.N + (self.N-1)
+          elif type_v == "HEX":
+             columns = self.N + int(np.amax(self.phi))
           else:
              columns = self.N
  
@@ -427,6 +430,8 @@ class redundant():
           rows = ((self.N*self.N) - self.N)/2
           if type_v == "RED":
              columns = self.N + (self.N-1)
+          elif type_v == "HEX":
+             columns = self.N + int(np.amax(self.phi))
           else:
              columns = self.N
           self.J2 = np.empty((rows,columns),dtype=object)
@@ -465,6 +470,8 @@ class redundant():
           rows = ((self.N*self.N) - self.N)
           if type_v == "RED":
              columns = 2*(self.N + (self.N-1))
+          elif type_v == "HEX":
+             columns = 2*(self.N + int(np.amax(self.phi)))
           else:
              columns = 2*(self.N)
           self.J = np.empty((rows,columns),dtype=object)
@@ -576,6 +583,105 @@ class redundant():
                   J_temp[r,c].conjugate()
           self.JH = J_temp.transpose()   
           
+      def hex_grid(self,hex_dim,l):
+          side = hex_dim + 1
+          ant_main_row = side + hex_dim
+        
+          elements = 1
+
+          #summing the antennas in hexactoganal rings 
+          for k in xrange(hex_dim):
+              elements = elements + (k+1)*6
+                
+          ant_x = np.zeros((elements,),dtype=float)
+          ant_y = np.zeros((elements,),dtype=float)
+          print "len(ant_x) = ",len(ant_x)
+          print "len(ant_y) = ",len(ant_y)
+          x = 0.0
+          y = 0.0
+
+          counter = 0
+        
+          for k in xrange(side):
+              x_row = x
+              y_row = y
+              for i in xrange(ant_main_row):
+                  if k == 0:
+                     ant_x[counter] = x_row 
+                     ant_y[counter] = y_row
+                     x_row = x_row + l
[…]

### 2016-07-06 expression_calc/H_gen.py
first attempt at general formulation

@@ -661,7 +661,7 @@ class redundant():
           print "phi = ",phi
           return phi
 
-      def create_hexagonal(self,hex_dim,l,print_pq=True):
+      def create_hexagonal(self,hex_dim,l,print_pq=False):
           self.regular_array=np.array([],dtype=object)
 
           ant_x,ant_y = self.hex_grid(hex_dim,l)
@@ -673,7 +673,7 @@ class redundant():
           for p in xrange(1,self.N):
               for q in xrange(p+1,self.N+1):
                   g_fact = factor("g",p,1,False,False,0,0,False) 
-                  y_fact = factor("y",self.phi[p-1,q-1],1,False,False,p,q,print_pq)          
+                  y_fact = factor("y",int(self.phi[p-1,q-1]),1,False,False,p,q,print_pq)          
                   gc_fact = factor("g",q,1,True,False,0,0,False)
                   t_temp = term()
                   t_temp.append_factor(g_fact)
@@ -743,6 +743,7 @@ class redundant():
           return string_out
 
       def compute_H(self,type_v="RED"):
+          print "ymax = ",np.amax(self.phi)
           if type_v == "RED":
              parameters = 2*(self.N + (self.N-1)) 
           elif type_v == "HEX":
@@ -1080,7 +1081,7 @@ if __name__ == "__main__":
    #print r.to_string_JH()
    print r.to_string_H()
    print r.to_string_H(simplify=True)
-   #r.to_latex_H(simplify=True)
+   r.to_latex_H(simplify=True)
    #r.to_latex_J()
    #r.to_latex_JH()
    

### 2016-07-07 expression_calc/hex_grid.py
newest phi computation

@@ -49,6 +49,76 @@ class Hgrid():
     
         return ant_x,ant_y
 
+    def hex_grid_ver2(self,hex_dim,l):
+        hex_dim = int(hex_dim)
+        side = int(hex_dim + 1)
+        ant_main_row = int(side + hex_dim)
+        
+        elements = 1
+
+        #summing the antennas in hexactoganal rings 
+        for k in xrange(hex_dim):
+            elements = elements + (k+1)*6
+                
+        ant_x = np.zeros((elements,),dtype=float)
+        ant_y = np.zeros((elements,),dtype=float)
+        print "len(ant_x) = ",len(ant_x)
+        print "len(ant_y) = ",len(ant_y)
+        x = 0.0
+        y = 0.0
+
+        counter = 0
+        
+        for k in xrange(side):
+            x_row = x
+            y_row = y
+            for i in xrange(ant_main_row):
+                if k == 0:
+                   ant_x[counter] = x_row 
+                   ant_y[counter] = y_row
+                   x_row = x_row + l
+                   counter = counter + 1 
+                else:
+                   ant_x[counter] = x_row
+                   ant_y[counter] = y_row
+                   counter = counter + 1
+                   
+                   ant_x[counter] = x_row
+                   ant_y[counter] = -1*y_row
+                   x_row = x_row + l
+                   counter = counter + 1   
+            x = x + l/2.0
+            y = y + (np.sqrt(3)/2.0)*l                 
+            ant_main_row = ant_main_row - 1
+       
+        y_idx = np.argsort(ant_y)
+        ant_y = ant_y[y_idx]
+        ant_x = ant_x[y_idx]
+
+        slice_value = int(side)
+        start_index = 0
+        add = True
+        ant_main_row = int(side + hex_dim)
+
+        for k in xrange(ant_main_row):
+            temp_vec_x = ant_x[start_index:start_index+slice_value]
+            x_idx = np.argsort(temp_vec_x)
+            temp_vec_x = temp_vec_x[x_idx]
+            ant_x[start_index:start_index+slice_value] = temp_vec_x
+            if slice_value == ant_main_row:
+               add = False
+            start_index = start_index+slice_value 
+            
+            if add:
+               slice_value = slice_value + 1
+            else:
[…]

### 2016-07-07 expression_calc/H_gen.py
fixed hex_grid

@@ -626,6 +626,76 @@ class redundant():
     
           return ant_x,ant_y
 
+      def hex_grid_ver2(self,hex_dim,l):
+          hex_dim = int(hex_dim)
+          side = int(hex_dim + 1)
+          ant_main_row = int(side + hex_dim)
+        
+          elements = 1
+
+          #summing the antennas in hexactoganal rings 
+          for k in xrange(hex_dim):
+              elements = elements + (k+1)*6
+                 
+          ant_x = np.zeros((elements,),dtype=float)
+          ant_y = np.zeros((elements,),dtype=float)
+          print "len(ant_x) = ",len(ant_x)
+          print "len(ant_y) = ",len(ant_y)
+          x = 0.0
+          y = 0.0
+
+          counter = 0
+        
+          for k in xrange(side):
+              x_row = x
+              y_row = y
+              for i in xrange(ant_main_row):
+                  if k == 0:
+                     ant_x[counter] = x_row 
+                     ant_y[counter] = y_row
+                     x_row = x_row + l
+                     counter = counter + 1 
+                  else:
+                     ant_x[counter] = x_row
+                     ant_y[counter] = y_row
+                     counter = counter + 1
+                   
+                     ant_x[counter] = x_row
+                     ant_y[counter] = -1*y_row
+                     x_row = x_row + l
+                     counter = counter + 1   
+              x = x + l/2.0
+              y = y + (np.sqrt(3)/2.0)*l                 
+              ant_main_row = ant_main_row - 1
+       
+          y_idx = np.argsort(ant_y)
+          ant_y = ant_y[y_idx]
+          ant_x = ant_x[y_idx]
+
+          slice_value = int(side)
+          start_index = 0
+          add = True
+          ant_main_row = int(side + hex_dim)
+
+          for k in xrange(ant_main_row):
+              temp_vec_x = ant_x[start_index:start_index+slice_value]
+              x_idx = np.argsort(temp_vec_x)
+              temp_vec_x = temp_vec_x[x_idx]
+              ant_x[start_index:start_index+slice_value] = temp_vec_x
+              if slice_value == ant_main_row:
+                 add = False
+              start_index = start_index+slice_value 
+            
+              if add:
+                 slice_value = slice_value + 1
+              else:
+                 slice_value = slice_value - 1  
+
[…]

### 2016-07-08 expression_calc/H_gen.py
added square grid

@@ -479,6 +479,34 @@ class redundant():
           self.J[:rows/2,columns/2:] = self.J2
           self.J[rows/2:,:columns/2] = self.Jc2
           self.J[rows/2:,columns/2:] = self.Jc1
+
+      #def create_J_L(self,type_v="RED"):
+      #    rows = ((self.N*self.N) - self.N)
+      #    if type_v == "RED":
+      #       columns = (self.N + (self.N-1))
+      #    elif type_v == "HEX":
+      #       columns = (self.N + int(np.amax(self.phi)))
+      #    else:
+      #       columns = (self.N)
+      #    self.J = np.empty((rows,columns),dtype=object)
+      #    self.J[:rows/2,:columns] = self.J1
+      #    self.J[:rows/2,columns/2:] = self.J2
+      #    self.J[rows/2:,:columns/2] = self.Jc2
+      #    self.J[rows/2:,columns/2:] = self.Jc1
+
+      #def create_special_J(self,type_v="RED"):
+      #    rows = ((self.N*self.N) - self.N)
+      #    if type_v == "RED":
+      #       columns = (self.N + (self.N-1))
+      #    elif type_v == "HEX":
+      #       columns = (self.N + int(np.amax(self.phi)))
+      #    else:
+      #       columns = (self.N)
+      #    self.J = np.empty((rows,columns),dtype=object)
+      #    self.J[:rows/2,:columns] = self.J1
+      #    self.J[:rows/2,columns/2:] = self.J2
+      #    self.J[rows/2:,:columns/2] = self.Jc2
+      #    self.J[rows/2:,columns/2:] = self.Jc1 
      
       def to_string_J1(self):
           string_out = " J1 = ["

### 2016-07-08 expression_calc/hex_grid.py

@@ -119,6 +119,29 @@ class Hgrid():
 
         return ant_x,ant_y
 
+    def square_grid(self,side,l):
+        elements = side*side
+                
+        ant_x = np.zeros((elements,),dtype=float)
+        ant_y = np.zeros((elements,),dtype=float)
+        print "len(ant_x) = ",len(ant_x)
+        print "len(ant_y) = ",len(ant_y)
+        x = 0.0
+        y = 0.0
+
+        counter = 0
+        
+        for k in xrange(side):
+            x = 0.0
+            for i in xrange(side):
+                ant_x[counter] = x
+                ant_y[counter] = y 
+                x = x + l
+                counter = counter + 1
+            y = y + l 
+ 
+        return ant_x,ant_y
+
     def determine_phi_value(self,red_vec_x,red_vec_y,ant_x_p,ant_x_q,ant_y_p,ant_y_q):
         red_x = ant_x_q - ant_x_p
         red_y = ant_y_q - ant_y_p
@@ -156,7 +179,8 @@ class Hgrid():
 
 if __name__ == "__main__":
    h = Hgrid()
-   ant_x,ant_y = h.hex_grid_ver2(1,20)
+   #ant_x,ant_y = h.hex_grid_ver2(1,20)
+   ant_x,ant_y = h.square_grid(3,20)
    print "ant_x = ",ant_x
    print "ant_y = ",ant_y
    plt.plot(ant_x,ant_y,"ro")

### 2016-07-08 expression_calc/hex_grid.py
added ew-array

@@ -142,6 +142,21 @@ class Hgrid():
  
         return ant_x,ant_y
 
+    def line_grid(self,elements,l):
+                       
+        ant_x = np.zeros((elements,),dtype=float)
+        ant_y = np.zeros((elements,),dtype=float)
+        print "len(ant_x) = ",len(ant_x)
+        print "len(ant_y) = ",len(ant_y)
+        x = 0.0
+        y = 0.0
+
+        for k in xrange(elements):
+            ant_x[k] = x
+            x = x + l
+
+        return ant_x,ant_y
+
     def determine_phi_value(self,red_vec_x,red_vec_y,ant_x_p,ant_x_q,ant_y_p,ant_y_q):
         red_x = ant_x_q - ant_x_p
         red_y = ant_y_q - ant_y_p
@@ -179,8 +194,9 @@ class Hgrid():
 
 if __name__ == "__main__":
    h = Hgrid()
-   #ant_x,ant_y = h.hex_grid_ver2(1,20)
-   ant_x,ant_y = h.square_grid(3,20)
+   #ant_x,ant_y = h.hex_grid_ver2(5,20)
+   ant_x,ant_y = h.square_grid(10,20)
+   #ant_x,ant_y = h.line_grid(100,20)
    print "ant_x = ",ant_x
    print "ant_y = ",ant_y
    plt.plot(ant_x,ant_y,"ro")

### 2016-07-11 expression_calc/solver.py
testing general formulation

@@ -95,6 +95,166 @@ class solver():
              plt.imshow(self.J_int,interpolation="nearest")
              plt.show() 
 
+
+      def hex_grid_ver2(self,hex_dim,l):
+          hex_dim = int(hex_dim)
+          side = int(hex_dim + 1)
+          ant_main_row = int(side + hex_dim)
+        
+          elements = 1
+
+          #summing the antennas in hexactoganal rings 
+          for k in xrange(hex_dim):
+              elements = elements + (k+1)*6
+                
+          ant_x = np.zeros((elements,),dtype=float)
+          ant_y = np.zeros((elements,),dtype=float)
+          print "len(ant_x) = ",len(ant_x)
+          print "len(ant_y) = ",len(ant_y)
+          x = 0.0
+          y = 0.0
+
+          counter = 0
+        
+          for k in xrange(side):
+              x_row = x
+              y_row = y
+              for i in xrange(ant_main_row):
+                  if k == 0:
+                     ant_x[counter] = x_row 
+                     ant_y[counter] = y_row
+                     x_row = x_row + l
+                     counter = counter + 1 
+                  else:
+                     ant_x[counter] = x_row
+                     ant_y[counter] = y_row
+                     counter = counter + 1
+                   
+                     ant_x[counter] = x_row
+                     ant_y[counter] = -1*y_row
+                     x_row = x_row + l
+                     counter = counter + 1   
+              x = x + l/2.0
+              y = y + (np.sqrt(3)/2.0)*l                 
+              ant_main_row = ant_main_row - 1
+       
+          y_idx = np.argsort(ant_y)
+          ant_y = ant_y[y_idx]
+          ant_x = ant_x[y_idx]
+
+          slice_value = int(side)
+          start_index = 0
+          add = True
+          ant_main_row = int(side + hex_dim)
+
+          for k in xrange(ant_main_row):
+              temp_vec_x = ant_x[start_index:start_index+slice_value]
+              x_idx = np.argsort(temp_vec_x)
+              temp_vec_x = temp_vec_x[x_idx]
+              ant_x[start_index:start_index+slice_value] = temp_vec_x
+              if slice_value == ant_main_row:
+                 add = False
+              start_index = start_index+slice_value 
+            
+              if add:
[…]

### 2016-07-11 expression_calc/solver.py
matrix C

@@ -210,11 +210,11 @@ class solver():
 
           for l in xrange(len(red_vec_x)):
               if (np.allclose(red_x,red_vec_x[l]) and np.allclose(red_y,red_vec_y[l])):
-                 return red_vec_x,red_vec_y,l+1
+                 return red_vec_x,red_vec_y,int(l+1)
 
           red_vec_x = np.append(red_vec_x,np.array([red_x]))
           red_vec_y = np.append(red_vec_y,np.array([red_y]))
-          return red_vec_x,red_vec_y,len(red_vec_x) 
+          return red_vec_x,red_vec_y,int(len(red_vec_x)) 
 
       def calculate_phi(self,ant_x,ant_y,plot=True):
           phi = np.zeros((len(ant_x),len(ant_y)))
@@ -242,16 +242,54 @@ class solver():
           print "phi = ",phi
           return phi,zeta
 
+      def xi_func_eval(self,i,j,phi):
+          column = phi[:,i]
+
+          for p in xrange(len(column)):
+              if column[p] == j:
+                 return p
+
+          return 0
+
+      def psi_func_eval(self,i,j,phi):
+          row = phi[i,:]
+
+          for q in xrange(len(row)):
+              if column[q] == j:
+                 return q
+
+          return 0
+
       '''
       INPUTS:
       z - Input vector.
       N - Number of elements in array.
       plot - Plot the resultant matrices if True.     
       '''
-      def generate_H_formula(self,z,N,plot=False):
+      def generate_H_formula(self,z,N,plot=False,dim=1,l=20,layout="HEX"):
           g = z[:N]
           y = z[N:]
-           
+          R = len(y)
+          if layout == "HEX":
+             ant_x,ant_y = self.hex_grid_ver2(dim,l)
+          elif layout == "LIN":
+             ant_x,ant_y = self.line_grid(dim,l)
+          else:
+             ant_x,ant_y = self.square_grid(dim,l)
+  
+          phi,zeta = self.calculate_phi(ant_x,ant_y,plot=False)
+          
+          #CONSTRUCTING C
+          #***************
+          C = np.zeros((N,N))
+          for i in xrange(N):
+              sum_v = 0
+              for k in xrange(N):
+                  if k != i:
+                     sum_v = np.absolute(g[k])**2*np.absolute(y[zeta[k,i]-1])**2 + sum_v
+              C[i,i] = sum_v
+
+          
 
          
 

### 2016-07-11 expression_calc/solver.py
matrix D,E

@@ -280,7 +280,7 @@ class solver():
           phi,zeta = self.calculate_phi(ant_x,ant_y,plot=False)
           
           #CONSTRUCTING C
-          #***************
+          #**************
           C = np.zeros((N,N))
           for i in xrange(N):
               sum_v = 0
@@ -289,7 +289,29 @@ class solver():
                      sum_v = np.absolute(g[k])**2*np.absolute(y[zeta[k,i]-1])**2 + sum_v
               C[i,i] = sum_v
 
-          
+
+          #CONSTRUCTING E
+          #**************
+          E = np.zeros((R,R))
+          for i in xrange(R):
+              pq = zip(*np.where(phi == (i+1))) #[(0, 0), (0, 1), (1, 2), (1, 3)]
+              sum_v = 0
+              for k in xrange(len(pq)):
+                  sum_v = np.absolute(g[pq[k][0]])^2*np.absolute(g[pq[k][1]])^2 + sum_v
+              E[i,i] = sum_v
+
+          #CONSTRUCTING D
+          #**************
+          D = np.zeros((N,R))
+          for i in xrange(N):
+              for j in xrange(R):
+                  psi = psi_func_eval(i,j+1,phi)
+                  if psi <> 0:
+                     D[i,j] = g[i]*y[j]*np.absolute(g[psi])**2 
+
+          DH = D.transpose.conj()   
+
+
 
          
 

### 2016-07-11 expression_calc/solver.py
G matrix

@@ -289,7 +289,6 @@ class solver():
                      sum_v = np.absolute(g[k])**2*np.absolute(y[zeta[k,i]-1])**2 + sum_v
               C[i,i] = sum_v
 
-
           #CONSTRUCTING E
           #**************
           E = np.zeros((R,R))
@@ -309,11 +308,32 @@ class solver():
                   if psi <> 0:
                      D[i,j] = g[i]*y[j]*np.absolute(g[psi])**2 
 
-          DH = D.transpose.conj()   
+          DH = D.transpose.conj()  
 
+          #CONSTRUCTING F
+          #**************
+          F = np.zeros((N,N))
+          for i in xrange(N):
+              for j in xrange(i,N):
+                  if i <> j:
+                     F[i,j] = g[i]*g[j]*y[phi[i,j]-1]
+                     F[j,i] = F[i,j] 
 
+          #CONSTRUCTING G
+          #**************
+          G = np.zeros((N,R))
+          for i in xrange(N):
+              for j in xrange(R):
+                  xi = xi_func_eval(i,j+1,phi)
+                  if xi <> 0:
+                     G[i,j] = g[j]*y[i]*np.absolute(g[xi])**2
 
-         
+           GT = G.transpose() 
+
+          #CONSTRUCTING Z
+          #**************
+          Z = np.zeros((R,R))
+        
 
       '''
       Creates the redundant Hessian from variables. Stores the result in self.H_sym

### 2016-07-11 expression_calc/solver.py
A,B matrix

@@ -306,7 +306,7 @@ class solver():
               for j in xrange(R):
                   psi = psi_func_eval(i,j+1,phi)
                   if psi <> 0:
-                     D[i,j] = g[i]*y[j]*np.absolute(g[psi])**2 
+                     D[i,j] = g[i]*y[j].conj()*np.absolute(g[psi])**2 
 
           DH = D.transpose.conj()  
 
@@ -326,13 +326,40 @@ class solver():
               for j in xrange(R):
                   xi = xi_func_eval(i,j+1,phi)
                   if xi <> 0:
-                     G[i,j] = g[j]*y[i]*np.absolute(g[xi])**2
+                     G[i,j] = g[i]*y[j]*np.absolute(g[xi])**2
 
            GT = G.transpose() 
 
           #CONSTRUCTING Z
           #**************
           Z = np.zeros((R,R))
+
+          #CONSTRUCTING A
+          #**************
+          A = np.zeros((N+R,N+R))
+          A[:N,:N] = C
+          A[:N,N:] = D
+          A[N:,N:] = E
+          A[N:,:N] = DH
+
+          #CONSTRUCTING B
+          #**************
+          B = np.zeros((N+R,N+R))
+          B[:N,:N] = F
+          B[:N,N:] = G
+          B[N:,N:] = Z
+          B[N:,:N] = GT
+
+          #CONSTRUCTING H
+          #**************
+          H = np.zeros((2*(N+R),2*(N+R)))
+          H[:N+R,:N+R] = A
+          H[:N+R,N+R:] = B
+          H[N+R:,:N+R] = B.conj()
+          H[N+R:,N+R:] = A.conj()
+
+          
+          
         
 
       '''

### 2016-07-11 expression_calc/solver.py
H formula works for EW array

@@ -217,14 +217,14 @@ class solver():
           return red_vec_x,red_vec_y,int(len(red_vec_x)) 
 
       def calculate_phi(self,ant_x,ant_y,plot=True):
-          phi = np.zeros((len(ant_x),len(ant_y)))
-          zeta = np.zeros((len(ant_x),len(ant_y)))
+          phi = np.zeros((len(ant_x),len(ant_y)),dtype=int)
+          zeta = np.zeros((len(ant_x),len(ant_y)),dtype=int)
           red_vec_x = np.array([])
           red_vec_y = np.array([])
           for k in xrange(len(ant_x)):
               for j in xrange(k+1,len(ant_x)):
                   red_vec_x,red_vec_y,phi[k,j]  = self.determine_phi_value(red_vec_x,red_vec_y,ant_x[k],ant_x[j],ant_y[k],ant_y[j])           
-                  zeta[k,j] phi[k,j]
+                  zeta[k,j] = phi[k,j]
                   zeta[j,k] = zeta[k,j]
                   #phi[j,k] = phi[k,j]
 
@@ -245,20 +245,23 @@ class solver():
       def xi_func_eval(self,i,j,phi):
           column = phi[:,i]
 
+          #print "column = ",column
+          #print "j = ",j
+
           for p in xrange(len(column)):
               if column[p] == j:
-                 return p
+                 return p,True
 
-          return 0
+          return 0,False
 
       def psi_func_eval(self,i,j,phi):
           row = phi[i,:]
 
           for q in xrange(len(row)):
-              if column[q] == j:
-                 return q
+              if row[q] == j:
+                 return q,True
 
-          return 0
+          return 0,False
 
       '''
       INPUTS:
@@ -266,7 +269,7 @@ class solver():
       N - Number of elements in array.
       plot - Plot the resultant matrices if True.     
       '''
-      def generate_H_formula(self,z,N,plot=False,dim=1,l=20,layout="HEX"):
+      def generate_H_formula(self,z,N,plot=False,dim=5,l=20,layout="LIN"):
           g = z[:N]
           y = z[N:]
           R = len(y)
@@ -281,9 +284,9 @@ class solver():
           
           #CONSTRUCTING C
           #**************
-          C = np.zeros((N,N))
[…]

### 2016-07-11 expression_calc/H_gen.py
test if formula works for general layout

@@ -778,7 +778,49 @@ class redundant():
                   t_temp.append_factor(y_fact)
                   t_temp.append_factor(gc_fact)
                   self.regular_array = np.append(self.regular_array,t_temp)
+
+      def square_grid(self,side,l):
+          elements = side*side
+                
+          ant_x = np.zeros((elements,),dtype=float)
+          ant_y = np.zeros((elements,),dtype=float)
+          print "len(ant_x) = ",len(ant_x)
+          print "len(ant_y) = ",len(ant_y)
+          x = 0.0
+          y = 0.0
+
+          counter = 0
+        
+          for k in xrange(side):
+              x = 0.0
+              for i in xrange(side):
+                  ant_x[counter] = x
+                  ant_y[counter] = y 
+                  x = x + l
+                  counter = counter + 1
+              y = y + l 
+ 
+          return ant_x,ant_y
                   
+      def create_square(self,side,l,print_pq=False):
+          self.regular_array=np.array([],dtype=object)
+
+          ant_x,ant_y = self.square_grid(side,l)
+          self.N = len(ant_x)
+          #plt.plot(ant_x,ant_y,"ro")
+          #plt.show()
+          self.phi = self.calculate_phi(ant_x,ant_y)
+          
+          for p in xrange(1,self.N):
+              for q in xrange(p+1,self.N+1):
+                  g_fact = factor("g",p,1,False,False,0,0,False) 
+                  y_fact = factor("y",int(self.phi[p-1,q-1]),1,False,False,p,q,print_pq)          
+                  gc_fact = factor("g",q,1,True,False,0,0,False)
+                  t_temp = term()
+                  t_temp.append_factor(g_fact)
+                  t_temp.append_factor(y_fact)
+                  t_temp.append_factor(gc_fact)
+                  self.regular_array = np.append(self.regular_array,t_temp)
 
       def create_regular(self,print_pq=False):
           spacing_vector = np.arange(self.N)[1:]  

### 2016-07-11 expression_calc/solver.py

@@ -66,6 +66,29 @@ class solver():
           r.conjugate_J1_J2()
           r.create_J()
           self.J_sym = r.J
+
+
+      '''
+      Creates the redundant Jacobian from variables. Stores the result in self.J_sym
+      
+      RETURNS:
+      N/A
+ 
+      INPUTS:
+      N - Number of elements in array.     
+      '''
+      def construct_sym_J_layout(self,N,dim=1,l=20,layout="HEX"):
+          r = H_gen.redundant(N)
+          if layout == "HEX":
+             r.create_hexagonal(hex_dim=dim,l=l)
+          else:
+             r.create_square(side=dim,l=l)
+
+          r.create_J1()
+          r.create_J2()
+          r.conjugate_J1_J2()
+          r.create_J()
+          self.J_sym = r.J
           
       '''
       Substitutes the vector z into the symbolic redundant Jacobian. Stores the result in self.J
@@ -262,6 +285,27 @@ class solver():
                  return q,True
 
           return 0,False
+
+
+      def calculate_R(self,dim=1,l=20,layout="HEX"):
+          if layout == "HEX":
+             ant_x,ant_y = self.hex_grid_ver2(dim,l)
+          elif layout == "LIN":
+             ant_x,ant_y = self.line_grid(dim,l)
+          else:
+             ant_x,ant_y = self.square_grid(dim,l)
+          phi,zeta = self.calculate_phi(ant_x,ant_y,plot=False)
+
+          return amax(phi)
+
+      def calculate_random_y(self,dim=1,l=20,layout="HEX"):
+          R = calculate_R(dim=dim,l=l,layout=layout)
+          y = np.zeros((R,),dtype=complex) 
+
+          for k in xrange(y):
+              y[k] = np.random.randn() + 1j*np.random.randn()
+
+          return y
 
       '''
       INPUTS:

### 2016-07-11 expression_calc/H_gen.py
Tested H formula for hexagonal layout

@@ -466,7 +466,9 @@ class redundant():
                    
               #print "###############"
 
-      def create_J(self,type_v="RED"):
+      def create_J(self,type_v="RED",phi=None):
+          if phi <> None:
+             self.phi = phi
           rows = ((self.N*self.N) - self.N)
           if type_v == "RED":
              columns = 2*(self.N + (self.N-1))
@@ -912,7 +914,9 @@ class redundant():
                   H_numerical[r,c] = self.H[r,c].substitute(g,y)   
           return H_numerical        
                   
-      def substitute_J(self,g,y,type_v="RED"):
+      def substitute_J(self,g,y,type_v="RED",phi=None):
+          if phi <> None:
+             self.phi = phi
           if type_v == "RED":
              parameters = 2*(self.N + (self.N-1))
              equations = (self.N**2 -self.N) 

### 2016-07-11 expression_calc/solver.py

@@ -84,10 +84,10 @@ class solver():
           else:
              r.create_square(side=dim,l=l)
 
-          r.create_J1()
-          r.create_J2()
+          r.create_J1(type_v=layout)
+          r.create_J2(type_v=layout)
           r.conjugate_J1_J2()
-          r.create_J()
+          r.create_J(type_v=layout)
           self.J_sym = r.J
           
       '''
@@ -101,12 +101,22 @@ class solver():
       N - Number of elements in array.
       plot - Plot the resultant matrices if True.     
       '''
-      def substitute_sym_J(self,z,N,plot=False):
+      def substitute_sym_J(self,z,N,plot=False,dim=1,l=20,layout="HEX"):
           g = z[:N]
           y = z[N:]
+          
+          if layout == "HEX":
+             ant_x,ant_y = self.hex_grid_ver2(dim,l)
+          elif layout == "LIN":
+             ant_x,ant_y = self.line_grid(dim,l)
+          else:
+             ant_x,ant_y = self.square_grid(dim,l)
+          
+          phi,zeta = self.calculate_phi(ant_x,ant_y,plot=False)
+
           r = H_gen.redundant(N)
           r.J = self.J_sym
-          self.J = r.substitute_J(g,y)
+          self.J = r.substitute_J(g,y,type_v=layout,phi=phi)
           temp_J = np.copy(self.J)
           temp_J[np.absolute(temp_J) > 1e-12] = 1.0
           self.J_int = temp_J.real
@@ -296,13 +306,13 @@ class solver():
              ant_x,ant_y = self.square_grid(dim,l)
           phi,zeta = self.calculate_phi(ant_x,ant_y,plot=False)
 
-          return amax(phi)
+          return np.amax(phi)
 
       def calculate_random_y(self,dim=1,l=20,layout="HEX"):
-          R = calculate_R(dim=dim,l=l,layout=layout)
+          R = self.calculate_R(dim=dim,l=l,layout=layout)
           y = np.zeros((R,),dtype=complex) 
 
-          for k in xrange(y):
+          for k in xrange(R):
               y[k] = np.random.randn() + 1j*np.random.randn()
 
           return y
[…]

### 2016-07-11 expression_calc/H_gen.py
Tested H formula for hexagonal layout and square layout

@@ -390,6 +390,8 @@ class redundant():
              columns = self.N + (self.N-1)
           elif type_v == "HEX":
              columns = self.N + int(np.amax(self.phi))
+          elif type_v == "SQR":
+             columns = self.N + int(np.amax(self.phi))
           else:
              columns = self.N
  
@@ -431,6 +433,8 @@ class redundant():
           if type_v == "RED":
              columns = self.N + (self.N-1)
           elif type_v == "HEX":
+             columns = self.N + int(np.amax(self.phi))
+          elif type_v == "SQR":
              columns = self.N + int(np.amax(self.phi))
           else:
              columns = self.N
@@ -474,6 +478,8 @@ class redundant():
              columns = 2*(self.N + (self.N-1))
           elif type_v == "HEX":
              columns = 2*(self.N + int(np.amax(self.phi)))
+          elif type_v == "SQR":
+             columns = 2*(self.N + int(np.amax(self.phi))) 
           else:
              columns = 2*(self.N)
           self.J = np.empty((rows,columns),dtype=object)
@@ -890,6 +896,8 @@ class redundant():
              parameters = 2*(self.N + (self.N-1)) 
           elif type_v == "HEX":
              parameters = 2*(self.N + int(np.amax(self.phi)))
+          elif type_v == "SQR":
+             parameters = 2*(self.N + int(np.amax(self.phi)))
           else:
              parameters = 2*(self.N) 
           self.H = np.empty((parameters,parameters),dtype=object)  
@@ -906,6 +914,8 @@ class redundant():
              parameters = 2*(self.N + (self.N-1)) 
           elif type_v == "HEX":
              parameters = 2*(self.N + int(np.amax(self.phi)))
+          elif type_v == "SQR":
+             parameters = 2*(self.N + int(np.amax(self.phi)))
           else:
              parameters = 2*(self.N) 
           H_numerical = np.zeros((parameters,parameters),dtype=complex)  
@@ -921,6 +931,9 @@ class redundant():
              parameters = 2*(self.N + (self.N-1))
              equations = (self.N**2 -self.N) 
           elif type_v == "HEX":
+             parameters = 2*(self.N + int(np.amax(self.phi)))
+             equations = (self.N**2 -self.N)
+          elif type_v == "SQR":
              parameters = 2*(self.N + int(np.amax(self.phi)))
              equations = (self.N**2 -self.N)
           else:

### 2016-07-11 expression_calc/solver.py

@@ -88,6 +88,7 @@ class solver():
           r.create_J2(type_v=layout)
           r.conjugate_J1_J2()
           r.create_J(type_v=layout)
+          print "self.J_sym.size = ",r.J.shape
           self.J_sym = r.J
           
       '''
@@ -1057,7 +1058,7 @@ if __name__ == "__main__":
    #print "itr = ",solver_object.itr
    '''
    s = simulator.sim()
-   N = 7
+   N = 9
    g = s.create_antenna_gains(N=N,max_amp=0.1,min_amp=0.05,freq_scale=5,time_steps=600,plot=True)
    #point_sources = s.create_point_sources(num_sources=10,fov=3,a=2)
    #u_m, v_m, w_m, b_m = s.create_uv_regular_line(N=N)
@@ -1067,14 +1068,14 @@ if __name__ == "__main__":
 
    g_0 = g[:,0]
    solver_object = solver()
-   y_0 = solver_object.calculate_random_y(dim=1,l=20,layout="HEX")
+   y_0 = solver_object.calculate_random_y(dim=3,l=20,layout="SQR")
    #y_0 = s.create_y_0_with_M(M,sig=0.01)
    z_0 = np.hstack([g_0,y_0])  
    print "#########################"
    print "len z_0 = ",len(z_0)
    
-   solver_object.construct_sym_J_layout(N,dim=1,l=20,layout="HEX")
-   solver_object.substitute_sym_J(z_0,N,plot=True)
+   solver_object.construct_sym_J_layout(N,dim=3,l=20,layout="SQR")
+   solver_object.substitute_sym_J(z_0,N,plot=True,dim=3,l=20,layout="SQR")
    #solver_object.substitute_sym_H(z_0,N,plot=True)
    #H_temp = np.copy(solver_object.H)
    
@@ -1084,7 +1085,7 @@ if __name__ == "__main__":
    plt.imshow(np.absolute(solver_object.H),interpolation="nearest")
    plt.colorbar()
    plt.show()
-   H_temp = solver_object.generate_H_formula(z_0,N,dim=1,l=20,layout="HEX")
+   H_temp = solver_object.generate_H_formula(z_0,N,dim=3,l=20,layout="SQR")
    plt.imshow(np.absolute(H_temp),interpolation="nearest")
    plt.colorbar()
    plt.show()

### 2016-07-12 paper.tex
started working on paper

@@
 %\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
-
+\newcommand{\bz}{\bmath{z}}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \section{Complex Optimization}
+The most standard way of tackling the calibration problem is to treat the real and imaginary parts of the residual (difference between observed and predicted visibilities) separately and then to independantly solve for the real and imaginary parts
+of the gains. \citet{Smirnov2015} have recently shown that calibration can be performed using complex optimization framework. The main difference between the traditional approach and the 
+complex optimization framework proposed by \citet{Smirnov2015} is that we no longer treat the real and imaginary parts of the residual separately which enables us to directly solve for the complex antenna gains (and their
+complex conjugate). The main reason why we can
+
+%A standard way to optimize a complex function $f(\bz)$ of $n$ complex variables $\bz\in\mathbb{Z}^n$ is to recast $f$ as  optimize the real and imaginary parts of  separately.
+
+The main difference between complex optimization and the more traditional
+approach to calibration is that in complex calibration we
+
 Describe complex optimization, Jacobian, Hessian etc... Benefits drawbacks compared to LINCAL... etc..
-
-\section{Conjugate Gradient method}
-brief description and lit review of the method.
-computational complexity of the algorithm.
-sparsity and condition number.
-
-\section{Results: EW regular-layout}
 
@@
 
+\section{Conjugate Gradient method}
+brief description and lit review of the method.
+computational complexity of the algorithm.
+sparsity and condition number.
 % \section{Methods, Observations, Simulations etc.}
@@
 
+\bibliographystyle{mnras}
+\bibliography{paper}
 % Alternatively you could enter them by hand, like this:
 % This method is tedious and prone to error if you have lots of references
-\begin{thebibliography}{99}
-\bibitem[\protect\citeauthoryear{Author}{2012}]{Author2012}
-Author A.~N., 2013, Journal of Improbable Astronomy, 1, 1
-\bibitem[\protect\citeauthoryear{Others}{2013}]{Others2013}
-Others S., 2012, Journal of Interesting Stuff, 17, 198
-\end{thebibliography}
+%\begin{thebibliography}{99}
+%\bibitem[\protect\citeauthoryear{Author}{2012}]{Author2012}
+%Author A.~N., 2013, Journal of Improbable Astronomy, 1, 1
+%\bibitem[\protect\citeauthoryear{Others}{2013}]{Others2013}
+%Others S., 2012, Journal of Interesting Stuff, 17, 198
+%\end{thebibliography}
 

### 2016-07-12 paper.tex
complex optimization section

@@
 \newcommand{\bz}{\bmath{z}}
+\newcommand{\conj}[1]{\overline{#1}}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \section{Complex Optimization}
-The most standard way of tackling the calibration problem is to treat the real and imaginary parts of the residual (difference between observed and predicted visibilities) separately and then to independantly solve for the real and imaginary parts
-of the gains. \citet{Smirnov2015} have recently shown that calibration can be performed using complex optimization framework. The main difference between the traditional approach and the 
-complex optimization framework proposed by \citet{Smirnov2015} is that we no longer treat the real and imaginary parts of the residual separately which enables us to directly solve for the complex antenna gains (and their
-complex conjugate). The main reason why we can
+The most standard way of tackling the calibration problem is to treat the real and imaginary parts of the residual (difference between observed and predicted visibilities) separately and then to simultaneously solve for the real and imaginary parts
+of the gains. \citet{Smirnov2015} have recently shown that calibration can be performed using a complex optimization framework. The main difference between the traditional approach and the 
+complex optimization framework proposed by \citet{Smirnov2015} is that in their framework they no longer treat the real and imaginary parts of the residual separately which enables them to directly solve for the complex antenna gains themselves (as well as their
+complex conjugates). Recall that most optimization techniques require a gradient operator to enable it to function properly. The main reason why we have traditionally treated 
+the real and imaginary parts of the residual separately is due to the fact that $f(z) = \conj{z}$ is not a differentiable complex function where $z\in\mathbb{C}$ and $\conj{(*)}$ denotes conjugation, i.e.
+$\frac{df(z)}{dz}$
+does not exist, if we apply the standard definition of differentiation:
+\begin{equation}
+x
+\end{equation}
+
+The main reason that we do not need to treat the real and imaginary parts of the residual separately any more is due to the 
+fact
 

### 2016-07-12 paper.tex
made little progress with complex optimization section

@@
 \newcommand{\bz}{\bmath{z}}
+\newcommand{\br}{\bmath{r}}
+\newcommand{\bg}{\bmath{g}}
+\newcommand{\bd}{\bmath{d}}
+\newcommand{\bv}{\bmath{v}}
 \newcommand{\conj}[1]{\overline{#1}}
@@
 \section{Complex Optimization}
+Scalar direction independent calibration is achieved through solving the following minimization problem:
+\begin{equation}
+\min_{\bg}\|\br\| = \min_{\bg}\|\bd - \bv(\bg)\|, 
+\end{equation}
+where
+\begin{enumerate}
+ \item $\bd$ is the observed visibility vector, i.e. 
+ \begin{equation}
+  \bd = [d_{pq}] \big \} ~ [pq] = 1\cdots N ~ (p<q).
+ \end{equation}
+ The vector entry $d_{pq}$ denotes the observed visibility associated with the baseline formed by antenna $p$ and $q$, while $N$ denotes the number or antennas
+ in the array. 
+ \item $\bg = [g_1,g_2,\cdots,g_N]$ represents the instrumental complex antenna gain vector.
+ \item $\bv(\bg)$ denotes the predicted visibility vector, i.e.
+ \begin{equation}
+  \bv(\bg) = [g_p y_{pq}\conj{g}_q]  \big \} ~ [pq] = 1\cdots N ~ (p<q).
+ \end{equation}
+ In Eq... $y_{pq}$ denotes the synthetic visibility which was generated by
+
+\end{enumerate}
+
 The most standard way of tackling the calibration problem is to treat the real and imaginary parts of the residual (difference between observed and predicted visibilities) separately and then to simultaneously solve for the real and imaginary parts

### 2016-07-12 paper.tex
started with an equation approach instead

@@
  \end{equation}
- In Eq... $y_{pq}$ denotes the synthetic visibility which was generated by
-
+ In Eq... $y_{pq}$ denotes the visibility which was created from a skymodel and is associated with antenna $p$ and $q$. [Need to rw this sentence] 
 \end{enumerate}
@@
 The most standard way of tackling the calibration problem is to treat the real and imaginary parts of the residual (difference between observed and predicted visibilities) separately and then to simultaneously solve for the real and imaginary parts
-of the gains. \citet{Smirnov2015} have recently shown that calibration can be performed using a complex optimization framework. The main difference between the traditional approach and the 
+of the gains. 
+\begin{equation}
+\min_{\hat{\bg}}\|\hat{\br}\| = \min_{\hat{\bg}}\|\hat{\bd} - \hat{\bv}(\hat{\bg})\|, 
+\end{equation}
+where $\hat{\bg} = \big[\mathcal{R}\{\bg\},\mathcal{I}\{\bg\}\big]^T$, still need to define the other vectors properly....
+
+\citet{Smirnov2015} have recently shown that calibration can be performed using a complex optimization framework. 
+\begin{equation}
+\min_{\breve{\bg}}\|\breve{\br}\| = \min_{\breve{\bg}}\|\breve{\bd} - \breve{\bv}(\breve{\bg})\|, 
+\end{equation}
+where $\breve{\bg} = \big[\bg,\conj{\bg}]^T$, still need to define the other vectors properly....
+
+The main difference between the traditional approach and the 
 complex optimization framework proposed by \citet{Smirnov2015} is that in their framework they no longer treat the real and imaginary parts of the residual separately which enables them to directly solve for the complex antenna gains themselves (as well as their

### 2016-07-12 paper.tex
more progress

@@
 
-\citet{Smirnov2015} have recently shown that calibration can be performed using a complex optimization framework. 
+We have to revert to the above formulation as a direct Taylor expansion of v is not possible as v is nonanalytic in its argument.
+
+\citet{Smirnov2015} have recently shown that calibration can be performed using the following complex framework 
 \begin{equation}
@@
 
-The main difference between the traditional approach and the 
-complex optimization framework proposed by \citet{Smirnov2015} is that in their framework they no longer treat the real and imaginary parts of the residual separately which enables them to directly solve for the complex antenna gains themselves (as well as their
-complex conjugates). Recall that most optimization techniques require a gradient operator to enable it to function properly. The main reason why we have traditionally treated 
+The above formulation is based on the fact that v breve is analytic in its argument and its complex conjugate as a whole.
+
+%The main difference between the traditional approach and the 
+%complex optimization framework proposed by \citet{Smirnov2015} is that in their framework they no longer treat the real and imaginary parts of the residual separately which enables them to directly solve for the complex antenna gains themselves (as well as their
+%complex conjugates).
+
+Recall that most optimization techniques require a gradient operator to enable it to function properly. The main reason why we have traditionally treated 
 the real and imaginary parts of the residual separately is due to the fact that $f(z) = \conj{z}$ is not a differentiable complex function where $z\in\mathbb{C}$ and $\conj{(*)}$ denotes conjugation, i.e.

### 2016-07-14 paper.tex
more progress
paper.tex

@@
 
-The most standard way of tackling the calibration problem is to treat the real and imaginary parts of the residual (difference between observed and predicted visibilities) separately and then to simultaneously solve for the real and imaginary parts
-of the gains. 
+One of the simplests approaches one could use to solve eq. is to use least-squares optimization which we can only employ if we are able to calculate the gradient of $\bv(\bg)$,
+implying that $\bv(\bg)$ needs to be analytic in $\bg$ (the Taylor series expansion of $\bv(\bg)$ exists).
+The problem we face if we try to naivly apply least-squares to eq is that $\bv(\bg)$ is not analytic in its argument, since $\frac{\partial \conj{z}}{\partial z}$ (where 
+$z \in \mathbb{C}$) does not exist.
+
+The easiest way to circumvent this problem is to recast our problem in such a way that the predicted visibility vector becomes analytic in its argument.
+The most widely used way to accomplish this is to restate eq as
 \begin{equation}
@@
 where $\hat{\bg} = \big[\mathcal{R}\{\bg\},\mathcal{I}\{\bg\}\big]^T$, still need to define the other vectors properly....
+Note that $\hat{\bv}(\hat{\bg})$ is now analytic in its argument and we can proceed and use least squares to estimate $\bg$.
 
-We have to revert to the above formulation as a direct Taylor expansion of v is not possible as v is nonanalytic in its argument.
-
-\citet{Smirnov2015} have recently shown that calibration can be performed using the following complex framework 
+\citet{Smirnov2015} have recently shown that calibration can also be performed using the following complex framework: 
 \begin{equation}
@@
 where $\breve{\bg} = \big[\bg,\conj{\bg}]^T$, still need to define the other vectors properly....
-
-The above formulation is based on the fact that v breve is analytic in its argument and its complex conjugate as a whole.
+The above formulation is valid and is in fact equivalent to the real/imaginary formulation from above, since v breve is analytic in its argument and its complex conjugate as a whole if we use the Wirtinger derivative
+to define our gradient operator.
 

### 2016-07-14 paper.tex
starting redundant index function

@@
 \section{Complex Optimization}
+\subsection{Scalar Calibration}
 Scalar direction independent calibration is achieved through solving the following minimization problem:
@@
 where $\breve{\bg} = \big[\bg,\conj{\bg}]^T$, still need to define the other vectors properly....
-The above formulation is valid and is in fact equivalent to the real/imaginary formulation from above, since v breve is analytic in its argument and its complex conjugate as a whole if we use the Wirtinger derivative
-to define our gradient operator.
+In the above formulation v breve is analytic in its argument and its complex conjugate as a whole under the assumption that the Wirtinger derivative
+is used to define the gradient operator. As v breve is analytic we will are also able to apply the least squares method to .
+
+The Wirtinger derivatives are defined as
+\begin{eqnarray}
+\frac{\partial}{\partial z} = \frac{1}{2}\left ( \frac{\partial}{\partial x} -  i \frac{\partial}{\partial y} \right ),&~&\frac{\partial}{\partial \conj{z}} = \frac{1}{2}\left ( \frac{\partial}{\partial x} +  i \frac{\partial}{\partial y} \right )
+\end{eqnarray}
+where $z = x + iy$ and $x,y \in \mathbb{R}$.
+It follows from the above that $\frac{\partial z}{\partial z} = 1,\frac{\partial z}{\partial \conj{z}} = 0,\frac{\partial \conj{z}}{\partial z} = 0$ and $\frac{\partial \conj{z}}{\partial \conj{z}} = 1$.
 
@@
 approach to calibration is that in complex calibration we
+
+\subsection{Redundant Calibration}
+In eq a sky model is required to generate the predicted visibilities $y_{pq}$ as we have too few equations to also solve for them and the gains. If we use a redundant layout
+we can reduce the number of unknowns so that we are able to solve for the sky and the gains simultaneously.
+
+\subsubsection{Antenna indices to redundancy index mapping}
+If an array is redundant its redundant baselines sample the exact same visibilites in the $uv$-plane. So even though we label the visibilies belonging to different redundant baselines with
+a different composite antenna index they actually represent the same visibilities. Another useful labelling strategy therefore would be to assign redundant indices to each unique
+redundant baseline of the array. Where the numbering can start from one and is incrementet every time a new redundant spacing is added to the existing baselines if we 
+add one antenna after another to our array in numerical order. A usefull mapping would then be the function that maps composite antenna indices to redundancy indices. 
+Let $\phi_{pq}:\mathbb{N}^2\rightarrow\mathbb{N}^+$ be the mapping that maps antenna indices to the redundant spacing indices of the array. Also note that this mapping is not symmetric as 
+$\phi_{pq} = 0$ if $p>q$. Moreover, let
+
+\begin{equation}
+\psi_{ij} = 
+\begin{cases}
+q~\textrm{if}~\exists! ~ q \in \mathbb{N} ~ s.t. ~(\phi_{iq} = j)\\
+0~\textrm{otherwise}
+\end{cases}
+\end{equation},
+
+\begin{equation}
+\xi_{ij} = 
+\begin{cases}
+p~\textrm{if}~\exists! ~ p \in \mathbb{N} ~ s.t. ~(\phi_{pi} = j)\\
+0~\textrm{otherwise}
+\end{cases}
+\end{equation}
+
+and
+
+\begin{equation}
+\zeta_{ij} = 
+\begin{cases}
+\phi_{ij}~\textrm{if}~i<j\\
+\phi_{ji}~\textrm{if}~i>j\\
+0~\textrm{otherwise}
+\end{cases}
+\end{equation}
 

### 2016-07-14 paper.tex
added beginnings of index to index mapping

@@
 \subsubsection{Antenna indices to redundancy index mapping}
-If an array is redundant its redundant baselines sample the exact same visibilites in the $uv$-plane. So even though we label the visibilies belonging to different redundant baselines with
+If an array is redundant its redundant baselines sample the exact same visibilites in the $uv$-plane. We denote the true observed visibility associated with baseline $pq$ and 
+baseline $rs$ with $y_{pq}$ and $y_{rs}$ respectively . However if baseline $pq$ and $rs$ represent the same redundant baselines then $y_{pq} = y_{rs}$. Another
+usefull indexing strategy would therefore be 
+, if they are redundant they are in fact the same value. belonging to different redundant baselines with
 a different composite antenna index they actually represent the same visibilities. Another useful labelling strategy therefore would be to assign redundant indices to each unique

### 2016-07-15 paper.tex
paper.tex

--- 2016-07-14 paper.tex
@@
 \subsubsection{Antenna indices to redundancy index mapping}
-If an array is redundant its redundant baselines sample the exact same visibilites in the $uv$-plane. So even though we label the visibilies belonging to different redundant baselines with
-a different composite antenna index they actually represent the same visibilities. Another useful labelling strategy therefore would be to assign redundant indices to each unique
-redundant baseline of the array. Where the numbering can start from one and is incrementet every time a new redundant spacing is added to the existing baselines if we 
-add one antenna after another to our array in numerical order. A usefull mapping would then be the function that maps composite antenna indices to redundancy indices. 
-Let $\phi_{pq}:\mathbb{N}^2\rightarrow\mathbb{N}^+$ be the mapping that maps antenna indices to the redundant spacing indices of the array. Also note that this mapping is not symmetric as 
-$\phi_{pq} = 0$ if $p>q$. Moreover, let
+If an array is redundant its redundant baselines sample the exact same visibilites in the $uv$-plane. Recall, that we denote the true observed visibility associated with baseline $pq$ and 
+baseline $rs$ with $y_{pq}$ and $y_{rs}$, respectively . However if baseline $pq$ and $rs$ represent the same redundant baseline then $y_{pq} = y_{rs}$. Another
+usefull indexing strategy would therefore be to use redundant baseline indices instead of antenna pairs to label the true observed visibilities so that we can identify 
+which baselines observe the same visibilities barring instrumental and environmental errors.
 
+Let $\phi_{ij}:\mathbb{N}^2\rightarrow\mathbb{N}^+$ be the mapping that maps antenna indices to the unique redundant baseline indices of the array. This mapping is not symmetric as 
+$\phi_{ij} = 0$ if $i>j$. For an east-west regular array we can construct a simple analytic expression for this mapping:
+\begin{equation}
+\widehat{\phi_{ij}} = 
+\begin{cases}
+j-i&\textrm{if}~i<j\\
+0&\textrm{otherwise}
+\end{cases}.
+\end{equation}
+As will become apparent later the following symmetric variant of $\phi_{ij}$ will turn out to be quite useful later on:
+\begin{equation}
+\zeta_{ij} = 
+\begin{cases}
+\phi_{ij}~\textrm{if}~i\leq j\\
+\phi_{ji}~\textrm{if}~i>j
+\end{cases}.
+\end{equation} 
+It is also crucial to be able to determine the antenna index of the antenna which was used to construct a specific redundant baseline if one is given the other antenna index and the redundant baseline index.
+This can be calculated via the following two functions:
+\begin{equation}
+\xi_{ij} = 
+\begin{cases}
+p~\textrm{if}~\exists! ~ p \in \mathbb{N} ~ s.t. ~(\phi_{pi} = j)\\
+0~\textrm{otherwise}
+\end{cases},
+\end{equation}
+and
 \begin{equation}
@@
 \end{cases}
-\end{equation},
-
-\begin{equation}
-\xi_{ij} = 
-\begin{cases}
-p~\textrm{if}~\exists! ~ p \in \mathbb{N} ~ s.t. ~(\phi_{pi} = j)\\
-0~\textrm{otherwise}
-\end{cases}
 \end{equation}
-
-and
-
-\begin{equation}
-\zeta_{ij} = 
-\begin{cases}
-\phi_{ij}~\textrm{if}~i<j\\
-\phi_{ji}~\textrm{if}~i>j\\
-0~\textrm{otherwise}
-\end{cases}
-\end{equation}
+We use $\xi_{ij}$ to retrieve the index of the second antenna if we know the redundant baseline index and the index of the first antenna used to construct the baseline in question, while we use $\psi_{ij}$ to return the index of the first antenna if we know the
+index of the second antenna instead. \textbf{NB::} I still need to plot $\zeta_{ij}$ for different antenna layouts.
 

### 2016-07-15 paper.tex
newest paper version

@@
 \newcommand{\bv}{\bmath{v}}
+\newcommand{\by}{\bmath{y}}
 \newcommand{\conj}[1]{\overline{#1}}
@@
 Describe complex optimization, Jacobian, Hessian etc... Benefits drawbacks compared to LINCAL... etc..
+
+\subsubsection{Redundant calibration}
+In the case of a redundant array we can not only solve for the gains $\bg$ but also the redundant visibilities $\by$ themselves. Where $\by = [y_1,\cdots,y_R]$, $y_i$ denotes 
+a unique redundant visibility and $R$ is the maximum value of the mapping $\phi_{ij}$. The reason we can now also solve for $\by$ is because we have reduced the number of 
+unknowns from $N+B$ to $N+R$, where $N+R$ is now less than $B$ due to the array redundancy. We can now reformulate the calibration problem for an redundant array as follow:
+\begin{equation}
+\min_{\bz}\|\br\| = \min_{\bz}\|\bd - \bv(\bz)\|, 
+\end{equation}
+where $\bz = [\bg,\by]^T$.
 

### 2016-07-18 paper.tex
some minor changes to document

@@
 \newcommand{\by}{\bmath{y}}
+\newcommand{\bJ}{\bmath{J}}
+\newcommand{\bD}{\bmath{D}}
 \newcommand{\conj}[1]{\overline{#1}}
@@
 \subsubsection{Redundant calibration}
-In the case of a redundant array we can not only solve for the gains $\bg$ but also the redundant visibilities $\by$ themselves. Where $\by = [y_1,\cdots,y_R]$, $y_i$ denotes 
+In the case of a redundant array we can not only solve for the gains $\bg$ but we can also solve for the redundant unknown visibilities $\by$. Where $\by = [y_1,\cdots,y_R]$, $y_i$ denotes 
 a unique redundant visibility and $R$ is the maximum value of the mapping $\phi_{ij}$. The reason we can now also solve for $\by$ is because we have reduced the number of 
@@
 
+If we now follow the same extension strategy of \citet{Smirnov2015} we can rewrite eq to
+\begin{equation}
+\min_{\breve{\bz}}\|\breve{\br}\| = \min_{\bz}\|\breve{\bd} - \breve{\bv}(\breve{\bz})\|, 
+\end{equation}
+where $\breve{\bz} = [\bz,\conj{\bz}]^T$ still need to define the remaining variables properly.
+
+We can now solve the above equation by using least-squares, which entails deriving the Gauss-Newton (GN) and Levenberg-Marquerdt (LM) update rules.
+(NEED TO WRITE THIS MUCH BETTER)
+
+\begin{equation}
+\bJ = \frac{\partial\bv}{\partial\bz}~\bJ^* = \frac{\partial \bv}{\partial \breve{\bz}} 
+\end{equation}
+
+\begin{equation}
+\bJ = \begin{bmatrix}
+       \bJ & \bJ^* \\
+       \conj{\bJ^*} & \conj{\bJ}  
+      \end{bmatrix}
+\end{equation}
+
+\begin{equation}
+ \Delta \breve{\bz} = \begin{bmatrix}\Delta \bz\\ \Delta \conj{\bz} \end{bmatrix} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\br_k}
+\end{equation}
+
+\begin{equation}
+ \Delta \breve{\bz} = \begin{bmatrix}\Delta \bz\\ \Delta \conj{\bz} \end{bmatrix} = (\bJ^H\bJ + \lambda \bD)^{-1}\bJ^H\breve{\br_k}
+\end{equation}
+
 \subsection{Jacobian derivation}
@@
 \boldsymbol{A} & \boldsymbol{B}\\
-\boldsymbol{B}^* & \boldsymbol{A}
+\boldsymbol{B}^* & \boldsymbol{A}^*
 \end{bmatrix}
@@
 \begin{cases}
- \sum_{k \neq i} \left | g_k \right |^2 \left | y_{|i-k|} \right |^2 & \textrm{if} ~ i=j\\
+ \sum_{k \neq i} \left | g_k \right |^2 \left | y_{\zeta_{ik}} \right |^2 & \textrm{if} ~ i=j\\
  0 & \textrm{otherwise}
@@
 \begin{cases}
- g_i y_j^*  \left | g_{i+j} \right |^2  & \textrm{if} ~ i<j\\
+ g_i y_j^*  \left | g_{\psi_{ij}} \right |^2  & \textrm{if} ~ \psi_{ij}\neq0\\
  0 & \textrm{otherwise}
@@
 \begin{cases}
- g_i g_j  \left | y_{|i-j|} \right |^2  & \textrm{if} ~ i \neq j\\
+ g_i g_j  \left | y_{\zeta_{ij}} \right |^2  & \textrm{if} ~ (i \neq j)\\
  0 & \textrm{otherwise}
@@
 \begin{cases}
- g_i y_j  \left | g_{i-j} \right |^2  & \textrm{if} ~ i > j\\
+ g_i y_j  \left | g_{\xi_{ij}} \right |^2  & \textrm{if} ~ (\xi_{ij}\neq0)\\
  0 & \textrm{otherwise}
@@
 \begin{equation}
-\mathcal{I}_i = \left\{pq|(q - p) = i \right\}
+\mathcal{I}_i = \left\{pq\in\mathbb{N}^2|(\phi_{pq} = i) \right\}.
 \end{equation}

### 2016-08-17 software/analytic.py
FIRST LINCAL ANALYTIC

@@ -728,11 +728,17 @@ class expression():
       RETURNS:
       string_out - The output string
       '''
-      def to_string(self,simplify=False,simplify_const=False):
+      def to_string(self,simplify=False,simplify_const=False,write_out_zeros=False):
           string_out = ""
           for term in self.terms:
-              if not term.zero: 
-                 string_out = string_out + " + " + term.to_string(simplify=simplify,simplify_constant=simplify_const)
+              if not write_out_zeros:
+                 if not term.zero: 
+                    string_out = string_out + " + " + term.to_string(simplify=simplify,simplify_constant=simplify_const)
+              else:
+                 if term.zero:
+                    string_out = string_out + "," + "0"
+                 else:
+                    string_out = string_out + " , " + term.to_string(simplify=simplify,simplify_constant=simplify_const)
 
           if string_out == "":
              string_out = "0" 
@@ -1744,14 +1750,6 @@ class redundant():
           file.write(string_out)
           file.write("\n\\end{eqnarray}\n") 
 
-          string_out = ""
-          file.write("\\begin{eqnarray}\n")
-          for r in xrange(H_temp.shape[0]):
-              string_out = string_out + "d_{"+str(r)+"} &=& " + H_temp[r,r].to_string(simplify,simplify_const)+"\\\\\n"
-          string_out = string_out[:-3]
-          file.write(string_out)
-          file.write("\n\\end{eqnarray}\n") 
-
           file.write("\\begin{equation}\n")
           file.write("\\boldsymbol{H}_5 = \n\\begin{bmatrix}\n")
           string_out = ""
@@ -1831,12 +1829,11 @@ class redundant():
           file.write(string_out)
           file.write("\n\\end{bmatrix}\n")
           file.write("\\end{equation}\n")
-         
-                     
+
           file.write("\\begin{equation}\n")
-          file.write("\\boldsymbol{H}_2 = \n\\begin{bmatrix}\n")
+          file.write("\\boldsymbol{H}_9 = \n\\begin{bmatrix}\n")
           string_out = ""
-          H_temp = deepcopy(self.H[:2*self.N-1,2*self.N-1:])
+          H_temp = deepcopy(self.H[2*self.N:2*self.N+self.L,:self.N])
           for r in xrange(H_temp.shape[0]):
               for c in xrange(H_temp.shape[1]):
                   #print "r = ",r
[…]

### 2016-08-18 software/LINCAL_test.py
LINCAL rinkle

@@ -1,0 +1,73 @@
+import numpy as np
+import scipy as sp
+import pylab as plt
+import simulator
+
+def LINCAL_J(z,N,L,phi):
+    g = z[:N]
+    y = z[N:]
+
+    rows = (N**2 - N)/2
+    columns = 2*N + 2*L
+
+    J = np.zeros((rows,columns),dtype=complex)
+
+    p_v = np.zeros((rows,),dtype=int)
+    q_v = np.zeros((rows,),dtype=int)
+          
+    counter = 0
+    for k in xrange(N):
+        for j in xrange(k+1,N):
+              p_v[counter] = k
+              q_v[counter] = j
+              counter = counter + 1
+
+    for r in xrange(rows):
+          p = p_v[r]
+          q = q_v[r]
+          phi_v = phi[p,q]
+          for c in xrange(columns):
+              if c == p:
+                 J[r,c] = g[p]*np.conj(g[q])*y[phi_v-1]  
+              if c == q:
+                 J[r,c] = g[p]*np.conj(g[q])*y[phi_v-1]
+              elif (c == (N + p)):
+                 J[r,c] = 1j*g[p]*np.conj(g[q])*y[phi_v-1]
+              elif (c == (N + q)):
+                 J[r,c] = -1j*g[p]*np.conj(g[q])*y[phi_v-1]
+              elif (c == (2*(N) - 1 + phi_v)):
+                 J[r,c] = g[p]*np.conj(g[q])*y[phi_v-1]
+              elif (c == (2*(N) + L - 1 + phi_v)):
+                 J[r,c] = 1j*g[p]*np.conj(g[q])*y[phi_v-1]  
+
+    J = np.vstack([J,np.conj(J)])     
+
+    return J
+
+
+if __name__ == "__main__":
+   s = simulator.sim(layout="REG",order=5)
+   s.generate_antenna_layout()
+   phi,zeta = s.calculate_phi(s.ant[:,0],s.ant[:,1])
+   s.plot_ant(title="REG")
+   s.plot_zeta(data=zeta,step1=1,step2=1,label_size=14,cs="cubehelix")
+   
+   g = np.random.randn(s.N)+1j*np.random.randn(s.N)
+   y = np.random.randn(s.L)+1j*np.random.randn(s.L)
+   z = np.hstack([g,y])
+
+   J = LINCAL_J(z,s.N,s.L,phi)
+
+   H = np.dot(J.conj().transpose(),J)
[…]

### 2016-08-18 software/analytic.py

@@ -945,7 +945,16 @@ class redundant():
                      #print "Hallo_zero"
                      #print "t.zero = ",t.zero
                   self.J[r,c].to_string()
-                  #print "self.J[r,c] = ",self.J[r,c]
+          
+          J_temp = deepcopy(self.J)
+          
+          for r in xrange(J_temp.shape[0]):
+              for c in xrange(J_temp.shape[1]): 
+                  J_temp[r,c].conjugate()
+          
+          self.J = np.vstack([deepcopy(self.J),J_temp])  
+          #print "self.J[r,c] = ",self.J[r,c]
+
 
        
       def create_J1(self,type_v="RED"):

### 2016-08-18 software/analytic.py
fixing LINCAL rinkle

@@ -2182,6 +2182,9 @@ if __name__ == "__main__":
    plt.imshow(H_int,interpolation="nearest")
    plt.show()
    print "H_int = ",H_int
+   #J_int = r.to_int_J()
+   #plt.imshow(J_int,interpolation="nearest")
+   #plt.show()
    r.to_latex_H2(simplify=True,simplify_const=True)
 
    #e1 = expression(r.J[:,0])

### 2016-08-19 software/LINCAL_test.py
Analytic expression and code are equal

@@ -44,30 +44,182 @@ def LINCAL_J(z,N,L,phi):
 
     return J
 
+def LINCAL_H_analytic(z,N,L,zeta):
+    g = z[:N]
+    y = z[N:] 
+
+    
+    #COMPUTING MATRIX A
+    ####################
+ 
+    A = np.zeros((2*N,2*N),dtype=float)
+
+    #COMPUTING MATRIX D
+    #*******************     
+    D = np.zeros((N,N),dtype=float)
+
+    T = 2*np.diag(np.ones((N,),dtype=float)) - np.ones((N,N),dtype=float)
+
+    for i in xrange(D.shape[0]):
+        for j in xrange(D.shape[1]):
+            #print "i = ",i
+            #print "j = ",j
+            if i <> j:
+               D[i,j] = 2*np.absolute(g[i])**2*np.absolute(g[j])**2*np.absolute(y[zeta[i,j]-1])**2   
+            else:
+               for k in xrange(D.shape[0]):
+                   print "k = ",k
+                   if k <> i:
+                      print "zeta[i,k] = ",zeta[i,k]
+                      D[i,i] = D[i,i] + 2*np.absolute(g[i])**2*np.absolute(g[k])**2*np.absolute(y[zeta[i,k]-1])**2  
+                   
+                       
+    A[:N,:N] = D
+    A[N:,N:] = T*D
+
+    #print "A = ",A
+    #plt.imshow(np.absolute(A),interpolation="nearest")
+    #plt.show()
+
+    #COMPUTING MATRIX B
+    ###################
+
+    B = np.zeros((2*N,2*L),dtype=float) 
+
+    #COMPUTING MATRIX E,F
+    #********************
+
+    E = np.zeros((N,L),dtype=float)
+    F = np.zeros((N,L),dtype=float)
+
+    for i in xrange(E.shape[0]):
+        for j in xrange(E.shape[1]):
+            p,q = create_index_ij(zeta,N,i,j)
+            #print "i = ",i
+            #print "j = ",j
+            #print "p = ",p
+            #print "q = ",q
+            for k in xrange(len(p)):
+                temp = 2*np.absolute(g[p[k]])**2*np.absolute(g[q[k]])**2*np.absolute(y[zeta[p[k],q[k]]-1])**2
+                E[i,j] = E[i,j] + temp 
[…]

### 2016-08-19 software/analytic.py
Started cleaning analytic.py

@@ -772,23 +772,142 @@ class expression():
               if not term.zero:
                  number = number + term.substitute(g_v,y_v)
           return number 
-           
-class redundant():
-      def __init__(self):
-          self.N = 0
-          self.L = 0
-          self.phi = np.array([],dtype=int)
-          self.zeta = np.array([],dtype=int)
-          self.regular_array = np.array([],dtype=object)
-          self.J1 = np.array([],dtype=object)
-          self.Jc1 = np.array([],dtype=object)
-          self.J2 = np.array([],dtype=object)
-          self.Jc2 = np.array([],dtype=object)
-          self.J = np.array([],dtype=object)
-          self.JH = np.array([],dtype=object)
-          self.H = np.array([],dtype=object)
-          self.phi = np.array([])
 
+###############################################################################
+
+'''
+Class that constructs the Jacobian and Hessian for a redundant layout analyticaly
+'''
+class redundant():
+      
+      '''
+      Consructer
+
+      INPUTS:
+      None
+
+      RETURNS:
+      None  
+      '''
+      def __init__(self):
+          
+          self.N = 0 #Number of antennas in the array
+          self.L = 0 #Number of unique redundant baselines 
+          self.phi = np.array([],dtype=int) #Mapping between antennas and redundant index
+          self.zeta = np.array([],dtype=int) #Symmetrical version of phi
+          self.regular_array = np.array([],dtype=object) #The array containing the model  
+          self.J1 = np.array([],dtype=object) #Upper right corner of complex Jacobian
+          self.Jc1 = np.array([],dtype=object) #Lower left corner of complex Jacobian
+          self.J2 = np.array([],dtype=object) #Uppper left corner of complex Jacobian
+          self.Jc2 = np.array([],dtype=object) #Lower right corner of complex Jacobian
+          self.J = np.array([],dtype=object) #Jacobian matrix
+          self.JH = np.array([],dtype=object) #Jacobian hermitian transpose
+          self.H = np.array([],dtype=object) #Hessian matrix
+
+#################################################################################
+# CREATING MODEL
+#################################################################################
+
+
+      '''
+      Creates the array and the assoiciated analytic expression for the model
+      INPUTS:
[…]

### 2016-08-19 software/analytic.py
cleaned up analytic a bit

@@ -805,9 +805,8 @@ class redundant():
           self.H = np.array([],dtype=object) #Hessian matrix
 
 #################################################################################
-# CREATING MODEL
+# CREATING MODEL, JOCOBIAN AND HESSIAN
 #################################################################################
-
 
       '''
       Creates the array and the assoiciated analytic expression for the model
@@ -1136,18 +1135,23 @@ class redundant():
                   #print "*********************"   
                    
               #print "###############"
+      '''
+      Creates the right upper corner of the complex Jacobian
 
-      def create_J2(self,type_v="RED"):
+      NB - CAN ONLY DO THE COMPLEX JACOBIAN NOT LINCAL OR LOGCAL
+
+      INPUTS:
+      None
+      
+      OUTPUTS:
+      None 
+      ''' 
+      def create_J2(self):
           
           rows = ((self.N*self.N) - self.N)/2
-          if type_v == "RED":
-             columns = self.N + (self.N-1)
-          elif type_v == "HEX":
-             columns = self.N + int(np.amax(self.phi))
-          elif type_v == "SQR":
-             columns = self.N + int(np.amax(self.phi))
-          else:
-             columns = self.N
+
+          columns = self.N+self.L
+
           self.J2 = np.empty((rows,columns),dtype=object)
 
           column_vector = np.array([],dtype=object)
@@ -1179,19 +1183,23 @@ class redundant():
                   #print "*********************"   
                    
               #print "###############"
+      '''
+      Creates the complex Jacobian
 
-      def create_J(self,type_v="RED",phi=None):
-          if phi <> None:
-             self.phi = phi
+      NB - CAN ONLY DO THE COMPLEX JACOBIAN NOT LINCAL OR LOGCAL
+
+      INPUTS:
+      None
+      
+      OUTPUTS:
+      None 
+      ''' 
+      def create_J(self):
+          
           rows = ((self.N*self.N) - self.N)
-          if type_v == "RED":
-             columns = 2*(self.N + (self.N-1))
-          elif type_v == "HEX":
-             columns = 2*(self.N + int(np.amax(self.phi)))
-          elif type_v == "SQR":
-             columns = 2*(self.N + int(np.amax(self.phi))) 
-          else:
-             columns = 2*(self.N)
+          
+          columns = self.N+self.L 
+           
           self.J = np.empty((rows,columns),dtype=object)
           self.J[:rows/2,:columns/2] = self.J1
[…]

### 2016-08-20 software/analytic.py
cleaned up a bit more

@@ -818,7 +818,7 @@ class redundant():
       OUTPUTS:
       None 
       '''
-      def create_redundant(self,layout="REG",order=5,print_f=False)
+      def create_redundant(self,layout="REG",order=5,print_pq=False):
           s = simulator.sim(layout=layout,order=order)
           s.generate_antenna_layout()
           phi,zeta = s.calculate_phi(s.ant[:,0],s.ant[:,1])
@@ -1060,7 +1060,7 @@ class redundant():
                      f2 = factor("g",p,1,False,ant_p=0,ant_q=0,print_f=False,value=0)
                      f3 = factor("g",q,1,True,ant_p=0,ant_q=0,print_f=False,value=0)
                      f4 = factor("y",phi_v,1,False,ant_p=p,ant_q=q,print_f=print_f,value=0)     
-                     t = term() OR 
+                     t = term() 
                      t.append_factor(f1)
                      t.append_factor(f2)
                      t.append_factor(f3) 
@@ -1074,13 +1074,13 @@ class redundant():
                      #print "t.zero = ",t.zero
                   self.J[r,c].to_string()
           
+          
+          J_temp = deepcopy(self.J)
           '''
           NB CODE BELOW CALCULATES THE LOWER PART OF JACOBIAN; DO NOT KNOW IF THIS IS THE CORRECT STRATEGY
           OR IF IT IS EVEN DOING WHAT I THINK. WANTED TO JUST REPEAT THE CONJUGATE OF THE MODEL AND DIFFERENTIATE TOWARDS 
           EACH VARIABLE. THIS IS A SIMPLE SHORTCUT WHICH I HOPE ACCOMPLISHES THIS
           '''
-          J_temp = deepcopy(self.J)
-          
           for r in xrange(J_temp.shape[0]):
               for c in xrange(J_temp.shape[1]): 
                   J_temp[r,c].conjugate()
@@ -1198,7 +1198,7 @@ class redundant():
           
           rows = ((self.N*self.N) - self.N)
           
-          columns = self.N+self.L 
+          columns = 2*(self.N+self.L) 
            
           self.J = np.empty((rows,columns),dtype=object)
           self.J[:rows/2,:columns/2] = self.J1
@@ -2236,8 +2236,50 @@ class redundant():
[…]

### 2016-08-20 software/analytic.py
newest cleanup

@@ -2272,150 +2272,77 @@ def Complex_example():
     print "H_int = ",H_int
     r.to_latex_H(simplify=True,simplify_const=True)    
 
+    #REORDERING OF HESSIAN OLD CODE
+    #H_int_old = np.copy(H_int)
+    #H_int[H_int <= 1] = 0
+    #G = nx.Graph(H_int)
+    #rcm = list(reverse_cuthill_mckee_ordering(G))
+    #rcm = list(reverse_cuthill_mckee_ordering(G))
+    #print "rcm = ",rcm
+    #A1 = H_int_old[rcm, :][:, rcm]
+    #plt.imshow(A1,interpolation='nearest')
+    #plt.colorbar()
+    #plt.show()
+
+
+'''
+SIMPLE EXAMPLE
+'''
+def Simple_example():
+    f1 = factor("g",1,1,True)
+    f2 = factor("y",1,1,False,print_f=False)
+    f3 = factor("a",3,1,False)
+    f4 = factor("c",3,1,False,ant_p=3,ant_q=6,print_f=True,value=1)  
+   
+    print "f1 = ",f1.to_string()
+    print "f2 = ",f2.to_string()
+    print "f3 = ",f3.to_string()
+    print "f4 = ",f4.to_string()
+
+    t1 = term()
+    t1.append_factor(f1)
+    t1.append_factor(f2)
+    t1.append_factor(f3)
+    t1.append_factor(f4)
+
+    f4 = factor("g",1,1,True)
+    f5 = factor("y",2,1,False,print_f=False)
+    f6 = factor("b",3,1,False) 
+    f7 = factor("c",3,1,False,ant_p=3,ant_q=5,print_f=True,value=1)
+
+    t2 = term()
+    t2.append_factor(f4)
+    t2.append_factor(f5)
+    t2.append_factor(f6) 
+    t2.append_factor(f7) 
+ 
+    print "t1 = ",t1.to_string()
+    print "t2 = ",t2.to_string()
+  
+    t1.multiply_terms(t2)
+
+    print "t1 = ",t1.to_string()
+
+    t3 = term()
+    t3.setZero()
+
+    print "t3 = ",t3.to_string()
+
+    e1 = expression(np.array([t1,t2],dtype=object)) 
+    e2 = expression(np.array([t1,t3],dtype=object)) 
+
+    print "e1 = ",e1.to_string()
+    print "e2 = ",e2.to_string()   
+
+    e1.dot(e2)   
+ 
[…]

### 2016-08-22 software/analytic.py
started looking into JHv

@@ -803,6 +803,8 @@ class redundant():
           self.J = np.array([],dtype=object) #Jacobian matrix
           self.JH = np.array([],dtype=object) #Jacobian hermitian transpose
           self.H = np.array([],dtype=object) #Hessian matrix
+          self.v = np.array([],dtype=object) #Model array - going to use it to calculate J^Hv - trying to derive redundant SteFCal.
+          self.JHv = np.array([],dtype=object) #J^H times v
 
 #################################################################################
 # CREATING MODEL, JOCOBIAN AND HESSIAN
@@ -837,7 +839,44 @@ class redundant():
                   t_temp.append_factor(y_fact)
                   t_temp.append_factor(gc_fact)
                   self.regular_array = np.append(self.regular_array,t_temp)
-      
+
+      '''
+      Creates the model vector v - will use it to compute J^Hv
+      INPUTS:
+      layout - Which layout to use (HEX, REG or SQR).
+      order - What layout order to use.
+      print_f - If true print the redundant index of factor; if false print the composite antenna index
+
+      OUTPUTS:
+      None 
+      '''
+      def create_v(self,layout="REG",order=5,print_pq=False):
+          s = simulator.sim(layout=layout,order=order)
+          s.generate_antenna_layout()
+          phi,zeta = s.calculate_phi(s.ant[:,0],s.ant[:,1])
+          self.N = s.N
+          self.L = s.L 
+          self.phi = phi
+          self.zeta = zeta
+          
+          for p in xrange(1,self.N):
+              for q in xrange(p+1,self.N+1):
+                  g_fact = factor("g",p,1,False,False,0,0,False) 
+                  y_fact = factor("y",int(self.phi[p-1,q-1]),1,False,False,p,q,print_pq)          
+                  gc_fact = factor("g",q,1,True,False,0,0,False)
+                  t_temp = term()
+                  t_temp.append_factor(g_fact)
+                  t_temp.append_factor(y_fact)
+                  t_temp.append_factor(gc_fact)
+                  self.v = np.append(self.regular_array,t_temp)
+          v_copy = deepcopy(self.v)
+
+          for k in xrange(len(v_copy)):
+              v_copy[k].conjugate()      
+
+          self.v = np.vstack([deepcopy(self.v),v_copy])
+
+
       '''
[…]

### 2016-08-22 software/analytic.py
started derivation of redundant stefcal

@@ -1340,16 +1340,15 @@ class redundant():
       None       
       ''' 
       def compute_JHv(self):
-          parameters = self.JH.shape[1]
-          
-        
+          parameters = self.JH.shape[0]
+                  
           self.JHv = np.empty((parameters,),dtype=object)  
           for r in xrange(parameters):
                 row = expression(self.JH[r,:])
                 row_temp = deepcopy(row)
-                column = expression(self.JHv)
-                  row_temp.dot(column)
-                  self.JHv[r] = row_temp
+                column = expression(self.v)
+                row_temp.dot(column)
+                self.JHv[r] = row_temp
 
       '''
       Substitute values into the HESSIAN
@@ -2160,6 +2159,30 @@ class redundant():
       OUTPUTS:
       None 
       '''    
+      def to_latex_JHv(self,simplify=False,simplify_const=True):
+          file = open("JHv_"+str(self.N)+".txt", "w")
+          file.write("\\boldsymbol{J}^Hv = \n\\begin{eqnarray}\n")
+          string_out = ""
+          JHv_temp = deepcopy(self.JHv)
+          for r in xrange(len(JHv_temp)):
+              string_out = string_out + JHv_temp[r].to_string(simplify,simplify_const)
+              string_out = string_out+"\\\\\n" 
+          string_out = string_out[:-3]
+          file.write(string_out)
+          file.write("\n\\end{eqnarray}\n")
+          file.close()
+
+      '''
+      Writes the hermitian transpose of the JACOBIAN to a latex txt file
+     
+      NB - MIGHT NOT BE COMPATABLE WITH NEW FACTOR PRINT INPUTS
+
+      INPUTS:
+      simplify - simplify the product of a variable and its conjugate
+           
+      OUTPUTS:
+      None 
+      '''    
       def to_latex_JH(self,simplify=False):
           file = open("JH_"+str(self.N)+".txt", "w")
           file.write("\\begin{equation}\n")
@@ -2331,23 +2354,15 @@ EXAMPLE OF HOW TO USE THE CODE TO CREATE ANALYTIC EXPRESSION OF JHv
 '''
 def JHv_example():
     r = redundant()
-    r.create_redundant(layout="HEX",order=1,print_pq=False) 
+    r.create_redundant(layout="REG",order=5,print_pq=True) 
     r.create_J1()
[…]

### 2016-08-23 software/LINCAL_test.py
Redundant StefCAL derivation

@@ -2,6 +2,10 @@ import numpy as np
 import scipy as sp
 import pylab as plt
 import simulator
+import analytic
+
+
+
 
 def LINCAL_J(z,N,L,phi):
     g = z[:N]
@@ -43,6 +47,23 @@ def LINCAL_J(z,N,L,phi):
     J = np.vstack([J,np.conj(J)])     
 
     return J
+
+def compute_v(z,N,phi):
+    g = z[:N]
+    y = z[N:]
+
+    rows = (N**2 - N)
+              
+    v = np.zeros((rows,),dtype=complex)
+
+    counter = 0
+    for k in xrange(N):
+        for j in xrange(k+1,N):
+              phi_v = phi[k,j]
+              v[counter] = g[k]*np.conjugate(g[j])*y[phi_v-1]
+              counter = counter + 1
+    v[counter:] = np.conjugate(v[:counter]) 
+    return v
 
 def LINCAL_H_analytic(z,N,L,zeta):
     g = z[:N]
@@ -188,8 +209,84 @@ def create_index_ij(zeta,N,i,j):
 
     return p,q 
   
+'''
+Creates the redundant complex Jacobian analytically. Returns the result
+      
+RETURNS:
+J - analytic Jacobian 
+ 
+INPUTS:
+dim - dimesnion of array
+layout - which layout ot choose     
+'''
+def construct_sym_J_layout(dim=5,layout="REG"):
+    r = analytic.redundant()
+    r.create_redundant(layout=layout,order=dim,print_pq=False)
+    r.create_J1()
+    r.create_J2()
+    r.conjugate_J1_J2()
+    r.create_J()
+    #print "self.J_sym.size = ",r.J.shape
+    return r.J,r.N,r.L
+          
+'''
+Substitutes the vector z into the symbolic redundant Jacobian. Stores the result in self.J
+      
+RETURNS:
+J - the numerical Jacobian
+ 
+INPUTS:
+z - Input vector.
+N - Number of elements in array.
+plot - Plot the resultant matrices if True.     
+'''
+def substitute_sym_J(z,J_sym,dim=5,layout="REG",plot=True):
+    r = analytic.redundant()
+    r.create_redundant(layout=layout,order=dim,print_pq=False)
+    g = z[:r.N]
+    y = z[r.N:]
+    r.J = J_sym
[…]

### 2016-08-23 software/analytic.py

@@ -73,6 +73,8 @@ class factor():
               string_out = self.type_factor + "_{" + str(self.ant_p)+str(self.ant_q)+"}"
            elif self.type_factor == "c":
               string_out = "("+str(self.value)+")" + "_{" + str(self.ant_p)+str(self.ant_q)+"}"
+           elif self.type_factor == "b":
+              string_out = self.type_factor + "_{" + str(self.ant_p)+str(self.ant_q)+"}"
            else:
               string_out = self.type_factor + "_" + str(self.index)
         else:   
@@ -350,6 +352,9 @@ class term():
 
           for k in xrange(len(self.constant_array)):
               self.constant_array[k].conjugate() 
+
+          for k in xrange(len(self.b_array)):
+              self.b_array[k].conjugate()
 
       '''
       Calculates the total constant factor multiplies all the factors in the term together
@@ -805,6 +810,10 @@ class redundant():
           self.H = np.array([],dtype=object) #Hessian matrix
           self.v = np.array([],dtype=object) #Model array - going to use it to calculate J^Hv - trying to derive redundant SteFCal.
           self.JHv = np.array([],dtype=object) #J^H times v
+          self.z = np.array([],dtype=object) #The parameter vector
+          self.d = np.array([],dtype=object) #The parameter vector
+          self.JHd = np.array([],dtype=object) #The parameter vector
+          self.Jz = np.array([],dtype=object) #The Jacobian times the parameter vector
 
 #################################################################################
 # CREATING MODEL, JOCOBIAN AND HESSIAN
@@ -868,15 +877,96 @@ class redundant():
                   t_temp.append_factor(g_fact)
                   t_temp.append_factor(y_fact)
                   t_temp.append_factor(gc_fact)
-                  self.v = np.append(self.regular_array,t_temp)
+                  self.v = np.append(self.v,t_temp)
           v_copy = deepcopy(self.v)
 
           for k in xrange(len(v_copy)):
-              v_copy[k].conjugate()      
+              v_copy[k].conjugate() 
+              self.v = np.append(self.v,v_copy[k]) #USING ALTERNATIVE APPROACH     
 
-          self.v = np.vstack([deepcopy(self.v),v_copy])
+          #self.v = np.vstack([deepcopy(self.v),v_copy]) - NOT WORKING NOT SURE WHY?
 
[…]

### 2016-08-24 software/analytic.py
deriving the final part of the derivations

@@ -900,7 +900,7 @@ class redundant():
       OUTPUTS:
       None 
       '''
-      def create_d(self,layout="REG",order=5,print_pq=False):
+      def create_d(self,layout="REG",order=5,print_pq=True):
           s = simulator.sim(layout=layout,order=order)
           s.generate_antenna_layout()
           phi,zeta = s.calculate_phi(s.ant[:,0],s.ant[:,1])
@@ -911,13 +911,13 @@ class redundant():
           
           for p in xrange(1,self.N):
               for q in xrange(p+1,self.N+1):
-                  b_fact = factor("b",p,1,False,False,p,q,False) 
+                  b_fact = factor("b",p,1,False,False,p,q,print_f=print_pq) 
                   t_temp = term()
                   t_temp.append_factor(b_fact)
-                  self.d = np.append(self.v,t_temp)
+                  self.d = np.append(self.d,t_temp)
           d_copy = deepcopy(self.d)
 
-          for k in xrange(len(v_copy)):
+          for k in xrange(len(d_copy)):
               d_copy[k].conjugate() 
               self.d = np.append(self.d,d_copy[k]) #USING ALTERNATIVE APPROACH     
 
@@ -1453,7 +1453,7 @@ class redundant():
           parameters = self.JH.shape[0]
           column = expression(self.d)
           #print "column = ",column.to_string()        
-          self.JHv = np.empty((parameters,),dtype=object)  
+          self.JHd = np.empty((parameters,),dtype=object)  
           for r in xrange(parameters):
                 row = expression(self.JH[r,:])
                 row_temp = deepcopy(row)
@@ -1731,7 +1731,7 @@ class redundant():
       output_string - output string
       '''      
       def to_string_JHd(self,simplify=False,simplify_const=False):
-          string_out = "JHv = ["
+          string_out = "JHd = ["
           for entry in self.JHd:
               string_out = string_out + entry.to_string(simplify)+","
           string_out = string_out[:-1]

### 2016-08-25 software/redundant_selfcal.py
first attempt at implementing redundant stefcal

@@ -1,0 +1,109 @@
+import numpy as np
+import scipy as sp
+import pylab as plt
+import simulator
+import analytic
+
+def redundant_stefcal(D,phi,tau=1e-6,alpha=0.333,max_itr=1000):
+    converged = False
+    N = D.shape[0]
+    D = D - D*np.eye(N)
+    L = np.amax(phi)
+
+    g_temp = np.ones((N,),dtype=complex)
+    y_temp = np.ones((L,),dtype=complex)
+    z_temp = np.hstack([g_temp,y_temp])
+
+    #EXTRACT BASELINE INDICES FOR EACH REDUNDANT SPACING
+    PQ = create_PQ(phi,L)
+    
+    for i in xrange(max_itr): #MAX NUMBER OF ITRS
+        g_old = np.copy(g_temp)
+        y_old = np.copy(y_temp)
+        z_old = np.copy(z_temp)
+ 
+        M = convert_y_to_M(PQ,y_old,N) #CONVERT y VECTOR TO M matrix
+
+        for p in xrange(N): #STEFCAL - update antenna gains
+            z = g_old*M[:,p]
+            g_temp[p] = np.sum(np.conj(D[:,p])*z)/(np.sum(np.conj(z)*z))
+        
+        for l in xrange(L): #UPDATE y
+            pq = PQ[str(l)]
+            num = 0
+            den = 0
+            for k in xrange(len(pq)): #loop through all baselines for each redundant spacing
+                p = pq[k,0]
+                q = pq[k,1]
+
+                num = num + np.conjugate(g_old[p])*g_old[q]*D[p,q]
+                den = den + np.absolute(g_old[p])**2*np.absolute(g_old[q])**2
+            y_temp[l] = num/den
+         
+        g_temp = alpha*g_temp + (1-alpha)*g_old
+        y_temp = alpha*y_temp + (1-alpha)*y_old
+        z_temp = np.hstack([g_temp,y_temp]) #final update
+
+        if (np.sqrt(np.sum(np.absolute(z_temp-z_old)**2))/np.sqrt(np.sum(np.absolute(z_temp)**2)) <= tau):
+           converged = True 
+           break
+
+    G = np.dot(np.diag(g_temp),temp)
+    G = np.dot(G,np.diag(g_temp.conj()))  
[…]

### 2016-08-25 software/redundant_selfcal.py
redundant_stefcal.py

@@ -4,18 +4,20 @@ import pylab as plt
 import simulator
 import analytic
 
-def redundant_stefcal(D,phi,tau=1e-6,alpha=0.333,max_itr=1000):
+def redundant_StEFCal(D,phi,tau=1e-3,alpha=0.3,max_itr=1000,PQ=None):
     converged = False
     N = D.shape[0]
     D = D - D*np.eye(N)
     L = np.amax(phi)
-
+    temp =np.ones((D.shape[0],D.shape[1]) ,dtype=complex)
+    
     g_temp = np.ones((N,),dtype=complex)
     y_temp = np.ones((L,),dtype=complex)
     z_temp = np.hstack([g_temp,y_temp])
 
     #EXTRACT BASELINE INDICES FOR EACH REDUNDANT SPACING
-    PQ = create_PQ(phi,L)
+    if PQ is not None:
+       PQ = create_PQ(phi,L)
     
     for i in xrange(max_itr): #MAX NUMBER OF ITRS
         g_old = np.copy(g_temp)
@@ -23,7 +25,11 @@ def redundant_stefcal(D,phi,tau=1e-6,alpha=0.333,max_itr=1000):
         z_old = np.copy(z_temp)
  
         M = convert_y_to_M(PQ,y_old,N) #CONVERT y VECTOR TO M matrix
-
+    
+        #plt.imshow(np.absolute(M))
+        #plt.show()
+        #from IPython import embed; embed()        
+        
         for p in xrange(N): #STEFCAL - update antenna gains
             z = g_old*M[:,p]
             g_temp[p] = np.sum(np.conj(D[:,p])*z)/(np.sum(np.conj(z)*z))
@@ -33,8 +39,8 @@ def redundant_stefcal(D,phi,tau=1e-6,alpha=0.333,max_itr=1000):
             num = 0
             den = 0
             for k in xrange(len(pq)): #loop through all baselines for each redundant spacing
-                p = pq[k,0]
-                q = pq[k,1]
+                p = pq[k][0]
+                q = pq[k][1]
 
                 num = num + np.conjugate(g_old[p])*g_old[q]*D[p,q]
                 den = den + np.absolute(g_old[p])**2*np.absolute(g_old[q])**2
[…]

### 2016-08-30 software/simulator.py
started implementing sparc

@@ -793,6 +793,48 @@ class sim():
                  return q,True
 
           return 0,False
+
+      '''
+      Returns the p if the redundant index and q are given
+      INPUTS:
+      i - q
+      j - redundant index
+      phi - mapping from pq to redundant index
+      
+      RETURNS:
+      p - p
+      Found - If p was found in phi?
+      '''
+      def xi_func_eval(self,i,j,phi):
+          column = phi[:,i]
+
+          #print "column = ",column
+          #print "j = ",j
+
+          for p in xrange(len(column)):
+              if column[p] == j:
+                 return p,True
+
+          return 0,False
+
+
+      '''
+      Returns the the dictionary PQ containing all the pq index sets associated with each redundant spacing
+      INPUTS:
+      phi - mapping from pq to redundant index
+      L - maximum entry in phi
+
+      RETURNS:
+      PQ - dictionary of pq index sets associated with each redundant spacing      
+     '''
+      def create_PQ(self,phi,L):
+          PQ = {}
+          for i in xrange(L):
+              pq = zip(*np.where(phi == (i+1)))
+              #print "pq2 = ",pq        
+              PQ[str(i)]=pq
+              #print "PQ = ",PQ    
+          return PQ
           
 """
 Function to verify the analytic expression for calculating the number of redundancies from the number of antennas in an HEXAGONAL 

### 2016-08-30 software/sparc.py

@@ -10,32 +10,198 @@ class SPARC():
 
           self.N = N
           self.L = L
+          self.B = (N**2-N)/2
           self.phi = phi
           self.PQ = PQ
  
           self.v = np.array([])
           self.r = np.array([])
 
-      def compute_v(self,z,phi):
-          g = z[:N] 
-          y = z[N:]
+      def compute_v(self,z):
+          g = z[:self.N] 
+          y = z[self.N:]
+           
+         
+          v = np.zeros((self.B,),dtype=complex)
 
           counter = 0
           for p in xrange(1,self.N):
               for q in xrange(p+1,self.N+1):
-                  v[counter] = g[p-1]*np.conjugate(g[q-1])*y[phi[p-1,q-1]-1]
-
-          v = np.vstack([v,v.conjugate()]]
+                  v[counter] = g[p-1]*np.conjugate(g[q-1])*y[self.phi[p-1,q-1]-1]
+                  counter = counter + 1
+          v = np.hstack([v,v.conjugate()])
         
           return v
 
-   
+      def vectorize_D(self,D):
+          d = np.zeros((self.B,),dtype=complex)  
+          
+          counter = 0
+          for p in xrange(1,self.N):
+              for q in xrange(p+1,self.N+1):
+                  d[counter] = D[p-1,q-1]
+                  counter = counter + 1
+          d = np.hstack([d,d.conjugate()])
+        
+          return d
+
+      def compute_r(self,d,v):
+          r = d-v
+          return r
+
+      def compute_J(self,z):
+          g = z[:self.N]
+          y = z[self.N:]
+
+          rows = (self.N**2 - self.N)/2
+          columns = self.N + self.L
+
+          J1 = np.zeros((rows,columns),dtype=complex)
+          J2 = np.zeros((rows,columns),dtype=complex)
+
+          J = np.zeros((2*rows,2*columns),dtype=complex)
+          
+          p_v = np.zeros((rows,),dtype=int)
+          q_v = np.zeros((rows,),dtype=int)
+          
+          counter = 0
+          for k in xrange(self.N):
+              for j in xrange(k+1,self.N):
+                  p_v[counter] = k
+                  q_v[counter] = j
+                  counter = counter + 1
+          
+          for r in xrange(rows):
+              p = p_v[r]
[…]

### 2016-08-31 software/simulator.py
fixed JHr bug

@@ -835,6 +835,33 @@ class sim():
               PQ[str(i)]=pq
               #print "PQ = ",PQ    
           return PQ
+
+      '''
+      Takes the y-vector and constructs an NxN matrix M
+
+      INPUTS:
+      PQ - dictionary of pq index sets associated with each redundant spacing 
+      y - vector from which M is constructed
+      N - number of antennas
+
+      OUTPUTS:
+      M - NxN matrix from y
+      '''
+      def convert_y_to_M(self,PQ,y,N):
+    
+          M = np.zeros((N,N),dtype=complex)
+          for i in xrange(len(y)):
+              pq = PQ[str(i)]
+              #print "pq = ",pq
+                          
+          for k in xrange(len(pq)):
+              p = pq[k][0]
+              q = pq[k][1]
+             
+              M[p,q] = y[i]
+              M[q,p] = np.conjugate(y[i])  
+              #from IPython import embed; embed() 
+          return M
           
 """
 Function to verify the analytic expression for calculating the number of redundancies from the number of antennas in an HEXAGONAL 

### 2016-08-31 software/sparc.py

@@ -3,15 +3,17 @@ import scipy as sp
 import pylab as plt
 import simulator
 import analytic
+from scipy.sparse import csr_matrix
 
 class SPARC():
 
-      def __init__(self,N,L,phi,PQ):
+      def __init__(self,N,L,phi,zeta,PQ):
 
           self.N = N
           self.L = L
           self.B = (N**2-N)/2
           self.phi = phi
+          self.zeta = zeta
           self.PQ = PQ
  
           self.v = np.array([])
@@ -48,6 +50,17 @@ class SPARC():
       def compute_r(self,d,v):
           r = d-v
           return r
+
+      def construct_R(self,r):
+          r_new = r[:self.B]
+          counter = 0
+          R = np.zeros((self.N,self.N),dtype=complex)
+          for k in xrange(self.N):
+              for j in xrange(k+1,self.N):
+                  R[k,j] = r_new[counter]
+                  R[j,k] = np.conjugate(r_new[counter])
+                  counter = counter + 1
+          return R 
 
       def compute_J(self,z):
           g = z[:self.N]
@@ -98,6 +111,41 @@ class SPARC():
       def compute_JHJ(self,J):
           return np.dot(J.transpose().conjugate(),J)
 
+      def compute_JHr(self,J,r):
+          return np.dot(J.transpose().conjugate(),r)
+
+      def compute_JHr_formula(self,z,D):
+          
+          g = z[:self.N]
+          y = z[self.N:]
+          
+          vec1 = np.zeros((self.N,),dtype=complex)
+          for i in xrange(len(vec1)):
+              sum_v = 0j
+              for k in xrange(len(vec1)):
+                  if k <> i:
+                     if k > i:
+                        x_ki = np.conjugate(y[self.zeta[i,k]-1])  
+                     else:
+                        x_ki = y[self.zeta[i,k]-1]
+                     sum_v = sum_v + g[k]*x_ki*D[i,k]
+              vec1[i] = sum_v
+
+          vec2 = np.zeros((self.L,),dtype=complex)
+          
+          for i in xrange(len(vec2)):
[…]

### 2016-08-31 software/sparc.py
added R

@@ -114,7 +114,7 @@ class SPARC():
       def compute_JHr(self,J,r):
           return np.dot(J.transpose().conjugate(),r)
 
-      def compute_JHr_formula(self,z,D):
+      def compute_JHr_formula(self,z,R):
           
           g = z[:self.N]
           y = z[self.N:]
@@ -128,7 +128,7 @@ class SPARC():
                         x_ki = np.conjugate(y[self.zeta[i,k]-1])  
                      else:
                         x_ki = y[self.zeta[i,k]-1]
-                     sum_v = sum_v + g[k]*x_ki*D[i,k]
+                     sum_v = sum_v + g[k]*x_ki*R[i,k]
               vec1[i] = sum_v
 
           vec2 = np.zeros((self.L,),dtype=complex)
@@ -140,7 +140,7 @@ class SPARC():
                   p = pq[k][0]
                   q = pq[k][1]
                   
-                  sum_v = sum_v + np.conjugate(g[p])*g[q]*D[p,q]
+                  sum_v = sum_v + np.conjugate(g[p])*g[q]*R[p,q]
               vec2[i] = sum_v
 
           return np.hstack([vec1,vec2,np.conjugate(vec1),np.conjugate(vec2)]) 

### 2016-09-01 software/simulator.py
buggy version

@@ -854,13 +854,13 @@ class sim():
               pq = PQ[str(i)]
               #print "pq = ",pq
                           
-          for k in xrange(len(pq)):
-              p = pq[k][0]
-              q = pq[k][1]
+              for k in xrange(len(pq)):
+                  p = pq[k][0]
+                  q = pq[k][1]
              
-              M[p,q] = y[i]
-              M[q,p] = np.conjugate(y[i])  
-              #from IPython import embed; embed() 
+                  M[p,q] = y[i]
+                  M[q,p] = np.conjugate(y[i])  
+                  #from IPython import embed; embed() 
           return M
           
 """
@@ -946,7 +946,7 @@ def example_usage():
     s.uv_tracks() #GENERATE UV TRACKS
     s.plot_uv_coverage(title="HEX") #PLOT THE UV TRACKS
     point_sources = s.create_point_sources(100,fov=3,a=2) #GENERATE RANDOM SKYMODEL
-    g=s.create_antenna_gains(s.N,0.9,0.1,50,1,5,s.nsteps,plot = True) #GENERATE GAINS
+    g=s.create_antenna_gains(s.N,0.99,0.95,2,1,5,s.nsteps,plot = True) #GENERATE GAINS
     D,sig = s.create_vis_mat(point_sources,s.u_m,s.v_m,g=g,SNR=20,w_m=None) #CREATE VIS MATRIX
     M,sig = s.create_vis_mat(point_sources,s.u_m,s.v_m,g=None,SNR=None,w_m=None) #PREDICTED VIS
     s.plot_visibilities([0,1],D,"b",s=False) #PLOT VIS

### 2016-09-01 software/sparc.py

@@ -4,6 +4,9 @@ import pylab as plt
 import simulator
 import analytic
 from scipy.sparse import csr_matrix
+from scipy.sparse.linalg import cg
+from scipy.linalg import pinv
+from scipy.sparse import dia_matrix
 
 class SPARC():
 
@@ -15,9 +18,8 @@ class SPARC():
           self.phi = phi
           self.zeta = zeta
           self.PQ = PQ
- 
-          self.v = np.array([])
-          self.r = np.array([])
+          self.itr = 0
+          self.itr_vec = np.array([],dtype=int)
 
       def compute_v(self,z):
           g = z[:self.N] 
@@ -145,13 +147,109 @@ class SPARC():
 
           return np.hstack([vec1,vec2,np.conjugate(vec1),np.conjugate(vec2)]) 
 
-             
+      def compute_inverse_cg(self,A,b,M=None,tol=1e-6):
+          self.itr = 0
+          x,info = cg(A=A,b=b,tol=tol,M=M,callback=self.report)
+          self.itr_vec = np.append(self.itr_vec,np.array([self.itr]))
+          return x,info
 
+      def report(self,xk):
+          self.itr = self.itr + 1
+
+      def compute_update_step_cg(self,d,r,z,psi_func_eval,xi_func_eval,lam=0,method="PCG",tol=1e-6):
+          R = sparc_object.construct_R(r)
+
+          H,H_sparse,D_l,D_sparse = self.generate_H_formula(z,psi_func_eval,xi_func_eval,lam=lam)
+          JHr_fun = self.compute_JHr_formula(z,R) #CAN STILL INCREASE THE SPEED BY DIRECTLY COMPUTING R form D and V
+          
+          if method == "PCG":
+             dz,info = sparc_object.compute_inverse_cg(A=H_sparse,b=JHr_fun,M=D_sparse,tol=tol)
+          else:
+             dz,info = sparc_object.compute_inverse_cg(A=H_sparse,b=JHr_fun,M=None,tol=tol)
+          return dz[:len(dz)/2],info
+
+
+      def levenberg_marquardt(self,D,psi_func_eval,xi_func_eval,convert_y_to_M,tol1=1e-6,tol2=1e-6,tol3=1e-6,lam=2,max_itr=2000,method="PCG"):
+
+          K=10
+
[…]

### 2016-09-19 software/plot_func.py
plot the function

@@ -241,12 +241,20 @@ def plot_outer_loop(SNR=10,k_upper1=5,k_upper2=5):
     
     for n in xrange(len(N)):
         outerloop_pcg_vec = outerloop_pcg_dic[str(N[n])]
+        
+        print "N = ",N[n]
+        print "len(outerloop_pcg_vec) = ",len(outerloop_pcg_vec[outerloop_pcg_vec<=9999])
+                  
         if len(outerloop_pcg_vec) == 0:
            break
         outerloop_pcg_mean[n] = np.median(outerloop_pcg_vec[outerloop_pcg_vec<=9999])
         outerloop_stef_vec = outerloop_stef_dic[str(N[n])]
+
+        print "len(outerloop_stef_vec) = ",len(outerloop_stef_vec[outerloop_stef_vec<=9999])         
+
         if len(outerloop_stef_vec) == 0:
            break
+        
         outerloop_stef_mean[n] = np.median(outerloop_stef_vec[outerloop_stef_vec<=9999])
         
         outerloop_pcg_std[n] = np.median(np.absolute(outerloop_pcg_vec[outerloop_pcg_vec<=9999] - np.median(outerloop_pcg_vec[outerloop_pcg_vec<=9999])))
@@ -453,6 +461,6 @@ def plot_sparsity():
 
 if __name__ == "__main__":
    #plot_kappa_itr(SNR=5)
-   plot_outer_loop(SNR=5)
+   plot_outer_loop(SNR=1000)
    #plot_time()
    #plot_sparsity()

### 2016-09-19 software/plot_func.py
plot the function

@@ -461,6 +461,6 @@ def plot_sparsity():
 
 if __name__ == "__main__":
    #plot_kappa_itr(SNR=5)
-   plot_outer_loop(SNR=1000)
+   plot_outer_loop(SNR=5)
    #plot_time()
    #plot_sparsity()

### 2016-09-19 software/plot_func.py
plot the function

@@ -250,7 +250,11 @@ def plot_outer_loop(SNR=10,k_upper1=5,k_upper2=5):
         outerloop_pcg_mean[n] = np.median(outerloop_pcg_vec[outerloop_pcg_vec<=9999])
         outerloop_stef_vec = outerloop_stef_dic[str(N[n])]
 
-        print "len(outerloop_stef_vec) = ",len(outerloop_stef_vec[outerloop_stef_vec<=9999])         
+        print "len(outerloop_stef_vec) = ",len(outerloop_stef_vec[outerloop_stef_vec<=9999])
+        temp_v = outerloop_stef_vec[outerloop_stef_vec<=9999]
+        temp_v = outerloop_stef_vec[temp_v>9000]
+        print "len(temp_v) = ",len(temp_v)              
+   
 
         if len(outerloop_stef_vec) == 0:
            break

### 2016-09-19 software/plot_func.py
plot the function

@@ -260,9 +260,12 @@ def plot_outer_loop(SNR=10,k_upper1=5,k_upper2=5):
            break
         
         outerloop_stef_mean[n] = np.median(outerloop_stef_vec[outerloop_stef_vec<=9999])
-        
+
+        print "outerloop_stef_mean = ",outerloop_stef_mean
+                
         outerloop_pcg_std[n] = np.median(np.absolute(outerloop_pcg_vec[outerloop_pcg_vec<=9999] - np.median(outerloop_pcg_vec[outerloop_pcg_vec<=9999])))
         outerloop_stef_std[n] = np.median(np.absolute(outerloop_stef_vec[outerloop_stef_vec<=9999] - np.median(outerloop_stef_vec[outerloop_stef_vec<=9999])))
+        print "outerloop_stef_std = ",outerloop_stef_std 
     
 
     #print "kappa_cg_mad = ",kappa_cg_mad

### 2016-09-19 software/plot_func.py
plot the function

@@ -243,15 +243,15 @@ def plot_outer_loop(SNR=10,k_upper1=5,k_upper2=5):
         outerloop_pcg_vec = outerloop_pcg_dic[str(N[n])]
         
         print "N = ",N[n]
-        print "len(outerloop_pcg_vec) = ",len(outerloop_pcg_vec[outerloop_pcg_vec<=9999])
+        print "len(outerloop_pcg_vec) = ",len(outerloop_pcg_vec[outerloop_pcg_vec<=9998])
                   
         if len(outerloop_pcg_vec) == 0:
-           break
-        outerloop_pcg_mean[n] = np.median(outerloop_pcg_vec[outerloop_pcg_vec<=9999])
+           break8
+        outerloop_pcg_mean[n] = np.median(outerloop_pcg_vec[outerloop_pcg_vec<=9998])
         outerloop_stef_vec = outerloop_stef_dic[str(N[n])]
 
-        print "len(outerloop_stef_vec) = ",len(outerloop_stef_vec[outerloop_stef_vec<=9999])
-        temp_v = outerloop_stef_vec[outerloop_stef_vec<=9999]
+        print "len(outerloop_stef_vec) = ",len(outerloop_stef_vec[outerloop_stef_vec<=9998])
+        temp_v = outerloop_stef_vec[outerloop_stef_vec<=9998]
         temp_v = outerloop_stef_vec[temp_v>9000]
         print "len(temp_v) = ",len(temp_v)              
    
@@ -259,12 +259,12 @@ def plot_outer_loop(SNR=10,k_upper1=5,k_upper2=5):
         if len(outerloop_stef_vec) == 0:
            break
         
-        outerloop_stef_mean[n] = np.median(outerloop_stef_vec[outerloop_stef_vec<=9999])
+        outerloop_stef_mean[n] = np.median(outerloop_stef_vec[outerloop_stef_vec<=9998])
 
         print "outerloop_stef_mean = ",outerloop_stef_mean
                 
-        outerloop_pcg_std[n] = np.median(np.absolute(outerloop_pcg_vec[outerloop_pcg_vec<=9999] - np.median(outerloop_pcg_vec[outerloop_pcg_vec<=9999])))
-        outerloop_stef_std[n] = np.median(np.absolute(outerloop_stef_vec[outerloop_stef_vec<=9999] - np.median(outerloop_stef_vec[outerloop_stef_vec<=9999])))
[…]

### 2016-09-19 software/plot_func.py
plot the function

@@ -243,14 +243,15 @@ def plot_outer_loop(SNR=10,k_upper1=5,k_upper2=5):
         outerloop_pcg_vec = outerloop_pcg_dic[str(N[n])]
         
         print "N = ",N[n]
-        print "len(outerloop_pcg_vec) = ",len(outerloop_pcg_vec[outerloop_pcg_vec<=9998])
+        print "len(outerloop_pcg_vec) = ",len(outerloop_pcg_vec)
                   
         if len(outerloop_pcg_vec) == 0:
-           break8
+           break
+
         outerloop_pcg_mean[n] = np.median(outerloop_pcg_vec[outerloop_pcg_vec<=9998])
         outerloop_stef_vec = outerloop_stef_dic[str(N[n])]
 
-        print "len(outerloop_stef_vec) = ",len(outerloop_stef_vec[outerloop_stef_vec<=9998])
+        print "len(outerloop_stef_vec) = ",len(outerloop_stef_vec)
         temp_v = outerloop_stef_vec[outerloop_stef_vec<=9998]
         temp_v = outerloop_stef_vec[temp_v>9000]
         print "len(temp_v) = ",len(temp_v)              

### 2016-09-19 software/plot_func.py
plot the function

@@ -216,6 +216,7 @@ def plot_outer_loop(SNR=10,k_upper1=5,k_upper2=5):
             c_cal = pickle.load(file_p)
             time_mat = pickle.load(file_p)
             outer_loop = pickle.load(file_p)
+            print "len(outer_loop) = ",len(outer_loop)
             outerloop_pcg_dic[str(N[i])] = np.append(outerloop_pcg_dic[str(N[i])],outer_loop)
             file_p.close()
     
@@ -236,6 +237,7 @@ def plot_outer_loop(SNR=10,k_upper1=5,k_upper2=5):
             c_cal = pickle.load(file_p)
             time_mat = pickle.load(file_p)
             outer_loop = pickle.load(file_p)
+            print "len(outer_loop) = ",len(outer_loop) 
             outerloop_stef_dic[str(N[i])] = np.append(outerloop_stef_dic[str(N[i])],outer_loop)
             file_p.close()
     
@@ -469,6 +471,6 @@ def plot_sparsity():
 
 if __name__ == "__main__":
    #plot_kappa_itr(SNR=5)
-   plot_outer_loop(SNR=1000)
+   plot_outer_loop(SNR=5)
    #plot_time()
    #plot_sparsity()

### 2016-09-19 software/plot_func.py
plot the function

@@ -218,6 +218,7 @@ def plot_outer_loop(SNR=10,k_upper1=5,k_upper2=5):
             outer_loop = pickle.load(file_p)
             print "len(outer_loop) = ",len(outer_loop)
             outerloop_pcg_dic[str(N[i])] = np.append(outerloop_pcg_dic[str(N[i])],outer_loop)
+            print "len(outerloop_pcg_dic[str(N[i])]) = ",len(outerloop_pcg_dic[str(N[i])])
             file_p.close()
     
     for k in xrange(k_upper2):

### 2016-09-19 software/plot_func.py
plot the function

@@ -472,6 +472,6 @@ def plot_sparsity():
 
 if __name__ == "__main__":
    #plot_kappa_itr(SNR=5)
-   plot_outer_loop(SNR=5)
+   plot_outer_loop(SNR=1000)
    #plot_time()
    #plot_sparsity()

### 2016-09-19 software/plot_func.py
plot the function

@@ -472,6 +472,6 @@ def plot_sparsity():
 
 if __name__ == "__main__":
    #plot_kappa_itr(SNR=5)
-   plot_outer_loop(SNR=1000)
+   plot_outer_loop(SNR=5)
    #plot_time()
    #plot_sparsity()

### 2016-09-19 software/plot_func.py
plot the function

@@ -472,6 +472,6 @@ def plot_sparsity():
 
 if __name__ == "__main__":
    #plot_kappa_itr(SNR=5)
-   plot_outer_loop(SNR=5)
+   plot_outer_loop(SNR=10)
    #plot_time()
    #plot_sparsity()

### 2016-09-19 software/plot_func.py
plot the function

@@ -472,6 +472,6 @@ def plot_sparsity():
 
 if __name__ == "__main__":
    #plot_kappa_itr(SNR=5)
-   plot_outer_loop(SNR=10)
+   plot_outer_loop(SNR=1000)
    #plot_time()
    #plot_sparsity()

### 2016-09-19 software/plot_func.py
trying time plot

@@ -289,11 +289,14 @@ def plot_outer_loop(SNR=10,k_upper1=5,k_upper2=5):
     ax.legend(loc=5)
     plt.show()
 
-def plot_time(SNR=1000,k_upper=1):
+def plot_time(SNR=1000,k_upper1=5,k_upper2=5,k_upper3=5):
     
-    N = np.array([7,19,37])
-    L = np.array([9,30,63])
+    N = np.array([7,19,37,61,91,108,169,217])
+    L = np.array([9,30,63,108,165,234,315,408])
+    
     P = 2*(N+L)
+ 
+     
 
     time_pcg_mean = np.zeros((len(N),),dtype=float)
     time_stef_mean = np.zeros((len(N),),dtype=float)
@@ -312,15 +315,11 @@ def plot_time(SNR=1000,k_upper=1):
         time_stef_dic[str(N[j])]=np.array([],dtype=float)
         time_svd_dic[str(N[j])]=np.array([],dtype=float)
 
-    for k in xrange(k_upper):
-        PCG_DIR = "./HEX_PCG_"+str(SNR)+"_True_"+str(k)
-        PCG_FILE_LIST = np.array([PCG_DIR+"/1_7_9_"+PCG_DIR[2:]+".p",PCG_DIR+"/2_19_30_"+PCG_DIR[2:]+".p",PCG_DIR+"/3_37_63_"+PCG_DIR[2:]+".p"]) 
-        #PCG_FILE_LIST = np.array([PCG_DIR+"/1_7_9_"+PCG_DIR[2:]+".p",PCG_DIR+"/2_19_30_"+PCG_DIR[2:]+".p"]) 
-        STEF_DIR = "./HEX_R_StEFCal_"+str(SNR)+"_"+str(k)
-        STEF_FILE_LIST = np.array([STEF_DIR+"/1_7_9_"+STEF_DIR[2:]+".p",STEF_DIR+"/2_19_30_"+STEF_DIR[2:]+".p",STEF_DIR+"/3_37_63_"+STEF_DIR[2:]+".p"])
-        #STEF_FILE_LIST = np.array([STEF_DIR+"/1_7_9_"+STEF_DIR[2:]+".p",STEF_DIR+"/2_19_30_"+STEF_DIR[2:]+".p"])
-        SVD_DIR = "./HEX_SVD_"+str(SNR)+"_True_"+str(k)
[…]

### 2016-09-19 software/plot_func.py
trying time plot

@@ -289,7 +289,7 @@ def plot_outer_loop(SNR=10,k_upper1=5,k_upper2=5):
     ax.legend(loc=5)
     plt.show()
 
-def plot_time(SNR=1000,k_upper1=5,k_upper2=5,k_upper3=5):
+def plot_time(SNR=5,k_upper1=5,k_upper2=5,k_upper3=5):
     
     N = np.array([7,19,37,61,91,108,169,217])
     L = np.array([9,30,63,108,165,234,315,408])

### 2016-09-20 software/plot_func.py
newest run

@@ -289,7 +289,7 @@ def plot_outer_loop(SNR=10,k_upper1=5,k_upper2=5):
     ax.legend(loc=5)
     plt.show()
 
-def plot_time(SNR=5,k_upper1=5,k_upper2=5,k_upper3=5):
+def plot_time(SNR=1000,k_upper1=5,k_upper2=5,k_upper3=5):
     
     N = np.array([7,19,37,61,91,108,169,217])
     L = np.array([9,30,63,108,165,234,315,408])

### 2016-09-20 software/plot_func.py
up to date

@@ -487,5 +487,5 @@ def plot_sparsity():
 if __name__ == "__main__":
    #plot_kappa_itr(SNR=5)
    #plot_outer_loop(SNR=5)
-   plot_time()
-   #plot_sparsity()
+   #plot_time()
+   plot_sparsity()

### 2016-09-21 paper.tex
added novelty discussion to paper

@@
 \section{Introduction}
+
+Main contributions of the paper:
+
+\begin{enumerate}
+ \item Present a general framework which unifies all the techniques developed so far showing they are all related (LINCAL, non-linear estimator, etc ...). They are all non-linear
+ least-squares techniques, i.e. they either employ Gauss-Newton or the Levenberg-Marquardt algorithms. We have to start with LINCAL showing that it is GN, then we have to motivate 
+ why we want to use Oleg's complex optimization framework.
+ \item Use Oleg's complex optimization framework to re-derive the non-linear technique proposed by Marthi and Chengular. The novelty lies, in the fact that by using Oleg's
+ framework we can find analytic expressions for the Jacobian, the Hessian and the Jacobian-residual product (which is not even the case for LINCAL). State that the algorithm is 
+ effectively related to SteFCal and is eff an independent rediscovery of SteFCal and an extension of it into the redundant domain.
+ \item Also we present the array geometry function to help us make the derivations from Marthi and Chengular easier to read and understand for a general layout.
+ \item We also mention at this point that in Oleg's paper the question is raised is there a fast way of computing the exact inverse, we then present this new method, which is called
+ conjugate gradient method. We discuss the algorithm and the two things which bound its execution time. Which is $kappa$ and $m$ (spectral condition number and its sparsity). Then
+ we explore both of these parameters.
+\end{enumerate}
 

### 2016-09-21 paper.tex
added novelty discussion to paper

@@
  we explore both of these parameters.
+
+ **************
+ FLOW OF PAPER
+ **************
+
+ \begin{enumerate}
+ \item We need the definition of visibilities as in Liu.
+ \item Introduce the array geometery function.
+ \item Write down logcal and lincal in matrix form... ?
+ \item Short discussion of least squares and jacobian and hessian's. General GN and LM update rules.
+ \item Introduce redundant calibration as a least squares problem.
+ \item Propose a possible solution witch leads to lincal. Mention that this approach works as in this form the function is differentiable (a taylor expansion in the 
+ parameters exists). Maybe mention that we can also divide the problem into real and imaginary etc...
+ \item Show how this relates to lincal for example ---> show lincal is GN.
+ \item Now introduce complex optimization ---> Main motivation for switching to the alternative framework is that the the differentials are very simplistic. We wish to show that
+ we can derive the method of Chengular.
+ \end{enumerate}
+
+ We
+
+ We neeed LOGCAL and LINCAL in matrix forms
+
 \end{enumerate}

### 2016-09-21 paper.tex
discussing the flow of the paper a bit

@@
  \item Short discussion of least squares and jacobian and hessian's. General GN and LM update rules.
- \item Introduce redundant calibration as a least squares problem.
+ \item Introduce redundant calibration as a least squares problem. We will use this general fact to derive both popular methods.
  \item Propose a possible solution witch leads to lincal. Mention that this approach works as in this form the function is differentiable (a taylor expansion in the 
@@
  we can derive the method of Chengular.
+ \item Do the derivation of Chengular. 
+ \item Brief discussion abouth Chengular and SteFCal and Complex Optimization. Here I show that in Oleg's paper he re-derives an algorithm called SteFCal. Stefcal works
+ on the basis of alternating direction implicit method. The first signs of achieving a similar algorithm already appeared in Stefan's conference paper in which the alternating
+ idea was first proposed. Then I mention Chegular re-derives the expression by using partial derivatives and extends it to redundant. One could also have used the linear alternating
+ approach. In an attempt to merge the terminology that has independently develop in the general calibration and redundant calibration literature and to emphasize the close
+ relationship between Stefcal and the approach derived by I will use the label Redunandat SteFCal (R-Stefcal) to refer to the NLS method proposed by ... Lastly we mention that similarly
+ to how Oleg re-derived stefcal in, we have achieved the same approach.
+ \item Faster Exact inverse. A question that Oleg poses in his paper is, does there exist a faster way to take the exact inverse of JhJ? One that is almost linear, and
+ implies that we can therefore implement the full LM algorithm. The aim here is of course to reduce the number of iterations that are required to converge by using the 
+ full inverse instead which would hopefully provide enough of a speedup to compensate for the more expensive full-inverse. The algorithm we propose is the conjugate 
+ gradient method.
+ \item We give the images of the HESSIAN of both lincal and the complex method (number of terms). So we can mention that both are sparse and contain diagonal entries that 
+ are more significant than 
  \end{enumerate}

### 2016-09-21 paper.tex
some of the flow of the paper has been specified

@@
  \begin{enumerate}
- \item We need the definition of visibilities as in Liu.
+ \item We need the definition of visibilities as in Liu. DEFINE SNR HERE ALREADY together maybe with the sigma value of the noise.
  \item Introduce the array geometery function.
@@
  \item We give the images of the HESSIAN of both lincal and the complex method (number of terms). So we can mention that both are sparse and contain diagonal entries that 
- are more significant than 
+ are more significant than the off-diagonal entries. What linear inversion approach can take advantage of both these phenomenon. One such technique is 
+ cg. 
+ \item Briefly discuss CG and how its computationally bounded by its condition number and its sparsity (how does the diagonal play a role).
+ \item Simulation description
+ \item Will CG improve things?
+ \item Now first discuss the condition
+ number of the Hessian before and after pre-conditioning (pre-codnitioning can only be applied if a good inverse of a matrix is known, if it is known then it can improve 
+ the spectral condition number of a matrix. We present here the kappa and iteration number graphs for the HEXAGONAL layout. Although the
+ \item Now we discuss the sparsity results. 
+ \item Provide a table that theoretical compares R-StEFCal and SPARC.
+ \item Number of outerloop iterations. 
+ \item Timing results.
+ \item Accuracy results.
+ \item Maybe some freq simulations.
  \end{enumerate}

### 2016-09-21 paper.tex
paper.tex

@@
 %and describes the problem the authors aim to solve \citep[e.g.][]{Author2012}.
+
+\section{Visibilities}
+The observed visibility $d_{pq}$ measured by the baseline formed by antenna $p$ and $q$ is described by
+\begin{equation}
+d_{pq} = g_{p}\conj{g}_{q}y_{pq} + n_{pq},
+\end{equation}
+where $g_{p}$ and $g_{q}$ are the direction independent gains respectively associated with antenna $p$ and $q$, $y_{pq}$ denotes the true visibility that baseline $pq$ measured
+and $n_{pq}$ is the thermal noise. Conjugation is denoted by $\conj{(*)}$. During an actual observation the true value of $g_p$, $g_q$ and $y_{pq}$ are unknown and are the quantities which have 
+to be estimated.
+
+The thermal noise has zero-mean and is normally distrubuted with standard deviation  
+\begin{equation}
+
+\end{equation}
+
+and $n_{pq}$ 
+zero-mean gaussian noise representing thermal noise.
 

### 2016-09-22 paper.tex
added noise

--- 2016-09-21 paper.tex
@@
 %and describes the problem the authors aim to solve \citep[e.g.][]{Author2012}.
+
+\section{Visibilities}
+The observed visibility $d_{pq}$ measured by the baseline formed by antenna $p$ and $q$ can be described as
+
+\begin{equation}
+d_{pq} = g_{p}\conj{g}_{q}y_{pq} + n_{pq},
+\end{equation}
+where $g_{p}$ and $g_{q}$ are the direction independent gains associated with antenna $p$ and $q$, $y_{pq}$ denotes the true visibility that baseline $pq$ measured
+and $n_{pq}$ is the thermal noise component.
+Conjugation is denoted by $\conj{(*)}$.
+During an actual observation the true value of $g_p$, $g_q$ and $y_{pq}$ are unknown and are the quantities which have
+to be estimated.
+
+The thermal noise is normally distributed with a mean of zero and a standard deviation that is equal to  
+\begin{equation}
+\sigma = \frac{\sqrt{2}k_{B}T_{\textrm{sys}}}{A\eta\sqrt{\Delta \nu \tau}}, 
+\end{equation}
+where $k_B$ is Boltzmann's constant, $T_{\textrm{sys}}$ is equal to the system temperature, $A$ is the effective collecting area of each element in the array, $\eta$ is a dimensionless
+efficiency factor, $\Delta \nu$ is the observational bandwidth and $\tau$ is the integration time per visibility. The
+
+and $n_{pq}$ 
+zero-mean gaussian noise representing thermal noise.
 

### 2016-09-23 paper.tex
newest paper version
added paper

@@
 \newcommand{\bv}{\bmath{v}}
+\newcommand{\bn}{\bmath{n}}
 \newcommand{\by}{\bmath{y}}
@@
 Conjugation is denoted by $\conj{(*)}$.
-During an actual observation the true value of $g_p$, $g_q$ and $y_{pq}$ are unknown and are the quantities which have
+During the course of an actual observation the true value of $g_p$, $g_q$ and $y_{pq}$ are unknown and are the quantities which infact have
 to be estimated.
@@
 where $k_B$ is Boltzmann's constant, $T_{\textrm{sys}}$ is equal to the system temperature, $A$ is the effective collecting area of each element in the array, $\eta$ is a dimensionless
-efficiency factor, $\Delta \nu$ is the observational bandwidth and $\tau$ is the integration time per visibility. The
+efficiency factor, $\Delta \nu$ is the observational bandwidth and $\tau$ is the integration time per visibility.
+
+Eq can also be restated in vector form 
+\begin{equation}
+\bd = \bv + \bn, 
+\end{equation}
+where 
+\begin{eqnarray}
+ \bd &=& [d_{pq}] \big \} ~ [pq] = 1\cdots N ~ (p<q),\nonumber\\
+ \bv &=& [g_p y_{pq}\conj{g}_q]  \big \} ~ [pq] = 1\cdots N ~ (p<q),\nonumber\\
+ \bn &=& [n_{pq}]  \big \} ~ [pq] = 1\cdots N ~ (p<q).
+\end{eqnarray}
+
+\begin{equation}
+  \bd = [d_{pq}] \big \} ~ [pq] = 1\cdots N ~ (p<q),
+\end{equation}
+\begin{equation}
+  \bv = [g_p y_{pq}\conj{g}_q]  \big \} ~ [pq] = 1\cdots N ~ (p<q)
+\end{equation}
+and
+\begin{equation}
+  \bn = [n_{pq}]  \big \} ~ [pq] = 1\cdots N ~ (p<q).
+\end{equation}
+
+\begin{enumerate}
+ \item $\bd$ is the observed visibility vector, i.e. 
+ \begin{equation}
+  \bd = [d_{pq}] \big \} ~ [pq] = 1\cdots N ~ (p<q).
+ \end{equation}
+ The vector entry $d_{pq}$ denotes the observed visibility associated with the baseline formed by antenna $p$ and $q$, while $N$ denotes the number or antennas
+ in the array. 
+ \item $\bg = [g_1,g_2,\cdots,g_N]$ represents the instrumental complex antenna gain vector.
+ \item $\bv(\bg)$ denotes the predicted visibility vector, i.e.
+ \begin{equation}
+  \bv(\bg) = [g_p y_{pq}\conj{g}_q]  \big \} ~ [pq] = 1\cdots N ~ (p<q).
+ \end{equation}
+ In Eq... $y_{pq}$ denotes the visibility which was created from a skymodel and is associated with antenna $p$ and $q$. [Need to rw this sentence] 
+\end{enumerate}
+
+The SNR is defined as 
+\begin{equation}
+\textrm{SNR} = \frac{<\bv\odot\conj{\bv}>_{t,pq}}{<\bn\odot\conj{\bn}>_{t,pq}}, 
+\end{equation}
+where $<*>_{t,pq}$ denotes averaging over time and baselines and $\odot$ denotes the Hadamard product.
 

### 2016-09-26 software/redundant_selfcal.py
newest version

@@ -215,15 +215,17 @@ def main(argv):
     return snr,l,max_order,min_order,exp_number
 
 if __name__ == "__main__":
+   '''
    snr,l,max_order,min_order,exp_number = main(sys.argv[1:])
    do_red_cal_experiment(SNR=snr,min_order=min_order,max_order=max_order,layout=l,exp_number=exp_number)
-
    '''
+   
    s = simulator.sim(nsteps=100,layout="HEX",order=1) #INSTANTIATE OBJECT
    #s.read_antenna_layout()
    s.generate_antenna_layout() #CREATE ANTENNA LAYOUT - DEFAULT IS HEXAGONAL
    s.plot_ant(title="HEX") #PLOT THE LAYOUT
    phi,zeta = s.calculate_phi(s.ant[:,0],s.ant[:,1])
+   s.plot_zeta(zeta,1,1,10,'cubehelix')
    s.uv_tracks() #GENERATE UV TRACKS
    s.plot_uv_coverage(title="HEX") #PLOT THE UV TRACKS
    point_sources = s.create_point_sources(100,fov=3,a=2) #GENERATE RANDOM SKYMODEL
@@ -232,11 +234,11 @@ if __name__ == "__main__":
    M,sig = s.create_vis_mat(point_sources,s.u_m,s.v_m,g=g,SNR=None,w_m=None) #PREDICTED VIS
    s.plot_visibilities([0,1],D,"b",s=False) #PLOT VIS
    s.plot_visibilities([0,1],M,"r",s=True)    
-   z_cal,c_cal,G_cal,M_cal = redundant_StEFCal_time(D,phi)
+   z_cal,c_cal,G_cal,M_cal,t,count_temp = redundant_StEFCal_time(D,phi)
    s.plot_visibilities([0,1],D,"b",s=False) #PLOT VIS
    s.plot_visibilities([0,1],M,"r",s=False)    
    s.plot_visibilities([0,1],G_cal*M_cal,"g",s=True)
-   '''
+  
     
    
 

### 2016-09-26 paper.tex
added some basics

@@
 \begin{equation}
+\label{eq:vis_definition}
 d_{pq} = g_{p}\conj{g}_{q}y_{pq} + n_{pq},
@@
 
-The thermal noise is normally distributed with a mean of zero and a standard deviation that is equal to  
+The real and imaginary components of the thermal noise is normally distributed with a mean of zero and a standard deviation that is equal to [NB - are the real and imaginary parts defined with same sigma?]
 \begin{equation}
@@
 
-Eq can also be restated in vector form 
+Eq.~\eqref{eq:vis_definition} can also be expressed in the following vector form 
 \begin{equation}
@@
  \bv &=& [g_p y_{pq}\conj{g}_q]  \big \} ~ [pq] = 1\cdots N ~ (p<q),\nonumber\\
- \bn &=& [n_{pq}]  \big \} ~ [pq] = 1\cdots N ~ (p<q).
+ \bn &=& [n_{pq}]  \big \} ~ [pq] = 1\cdots N ~ (p<q).\label{eq:vec_definitions}
 \end{eqnarray}
+In Eq.~\eqref{eq:vec_definitions}, $N$ denotes the number of antennas in the array.
 
-\begin{equation}
-  \bd = [d_{pq}] \big \} ~ [pq] = 1\cdots N ~ (p<q),
-\end{equation}
-\begin{equation}
-  \bv = [g_p y_{pq}\conj{g}_q]  \big \} ~ [pq] = 1\cdots N ~ (p<q)
-\end{equation}
-and
-\begin{equation}
-  \bn = [n_{pq}]  \big \} ~ [pq] = 1\cdots N ~ (p<q).
-\end{equation}
-
-\begin{enumerate}
- \item $\bd$ is the observed visibility vector, i.e. 
- \begin{equation}
-  \bd = [d_{pq}] \big \} ~ [pq] = 1\cdots N ~ (p<q).
- \end{equation}
- The vector entry $d_{pq}$ denotes the observed visibility associated with the baseline formed by antenna $p$ and $q$, while $N$ denotes the number or antennas
- in the array. 
- \item $\bg = [g_1,g_2,\cdots,g_N]$ represents the instrumental complex antenna gain vector.
- \item $\bv(\bg)$ denotes the predicted visibility vector, i.e.
- \begin{equation}
-  \bv(\bg) = [g_p y_{pq}\conj{g}_q]  \big \} ~ [pq] = 1\cdots N ~ (p<q).
- \end{equation}
- In Eq... $y_{pq}$ denotes the visibility which was created from a skymodel and is associated with antenna $p$ and $q$. [Need to rw this sentence] 
-\end{enumerate}
-
-The SNR is defined as 
+We will be using the following SNR (signal-to-noise ratio) definition in this paper  
 \begin{equation}
@@
 \end{equation}
-where $<*>_{t,pq}$ denotes averaging over time and baselines and $\odot$ denotes the Hadamard product.
+where $<*>_{t,pq}$ denotes averaging over time and baselines and $\odot$ denotes the Hadamard product. The definition we use here is similar 
+to the SNR definitions that were used in ??,??.
 

### 2016-09-26 paper.tex
added some references
started adding geometry mapping

@@
 where $<*>_{t,pq}$ denotes averaging over time and baselines and $\odot$ denotes the Hadamard product. The definition we use here is similar 
-to the SNR definitions that were used in ??,??.
+to the SNR definitions of \citet{Liu2010,Marthi2014}.
 
-and $n_{pq}$ 
-zero-mean gaussian noise representing thermal noise.
+\subsection{Redundant Array Geometry Mapping}
+If an array is redundant then its redundant baselines sample the exact same visibilites in the $uv$-plane, i.e. if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$. We can
+therefore also use the following useful indexing strategy: we can use singular redundant baseline indexes instead of composite antenna pairs to label the true observed visibilities.
+
+Let $\phi_{ij}:\mathbb{N}^2\rightarrow\mathbb{N}^+$ be the mapping that maps antenna indices to the unique redundant baseline indices of the array. This mapping is not symmetric as 
+$\phi_{ij} = 0$ if $i>j$. For an east-west regular array we can construct a simple analytic expression for this mapping:
+\begin{equation}
+\widehat{\phi_{ij}} = 
+\begin{cases}
+j-i&\textrm{if}~i<j\\
+0&\textrm{otherwise}
+\end{cases}.
+\end{equation}
+As will become apparent later the following symmetric variant of $\phi_{ij}$ will turn out to be quite useful later on:
+\begin{equation}
+\zeta_{ij} = 
+\begin{cases}
+\phi_{ij}~\textrm{if}~i\leq j\\
+\phi_{ji}~\textrm{if}~i>j
+\end{cases}.
+\end{equation} 
+It is also crucial to be able to determine the antenna index of the antenna which was used to construct a specific redundant baseline if one is given the other antenna index and the redundant baseline index.
+This can be calculated via the following two functions:
+\begin{equation}
+\xi_{ij} = 
+\begin{cases}
+p~\textrm{if}~\exists! ~ p \in \mathbb{N} ~ s.t. ~(\phi_{pi} = j)\\
+0~\textrm{otherwise}
+\end{cases},
+\end{equation}
+and
+\begin{equation}
+\psi_{ij} = 
+\begin{cases}
+q~\textrm{if}~\exists! ~ q \in \mathbb{N} ~ s.t. ~(\phi_{iq} = j)\\
+0~\textrm{otherwise}
+\end{cases}
+\end{equation}
+We use $\xi_{ij}$ to retrieve the index of the second antenna if we know the redundant baseline index and the index of the first antenna used to construct the baseline in question, while we use $\psi_{ij}$ to return the index of the first antenna if we know the
+index of the second antenna instead. \textbf{NB::} I still need to plot $\zeta_{ij}$ for different antenna layouts.
+
+Describe complex optimization, Jacobian, Hessian etc... Benefits drawbacks compared to LINCAL... etc..
 

### 2016-09-26 paper.tex
red mapping section

@@
 
-Eq.~\eqref{eq:vis_definition} can also be expressed in the following vector form 
-\begin{equation}
-\bd = \bv + \bn, 
-\end{equation}
-where 
-\begin{eqnarray}
- \bd &=& [d_{pq}] \big \} ~ [pq] = 1\cdots N ~ (p<q),\nonumber\\
- \bv &=& [g_p y_{pq}\conj{g}_q]  \big \} ~ [pq] = 1\cdots N ~ (p<q),\nonumber\\
- \bn &=& [n_{pq}]  \big \} ~ [pq] = 1\cdots N ~ (p<q).\label{eq:vec_definitions}
-\end{eqnarray}
-In Eq.~\eqref{eq:vec_definitions}, $N$ denotes the number of antennas in the array.
-
-We will be using the following SNR (signal-to-noise ratio) definition in this paper  
-\begin{equation}
-\textrm{SNR} = \frac{<\bv\odot\conj{\bv}>_{t,pq}}{<\bn\odot\conj{\bn}>_{t,pq}}, 
-\end{equation}
-where $<*>_{t,pq}$ denotes averaging over time and baselines and $\odot$ denotes the Hadamard product. The definition we use here is similar 
-to the SNR definitions of \citet{Liu2010,Marthi2014}.
-
 \subsection{Redundant Array Geometry Mapping}
 If an array is redundant then its redundant baselines sample the exact same visibilites in the $uv$-plane, i.e. if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$. We can
-therefore also use the following useful indexing strategy: we can use singular redundant baseline indexes instead of composite antenna pairs to label the true observed visibilities.
+therefore also use the following useful indexing strategy: We can use singular redundant baseline indexes instead of composite antenna pairs to label the true observed visibilities.
 
-Let $\phi_{ij}:\mathbb{N}^2\rightarrow\mathbb{N}^+$ be the mapping that maps antenna indices to the unique redundant baseline indices of the array. This mapping is not symmetric as 
-$\phi_{ij} = 0$ if $i>j$. For an east-west regular array we can construct a simple analytic expression for this mapping:
+Let $\phi_{ij}:\mathbb{N}^2\rightarrow\mathbb{N}^+$ be the mapping that maps composite antenna pairs to the unique redundant baseline indexes of an array. This mapping is not symmetric as 
+$\phi_{ij} = 0$ if $i>j$. It is possible to construct a simple analytic expression for $\phi_{ij}$ when our array is in an east-west regular configuration:
 \begin{equation}
-\widehat{\phi_{ij}} = 
+\phi_{ij} = 
 \begin{cases}
@@
 \end{equation}
-As will become apparent later the following symmetric variant of $\phi_{ij}$ will turn out to be quite useful later on:
+It becomes increasingly more difficult to construct analytic expressions of $\phi_{ij}$ for other more complicated array layouts.
+
+We can also define the following symmetric variant of $\phi_{ij}$:
 \begin{equation}
@@
 \end{equation} 
-It is also crucial to be able to determine the antenna index of the antenna which was used to construct a specific redundant baseline if one is given the other antenna index and the redundant baseline index.
-This can be calculated via the following two functions:
+If the mapping $\phi_{ij}$ is known and one is given one of the two antenna indexes that together form a redundant baseline as well as the redundant baseline index itself then it is possible 
+to calculate the unknown antenna index. This can be calculated via the following two expressions:
 \begin{equation}
@@
 \end{equation}
-We use $\xi_{ij}$ to retrieve the index of the second antenna if we know the redundant baseline index and the index of the first antenna used to construct the baseline in question, while we use $\psi_{ij}$ to return the index of the first antenna if we know the
-index of the second antenna instead. \textbf{NB::} I still need to plot $\zeta_{ij}$ for different antenna layouts.
 
-Describe complex optimization, Jacobian, Hessian etc... Benefits drawbacks compared to LINCAL... etc..
+We use $\xi_{ij}$ to retrieve the first antenna index of the composite antenna pair associated with a particular baseline if the index of the first antenna in the composite antenna pair and the unique redundant baseline index are known, while we use $\psi_{ij}$ to obtain
+a similar result; the second antenna index of the composite antenna pair. \textbf{NB::} I still need to plot $\zeta_{ij}$ for different antenna layouts.
+
+If the array is redundant and the mapping $\phi_{ij}$ is known we can rewrite Eq.~\eqref{eq:vis_definition} as
+\begin{equation}
+\label{eq:vis_red}
+d_{pq} = g_{p}\conj{g}_{q}y_{\phi_{pq}} + n_{pq}.
+\end{equation}
+Eq.~\eqref{eq:vis_red} can also be expressed in the following vector form 
+\begin{equation}
+\bd = \bv + \bn, 
+\end{equation}
+where 
+\begin{eqnarray}
+ \bd &=& [d_{pq}] \big \} ~ [pq] = 1\cdots N ~ (p<q),\nonumber\\
+ \bv &=& [g_p y_{\phi_{pq}} \conj{g}_q]  \big \} ~ [pq] = 1\cdots N ~ (p<q),\nonumber\\
+ \bn &=& [n_{pq}]  \big \} ~ [pq] = 1\cdots N ~ (p<q).\label{eq:vec_definitions}
+\end{eqnarray}
+In Eq.~\eqref{eq:vec_definitions}, $N$ denotes the number of antennas in the array. NB --- above equation is wrong, needs to be baselines ...
+
+We will be using the following SNR (signal-to-noise ratio) definition in this paper  
+\begin{equation}
+\textrm{SNR} = \frac{<\bv\odot\conj{\bv}>_{t,pq}}{<\bn\odot\conj{\bn}>_{t,pq}}, 
+\end{equation}
+where $<*>_{t,pq}$ denotes averaging over time and baselines and $\odot$ denotes the Hadamard product. The definition we use here is similar 
+to the SNR definitions of \citet{Liu2010,Marthi2014}.
 

### 2016-09-26 software/simulator.py
added plotting of phi

@@ -304,6 +304,7 @@ class sim():
     
           plt.ylim(-1.1*m,1.1*m)
           plt.xlim(-1.1*m,1.1*m)
+          plt.tight_layout()
           plt.show() 
 
       '''
@@ -953,8 +954,34 @@ def example_usage():
     s.plot_visibilities([0,1],D,"b",s=False) #PLOT VIS
     s.plot_visibilities([0,1],M,"r",s=True)
 
+def plot_paper_layouts():
+    s = sim(layout="HEX",order=5) #INSTANTIATE OBJECT
+    s.generate_antenna_layout()
+    s.plot_ant(title="HEX")
+    phi,zeta = s.calculate_phi(s.ant[:,0],s.ant[:,1])
+    s.plot_zeta(zeta,10,10,14,'jet') 
+    print "s.L = ",s.L
+    print "s.N = ",s.N
+
+    s = sim(layout="SQR",order=10) #INSTANTIATE OBJECT
+    s.generate_antenna_layout()
+    s.plot_ant(title="SQR")
+    phi,zeta = s.calculate_phi(s.ant[:,0],s.ant[:,1])
+    s.plot_zeta(zeta,10,10,14,'jet')
+    print "s.L = ",s.L
+    print "s.N = ",s.N
+
+    s = sim(layout="REG",order=100) #INSTANTIATE OBJECT
+    s.generate_antenna_layout()
+    s.plot_ant(title="REG")
+    phi,zeta = s.calculate_phi(s.ant[:,0],s.ant[:,1])
+    s.plot_zeta(zeta,10,10,14,'jet')
+    print "s.L = ",s.L
+    print "s.N = ",s.N
+
 if __name__ == "__main__":
-   example_usage()
+   #example_usage()
+   plot_paper_layouts()
    
    
 

### 2016-09-26 paper.tex
complete reworks of basic template

@@
-% mnras_template.tex
+% mn2esample.tex
 %
-% LaTeX template for creating an MNRAS paper
+% v2.1 released 22nd May 2002 (G. Hutton)
 %
-% v3.0 released 14 May 2015
-% (version numbers match those of mnras.cls)
+% The mnsample.tex file has been amended to highlight
+% the proper use of LaTeX2e code with the class file
+% and using natbib cross-referencing. These changes
+% do not reflect the original paper by A. V. Raveendran.
 %
-% Copyright (C) Royal Astronomical Society 2015
-% Authors:
-% Keith T. Smith (Royal Astronomical Society)
+% Previous versions of this sample document were
+% compatible with the LaTeX 2.09 style file mn.sty
+% v1.2 released 5th September 1994 (M. Reed)
+% v1.1 released 18th July 1994
+% v1.0 released 28th January 1994
 
-% Change log
+\documentclass[useAMS,usenatbib]{mn2e}
+
+% If your system does not have the AMS fonts version 2.0 installed, then
+% remove the useAMS option.
 %
-% v3.0 May 2015
-%    Renamed to match the new package name
-%    Version number matches mnras.cls
-%    A few minor tweaks to wording
-% v1.0 September 2013
-%    Beta testing only - never publicly released
-%    First version: a simple (ish) template for creating an MNRAS paper
+% useAMS allows you to obtain upright Greek characters.
+% e.g. \umu, \upi etc.  See the section on "Upright Greek characters" in
+% this guide for further information.
+%
+% If you are using AMS 2.0 fonts, bold math letters/symbols are available
+% at a larger range of sizes for NFSS release 1 and 2 (using \boldmath or
+% preferably \bmath).
+%
+% The usenatbib command allows the use of Patrick Daly's natbib.sty for
+% cross-referencing.
+%
+% If you wish to typeset the paper in Times font (if you do not have the
+% PostScript Type 1 Computer Modern fonts you will need to do this to get
+% smoother fonts in a PDF file) then uncomment the next line
+% \usepackage{Times}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Basic setup. Most papers should leave these options alone.
-\documentclass[a4paper,fleqn,usenatbib]{mnras}
-
-% MNRAS is set in Times font. If you don't have this installed (most LaTeX
-% installations will be fine) or prefer the old Computer Modern fonts, comment
-% out the following line
-%\usepackage{newtxtext,newtxmath}
-% Depending on your LaTeX fonts installation, you might get better results with one of these:
-%\usepackage{mathptmx}
-\usepackage{txfonts}
-
-% Use vector fonts, so it zooms properly in on-screen viewing software
-% Don't change these lines unless you know what you are doing
-\usepackage[T1]{fontenc}
-\usepackage{ae,aecompl}
-
-%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%
-
-% Only include extra packages if you really need them. Common packages are:
-\usepackage{graphicx}	% Including figure files
-%\usepackage{wasysym}
-\usepackage{savesym}
-\savesymbol{iint}
-\savesymbol{iiint}
-\savesymbol{iiiint}
-\savesymbol{idotsint}
+%%%%% AUTHORS - PLACE YOUR OWN MACROS HERE %%%%%
+\usepackage{subfigure}
 \usepackage{amsmath}	% Advanced maths commands
 \usepackage{amssymb}	% Extra maths symbols
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
-
-% Please keep new commands to a minimum, and use \newcommand not \def to avoid
-% overwriting existing commands. Example:
-%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
+\usepackage{graphicx}
 \newcommand{\bz}{\bmath{z}}
@@
 \newcommand{\conj}[1]{\overline{#1}}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%
+\title[Dust envelopes around RV Tauri stars]{Dust envelopes around RV Tauri stars}
+\author[A. V. Raveendran and A. N. Other]{A. V. Raveendran$^{1}$\thanks{E-mail:
+email@address (AVR); otheremail@otheraddress (ANO)} and A. N.
+Other$^{2}$\footnotemark[1]\thanks{This file has been amended to
+highlight the proper use of \LaTeXe\ code with the class file.
+These changes are for illustrative purposes and do not reflect the
+original paper by A. V. Raveendran.}\\
+$^{1}$Indian Institute of Astrophysics, Bangalore 560034, India\\
+$^{2}$Building, Institute, Street Address, City, Code, Country}
+\begin{document}
 
-% Title of the paper, and the short title which is used in the headers.
-% Keep the title short and informative.
-\title[Sparse Redundant Calibration]{SpARC: Sparse Redundant Calibration}
+\date{Accepted 1988 December 15. Received 1988 December 14; in original form 1988 October 11}
 
-% The list of authors, and the short list which is used in the headers.
-% If you need two or more lines of authors, add an extra line using \newauthor
-\author[K. T. Smith et al.]{
-Keith T. Smith,$^{1}$\thanks{E-mail: mn@ras.org.uk (KTS)}
-A. N. Other,$^{2}$
-Third Author$^{2,3}$
-and Fourth Author$^{3}$
-\\
-% List of institutions
-$^{1}$Royal Astronomical Society, Burlington House, Piccadilly, London W1J 0BQ, UK\\
-$^{2}$Department, Institution, Street Address, City Postal Code, Country\\
-$^{3}$Another Department, Different Institution, Street Address, City Postal Code, Country
-}
+\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2002}
 
-% These dates will be filled out by the publisher
-\date{Accepted XXX. Received YYY; in original form ZZZ}
-
-% Enter the current year, for the copyright statements etc.
-\pubyear{2015}
-
-% Don't change these lines
-\begin{document}
-\label{firstpage}
-\pagerange{\pageref{firstpage}--\pageref{lastpage}}
 \maketitle
 
-% Abstract of the paper
+\label{firstpage}
+
 \begin{abstract}
-This is a simple template for authors to write new MNRAS papers.
-The abstract should briefly describe the aims, methods, and main results of the paper.
-It should be a single paragraph not more than 250 words (200 words for Letters).
-No references should appear in the abstract.
+In the {\it IRAS\/} [12]--[25],  [25]--[60] colour--colour diagram, RV
+Tauri stars are found to populate cooler temperature regions
+$(T<600\,\rmn{K})$, distinctly different from those occupied by  the
+oxygen and carbon Miras. The {\it IRAS\/} fluxes are  consistent with the
+dust density in the envelope varying as the inverse square of the
+radial distance, implying that the grain formation processes in these
+objects are most probably continuous and not sporadic. It is found that
+the spectroscopic subgroups A and B are well separated in the
+far-infrared two-colour diagram, with group B objects having
+systematically cooler dust envelopes. We interpret this as being due to
+a difference in the nature of grains, including the chemical
+composition, in the two cases.
 \end{abstract}
 
-% Select between one and six entries from the list of approved keywords.
-% Don't make up new ones.
 \begin{keywords}
-keyword1 -- keyword2 -- keyword3
+circumstellar matter -- infrared: stars.
 \end{keywords}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%
 
@@
  we explore both of these parameters.
+ \end{enumerate}
 
@@
 
- We neeed LOGCAL and LINCAL in matrix forms
-
-\end{enumerate}
-
-%All papers should start with an Introduction section, which sets the work
-%in context, cites relevant earlier studies in the field by \citet{Others2013},
-%and describes the problem the authors aim to solve \citep[e.g.][]{Author2012}.
-
  \section{Visibilities}
@@
 \end{equation}
+
+\begin{figure*}
+\centering
+\subfigure[full-complex:   Real]
+{\includegraphics[width=0.33\textwidth]{./HEX_lay.pdf}\label{fig:stef_r}}
+\subfigure[full-complex:  Imag]
+{\includegraphics[width=0.33\textwidth]{./SQR_lay.pdf}\label{fig:stef_i}}
+\subfigure[full-complex: ghost pattern]
+{\includegraphics[width=0.33\textwidth]{./REG_lay.pdf}\label{fig:full_pat_45}}
+
+\subfigure[phase-only:   Real]
+{\includegraphics[width=0.33\textwidth]{./HEX_phi.pdf}\label{fig:phase_r}}
+\subfigure[phase-only:   Imag]
+{\includegraphics[width=0.33\textwidth]{./SQR_phi.pdf}\label{fig:phase_i}}
+\subfigure[phase-only: ghost pattern]
+{\includegraphics[width=0.33\textwidth]{./REG_phi.pdf}\label{fig:phase_pat_45}}
+\caption{scenario. \label{fig:ghost_pattern_bas_45}}
+\end{figure*}
+
 If the mapping $\phi_{ij}$ is known and one is given one of the two antenna indexes that together form a redundant baseline as well as the redundant baseline index itself then it is possible 
@@
 \begin{eqnarray}
- \bd &=& [d_{pq}] \big \} ~ [pq] = 1\cdots N ~ (p<q),\nonumber\\
- \bv &=& [g_p y_{\phi_{pq}} \conj{g}_q]  \big \} ~ [pq] = 1\cdots N ~ (p<q),\nonumber\\
- \bn &=& [n_{pq}]  \big \} ~ [pq] = 1\cdots N ~ (p<q).\label{eq:vec_definitions}
+ \bd &=& [d_{pq}] \big \} ~ [pq] = 1\cdots B ~ (p<q),\nonumber\\
+ \bv &=& [g_p y_{\phi_{pq}} \conj{g}_q]  \big \} ~ [pq] = 1\cdots B ~ (p<q),\nonumber\\
+ \bn &=& [n_{pq}]  \big \} ~ [pq] = 1\cdots B ~ (p<q).\label{eq:vec_definitions}
 \end{eqnarray}
-In Eq.~\eqref{eq:vec_definitions}, $N$ denotes the number of antennas in the array. NB --- above equation is wrong, needs to be baselines ...
+In Eq.~\eqref{eq:vec_definitions}, $B$ denotes the number of baselines in the array.
 
@@
 
-\section{Historical Overview}
+\bibliographystyle{mn2e}
+\bibliography{paper}
 
-Write a little about how redundant calibration developed, a basic lit review.
-Maybe write a little bit about LOGCAL here as well.
+\label{lastpage}
 
-\subsection{EW regular-layout}
-Describe the EW-layout and its redundant baselines
-
-\subsection{LINCAL}
-\begin{equation}
-c_{ij} = y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}
-\end{equation}
-
-\begin{eqnarray}
-\frac{c_{ij}}{\partial \eta_i} &=& y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\ 
-\frac{c_{ij}}{\partial \varphi_i} &=&  -i y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\
-\frac{c_{ij}}{\partial \eta_j} &=& y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\ 
-\frac{c_{ij}}{\partial \varphi_j} &=&  i y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\
-\frac{y_{i-j}}{\partial \varphi_j} &=&  e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}
-\end{eqnarray}
-
-The Wirtinger derivative is used in the last equation.
-
-\begin{eqnarray}
-c_{ij} &\approx& c_{ij}^0 + \Delta \eta_i(y_{i-j}^0 e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}) + \Delta \eta_j(y_{i-j}^0 e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0})\\
-&& -i\Delta\varphi_i (y_{i-j}^0e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}) +i\Delta\varphi_j (y_{i-j}^0e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0})\\
-&& y_{i-j}^1(y_{i-j}^0e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0})\\
-&\approx&  c_{ij}^0 + e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_i+ \Delta \eta_j - i\Delta\varphi_i + i\Delta\varphi_i))
-\end{eqnarray}
-
-\begin{equation}
-\label{eq:residual}
-r_{ij} = \delta_{ij} = c_{ij}-c_{ij}^0 = e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_i+ \Delta \eta_j - i\Delta\varphi_i + i\Delta\varphi_i)) 
-\end{equation}
-
-Eq.~\ref{eq:residual} allows us to construct the following linear system:
-
-\begin{equation}
-\boldsymbol{J}[\boldsymbol{\Delta \eta},\boldsymbol{\Delta \varphi},\boldsymbol{\Delta y}]^T = \boldsymbol{r}, 
-\end{equation}
-
-where $\boldsymbol{J}$ is equal to 
-\begin{equation}
-\boldsymbol{J} = \bigg [\overbrace{\frac{c_{pq}}{\partial \eta_i}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial \varphi_i}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial y_{t}}}^{t=1\cdots r_s} \bigg ] \bigg \} [pq] = 1\cdots N_{\textrm{b}} (p<q)
-\end{equation}
-or
-\begin{equation}
-\boldsymbol{J} = \bigg [\frac{c_{pq}}{\partial \boldsymbol{\eta}},~\frac{c_{pq}}{\partial \boldsymbol{\varphi}},~\frac{c_{pq}}{\partial \boldsymbol{y}} \bigg ] \bigg \} [pq] = 1\cdots N_{\textrm{b}} (p<q).
-\end{equation}
-
-The last column is again a Wirtinger derivative.
-
-Which means we can obtain the least-squares estimate as follows:
-
-\begin{equation}
-[\boldsymbol{\Delta \eta},\boldsymbol{\Delta \varphi},\boldsymbol{\Delta y}]^T = [\boldsymbol{J}^H\boldsymbol{J}]^{-1}\boldsymbol{J}^H\boldsymbol{r}.
-\end{equation}
-
-What is interesting about LINCAL is that it seems to be a combination of standard least-squares (splitting the problem into amplitude and phase) and 
-complex least-squares.
-The gains are solved with a standard approach (splitting into amp and phase) and the redundant visibilities use a Wirtinger derivative.
-What is strange though is that in normal complex optimization, if we optimize towards a complex variable x we also need to optimize towards
-its complex conjugate.
-My best guess is that LINCAL works in any case since the conjugate redundant baselines are not used to calibrate ($p<q$).
-The only difference between the framework I was using is that
-we regard the whole problem as complex, we construct our Jacobian by calculating the derivative towards $g$ and its conjugate as well as $y$ and its conjugate - Smirnov \& Tasse (2016).
-
-Now that the relation between LINCAL and GN has been shown we can start focusing on the speedup that the conjugate gradient method provides as Liu (2010) originally proposed.
-Also worth checking: whether using an LM step instead of a GN step improves your convergence properties.
-Briefly review LINCAL here and show that it is equivalent to the GN algorithm.
-
-\section{Complex Optimization}
-\subsection{Scalar Calibration}
-Scalar direction independent calibration is achieved through solving the following minimization problem:
-\begin{equation}
-\min_{\bg}\|\br\| = \min_{\bg}\|\bd - \bv(\bg)\|, 
-\end{equation}
-where
-\begin{enumerate}
- \item $\bd$ is the observed visibility vector, i.e. 
- \begin{equation}
-  \bd = [d_{pq}] \big \} ~ [pq] = 1\cdots N ~ (p<q).
- \end{equation}
- The vector entry $d_{pq}$ denotes the observed visibility associated with the baseline formed by antenna $p$ and $q$, while $N$ denotes the number or antennas
- in the array. 
- \item $\bg = [g_1,g_2,\cdots,g_N]$ represents the instrumental complex antenna gain vector.
- \item $\bv(\bg)$ denotes the predicted visibility vector, i.e.
- \begin{equation}
-  \bv(\bg) = [g_p y_{pq}\conj{g}_q]  \big \} ~ [pq] = 1\cdots N ~ (p<q).
- \end{equation}
- In Eq... $y_{pq}$ denotes the visibility which was created from a skymodel and is associated with antenna $p$ and $q$. [Need to rw this sentence] 
-\end{enumerate}
-
-One of the simplests approaches one could use to solve eq. is to use least-squares optimization which we can only employ if we are able to calculate the gradient of $\bv(\bg)$,
-implying that $\bv(\bg)$ needs to be analytic in $\bg$ (the Taylor series expansion of $\bv(\bg)$ exists).
-The problem we face if we try to naivly apply least-squares to eq is that $\bv(\bg)$ is not analytic in its argument, since $\frac{\partial \conj{z}}{\partial z}$ (where 
-$z \in \mathbb{C}$) does not exist.
-
-The easiest way to circumvent this problem is to recast our problem in such a way that the predicted visibility vector becomes analytic in its argument.
-The most widely used way to accomplish this is to restate eq as
-\begin{equation}
-\min_{\hat{\bg}}\|\hat{\br}\| = \min_{\hat{\bg}}\|\hat{\bd} - \hat{\bv}(\hat{\bg})\|, 
-\end{equation}
-where $\hat{\bg} = \big[\mathcal{R}\{\bg\},\mathcal{I}\{\bg\}\big]^T$, still need to define the other vectors properly....
-Note that $\hat{\bv}(\hat{\bg})$ is now analytic in its argument and we can proceed and use least squares to estimate $\bg$.
-
-\citet{Smirnov2015} have recently shown that calibration can also be performed using the following complex framework: 
-\begin{equation}
-\min_{\breve{\bg}}\|\breve{\br}\| = \min_{\breve{\bg}}\|\breve{\bd} - \breve{\bv}(\breve{\bg})\|, 
-\end{equation}
-where $\breve{\bg} = \big[\bg,\conj{\bg}]^T$, still need to define the other vectors properly....
-In the above formulation v breve is analytic in its argument and its complex conjugate as a whole under the assumption that the Wirtinger derivative
-is used to define the gradient operator. As v breve is analytic we will are also able to apply the least squares method to .
-
-The Wirtinger derivatives are defined as
-\begin{eqnarray}
-\frac{\partial}{\partial z} = \frac{1}{2}\left ( \frac{\partial}{\partial x} -  i \frac{\partial}{\partial y} \right ),&~&\frac{\partial}{\partial \conj{z}} = \frac{1}{2}\left ( \frac{\partial}{\partial x} +  i \frac{\partial}{\partial y} \right )
-\end{eqnarray}
-where $z = x + iy$ and $x,y \in \mathbb{R}$.
-It follows from the above that $\frac{\partial z}{\partial z} = 1,\frac{\partial z}{\partial \conj{z}} = 0,\frac{\partial \conj{z}}{\partial z} = 0$ and $\frac{\partial \conj{z}}{\partial \conj{z}} = 1$.
-
-%The main difference between the traditional approach and the 
-%complex optimization framework proposed by \citet{Smirnov2015} is that in their framework they no longer treat the real and imaginary parts of the residual separately which enables them to directly solve for the complex antenna gains themselves (as well as their
-%complex conjugates).
-
-Recall that most optimization techniques require a gradient operator to enable it to function properly. The main reason why we have traditionally treated 
-the real and imaginary parts of the residual separately is due to the fact that $f(z) = \conj{z}$ is not a differentiable complex function where $z\in\mathbb{C}$ and $\conj{(*)}$ denotes conjugation, i.e.
-$\frac{df(z)}{dz}$
-does not exist, if we apply the standard definition of differentiation:
-\begin{equation}
-x
-\end{equation}
-
-The main reason that we do not need to treat the real and imaginary parts of the residual separately any more is due to the 
-fact
-
-%A standard way to optimize a complex function $f(\bz)$ of $n$ complex variables $\bz\in\mathbb{Z}^n$ is to recast $f$ as  optimize the real and imaginary parts of  separately.
-
-The main difference between complex optimization and the more traditional
-approach to calibration is that in complex calibration we
-
-\subsection{Redundant Calibration}
-In eq a sky model is required to generate the predicted visibilities $y_{pq}$ as we have too few equations to also solve for them and the gains. If we use a redundant layout
-we can reduce the number of unknowns so that we are able to solve for the sky and the gains simultaneously.
-
-\subsubsection{Antenna indices to redundancy index mapping}
-If an array is redundant its redundant baselines sample the exact same visibilites in the $uv$-plane. Recall, that we denote the true observed visibility associated with baseline $pq$ and 
-baseline $rs$ with $y_{pq}$ and $y_{rs}$, respectively . However if baseline $pq$ and $rs$ represent the same redundant baseline then $y_{pq} = y_{rs}$. Another
-usefull indexing strategy would therefore be to use redundant baseline indices instead of antenna pairs to label the true observed visibilities so that we can identify 
-which baselines observe the same visibilities barring instrumental and environmental errors.
-
-Let $\phi_{ij}:\mathbb{N}^2\rightarrow\mathbb{N}^+$ be the mapping that maps antenna indices to the unique redundant baseline indices of the array. This mapping is not symmetric as 
-$\phi_{ij} = 0$ if $i>j$. For an east-west regular array we can construct a simple analytic expression for this mapping:
-\begin{equation}
-\widehat{\phi_{ij}} = 
-\begin{cases}
-j-i&\textrm{if}~i<j\\
-0&\textrm{otherwise}
-\end{cases}.
-\end{equation}
-As will become apparent later the following symmetric variant of $\phi_{ij}$ will turn out to be quite useful later on:
-\begin{equation}
-\zeta_{ij} = 
-\begin{cases}
-\phi_{ij}~\textrm{if}~i\leq j\\
-\phi_{ji}~\textrm{if}~i>j
-\end{cases}.
-\end{equation} 
-It is also crucial to be able to determine the antenna index of the antenna which was used to construct a specific redundant baseline if one is given the other antenna index and the redundant baseline index.
-This can be calculated via the following two functions:
-\begin{equation}
-\xi_{ij} = 
-\begin{cases}
-p~\textrm{if}~\exists! ~ p \in \mathbb{N} ~ s.t. ~(\phi_{pi} = j)\\
-0~\textrm{otherwise}
-\end{cases},
-\end{equation}
-and
-\begin{equation}
-\psi_{ij} = 
-\begin{cases}
-q~\textrm{if}~\exists! ~ q \in \mathbb{N} ~ s.t. ~(\phi_{iq} = j)\\
-0~\textrm{otherwise}
-\end{cases}
-\end{equation}
-We use $\xi_{ij}$ to retrieve the index of the second antenna if we know the redundant baseline index and the index of the first antenna used to construct the baseline in question, while we use $\psi_{ij}$ to return the index of the first antenna if we know the
-index of the second antenna instead. \textbf{NB::} I still need to plot $\zeta_{ij}$ for different antenna layouts.
-
-Describe complex optimization, Jacobian, Hessian etc... Benefits drawbacks compared to LINCAL... etc..
-
-\subsubsection{Redundant calibration}
-In the case of a redundant array we can not only solve for the gains $\bg$ but we can also solve for the redundant unknown visibilities $\by$. Where $\by = [y_1,\cdots,y_R]$, $y_i$ denotes 
-a unique redundant visibility and $R$ is the maximum value of the mapping $\phi_{ij}$. The reason we can now also solve for $\by$ is because we have reduced the number of 
-unknowns from $N+B$ to $N+R$, where $N+R$ is now less than $B$ due to the array redundancy. We can now reformulate the calibration problem for an redundant array as follow:
-\begin{equation}
-\min_{\bz}\|\br\| = \min_{\bz}\|\bd - \bv(\bz)\|, 
-\end{equation}
-where $\bz = [\bg,\by]^T$.
-
-If we now follow the same extension strategy of \citet{Smirnov2015} we can rewrite eq to
-\begin{equation}
-\min_{\breve{\bz}}\|\breve{\br}\| = \min_{\bz}\|\breve{\bd} - \breve{\bv}(\breve{\bz})\|, 
-\end{equation}
-where $\breve{\bz} = [\bz,\conj{\bz}]^T$ still need to define the remaining variables properly.
-
-We can now solve the above equation by using least-squares, which entails deriving the Gauss-Newton (GN) and Levenberg-Marquerdt (LM) update rules.
-(NEED TO WRITE THIS MUCH BETTER)
-
-\begin{equation}
-\bJ = \frac{\partial\bv}{\partial\bz}~\bJ^* = \frac{\partial \bv}{\partial \breve{\bz}} 
-\end{equation}
-
-\begin{equation}
-\bJ = \begin{bmatrix}
-       \bJ & \bJ^* \\
-       \conj{\bJ^*} & \conj{\bJ}  
-      \end{bmatrix}
-\end{equation}
-
-\begin{equation}
- \Delta \breve{\bz} = \begin{bmatrix}\Delta \bz\\ \Delta \conj{\bz} \end{bmatrix} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\br_k}
-\end{equation}
-
-\begin{equation}
- \Delta \breve{\bz} = \begin{bmatrix}\Delta \bz\\ \Delta \conj{\bz} \end{bmatrix} = (\bJ^H\bJ + \lambda \bD)^{-1}\bJ^H\breve{\br_k}
-\end{equation}
-
-\subsection{Jacobian derivation}
-\subsection{Hessian derivation}
-\begin{equation}
-\boldsymbol{H} = 
-\begin{bmatrix}
-\boldsymbol{A} & \boldsymbol{B}\\
-\boldsymbol{B}^* & \boldsymbol{A}^*
-\end{bmatrix}
-\end{equation}
-
-With
-
-\begin{equation}
-\boldsymbol{A} = 
-\begin{bmatrix}
-\boldsymbol{C} & \boldsymbol{D}\\
-\boldsymbol{D^H} & \boldsymbol{E}
-\end{bmatrix}
-\end{equation}
-
-\begin{equation}
-\boldsymbol{B} = 
-\begin{bmatrix}
-\boldsymbol{F} & \boldsymbol{G}\\
-\boldsymbol{G}^T & \boldsymbol{0}
-\end{bmatrix}
-\end{equation}
-
-\begin{equation}
-[\boldsymbol{C}]_{ij} = 
-\begin{cases}
- \sum_{k \neq i} \left | g_k \right |^2 \left | y_{\zeta_{ik}} \right |^2 & \textrm{if} ~ i=j\\
- 0 & \textrm{otherwise}
-\end{cases}
-\end{equation}
-
-\begin{equation}
-[\boldsymbol{D}]_{ij} = 
-\begin{cases}
- g_i y_j^*  \left | g_{\psi_{ij}} \right |^2  & \textrm{if} ~ \psi_{ij}\neq0\\
- 0 & \textrm{otherwise}
-\end{cases}
-\end{equation}
-
-\begin{equation}
-[\boldsymbol{E}]_{ij} = 
-\begin{cases}
- \sum_{pq \in \mathcal{I}_i} \left | g_p \right |^2 \left | g_q \right |^2  & \textrm{if} ~ i=j\\
- 0 & \textrm{otherwise}
-\end{cases}
-\end{equation}
-
-\begin{equation}
-[\boldsymbol{F}]_{ij} = 
-\begin{cases}
- g_i g_j  \left | y_{\zeta_{ij}} \right |^2  & \textrm{if} ~ (i \neq j)\\
- 0 & \textrm{otherwise}
-\end{cases}
-\end{equation}
-
-\begin{equation}
-[\boldsymbol{G}]_{ij} = 
-\begin{cases}
- g_i y_j  \left | g_{\xi_{ij}} \right |^2  & \textrm{if} ~ (\xi_{ij}\neq0)\\
- 0 & \textrm{otherwise}
-\end{cases}
-\end{equation}
-
-Moreover,
-
-\begin{equation}
-\mathcal{I}_i = \left\{pq\in\mathbb{N}^2|(\phi_{pq} = i) \right\}.
-\end{equation}
-
-The dimensions of the matrices:
-
-\begin{enumerate}
-\item $\boldsymbol{H}$: is a $(4N-2)\times(4N-2)$ or a $P \times P$ matrix. $N$ denotes the number of antennas and $P$ denotes the number of parameters. This matrix
-contains $6N^2 - 2N - 2$ non-zero entries. It contains $16N^2 - 16N + 4$ entries. We can construct this matrix with $4N^2 -  4N$ or $\frac{1}{4} P^2 -1$ elementary operations.
-\item $\boldsymbol{A}$: is a $(2N-1)\times(2N-1)$ matrix. This matrix has $N^2 + N - 1$ non-zero entries. We can construct this matrix with $2\frac{1}{2} (N^2 -  N)$ elementary operations.  
-\item $\boldsymbol{B}$: is a $(2N-1)\times(2N-1)$ matrix. This matrix has $2N^2 - 2N$ non-zero entries. We can construct this matrix with $1\frac{1}{2} (N^2 -  N)$ elementary operations.
-\item $\boldsymbol{C}$: is a $N\times N$ matrix. This matrix has $N$ non-zero entries. We can construct this matrix with $N^2-N$ elementary operations.
-\item $\boldsymbol{D}$: is a $N \times (N-1)$ matrix. This matrix has $\frac{1}{2} (N^2 -  N)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations. 
-\item $\boldsymbol{E}$: is a $(N-1) \times (N-1)$ matrix. This matrix has $(N-1)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.  
-\item $\boldsymbol{F}$: is a $N \times N$ matrix. This matrix has $N^2 - N$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.
-\item $\boldsymbol{G}$: is a $N \times (N-1)$ matrix. This matrix has $\frac{1}{2} (N^2 -  N)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.
-\item $\boldsymbol{0}$: is a $(N-1) \times (N-1)$ all zero matrix. 
-\end{enumerate}
-
-We can construct $\boldsymbol{C}-\boldsymbol{G}$ in $O(N^2)$
-
-The sparsity ratio of $\boldsymbol{H}$ is equal to
-\begin{equation}
-\gamma_N = \frac{10N^2-14N+6}{16N^2-16N+4} 
-\end{equation}
-
-The asymptotic sparsity ratio of $\boldsymbol{H}$ is equal to
-
-\begin{equation}
-\gamma = \lim_{N\rightarrow \infty}\frac{5N^2-7N+3}{8N^2-8N+2} = \frac{5}{8}
-\end{equation}
-
-Asymptotically the computational cost of the matrix-vector product between $\boldsymbol{H}$ and a vector of appropriate size is $O(P^2)$. However, the computational cost converges 
-very slowly to its asymptotic value, and in general is equal to $O(P^{k_N})$, where $k_N < 2$.
-
-\begin{eqnarray}
-P^{k_N}_N &=& (1 - \gamma_N)P^2_N\\
-k_N &=& \log_{P_N}(1 - \gamma_N) + 2\\
-k &=& \lim_{N\rightarrow \infty} \log_{P_N}(1 - \gamma_N) + 2 = 2
-\end{eqnarray}
-
-Moreover,
-
-\begin{eqnarray}
-P^{c_N}_N &=& \frac{1}{4} P_N^2 -1\\
-c_N &=& \log_{P_N}\left (\frac{1}{4} P_N^2-1 \right )\\
-c &=& \lim_{N\rightarrow \infty} \log_{P_N}\left (\frac{1}{4} P_N^2-1\right ) = 2
-\end{eqnarray}
-
-NB $P_c$ IS WRONG HERE DID NOT TAKE INTO ACCOUNT THE CONJUGANTION OF $\boldsymbol{B}$ (need multiply its computational complexity by two).
-
-\subsection{Jacobian-residual product}
-
-\subsection{Computational complexity using CG}
-\subsubsection{Sparsity}
-\subsubsection{Condition number}
-
-\begin{enumerate}
- \item Plot eigenvalue distribution for different SNR and antenna numbers and different lambda values
- \item Plot computational complexity bound plus error reduction for different setups
- \item Plot number of iterations required by CG as a function of number of antennas
- \item Plot redundant baseline visibilities to show that they are approximately equal.
-\end{enumerate}
-
-NB:: Repeat the second last plot for the positive-semi definite case (GN), showing that even if the Hessian is singular we get fast convergence.
-
-\section{Results: Regualar square grid}
-If there is time
-
-\section{Results: Hexanol grid}
-NB:: Need to test if it works for this redundant layout.
-
-\section{Conjugate Gradient method}
-brief description and lit review of the method.
-computational complexity of the algorithm.
-sparsity and condition number.
-% \section{Methods, Observations, Simulations etc.}
-% 
-% Normally the next section describes the techniques the authors used.
-% It is frequently split into subsections, such as Section~\ref{sec:maths} below.
-% 
-% \subsection{Maths}
-% \label{sec:maths} % used for referring to this section from elsewhere
-% 
-% Simple mathematics can be inserted into the flow of the text e.g. $2\times3=6$
-% or $v=220$\,km\,s$^{-1}$, but more complicated expressions should be entered
-% as a numbered equation:
-% 
-% \begin{equation}
-%     x=\frac{-b\pm\sqrt{b^2-4ac}}{2a}.
-% 	\label{eq:quadratic}
-% \end{equation}
-% 
-% Refer back to them as e.g. equation~(\ref{eq:quadratic}).
-% 
-% \subsection{Figures and tables}
-% 
-% Figures and tables should be placed at logical positions in the text. Don't
-% worry about the exact layout, which will be handled by the publishers.
-% 
-% Figures are referred to as e.g. Fig.~\ref{fig:example_figure}, and tables as
-% e.g. Table~\ref{tab:example_table}.
-% 
-% % Example figure
-% \begin{figure}
-% 	% To include a figure from a file named example.*
-% 	% Allowable file formats are eps or ps if compiling using latex
-% 	% or pdf, png, jpg if compiling using pdflatex
-% 	\includegraphics[width=\columnwidth]{example}
-%     \caption{This is an example figure. Captions appear below each figure.
-% 	Give enough detail for the reader to understand what they're looking at,
-% 	but leave detailed discussion to the main body of the text.}
-%     \label{fig:example_figure}
-% \end{figure}
-% 
-% % Example table
-% \begin{table}
-% 	\centering
-% 	\caption{This is an example table. Captions appear above each table.
-% 	Remember to define the quantities, symbols and units used.}
-% 	\label{tab:example_table}
-% 	\begin{tabular}{lccr} % four columns, alignment for each
-% 		\hline
-% 		A & B & C & D\\
-% 		\hline
-% 		1 & 2 & 3 & 4\\
-% 		2 & 4 & 6 & 8\\
-% 		3 & 5 & 7 & 9\\
-% 		\hline
-% 	\end{tabular}
-% \end{table}
-% 
-% 
-% \section{Conclusions}
-% 
-% The last numbered section should briefly summarise what has been done, and describe
-% the final conclusions which the authors draw from their work.
-
-\section*{Acknowledgements}
-
-The Acknowledgements section is not numbered. Here you can thank helpful
-colleagues, acknowledge funding agencies, telescopes and facilities used etc.
-Try to keep it short.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%
-
-% The best way to enter references is to use BibTeX:
-
-%\bibliographystyle{mnras}
-%\bibliography{example} % if your bibtex file is called example.bib
-
-\bibliographystyle{mnras}
-\bibliography{paper}
-% Alternatively you could enter them by hand, like this:
-% This method is tedious and prone to error if you have lots of references
-%\begin{thebibliography}{99}
-%\bibitem[\protect\citeauthoryear{Author}{2012}]{Author2012}
-%Author A.~N., 2013, Journal of Improbable Astronomy, 1, 1
-%\bibitem[\protect\citeauthoryear{Others}{2013}]{Others2013}
-%Others S., 2012, Journal of Interesting Stuff, 17, 198
-%\end{thebibliography}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%
-
-\appendix
-
-\section{Some extra material}
-
-If you want to present additional material which would interrupt the flow of the main paper,
-it can be placed in an Appendix which appears after the list of references.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Don't change these lines
-\bsp	% typesetting comment
-\label{lastpage}
 \end{document}
-
-% End of mnras_template.tex

### 2016-09-26 paper_old.tex
complete reworks of basic template

% mnras_template.tex
%
% LaTeX template for creating an MNRAS paper
%
% v3.0 released 14 May 2015
% (version numbers match those of mnras.cls)
%
% Copyright (C) Royal Astronomical Society 2015
% Authors:
% Keith T. Smith (Royal Astronomical Society)

% Change log
%
% v3.0 May 2015
%    Renamed to match the new package name
%    Version number matches mnras.cls
%    A few minor tweaks to wording
% v1.0 September 2013
%    Beta testing only - never publicly released
%    First version: a simple (ish) template for creating an MNRAS paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[a4paper,fleqn,usenatbib]{mnras}

% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
%\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
\usepackage{txfonts}

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}
%\usepackage{subfig,multirow}
\usepackage{pdflscape}

%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
%\usepackage{wasysym}
\usepackage{savesym}
\savesymbol{iint}
\savesymbol{iiint}
\savesymbol{iiiint}
\savesymbol{idotsint}
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{amsmath,color}
\usepackage{graphicx}
%\usepackage[showframe]{geometry}% http://ctan.org/pkg/geometry
\usepackage{lipsum}% http://ctan.org/pkg/lipsum
\usepackage{graphicx}% http://ctan.org/pkg/graphicx
%\usepackage[demo]{graphicx}
%\usepackage{subfigure}
\usepackage{marvosym}
\usepackage{mathtools}
\usepackage{cancel}
\usepackage{bm}
\usepackage[T1]{fontenc}
\usepackage{aecompl} 
\usepackage{arydshln}
\usepackage{subfig,multirow}
\usepackage{mathrsfs}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
\newcommand{\bz}{\bmath{z}}
\newcommand{\br}{\bmath{r}}
\newcommand{\bg}{\bmath{g}}
\newcommand{\bd}{\bmath{d}}
\newcommand{\bv}{\bmath{v}}
\newcommand{\bn}{\bmath{n}}
\newcommand{\by}{\bmath{y}}
\newcommand{\bJ}{\bmath{J}}
\newcommand{\bD}{\bmath{D}}
\newcommand{\conj}[1]{\overline{#1}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Sparse Redundant Calibration]{SpARC: Sparse Redundant Calibration}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[K. T. Smith et al.]{
Keith T. Smith,$^{1}$\thanks{E-mail: mn@ras.org.uk (KTS)}
A. N. Other,$^{2}$
Third Author$^{2,3}$
and Fourth Author$^{3}$
\\
% List of institutions
$^{1}$Royal Astronomical Society, Burlington House, Piccadilly, London W1J 0BQ, UK\\
$^{2}$Department, Institution, Street Address, City Postal Code, Country\\
$^{3}$Another Department, Different Institution, Street Address, City Postal Code, Country
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2015}

% Don't change these lines
\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle

% Abstract of the paper
\begin{abstract}
This is a simple template for authors to write new MNRAS papers.
The abstract should briefly describe the aims, methods, and main results of the paper.
It should be a single paragraph not more than 250 words (200 words for Letters).
No references should appear in the abstract.
\end{abstract}

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
keyword1 -- keyword2 -- keyword3
\end{keywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%

\section{Introduction}

Main contributions of the paper:

\begin{enumerate}
 \item Present a general framework which unifies all the techniques developed so far showing they are all related (LINCAL, non-linear estimator, etc ...). They are all non-linear
 least-squares techniques, i.e. they either employ Gauss-Newton or the Levenberg-Marquardt algorithms. We have to start with LINCAL showing that it is GN, then we have to motivate 
 why we want to use Oleg's complex optimization framework.
 \item Use Oleg's complex optimization framework to re-derive the non-linear technique proposed by Marthi and Chengular. The novelty lies, in the fact that by using Oleg's
 framework we can find analytic expressions for the Jacobian, the Hessian and the Jacobian-residual product (which is not even the case for LINCAL). State that the algorithm is 
 effectively related to SteFCal and is eff an independent rediscovery of SteFCal and an extension of it into the redundant domain.
 \item Also we present the array geometry function to help us make the derivations from Marthi and Chengular easier to read and understand for a general layout.
 \item We also mention at this point that in Oleg's paper the question is raised is there a fast way of computing the exact inverse, we then present this new method, which is called
 conjugate gradient method. We discuss the algorithm and the two things which bound its execution time. Which is $kappa$ and $m$ (spectral condition number and its sparsity). Then
 we explore both of these parameters.
  \end{enumerate} 
 **************
 FLOW OF PAPER
 **************

 \begin{enumerate}
 \item We need the definition of visibilities as in Liu. DEFINE SNR HERE ALREADY together maybe with the sigma value of the noise.
 \item Introduce the array geometery function.
 \item Write down logcal and lincal in matrix form... ?
 \item Short discussion of least squares and jacobian and hessian's. General GN and LM update rules.
 \item Introduce redundant calibration as a least squares problem. We will use this general fact to derive both popular methods.
 \item Propose a possible solution witch leads to lincal. Mention that this approach works as in this form the function is differentiable (a taylor expansion in the 
 parameters exists). Maybe mention that we can also divide the problem into real and imaginary etc...
 \item Show how this relates to lincal for example ---> show lincal is GN.
 \item Now introduce complex optimization ---> Main motivation for switching to the alternative framework is that the the differentials are very simplistic. We wish to show that
 we can derive the method of Chengular.
 \item Do the derivation of Chengular. 
 \item Brief discussion abouth Chengular and SteFCal and Complex Optimization. Here I show that in Oleg's paper he re-derives an algorithm called SteFCal. Stefcal works
 on the basis of alternating direction implicit method. The first signs of achieving a similar algorithm already appeared in Stefan's conference paper in which the alternating
 idea was first proposed. Then I mention Chegular re-derives the expression by using partial derivatives and extends it to redundant. One could also have used the linear alternating
 approach. In an attempt to merge the terminology that has independently develop in the general calibration and redundant calibration literature and to emphasize the close
 relationship between Stefcal and the approach derived by I will use the label Redunandat SteFCal (R-Stefcal) to refer to the NLS method proposed by ... Lastly we mention that similarly
 to how Oleg re-derived stefcal in, we have achieved the same approach.
 \item Faster Exact inverse. A question that Oleg poses in his paper is, does there exist a faster way to take the exact inverse of JhJ? One that is almost linear, and
 implies that we can therefore implement the full LM algorithm. The aim here is of course to reduce the number of iterations that are required to converge by using the 
 full inverse instead which would hopefully provide enough of a speedup to compensate for the more expensive full-inverse. The algorithm we propose is the conjugate 
 gradient method.
 \item We give the images of the HESSIAN of both lincal and the complex method (number of terms). So we can mention that both are sparse and contain diagonal entries that 
 are more significant than the off-diagonal entries. What linear inversion approach can take advantage of both these phenomenon. One such technique is 
 cg. 
 \item Briefly discuss CG and how its computationally bounded by its condition number and its sparsity (how does the diagonal play a role).
 \item Simulation description
 \item Will CG improve things?
 \item Now first discuss the condition
 number of the Hessian before and after pre-conditioning (pre-codnitioning can only be applied if a good inverse of a matrix is known, if it is known then it can improve 
 the spectral condition number of a matrix. We present here the kappa and iteration number graphs for the HEXAGONAL layout. Although the
 \item Now we discuss the sparsity results. 
 \item Provide a table that theoretical compares R-StEFCal and SPARC.
 \item Number of outerloop iterations. 
 \item Timing results.
 \item Accuracy results.
 \item Maybe some freq simulations.
 \end{enumerate}

 We

 We neeed LOGCAL and LINCAL in matrix forms

%All papers should start with an Introduction section, which sets the work
%in context, cites relevant earlier studies in the field by \citet{Others2013},
%and describes the problem the authors aim to solve \citep[e.g.][]{Author2012}.

\section{Visibilities}
The observed visibility $d_{pq}$ measured by the baseline formed by antenna $p$ and $q$ can be described as

\begin{equation}
\label{eq:vis_definition}
d_{pq} = g_{p}\conj{g}_{q}y_{pq} + n_{pq},
\end{equation}
where $g_{p}$ and $g_{q}$ are the direction independent gains associated with antenna $p$ and $q$, $y_{pq}$ denotes the true visibility that baseline $pq$ measured
and $n_{pq}$ is the thermal noise component. Conjugation is denoted by $\conj{(*)}$. During the course of an actual observation the true value of $g_p$, $g_q$ and $y_{pq}$ are unknown and are the quantities which infact have 
to be estimated.

The real and imaginary components of the thermal noise is normally distributed with a mean of zero and a standard deviation that is equal to [NB - are the real and imaginary parts defined with same sigma?]  
\begin{equation}
\sigma = \frac{\sqrt{2}k_{B}T_{\textrm{sys}}}{A\eta\sqrt{\Delta \nu \tau}}, 
\end{equation}
where $k_B$ is Boltzmann's constant, $T_{\textrm{sys}}$ is equal to the system temperature, $A$ is the effective collecting area of each element in the array, $\eta$ is a dimensionless
efficiency factor, $\Delta \nu$ is the observational bandwidth and $\tau$ is the integration time per visibility.

\subsection{Redundant Array Geometry Mapping}
If an array is redundant then its redundant baselines sample the exact same visibilites in the $uv$-plane, i.e. if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$. We can
therefore also use the following useful indexing strategy: We can use singular redundant baseline indexes instead of composite antenna pairs to label the true observed visibilities.

Let $\phi_{ij}:\mathbb{N}^2\rightarrow\mathbb{N}^+$ be the mapping that maps composite antenna pairs to the unique redundant baseline indexes of an array. This mapping is not symmetric as 
$\phi_{ij} = 0$ if $i>j$. It is possible to construct a simple analytic expression for $\phi_{ij}$ when our array is in an east-west regular configuration:
\begin{equation}
\phi_{ij} = 
\begin{cases}
j-i&\textrm{if}~i<j\\
0&\textrm{otherwise}
\end{cases}.
\end{equation}
It becomes increasingly more difficult to construct analytic expressions of $\phi_{ij}$ for other more complicated array layouts.

We can also define the following symmetric variant of $\phi_{ij}$:
\begin{equation}
\zeta_{ij} = 
\begin{cases}
\phi_{ij}~\textrm{if}~i\leq j\\
\phi_{ji}~\textrm{if}~i>j
\end{cases}.
\end{equation}

%\begin{figure*}
%\centering
%\subfig[full-complex:   Real]
%{\includegraphics[width=0.34\textwidth]{./HEX_lay.pdf}\label{fig:stef_r}}
% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
%\subfig[full-complex:  Imag]
%{\includegraphics[width=0.34\textwidth]{./SQR_lay.pdf}\label{fig:stef_i}}
% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
%\subfig[full-complex: ghost pattern]
%{\includegraphics[width=0.34\textwidth]{./REG_lay.pdf}\label{fig:full_pat_45}}\\
% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
%\subfig[phase-only:   Real]
%{\includegraphics[width=0.34\textwidth]{./HEX_phi.pdf}\label{fig:phase_r}}
% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
%\subfig[phase-only:   Imag]
%{\includegraphics[width=0.34\textwidth]{./SQR_phi.pdf}\label{fig:phase_i}}
% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
%\subfig[phase-only: ghost pattern]
%{\includegraphics[width=0.34\textwidth]{./REG_phi.pdf}\label{fig:phase_pat_45}}
% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
%\caption{scenario. \label{fig:ghost_pattern_bas_45}}
%\end{figure*}

\begin{figure*}%
\centering
\subfloat[./HEX_lay.pdf]{}%
\qquad
\subfloat[./SQR_lay.pdf]{}
\caption{Here are the first two figures of a continued figure.}%
\label{fig:cont}%
\end{figure*}

If the mapping $\phi_{ij}$ is known and one is given one of the two antenna indexes that together form a redundant baseline as well as the redundant baseline index itself then it is possible 
to calculate the unknown antenna index. This can be calculated via the following two expressions:
\begin{equation}
\xi_{ij} = 
\begin{cases}
p~\textrm{if}~\exists! ~ p \in \mathbb{N} ~ s.t. ~(\phi_{pi} = j)\\
0~\textrm{otherwise}
\end{cases},
\end{equation}
and
\begin{equation}
\psi_{ij} = 
\begin{cases}
q~\textrm{if}~\exists! ~ q \in \mathbb{N} ~ s.t. ~(\phi_{iq} = j)\\
0~\textrm{otherwise}
\end{cases}
\end{equation}

We use $\xi_{ij}$ to retrieve the first antenna index of the composite antenna pair associated with a particular baseline if the index of the first antenna in the composite antenna pair and the unique redundant baseline index are known, while we use $\psi_{ij}$ to obtain 
a similar result; the second antenna index of the composite antenna pair. \textbf{NB::} I still need to plot $\zeta_{ij}$ for different antenna layouts.

If the array is redundant and the mapping $\phi_{ij}$ is known we can rewrite Eq.~\eqref{eq:vis_definition} as
\begin{equation}
\label{eq:vis_red}
d_{pq} = g_{p}\conj{g}_{q}y_{\phi_{pq}} + n_{pq}.
\end{equation}
Eq.~\eqref{eq:vis_red} can also be expressed in the following vector form 
\begin{equation}
\bd = \bv + \bn, 
\end{equation}
where 
\begin{eqnarray}
 \bd &=& [d_{pq}] \big \} ~ [pq] = 1\cdots B ~ (p<q),\nonumber\\
 \bv &=& [g_p y_{\phi_{pq}} \conj{g}_q]  \big \} ~ [pq] = 1\cdots B ~ (p<q),\nonumber\\
 \bn &=& [n_{pq}]  \big \} ~ [pq] = 1\cdots B ~ (p<q).\label{eq:vec_definitions}
\end{eqnarray}
In Eq.~\eqref{eq:vec_definitions}, $B$ denotes the number of baselines in the array.

We will be using the following SNR (signal-to-noise ratio) definition in this paper  
\begin{equation}
\textrm{SNR} = \frac{<\bv\odot\conj{\bv}>_{t,pq}}{<\bn\odot\conj{\bn}>_{t,pq}}, 
\end{equation}
where $<*>_{t,pq}$ denotes averaging over time and baselines and $\odot$ denotes the Hadamard product. The definition we use here is similar 
to the SNR definitions of \citet{Liu2010,Marthi2014}.

\section{Historical Overview}

Write a little about how redundant calibration developed, a basic lit review.
Maybe write a little bit about LOGCAL here as well.

\subsection{EW regular-layout}
Describe the EW-layout and its redundant baselines

\subsection{LINCAL}
\begin{equation}
c_{ij} = y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}
\end{equation}

\begin{eqnarray}
\frac{c_{ij}}{\partial \eta_i} &=& y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\ 
\frac{c_{ij}}{\partial \varphi_i} &=&  -i y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\
\frac{c_{ij}}{\partial \eta_j} &=& y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\ 
\frac{c_{ij}}{\partial \varphi_j} &=&  i y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\
\frac{y_{i-j}}{\partial \varphi_j} &=&  e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}
\end{eqnarray}

The Wirtinger derivative is used in the last equation.

\begin{eqnarray}
c_{ij} &\approx& c_{ij}^0 + \Delta \eta_i(y_{i-j}^0 e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}) + \Delta \eta_j(y_{i-j}^0 e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0})\\
&& -i\Delta\varphi_i (y_{i-j}^0e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}) +i\Delta\varphi_j (y_{i-j}^0e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0})\\
&& y_{i-j}^1(y_{i-j}^0e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0})\\
&\approx&  c_{ij}^0 + e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_i+ \Delta \eta_j - i\Delta\varphi_i + i\Delta\varphi_i))
\end{eqnarray}

\begin{equation}
\label{eq:residual}
r_{ij} = \delta_{ij} = c_{ij}-c_{ij}^0 = e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_i+ \Delta \eta_j - i\Delta\varphi_i + i\Delta\varphi_i)) 
\end{equation}

Eq.~\ref{eq:residual} allows us to construct the following linear system:

\begin{equation}
\boldsymbol{J}[\boldsymbol{\Delta \eta},\boldsymbol{\Delta \varphi},\boldsymbol{\Delta y}]^T = \boldsymbol{r}, 
\end{equation}

where $\boldsymbol{J}$ is equal to 
\begin{equation}
\boldsymbol{J} = \bigg [\overbrace{\frac{c_{pq}}{\partial \eta_i}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial \varphi_i}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial y_{t}}}^{t=1\cdots r_s} \bigg ] \bigg \} [pq] = 1\cdots N_{\textrm{b}} (p<q) 
\end{equation}
or
\begin{equation}
\boldsymbol{J} = \bigg [\frac{c_{pq}}{\partial \boldsymbol{\eta}},~\frac{c_{pq}}{\partial \boldsymbol{\varphi}},~\frac{c_{pq}}{\partial \boldsymbol{y}} \bigg ] \bigg \} [pq] = 1\cdots N_{\textrm{b}} (p<q). 
\end{equation}

The last column is again a Wirtinger derivative.

Which means we can obtain the least-squares estimate as follows:

\begin{equation}
[\boldsymbol{\Delta \eta},\boldsymbol{\Delta \varphi},\boldsymbol{\Delta y}]^T = [\boldsymbol{J}^H\boldsymbol{J}]^{-1}\boldsymbol{J}^H\boldsymbol{r}.
\end{equation}

What is interesting about LINCAL is that it seems to be a combination of standard least-squares (splitting the problem into amplitude and phase) and 
complex least-squares. The gains are solved with a standard approach (splitting into amp and phase) and the redundant visibilities use a Wirtinger derivative. What is strange though is that in normal complex optimization, if we optimize towards a complex variable x we also need to optimize towards
its complex conjugate. My best guess is that LINCAL works in any case since the conjugate redundant baselines are not used to calibrate ($p<q$). The only difference between the framework I was using is that
we regard the whole problem as complex, we construct our Jacobian by calculating the derivative towards $g$ and its conjugate as well as $y$ and its conjugate - Smirnov \& Tasse (2016).

Now that the relation between LINCAL and GN has been shown we can start focusing on the speedup that the conjugate gradient method provides as Liu (2010) originally proposed.
Also worth checking: whether using an LM step instead of a GN step improves your convergence properties.
Briefly review LINCAL here and show that it is equivalent to the GN algorithm.

\section{Complex Optimization}
\subsection{Scalar Calibration}
Scalar direction independent calibration is achieved through solving the following minimization problem:
\begin{equation}
\min_{\bg}\|\br\| = \min_{\bg}\|\bd - \bv(\bg)\|, 
\end{equation}
where
\begin{enumerate}
 \item $\bd$ is the observed visibility vector, i.e. 
 \begin{equation}
  \bd = [d_{pq}] \big \} ~ [pq] = 1\cdots N ~ (p<q).
 \end{equation}
 The vector entry $d_{pq}$ denotes the observed visibility associated with the baseline formed by antenna $p$ and $q$, while $N$ denotes the number or antennas
 in the array. 
 \item $\bg = [g_1,g_2,\cdots,g_N]$ represents the instrumental complex antenna gain vector.
 \item $\bv(\bg)$ denotes the predicted visibility vector, i.e.
 \begin{equation}
  \bv(\bg) = [g_p y_{pq}\conj{g}_q]  \big \} ~ [pq] = 1\cdots N ~ (p<q).
 \end{equation}
 In Eq... $y_{pq}$ denotes the visibility which was created from a skymodel and is associated with antenna $p$ and $q$. [Need to rw this sentence] 
\end{enumerate}

One of the simplests approaches one could use to solve eq. is to use least-squares optimization which we can only employ if we are able to calculate the gradient of $\bv(\bg)$,
implying that $\bv(\bg)$ needs to be analytic in $\bg$ (the Taylor series expansion of $\bv(\bg)$ exists).
The problem we face if we try to naivly apply least-squares to eq is that $\bv(\bg)$ is not analytic in its argument, since $\frac{\partial \conj{z}}{\partial z}$ (where 
$z \in \mathbb{C}$) does not exist.

The easiest way to circumvent this problem is to recast our problem in such a way that the predicted visibility vector becomes analytic in its argument.
The most widely used way to accomplish this is to restate eq as
\begin{equation}
\min_{\hat{\bg}}\|\hat{\br}\| = \min_{\hat{\bg}}\|\hat{\bd} - \hat{\bv}(\hat{\bg})\|, 
\end{equation}
where $\hat{\bg} = \big[\mathcal{R}\{\bg\},\mathcal{I}\{\bg\}\big]^T$, still need to define the other vectors properly....
Note that $\hat{\bv}(\hat{\bg})$ is now analytic in its argument and we can proceed and use least squares to estimate $\bg$.

\citet{Smirnov2015} have recently shown that calibration can also be performed using the following complex framework: 
\begin{equation}
\min_{\breve{\bg}}\|\breve{\br}\| = \min_{\breve{\bg}}\|\breve{\bd} - \breve{\bv}(\breve{\bg})\|, 
\end{equation}
where $\breve{\bg} = \big[\bg,\conj{\bg}]^T$, still need to define the other vectors properly....
In the above formulation v breve is analytic in its argument and its complex conjugate as a whole under the assumption that the Wirtinger derivative
is used to define the gradient operator. As v breve is analytic we will are also able to apply the least squares method to .

The Wirtinger derivatives are defined as
\begin{eqnarray}
\frac{\partial}{\partial z} = \frac{1}{2}\left ( \frac{\partial}{\partial x} -  i \frac{\partial}{\partial y} \right ),&~&\frac{\partial}{\partial \conj{z}} = \frac{1}{2}\left ( \frac{\partial}{\partial x} +  i \frac{\partial}{\partial y} \right ) 
\end{eqnarray}
where $z = x + iy$ and $x,y \in \mathbb{R}$. It follows from the above that $\frac{\partial z}{\partial z} = 1,\frac{\partial z}{\partial \conj{z}} = 0,\frac{\partial \conj{z}}{\partial z} = 0$ and $\frac{\partial \conj{z}}{\partial \conj{z}} = 1$.

%The main difference between the traditional approach and the 
%complex optimization framework proposed by \citet{Smirnov2015} is that in their framework they no longer treat the real and imaginary parts of the residual separately which enables them to directly solve for the complex antenna gains themselves (as well as their
%complex conjugates).

Recall that most optimization techniques require a gradient operator to enable it to function properly. The main reason why we have traditionally treated 
the real and imaginary parts of the residual separately is due to the fact that $f(z) = \conj{z}$ is not a differentiable complex function where $z\in\mathbb{C}$ and $\conj{(*)}$ denotes conjugation, i.e. $\frac{df(z)}{dz}$
does not exist, if we apply the standard definition of differentiation:
\begin{equation}
x
\end{equation}

The main reason that we do not need to treat the real and imaginary parts of the residual separately any more is due to the 
fact

%A standard way to optimize a complex function $f(\bz)$ of $n$ complex variables $\bz\in\mathbb{Z}^n$ is to recast $f$ as  optimize the real and imaginary parts of  separately.

The main difference between complex optimization and the more traditional
approach to calibration is that in complex calibration we

\subsection{Redundant Calibration}
In eq a sky model is required to generate the predicted visibilities $y_{pq}$ as we have too few equations to also solve for them and the gains. If we use a redundant layout
we can reduce the number of unknowns so that we are able to solve for the sky and the gains simultaneously.

\subsubsection{Antenna indices to redundancy index mapping}
If an array is redundant its redundant baselines sample the exact same visibilites in the $uv$-plane. Recall, that we denote the true observed visibility associated with baseline $pq$ and 
baseline $rs$ with $y_{pq}$ and $y_{rs}$, respectively . However if baseline $pq$ and $rs$ represent the same redundant baseline then $y_{pq} = y_{rs}$. Another
usefull indexing strategy would therefore be to use redundant baseline indices instead of antenna pairs to label the true observed visibilities so that we can identify 
which baselines observe the same visibilities barring instrumental and environmental errors.

Let $\phi_{ij}:\mathbb{N}^2\rightarrow\mathbb{N}^+$ be the mapping that maps antenna indices to the unique redundant baseline indices of the array. This mapping is not symmetric as 
$\phi_{ij} = 0$ if $i>j$. For an east-west regular array we can construct a simple analytic expression for this mapping:
\begin{equation}
\widehat{\phi_{ij}} = 
\begin{cases}
j-i&\textrm{if}~i<j\\
0&\textrm{otherwise}
\end{cases}.
\end{equation}
As will become apparent later the following symmetric variant of $\phi_{ij}$ will turn out to be quite useful later on:
\begin{equation}
\zeta_{ij} = 
\begin{cases}
\phi_{ij}~\textrm{if}~i\leq j\\
\phi_{ji}~\textrm{if}~i>j
\end{cases}.
\end{equation} 
It is also crucial to be able to determine the antenna index of the antenna which was used to construct a specific redundant baseline if one is given the other antenna index and the redundant baseline index. This can be calculated via the following two functions:
\begin{equation}
\xi_{ij} = 
\begin{cases}
p~\textrm{if}~\exists! ~ p \in \mathbb{N} ~ s.t. ~(\phi_{pi} = j)\\
0~\textrm{otherwise}
\end{cases},
\end{equation}
and
\begin{equation}
\psi_{ij} = 
\begin{cases}
q~\textrm{if}~\exists! ~ q \in \mathbb{N} ~ s.t. ~(\phi_{iq} = j)\\
0~\textrm{otherwise}
\end{cases}
\end{equation}
We use $\xi_{ij}$ to retrieve the index of the second antenna if we know the redundant baseline index and the index of the first antenna used to construct the baseline in question, while we use $\psi_{ij}$ to return the index of the first antenna if we know the 
index of the second antenna instead. \textbf{NB::} I still need to plot $\zeta_{ij}$ for different antenna layouts.

Describe complex optimization, Jacobian, Hessian etc... Benefits drawbacks compared to LINCAL... etc..

\subsubsection{Redundant calibration}
In the case of a redundant array we can not only solve for the gains $\bg$ but we can also solve for the redundant unknown visibilities $\by$. Where $\by = [y_1,\cdots,y_R]$, $y_i$ denotes 
a unique redundant visibility and $R$ is the maximum value of the mapping $\phi_{ij}$. The reason we can now also solve for $\by$ is because we have reduced the number of 
unknowns from $N+B$ to $N+R$, where $N+R$ is now less than $B$ due to the array redundancy. We can now reformulate the calibration problem for an redundant array as follow:
\begin{equation}
\min_{\bz}\|\br\| = \min_{\bz}\|\bd - \bv(\bz)\|, 
\end{equation}
where $\bz = [\bg,\by]^T$.

If we now follow the same extension strategy of \citet{Smirnov2015} we can rewrite eq to
\begin{equation}
\min_{\breve{\bz}}\|\breve{\br}\| = \min_{\bz}\|\breve{\bd} - \breve{\bv}(\breve{\bz})\|, 
\end{equation}
where $\breve{\bz} = [\bz,\conj{\bz}]^T$ still need to define the remaining variables properly.

We can now solve the above equation by using least-squares, which entails deriving the Gauss-Newton (GN) and Levenberg-Marquerdt (LM) update rules.
(NEED TO WRITE THIS MUCH BETTER)

\begin{equation}
\bJ = \frac{\partial\bv}{\partial\bz}~\bJ^* = \frac{\partial \bv}{\partial \breve{\bz}} 
\end{equation}

\begin{equation}
\bJ = \begin{bmatrix}
       \bJ & \bJ^* \\
       \conj{\bJ^*} & \conj{\bJ}  
      \end{bmatrix}
\end{equation}

\begin{equation}
 \Delta \breve{\bz} = \begin{bmatrix}\Delta \bz\\ \Delta \conj{\bz} \end{bmatrix} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\br_k}
\end{equation}

\begin{equation}
 \Delta \breve{\bz} = \begin{bmatrix}\Delta \bz\\ \Delta \conj{\bz} \end{bmatrix} = (\bJ^H\bJ + \lambda \bD)^{-1}\bJ^H\breve{\br_k}
\end{equation}

\subsection{Jacobian derivation}
\subsection{Hessian derivation}
\begin{equation}
\boldsymbol{H} = 
\begin{bmatrix}
\boldsymbol{A} & \boldsymbol{B}\\
\boldsymbol{B}^* & \boldsymbol{A}^*
\end{bmatrix}
\end{equation}

With

\begin{equation}
\boldsymbol{A} = 
\begin{bmatrix}
\boldsymbol{C} & \boldsymbol{D}\\
\boldsymbol{D^H} & \boldsymbol{E}
\end{bmatrix}
\end{equation}

\begin{equation}
\boldsymbol{B} = 
\begin{bmatrix}
\boldsymbol{F} & \boldsymbol{G}\\
\boldsymbol{G}^T & \boldsymbol{0}
\end{bmatrix}
\end{equation}

\begin{equation}
[\boldsymbol{C}]_{ij} = 
\begin{cases}
 \sum_{k \neq i} \left | g_k \right |^2 \left | y_{\zeta_{ik}} \right |^2 & \textrm{if} ~ i=j\\
 0 & \textrm{otherwise}
\end{cases}
\end{equation}

\begin{equation}
[\boldsymbol{D}]_{ij} = 
\begin{cases}
 g_i y_j^*  \left | g_{\psi_{ij}} \right |^2  & \textrm{if} ~ \psi_{ij}\neq0\\
 0 & \textrm{otherwise}
\end{cases}
\end{equation}

\begin{equation}
[\boldsymbol{E}]_{ij} = 
\begin{cases}
 \sum_{pq \in \mathcal{I}_i} \left | g_p \right |^2 \left | g_q \right |^2  & \textrm{if} ~ i=j\\
 0 & \textrm{otherwise}
\end{cases}
\end{equation}

\begin{equation}
[\boldsymbol{F}]_{ij} = 
\begin{cases}
 g_i g_j  \left | y_{\zeta_{ij}} \right |^2  & \textrm{if} ~ (i \neq j)\\
 0 & \textrm{otherwise}
\end{cases}
\end{equation}

\begin{equation}
[\boldsymbol{G}]_{ij} = 
\begin{cases}
 g_i y_j  \left | g_{\xi_{ij}} \right |^2  & \textrm{if} ~ (\xi_{ij}\neq0)\\
 0 & \textrm{otherwise}
\end{cases}
\end{equation}

Moreover,

\begin{equation}
\mathcal{I}_i = \left\{pq\in\mathbb{N}^2|(\phi_{pq} = i) \right\}.
\end{equation}

The dimensions of the matrices:

\begin{enumerate}
\item $\boldsymbol{H}$: is a $(4N-2)\times(4N-2)$ or a $P \times P$ matrix. $N$ denotes the number of antennas and $P$ denotes the number of parameters. This matrix
contains $6N^2 - 2N - 2$ non-zero entries. It contains $16N^2 - 16N + 4$ entries. We can construct this matrix with $4N^2 -  4N$ or $\frac{1}{4} P^2 -1$ elementary operations.
\item $\boldsymbol{A}$: is a $(2N-1)\times(2N-1)$ matrix. This matrix has $N^2 + N - 1$ non-zero entries. We can construct this matrix with $2\frac{1}{2} (N^2 -  N)$ elementary operations.  
\item $\boldsymbol{B}$: is a $(2N-1)\times(2N-1)$ matrix. This matrix has $2N^2 - 2N$ non-zero entries. We can construct this matrix with $1\frac{1}{2} (N^2 -  N)$ elementary operations.
\item $\boldsymbol{C}$: is a $N\times N$ matrix. This matrix has $N$ non-zero entries. We can construct this matrix with $N^2-N$ elementary operations.
\item $\boldsymbol{D}$: is a $N \times (N-1)$ matrix. This matrix has $\frac{1}{2} (N^2 -  N)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations. 
\item $\boldsymbol{E}$: is a $(N-1) \times (N-1)$ matrix. This matrix has $(N-1)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.  
\item $\boldsymbol{F}$: is a $N \times N$ matrix. This matrix has $N^2 - N$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.
\item $\boldsymbol{G}$: is a $N \times (N-1)$ matrix. This matrix has $\frac{1}{2} (N^2 -  N)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.
\item $\boldsymbol{0}$: is a $(N-1) \times (N-1)$ all zero matrix. 
\end{enumerate}

We can construct $\boldsymbol{C}-\boldsymbol{G}$ in $O(N^2)$

The sparsity ratio of $\boldsymbol{H}$ is equal to
\begin{equation}
\gamma_N = \frac{10N^2-14N+6}{16N^2-16N+4} 
\end{equation}

The asymptotic sparsity ratio of $\boldsymbol{H}$ is equal to

\begin{equation}
\gamma = \lim_{N\rightarrow \infty}\frac{5N^2-7N+3}{8N^2-8N+2} = \frac{5}{8}
\end{equation}

Asymptotically the computational cost of the matrix-vector product between $\boldsymbol{H}$ and a vector of appropriate size is $O(P^2)$. However, the computational cost converges 
very slowly to its asymptotic value, and in general is equal to $O(P^{k_N})$, where $k_N < 2$.

\begin{eqnarray}
P^{k_N}_N &=& (1 - \gamma_N)P^2_N\\
k_N &=& \log_{P_N}(1 - \gamma_N) + 2\\
k &=& \lim_{N\rightarrow \infty} \log_{P_N}(1 - \gamma_N) + 2 = 2
\end{eqnarray}

Moreover,

\begin{eqnarray}
P^{c_N}_N &=& \frac{1}{4} P_N^2 -1\\
c_N &=& \log_{P_N}\left (\frac{1}{4} P_N^2-1 \right )\\
c &=& \lim_{N\rightarrow \infty} \log_{P_N}\left (\frac{1}{4} P_N^2-1\right ) = 2
\end{eqnarray}

NB $P_c$ IS WRONG HERE DID NOT TAKE INTO ACCOUNT THE CONJUGANTION OF $\boldsymbol{B}$ (need multiply its computational complexity by two).

\subsection{Jacobian-residual product}

\subsection{Computational complexity using CG}
\subsubsection{Sparsity}
\subsubsection{Condition number}

\begin{enumerate}
 \item Plot eigenvalue distribution for different SNR and antenna numbers and different lambda values
 \item Plot computational complexity bound plus error reduction for different setups
 \item Plot number of iterations required by CG as a function of number of antennas
 \item Plot redundant baseline visibilities to show that they are approximately equal.
\end{enumerate}

NB:: Repeat the second last plot for the positive-semi definite case (GN), showing that even if the Hessian is singular we get fast convergence.

\section{Results: Regualar square grid}
If there is time

\section{Results: Hexanol grid}
NB:: Need to test if it works for this redundant layout.

\section{Conjugate Gradient method}
brief description and lit review of the method.
computational complexity of the algorithm.
sparsity and condition number.
% \section{Methods, Observations, Simulations etc.}
% 
% Normally the next section describes the techniques the authors used.
% It is frequently split into subsections, such as Section~\ref{sec:maths} below.
% 
% \subsection{Maths}
% \label{sec:maths} % used for referring to this section from elsewhere
% 
% Simple mathematics can be inserted into the flow of the text e.g. $2\times3=6$
% or $v=220$\,km\,s$^{-1}$, but more complicated expressions should be entered
% as a numbered equation:
% 
% \begin{equation}
%     x=\frac{-b\pm\sqrt{b^2-4ac}}{2a}.
% 	\label{eq:quadratic}
% \end{equation}
% 
% Refer back to them as e.g. equation~(\ref{eq:quadratic}).
% 
% \subsection{Figures and tables}
% 
% Figures and tables should be placed at logical positions in the text. Don't
% worry about the exact layout, which will be handled by the publishers.
% 
% Figures are referred to as e.g. Fig.~\ref{fig:example_figure}, and tables as
% e.g. Table~\ref{tab:example_table}.
% 
% % Example figure
% \begin{figure}
% 	% To include a figure from a file named example.*
% 	% Allowable file formats are eps or ps if compiling using latex
% 	% or pdf, png, jpg if compiling using pdflatex
% 	\includegraphics[width=\columnwidth]{example}
%     \caption{This is an example figure. Captions appear below each figure.
% 	Give enough detail for the reader to understand what they're looking at,
% 	but leave detailed discussion to the main body of the text.}
%     \label{fig:example_figure}
% \end{figure}
% 
% % Example table
% \begin{table}
% 	\centering
% 	\caption{This is an example table. Captions appear above each table.
% 	Remember to define the quantities, symbols and units used.}
% 	\label{tab:example_table}
% 	\begin{tabular}{lccr} % four columns, alignment for each
% 		\hline
% 		A & B & C & D\\
% 		\hline
% 		1 & 2 & 3 & 4\\
% 		2 & 4 & 6 & 8\\
% 		3 & 5 & 7 & 9\\
% 		\hline
% 	\end{tabular}
% \end{table}
% 
% 
% \section{Conclusions}
% 
% The last numbered section should briefly summarise what has been done, and describe
% the final conclusions which the authors draw from their work.

\section*{Acknowledgements}

The Acknowledgements section is not numbered. Here you can thank helpful
colleagues, acknowledge funding agencies, telescopes and facilities used etc.
Try to keep it short.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%

% The best way to enter references is to use BibTeX:

%\bibliographystyle{mnras}
%\bibliography{example} % if your bibtex file is called example.bib

\bibliographystyle{mnras}
\bibliography{paper}
% Alternatively you could enter them by hand, like this:
% This method is tedious and prone to error if you have lots of references
%\begin{thebibliography}{99}
%\bibitem[\protect\citeauthoryear{Author}{2012}]{Author2012}
%Author A.~N., 2013, Journal of Improbable Astronomy, 1, 1
%\bibitem[\protect\citeauthoryear{Others}{2013}]{Others2013}
%Others S., 2012, Journal of Interesting Stuff, 17, 198
%\end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%

\appendix

\section{Some extra material}

If you want to present additional material which would interrupt the flow of the main paper,
it can be placed in an Appendix which appears after the list of references.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}
\end{document}

% End of mnras_template.tex

### 2016-09-26 paper.tex
Started to load antenna layout figures

@@
 
-\title[Dust envelopes around RV Tauri stars]{Dust envelopes around RV Tauri stars}
+\title[Unknown]{TBD}
 \author[A. V. Raveendran and A. N. Other]{A. V. Raveendran$^{1}$\thanks{E-mail:
@@
 \begin{abstract}
-In the {\it IRAS\/} [12]--[25],  [25]--[60] colour--colour diagram, RV
-Tauri stars are found to populate cooler temperature regions
-$(T<600\,\rmn{K})$, distinctly different from those occupied by  the
-oxygen and carbon Miras. The {\it IRAS\/} fluxes are  consistent with the
-dust density in the envelope varying as the inverse square of the
-radial distance, implying that the grain formation processes in these
-objects are most probably continuous and not sporadic. It is found that
-the spectroscopic subgroups A and B are well separated in the
-far-infrared two-colour diagram, with group B objects having
-systematically cooler dust envelopes. We interpret this as being due to
-a difference in the nature of grains, including the chemical
-composition, in the two cases.
+TBD
 \end{abstract}
@@
 \centering
-\subfigure[full-complex:   Real]
-{\includegraphics[width=0.33\textwidth]{./HEX_lay.pdf}\label{fig:stef_r}}
-\subfigure[full-complex:  Imag]
-{\includegraphics[width=0.33\textwidth]{./SQR_lay.pdf}\label{fig:stef_i}}
-\subfigure[full-complex: ghost pattern]
-{\includegraphics[width=0.33\textwidth]{./REG_lay.pdf}\label{fig:full_pat_45}}
+\subfigure[Hexagonal layout]
+{\includegraphics[width=0.33\textwidth]{./HEX_lay.pdf}\label{fig:HEX_lay}}
+\subfigure[Square layout]
+{\includegraphics[width=0.33\textwidth]{./SQR_lay.pdf}\label{fig:SQR_lay}}
+\subfigure[Regular east-west layout]
+{\includegraphics[width=0.33\textwidth]{./REG_lay.pdf}\label{fig:REG_lay}}
 
-\subfigure[phase-only:   Real]
-{\includegraphics[width=0.33\textwidth]{./HEX_phi.pdf}\label{fig:phase_r}}
-\subfigure[phase-only:   Imag]
-{\includegraphics[width=0.33\textwidth]{./SQR_phi.pdf}\label{fig:phase_i}}
-\subfigure[phase-only: ghost pattern]
-{\includegraphics[width=0.33\textwidth]{./REG_phi.pdf}\label{fig:phase_pat_45}}
-\caption{scenario. \label{fig:ghost_pattern_bas_45}}
+\subfigure[Hexagonal: $\zeta_{pq}$]
+{\includegraphics[width=0.33\textwidth]{./HEX_phi.pdf}\label{fig:HEX_phi}}
+\subfigure[Square: $\zeta_{pq}$]
+{\includegraphics[width=0.33\textwidth]{./SQR_phi.pdf}\label{fig:SQR_phi}}
+\subfigure[Regular east-west: $\zeta_{pq}$]
+{\includegraphics[width=0.33\textwidth]{./REG_phi.pdf}\label{fig:REG_phi}}
+\caption{In the top panel we have three different redundant antenna layouts. In the bottom panel we have the symmetric redundancy geometry mapping $\zeta_{ij}$
+of each of the layouts in the top panel.}
 \end{figure*}

### 2016-09-27 paper.tex
added more to geometry matrix text

@@
 
-\subsection{Redundant Array Geometry Mapping}
+\subsection{Redundant Array Geometry Function}
 If an array is redundant then its redundant baselines sample the exact same visibilites in the $uv$-plane, i.e. if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$. We can
@@
 
-Let $\phi_{ij}:\mathbb{N}^2\rightarrow\mathbb{N}^+$ be the mapping that maps composite antenna pairs to the unique redundant baseline indexes of an array. This mapping is not symmetric as 
-$\phi_{ij} = 0$ if $i>j$. It is possible to construct a simple analytic expression for $\phi_{ij}$ when our array is in an east-west regular configuration:
-\begin{equation}
-\phi_{ij} = 
-\begin{cases}
-j-i&\textrm{if}~i<j\\
-0&\textrm{otherwise}
-\end{cases}.
-\end{equation}
-It becomes increasingly more difficult to construct analytic expressions of $\phi_{ij}$ for other more complicated array layouts.
-
-We can also define the following symmetric variant of $\phi_{ij}$:
+Let $\phi_{ij}:\mathbb{N}^2\rightarrow\mathbb{N}^+$ be the function that maps composite antenna pairs to the unique redundant baseline indexes of an array. We refer to this function 
+as the redundant array geometry function in this paper. This mapping is not symmetric as 
+$\phi_{ij} = 0$ if $i>j$. We can also define the following symmetric variant of $\phi_{ij}$:
 \begin{equation}
@@
 \end{equation} 
+It is possible to construct a simple analytic expression for $\zeta_{ij}$ when our array is in an east-west regular configuration: $\zeta_{ij} = |j-i|$. 
+It becomes increasingly more difficult to construct analytic expressions of $\zeta_{ij}$ for other more complicated array layouts. The empirically constructed symmetric geometry
+functions of three different redundant layouts are depicted in Fig.~\ref{fig:geometry_function}. We denote the range of $\zeta_{ij}$ with $\mathcal{R}(\zeta_{ij})$. The maximal element 
+that $\zeta_{ij}$ can ascertain is denoted by $L$ and can be construed as the maximal number of unique redundant baselines which can be formed for a given 
+array layout. Alternatively, we can interpret the symmetric geometry functions 
+in Fig.~\ref{fig:geometry_function} as matrices. In this alternative paradigm, $\zeta_{ij}$ denotes a matrix entry instead. The dimension of these so called geometry 
+matrices are $N\times N$, where $N$ denotes the number of antennas in the array.
 
@@
 {\includegraphics[width=0.33\textwidth]{./REG_phi.pdf}\label{fig:REG_phi}}
-\caption{In the top panel we have three different redundant antenna layouts. In the bottom panel we have the symmetric redundancy geometry mapping $\zeta_{ij}$
-of each of the layouts in the top panel.}
+\caption{In the top panel we have three different redundant antenna layouts, namely a hexagonal (left), square (middle) and a regular east-west (right) layout. 
+In the bottom panel we have the respective symmetric redundancy geometry functions for the layouts on the top panel. We used 91 antennas
+to construct the the hexagonal layout, while a 100 antennas where used in the square and east-west layouts. The maximal amount of redundant baselines that can be formed for 
+the hexagonal, square and east-west layouts which are depicted in the top panel are 165, 180 and 99 respectively.\label{fig:geometry_function}}
 \end{figure*}
@@
 \end{equation}
-
-We use $\xi_{ij}$ to retrieve the first antenna index of the composite antenna pair associated with a particular baseline if the index of the first antenna in the composite antenna pair and the unique redundant baseline index are known, while we use $\psi_{ij}$ to obtain
+We use $\xi_{ij}$ to retrieve the first antenna index of the composite antenna pair associated with a particular baseline, if the index of the first antenna in the composite antenna pair and the unique redundant baseline index are known, while we use $\psi_{ij}$ to obtain
 a similar result; the second antenna index of the composite antenna pair. \textbf{NB::} I still need to plot $\zeta_{ij}$ for different antenna layouts.

### 2016-09-27 paper.tex
added paper

@@
 \usepackage{graphicx}
+\usepackage{mathtools}
+
 \newcommand{\bz}{\bmath{z}}
+\newcommand{\bA}{\bmath{A}}
 \newcommand{\br}{\bmath{r}}
@@
 \newcommand{\conj}[1]{\overline{#1}}
+
+\newcommand\coolover[2]{\mathrlap{\smash{\overbrace{\phantom{%
+    \begin{matrix} #2 \end{matrix}}}^{\mbox{$#1$}}}}#2}
+
+\newcommand\coolunder[2]{\mathrlap{\smash{\underbrace{\phantom{%
+    \begin{matrix} #2 \end{matrix}}}_{\mbox{$#1$}}}}#2}
+
+\newcommand\coolleftbrace[2]{%
+#1\left\{\vphantom{\begin{matrix} #2 \end{matrix}}\right.}
+
+\newcommand\coolrightbrace[2]{%
+\left.\vphantom{\begin{matrix} #1 \end{matrix}}\right\}#2}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 If an array is redundant then its redundant baselines sample the exact same visibilites in the $uv$-plane, i.e. if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$. We can
-therefore also use the following useful indexing strategy: We can use singular redundant baseline indexes instead of composite antenna pairs to label the true observed visibilities.
+therefore also use the following useful indexing strategy: we can use singular redundant baseline indexes instead of composite antenna pairs to label the true observed visibilities.
 
@@
 \end{equation} 
-It is possible to construct a simple analytic expression for $\zeta_{ij}$ when our array is in an east-west regular configuration: $\zeta_{ij} = |j-i|$. 
+It is possible to construct a simple analytic expression for $\zeta_{ij}$ when our array is in an east-west regular configuration, namely $\zeta_{ij} = |j-i|$. 
 It becomes increasingly more difficult to construct analytic expressions of $\zeta_{ij}$ for other more complicated array layouts. The empirically constructed symmetric geometry
@@
 to construct the the hexagonal layout, while a 100 antennas where used in the square and east-west layouts. The maximal amount of redundant baselines that can be formed for 
-the hexagonal, square and east-west layouts which are depicted in the top panel are 165, 180 and 99 respectively.\label{fig:geometry_function}}
+the hexagonal, square and east-west layouts in the top panel are 165, 180 and 99 baselines respectively.\label{fig:geometry_function}}
 \end{figure*}
 
-If the mapping $\phi_{ij}$ is known and one is given one of the two antenna indexes that together form a redundant baseline as well as the redundant baseline index itself then it is possible 
-to calculate the unknown antenna index. This can be calculated via the following two expressions:
+If the function $\phi_{ij}$ is known and one is given one of the two antenna indexes that together form a redundant baseline as well as the redundant baseline index itself then it is possible 
+to calculate the unknown antenna index. This can be computed via the following two expressions:
 \begin{equation}
@@
 
-If the array is redundant and the mapping $\phi_{ij}$ is known we can rewrite Eq.~\eqref{eq:vis_definition} as
+\subsection{Signal-to-noise Ratio}
+If the array is redundant and the function $\phi_{ij}$ is known we can rewrite Eq.~\eqref{eq:vis_definition} as
 \begin{equation}
@@
 \end{equation}
-where $<*>_{t,pq}$ denotes averaging over time and baselines and $\odot$ denotes the Hadamard product. The definition we use here is similar 
-to the SNR definitions of \citet{Liu2010,Marthi2014}.
+where $<*>_{t,pq}$ denotes averaging over time and baselines and $\odot$ denotes the Hadamard product.
+The definition we use here is based on the SNR definitions used in \citet{Liu2010} and \citet{Marthi2014}.
+
+\subsection{LOGCAL}
+One of the best known algorithms that is currently used to perform redundant calibration involves the use of logarithms. If we ignore the noise term then we can 
+rewrite Eq.~\eqref{eq:vis_red} as
+\begin{eqnarray}
+\ln |d_{ij}| &=& \ln |g_i| + \ln |g_j| + \ln |y_{\phi_{ij}}| \label{eq:logcal_amp}\\
+\angle d_{ij} &=& \angle g_i - \angle g_j + \angle \phi_{ij} \label{eq:logcal_phase}
+\end{eqnarray}
+
+We can easily express Eq.~\eqref{eq:logcal_amp} with the following linear system
+\begin{equation}
+\widetilde{\bd} = \bA\bz, 
+\end{equation}
+which has the following least-squares solution
+\begin{equation}
+\bz = (\bA^T\bA)^{-1}\bA^T\widetilde{\bd},
+\end{equation}
+where
+\begin{equation}
+\widetilde{\bd} = [\ln |d_{pq}|] \big \} ~ [pq] = 1\cdots B ~ (p<q),\nonumber\\ 
+\end{equation}
+and
+\begin{equation}
+\bA = 
+\begin{bmatrix}
+\coolover{{\scriptstyle j=1,\ldots,N}}{1^j_p\vee1^j_q} & \coolover{{\scriptstyle k=1,\ldots,L}\!\!\!\!}{~1_{\phi_{pq}}^k}
+\end{bmatrix}
+\begin{matrix}% matrix for right braces 
+\coolrightbrace{1_{\phi_{pq}}^k}{[pq] = 1,\ldots,B~(p<q)}
+\end{matrix}.
+\end{equation}
 

### 2016-09-27 paper.tex
LOGCAL
LINCAL

@@
 \begin{equation}
-\widetilde{\bd} = [\ln |d_{pq}|] \big \} ~ [pq] = 1\cdots B ~ (p<q),\nonumber\\ 
+\widetilde{\bd} = [\ln |d_{pq}|] \big \} ~ [pq] = 1,\cdots,B ~ (p<q),\nonumber\\ 
 \end{equation}
@@
 \begin{bmatrix}
-\coolover{{\scriptstyle j=1,\ldots,N}}{1^j_p\vee1^j_q} & \coolover{{\scriptstyle k=1,\ldots,L}\!\!\!\!}{~1_{\phi_{pq}}^k}
+\coolover{{\scriptstyle j=1,\ldots,N}}{\delta^j_p\vee\delta^j_q} & \coolover{{\scriptstyle k=1,\ldots,L}\!\!\!\!}{~\delta_{\phi_{pq}}^k}
 \end{bmatrix}
 \begin{matrix}% matrix for right braces 
-\coolrightbrace{1_{\phi_{pq}}^k}{[pq] = 1,\ldots,B~(p<q)}
+\coolrightbrace{\delta_{\phi_{pq}}^k}{[pq] = 1,\ldots,B~(p<q)}
 \end{matrix}.
+\end{equation}
+
+Moreover, $(*)^T$ denotes the transpose of its operand, $(*)^{-1}$ denotes matrix inversion, $\vee$ denotes the logical or operator and
+\begin{equation}
+\delta_y^x  = \begin{cases}
+          1 & \textrm{if}~x=y\\
+          0 & \textrm{otherwise}
+         \end{cases}.
+\end{equation}
+A similar linear system can be constructed for Eq.~\eqref{eq:logcal_phase}.
+
+\subsection{LINCAL}
+\begin{equation}
+c_{ij} = y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}
+\end{equation}
+
+\begin{eqnarray}
+\frac{c_{ij}}{\partial \eta_i} &=& y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\ 
+\frac{c_{ij}}{\partial \varphi_i} &=&  -i y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\
+\frac{c_{ij}}{\partial \eta_j} &=& y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\ 
+\frac{c_{ij}}{\partial \varphi_j} &=&  i y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\
+\frac{y_{i-j}}{\partial \varphi_j} &=&  e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}
+\end{eqnarray}
+
+The Wirtinger derivative is used in the last equation.
+
+\begin{eqnarray}
+c_{ij} &\approx& c_{ij}^0 + \Delta \eta_i(y_{i-j}^0 e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}) + \Delta \eta_j(y_{i-j}^0 e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0})\\
+&& -i\Delta\varphi_i (y_{i-j}^0e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}) +i\Delta\varphi_j (y_{i-j}^0e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0})\\
+&& y_{i-j}^1(y_{i-j}^0e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0})\\
+&\approx&  c_{ij}^0 + e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_i+ \Delta \eta_j - i\Delta\varphi_i + i\Delta\varphi_i))
+\end{eqnarray}
+
+\begin{equation}
+\label{eq:residual}
+r_{ij} = \delta_{ij} = c_{ij}-c_{ij}^0 = e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_i+ \Delta \eta_j - i\Delta\varphi_i + i\Delta\varphi_i)) 
+\end{equation}
+
+Eq.~\ref{eq:residual} allows us to construct the following linear system:
+
+\begin{equation}
+\boldsymbol{J}[\boldsymbol{\Delta \eta},\boldsymbol{\Delta \varphi},\boldsymbol{\Delta y}]^T = \boldsymbol{r}, 
+\end{equation}
+
+where $\boldsymbol{J}$ is equal to 
+\begin{equation}
+\boldsymbol{J} = \bigg [\overbrace{\frac{c_{pq}}{\partial \eta_i}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial \varphi_i}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial y_{t}}}^{t=1\cdots r_s} \bigg ] \bigg \} [pq] = 1\cdots N_{\textrm{b}} (p<q)
+\end{equation}
+or
+\begin{equation}
+\boldsymbol{J} = \bigg [\frac{c_{pq}}{\partial \boldsymbol{\eta}},~\frac{c_{pq}}{\partial \boldsymbol{\varphi}},~\frac{c_{pq}}{\partial \boldsymbol{y}} \bigg ] \bigg \} [pq] = 1\cdots N_{\textrm{b}} (p<q).
+\end{equation}
+
+The last column is again a Wirtinger derivative.
+
+Which means we can obtain the least-squares estimate as follows:
+
+\begin{equation}
+[\boldsymbol{\Delta \eta},\boldsymbol{\Delta \varphi},\boldsymbol{\Delta y}]^T = [\boldsymbol{J}^H\boldsymbol{J}]^{-1}\boldsymbol{J}^H\boldsymbol{r}.
 \end{equation}

### 2016-09-27 paper.tex
LINCAL revamp

@@
 \subsection{LINCAL}
+Assume that $\eta_i,\eta_j,\eta_{\phi_{ij}},\varphi_i,\varphi_j$ and $\varphi_{\phi_{ij}}$ are real valued variables.
+Let us once again ignore the noise term of Eq.~\eqref{eq:vis_red}.
+Substituting $g_i$ with $e^{\eta_i - i \varphi_i}$, $g_j$ with $e^{\eta_j - i \varphi_j}$ and $y_{\phi_{ij}}$ with $e^{\eta_{\phi_{ij}}- i \varphi_{\phi_{ij}}}$ in Eq.~\eqref{eq:vis_red} results in
 \begin{equation}
-c_{ij} = y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}
+d_{ij} =  e^{\eta_i - i \varphi_i} e^{\eta_{\phi_{ij}}- i \varphi_{\phi_{ij}}} e^{\eta_j + i \varphi_j}.
 \end{equation}
+The following partial derivatives are easily computed
 

### 2016-09-28 paper.tex
LINCAL partial
LINCAL again

--- 2016-09-27 paper.tex
@@
 \newcommand{\bD}{\bmath{D}}
+\newcommand{\bN}{\bmath{N}}
+\newcommand{\bM}{\bmath{M}}
 \newcommand{\conj}[1]{\overline{#1}}
@@
 \subsection{LINCAL}
+Assume that $\eta_i,\eta_j,\widetilde{\eta}_{\phi_{ij}},\varphi_i,\varphi_j$ and $\widetilde{\varphi}_{\phi_{ij}}$ are real valued variables.
+Substituting $g_i$ with $e^{\eta_i - i \varphi_i}$, $g_j$ with $e^{\eta_j - i \varphi_j}$ and $y_{\phi_{ij}}$ with $e^{\eta_{\widetilde{\phi}_{ij}}- i \widetilde{\varphi}_{\phi_{ij}}}$ in Eq.~\eqref{eq:vis_red} results in
 \begin{equation}
-c_{ij} = y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}
+\label{eq:lincal}
+d_{ij} =  e^{\eta_i - i \varphi_i} e^{\widetilde{\eta}_{\phi_{ij}}- i \widetilde{\varphi}_{\phi_{ij}}} e^{\eta_j + i \varphi_j} + n_{ij}.
 \end{equation}
+
+A possible approach one can use to solve the unknowns in Eq.~\eqref{eq:lincal} is to write down the Taylor expansion of the first term of the right hand side of Eq.~\eqref{eq:lincal} around some 
+initial fiducial guess of $\eta_i,\eta_j,\widetilde{\eta}_{\phi_{ij}},\varphi_i,\varphi_j$ and $\widetilde{\varphi}_{\phi_{ij}}$, which we denote with 
+$\eta_i^0,\eta_j^0,\widetilde{\eta}_{\phi_{ij}}^0,\varphi_i^0,\varphi_j^0$ and $\widetilde{\varphi}_{\phi_{ij}}^0$, and then to solve the perturbations from this 
+initial guess that minimize the difference between the right and left hand side of Eq.~\eqref{eq:lincal}. Once the perturbation values 
+have been computed it is trivial to obtain an estimate of $g_i$, $g_j$ and $y_{\phi_{ij}}$.
+
+If we Taylor expand the first term on the right hand side of Eq.~\eqref{eq:lincal} around a fiducial initial guess we obtain
+\begin{equation}
+\label{eq:taylor_expansion}
+d_{ij} = v_{ij}^0[1+\Delta \eta_i + \Delta \eta_j + i \varphi_i - i \Delta \varphi_j + \Delta \widetilde{\eta}_{\phi_{ij}} + i\Delta \widetilde{\eta}_{\phi_{ij}}] + n_{ij}, 
+\end{equation}
+where
+\begin{equation}
+v_{ij}^0 = e^{\eta_i^0 - i \varphi_i^0} e^{\widetilde{\eta}_{\phi_{ij}}^0 - i \widetilde{\varphi}_{\phi_{ij}}^0} e^{\eta_j^0 + i \varphi_j^0}, 
+\end{equation}
+and $\eta_i^0,\eta_j^0,\widetilde{\eta}_{\phi_{ij}}^0,\varphi_i^0,\varphi_j^0$ and $\widetilde{\varphi}_{\phi_{ij}}^0$ respectively denote the initial guess for $\eta_i,\eta_j,\widetilde{\eta}_{\phi_{ij}},\varphi_i,\varphi_j$ and $\widetilde{\varphi}_{\phi_{ij}}$.
+Eq.~\eqref{eq:taylor_expansion} can be used to construct the following linear system
+\begin{equation}
+\breve{\br} = \bJ\Delta\widetilde{\bz}, 
+\end{equation}
+where 
+\begin{equation}
+\breve{\br} = [d_{ij}-v_{ij}^0] \big \} ~ [ij] = 1\cdots 2B, 
+\end{equation}
+\begin{equation}
+\label{eq:jac_lincal}
+\bJ = \begin{bmatrix}
+      \bN & \bM\\
+      \conj{\bN} & \conj{\bM}
+      \end{bmatrix}
+\end{equation}
+and
+\begin{equation}
+\Delta\widetilde{\bz} = 
+\begin{bmatrix}
+\Delta \eta_i \\
+\Delta \varphi_j \\
+\Delta \widetilde{\eta}_k \\
+\Delta \widetilde{\varphi}_l \\
+\end{bmatrix}
+\begin{matrix}% matrix for right braces 
+\coolrightbrace{\eta_i}{i = 1\cdots N}\\
+\coolrightbrace{\eta_i}{j = 1\cdots N}\\
+\coolrightbrace{\eta_i}{k = 1\cdots L}\\
+\coolrightbrace{\eta_i}{l = 1 \cdots L}
+\end{matrix}.
+\end{equation}
+
+In Eq.~\eqref{eq:jac_lincal}
+
+\begin{eqnarray}
+\bN &=& 
+\begin{bmatrix}
+\coolover{\!\!\!\!\!\!\!{\scriptstyle j=1,\ldots,N}}{\frac{\partial v_{pq}}{\partial \eta_j}} & \coolover{\!\!\!\!{\scriptstyle k=1,\ldots,N}}{~\frac{\partial v_{pq}}{\partial \varphi_k}}
+\end{bmatrix}
+\begin{matrix}% matrix for right braces 
+\coolrightbrace{\frac{\partial v_{ij}}{\partial \eta_j}}{\scriptstyle [pq] = 1,\ldots,B~(p<q)}
+\end{matrix}\\
+&=& \begin{bmatrix}
+\coolunder{\!\!\!\!\!\!\!{\scriptstyle j=1,\ldots,N}}{n_{j,pq}} & \coolunder{\!\!\!\!{\scriptstyle k=1,\ldots,N}}{(-1)^{\delta_q^k}i n_{k,pq}}
+\end{bmatrix}
+\begin{matrix}% matrix for right braces 
+\coolrightbrace{\frac{\partial v_{ij}}{\partial \eta_j}}{\scriptstyle [pq] = 1,\ldots,B~(p<q)}
+\end{matrix}\\
+\end{eqnarray}
+
+Moreover,
+\begin{equation}
+n_{j,pq} = (\delta_{p}^{j}\vee\delta_{q}^{j}) v_{pq} 
+\end{equation}
+
+fiducial initial guess
+
+The following partial derivatives are now easily computed
 

### 2016-09-28 paper.tex
LINCAL

@@
 
+\begin{eqnarray}
+\bM &=& 
+\begin{bmatrix}
+\coolover{\!\!\!\!\!\!\!{\scriptstyle j=1,\ldots,L}}{\frac{\partial v_{pq}}{\partial \widetilde{\eta}_j}} & \coolover{\!\!\!\!{\scriptstyle k=1,\ldots,L}}{~\frac{\partial v_{pq}}{\partial \widetilde{\varphi}_k}}
+\end{bmatrix}
+\begin{matrix}% matrix for right braces 
+\coolrightbrace{\frac{\partial v_{ij}}{\partial \eta_j}}{\scriptstyle [pq] = 1,\ldots,B~(p<q)}
+\end{matrix}\\
+&=& \begin{bmatrix}
+\coolunder{\!\!\!\!\!\!\!{\scriptstyle j=1,\ldots,N}}{m_{j,pq}} & \coolunder{\!\!\!\!{\scriptstyle k=1,\ldots,N}}{i m_{k,pq}}
+\end{bmatrix}
+\begin{matrix}% matrix for right braces 
+\coolrightbrace{\frac{\partial v_{ij}}{\partial \eta_j}}{\scriptstyle [pq] = 1,\ldots,B~(p<q)}
+\end{matrix}\\
+\end{eqnarray}
+
 Moreover,
@@
 \end{equation}
-
-fiducial initial guess
-
-The following partial derivatives are now easily computed
-
-\begin{eqnarray}
-\frac{c_{ij}}{\partial \eta_i} &=& y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\ 
-\frac{c_{ij}}{\partial \varphi_i} &=&  -i y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\
-\frac{c_{ij}}{\partial \eta_j} &=& y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\ 
-\frac{c_{ij}}{\partial \varphi_j} &=&  i y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\
-\frac{y_{i-j}}{\partial \varphi_j} &=&  e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}
-\end{eqnarray}
-
-The Wirtinger derivative is used in the last equation.
-
-\begin{eqnarray}
-c_{ij} &\approx& c_{ij}^0 + \Delta \eta_i(y_{i-j}^0 e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}) + \Delta \eta_j(y_{i-j}^0 e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0})\\
-&& -i\Delta\varphi_i (y_{i-j}^0e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}) +i\Delta\varphi_j (y_{i-j}^0e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0})\\
-&& y_{i-j}^1(y_{i-j}^0e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0})\\
-&\approx&  c_{ij}^0 + e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_i+ \Delta \eta_j - i\Delta\varphi_i + i\Delta\varphi_i))
-\end{eqnarray}
-
+and
 \begin{equation}
-\label{eq:residual}
-r_{ij} = \delta_{ij} = c_{ij}-c_{ij}^0 = e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_i+ \Delta \eta_j - i\Delta\varphi_i + i\Delta\varphi_i)) 
+m_{j,pq} = \delta_{\phi_{pq}}^{j} v_{pq}.
 \end{equation}
 
-Eq.~\ref{eq:residual} allows us to construct the following linear system:
-
-\begin{equation}
-\boldsymbol{J}[\boldsymbol{\Delta \eta},\boldsymbol{\Delta \varphi},\boldsymbol{\Delta y}]^T = \boldsymbol{r}, 
-\end{equation}
-
-where $\boldsymbol{J}$ is equal to 
-\begin{equation}
-\boldsymbol{J} = \bigg [\overbrace{\frac{c_{pq}}{\partial \eta_i}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial \varphi_i}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial y_{t}}}^{t=1\cdots r_s} \bigg ] \bigg \} [pq] = 1\cdots N_{\textrm{b}} (p<q)
-\end{equation}
-or
-\begin{equation}
-\boldsymbol{J} = \bigg [\frac{c_{pq}}{\partial \boldsymbol{\eta}},~\frac{c_{pq}}{\partial \boldsymbol{\varphi}},~\frac{c_{pq}}{\partial \boldsymbol{y}} \bigg ] \bigg \} [pq] = 1\cdots N_{\textrm{b}} (p<q).
-\end{equation}
-
-The last column is again a Wirtinger derivative.
-
-Which means we can obtain the least-squares estimate as follows:
-
-\begin{equation}
-[\boldsymbol{\Delta \eta},\boldsymbol{\Delta \varphi},\boldsymbol{\Delta y}]^T = [\boldsymbol{J}^H\boldsymbol{J}]^{-1}\boldsymbol{J}^H\boldsymbol{r}.
-\end{equation}
+% fiducial initial guess 
+% 
+% The following partial derivatives are now easily computed
+% 
+% \begin{eqnarray}
+% \frac{c_{ij}}{\partial \eta_i} &=& y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\ 
+% \frac{c_{ij}}{\partial \varphi_i} &=&  -i y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\
+% \frac{c_{ij}}{\partial \eta_j} &=& y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\ 
+% \frac{c_{ij}}{\partial \varphi_j} &=&  i y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\
+% \frac{y_{i-j}}{\partial \varphi_j} &=&  e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}
+% \end{eqnarray}
+% 
+% 
+% The Wirtinger derivative is used in the last equation.
+% 
+% \begin{eqnarray}
+% c_{ij} &\approx& c_{ij}^0 + \Delta \eta_i(y_{i-j}^0 e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}) + \Delta \eta_j(y_{i-j}^0 e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0})\\
+% && -i\Delta\varphi_i (y_{i-j}^0e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}) +i\Delta\varphi_j (y_{i-j}^0e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0})\\
+% && y_{i-j}^1(y_{i-j}^0e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0})\\
+% &\approx&  c_{ij}^0 + e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_i+ \Delta \eta_j - i\Delta\varphi_i + i\Delta\varphi_i))
+% \end{eqnarray}
+% 
+% \begin{equation}
+% \label{eq:residual}
+% r_{ij} = \delta_{ij} = c_{ij}-c_{ij}^0 = e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_i+ \Delta \eta_j - i\Delta\varphi_i + i\Delta\varphi_i)) 
+% \end{equation}
+% 
+% Eq.~\ref{eq:residual} allows us to construct the following linear system:
+% 
+% \begin{equation}
+% \boldsymbol{J}[\boldsymbol{\Delta \eta},\boldsymbol{\Delta \varphi},\boldsymbol{\Delta y}]^T = \boldsymbol{r}, 
+% \end{equation}
+% 
+% where $\boldsymbol{J}$ is equal to 
+% \begin{equation}
+% \boldsymbol{J} = \bigg [\overbrace{\frac{c_{pq}}{\partial \eta_i}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial \varphi_i}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial y_{t}}}^{t=1\cdots r_s} \bigg ] \bigg \} [pq] = 1\cdots N_{\textrm{b}} (p<q)
+% \end{equation}
+% or
+% \begin{equation}
+% \boldsymbol{J} = \bigg [\frac{c_{pq}}{\partial \boldsymbol{\eta}},~\frac{c_{pq}}{\partial \boldsymbol{\varphi}},~\frac{c_{pq}}{\partial \boldsymbol{y}} \bigg ] \bigg \} [pq] = 1\cdots N_{\textrm{b}} (p<q).
+% \end{equation}
+% 
+% The last column is again a Wirtinger derivative.
+% 
+% Which means we can obtain the least-squares estimate as follows:
+% 
+% \begin{equation}
+% [\boldsymbol{\Delta \eta},\boldsymbol{\Delta \varphi},\boldsymbol{\Delta y}]^T = [\boldsymbol{J}^H\boldsymbol{J}]^{-1}\boldsymbol{J}^H\boldsymbol{r}.
+% \end{equation}
 

### 2016-10-03 paper.tex
added some intro tex in the derivation of redundant stefcal

@@
 
+\section{Least Squares}
+Scalar redundant calibration can be achieved by solving the following optimization problem
+\begin{equation}
+\min_{\bg,\by} \|\bd - \bv(\bg,\by)\|. 
+\end{equation}
+The most standard way of solving a least squares method is to use a gradient type of minimization algorithm to iteratively perturb the parameters so that the ... minimize the 
+difference between observed and predicted visibilities. The gradient based minimization algorithms generally requires the model $\bv$ to be differentiable
+towards each unknown. When the problem is complex, which is the case for Eq..., it becomes difficult to directly apply gradient methods to such problems. The reason being,
+many complex functions are not analytic nor differentiable (no Taylor expansion exists) if we use classical definitions. For instance if the 
+classical definition of differentiaton is used then $\frac{\partial z}{\partial \conj{z}}$, where $z \in \mathbb{C}$, does not exist.
+
+The standard way of avoiding such complications is to divide the problem into its real and imaginary parts and then to solve for the real and imaginary parts of the parameters separately.
+Recently cite the complex paper has proposed to extend the problem into [RW HERE NEED TO STATE BETTER] its conjugate domain and then to solve for the parameters and there complex conjugates.
+This can be accomplished
+if we introduce Wirtinger derivatives. (I have to look into how to write this better).
+
+The solution to Eq. can now be obtained by using a gradient type of minimization algorithm, like GN and LM.
+
+The GN update rule is given by\\
+
+GN UPDATE RULE
+
 % fiducial initial guess 

### 2016-10-04 paper.tex
developed new mapping to make mathtools redundant

@@
 \left.\vphantom{\begin{matrix} #1 \end{matrix}}\right\}#2}
+
+%MATRIX
+% Load TikZ
+\usepackage{tikz}
+\usetikzlibrary{matrix,decorations.pathreplacing,calc}
+
+% Set various styles for the matrices and braces. It might pay off to fiddle around with the values a little bit
+\pgfkeys{tikz/mymatrixenv/.style={decoration=brace,every left delimiter/.style={xshift=3pt},every right delimiter/.style={xshift=-3pt}}}
+\pgfkeys{tikz/mymatrix/.style={matrix of math nodes,left delimiter=[,right delimiter={]},inner sep=2pt,column sep=1em,row sep=0.5em,nodes={inner sep=0pt}}}
+\pgfkeys{tikz/mymatrixbrace/.style={decorate,thick}}
+\newcommand\mymatrixbraceoffseth{0.5em}
+\newcommand\mymatrixbraceoffsetv{0.2em}
+
+% Now the commands to produce the braces. (I'll explain below how to use them.)
+\newcommand*\mymatrixbraceright[4][m]{
+    \draw[mymatrixbrace] ($(#1.north west)!(#1-#3-1.south west)!(#1.south west)-(\mymatrixbraceoffseth,0)$)
+        -- node[left=2pt] {#4} 
+        ($(#1.north west)!(#1-#2-1.north west)!(#1.south west)-(\mymatrixbraceoffseth,0)$);
+}
+\newcommand*\mymatrixbraceleft[4][m]{
+    \draw[mymatrixbrace] ($(#1.north east)!(#1-#2-1.north east)!(#1.south east)+(\mymatrixbraceoffseth,0)$)
+        -- node[right=2pt] {#4} 
+        ($(#1.north east)!(#1-#3-1.south east)!(#1.south east)+(\mymatrixbraceoffseth,0)$);
+}
+\newcommand*\mymatrixbracetop[4][m]{
+    \draw[mymatrixbrace] ($(#1.north west)!(#1-1-#2.north west)!(#1.north east)+(0,\mymatrixbraceoffsetv)$)
+        -- node[above=2pt] {#4} 
+        ($(#1.north west)!(#1-1-#3.north east)!(#1.north east)+(0,\mymatrixbraceoffsetv)$);
+}
+\newcommand*\mymatrixbracebottom[4][m]{
+    \draw[mymatrixbrace] ($(#1.south west)!(#1-1-#3.south east)!(#1.south east)-(0,\mymatrixbraceoffsetv)$)
+        -- node[below=2pt] {#4} 
+        ($(#1.south west)!(#1-1-#2.south west)!(#1.south east)-(0,\mymatrixbraceoffsetv)$);
+}
+
+%%END MATRIX
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 where 
-\begin{eqnarray}
- \bd &=& [d_{pq}] \big \} ~ [pq] = 1\cdots B ~ (p<q),\nonumber\\
- \bv &=& [g_p y_{\phi_{pq}} \conj{g}_q]  \big \} ~ [pq] = 1\cdots B ~ (p<q),\nonumber\\
- \bn &=& [n_{pq}]  \big \} ~ [pq] = 1\cdots B ~ (p<q).\label{eq:vec_definitions}
-\end{eqnarray}
-In Eq.~\eqref{eq:vec_definitions}, $B$ denotes the number of baselines in the array.
+\begin{align}
+ \left [ \bd \right]_{\alpha_{pq}} &= d_{pq}, & \left [ \bv \right ]_{\alpha_{pq}} &= g_p y_{\phi_{pq}} \conj{g}_q\nonumber\\
+ \left [ \bn \right ]_{\alpha_{pq}} &= n_{pq}. &  &\label{eq:vec_definitions}
+\end{align}
+In Eq.~\eqref{eq:vec_definitions}, the size of the three vectors are equal to $B$, which denotes the number of baselines in the array. Moreover,
+\begin{equation}
+\alpha_{pq} =
+\begin{cases}
+(q-p) + (p-1)\left (N-\frac{1}{2}p \right ) & \textrm{if}~p<q\\
+0 & \textrm{otherwise}
+\end{cases}.
+\end{equation}
+The easiest way to construct the vectors in Eq.~\ref{eq:vec_definitions} is to enumerate through the composite indexes $pq$ in the following order $12, 13,\cdots,1N,23,\cdots$.
+For each composite index we then compute the right hand side of the appropriate equation in Eq.~\ref{eq:vec_definitions} which we then need to assign to the 
+correct position in the vector.
+The correct position is calculated with $\alpha_{pq}$.
+Using the aforementioned order, however, allows you to fill the array in numerical order, i.e.
+$12,13,\cdots,1N,23,\cdots \xrightarrow{\alpha_{pq}} 1,2,3,\cdots,B$.
 
@@
 
+\begin{equation}
+\begin{tikzpicture}[mymatrixenv]
+    \matrix[mymatrix] (m)  {
+        a & b & c & d \\
+        e & f & g & h \\
+        i & j & k & l \\
+        m & n & o & p \\
+    };
+    \mymatrixbraceright{1}{3}{$A$}
+    \mymatrixbracetop{2}{4}{$R$}
+\end{tikzpicture}
+\end{equation}
+
+\begin{equation}
+\begin{tikzpicture}[mymatrixenv]
+    \matrix[mymatrix] (m)  {
+        \frac{\partial v_{pq}}{\partial \eta_j} & \frac{\partial v_{pq}}{\partial \varphi_k}\\
+    };
+    \mymatrixbraceleft{1}{1}{A}
+    \mymatrixbracetop{2}{2}{$j=1,\ldots,N$}
+\end{tikzpicture}
+\end{equation}
+
+\begin{eqnarray}
+\bN &=& 
+\begin{bmatrix}
+\coolover{\!\!\!\!\!\!\!{\scriptstyle j=1,\ldots,N}}{\frac{\partial v_{pq}}{\partial \eta_j}} & \coolover{\!\!\!\!{\scriptstyle k=1,\ldots,N}}{~\frac{\partial v_{pq}}{\partial \varphi_k}}
+\end{bmatrix}
+\begin{matrix}% matrix for right braces 
+\coolrightbrace{\frac{\partial v_{ij}}{\partial \eta_j}}{\scriptstyle [pq] = 1,\ldots,B~(p<q)}
+\end{matrix}\\
+&=& \begin{bmatrix}
+\coolunder{\!\!\!\!\!\!\!{\scriptstyle j=1,\ldots,N}}{n_{j,pq}} & \coolunder{\!\!\!\!{\scriptstyle k=1,\ldots,N}}{(-1)^{\delta_q^k}i n_{k,pq}}
+\end{bmatrix}
+\begin{matrix}% matrix for right braces 
+\coolrightbrace{\frac{\partial v_{ij}}{\partial \eta_j}}{\scriptstyle [pq] = 1,\ldots,B~(p<q)}
+\end{matrix}\\
+\end{eqnarray}
+
 % fiducial initial guess 

### 2016-10-04 paper.tex
implemented workaround for mathtools

--- 2016-10-03 paper.tex
@@
 \usepackage{graphicx}
-\usepackage{mathtools}
 
@@
 
-\newcommand\coolover[2]{\mathrlap{\smash{\overbrace{\phantom{%
-    \begin{matrix} #2 \end{matrix}}}^{\mbox{$#1$}}}}#2}
-
-\newcommand\coolunder[2]{\mathrlap{\smash{\underbrace{\phantom{%
-    \begin{matrix} #2 \end{matrix}}}_{\mbox{$#1$}}}}#2}
-
-\newcommand\coolleftbrace[2]{%
-#1\left\{\vphantom{\begin{matrix} #2 \end{matrix}}\right.}
-
-\newcommand\coolrightbrace[2]{%
-\left.\vphantom{\begin{matrix} #1 \end{matrix}}\right\}#2}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 where 
-\begin{eqnarray}
- \bd &=& [d_{pq}] \big \} ~ [pq] = 1\cdots B ~ (p<q),\nonumber\\
- \bv &=& [g_p y_{\phi_{pq}} \conj{g}_q]  \big \} ~ [pq] = 1\cdots B ~ (p<q),\nonumber\\
- \bn &=& [n_{pq}]  \big \} ~ [pq] = 1\cdots B ~ (p<q).\label{eq:vec_definitions}
-\end{eqnarray}
-In Eq.~\eqref{eq:vec_definitions}, $B$ denotes the number of baselines in the array.
+\begin{align}
+ \left [ \bd \right]_{\alpha_{pq}} &= d_{pq}, & \left [ \bv \right ]_{\alpha_{pq}} &= g_p y_{\phi_{pq}} \conj{g}_q\nonumber\\
+ \left [ \bn \right ]_{\alpha_{pq}} &= n_{pq}. &  &\label{eq:vec_definitions}
+\end{align}
+In Eq.~\eqref{eq:vec_definitions}, the size of the three vectors are equal to $B$ (i.e. $p<q$), which denotes the number of baselines in the array. Moreover,
+\begin{equation}
+\alpha_{pq} =
+\begin{cases}
+(q-p) + (p-1)\left (N-\frac{1}{2}p \right ) & \textrm{if}~p<q\\
+B + (p-q) + (q-1)\left (N-\frac{1}{2}q \right )) & \textrm{if}~p>q\\
+%B+q + \frac{1}{2}(p-1)(p-2) & \textrm{if}~p>q\\
+0 & \textrm{otherwise}
+\end{cases}.
+\end{equation}
+The easiest way to construct the vectors in Eq.~\ref{eq:vec_definitions} is to enumerate through the composite indexes $pq$ in the following order $12, 13,\cdots,1N,23,\cdots$.
+For each composite index we then compute the right hand side of the appropriate vector element definition which we then need to assign to the 
+correct vector position.
+The correct position is calculated with $\alpha_{pq}$.
+Using the aforementioned order, however, allows you to fill the vector in numerical order, i.e.
+$12,13,\cdots,1N,23,\cdots \xrightarrow{\alpha_{pq}} 1,2,3,\cdots,B$.
 
@@
 \subsection{LOGCAL}
+NEED BETTER INTRO\\
 One of the best known algorithms that is currently used to perform redundant calibration involves the use of logarithms. If we ignore the noise term then we can 
@@
 \begin{eqnarray}
-\ln |d_{ij}| &=& \ln |g_i| + \ln |g_j| + \ln |y_{\phi_{ij}}| \label{eq:logcal_amp}\\
-\angle d_{ij} &=& \angle g_i - \angle g_j + \angle \phi_{ij} \label{eq:logcal_phase}
+\ln |d_{pq}| &=& \ln |g_p| + \ln |g_q| + \ln |y_{\phi_{pq}}| \label{eq:logcal_amp}\\
+\angle d_{pq} &=& \angle g_p - \angle g_q + \angle \phi_{pq} \label{eq:logcal_phase}
 \end{eqnarray}
@@
 \begin{equation}
-\widetilde{\bd} = \bA\bz, 
-\end{equation}
-which has the following least-squares solution
-\begin{equation}
-\bz = (\bA^T\bA)^{-1}\bA^T\widetilde{\bd},
+\widetilde{\bd} = \bA\widetilde{\bz}, 
 \end{equation}
@@
 \begin{equation}
-\widetilde{\bd} = [\ln |d_{pq}|] \big \} ~ [pq] = 1,\cdots,B ~ (p<q),\nonumber\\ 
+\left [ \widetilde{\bd}\right ]_{\alpha_{pq}} = \ln |d_{pq}|,\nonumber\\ 
+\end{equation}
+\begin{equation}
+\widetilde{\bz} = \begin{bmatrix}
+                  \ln | \bg |\\
+                  \ln | \by |
+                  \end{bmatrix},
+\end{equation}
+with $\bg = [g_1,g_2,\cdots,g_N]^T$ and $\by = [y_1,y_2,\cdots,y_L]^T$.
+
+Furthermore,
+\begin{equation}
+\label{eq:A}
+\bA = \begin{bmatrix}
+       \bA_1 & \bA_2
+      \end{bmatrix}
 \end{equation}
@@
 \begin{equation}
-\bA = 
-\begin{bmatrix}
-\coolover{{\scriptstyle j=1,\ldots,N}}{\delta^j_p\vee\delta^j_q} & \coolover{{\scriptstyle k=1,\ldots,L}\!\!\!\!}{~\delta_{\phi_{pq}}^k}
-\end{bmatrix}
-\begin{matrix}% matrix for right braces 
-\coolrightbrace{\delta_{\phi_{pq}}^k}{[pq] = 1,\ldots,B~(p<q)}
-\end{matrix}.
+\left [ \bA_1 \right ]_{\alpha_{pq},j} = \delta^j_p\vee\delta^j_q,  ~~~ \left [ \bA_2 \right ]_{\alpha_{pq},j} = \delta_{\phi_{pq}}^j. 
 \end{equation}
 
-Moreover, $(*)^T$ denotes the transpose of its operand, $(*)^{-1}$ denotes matrix inversion, $\vee$ denotes the logical or operator and
+The dimensions of $\bA_1$ and $\bA_2$ are $B\times N$ and $B\times L$.
+Moreover, $\vee$ denotes the logical or operator and
 \begin{equation}
@@
 \end{equation}
-A similar linear system can be constructed for Eq.~\eqref{eq:logcal_phase}.
+A similar linear system can be constructed for Eq.~\eqref{eq:logcal_phase}.\\
+
+NEED TO ROUND OF HERE STILL
 
 \subsection{LINCAL}
-Assume that $\eta_i,\eta_j,\widetilde{\eta}_{\phi_{ij}},\varphi_i,\varphi_j$ and $\widetilde{\varphi}_{\phi_{ij}}$ are real valued variables.
-Substituting $g_i$ with $e^{\eta_i - i \varphi_i}$, $g_j$ with $e^{\eta_j - i \varphi_j}$ and $y_{\phi_{ij}}$ with $e^{\eta_{\widetilde{\phi}_{ij}}- i \widetilde{\varphi}_{\phi_{ij}}}$ in Eq.~\eqref{eq:vis_red} results in
+Assume that $\eta_p,\eta_q,\widetilde{\eta}_{\phi_{pq}},\varphi_p,\varphi_q$ and $\widetilde{\varphi}_{\phi_{pq}}$ are real valued variables.
+Substituting $g_p$ with $e^{\eta_p - i \varphi_p}$, $g_q$ with $e^{\eta_q - i \varphi_q}$ and $y_{\phi_{pq}}$ with $e^{\eta_{\widetilde{\phi}_{pq}}- i \widetilde{\varphi}_{\phi_{pq}}}$ in Eq.~\eqref{eq:vis_red} results in
 \begin{equation}
 \label{eq:lincal}
-d_{ij} =  e^{\eta_i - i \varphi_i} e^{\widetilde{\eta}_{\phi_{ij}}- i \widetilde{\varphi}_{\phi_{ij}}} e^{\eta_j + i \varphi_j} + n_{ij}.
+d_{pq} =  e^{\eta_p - i \varphi_p} e^{\widetilde{\eta}_{\phi_{pq}}- i \widetilde{\varphi}_{\phi_{pq}}} e^{\eta_q + i \varphi_q} + n_{pq}.
 \end{equation}
@@
 A possible approach one can use to solve the unknowns in Eq.~\eqref{eq:lincal} is to write down the Taylor expansion of the first term of the right hand side of Eq.~\eqref{eq:lincal} around some 
-initial fiducial guess of $\eta_i,\eta_j,\widetilde{\eta}_{\phi_{ij}},\varphi_i,\varphi_j$ and $\widetilde{\varphi}_{\phi_{ij}}$, which we denote with 
-$\eta_i^0,\eta_j^0,\widetilde{\eta}_{\phi_{ij}}^0,\varphi_i^0,\varphi_j^0$ and $\widetilde{\varphi}_{\phi_{ij}}^0$, and then to solve the perturbations from this 
+initial fiducial guess of $\eta_p,\eta_q,\widetilde{\eta}_{\phi_{pq}},\varphi_p,\varphi_q$ and $\widetilde{\varphi}_{\phi_{pq}}$, which we denote with 
+$\eta_p^0,\eta_q^0,\widetilde{\eta}_{\phi_{pq}}^0,\varphi_p^0,\varphi_q^0$ and $\widetilde{\varphi}_{\phi_{pq}}^0$, and then to solve the perturbations from this 
 initial guess that minimize the difference between the right and left hand side of Eq.~\eqref{eq:lincal}. Once the perturbation values 
@@
 \label{eq:taylor_expansion}
-d_{ij} = v_{ij}^0[1+\Delta \eta_i + \Delta \eta_j + i \varphi_i - i \Delta \varphi_j + \Delta \widetilde{\eta}_{\phi_{ij}} + i\Delta \widetilde{\eta}_{\phi_{ij}}] + n_{ij}, 
+d_{pq} = v_{pq}^0[1+\Delta \eta_p + \Delta \eta_q + i \varphi_p - i \Delta \varphi_q + \Delta \widetilde{\eta}_{\phi_{pq}} + i\Delta \widetilde{\eta}_{\phi_{pq}}] + n_{pq}, 
 \end{equation}
@@
 \begin{equation}
-v_{ij}^0 = e^{\eta_i^0 - i \varphi_i^0} e^{\widetilde{\eta}_{\phi_{ij}}^0 - i \widetilde{\varphi}_{\phi_{ij}}^0} e^{\eta_j^0 + i \varphi_j^0}, 
+v_{pq}^0 = e^{\eta_p^0 - i \varphi_p^0} e^{\widetilde{\eta}_{\phi_{pq}}^0 - i \widetilde{\varphi}_{\phi_{pq}}^0} e^{\eta_q^0 + i \varphi_q^0}, 
 \end{equation}
-and $\eta_i^0,\eta_j^0,\widetilde{\eta}_{\phi_{ij}}^0,\varphi_i^0,\varphi_j^0$ and $\widetilde{\varphi}_{\phi_{ij}}^0$ respectively denote the initial guess for $\eta_i,\eta_j,\widetilde{\eta}_{\phi_{ij}},\varphi_i,\varphi_j$ and $\widetilde{\varphi}_{\phi_{ij}}$.
+and $\eta_p^0,\eta_q^0,\widetilde{\eta}_{\phi_{pq}}^0,\varphi_p^0,\varphi_q^0$ and $\widetilde{\varphi}_{\phi_{pq}}^0$ respectively denote the initial guess for $\eta_p,\eta_q,\widetilde{\eta}_{\phi_{pq}},\varphi_p,\varphi_q$ and $\widetilde{\varphi}_{\phi_{pq}}$.
 Eq.~\eqref{eq:taylor_expansion} can be used to construct the following linear system
 \begin{equation}
-\breve{\br} = \bJ\Delta\widetilde{\bz}, 
+\breve{\br} = \bJ\Delta\widehat{\bz}, 
 \end{equation}
@@
 \begin{equation}
-\breve{\br} = [d_{ij}-v_{ij}^0] \big \} ~ [ij] = 1\cdots 2B, 
+\left [ \breve{\br} \right ]_{\alpha_{pq}} = d_{pq}-v_{pq}^0, 
 \end{equation}
@@
 \begin{equation}
-\Delta\widetilde{\bz} = 
+\Delta\widehat{\bz} = 
 \begin{bmatrix}
-\Delta \eta_i \\
-\Delta \varphi_j \\
-\Delta \widetilde{\eta}_k \\
-\Delta \widetilde{\varphi}_l \\
+\Delta \bmath{\eta} \\
+\Delta \bmath{\varphi} \\
+\Delta \widetilde{\bmath{\eta}} \\
+\Delta \widetilde{\bmath{\varphi}},
 \end{bmatrix}
-\begin{matrix}% matrix for right braces 
-\coolrightbrace{\eta_i}{i = 1\cdots N}\\
-\coolrightbrace{\eta_i}{j = 1\cdots N}\\
-\coolrightbrace{\eta_i}{k = 1\cdots L}\\
-\coolrightbrace{\eta_i}{l = 1 \cdots L}
-\end{matrix}.
 \end{equation}
+with
+\begin{align}
+ \Delta \bmath{\eta} &= [\Delta \eta_1,\Delta \eta_2,\cdots \Delta \eta_N]^T\nonumber\\
+ \Delta \bmath{\varphi} &= [\Delta \varphi_1,\Delta \varphi_2,\cdots \Delta \varphi_N]^T\nonumber\\
+ \Delta \widetilde{\bmath{\eta}} &= [\Delta \widetilde{\eta_1},\Delta \widetilde{\eta_2},\cdots \Delta \widetilde{\eta_N}]^T\nonumber\\
+ \Delta \widetilde{\bmath{\varphi}} &= [\Delta \widetilde{\varphi_1},\Delta \widetilde{\varphi_2},\cdots \Delta \widetilde{\varphi_N}]^T.
+\end{align}
 
-In Eq.~\eqref{eq:jac_lincal}
-
-\begin{eqnarray}
-\bN &=& 
-\begin{bmatrix}
-\coolover{\!\!\!\!\!\!\!{\scriptstyle j=1,\ldots,N}}{\frac{\partial v_{pq}}{\partial \eta_j}} & \coolover{\!\!\!\!{\scriptstyle k=1,\ldots,N}}{~\frac{\partial v_{pq}}{\partial \varphi_k}}
-\end{bmatrix}
-\begin{matrix}% matrix for right braces 
-\coolrightbrace{\frac{\partial v_{ij}}{\partial \eta_j}}{\scriptstyle [pq] = 1,\ldots,B~(p<q)}
-\end{matrix}\\
-&=& \begin{bmatrix}
-\coolunder{\!\!\!\!\!\!\!{\scriptstyle j=1,\ldots,N}}{n_{j,pq}} & \coolunder{\!\!\!\!{\scriptstyle k=1,\ldots,N}}{(-1)^{\delta_q^k}i n_{k,pq}}
-\end{bmatrix}
-\begin{matrix}% matrix for right braces 
-\coolrightbrace{\frac{\partial v_{ij}}{\partial \eta_j}}{\scriptstyle [pq] = 1,\ldots,B~(p<q)}
-\end{matrix}\\
-\end{eqnarray}
-
-\begin{eqnarray}
-\bM &=& 
-\begin{bmatrix}
-\coolover{\!\!\!\!\!\!\!{\scriptstyle j=1,\ldots,L}}{\frac{\partial v_{pq}}{\partial \widetilde{\eta}_j}} & \coolover{\!\!\!\!{\scriptstyle k=1,\ldots,L}}{~\frac{\partial v_{pq}}{\partial \widetilde{\varphi}_k}}
-\end{bmatrix}
-\begin{matrix}% matrix for right braces 
-\coolrightbrace{\frac{\partial v_{ij}}{\partial \eta_j}}{\scriptstyle [pq] = 1,\ldots,B~(p<q)}
-\end{matrix}\\
-&=& \begin{bmatrix}
-\coolunder{\!\!\!\!\!\!\!{\scriptstyle j=1,\ldots,N}}{m_{j,pq}} & \coolunder{\!\!\!\!{\scriptstyle k=1,\ldots,N}}{i m_{k,pq}}
-\end{bmatrix}
-\begin{matrix}% matrix for right braces 
-\coolrightbrace{\frac{\partial v_{ij}}{\partial \eta_j}}{\scriptstyle [pq] = 1,\ldots,B~(p<q)}
-\end{matrix}\\
-\end{eqnarray}
+Furthermore, 
+\begin{equation}
+\bN = \begin{bmatrix}
+      \bN_1 & \bN_2 
+      \end{bmatrix},
+\end{equation}
+with 
+\begin{align}
+[\bN_1]_{\alpha_{pq},j} &= \frac{\partial v_{pq}}{\partial \eta_j} {\bigg |}_{\widehat{\bz}=\widehat{\bz}^0}\\
+&= (\delta_{p}^{j}\vee\delta_{q}^{j}) v_{pq}^0,
+\end{align}
+and
+\begin{align}
+ [\bN_2]_{\alpha_{pq},j} &= \frac{\partial v_{pq}}{\partial \varphi_j} {\bigg |}_{\widehat{\bz}=\widehat{\bz}^0}\\
+ &= (-1)^{\delta_{q}^{j}}i(\delta_{p}^{j}\vee\delta_{q}^{j}) v_{pq}^0. 
+\end{align}
 
@@
 \begin{equation}
-n_{j,pq} = (\delta_{p}^{j}\vee\delta_{q}^{j}) v_{pq} 
+\bM = \begin{bmatrix}
+      \bM_1 & \bM_2 
+      \end{bmatrix},
 \end{equation}
+with
+\begin{align}
+[\bM_1]_{\alpha_{pq},j} &= \frac{\partial v_{pq}}{\partial \widetilde{\eta}_j} {\bigg |}_{\widehat{\bz}=\widehat{\bz}^0}\\
+&=  \delta_{\phi_{pq}}^{j} v_{pq}^0,
+\end{align}
 and
-\begin{equation}
-m_{j,pq} = \delta_{\phi_{pq}}^{j} v_{pq}.
-\end{equation}
+\begin{align}
+[\bM_2]_{\alpha_{pq},j} &= \frac{\partial v_{pq}}{\partial \widetilde{\varphi}_j} {\bigg |}_{\widehat{\bz}=\widehat{\bz}^0}\\
+&= i\delta_{\phi_{pq}}^{j} v_{pq}^0,
+\end{align}
+
+STILL NEEDS LOTS OF WORK.
 
@@
 % \begin{eqnarray}
-% \frac{c_{ij}}{\partial \eta_i} &=& y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\ 
-% \frac{c_{ij}}{\partial \varphi_i} &=&  -i y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\
-% \frac{c_{ij}}{\partial \eta_j} &=& y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\ 
-% \frac{c_{ij}}{\partial \varphi_j} &=&  i y_{i-j} e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}\\
-% \frac{y_{i-j}}{\partial \varphi_j} &=&  e^{\eta_i - i \varphi_i} e^{\eta_j - i \varphi_j}
+% \frac{c_{ij}}{\partial \eta_p} &=& y_{i-j} e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}\\ 
+% \frac{c_{ij}}{\partial \varphi_p} &=&  -i y_{i-j} e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}\\
+% \frac{c_{ij}}{\partial \eta_q} &=& y_{i-j} e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}\\ 
+% \frac{c_{ij}}{\partial \varphi_q} &=&  i y_{i-j} e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}\\
+% \frac{y_{i-j}}{\partial \varphi_q} &=&  e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}
 % \end{eqnarray}
@@
 % \begin{eqnarray}
-% c_{ij} &\approx& c_{ij}^0 + \Delta \eta_i(y_{i-j}^0 e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}) + \Delta \eta_j(y_{i-j}^0 e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0})\\
-% && -i\Delta\varphi_i (y_{i-j}^0e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}) +i\Delta\varphi_j (y_{i-j}^0e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0})\\
-% && y_{i-j}^1(y_{i-j}^0e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0})\\
-% &\approx&  c_{ij}^0 + e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_i+ \Delta \eta_j - i\Delta\varphi_i + i\Delta\varphi_i))
+% c_{ij} &\approx& c_{ij}^0 + \Delta \eta_p(y_{i-j}^0 e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0}) + \Delta \eta_q(y_{i-j}^0 e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0})\\
+% && -i\Delta\varphi_p (y_{i-j}^0e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0}) +i\Delta\varphi_q (y_{i-j}^0e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0})\\
+% && y_{i-j}^1(y_{i-j}^0e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0})\\
+% &\approx&  c_{ij}^0 + e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_p+ \Delta \eta_q - i\Delta\varphi_p + i\Delta\varphi_p))
 % \end{eqnarray}
@@
 % \label{eq:residual}
-% r_{ij} = \delta_{ij} = c_{ij}-c_{ij}^0 = e^{\eta_i^0 - i \varphi_i^0} e^{\eta_j^0 - i \varphi_j^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_i+ \Delta \eta_j - i\Delta\varphi_i + i\Delta\varphi_i)) 
+% r_{ij} = \delta_{ij} = c_{ij}-c_{ij}^0 = e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_p+ \Delta \eta_q - i\Delta\varphi_p + i\Delta\varphi_p)) 
 % \end{equation}
@@
 % \begin{equation}
-% \boldsymbol{J} = \bigg [\overbrace{\frac{c_{pq}}{\partial \eta_i}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial \varphi_i}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial y_{t}}}^{t=1\cdots r_s} \bigg ] \bigg \} [pq] = 1\cdots N_{\textrm{b}} (p<q)
+% \boldsymbol{J} = \bigg [\overbrace{\frac{c_{pq}}{\partial \eta_p}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial \varphi_p}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial y_{t}}}^{t=1\cdots r_s} \bigg ] \bigg \} [pq] = 1\cdots N_{\textrm{b}} (p<q)
 % \end{equation}

### 2016-10-05 paper.tex
added sme text to complex optimization

@@
 \begin{equation}
-\min_{\bg,\by} \|\bd - \bv(\bg,\by)\|. 
+\label{eq:red_least_squares}
+\min_{\bz} \|\bd - \bv(\bz)\|, 
 \end{equation}
-The most standard way of solving a least squares method is to use a gradient type of minimization algorithm to iteratively perturb the parameters so that the ... minimize the 
+where $\bz = [\bg,\by]^T$.
+The most standard way of solving a least squares problem is to use a gradient type of minimization algorithm to iteratively perturb the parameters so that the ... minimize the 
 difference between observed and predicted visibilities. The gradient based minimization algorithms generally requires the model $\bv$ to be differentiable
@@
 
-The standard way of avoiding such complications is to divide the problem into its real and imaginary parts and then to solve for the real and imaginary parts of the parameters separately.
-Recently cite the complex paper has proposed to extend the problem into [RW HERE NEED TO STATE BETTER] its conjugate domain and then to solve for the parameters and there complex conjugates.
+The only way to circumvent the differentiability conundrum is to recast our optimization problem in such a way that $\bv$ becomes analytic in its argument.
+The standard way of achieving this to divide the optimization problem into its real and imaginary parts and then to solve for the real and imaginary parts of the parameters separately.
+
+Recently Kreutz-Delgado and Sorber have proposed an alternative strategy, which involves the use of Wirtinger calculus (Wirtinger 1927). The Wirtinger derivatives 
+are defined to be 
+\begin{align}
+\frac{\partial}{\partial z} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} -  i \frac{\partial}{\partial y} \right ),&\frac{\partial}{\partial \conj{z}} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} +  i \frac{\partial}{\partial y} \right ).
+\end{align}
+
+\cite{Smirnov2015} recently proposed to use this alternative framework instead of the classic real and imaginary approach to perform normal calibration. We will now show
+that this framework can also be applied to redundant calibration. The first step is to write down the conjugated augmented version of Eq.~\eqref{eq:red_least_squares}, which is equal 
+to
+\begin{equation}
+\label{eq:red_least_squares_complex}
+\min_{\breve{\bz}} \|\breve{\bd} - \breve{\bv}(\breve{\bz})\|, 
+\end{equation}
+where $\breve{\bz} = [\bz,\conj{\bz}]^T$, $\breve{\bd} = [\bd,\conj{\bd}]^T$, $\breve{\bv} = [\bv,\conj{\bv}]$. 
+In Eq.~\eqref{eq:red_least_squares_complex} $\breve{\bv}$ is analytic in its argument and its complex conjugate as a whole under the
+assumption that the Wirtinger derivative is used to define the gradient operator.
+
+cite the complex paper has proposed to extend the problem into [RW HERE NEED TO STATE BETTER] its conjugate domain and then to solve for the parameters and there complex conjugates.
 This can be accomplished

### 2016-10-06 paper.tex
started to write the least squares section

@@
 \section{Least Squares}
-Scalar redundant calibration can be achieved by solving the following optimization problem
+Redundant calibration can be achieved by solving the following optimization problem
 \begin{equation}
@@
 \end{equation}
-where $\bz = [\bg,\by]^T$.
-The most standard way of solving a least squares problem is to use a gradient type of minimization algorithm to iteratively perturb the parameters so that the ... minimize the 
-difference between observed and predicted visibilities. The gradient based minimization algorithms generally requires the model $\bv$ to be differentiable
-towards each unknown. When the problem is complex, which is the case for Eq..., it becomes difficult to directly apply gradient methods to such problems. The reason being,
-many complex functions are not analytic nor differentiable (no Taylor expansion exists) if we use classical definitions. For instance if the 
-classical definition of differentiaton is used then $\frac{\partial z}{\partial \conj{z}}$, where $z \in \mathbb{C}$, does not exist.
+where $\bz = [\bg,\by]^T$. Eq.~\eqref{eq:red_least_squares} is a nonlinear least squares problem.
 
-The only way to circumvent the differentiability conundrum is to recast our optimization problem in such a way that $\bv$ becomes analytic in its argument.
-The standard way of achieving this to divide the optimization problem into its real and imaginary parts and then to solve for the real and imaginary parts of the parameters separately.
+The most standard way of solving a general least squares problem is to use a gradient type of minimization algorithm whose main purpose is to find the model parameters that 
+minimizes the difference between the observed and model data.
+In this paper we will be focusing on the following gradient minimization algorithms: Gauss-Newton and Levenberg-Marquardt (cite LM here).
+For the redundant calibration problem in Eq.~\eqref{eq:red_least_squares} $\bd$, which is the corrupted observed visibilities, is the observed data,
+while $\bv$, which is the predicted visibilities, represents the model data.\footnote{In this paper we use the same symbol to denote predicted visibilities and true observed visibilities as it
+is trivial to distinguish which one is implied from the context.}
 
-Recently Kreutz-Delgado and Sorber have proposed an alternative strategy, which involves the use of Wirtinger calculus (Wirtinger 1927). The Wirtinger derivatives 
+The gradient based minimization algorithms that are generally used to solve least squares problems generally require the model $\bv$ to be differentiable
+towards each model parameter.
+When the least squares problem is complex, which is the case for Eq.~\eqref{eq:red_least_squares}, it becomes less straightforward to apply these gradient based minimization methods.
+The reason being,
+many complex functions are not analytic (no Taylor expansion around a fiducial point exists) and therefore not differentiable if we make use of classic differentiation. For instance if the 
+classical definition of differentiation is used then $\frac{\partial z}{\partial \conj{z}}$, where $z \in \mathbb{C}$, does not exist.
+
+The only way to circumvent the differentiability conundrum that plaques complex least square problems is to recast our complex optimization problem in such a way that $\bv$ becomes analytic in its argument.
+The standard way of achieving this is to divide the complex optimization problem into its real and imaginary parts and then to solve for the real and imaginary parts of the parameters separately.
+
+Recently Kreutz-Delgado and Sorber proposed an alternative strategy, which involves the use of Wirtinger calculus (Wirtinger 1927). The Wirtinger derivatives 
 are defined to be 
@@
 
-\cite{Smirnov2015} recently proposed to use this alternative framework instead of the classic real and imaginary approach to perform normal calibration. We will now show
-that this framework can also be applied to redundant calibration. The first step is to write down the conjugated augmented version of Eq.~\eqref{eq:red_least_squares}, which is equal 
-to
+\cite{Smirnov2015} recently proposed to use this Wirtinger framework instead of the classic real and imaginary approach to perform normal calibration. We will now show
+that this framework can also be applied to redundant calibration. Analogous to Smirnov and Tasse, we recast Eq.~\eqref{eq:red_least_squares} to 
 \begin{equation}
@@
 where $\breve{\bz} = [\bz,\conj{\bz}]^T$, $\breve{\bd} = [\bd,\conj{\bd}]^T$, $\breve{\bv} = [\bv,\conj{\bv}]$.
-In Eq.~\eqref{eq:red_least_squares_complex} $\breve{\bv}$ is analytic in its argument and its complex conjugate as a whole under the
-assumption that the Wirtinger derivative is used to define the gradient operator.
+Eq.~\eqref{eq:red_least_squares_complex} is the augmented counterpart of Eq.~\eqref{eq:red_least_squares}.
+If we use the Wirtinger derivative instead of the standard derivative to define the gradient operator then 
+$\breve{\bv}$ becomes analytic in $\bz$ and $\conj{\bz}$ as a whole (cite smirnov). The fact that $\breve{\bv}$ is now analytic in its argument implies that we can use the complex Gauss-Newton
+and Levenberg-Marquardt equivalent algorithms to so solve Eq.~\eqref{eq:red_least_squares_complex}.
 

### 2016-10-06 paper.tex
made section even better

@@
 The reason being,
-many complex functions are not analytic (no Taylor expansion around a fiducial point exists) and therefore not differentiable if we make use of classic differentiation. For instance if the 
+many complex functions are not analytic (no Taylor expansion around a fiducial point exists) and therefore not differentiable if the classic notion of differentiation is used. For instance if the 
 classical definition of differentiation is used then $\frac{\partial z}{\partial \conj{z}}$, where $z \in \mathbb{C}$, does not exist.
@@
 \end{align}
+Using the above definitions we can now easily compute the following:
+\begin{align}
+\frac{\partial z}{\partial z} & = 1, & \frac{\partial \conj{z}}{\partial z}&=0, & \frac{\partial z}{\partial \conj{z}} & = 0, & \frac{\partial \conj{z}}{\partial \conj{z}}&=1.
+\end{align}
 
-\cite{Smirnov2015} recently proposed to use this Wirtinger framework instead of the classic real and imaginary approach to perform normal calibration. We will now show
-that this framework can also be applied to redundant calibration. Analogous to Smirnov and Tasse, we recast Eq.~\eqref{eq:red_least_squares} to 
+\citet{Smirnov2015} recently proposed to use this Wirtinger framework instead of the classic real and imaginary approach to perform normal calibration. We will now show
+that the Wirtinger framework can also be applied to redundant calibration. Analogous to \cite{Smirnov2015}, we recast Eq.~\eqref{eq:red_least_squares} to 
 \begin{equation}
@@
 Eq.~\eqref{eq:red_least_squares_complex} is the augmented counterpart of Eq.~\eqref{eq:red_least_squares}.
-If we use the Wirtinger derivative instead of the standard derivative to define the gradient operator then 
-$\breve{\bv}$ becomes analytic in $\bz$ and $\conj{\bz}$ as a whole (cite smirnov). The fact that $\breve{\bv}$ is now analytic in its argument implies that we can use the complex Gauss-Newton
-and Levenberg-Marquardt equivalent algorithms to so solve Eq.~\eqref{eq:red_least_squares_complex}.
+If we use the Wirtinger derivative instead of the classic notion of a derivative to define the gradient operator then 
+$\bv$ becomes analytic in $\bz$ and $\conj{\bz}$ as a whole \citep{Smirnov2015}.
+The fact that $\breve{\bv}$ is now analytic in its argument has the immediate consequence that we can use the complex variants of the Gauss-Newton
+and the Levenberg-Marquardt algorithm to solve Eq.~\eqref{eq:red_least_squares_complex}.
 

### 2016-10-06 paper.tex
started adding Jacobian and Hessian

@@
 \newcommand{\bM}{\bmath{M}}
+\newcommand{\bI}{\bmath{I}}
+\newcommand{\bzero}{\bmath{0}}
 \newcommand{\conj}[1]{\overline{#1}}
@@
 \label{eq:red_least_squares}
-\min_{\bz} \|\bd - \bv(\bz)\|, 
+\min_{\bz} \|\br\| = \min_{\bz} \|\bd - \bv(\bz)\|, 
 \end{equation}
@@
 \label{eq:red_least_squares_complex}
-\min_{\breve{\bz}} \|\breve{\bd} - \breve{\bv}(\breve{\bz})\|, 
+\min_{\breve{\bz}} \|\breve{\br}\| = \min_{\breve{\bz}} \|\breve{\bd} - \breve{\bv}(\breve{\bz})\|, 
 \end{equation}
@@
 If we use the Wirtinger derivative instead of the classic notion of a derivative to define the gradient operator then 
-$\bv$ becomes analytic in $\bz$ and $\conj{\bz}$ as a whole \citep{Smirnov2015}.
-The fact that $\breve{\bv}$ is now analytic in its argument has the immediate consequence that we can use the complex variants of the Gauss-Newton
-and the Levenberg-Marquardt algorithm to solve Eq.~\eqref{eq:red_least_squares_complex}.
+$\bv$ becomes analytic in $\bz$ and $\conj{\bz}$ as a whole. The fact that $\bv$ is now analytic in its argument has the immediate consequence that it becomes possible to use the Gauss-Newton
+and the Levenberg-Marquardt algorithm to solve Eq.~\eqref{eq:red_least_squares_complex} \citep{Smirnov2015}.
+
+\subsection{Complex GN and LM}
+In this section we discuss how we can use the GN and LM algorithms to solve Eq.~\eqref{eq:red_least_squares_complex} (i.e. estimate $\breve{\bz}$). The complex Jacobian is defined to be
+\begin{equation}
+\label{eq:Jacobian}
+\bJ = \begin{bmatrix}
+       \bJ & \bJ^*\\
+       \conj{\bJ}^* & \conj{\bJ} 
+      \end{bmatrix},
+\end{equation}
+where 
+\begin{align}
+\bJ &= \frac{\partial \bv}{\partial \bz}, & \bJ^* &= \frac{\partial \bv}{\partial \conj{\bz}}. 
+\end{align}
+In Eq.~\eqref{eq:Jacobian}, $\frac{\partial}{\partial \bz}$ and $\frac{\partial}{\partial \conj{\bz}}$ respectively denote the cogradient operator and the conjugate cogradient operator (cite Sorber).
+The GN update step is now defined as:
+\begin{equation}
+ \Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\br}.
+\end{equation}
+The LM update rule is very similar, the major difference being a damping factor $\lambda$ is introduced:
+\begin{equation}
+\Delta \breve{\bz} = (\bJ^H\bJ + \lambda\bD)^{-1}\bJ^H\breve{\br},
+\end{equation}
+where $\bD=\bI\odot\bJ^H\bJ$. 
+NEED TO MENTION THE SILENT INDEX $k$ AND HOW THE UPDATE RULE IS ITERATIVE. ALSO MENTION THAT $\lambda$ IMPROVES THE CONVERGENCE PROPERTIES. WE WILL REFER TO $\bJ^H\bJ$ AS THE 
+HESSIAN.
+
+\subsection{Redundant Jacobian and Hessian}
+If we apply the definition in Eq.~\eqref{eq:Jacobian} to Eq.~\eqref{eq:red_least_squares_complex} we obtain the following analytic result:
+\begin{equation}
+\label{eq:Jacobian_red}
+\bJ = \begin{bmatrix}
+       \bM & \bN\\
+       \conj{\bN} & \conj{\bM}
+      \end{bmatrix},
+\end{equation}
+where
+\begin{equation}
+\bM =\begin{bmatrix}
+      \bM_1 & \bM_2
+     \end{bmatrix},
+\end{equation}
+and 
+\begin{equation}
+\bN = \begin{bmatrix}
+       \bN_1 & \bzero
+      \end{bmatrix}.
+\end{equation}
+Moreover,
+\begin{align}
+\left [ \bM_1 \right ]_{\alpha_{pq},j} &= y_{\phi_{pq}}\conj{g}_q\delta_p^j, & \left [ \bM_2 \right ]_{\alpha_{pq},j} &= g_p\conj{g}_q\delta_{\phi_{pq}}^j 
+\end{align}
+and
+\begin{equation}
+\left [ \bN_1 \right ]_{\alpha_{pq},j} = g_py_{\phi_{pq}}\delta^j_q. 
+\end{equation}
+
+It is now trivial to compute the Hessian $\bJ^H\bJ$ by using Eq.~\eqref{eq:Jacobian_red}. If we substitute Eq.~\eqref{eq:Jacobian_red} into $\bJ^H\bJ$
+we obtain
 

### 2016-10-06 paper.tex
started adding Jacobian and Hessian
added simplify  GN section

@@
 \newcommand{\bA}{\bmath{A}}
+\newcommand{\bB}{\bmath{B}}
+\newcommand{\bC}{\bmath{C}}
+\newcommand{\bE}{\bmath{E}}
+\newcommand{\bF}{\bmath{F}}
+\newcommand{\bG}{\bmath{G}}
 \newcommand{\br}{\bmath{r}}
@@
 \newcommand{\bD}{\bmath{D}}
+\newcommand{\bH}{\bmath{H}}
 \newcommand{\bN}{\bmath{N}}
@@
 
+STILL NEEDS LOADS OF WORK. REFERENCES.
+
 \subsection{Complex GN and LM}
@@
 \begin{equation}
+\label{eq:GN_update}
  \Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\br}.
@@
 we obtain 
+\begin{equation}
+\bH = \bJ^H\bJ = 
+\begin{bmatrix}
+\bA & \bB\\
+\conj{\bB} & \conj{\bA}
+\end{bmatrix},
+\end{equation}
+where
+
+\begin{align}
+\bA &= \begin{bmatrix} \bC & \bD\\ \bD^H & \bE \end{bmatrix}, & \bB &= \begin{bmatrix} \bF & \bG\\ \bG^T & \bzero \end{bmatrix},
+\end{align}
+
+\begin{equation}
+[\bC]_{ij} = 
+\begin{cases}
+ \sum_{k \neq i} \left | g_k \right |^2 \left | y_{\zeta_{ik}} \right |^2 & \textrm{if} ~ i=j\\
+ 0 & \textrm{otherwise}
+\end{cases},
+\end{equation}
+\begin{equation}
+[\bD]_{ij} = 
+\begin{cases}
+ g_i y_j^*  \left | g_{\psi_{ij}} \right |^2  & \textrm{if} ~ \psi_{ij}\neq0\\
+ 0 & \textrm{otherwise}
+\end{cases},
+\end{equation}
+
+\begin{equation}
+[\bE]_{ij} = 
+\begin{cases}
+ \sum_{pq \in \mathcal{PQ}_i} \left | g_p \right |^2 \left | g_q \right |^2  & \textrm{if} ~ i=j\\
+ 0 & \textrm{otherwise}
+\end{cases},
+\end{equation}
+\begin{equation}
+[\bF]_{ij} = 
+\begin{cases}
+ g_i g_j  \left | y_{\zeta_{ij}} \right |^2  & \textrm{if} ~ i \neq j\\
+ 0 & \textrm{otherwise}
+\end{cases},
+\end{equation}
+and
+\begin{equation}
+[\boldsymbol{G}]_{ij} = 
+\begin{cases}
+ g_i y_j  \left | g_{\xi_{ij}} \right |^2  & \textrm{if} ~ \xi_{ij}\neq0\\
+ 0 & \textrm{otherwise}
+\end{cases}.
+\end{equation}
+
+Moreover, $(*)^T$ denotes matrix transposition and
+
+\begin{equation}
+\mathcal{PQ}_i = \left\{pq\in\mathbb{N}^2|(\phi_{pq} = i) \right\}.
+\end{equation}
+
+\subsection{Simplifying GN}
+
+The following identities,
+\begin{align}
+\label{eq:identities}
+\frac{1}{3}\bJ\breve{\bz} &= \breve{\bv}, & \bJ^H\breve{\bv} &= (\bI\odot\bH)\breve{\bz} 
+\end{align}
+are trivially established by mechanically showing that the left hand side of each identity in Eq.~\eqref{eq:identities} is equal to its right hand side.
+
+The first identity in Eq.~\eqref{eq:identities} can now be used to simplify Eq.~\eqref{eq:GN_update} as follows:
+\begin{align}
+\Delta \breve{\bz} &= (\bJ^H\bJ)^{-1}\bJ^H(\breve{\bd}-\breve{\bv})\\
+&=(\bJ^H\bJ)^{-1}\bJ^H\breve{\bd}-\frac{1}{3}\breve{\bz}
+\end{align}
+
+HAVE TO MAKE SURE I DEFINE ALL THE MATH SYMBOLS THAT I USE. IF NOT HERE THEN EARLIER....
 

### 2016-10-06 paper.tex
JHr
added GN and LINCAL section

@@
 \newcommand{\bI}{\bmath{I}}
+\newcommand{\ba}{\bmath{a}}
+\newcommand{\bb}{\bmath{b}}
 \newcommand{\bzero}{\bmath{0}}
+%\newcommand{\bvarrho}{\bmath{\varrho}}
+%\newcommand{\bnu}{\bmath{\nu}}
+%\newcommand{\bvarphi}{\bmath{\varphi}}
 \newcommand{\conj}[1]{\overline{#1}}
@@
 \begin{cases}
- g_i y_j^*  \left | g_{\psi_{ij}} \right |^2  & \textrm{if} ~ \psi_{ij}\neq0\\
+ g_i \conj{y}_j  \left | g_{\psi_{ij}} \right |^2  & \textrm{if} ~ \psi_{ij}\neq0\\
  0 & \textrm{otherwise}
@@
 
+HAVE TO MAKE SURE I DEFINE ALL THE MATH SYMBOLS THAT I USE. IF NOT HERE THEN EARLIER....
+
+Furthermore, substituting Eq.~\eqref{eq:Jacobian_red} into $\bJ^H\breve{\br}$ results in
+\begin{equation}
+\bJ^H\breve{\br} = \begin{bmatrix}
+                   \ba & \bb & \conj{\ba} & \conj{\bb}\\
+                   \end{bmatrix}^T,
+\end{equation}
+where
+\begin{align}
+\left [ \ba \right ]_i &= \sum_{k\neq i} g_k \widetilde{y}_{ik}r_{ik},  & \left [ \bb \right ]_i &= \sum_{pq\in\mathcal{PQ}_i}\conj{g}_p g_q r_{pq},
+\end{align}
+and
+\begin{equation}
+\widetilde{y}_{ik} = 
+\begin{cases}
+\conj{y}_{\zeta_{ik}} & \textrm{if}~k > i\\
+y_{\zeta_{ik}} & \textrm{otherwise}
+\end{cases}.
+\end{equation}
+
+HAVE TO RECHECK THE ABOVE EQUATIONS DIRECTLY FROM THE CODE.
+
+HAVE TO STATE WHETHER COLUMN OR ROW VECTOR.
+
+% \begin{equation}
+% \boldsymbol{J}^H\breve{\boldsymbol{d}} = 
+% \begin{bmatrix}
+% \sum_{k\neq i } g_k x_{ik}d_{ik}\\
+% \sum_{pq \in \mathcal{I}_j} \conj{g}_p g_q d_{pq}\\
+% \downarrow^{*}
+% \end{bmatrix}
+% \begin{matrix}% matrix for right braces 
+% \coolrightbrace{\sum g_k x_{\zeta_{ik}}d_{ik}}{i = 1\cdots N}\\
+% \coolrightbrace{\sum \conj{g}_p g_q d_{pq}}{j = 1\cdots L}\\
+% \vphantom{\downarrow^{*}}
+% \end{matrix}
+% \end{equation}
+
 \subsection{Simplifying GN}
@@
 
-HAVE TO MAKE SURE I DEFINE ALL THE MATH SYMBOLS THAT I USE. IF NOT HERE THEN EARLIER....
+STILL NEED TO ADD z k+1 EXPRESSION.
+
+\subsection{GN and LINCAL}
+As we mentioned at the start of this section, the Wirtinger framework is not the only way in which we can ensure that $\bv$ is analytic in its argument. We could also 
+have divided the problem in Eq ... into its real and imaginary components and then solver for the real and imaginary parts of the parameters separately. There is a third way
+of dealing with the issue. We can split the problem in Eq into its phase and amplitude components and then solve for the amplitude and phase of the parameters separately.
+This can be achieved by first rewriting Eq. as
+
+\begin{equation}
+\label{eq:red_least_squares_lincal}
+\min_{\bmath{\varrho}} \|\breve{\br}\| = \min_{\bmath{\varrho}} \|\breve{\bd} - \breve{\bv}(\bmath{\varrho})\|, 
+\end{equation}
+where $\bmath{\varrho} = \left [\bmath{\eta},\bmath{\varphi},\widetilde{\bmath{\eta}}, \widetilde{\bmath{\varphi}}\right ]^T$. The Jacobian associated with 
+Eq.~\eqref{eq:red_least_squares_lincal} is equal to the Eq xxx in the LINCAL section.
+We have therefore shown that doing one iteration of LINCAL is equivalent to a basic GN update step; the GN update step was just re-derived in the Appendix from first principles.
+An easy why of improving the convergence properties of LINCAL would be to simply add the LM damping factor to the basic LINCAL update step.
+
+\subsection{LM and StEFCal}
 

### 2016-10-07 paper.tex
started stefcal section

@@
 
-\subsection{LM and StEFCal}
+\subsection{Diagonalization and StEFCal}
+StEFCal (Statistically Efficient Calibration) is an alternating direction implicit method, in which the measurement equation (cite the equation earlier in paper) is linearized by assuming that the gains are known, but
+that their conjugates are unknown (REMEMBER TO CITE ALL THE STEFCAL PAPERS MITCHELL ETC...).
+Under this assumption, the conjugates of the gains can be obtained trivially.
+Once the conjugates of the gains have been computed, the gains can be solved for.
+One can then alternate between computing the gains and their conjugates until convergence
+is reached.
+By correctly exploiting the symmetry inherent in the calibration problem this alternating approach can be replaced by a single iterative update step [MUST IMPROVE HERE].
+\cite{Salvini2014} also show that
+the convergence of the algorithm can be improved if for every even iteration the basic update step is modified as follows: the current gain estimate is replaced by the average of the current gain solution and the gain solution of the previous odd iteration.
+StEFCal lowers the computational complexity of normal skymodel-based calibration from O$(P^3)$ to O$(P^2)$.
+
+\citet{Smirnov2015} recently showed that one can also derive StEFCal by viewing normal skymodel-based calibration as a complex optimization problem.
+They first compute the skymodel-based calibration equivalent analytic expressions of $\bJ$,
+$\bH$ and $\bJ^H\breve{\br}$. While inspecting the skymodel-based $\bH$ it becomes apparent that its diagonal entries are much more significant than its off-diagonal entries.
+By approximating $\bH$ by its diagonal and substituting this approximate Hessian back into the GN update step they are able to re-derive the fundamental StEFCal update step. They then show that the 
+averaging update step used on even iterations can be derived by substituting the approximate $H$ into the LM update step, whilst setting $\lambda$ to one.
+
+The natural question now arises: is it possible to derive a redundant calibration variant of the SteFCal algorithm if we follow a similar approach to what \citet{Smirnov2015}
+proposed? We now attempt to try and answer that exact question:
+
+a very fast computational calibration This linearization trick improves
 

### 2016-10-07 paper.tex
nearing the end of redirivation of redundant stefcal

@@
 \begin{equation}
+\label{eq:LM_update}
 \Delta \breve{\bz} = (\bJ^H\bJ + \lambda\bD)^{-1}\bJ^H\breve{\br},
@@
 \begin{equation}
+\label{eq:red_H}
 \bH = \bJ^H\bJ = 
@@
 \begin{align}
+\label{eq:ab}
 \left [ \ba \right ]_i &= \sum_{k\neq i} g_k \widetilde{y}_{ik}r_{ik},  & \left [ \bb \right ]_i &= \sum_{pq\in\mathcal{PQ}_i}\conj{g}_p g_q r_{pq},
@@
 \label{eq:identities}
-\frac{1}{3}\bJ\breve{\bz} &= \breve{\bv}, & \bJ^H\breve{\bv} &= (\bI\odot\bH)\breve{\bz} 
+\frac{1}{3}\bJ\breve{\bz} &= \breve{\bv}, & \bJ^H\breve{\bv} &= \widetilde{\bH}\breve{\bz} 
 \end{align}
 are trivially established by mechanically showing that the left hand side of each identity in Eq.~\eqref{eq:identities} is equal to its right hand side.
+In Eq.~\eqref{eq:identities}, $\widetilde{\bH}$ denotes $\bI\odot\bH$.
 
-The first identity in Eq.~\eqref{eq:identities} can now be used to simplify Eq.~\eqref{eq:GN_update} as follows:
-\begin{align}
-\Delta \breve{\bz} &= (\bJ^H\bJ)^{-1}\bJ^H(\breve{\bd}-\breve{\bv})\\
-&=(\bJ^H\bJ)^{-1}\bJ^H\breve{\bd}-\frac{1}{3}\breve{\bz}
-\end{align}
-
-STILL NEED TO ADD z k+1 EXPRESSION.
+Replacing $\breve{\br}$ with $\breve{\bd}-\breve{\bv}$ in Eq.~\eqref{eq:GN_update} results in
+\begin{equation}
+\label{eq:temp_GN_eq}
+\Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H(\breve{\bd}-\breve{\bv}), 
+\end{equation}
+Substituting the first identity of Eq.~\eqref{eq:identities} into Eq.~\eqref{eq:temp_GN_eq} and simplifying we obtain
+\begin{equation}
+\label{eq:two_thrids}
+\Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd}-\frac{1}{3}\breve{\bz}.
+\end{equation}
+It now trivially follows from Eq.~\eqref{eq:two_thrids} that:
+\begin{equation}
+\breve{\bz}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{2}{3}\breve{\bz}_{k}. 
+\end{equation}
+%STILL NEED TO ADD z k+1 EXPRESSION.
+THE TAKE HOME HERE IS WE CAN CALCULATE THE UPDATE STEP WITHOUT CALCULATING THE RESIDUAL.
 
@@
 The natural question now arises: is it possible to derive a redundant calibration variant of the SteFCal algorithm if we follow a similar approach to what \citet{Smirnov2015}
-proposed? We now attempt to try and answer that exact question:
+proposed?
+
+Inspecting Eq.~\eqref{eq:red_H} reveals that in the case of redundant calibration the Hessian's diagonal entries are also more significant than its off-diagonal entries.
+If we substitute $\bJ^H\bJ$ with $\widetilde{\bH}$ (by its diagonal approximation) and replace $\breve{\br}$ with $\breve{\bd} - \breve{\bv}$ in Eq.~\eqref{eq:GN_update} we obtain
+\begin{equation}
+\label{eq:appr_GN}
+ \Delta \breve{\bz} \approx \widetilde{\bH}^{-1}\bJ^H(\breve{\bd}-\breve{\bv})
+\end{equation}
+Utilizing the second identity in Eq.~\eqref{eq:identities} allows us to simplify Eq.~\eqref{eq:appr_GN} to
+\begin{equation}
+  \Delta \breve{\bz} \approx \widetilde{\bH}^{-1}\bJ^H\breve{\bd}-\breve{\bz}.
+\end{equation}
+It now trivially follows that
+\begin{equation}
+ \breve{\bz}_{k+1} \approx \widetilde{\bH}^{-1}\bJ^H\breve{\bd}.
+\end{equation}
+We obtain a similar result if we repeat the above procedure, only in this case we use Eq.~\eqref{eq:LM_update} instead,
+namely
+\begin{align}
+\breve{\bz}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bz}_k,\label{eq:lambda}\\
+ &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bz}_k, \label{eq:alpha}  
+\end{align}
+where $\alpha = \frac{1}{1+\lambda}$. Moreover, $\lambda = \frac{1-\alpha}{\alpha}$. The analytic expression of $\bJ^H\breve{\bd}$ will be very similar to the analytic 
+expression of $\bJ^H\breve{\br}$, the only difference being that in Eq.~\eqref{eq:ab} the letter $r$ would be replaced by a $d$. If we substitute the analytic expression
+of $\bJ^H\breve{\bd}$ and $\widetilde{\bH}^{-1}$ (which is easily computed as the matrix is a diagonal matrix) into Eq.~\eqref{eq:alpha} we obtain the following two update rules:
+\begin{equation}
+\label{eq:g_update}
+g_{i}^{k+1} = \alpha \frac{\sum_{j\neq i} g_j^k \widetilde{y}_{ij}^{~\!\!k} d_{ij}}{\sum_{j\neq i} |g_j^k|^2|y_{\zeta_{ij}}^k|^2} + (1-\alpha) g_i^k, 
+\end{equation}
+and
+\begin{equation}
+\label{eq:y_update}
+y_{i}^{k+1} = \alpha \frac{\sum_{pq \in \mathcal{PQ}_i} \conj{g}_p^k g_q^k d_{pq}}{\sum_{pq \in \mathcal{PQ}_i}|g_p^k|^2|g_q^k|^2} + (1-\alpha) y_i^k 
+\end{equation}
+
+Interestingly enough, if we substitute $\alpha = 1$ into Eq.~\ref{eq:g_update} then it reduces to the skymodel-based StEFCal update rule .
+
+\begin{eqnarray}
+\begin{bmatrix} \Delta \boldsymbol{z}\\ \Delta \conj{\boldsymbol{z}} \end{bmatrix} &\approx& \widetilde{\boldsymbol{H}}^{-1}\boldsymbol{J}^H(\breve{\boldsymbol{d}}-\breve{\boldsymbol{v}})\\ 
+&=& \widetilde{\boldsymbol{H}}^{-1}\boldsymbol{J}^H\breve{\boldsymbol{d}} -  \widetilde{\boldsymbol{H}}^{-1}\widetilde{\boldsymbol{H}}\breve{\boldsymbol{z}}\\
+\breve{\boldsymbol{z}}_{k+1} &\approx& \widetilde{\boldsymbol{H}}^{-1}\boldsymbol{J}^H\breve{\boldsymbol{d}}
+\end{eqnarray}
+
+Levenberg-Marquardt:
+
+\begin{eqnarray}
+\breve{\boldsymbol{z}}_{k+1} &\approx& \frac{1}{1+\lambda}\widetilde{\boldsymbol{H}}^{-1}\boldsymbol{J}^H\breve{\boldsymbol{d}} + \frac{\lambda}{1+\lambda} \breve{\boldsymbol{z}}_k\\
+ &=& \alpha \widetilde{\boldsymbol{H}}^{-1}\boldsymbol{J}^H\breve{\boldsymbol{d}} + (1-\alpha)\breve{\boldsymbol{z}}_k 
+\end{eqnarray}
+
+with $\alpha = \frac{1}{1+\lambda}$. Moreover $\lambda = \frac{1-\alpha}{\alpha}$. So in Marthi and Chengular (2014) they use an $\alpha$ of 0.3, which corresponds to an LM damping factor $\lambda$
+of $2\frac{1}{3}$.
 

### 2016-10-07 paper.tex
interesting observation

@@
 \begin{equation}
-\label{eq:two_thrids}
+\label{eq:one_thrid}
 \Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd}-\frac{1}{3}\breve{\bz}.
 \end{equation}
-It now trivially follows from Eq.~\eqref{eq:two_thrids} that:
+It now trivially follows from Eq.~\eqref{eq:one_thrid} that:
 \begin{equation}
+\label{eq:two_thrids}
 \breve{\bz}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{2}{3}\breve{\bz}_{k}. 
@@
 \label{eq:y_update}
-y_{i}^{k+1} = \alpha \frac{\sum_{pq \in \mathcal{PQ}_i} \conj{g}_p^k g_q^k d_{pq}}{\sum_{pq \in \mathcal{PQ}_i}|g_p^k|^2|g_q^k|^2} + (1-\alpha) y_i^k 
+y_{i}^{k+1} = \alpha \frac{\sum_{pq \in \mathcal{PQ}_i} \conj{g}_p^k g_q^k d_{pq}}{\sum_{pq \in \mathcal{PQ}_i}|g_p^k|^2|g_q^k|^2} + (1-\alpha) y_i^k. 
 \end{equation}
+
+The question now arises what value should we use for $\alpha$? Although an exact mathematical proof for the optimal choice of $\alpha$ still eludes us, it is still possible to
+venture a clever guess by studying what the choice of $\alpha$ is in the case of StEFCal. The StEFCal equivalent of Eq.~\eqref{eq:two_thrids}
+is equal to
+\begin{equation}
+\label{eq:one_half}
+\breve{\bg}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{1}{2}\breve{\bg}_{k}. 
+\end{equation}
+Note that in Eq.~\eqref{eq:one_half} the influence that the estimate of the parameters from the previous iteration has on the current iteration is scaled by a half, while
+in Eq.~\eqref{eq:two_thrids} it is scaled by two thirds. Moreover, the StEFCal equivalent of Eq.~\eqref{eq:lambda} and Eq.~\eqref{eq:alpha} are
+\begin{align}
+\breve{\bg}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bg}_k,\label{eq:stef_lambda}\\
+ &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bg}_k, \label{eq:stef_alpha}  
+\end{align}
+
+On the odd iterations $\lambda$ is chosen to be zero, which corresponds to an $\alpha$ equal to one. On the even iteration $\lambda$ is chosen to be equal to one, which corresponds
+to an $\alpha$ equal to a half. On the odd iteration
+
+\begin{align}
+\breve{\bz}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bz}_k,\label{eq:lambda}\\
+ &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bz}_k, \label{eq:alpha}  
+\end{align}
 

### 2016-10-07 paper.tex
need to rewrite stefcal section

@@
 On the odd iterations $\lambda$ is chosen to be zero, which corresponds to an $\alpha$ equal to one. On the even iteration $\lambda$ is chosen to be equal to one, which corresponds
-to an $\alpha$ equal to a half. On the odd iteration
+to an $\alpha$ equal to a half. Note that, on the odd iteration the influence that the estimate of the parameters of the previous iteration has on the current estimate is scaled
+by a half. Which is exactly the same as in, Eq.~\eqref{eq:one_half}.
 
-\begin{align}
-\breve{\bz}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bz}_k,\label{eq:lambda}\\
- &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bz}_k, \label{eq:alpha}  
-\end{align}
+If we wish to scale $\breve{\bz}$ by two thirds in Eq.~\eqref{eq:alpha} we should choose $\alpha$ equal to a third, which corresponds to a $\lambda$ equal to two.
 
-Interestingly enough, if we substitute $\alpha = 1$ into Eq.~\ref{eq:g_update} then it reduces to the skymodel-based StEFCal update rule .
+At this point there is no way of knowing if this is mere coincidence. DEF REWRITE THIS.
+
+REWRTIE THIS IN A DIFFERENT WAY. START WITH CHENGULAR AND JUST MENTION THE LINK TO STEFCAL'S CHOICE OF ALPHA.
+
+The update rules in Eq.~\eqref{eq:lambda} and Eq.~\eqref{eq:alpha} have been derived before, by Chengular... The approach they followed is discussed in the Appendix.
+
+Interestingly enough, if we substitute $\alpha = 1$ into Eq.~\ref{eq:g_update} then it reduces to the skymodel-based StEFCal update rule.
+
+In an attempt to merge the terminology that has independently develop in the general calibration and redundant calibration literature and to emphasize the close
+relationship between Stefcal and the approach derived by I will use the label Redunandat SteFCal (R-Stefcal) to refer to the NLS method proposed by ...
 

### 2016-10-11 paper.tex
need to rewrite stefcal section
started conjugate gradient section

@@
 \usepackage{graphicx}
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
 
@@
 \newcommand{\bb}{\bmath{b}}
+\newcommand{\bx}{\bmath{x}}
+\newcommand{\bp}{\bmath{p}}
 \newcommand{\bzero}{\bmath{0}}
@@
 
+\section{Conjugate Gradient Method}
+
+The conjugate gradient method is an iterative method which is used to solve large linear systems, i.e.
+\begin{equation}
+\label{eq:linear_system}
+\bb = \bA\bx,
+\end{equation}
+where $\bb$ is a known complex column vector, $\bx$ is the unknown complex column vector with the same length as $\bb$, and $\bA$ is a square complex matrix with appropriate dimensions.  
+The conjugate gradient method can in general only be applied to the above linear system if $\bA$ is a square positive-definite or a positive semi-definite Hermitian matrix. If
+$\bA$ is a square positive semi-definite Hermitian matrix, then the conjugate gradient method only converges if $\bb$ is in the column range of $\bA$ (cite Lu2016). Moreover, using the 
+conjugate gradient method to solve Eq.~\eqref{eq:linear_system} will only be computationally advantageous if $\bA$ is also sparse (contain many zero entries).
+
+The Conjugate Gradient method was independently discovered by Hestenes and Stiefel. They later published a joint paper, n, which is now considered as the seminal
+reference on CG. Kaniel was able to determined convergence bounds for CG using Chebyshev polynomial. A much more in depth analysis of CG convergence was conducted 
+by van erSluis and van der Vorst. Reid is credited for popularizing CG as an iterative method that can be used to solve sparse linear systems.
+
+In the early fifties, the method of Conjugate Gradients
+was discovered independently by Hestenes [10] and Stiefel [15]; shortly thereafter, they jointly published
+what is considered the seminal reference on CG [11]. Convergence bounds for CG in terms of Chebyshev
+polynomials were developed by Kaniel [12]. A more thorough analysis of CG convergence is provided by
+van der Sluis and van der Vorst [16]. CG was popularized as an iterative method for large, sparse matrices
+by Reid [13] in 1971.
+
+\begin{algorithm}
+\caption{Conjugate Gradient Method}\label{algo:extrap_algo}
+\begin{algorithmic}[1]
+\State $\br_0 \gets \bA\bx_0 - \bb$
+\State $\bp_0 \gets -\br_0$
+\State $\bmath{u} \gets \text{linspace}(u_{\text{min}},u_{\text{max}},u_{\text{dim}})$
+\State $\bmath{v} \gets \text{linspace}(v_{\text{min}},v_{\text{max}},v_{\text{dim}})$ 
+\State $g_{pq} \gets \text{zeros}((u_{\text{dim}},v_{\text{dim}}),\text{dtype}=\text{complex})$ \Comment {define your 2D extrapolated regular $uv$-grid}
+\State $i \gets 1$
+\For{$i\leq u_{\text{dim}}$}  
+\State $j \gets 1$
+\For{$j\leq v_{\text{dim}}$}  
+\State $\bb \gets (\bmath{u}_i,\bmath{v}_j)^T$
+%\State $\Rcal \gets \left\{ \sum_{k\in\mathcal{K}}A_k e^{-2\pi i [\widehat{X}_{rs} \circ \widehat{X}^{-1}_{pq}(\bb)]^T\cdot\bs_k} \right \}_{rs}$
+%\State $\Mcal \gets \left\{ \sum_{h\in\mathcal{H}}A_h e^{-2\pi i [\widehat{X}_{rs} \circ \widehat{X}^{-1}_{pq}(\bb)]^T\cdot\bs_h} \right \}_{rs}$
+%\Comment {with $\mathcal{K}$ containing the index values of all the sources in the field of view and $\mathcal{L}$ containing the index values of the sources that are included in the
+%calibration model}
+%\State $\Gcal \gets \text{calibrate}(\Rcal,\Mcal)$
+%\Comment {any calibration algorithm can be used here (we assume here that the algorithm outputs $\Gcal$)}
+%\State $g_{pq}^{ij} \gets [\Gcal]_{pq}$
+\State $j \gets j + 1$
+\EndFor
+\State $i \gets i + 1$
+\EndFor
+\end{algorithmic}
+\end{algorithm}
+
 \section{Least Squares}
@@
 
-Interestingly enough, if we substitute $\alpha = 1$ into Eq.~\ref{eq:g_update} then it reduces to the skymodel-based StEFCal update rule.
+Interestingly enough, if we substitute $\alpha = 1$ into Eq.~\ref{eq:g_update} then it reduces to the skymodel-based StEFCal update rule. In Chengular they propose 
+an $\alpha = 0.3$.
 
@@
 
-\begin{eqnarray}
-\begin{bmatrix} \Delta \boldsymbol{z}\\ \Delta \conj{\boldsymbol{z}} \end{bmatrix} &\approx& \widetilde{\boldsymbol{H}}^{-1}\boldsymbol{J}^H(\breve{\boldsymbol{d}}-\breve{\boldsymbol{v}})\\ 
-&=& \widetilde{\boldsymbol{H}}^{-1}\boldsymbol{J}^H\breve{\boldsymbol{d}} -  \widetilde{\boldsymbol{H}}^{-1}\widetilde{\boldsymbol{H}}\breve{\boldsymbol{z}}\\
-\breve{\boldsymbol{z}}_{k+1} &\approx& \widetilde{\boldsymbol{H}}^{-1}\boldsymbol{J}^H\breve{\boldsymbol{d}}
-\end{eqnarray}
-
-Levenberg-Marquardt:
-
-\begin{eqnarray}
-\breve{\boldsymbol{z}}_{k+1} &\approx& \frac{1}{1+\lambda}\widetilde{\boldsymbol{H}}^{-1}\boldsymbol{J}^H\breve{\boldsymbol{d}} + \frac{\lambda}{1+\lambda} \breve{\boldsymbol{z}}_k\\
- &=& \alpha \widetilde{\boldsymbol{H}}^{-1}\boldsymbol{J}^H\breve{\boldsymbol{d}} + (1-\alpha)\breve{\boldsymbol{z}}_k 
-\end{eqnarray}
+****\\
 
@@
 of $2\frac{1}{3}$.
-
-a very fast computational calibration This linearization trick improves
-
-cite the complex paper has proposed to extend the problem into [RW HERE NEED TO STATE BETTER] its conjugate domain and then to solve for the parameters and there complex conjugates.
-This can be accomplished
-if we introduce Wirtinger derivatives. (I have to look into how to write this better).
-
-The solution to Eq. can now be obtained by using a gradient type of minimization algorithm, like GN and LM.
-
-The GN update rule is given by\\
-
-GN UPDATE RULE
 

### 2016-10-11 paper.tex
added more text to the conjugate gradient section

@@
 The Conjugate Gradient method was independently discovered by Hestenes and Stiefel. They later published a joint paper, n, which is now considered as the seminal
-reference on CG. Kaniel was able to determined convergence bounds for CG using Chebyshev polynomial. A much more in depth analysis of CG convergence was conducted 
-by van erSluis and van der Vorst. Reid is credited for popularizing CG as an iterative method that can be used to solve sparse linear systems.
-
-In the early fifties, the method of Conjugate Gradients
-was discovered independently by Hestenes [10] and Stiefel [15]; shortly thereafter, they jointly published
-what is considered the seminal reference on CG [11]. Convergence bounds for CG in terms of Chebyshev
-polynomials were developed by Kaniel [12]. A more thorough analysis of CG convergence is provided by
-van der Sluis and van der Vorst [16]. CG was popularized as an iterative method for large, sparse matrices
-by Reid [13] in 1971.
+reference on CG. Kaniel was able to determine convergence bounds for CG using Chebyshev polynomials. A much more in depth analysis of CG convergence was conducted 
+by van erSluis and van der Vorst. [Reid is credited for popularizing CG as an iterative method that can efficiently solve large sparse linear systems] RW.
 
@@
 \begin{algorithmic}[1]
-\State $\br_0 \gets \bA\bx_0 - \bb$
-\State $\bp_0 \gets -\br_0$
-\State $\bmath{u} \gets \text{linspace}(u_{\text{min}},u_{\text{max}},u_{\text{dim}})$
-\State $\bmath{v} \gets \text{linspace}(v_{\text{min}},v_{\text{max}},v_{\text{dim}})$ 
-\State $g_{pq} \gets \text{zeros}((u_{\text{dim}},v_{\text{dim}}),\text{dtype}=\text{complex})$ \Comment {define your 2D extrapolated regular $uv$-grid}
-\State $i \gets 1$
-\For{$i\leq u_{\text{dim}}$}  
-\State $j \gets 1$
-\For{$j\leq v_{\text{dim}}$}  
-\State $\bb \gets (\bmath{u}_i,\bmath{v}_j)^T$
-%\State $\Rcal \gets \left\{ \sum_{k\in\mathcal{K}}A_k e^{-2\pi i [\widehat{X}_{rs} \circ \widehat{X}^{-1}_{pq}(\bb)]^T\cdot\bs_k} \right \}_{rs}$
-%\State $\Mcal \gets \left\{ \sum_{h\in\mathcal{H}}A_h e^{-2\pi i [\widehat{X}_{rs} \circ \widehat{X}^{-1}_{pq}(\bb)]^T\cdot\bs_h} \right \}_{rs}$
-%\Comment {with $\mathcal{K}$ containing the index values of all the sources in the field of view and $\mathcal{L}$ containing the index values of the sources that are included in the
-%calibration model}
-%\State $\Gcal \gets \text{calibrate}(\Rcal,\Mcal)$
-%\Comment {any calibration algorithm can be used here (we assume here that the algorithm outputs $\Gcal$)}
-%\State $g_{pq}^{ij} \gets [\Gcal]_{pq}$
-\State $j \gets j + 1$
-\EndFor
-\State $i \gets i + 1$
+\State $\br \gets \bA\bx_0 - \bb$
+\State $\bp \gets -\br_0$
+\For{$k\leq k_{\textrm{max}}$}
+\State $\alpha_k \gets \frac{\br_k^H\br_k}{\bp_k^H\bA\bp_k}$
+\State $\bx_{k+1} \gets \bx_k + \alpha_k\bp_k$
+\State $\br_{k+1} \gets \br_k - \alpha_k\bA\bp_k$
+\If{$\|\br_{k+1}\|<\epsilon\|\br_0\|$}
+\State break
+\EndIf
+\State $\beta_{k} \gets \frac{\br_{k+1}^H\br_{k+1}}{\br_k^H\br_k}$
+\State $\bp_{k+1} \gets \br_{k+1} + \beta_{k}\bp_k$
+\State $k \gets k + 1$
 \EndFor
@@
 \end{algorithm}
+
+In general, when $\bA$ is neither square, positive semi-definite or Hermitian we can use the normal equation 
+\begin{equation}
+\label{eq:normal_equation}
+\bA^H\bA\bx = \bA^H\bb 
+\end{equation}
+to estimate the $\bx$ that minimizes
+\begin{equation}
+\|\bb-\bA\bx\|. 
+\end{equation}
+The CG method is gaurenteed to converge if we use it to solve Eq.~\eqref{eq:normal_equation}.
 

### 2016-10-11 paper.tex
paper.tex

@@
 
-In general, when $\bA$ is neither square, positive semi-definite or Hermitian we can use the normal equation 
+In general, when $\bA$ has dimensions $m\times n$, with $m > n$ we can use the normal equation 
 \begin{equation}
@@
 \begin{equation}
+\label{eq:norm_ls}
 \|\bb-\bA\bx\|. 
+\end{equation}
+If $\bA^H\bA$ is invertiable then the solution that minimizes Eq.~\eqref{eq:norm_ls} is unique. In contrast, if $\bA^H\bA$ is singular there are an infinite amount of solutions
+that minimizes Eq.~\eqref{eq:norm_ls}. Moreover, we can use the CG method to solve Eq.~\eqref{eq:normal_equation}, since $\bA^H\bA$ is Hermitian by construction 
+and $\bA^H\bb$ is in the column range of $\bA^H\bA$.
+\begin{equation}
+\bA^H\bb \in \mathcal{R}(\bA^H) \implies \bA^H\bb \in \mathcal{R}(\bA^H\bA).   
 \end{equation}

### 2016-10-11 paper.tex
paper.tex

@@
 \end{equation}
-The CG method is gaurenteed to converge if we use it to solve Eq.~\eqref{eq:normal_equation}.
+
+The computational complexity of CG is $\mathbb{O}(\sqrt{\kappa}m)$, where $\kappa$ is the generalized specctral condition number of the matrix $\bA$ and $m$ is the number of
+non-zero entries in the matrix $\bA$. The mathematical
+definition of $\kappa(\bA)$ is
+\begin{equation}
+\kappa(\bA) = \frac{\lambda_{\textrm{max}}}{\lambda^{+}_{\textrm{min}}}, 
+\end{equation}
+where $\lambda_{\textrm{max}}$ denotes the largest eigenvalue of $\bA$ and $\lambda^{+}_{\textrm{min}}$ presents the smallest positive eigenvalue. 
+Interpreting the computational complexity equation further.
+Note from Algorithm 1 that the CG method has one major outer loop.
+In the worst possible case, the amount of iterations needed to converge is
+proportional to $\sqrt{\kappa}$. Also notice that the most expensive 
+operation that is conducted within each loop is the vector-matrix product, which has complexity $\mathbb{O}(m)$, where $m$ is the number of non-zero entries in the matrix.
+The computational complexity of the CG method is therefore 
+\begin{equation}
+\mathbb{O}(\sqrt{\kappa}m) 
+\end{equation}
+
+It should now be clear to
+the reader why the CG method is espicially applicable to sparse linear systems, since in those cases $m < N^2$. 
+We can improve the computational complexity of the CG method by preconditioning. 
+[Preconditioning is a technique for improving the condition number of a matrix. Suppose that ' is a
+symmetric, positive-definite matrix that approximates, but is easier to invert. We can solve indirectly by solving] - RW directly coppied from source
+\begin{equation}
+\bM^{-1}\bA\bx = \bM^{-1}\bb
+\end{equation}
+
+The best precondtioner would be the preconditioner which makes $\kappa$ unity, i.e. completly
+eliminating it from the computational complexity equation.
+The only choice of $\bM$ which is capable of best way to do this is to chooose as preconditioner the matrix $(\bA)^{-1}$, which is what we would like to determine in the frist place.
+If the diagonal entries of $\bA$ are more significant than its off-diagonal entries, then the Jacobian preconditioner is a good preconditioner to use.
 

### 2016-10-13 paper.tex
very rough conjugate gradient section
started cleaning up CG section

@@
 \section{Conjugate Gradient Method}
-
-The conjugate gradient method is an iterative method which is used to solve large linear systems, i.e.
+The Conjugate Gradient (CG) method was independently discovered by \citet{Hestenes1973} and \citet{Stiefel1952}. They later published a joint paper, which is now considered as the seminal
+reference on CG \citep{Hestenes1952}. The iterative CG method is used to solve a particular class of linear systems. The CG method is generally used to solve
 \begin{equation}
@@
 \end{equation}
-where $\bb$ is a known complex column vector, $\bx$ is the unknown complex column vector with the same length as $\bb$, and $\bA$ is a square complex matrix with appropriate dimensions.  
-The conjugate gradient method can in general only be applied to the above linear system if $\bA$ is a square positive-definite or a positive semi-definite Hermitian matrix. If
-$\bA$ is a square positive semi-definite Hermitian matrix, then the conjugate gradient method only converges if $\bb$ is in the column range of $\bA$ (cite Lu2016). Moreover, using the 
-conjugate gradient method to solve Eq.~\eqref{eq:linear_system} will only be computationally advantageous if $\bA$ is also sparse (contain many zero entries).
+where $\bb\in\mathbb{C}^P$ is a known complex column vector of size $P$, $\bx\in\mathbb{C}^P$ is an unknown complex column vector with the same length as $\bb$, and $\bA\in\mathbb{C}^{P\times P}$ is a square positive-definite Hermitian matrix with the appropriate dimensions.
+Moreover, using the CG method to solve Eq.~\eqref{eq:linear_system} will only be computationally advantageous if $\bA$ is also sparse (contain many zero entries).
+Using the iterative CG method to solve large sparse linear sytems was popularized by \citet{Reid1971}. A good tutorial on the CG method can be found in \citep{Shewchuk1994}.
 
-The Conjugate Gradient method was independently discovered by Hestenes and Stiefel. They later published a joint paper, n, which is now considered as the seminal
-reference on CG. Kaniel was able to determine convergence bounds for CG using Chebyshev polynomials. A much more in depth analysis of CG convergence was conducted 
-by van erSluis and van der Vorst. [Reid is credited for popularizing CG as an iterative method that can efficiently solve large sparse linear systems] RW.
+%If $\bA$ is a square positive semi-definite Hermitian matrix, then the conjugate gradient method only converges if $\bb$ is in the column range of $\bA$ (cite Lu2016).
 
 \begin{algorithm}
-\caption{Conjugate Gradient Method}\label{algo:extrap_algo}
+\caption{Conjugate Gradient Method}\label{algo:CG}
 \begin{algorithmic}[1]
@@
 
-In general, when $\bA$ has dimensions $m\times n$, with $m > n$ we can use the normal equation 
-\begin{equation}
-\label{eq:normal_equation}
-\bA^H\bA\bx = \bA^H\bb 
-\end{equation}
-to estimate the $\bx$ that minimizes
-\begin{equation}
-\label{eq:norm_ls}
-\|\bb-\bA\bx\|. 
-\end{equation}
-If $\bA^H\bA$ is invertiable then the solution that minimizes Eq.~\eqref{eq:norm_ls} is unique. In contrast, if $\bA^H\bA$ is singular there are an infinite amount of solutions
-that minimizes Eq.~\eqref{eq:norm_ls}. Moreover, we can use the CG method to solve Eq.~\eqref{eq:normal_equation}, since $\bA^H\bA$ is Hermitian by construction 
-and $\bA^H\bb$ is in the column range of $\bA^H\bA$.
-\begin{equation}
-\bA^H\bb \in \mathcal{R}(\bA^H) \implies \bA^H\bb \in \mathcal{R}(\bA^H\bA).   
-\end{equation}
-
+\subsection{Computational Complexity}
+\citet{Kaniel1966} was able to determine convergence bounds for CG using Chebyshev polynomials. A much more in depth analysis of CG convergence can be found in
+\citep{Sluis1986}.
 The computational complexity of CG is $\mathbb{O}(\sqrt{\kappa}m)$, where $\kappa$ is the generalized specctral condition number of the matrix $\bA$ and $m$ is the number of
@@
 If the diagonal entries of $\bA$ are more significant than its off-diagonal entries, then the Jacobian preconditioner is a good preconditioner to use. 
+State better that the choice of preconditioner, the value of kappa and what the effect of this would be on the computational complexity of the method.
+
+In general, when $\bA$ has dimensions $m\times n$, with $m > n$ we can use the normal equation 
+\begin{equation}
+\label{eq:normal_equation}
+\bA^H\bA\bx = \bA^H\bb 
+\end{equation}
+to estimate the $\bx$ that minimizes
+\begin{equation}
+\label{eq:norm_ls}
+\|\bb-\bA\bx\|. 
+\end{equation}
+If $\bA^H\bA$ is invertiable then the solution that minimizes Eq.~\eqref{eq:norm_ls} is unique. In contrast, if $\bA^H\bA$ is singular there are an infinite amount of solutions
+that minimizes Eq.~\eqref{eq:norm_ls}. Moreover, we can use the CG method to solve Eq.~\eqref{eq:normal_equation}, since $\bA^H\bA$ is Hermitian by construction 
+and $\bA^H\bb$ is in the column range of $\bA^H\bA$.
+\begin{equation}
+\bA^H\bb \in \mathcal{R}(\bA^H) \implies \bA^H\bb \in \mathcal{R}(\bA^H\bA).   
+\end{equation}
 

### 2016-10-13 paper.tex
more progress on CG section

@@
 The Conjugate Gradient (CG) method was independently discovered by \citet{Hestenes1973} and \citet{Stiefel1952}. They later published a joint paper, which is now considered as the seminal
-reference on CG \citep{Hestenes1952}. The iterative CG method is used to solve a particular class of linear systems. The CG method is generally used to solve
+reference on CG \citep{Hestenes1952}. The iterative CG method is used to solve a particular class of linear system. The CG method is generally used to solve
 \begin{equation}
@@
 Using the iterative CG method to solve large sparse linear sytems was popularized by \citet{Reid1971}. A good tutorial on the CG method can be found in \citep{Shewchuk1994}.
+The iterative CG method is presented in Algorithm \ref{algo:CG}.
 
@@
 \begin{algorithm}
-\caption{Conjugate Gradient Method}\label{algo:CG}
+\caption{Conjugate Gradient Method.
+Inputs: $\bA$, $\bb$, a starting value $x$, maximum number of iterations $k_{\textrm{max}}$ and an error tolerance $\epsilon<1$ Output: $\bx$ the solution to Eq.~\eqref{eq:linear_system}.
+\citep{Shewchuk1994}.}\label{algo:CG}
 \begin{algorithmic}[1]
-\State $\br \gets \bA\bx_0 - \bb$
-\State $\bp \gets -\br_0$
-\For{$k\leq k_{\textrm{max}}$}
-\State $\alpha_k \gets \frac{\br_k^H\br_k}{\bp_k^H\bA\bp_k}$
-\State $\bx_{k+1} \gets \bx_k + \alpha_k\bp_k$
-\State $\br_{k+1} \gets \br_k - \alpha_k\bA\bp_k$
-\If{$\|\br_{k+1}\|<\epsilon\|\br_0\|$}
-\State break
-\EndIf
-\State $\beta_{k} \gets \frac{\br_{k+1}^H\br_{k+1}}{\br_k^H\br_k}$
-\State $\bp_{k+1} \gets \br_{k+1} + \beta_{k}\bp_k$
+\State $k \gets 0$
+\State $\br \gets \bb - \bA\bx$
+\State $\bp \gets \br$
+\State $\delta_{\textrm{new}} \gets \br^H\br$
+\State $\delta_0 \gets \delta_{\textrm{new}}$
+\While{$k\leq k_{\textrm{max}}$ and $\delta_{\textrm{new}} > \epsilon^2\delta_0$}
+\State $\alpha_k \gets \frac{\delta_{\textrm{new}}}{\bp^H\bA\bp}$
+\State $\bx \gets \bx + \alpha_k\bp$
+\State $\br \gets \br - \alpha\bA\bp$
+\State $\delta_{\textrm{old}} \gets \delta_{\textrm{new}}$ 
+\State $\delta_{\textrm{new}} \gets \br^H\br$
+\State $\beta_{k} \gets \frac{\delta_{\textrm{new}}}{\delta_{\textrm{old}}}$
+\State $\bp \gets \br + \beta\bp$
 \State $k \gets k + 1$
-\EndFor
+\EndWhile
 \end{algorithmic}
@@
 \citet{Kaniel1966} was able to determine convergence bounds for CG using Chebyshev polynomials. A much more in depth analysis of CG convergence can be found in
-\citep{Sluis1986}.
+\citep{Sluis1986}. As can be seen from Algorithm \ref{algo:CG}, the CG method makes use of only one major outer loop. It turns out that in the worst possible case 
+the number of major iterations that is required by CG to converge is proportional to $\sqrt{\kappa}$.
+
 The computational complexity of CG is $\mathbb{O}(\sqrt{\kappa}m)$, where $\kappa$ is the generalized specctral condition number of the matrix $\bA$ and $m$ is the number of

### 2016-10-13 paper.tex
improved CG section

@@
 \citet{Kaniel1966} was able to determine convergence bounds for CG using Chebyshev polynomials. A much more in depth analysis of CG convergence can be found in
-\citep{Sluis1986}. As can be seen from Algorithm \ref{algo:CG}, the CG method makes use of only one major outer loop. It turns out that in the worst possible case 
-the number of major iterations that is required by CG to converge is proportional to $\sqrt{\kappa}$.
-
-The computational complexity of CG is $\mathbb{O}(\sqrt{\kappa}m)$, where $\kappa$ is the generalized specctral condition number of the matrix $\bA$ and $m$ is the number of
-non-zero entries in the matrix $\bA$. The mathematical
-definition of $\kappa(\bA)$ is
+\citep{Sluis1986}. As can be seen from Algorithm \ref{algo:CG}, the CG method makes use of only one major outer loop (line 6 of Algorithm \ref{algo:CG}). It turns out that at worst 
+the number of major outer iterations that CG requires to converge\footnote{If the $\bA$-norm is used to calculate the error.} is proportional to $\sqrt{\kappa}$, where $\kappa$ is the generalized specctral condition number of the matrix $\bA$ \citep{Lu2015}.
+The mathematical definition of $\kappa(\bA)$ is
 \begin{equation}
+\label{eq:kappa}
 \kappa(\bA) = \frac{\lambda_{\textrm{max}}}{\lambda^{+}_{\textrm{min}}}, 
 \end{equation}
-where $\lambda_{\textrm{max}}$ denotes the largest eigenvalue of $\bA$ and $\lambda^{+}_{\textrm{min}}$ presents the smallest positive eigenvalue. 
-Interpreting the computational complexity equation further.
-Note from Algorithm 1 that the CG method has one major outer loop.
-In the worst possible case, the amount of iterations needed to converge is
-proportional to $\sqrt{\kappa}$. Also notice that the most expensive 
-operation that is conducted within each loop is the vector-matrix product, which has complexity $\mathbb{O}(m)$, where $m$ is the number of non-zero entries in the matrix.
-The computational complexity of the CG method is therefore 
+where $\lambda_{\textrm{max}}$ and $\lambda^{+}_{\textrm{min}}$ respectively denote the largest eigenvalue and the smallest positive eigenvalue of $\bA$.
+Note that, since $\bA$ is positive-definite all its eigenvalues will be positive. The reason why it is required to spcifically add the positivity requirement to Eq.~\eqref{eq:kappa} will become
+apparent after reading Sec \textbf{??}.
+Furthermore, the most expensive operation that is conducted within each major loop of Algorithm~\ref{algo:CG} is the vector-matrix product $\bA\bp$, which has a complexity of $\mathbb{O}(m)$, where $m$ is the number of non-zero entries in $\bA$.
+The computational complexity of CG is therefore 
 \begin{equation}
-\mathbb{O}(\sqrt{\kappa}m) 
+\label{eq:cg_bound}
+\mathbb{O}(\sqrt{\kappa}m). 
 \end{equation}
+It should now also be clear to the reader why the CG method is espicially applicable to sparse linear systems, since in those cases $m < P^2$.
 
-It should now be clear to
-the reader why the CG method is espicially applicable to sparse linear systems, since in those cases $m < N^2$. 
-We can improve the computational complexity of the CG method by preconditioning. 
-[Preconditioning is a technique for improving the condition number of a matrix. Suppose that ' is a
-symmetric, positive-definite matrix that approximates, but is easier to invert. We can solve indirectly by solving] - RW directly coppied from source
+\subsection{Preconditioning}
+Preconditioning is a technique which allows us to improve the condition number of a matrix. Let $\bM$ be a positive-definite Hermitian matrix which approximates $\bA$.
+Moreover, the inverse of $\bM$ must also be much easier to compute than the inverse of $\bA$. We can now indirectly solve for $\bx$ by using 
 \begin{equation}
-\bM^{-1}\bA\bx = \bM^{-1}\bb
+\label{eq:prec}
+\bM^{-1}\bA\bx = \bM^{-1}\bb.
 \end{equation}
+The logic behind using Eq.~\eqref{eq:prec} rather than Eq.~\eqref{eq:linear_system} to obtain $\bx$ is that if $\bM$ is a good preconditioner then $\kappa(\bM^{-1}\bA) \ll \kappa(\bA)$, i.e.
+by preconditioning we reduce $\kappa$ which improves the convergence rate of CG. The best choice of $\bM$ would be $\bA$, which would reduce $\kappa$ to unity which reduces
+Eq.~\eqref{eq:cg_bound} to $\mathbb{O}(m)$. But of course, if we knew the inverse of $\bA$ then obtaining $\bx$ would be trivial and we would not need CG or any other type of 
+linear solver for that mattter to compute it in the first place. If the diagonal entries of $\bA$ is much more significant than its off-diagonal entries then the 
+Jacobian preconditioner is a good choice of $\bM$. The Jacobian preconditioner is computed as follows:
+\begin{equation}
+\bM = \bA\odot\bI, 
+\end{equation}
+where $\bI$ is the identity matrix. Since the Jacobian preconditioner is a diagonal matrix is inverse can be computed trivially.
+
+Using preconditioning will therefore improve the 
+, which 
+implies that the computational complexity of the CG algorithm will be less in the case of  Eq.~\eqref{eq:prec} than in the case of Eq.~\eqref{eq:linear_system}.
 

### 2016-10-13 paper.tex
improved CG section

@@
 The logic behind using Eq.~\eqref{eq:prec} rather than Eq.~\eqref{eq:linear_system} to obtain $\bx$ is that if $\bM$ is a good preconditioner then $\kappa(\bM^{-1}\bA) \ll \kappa(\bA)$, i.e.
-by preconditioning we reduce $\kappa$ which improves the convergence rate of CG. The best choice of $\bM$ would be $\bA$, which would reduce $\kappa$ to unity which reduces
-Eq.~\eqref{eq:cg_bound} to $\mathbb{O}(m)$. But of course, if we knew the inverse of $\bA$ then obtaining $\bx$ would be trivial and we would not need CG or any other type of 
-linear solver for that mattter to compute it in the first place. If the diagonal entries of $\bA$ is much more significant than its off-diagonal entries then the 
+by preconditioning we reduce $\kappa$ which improves the convergence rate of CG.
+The best choice of $\bM$ would be $\bA$, which would reduce $\kappa$ (as the largest and smallest eignevalue of the Identity matrix is equal to one) to unity which reduces
+Eq.~\eqref{eq:cg_bound} to $\mathbb{O}(m)$. But of course, if we knew the inverse of $\bA$ then obtaining $\bx$ would be trivial and there would be no need for the CG method in the first place. 
+The better $\bM$ approximates $\bA$ the better preconditioner it will be. However, if it is not much less expensive to 
+compute the inverse of $\bM$ thant that of $\bA$ we gain nothing, as we still need to
+
+If the diagonal entries of $\bA$ is much more significant than its off-diagonal entries then the 
 Jacobian preconditioner is a good choice of $\bM$. The Jacobian preconditioner is computed as follows:

### 2016-10-14 paper.tex
started to clean up normal equation

--- 2016-10-13 paper.tex
@@
 \citet{Kaniel1966} was able to determine convergence bounds for CG using Chebyshev polynomials. A much more in depth analysis of CG convergence can be found in
-\citep{Sluis1986}. As can be seen from Algorithm \ref{algo:CG}, the CG method makes use of only one major outer loop. It turns out that in the worst possible case 
-the number of major iterations that is required by CG to converge is proportional to $\sqrt{\kappa}$.
+\citep{Sluis1986}. As can be seen from Algorithm \ref{algo:CG}, the CG method makes use of a major outer loop (line 6 of Algorithm \ref{algo:CG}). It turns out that at worst 
+the number of major iterations that CG requires to converge\footnote{If the $\bA$-norm is used to calculate the error.} is proportional to $\sqrt{\kappa}$, where $\kappa$ is the specctral condition number of the matrix $\bA$.
+The mathematical definition of $\kappa(\bA)$ is
+\begin{equation}
+\label{eq:kappa}
+\kappa(\bA) = \frac{\lambda_{\textrm{max}}}{\lambda_{\textrm{min}}}, 
+\end{equation}
+where $\lambda_{\textrm{max}}$ and $\lambda_{\textrm{min}}$ respectively denote the largest and the smallest eigenvalue of $\bA$.
+Moreover, the most expensive operation that is conducted within each major loop of Algorithm~\ref{algo:CG} is the vector-matrix product $\bA\bp$, which has a complexity of $\mathbb{O}(m)$, where $m$ is the number of non-zero entries in $\bA$.
+The computational complexity of CG is therefore 
+\begin{equation}
+\label{eq:cg_bound}
+\mathbb{O}(\sqrt{\kappa}m). 
+\end{equation}
+It should now also be clear to the reader why the CG method is especially applicable to sparse linear systems, since in those cases $m < P^2$.
 
-The computational complexity of CG is $\mathbb{O}(\sqrt{\kappa}m)$, where $\kappa$ is the generalized specctral condition number of the matrix $\bA$ and $m$ is the number of
-non-zero entries in the matrix $\bA$. The mathematical
-definition of $\kappa(\bA)$ is
+\subsection{Preconditioning}
+Preconditioning is a technique which allows us to improve the condition number of a matrix. Let $\bM$ be a positive-definite Hermitian matrix.
+We can now indirectly solve for $\bx$ by using 
 \begin{equation}
-\kappa(\bA) = \frac{\lambda_{\textrm{max}}}{\lambda^{+}_{\textrm{min}}}, 
+\label{eq:prec}
+\bM^{-1}\bA\bx = \bM^{-1}\bb.
 \end{equation}
-where $\lambda_{\textrm{max}}$ denotes the largest eigenvalue of $\bA$ and $\lambda^{+}_{\textrm{min}}$ presents the smallest positive eigenvalue. 
-Interpreting the computational complexity equation further.
-Note from Algorithm 1 that the CG method has one major outer loop.
-In the worst possible case, the amount of iterations needed to converge is
-proportional to $\sqrt{\kappa}$. Also notice that the most expensive 
-operation that is conducted within each loop is the vector-matrix product, which has complexity $\mathbb{O}(m)$, where $m$ is the number of non-zero entries in the matrix.
-The computational complexity of the CG method is therefore 
+If $\bM$ is a good preconditioner then $\kappa(\bM^{-1}\bA) \ll \kappa(\bA)$, i.e. a good preconditioner lowers the condition number of a matrix which in turn increases the convergence speed 
+of CG (see Eq.~\eqref{eq:cg_bound}).
+The best possible choice of $\bM$ would therefore be $\bA$, which would lead to $\kappa$ (as the largest and smallest eigenvalue of the Identity matrix is equal to one) becoming unity.
+If $\kappa$ is one then Eq.~\eqref{eq:cg_bound} simplifies and becomes $\mathbb{O}(m)$.
+However, if we knew the inverse of $\bA$ then obtaining $\bx$ would be trivial and there would be no need for the CG method in the first place.
+In short, a good preconditioner $\bM$ must approximate $\bA$ fairly accurately. However, if it is not much less expensive to 
+compute the inverse of $\bM$ than that of $\bA$ we actually gain nothing (see Eq.~\eqref{eq:prec}).
+
+If the diagonal entries of $\bA$ is much more significant than its off-diagonal entries then the 
+Jacobian preconditioner is a good choice of $\bM$. The Jacobian preconditioner is computed as follows:
 \begin{equation}
-\mathbb{O}(\sqrt{\kappa}m) 
+\bM = \bA\odot\bI, 
 \end{equation}
+where $\bI$ is the identity matrix. Since the Jacobian preconditioner is a diagonal matrix its inverse can be obtained trivially.
 
-It should now be clear to
-the reader why the CG method is espicially applicable to sparse linear systems, since in those cases $m < N^2$. 
-We can improve the computational complexity of the CG method by preconditioning. 
-[Preconditioning is a technique for improving the condition number of a matrix. Suppose that ' is a
-symmetric, positive-definite matrix that approximates, but is easier to invert. We can solve indirectly by solving] - RW directly coppied from source
+\subsection{Normal Equation}
+Lets assume that instead of Eq.~\eqref{eq:linear_system} we have a more general linear system, i.e.
 \begin{equation}
-\bM^{-1}\bA\bx = \bM^{-1}\bb
+ \bb = \bB\bx,
 \end{equation}
-
-The best precondtioner would be the preconditioner which makes $\kappa$ unity, i.e. completly
-eliminating it from the computational complexity equation.
-The only choice of $\bM$ which is capable of best way to do this is to chooose as preconditioner the matrix $(\bA)^{-1}$, which is what we would like to determine in the frist place.
-If the diagonal entries of $\bA$ are more significant than its off-diagonal entries, then the Jacobian preconditioner is a good preconditioner to use. 
-State better that the choice of preconditioner, the value of kappa and what the effect of this would be on the computational complexity of the method.
-
-In general, when $\bA$ has dimensions $m\times n$, with $m > n$ we can use the normal equation 
+where $\bb\in\mathbb{C}^K$, $\bx\in\mathbb{C}^P$  and $\bB\in\mathbb{C}^{K \times P}$, with $K > P$. We can use the normal equation 
 \begin{equation}
 \label{eq:normal_equation}
-\bA^H\bA\bx = \bA^H\bb 
+\bB^H\bB\bx = \bB^H\bb 
 \end{equation}
@@
 \label{eq:norm_ls}
-\|\bb-\bA\bx\|. 
+\|\bb-\bB\bx\|_F^2. 
 \end{equation}
-If $\bA^H\bA$ is invertiable then the solution that minimizes Eq.~\eqref{eq:norm_ls} is unique. In contrast, if $\bA^H\bA$ is singular there are an infinite amount of solutions
-that minimizes Eq.~\eqref{eq:norm_ls}. Moreover, we can use the CG method to solve Eq.~\eqref{eq:normal_equation}, since $\bA^H\bA$ is Hermitian by construction 
+There are a few important observations that we can make while inspecting Eq.~\eqref{eq:normal_equation}:
+\begin{enumerate}
+\item $\bB^H\bB$ is an Hermitian positive semi-definite square matrix by construction.
+\item $\bB^H\bb$ is in the column range of $\bB^H\bB$; this follows trivially from 
+\begin{equation}
+\bB^H\bb \in \mathcal{R}(\bB^H) \implies \bB^H\bb \in \mathcal{R}(\bB^H\bB).   
+\end{equation}
+\end{enumerate}
+
+Recall that we mentioned that in general we can only solve Eq.~\eqref{eq:linear_system} with CG if $\bA$ is an Hermitian positive-definite square matrix. Recently,
+\citet{Lu2015} showed that the CG method has an even broader applicability, it can be used to solve Eq.~\eqref{eq:linear_system} even if $\bA$ is an Hermitian positive semi-definite
+square Hermitian matrix as long as $\bb$ is in the column range of $\bA$. Combining the result from Lu with the aformentioned two observations we made in regards to Eq.~\eqref{}
+
+It now trivially follows \citet{Lu2015} from follows that we can apply the CG method to solve any normal equation. We have already
+
+%It is a well known fact that the normal equation has at least one solution.
+If $\bB^H\bB$ is invertible then the solution that minimizes Eq.~\eqref{eq:norm_ls} is unique.
+In contrast, if $\bB^H\bB$ is singular there are an infinite amount of solutions
+%that minimizes Eq.~\eqref{eq:norm_ls}. 
+%Interestingly enough, we can use the CG method to solve Eq.~\eqref{eq:normal_equation} in either one of the previously mentioned scenarios.
+
+Moreover, we can use the CG method to solve Eq.~\eqref{eq:normal_equation}, since $\bA^H\bA$ is Hermitian by construction 
 and $\bA^H\bb$ is in the column range of $\bA^H\bA$.

### 2016-10-14 paper.tex
polished off theCombining the result from Lu with the aformentioned two observations we made in regards to Eq.~\eqref{}
It now trivially follows \citet{Lu2015} from follows that we can apply the CG method to solve any normal equation. We have already

%It is a well known fact that the normal equation has at least one solution. If $\bB^H\bB$ is invertible then the solution that minimizes Eq.~\eqref{eq:norm_ls} is unique. In contrast, if $\bB^H\bB$ is singular there are an infinite amount of solutions
%that minimizes Eq.~\eqref{eq:norm_ls}.
%Interestingly enough, we can use the CG method to solv

@@
 \citep{Sluis1986}. As can be seen from Algorithm \ref{algo:CG}, the CG method makes use of a major outer loop (line 6 of Algorithm \ref{algo:CG}). It turns out that at worst 
-the number of major iterations that CG requires to converge\footnote{If the $\bA$-norm is used to calculate the error.} is proportional to $\sqrt{\kappa}$, where $\kappa$ is the specctral condition number of the matrix $\bA$.
+the number of major iterations that CG requires to converge\footnote{If the $\bA$-norm is used to calculate the error.} is proportional to $\sqrt{\kappa}$, where $\kappa$ is the spectral condition number of the matrix $\bA$.
 The mathematical definition of $\kappa(\bA)$ is
@@
 \begin{enumerate}
-\item $\bB^H\bB$ is an Hermitian positive semi-definite square matrix by construction.
+\item $\bB^H\bB$ is a square positive semi-definite Hermitian matrix by construction.
 \item $\bB^H\bb$ is in the column range of $\bB^H\bB$; this follows trivially from 
@@
 
-Recall that we mentioned that in general we can only solve Eq.~\eqref{eq:linear_system} with CG if $\bA$ is an Hermitian positive-definite square matrix. Recently,
+Recall that we mentioned that in general we can only solve Eq.~\eqref{eq:linear_system} with CG if $\bA$ is a square positive-definite Hermitian matrix. Recently,
 \citet{Lu2015} showed that the CG method has an even broader applicability, it can be used to solve Eq.~\eqref{eq:linear_system} even if $\bA$ is an Hermitian positive semi-definite
-square Hermitian matrix as long as $\bb$ is in the column range of $\bA$. Combining the result from Lu with the aformentioned two observations we made in regards to Eq.~\eqref{}
+square matrix as long as $\bb$ is in the column range of $\bA$.
 
-It now trivially follows \citet{Lu2015} from follows that we can apply the CG method to solve any normal equation. We have already
+In light of \citet{Lu2015}, the two observations in the numbered list we presented earlier in this section imply that Eq.~\eqref{eq:normal_equation}
+can be solved by employing the CG method. The solution returned by CG is however only unique if $\bB^H\bB$ is positive-definite (invertible). The solution
+obtained by CG, whether it be unique or not, does however minimize Eq.~\eqref{eq:norm_ls} in a least squares sense.
 
-%It is a well known fact that the normal equation has at least one solution.
-If $\bB^H\bB$ is invertible then the solution that minimizes Eq.~\eqref{eq:norm_ls} is unique.
-In contrast, if $\bB^H\bB$ is singular there are an infinite amount of solutions
-%that minimizes Eq.~\eqref{eq:norm_ls}. 
-%Interestingly enough, we can use the CG method to solve Eq.~\eqref{eq:normal_equation} in either one of the previously mentioned scenarios.
-
-Moreover, we can use the CG method to solve Eq.~\eqref{eq:normal_equation}, since $\bA^H\bA$ is Hermitian by construction 
-and $\bA^H\bb$ is in the column range of $\bA^H\bA$.
-\begin{equation}
-\bA^H\bb \in \mathcal{R}(\bA^H) \implies \bA^H\bb \in \mathcal{R}(\bA^H\bA).   
-\end{equation}
+Note that if $\bA$ is positive semi-definite its smallest eigenvalue is zero, which implies that Eq.~\eqref{eq:kappa} is not defined. \citet{Lu2015} show that when $\bA$
+is a square positive semi-definite Hermitian matrix we do not use the spectral condition number to bound the convergence complexity of CG, but the general spectral condition
+number instead. The general spectral condition number is very similar to the spectral condition number, the only difference being that in Eq.~\eqref{eq:kappa} we replace  
+$\lambda_{\textrm{min}}$ with $\lambda_{\textrm{min}}^+$, which denotes the smallest positive eigenvalue.
 

### 2016-10-14 paper.tex
cg

@@
 
-Recall that we mentioned that in general we can only solve Eq.~\eqref{eq:linear_system} with CG if $\bA$ is a square positive-definite Hermitian matrix. Recently,
-\citet{Lu2015} showed that the CG method has an even broader applicability, it can be used to solve Eq.~\eqref{eq:linear_system} even if $\bA$ is an Hermitian positive semi-definite
-square matrix as long as $\bb$ is in the column range of $\bA$.
+Recall that we mentioned that in general we can only solve Eq.~\eqref{eq:linear_system} with CG if $\bA$ is a square positive-definite Hermitian matrix.
+The CG method actually has an even broader applicability, it can be used to solve Eq.~\eqref{eq:linear_system} even if $\bA$ is an Hermitian positive semi-definite
+square matrix as long as $\bb$ is in the column range of $\bA$ \citep{Lu2015}.
 
@@
 is a square positive semi-definite Hermitian matrix we do not use the spectral condition number to bound the convergence complexity of CG, but the general spectral condition
-number instead. The general spectral condition number is very similar to the spectral condition number, the only difference being that in Eq.~\eqref{eq:kappa} we replace  
+number instead. The general spectral condition number is very similar to the spectral condition number; the only difference being that in Eq.~\eqref{eq:kappa} we replace  
 $\lambda_{\textrm{min}}$ with $\lambda_{\textrm{min}}^+$, which denotes the smallest positive eigenvalue.

### 2016-10-17 paper.tex
started to improve least squares section

@@
 \section{Least Squares}
-Redundant calibration can be achieved by solving the following optimization problem
+The method of nonlinear least squares is tanatamount to solving the following optimization problem  
 \begin{equation}
 \label{eq:red_least_squares}
-\min_{\bz} \|\br\| = \min_{\bz} \|\bd - \bv(\bz)\|, 
+\min_{\bz} \Lambda(\bz) = \min_{\bz} \|\br\|_F^2 = \min_{\bz} \|\bd - \bv(\bz)\|_F^2, 
 \end{equation}
-where $\bz = [\bg,\by]^T$. Eq.~\eqref{eq:red_least_squares} is a nonlinear least squares problem.
+where $\Lambda$ is known as the objective function, $\br$ is the residual vector, $\bd$ is the observed data vector, $\bv$ is the model data vector and $\bz$ is the solution which minimizes Eq.~\eqref{eq:red_least_squares}.
+The main objective of least squares is to find the model parameters which minimizes the squared difference between the observed data and the model. Moreover, $\|*\|_F$ denotes the 
+Frobenius norm.
 
-The most standard way of solving a general least squares problem is to use a gradient type of minimization algorithm whose main purpose is to find the model parameters that 
-minimizes the difference between the observed and model data.
-In this paper we will be focusing on the following gradient minimization algorithms: Gauss-Newton and Levenberg-Marquardt (cite LM here).
-For the redundant calibration problem in Eq.~\eqref{eq:red_least_squares} $\bd$, which is the corrupted observed visibilities, is the observed data,
-while $\bv$, which is the predicted visibilities, represents the model data.\footnote{In this paper we use the same symbol to denote predicted visibilities and true observed visibilities as it
-is trivial to distinguish which one is implied from the context.}
+The most standard way of solving Eq.~\eqref{eq:red_least_squares} is to use a gradient type of minimization algorithm.
+In this paper we will be focusing on the following gradient minimization algorithms: Gauss-Newton and Levenberg-Marquardt \citep{Levenberg1944,Marquardt1963}.
 
@@
 
-Recently Kreutz-Delgado and Sorber proposed an alternative strategy, which involves the use of Wirtinger calculus (Wirtinger 1927). The Wirtinger derivatives 
+Recently, \citet{Sorber2012} proposed an alternative strategy, which involves the use of Wirtinger calculus \citep{Wirtinger1927}. The Wirtinger derivatives 
 are defined to be 
@@
 
+\subsection{Redundant Calibration}
 \citet{Smirnov2015} recently proposed to use this Wirtinger framework instead of the classic real and imaginary approach to perform normal calibration. We will now show
-that the Wirtinger framework can also be applied to redundant calibration. Analogous to \cite{Smirnov2015}, we recast Eq.~\eqref{eq:red_least_squares} to 
+that the Wirtinger framework can also be applied to redundant calibration.
+
+For the redundant calibration problem in Eq.~\eqref{eq:red_least_squares} $\bd$, which is the corrupted observed visibilities, is the observed data,
+while $\bv$, which is the predicted visibilities, represents the model data.\footnote{In this paper we use the same symbol to denote predicted visibilities and true observed visibilities as it
+is trivial to distinguish which one is implied from the context.}
+
+ Analogous to \cite{Smirnov2015}, we recast Eq.~\eqref{eq:red_least_squares} to 
 \begin{equation}

### 2016-10-17 paper.tex
polished lq section

@@
 \newcommand{\bM}{\bmath{M}}
+\newcommand{\bO}{\bmath{O}}
+\newcommand{\bP}{\bmath{P}}
+\newcommand{\bQ}{\bmath{Q}}
 \newcommand{\bI}{\bmath{I}}
@@
 \newcommand{\bp}{\bmath{p}}
+\newcommand{\bmJ}{\bmath{\mathcal{J}}}
 \newcommand{\bzero}{\bmath{0}}
@@
 \section{Least Squares}
-The method of nonlinear least squares is tanatamount to solving the following optimization problem  
+The method of nonlinear least squares is tanatamount to solving the following optimization problem:  
 \begin{equation}
-\label{eq:red_least_squares}
-\min_{\bz} \Lambda(\bz) = \min_{\bz} \|\br\|_F^2 = \min_{\bz} \|\bd - \bv(\bz)\|_F^2, 
+\label{eq:least_squares}
+\min_{\bzeta} \Lambda(\bzeta) = \min_{\bzeta} \|\brho\|_F^2 = \min_{\bzeta} \|\bdelta - \bnu(\bzeta)\|_F^2, 
 \end{equation}
-where $\Lambda$ is known as the objective function, $\br$ is the residual vector, $\bd$ is the observed data vector, $\bv$ is the model data vector and $\bz$ is the solution which minimizes Eq.~\eqref{eq:red_least_squares}.
-The main objective of least squares is to find the model parameters which minimizes the squared difference between the observed data and the model. Moreover, $\|*\|_F$ denotes the 
+where $\Lambda$ is known as the objective function, $\brho$ is the residual vector, $\bdelta$ is the observed data vector, $\bnu$ is the model data vector and $\bzeta$ is the solution which minimizes Eq.~\eqref{eq:least_squares}.
+The main objective of least squares is to find the model parameters which minimizes the squared difference between the observed data and the model predicted data. Moreover, $\|*\|_F$ denotes the 
 Frobenius norm.
 
-The most standard way of solving Eq.~\eqref{eq:red_least_squares} is to use a gradient type of minimization algorithm.
+The most standard way of solving Eq.~\eqref{eq:least_squares} is to use a gradient type of minimization algorithm.
 In this paper we will be focusing on the following gradient minimization algorithms: Gauss-Newton and Levenberg-Marquardt \citep{Levenberg1944,Marquardt1963}.
 
-The gradient based minimization algorithms that are generally used to solve least squares problems generally require the model $\bv$ to be differentiable
-towards each model parameter.
-When the least squares problem is complex, which is the case for Eq.~\eqref{eq:red_least_squares}, it becomes less straightforward to apply these gradient based minimization methods.
-The reason being,
+The gradient based minimization algorithms that are generally used to solve least squares problems generally require the model $\bnu$ to be differentiable
+towards each model parameter. When the least squares problem is complex it becomes less straightforward to apply these gradient based minimization methods. The reason being,
 many complex functions are not analytic (no Taylor expansion around a fiducial point exists) and therefore not differentiable if the classic notion of differentiation is used. For instance if the 
@@
 
-The only way to circumvent the differentiability conundrum that plaques complex least square problems is to recast our complex optimization problem in such a way that $\bv$ becomes analytic in its argument.
+Redundant calibration is achieved by solving the following optimization problem:
+
+The only way to circumvent the differentiability conundrum that plaques complex least square problems is to recast our complex optimization problem in such a way that $\bnu$ becomes analytic in its argument.
 The standard way of achieving this is to divide the complex optimization problem into its real and imaginary parts and then to solve for the real and imaginary parts of the parameters separately.
@@
 \begin{align}
+\label{eq:wir}
 \frac{\partial}{\partial z} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} -  i \frac{\partial}{\partial y} \right ),&\frac{\partial}{\partial \conj{z}} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} +  i \frac{\partial}{\partial y} \right ).
@@
 \subsection{Redundant Calibration}
-\citet{Smirnov2015} recently proposed to use this Wirtinger framework instead of the classic real and imaginary approach to perform normal calibration. We will now show
+\citet{Smirnov2015} recently proposed to use this Wirtinger framework instead of the classic real and imaginary approach to perform normal skymodel-based calibration. We will now show
 that the Wirtinger framework can also be applied to redundant calibration.
 
-For the redundant calibration problem in Eq.~\eqref{eq:red_least_squares} $\bd$, which is the corrupted observed visibilities, is the observed data,
-while $\bv$, which is the predicted visibilities, represents the model data.\footnote{In this paper we use the same symbol to denote predicted visibilities and true observed visibilities as it
-is trivial to distinguish which one is implied from the context.}
+We can perform redundant calibration by solving the following optimization problem: 
+\begin{equation}
+\label{eq:least_squares_red}
+\min_{\bz} \Lambda(\bz) = \min_{\bz} \|\br\|_F^2 = \min_{\bz} \|\bd - \bv(\bz)\|_F^2, 
+\end{equation}
+where $\bz = [\bg,\by]^T$.
+The other vectors in Eq.~\eqref{eq:least_squares_red} were defined in Section~\ref{} \footnote{In this paper we use the same symbol to denote corrupted predicted visibilities and true observed corrupted visibilities as it
+is trivial to distinguish which one is implied from the context.
+Using the same symbol for these two quantities improves the readability of the paper.}.
+Note the similarity between Eq.~\eqref{eq:least_squares} and Eq.~\eqref{eq:least_squares_red},
+i.e. redundant calibration can be cast as a non-linear least squares problem.
 
- Analogous to \cite{Smirnov2015}, we recast Eq.~\eqref{eq:red_least_squares} to 
+Analogous to \cite{Smirnov2015}, we now recast Eq.~\eqref{eq:least_squares_red} to 
 \begin{equation}
-\label{eq:red_least_squares_complex}
+\label{eq:least_squares_complex}
 \min_{\breve{\bz}} \|\breve{\br}\| = \min_{\breve{\bz}} \|\breve{\bd} - \breve{\bv}(\breve{\bz})\|, 
@@
 where $\breve{\bz} = [\bz,\conj{\bz}]^T$, $\breve{\bd} = [\bd,\conj{\bd}]^T$, $\breve{\bv} = [\bv,\conj{\bv}]$.
-Eq.~\eqref{eq:red_least_squares_complex} is the augmented counterpart of Eq.~\eqref{eq:red_least_squares}.
-If we use the Wirtinger derivative instead of the classic notion of a derivative to define the gradient operator then 
-$\bv$ becomes analytic in $\bz$ and $\conj{\bz}$ as a whole. The fact that $\bv$ is now analytic in its argument has the immediate consequence that it becomes possible to use the Gauss-Newton
-and the Levenberg-Marquardt algorithm to solve Eq.~\eqref{eq:red_least_squares_complex} \citep{Smirnov2015}.
-
-STILL NEEDS LOADS OF WORK. REFERENCES.
+Eq.~\eqref{eq:least_squares_complex} is the augmented counterpart of Eq.~\eqref{eq:least_squares_red}.
+If we use the Wirtinger derivatives, Eq.~\eqref{eq:wir}, instead of the classic notion of a derivative to define the gradient operator then 
+$\bv$ becomes analytic in $\bz$ and $\conj{\bz}$ as a whole \citep{Smirnov2015}.
+The fact that $\bv$ is now analytic in its argument has the immediate consequence: it becomes possible to apply complex GN
+and complex LM to Eq.~\eqref{eq:least_squares_complex}.
 
 \subsection{Complex GN and LM}
-In this section we discuss how we can use the GN and LM algorithms to solve Eq.~\eqref{eq:red_least_squares_complex} (i.e. estimate $\breve{\bz}$). The complex Jacobian is defined to be
+In this section we discuss how we can use the GN and LM algorithms to solve Eq.~\eqref{eq:least_squares_complex} (i.e. estimate $\breve{\bz}$). The complex Jacobian is defined to be
 \begin{equation}
@@
 \bJ = \begin{bmatrix}
-       \bJ & \bJ^*\\
-       \conj{\bJ}^* & \conj{\bJ} 
+       \bmJ & \bmJ^*\\
+       \conj{\bmJ}^* & \conj{\bmJ} 
       \end{bmatrix},
@@
 \begin{align}
-\bJ &= \frac{\partial \bv}{\partial \bz}, & \bJ^* &= \frac{\partial \bv}{\partial \conj{\bz}}. 
+\bmJ &= \frac{\partial \bv}{\partial \bz}, & \bmJ^* &= \frac{\partial \bv}{\partial \conj{\bz}}. 
 \end{align}
-In Eq.~\eqref{eq:Jacobian}, $\frac{\partial}{\partial \bz}$ and $\frac{\partial}{\partial \conj{\bz}}$ respectively denote the cogradient operator and the conjugate cogradient operator (cite Sorber).
+In Eq.~\eqref{eq:Jacobian}, $\frac{\partial}{\partial \bz}$ and $\frac{\partial}{\partial \conj{\bz}}$ respectively denote the cogradient operator and the conjugate cogradient operator \citep{Sorber2012}.
 The GN update step is now defined as:
@@
 where $\bD=\bI\odot\bJ^H\bJ$.
-NEED TO MENTION THE SILENT INDEX $k$ AND HOW THE UPDATE RULE IS ITERATIVE. ALSO MENTION THAT $\lambda$ IMPROVES THE CONVERGENCE PROPERTIES. WE WILL REFER TO $\bJ^H\bJ$ AS THE 
-HESSIAN.
+
+We use Eq.~\eqref{eq:GN_update} or Eq.~\eqref{eq:LM_update} to iteratively update our 
+parameter vector as follows:
+\begin{equation}
+\label{eq:update}
+\breve{\bz}_{k+1} \gets \breve{\bz}_{k} + \Delta \breve{\bz}_{k}. 
+\end{equation}
+We update our parameter vector by using Eq.~\eqref{eq:update} until convergence is reached. To improve readability we do not explicitly denote the iteration subscript $k$ in 
+Eq.~\eqref{eq:GN_update} and Eq.~\eqref{eq:LM_update}.
 
 \subsection{Redundant Jacobian and Hessian}
-If we apply the definition in Eq.~\eqref{eq:Jacobian} to Eq.~\eqref{eq:red_least_squares_complex} we obtain the following analytic result:
+If we apply the definition in Eq.~\eqref{eq:Jacobian} to Eq.~\eqref{eq:least_squares_complex} we obtain the following analytic result:
 \begin{equation}
@@
 \bM =\begin{bmatrix}
-      \bM_1 & \bM_2
+      \bO & \bP
      \end{bmatrix},
@@
 \bN = \begin{bmatrix}
-       \bN_1 & \bzero
+       \bQ & \bzero
       \end{bmatrix}.
@@
 \begin{equation}
-\label{eq:red_least_squares_lincal}
+\label{eq:least_squares_lincal}
 \min_{\bmath{\varrho}} \|\breve{\br}\| = \min_{\bmath{\varrho}} \|\breve{\bd} - \breve{\bv}(\bmath{\varrho})\|, 
@@
 where $\bmath{\varrho} = \left [\bmath{\eta},\bmath{\varphi},\widetilde{\bmath{\eta}}, \widetilde{\bmath{\varphi}}\right ]^T$. The Jacobian associated with 
-Eq.~\eqref{eq:red_least_squares_lincal} is equal to the Eq xxx in the LINCAL section.
+Eq.~\eqref{eq:least_squares_lincal} is equal to the Eq xxx in the LINCAL section.
 We have therefore shown that doing one iteration of LINCAL is equivalent to a basic GN update step; the GN update step was just re-derived in the Appendix from first principles.

### 2016-10-17 paper.tex
improved LS section

@@
 \label{eq:vis_definition}
-d_{pq} = g_{p}\conj{g}_{q}y_{pq} + n_{pq},
+d_{pq} = g_{p}\conj{g_q}y_{pq} + n_{pq},
 \end{equation}
@@
 \label{eq:vis_red}
-d_{pq} = g_{p}\conj{g}_{q}y_{\phi_{pq}} + n_{pq}.
+d_{pq} = g_{p}\conj{g_q}y_{\phi_{pq}} + n_{pq}.
 \end{equation}
@@
 \begin{align}
- \left [ \bd \right]_{\alpha_{pq}} &= d_{pq}, & \left [ \bv \right ]_{\alpha_{pq}} &= g_p y_{\phi_{pq}} \conj{g}_q\nonumber\\
+ \left [ \bd \right]_{\alpha_{pq}} &= d_{pq}, & \left [ \bv \right ]_{\alpha_{pq}} &= g_p y_{\phi_{pq}} \conj{g_q}\nonumber\\
  \left [ \bn \right ]_{\alpha_{pq}} &= n_{pq}. &  &\label{eq:vec_definitions}
@@
 \label{eq:update}
-\breve{\bz}_{k+1} \gets \breve{\bz}_{k} + \Delta \breve{\bz}_{k}. 
+\breve{\bz}_{k+1} = \breve{\bz}_{k} + \Delta \breve{\bz}_{k}. 
 \end{equation}
@@
 
-\subsection{Redundant Jacobian and Hessian}
+\subsection{Analytic Expressions}
 If we apply the definition in Eq.~\eqref{eq:Jacobian} to Eq.~\eqref{eq:least_squares_complex} we obtain the following analytic result:
@@
 Moreover,
-\begin{align}
-\left [ \bM_1 \right ]_{\alpha_{pq},j} &= y_{\phi_{pq}}\conj{g}_q\delta_p^j, & \left [ \bM_2 \right ]_{\alpha_{pq},j} &= g_p\conj{g}_q\delta_{\phi_{pq}}^j 
-\end{align}
+\begin{equation}
+\left [ \bO  \right ]_{\alpha_{pq},j} = \begin{cases}
+                                         y_{\phi_{pq}}\conj{g_q} & \textrm{if}~p=j\\
+                                         0  & \textrm{otherwise} 
+                                        \end{cases},
+\end{equation}
+
+\begin{equation}
+\left [ \bP  \right ]_{\alpha_{pq},j} = \begin{cases}
+                                         g_p\conj{g_q} & \textrm{if}~\phi_{pq}=j\\
+                                         0  & \textrm{otherwise} 
+                                        \end{cases}
+\end{equation}
 and
 \begin{equation}
-\left [ \bN_1 \right ]_{\alpha_{pq},j} = g_py_{\phi_{pq}}\delta^j_q. 
+\left [ \bQ  \right ]_{\alpha_{pq},j} = \begin{cases}
+                                         g_py_{\phi_{pq}} & \textrm{if}~q=j\\
+                                         0  & \textrm{otherwise} 
+                                        \end{cases}
 \end{equation}
 
-It is now trivial to compute the Hessian $\bJ^H\bJ$ by using Eq.~\eqref{eq:Jacobian_red}. If we substitute Eq.~\eqref{eq:Jacobian_red} into $\bJ^H\bJ$
+%\begin{align}
+%\left [ \bM_1 \right ]_{\alpha_{pq},j} &= y_{\phi_{pq}}\conj{g_q}\delta_p^j, & \left [ \bM_2 \right ]_{\alpha_{pq},j} &= g_p\conj{g_q}\delta_{\phi_{pq}}^j 
+%\end{align}
+%\begin{equation}
+%\left [ \bN_1 \right ]_{\alpha_{pq},j} = g_py_{\phi_{pq}}\delta^j_q. 
+%\end{equation}
+
+It is now trivial to compute the Hessian $\bH$ by using Eq.~\eqref{eq:Jacobian_red}. If we substitute Eq.~\eqref{eq:Jacobian_red} into $\bJ^H\bJ$
 we obtain 
@@
 \end{equation}
-
 Moreover, $(*)^T$ denotes matrix transposition and
-
 \begin{equation}
@@
 
-HAVE TO MAKE SURE I DEFINE ALL THE MATH SYMBOLS THAT I USE. IF NOT HERE THEN EARLIER....
-
 Furthermore, substituting Eq.~\eqref{eq:Jacobian_red} into $\bJ^H\breve{\br}$ results in
@@
 \bJ^H\breve{\br} = \begin{bmatrix}
-                   \ba & \bb & \conj{\ba} & \conj{\bb}\\
-                   \end{bmatrix}^T,
+                   \ba \\
+                   \bb \\
+                   \conj{\ba}\\
+                   \conj{\bb}
+                   \end{bmatrix},
 \end{equation}
@@
 \end{equation}
+Moreover $\ba$ and $\bb$ are both column vectors.
 
 HAVE TO RECHECK THE ABOVE EQUATIONS DIRECTLY FROM THE CODE.
-
-HAVE TO STATE WHETHER COLUMN OR ROW VECTOR.
+DEFINED ALL OPERATORS AND SYMBOLS?
 
@@
 \end{equation}
-Substituting the first identity of Eq.~\eqref{eq:identities} into Eq.~\eqref{eq:temp_GN_eq} and simplifying we obtain
+Substituting the first identity of Eq.~\eqref{eq:identities} into Eq.~\eqref{eq:temp_GN_eq} and simplifying the result 
 \begin{equation}
@@
 \end{equation}
-It now trivially follows from Eq.~\eqref{eq:one_thrid} that:
+It now trivially follows that:
 \begin{equation}
@@
 \end{equation}
-%STILL NEED TO ADD z k+1 EXPRESSION.
-THE TAKE HOME HERE IS WE CAN CALCULATE THE UPDATE STEP WITHOUT CALCULATING THE RESIDUAL.
+Eq.~\eqref{eq:two_thrids} shows us that in the case of redundant calibration we can calculate the GN update step without calculating the residual. The 
+normal skymodel-based equivalent of Eq.~\eqref{eq:two_thrids} is equal to \citep{Smirnov2015}:
+\begin{equation}
+\label{eq:one_half}
+\breve{\bg}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{1}{2}\breve{\bg}_{k}, 
+\end{equation}
+where $\breve{\bg} = [\bg^T,\conj{\bg}^T]^T$
+
+FIX TRANSPOSE PROBLEM
 

### 2016-10-17 paper.tex
started with LINCAL section

@@
 \begin{equation}
+\label{eq:vis_red_vec}
 \bd = \bv + \bn, 
@@
 \subsection{LINCAL}
-Assume that $\eta_p,\eta_q,\widetilde{\eta}_{\phi_{pq}},\varphi_p,\varphi_q$ and $\widetilde{\varphi}_{\phi_{pq}}$ are real valued variables.
-Substituting $g_p$ with $e^{\eta_p - i \varphi_p}$, $g_q$ with $e^{\eta_q - i \varphi_q}$ and $y_{\phi_{pq}}$ with $e^{\eta_{\widetilde{\phi}_{pq}}- i \widetilde{\varphi}_{\phi_{pq}}}$ in Eq.~\eqref{eq:vis_red} results in
-\begin{equation}
-\label{eq:lincal}
-d_{pq} =  e^{\eta_p - i \varphi_p} e^{\widetilde{\eta}_{\phi_{pq}}- i \widetilde{\varphi}_{\phi_{pq}}} e^{\eta_q + i \varphi_q} + n_{pq}.
-\end{equation}
-
-A possible approach one can use to solve the unknowns in Eq.~\eqref{eq:lincal} is to write down the Taylor expansion of the first term of the right hand side of Eq.~\eqref{eq:lincal} around some 
-initial fiducial guess of $\eta_p,\eta_q,\widetilde{\eta}_{\phi_{pq}},\varphi_p,\varphi_q$ and $\widetilde{\varphi}_{\phi_{pq}}$, which we denote with 
-$\eta_p^0,\eta_q^0,\widetilde{\eta}_{\phi_{pq}}^0,\varphi_p^0,\varphi_q^0$ and $\widetilde{\varphi}_{\phi_{pq}}^0$, and then to solve the perturbations from this 
-initial guess that minimize the difference between the right and left hand side of Eq.~\eqref{eq:lincal}. Once the perturbation values 
-have been computed it is trivial to obtain an estimate of $g_i$, $g_j$ and $y_{\phi_{ij}}$.
-
-If we Taylor expand the first term on the right hand side of Eq.~\eqref{eq:lincal} around a fiducial initial guess we obtain
-\begin{equation}
-\label{eq:taylor_expansion}
-d_{pq} = v_{pq}^0[1+\Delta \eta_p + \Delta \eta_q + i \varphi_p - i \Delta \varphi_q + \Delta \widetilde{\eta}_{\phi_{pq}} + i\Delta \widetilde{\eta}_{\phi_{pq}}] + n_{pq}, 
-\end{equation}
-where
-\begin{equation}
-v_{pq}^0 = e^{\eta_p^0 - i \varphi_p^0} e^{\widetilde{\eta}_{\phi_{pq}}^0 - i \widetilde{\varphi}_{\phi_{pq}}^0} e^{\eta_q^0 + i \varphi_q^0}, 
-\end{equation}
-and $\eta_p^0,\eta_q^0,\widetilde{\eta}_{\phi_{pq}}^0,\varphi_p^0,\varphi_q^0$ and $\widetilde{\varphi}_{\phi_{pq}}^0$ respectively denote the initial guess for $\eta_p,\eta_q,\widetilde{\eta}_{\phi_{pq}},\varphi_p,\varphi_q$ and $\widetilde{\varphi}_{\phi_{pq}}$.
-Eq.~\eqref{eq:taylor_expansion} can be used to construct the following linear system
-\begin{equation}
-\breve{\br} = \bJ\Delta\widehat{\bz}, 
-\end{equation}
-where 
-\begin{equation}
-\left [ \breve{\br} \right ]_{\alpha_{pq}} = d_{pq}-v_{pq}^0, 
-\end{equation}
-\begin{equation}
-\label{eq:jac_lincal}
-\bJ = \begin{bmatrix}
-      \bN & \bM\\
-      \conj{\bN} & \conj{\bM}
-      \end{bmatrix}
-\end{equation}
-and
-\begin{equation}
-\Delta\widehat{\bz} = 
-\begin{bmatrix}
-\Delta \bmath{\eta} \\
-\Delta \bmath{\varphi} \\
-\Delta \widetilde{\bmath{\eta}} \\
-\Delta \widetilde{\bmath{\varphi}},
-\end{bmatrix}
-\end{equation}
-with
-\begin{align}
- \Delta \bmath{\eta} &= [\Delta \eta_1,\Delta \eta_2,\cdots \Delta \eta_N]^T\nonumber\\
- \Delta \bmath{\varphi} &= [\Delta \varphi_1,\Delta \varphi_2,\cdots \Delta \varphi_N]^T\nonumber\\
- \Delta \widetilde{\bmath{\eta}} &= [\Delta \widetilde{\eta_1},\Delta \widetilde{\eta_2},\cdots \Delta \widetilde{\eta_N}]^T\nonumber\\
- \Delta \widetilde{\bmath{\varphi}} &= [\Delta \widetilde{\varphi_1},\Delta \widetilde{\varphi_2},\cdots \Delta \widetilde{\varphi_N}]^T.
-\end{align}
-
-Furthermore, 
-\begin{equation}
-\bN = \begin{bmatrix}
-      \bN_1 & \bN_2 
-      \end{bmatrix},
-\end{equation}
-with 
-\begin{align}
-[\bN_1]_{\alpha_{pq},j} &= \frac{\partial v_{pq}}{\partial \eta_j} {\bigg |}_{\widehat{\bz}=\widehat{\bz}^0}\\
-&= (\delta_{p}^{j}\vee\delta_{q}^{j}) v_{pq}^0,
-\end{align}
-and
-\begin{align}
- [\bN_2]_{\alpha_{pq},j} &= \frac{\partial v_{pq}}{\partial \varphi_j} {\bigg |}_{\widehat{\bz}=\widehat{\bz}^0}\\
- &= (-1)^{\delta_{q}^{j}}i(\delta_{p}^{j}\vee\delta_{q}^{j}) v_{pq}^0. 
-\end{align}
-
-Moreover,
-\begin{equation}
-\bM = \begin{bmatrix}
-      \bM_1 & \bM_2 
-      \end{bmatrix},
-\end{equation}
-with
-\begin{align}
-[\bM_1]_{\alpha_{pq},j} &= \frac{\partial v_{pq}}{\partial \widetilde{\eta}_j} {\bigg |}_{\widehat{\bz}=\widehat{\bz}^0}\\
-&=  \delta_{\phi_{pq}}^{j} v_{pq}^0,
-\end{align}
-and
-\begin{align}
-[\bM_2]_{\alpha_{pq},j} &= \frac{\partial v_{pq}}{\partial \widetilde{\varphi}_j} {\bigg |}_{\widehat{\bz}=\widehat{\bz}^0}\\
-&= i\delta_{\phi_{pq}}^{j} v_{pq}^0,
-\end{align}
-
-STILL NEEDS LOTS OF WORK.
 
@@
 
+\appendix
+
+\section{LINCAL}
+%Assume that $\eta_p,\eta_q,\widetilde{\eta}_{\phi_{pq}},\varphi_p,\varphi_q$ and $\widetilde{\varphi}_{\phi_{pq}}$ are real valued variables.
+We can rewrite Eq.~\eqref{eq:vis_red_vec} to
+\begin{equation}
+\label{eq:lincal_vec}
+\bd = \bv(\bvarrho) + \bn, 
+\end{equation}
+where 
+\begin{equation}
+\left [ \bv(\bvarrho) \right ]_{\alpha_{pq}} = e^{\eta_p - i \varphi_p} e^{\widetilde{\eta}_{\phi_{pq}}- i \widetilde{\varphi}_{\phi_{pq}}} e^{\eta_q + i \varphi_q},
+\end{equation}
+\begin{equation}
+ \bvarrho = 
+\begin{bmatrix}
+\bmath{\eta} \\
+\bvarphi \\
+\widetilde{\bmath{\eta}} \\
+\widetilde{\bvarphi}
+\end{bmatrix},
+\end{equation}
+with
+\begin{align}
+ \bmath{\eta} &= [\eta_1, \eta_2,\cdots \eta_N]^T\nonumber\\
+ \bmath{\varphi} &= [\varphi_1, \varphi_2,\cdots \varphi_N]^T\nonumber\\
+ \widetilde{\bmath{\eta}} &= [\widetilde{\eta_1}, \widetilde{\eta_2},\cdots \widetilde{\eta_L}]^T\nonumber\\
+ \widetilde{\bmath{\varphi}} &= [\widetilde{\varphi_1}, \widetilde{\varphi_2},\cdots \widetilde{\varphi_L}]^T.
+\end{align}
+Moreover, $\bmath{\eta},~\bvarphi,~\widetilde{\bmath{\eta}}$ and $\widetilde{\bvarphi}$ are real valued vectors.
+If we Taylor expand Eq.~\eqref{eq:lincal_vec} we obtain
+\begin{align}
+\bd &= \bv + \Delta \bvarrho \bmJ + \bn\\
+\br &= \Delta \bvarrho \bmJ + \bn,
+\end{align}
+where
+\begin{equation}
+\label{eq:jac_upper_lincal}
+\bmJ = \begin{bmatrix}
+       \bN & \bM 
+       \end{bmatrix}
+\end{equation}
+
+\begin{align}
+ \left [ \bd \right]_{\alpha_{pq}} &= d_{pq}, & \left [ \bv \right ]_{\alpha_{pq}} &= g_p y_{\phi_{pq}} \conj{g_q}\nonumber\\
+ \left [ \bn \right ]_{\alpha_{pq}} &= n_{pq}. &  &\label{eq:lincal_definitions}
+\end{align}
+
+A possible approach one can use to solve the unknowns in Eq.~\eqref{eq:lincal} is to write down the Taylor expansion of the first term of the right hand side of Eq.~\eqref{eq:lincal} around some 
+initial fiducial guess of $\eta_p,\eta_q,\widetilde{\eta}_{\phi_{pq}},\varphi_p,\varphi_q$ and $\widetilde{\varphi}_{\phi_{pq}}$, which we denote with 
+$\eta_p^0,\eta_q^0,\widetilde{\eta}_{\phi_{pq}}^0,\varphi_p^0,\varphi_q^0$ and $\widetilde{\varphi}_{\phi_{pq}}^0$, and then to solve the perturbations from this 
+initial guess that minimize the difference between the right and left hand side of Eq.~\eqref{eq:lincal}. Once the perturbation values 
+have been computed it is trivial to obtain an estimate of $g_i$, $g_j$ and $y_{\phi_{ij}}$.
+
+If we Taylor expand the first term on the right hand side of Eq.~\eqref{eq:lincal} around a fiducial initial guess we obtain
+\begin{equation}
+\label{eq:taylor_expansion}
+d_{pq} = v_{pq}^0[1+\Delta \eta_p + \Delta \eta_q + i \varphi_p - i \Delta \varphi_q + \Delta \widetilde{\eta}_{\phi_{pq}} + i\Delta \widetilde{\eta}_{\phi_{pq}}] + n_{pq}, 
+\end{equation}
+where
+\begin{equation}
+v_{pq}^0 = e^{\eta_p^0 - i \varphi_p^0} e^{\widetilde{\eta}_{\phi_{pq}}^0 - i \widetilde{\varphi}_{\phi_{pq}}^0} e^{\eta_q^0 + i \varphi_q^0}, 
+\end{equation}
+and $\eta_p^0,\eta_q^0,\widetilde{\eta}_{\phi_{pq}}^0,\varphi_p^0,\varphi_q^0$ and $\widetilde{\varphi}_{\phi_{pq}}^0$ respectively denote the initial guess for $\eta_p,\eta_q,\widetilde{\eta}_{\phi_{pq}},\varphi_p,\varphi_q$ and $\widetilde{\varphi}_{\phi_{pq}}$.
+Eq.~\eqref{eq:taylor_expansion} can be used to construct the following linear system
+\begin{equation}
+\breve{\br} = \bJ\Delta\widehat{\bz}, 
+\end{equation}
+where 
+\begin{equation}
+\left [ \breve{\br} \right ]_{\alpha_{pq}} = d_{pq}-v_{pq}^0, 
+\end{equation}
+\begin{equation}
+\label{eq:jac_lincal}
+\bJ = \begin{bmatrix}
+      \bN & \bM\\
+      \conj{\bN} & \conj{\bM}
+      \end{bmatrix}
+\end{equation}
+and
+\begin{equation}
+\Delta\widehat{\bz} = 
+\begin{bmatrix}
+\Delta \bmath{\eta} \\
+\Delta \bmath{\varphi} \\
+\Delta \widetilde{\bmath{\eta}} \\
+\Delta \widetilde{\bmath{\varphi}},
+\end{bmatrix}
+\end{equation}
+with
+\begin{align}
+ \Delta \bmath{\eta} &= [\Delta \eta_1,\Delta \eta_2,\cdots \Delta \eta_N]^T\nonumber\\
+ \Delta \bmath{\varphi} &= [\Delta \varphi_1,\Delta \varphi_2,\cdots \Delta \varphi_N]^T\nonumber\\
+ \Delta \widetilde{\bmath{\eta}} &= [\Delta \widetilde{\eta_1},\Delta \widetilde{\eta_2},\cdots \Delta \widetilde{\eta_N}]^T\nonumber\\
+ \Delta \widetilde{\bmath{\varphi}} &= [\Delta \widetilde{\varphi_1},\Delta \widetilde{\varphi_2},\cdots \Delta \widetilde{\varphi_N}]^T.
+\end{align}
+
+Furthermore, 
+\begin{equation}
+\bN = \begin{bmatrix}
+      \bN_1 & \bN_2 
+      \end{bmatrix},
+\end{equation}
+with 
+\begin{align}
+[\bN_1]_{\alpha_{pq},j} &= \frac{\partial v_{pq}}{\partial \eta_j} {\bigg |}_{\widehat{\bz}=\widehat{\bz}^0}\\
+&= (\delta_{p}^{j}\vee\delta_{q}^{j}) v_{pq}^0,
+\end{align}
+and
+\begin{align}
+ [\bN_2]_{\alpha_{pq},j} &= \frac{\partial v_{pq}}{\partial \varphi_j} {\bigg |}_{\widehat{\bz}=\widehat{\bz}^0}\\
+ &= (-1)^{\delta_{q}^{j}}i(\delta_{p}^{j}\vee\delta_{q}^{j}) v_{pq}^0. 
+\end{align}
+
+Moreover,
+\begin{equation}
+\bM = \begin{bmatrix}
+      \bM_1 & \bM_2 
+      \end{bmatrix},
+\end{equation}
+with
+\begin{align}
+[\bM_1]_{\alpha_{pq},j} &= \frac{\partial v_{pq}}{\partial \widetilde{\eta}_j} {\bigg |}_{\widehat{\bz}=\widehat{\bz}^0}\\
+&=  \delta_{\phi_{pq}}^{j} v_{pq}^0,
+\end{align}
+and
+\begin{align}
+[\bM_2]_{\alpha_{pq},j} &= \frac{\partial v_{pq}}{\partial \widetilde{\varphi}_j} {\bigg |}_{\widehat{\bz}=\widehat{\bz}^0}\\
+&= i\delta_{\phi_{pq}}^{j} v_{pq}^0,
+\end{align}
+
+STILL NEEDS LOTS OF WORK.
+
 \label{lastpage}

### 2016-10-18 paper.tex
lincal

@@
 \newcommand{\bQ}{\bmath{Q}}
+\newcommand{\bR}{\bmath{R}}
 \newcommand{\bI}{\bmath{I}}
@@
 where 
-\begin{equation}
-\left [ \bv(\bvarrho) \right ]_{\alpha_{pq}} = e^{\eta_p - i \varphi_p} e^{\widetilde{\eta}_{\phi_{pq}}- i \widetilde{\varphi}_{\phi_{pq}}} e^{\eta_q + i \varphi_q},
-\end{equation}
+\begin{align}
+\left [ \bv(\bvarrho) \right ]_{\alpha_{pq}} &= e^{\eta_p - i \varphi_p} e^{\widetilde{\eta}_{\phi_{pq}}- i \widetilde{\varphi}_{\phi_{pq}}} e^{\eta_q + i \varphi_q}\\
+&= v_{pq},
+\end{align}
+and
 \begin{equation}
@@
 \widetilde{\bvarphi}
-\end{bmatrix},
-\end{equation}
-with
-\begin{align}
- \bmath{\eta} &= [\eta_1, \eta_2,\cdots \eta_N]^T\nonumber\\
- \bmath{\varphi} &= [\varphi_1, \varphi_2,\cdots \varphi_N]^T\nonumber\\
- \widetilde{\bmath{\eta}} &= [\widetilde{\eta_1}, \widetilde{\eta_2},\cdots \widetilde{\eta_L}]^T\nonumber\\
- \widetilde{\bmath{\varphi}} &= [\widetilde{\varphi_1}, \widetilde{\varphi_2},\cdots \widetilde{\varphi_L}]^T.
-\end{align}
-Moreover, $\bmath{\eta},~\bvarphi,~\widetilde{\bmath{\eta}}$ and $\widetilde{\bvarphi}$ are real valued vectors.
-If we Taylor expand Eq.~\eqref{eq:lincal_vec} we obtain
-\begin{align}
-\bd &= \bv + \Delta \bvarrho \bmJ + \bn\\
-\br &= \Delta \bvarrho \bmJ + \bn,
-\end{align}
-where
+\end{bmatrix},\qquad
+\begin{aligned}
+\bmath{\eta} &= [\eta_1, \eta_2,\cdots \eta_N]^T\\
+ \bmath{\varphi} &= [\varphi_1, \varphi_2,\cdots \varphi_N]^T\\
+ \widetilde{\bmath{\eta}} &= [\widetilde{\eta_1}, \widetilde{\eta_2},\cdots \widetilde{\eta_L}]^T\\
+ \widetilde{\bmath{\varphi}} &= [\widetilde{\varphi_1}, \widetilde{\varphi_2},\cdots \widetilde{\varphi_L}]^T
+\end{aligned}.
+\label{eq:help}
+\end{equation}
+Note that the vectors $\bmath{\eta},~\bvarphi,~\widetilde{\bmath{\eta}}$ and $\widetilde{\bvarphi}$ are real valued.
+
+The first oreder Taylor expansion of Eq.~\eqref{eq:lincal_vec} around the fiducial point $\bvarrho_0$ is equal to
+\begin{equation}
+\label{eq:d_lin}
+\bd = \bv_0 + \bmJ_0 \Delta \bvarrho  + \bn.
+\end{equation}
+Calculating the residual and dropping the reference to a specific fiducial point, we obtain 
+\begin{equation}
+\label{eq:r_lin}
+\br = \bd - \bv = \bmJ \Delta \bvarrho + \bn.
+\end{equation}
+Moreover,
 \begin{equation}
@@
        \bN & \bM 
-       \end{bmatrix}
-\end{equation}
-
+       \end{bmatrix},
+\end{equation}
+where
 \begin{align}
- \left [ \bd \right]_{\alpha_{pq}} &= d_{pq}, & \left [ \bv \right ]_{\alpha_{pq}} &= g_p y_{\phi_{pq}} \conj{g_q}\nonumber\\
- \left [ \bn \right ]_{\alpha_{pq}} &= n_{pq}. &  &\label{eq:lincal_definitions}
+\bN &= \begin{bmatrix} \bO & \bP \end{bmatrix}, & \bM &= \begin{bmatrix} \bQ & \bR \end{bmatrix}, 
 \end{align}
 
-A possible approach one can use to solve the unknowns in Eq.~\eqref{eq:lincal} is to write down the Taylor expansion of the first term of the right hand side of Eq.~\eqref{eq:lincal} around some 
-initial fiducial guess of $\eta_p,\eta_q,\widetilde{\eta}_{\phi_{pq}},\varphi_p,\varphi_q$ and $\widetilde{\varphi}_{\phi_{pq}}$, which we denote with 
-$\eta_p^0,\eta_q^0,\widetilde{\eta}_{\phi_{pq}}^0,\varphi_p^0,\varphi_q^0$ and $\widetilde{\varphi}_{\phi_{pq}}^0$, and then to solve the perturbations from this 
-initial guess that minimize the difference between the right and left hand side of Eq.~\eqref{eq:lincal}. Once the perturbation values 
-have been computed it is trivial to obtain an estimate of $g_i$, $g_j$ and $y_{\phi_{ij}}$.
-
-If we Taylor expand the first term on the right hand side of Eq.~\eqref{eq:lincal} around a fiducial initial guess we obtain
-\begin{equation}
-\label{eq:taylor_expansion}
-d_{pq} = v_{pq}^0[1+\Delta \eta_p + \Delta \eta_q + i \varphi_p - i \Delta \varphi_q + \Delta \widetilde{\eta}_{\phi_{pq}} + i\Delta \widetilde{\eta}_{\phi_{pq}}] + n_{pq}, 
-\end{equation}
-where
-\begin{equation}
-v_{pq}^0 = e^{\eta_p^0 - i \varphi_p^0} e^{\widetilde{\eta}_{\phi_{pq}}^0 - i \widetilde{\varphi}_{\phi_{pq}}^0} e^{\eta_q^0 + i \varphi_q^0}, 
-\end{equation}
-and $\eta_p^0,\eta_q^0,\widetilde{\eta}_{\phi_{pq}}^0,\varphi_p^0,\varphi_q^0$ and $\widetilde{\varphi}_{\phi_{pq}}^0$ respectively denote the initial guess for $\eta_p,\eta_q,\widetilde{\eta}_{\phi_{pq}},\varphi_p,\varphi_q$ and $\widetilde{\varphi}_{\phi_{pq}}$.
-Eq.~\eqref{eq:taylor_expansion} can be used to construct the following linear system
-\begin{equation}
-\breve{\br} = \bJ\Delta\widehat{\bz}, 
-\end{equation}
-where 
-\begin{equation}
-\left [ \breve{\br} \right ]_{\alpha_{pq}} = d_{pq}-v_{pq}^0, 
-\end{equation}
-\begin{equation}
-\label{eq:jac_lincal}
-\bJ = \begin{bmatrix}
-      \bN & \bM\\
-      \conj{\bN} & \conj{\bM}
-      \end{bmatrix}
-\end{equation}
-and
-\begin{equation}
-\Delta\widehat{\bz} = 
+\begin{equation}
+ [\bO]_{\alpha_{pq},j} =  \frac{\partial v_{pq}}{\partial \eta_j} = \begin{cases} 
+    v_{pq} &\textrm{if}~(p=j)~\textrm{or}~(q=j)\\
+    0 & \textrm{otherwise}
+   \end{cases},
+\end{equation}
+
+\begin{equation}
+ [\bP]_{\alpha_{pq},j} =  \frac{\partial v_{pq}}{\partial \varphi_j} = \begin{cases}
+                                                                        i v_{pq} &\textrm{if}~ p=j\\
+                                                                        -i v_{pq} &\textrm{if}~ q=j \\
+                                                                        0 & \textrm{otherwise} 
+                                                                       \end{cases}
+\end{equation}
+
+\begin{equation}
+ [\bQ]_{\alpha_{pq},j} =  \frac{\partial v_{pq}}{\partial \widetilde{\eta}_j} = \begin{cases} 
+    v_{pq} &\textrm{if}~\phi_{pq}=j\\
+    0 & \textrm{otherwise}
+   \end{cases},
+\end{equation}
+
+\begin{equation}
+ [\bR]_{\alpha_{pq},j} =  \frac{\partial v_{pq}}{\partial \widetilde{\varphi}_j} = \begin{cases} 
+    i v_{pq} &\textrm{if}~\phi_{pq}=j\\
+    0 & \textrm{otherwise}
+   \end{cases}.
+\end{equation}
+
+The augmented residual is equal to 
+\begin{equation}
+\breve{\br} = \bJ \Delta \bvarrho + \bn, 
+\end{equation}
+
+where $\breve{\br} = [r,\conj{r}]^T$ and 
+\begin{equation}
+\bJ =
 \begin{bmatrix}
-\Delta \bmath{\eta} \\
-\Delta \bmath{\varphi} \\
-\Delta \widetilde{\bmath{\eta}} \\
-\Delta \widetilde{\bmath{\varphi}},
-\end{bmatrix}
-\end{equation}
-with
-\begin{align}
- \Delta \bmath{\eta} &= [\Delta \eta_1,\Delta \eta_2,\cdots \Delta \eta_N]^T\nonumber\\
- \Delta \bmath{\varphi} &= [\Delta \varphi_1,\Delta \varphi_2,\cdots \Delta \varphi_N]^T\nonumber\\
- \Delta \widetilde{\bmath{\eta}} &= [\Delta \widetilde{\eta_1},\Delta \widetilde{\eta_2},\cdots \Delta \widetilde{\eta_N}]^T\nonumber\\
- \Delta \widetilde{\bmath{\varphi}} &= [\Delta \widetilde{\varphi_1},\Delta \widetilde{\varphi_2},\cdots \Delta \widetilde{\varphi_N}]^T.
-\end{align}
-
-Furthermore, 
-\begin{equation}
-\bN = \begin{bmatrix}
-      \bN_1 & \bN_2 
-      \end{bmatrix},
-\end{equation}
-with 
-\begin{align}
-[\bN_1]_{\alpha_{pq},j} &= \frac{\partial v_{pq}}{\partial \eta_j} {\bigg |}_{\widehat{\bz}=\widehat{\bz}^0}\\
-&= (\delta_{p}^{j}\vee\delta_{q}^{j}) v_{pq}^0,
-\end{align}
-and
-\begin{align}
- [\bN_2]_{\alpha_{pq},j} &= \frac{\partial v_{pq}}{\partial \varphi_j} {\bigg |}_{\widehat{\bz}=\widehat{\bz}^0}\\
- &= (-1)^{\delta_{q}^{j}}i(\delta_{p}^{j}\vee\delta_{q}^{j}) v_{pq}^0. 
-\end{align}
-
-Moreover,
-\begin{equation}
-\bM = \begin{bmatrix}
-      \bM_1 & \bM_2 
-      \end{bmatrix},
-\end{equation}
-with
-\begin{align}
-[\bM_1]_{\alpha_{pq},j} &= \frac{\partial v_{pq}}{\partial \widetilde{\eta}_j} {\bigg |}_{\widehat{\bz}=\widehat{\bz}^0}\\
-&=  \delta_{\phi_{pq}}^{j} v_{pq}^0,
-\end{align}
-and
-\begin{align}
-[\bM_2]_{\alpha_{pq},j} &= \frac{\partial v_{pq}}{\partial \widetilde{\varphi}_j} {\bigg |}_{\widehat{\bz}=\widehat{\bz}^0}\\
-&= i\delta_{\phi_{pq}}^{j} v_{pq}^0,
-\end{align}
+ \bmJ\\
+ \conj{\bmJ}
+\end{bmatrix}.
+\end{equation}
 

### 2016-10-19 paper.tex
LINCAL sort of complete

@@
 \section{LINCAL}
+In this section we derive LINCAL \citep{Liu2010} from first principles.
 %Assume that $\eta_p,\eta_q,\widetilde{\eta}_{\phi_{pq}},\varphi_p,\varphi_q$ and $\widetilde{\varphi}_{\phi_{pq}}$ are real valued variables.
@@
 
-The first oreder Taylor expansion of Eq.~\eqref{eq:lincal_vec} around the fiducial point $\bvarrho_0$ is equal to
+The first order Taylor expansion of Eq.~\eqref{eq:lincal_vec} around the fiducial point $\bvarrho_0$ is equal to
 \begin{equation}
@@
 \end{equation}
-Calculating the residual and dropping the reference to a specific fiducial point, we obtain 
+If we calculate the residual, dropping the reference to a specific fiducial point, we obtain: 
 \begin{equation}
@@
 \end{align}
-
 \begin{equation}
@@
 \end{equation}
-
 \begin{equation}
@@
 \end{equation}
-
-The augmented residual is equal to 
+Augmenting the residual by taking into account its conjugate results in 
 \begin{equation}
@@
 \end{equation}
-
-where $\breve{\br} = [r,\conj{r}]^T$ and 
+where $\breve{\br} = \begin{bmatrix}\br^T & \conj{\br}^T \end{bmatrix}^T$ and 
 \begin{equation}
@@
 \end{equation}
-
-STILL NEEDS LOTS OF WORK.
-
+The solution to 
+\begin{equation}
+ \min_{\Delta \bvarrho} \| \breve{\br} - \bJ \Delta \bvarrho\|_F^2,
+\end{equation}
+is equal to 
+\begin{equation}
+\label{eq:LINCAL_update}
+\Delta \bvarrho = (\bJ^H\bJ)^{-1} \bJ^H\breve{\br}. 
+\end{equation}
+Eq.~\eqref{eq:LINCAL_update} is used to iteratively determine a new estimate of the parameter vector $\bvarrho$ as follows
+\begin{equation}
+\label{eq:new_rho}
+\bvarrho_{k+1} = \bvarrho_k + \Delta \bvarrho_k. 
+\end{equation}
+Note that we do not explicitly denote the $k$ index in Eq.~\eqref{eq:LINCAL_update}. We update our parameter vector until convergence is reached.   
 \label{lastpage}
-
 \end{document}

### 2016-10-19 paper.tex
LOGCAL sort of complete

@@
 \usepackage[noend]{algpseudocode}
+\usepackage{mathrsfs}
 
 \newcommand{\bz}{\bmath{z}}
+\newcommand{\bs}{\bmath{s}}
 \newcommand{\bA}{\bmath{A}}
@@
 
-\subsection{LOGCAL}
-NEED BETTER INTRO\\
-One of the best known algorithms that is currently used to perform redundant calibration involves the use of logarithms. If we ignore the noise term then we can 
-rewrite Eq.~\eqref{eq:vis_red} as
-\begin{eqnarray}
-\ln |d_{pq}| &=& \ln |g_p| + \ln |g_q| + \ln |y_{\phi_{pq}}| \label{eq:logcal_amp}\\
-\angle d_{pq} &=& \angle g_p - \angle g_q + \angle \phi_{pq} \label{eq:logcal_phase}
-\end{eqnarray}
-
-We can easily express Eq.~\eqref{eq:logcal_amp} with the following linear system
-\begin{equation}
-\widetilde{\bd} = \bA\widetilde{\bz}, 
-\end{equation}
-where
-\begin{equation}
-\left [ \widetilde{\bd}\right ]_{\alpha_{pq}} = \ln |d_{pq}|,\nonumber\\ 
-\end{equation}
-\begin{equation}
-\widetilde{\bz} = \begin{bmatrix}
-                  \ln | \bg |\\
-                  \ln | \by |
-                  \end{bmatrix},
-\end{equation}
-with $\bg = [g_1,g_2,\cdots,g_N]^T$ and $\by = [y_1,y_2,\cdots,y_L]^T$.
-
-Furthermore,
-\begin{equation}
-\label{eq:A}
-\bA = \begin{bmatrix}
-       \bA_1 & \bA_2
-      \end{bmatrix}
-\end{equation}
-and
-\begin{equation}
-\left [ \bA_1 \right ]_{\alpha_{pq},j} = \delta^j_p\vee\delta^j_q,  ~~~ \left [ \bA_2 \right ]_{\alpha_{pq},j} = \delta_{\phi_{pq}}^j. 
-\end{equation}
-
-The dimensions of $\bA_1$ and $\bA_2$ are $B\times N$ and $B\times L$.
-Moreover, $\vee$ denotes the logical or operator and
-\begin{equation}
-\delta_y^x  = \begin{cases}
-          1 & \textrm{if}~x=y\\
-          0 & \textrm{otherwise}
-         \end{cases}.
-\end{equation}
-A similar linear system can be constructed for Eq.~\eqref{eq:logcal_phase}.\\
-
-NEED TO ROUND OF HERE STILL
-
-\subsection{LINCAL}
-
 \section{Conjugate Gradient Method}
@@
 \subsection{GN and LINCAL}
-As we mentioned at the start of this section, the Wirtinger framework is not the only way in which we can ensure that $\bv$ is analytic in its argument. We could also 
-have divided the problem in Eq ... into its real and imaginary components and then solver for the real and imaginary parts of the parameters separately. There is a third way
-of dealing with the issue. We can split the problem in Eq into its phase and amplitude components and then solve for the amplitude and phase of the parameters separately.
-This can be achieved by first rewriting Eq. as
-
+As we mentioned in the beginning of this section, the Wirtinger framework is not the only way of ensuring that $\bv$ is analytic in its argument. We could also 
+have divided the problem in Eq.~\eqref{eq:least_squares_red} into its real and imaginary components after-which we solve for the real and imaginary parts of the parameters separately.
+There is a third way
+of dealing with the issue.
+We can split the problem in Eq.~\eqref{eq:least_squares_complex} into its phase and amplitude components and then solve for the amplitude and phase of the parameters separately.
+This can be achieved by first rewriting Eq.~\eqref{eq:least_squares_complex} to
 \begin{equation}
@@
 \end{equation}
-where $\bmath{\varrho} = \left [\bmath{\eta},\bmath{\varphi},\widetilde{\bmath{\eta}}, \widetilde{\bmath{\varphi}}\right ]^T$. The Jacobian associated with 
-Eq.~\eqref{eq:least_squares_lincal} is equal to the Eq xxx in the LINCAL section.
-We have therefore shown that doing one iteration of LINCAL is equivalent to a basic GN update step; the GN update step was just re-derived in the Appendix from first principles.
-An easy why of improving the convergence properties of LINCAL would be to simply add the LM damping factor to the basic LINCAL update step.
+where $\bmath{\varrho}$ is the phase and amplitude parameter vector and is defined in Appendix~\ref{sec:lincal}. The Jacobian associated with 
+Eq.~\eqref{eq:least_squares_lincal} is equal to Eq.~\eqref{eq:Jac_lin}. It should therefore be obvious to the reader that 
+performing one iteration of LINCAL is equivalent to a basic GN update step; the GN update step was just re-derived in Appendix~\ref{sec:lincal} from first principles.
+Therefore, a straightforward way of improving the convergence properties of stock standard LINCAL would be to simply add the LM damping factor to the basic LINCAL update step.
 
@@
 
+\section{LOGCAL}
+\label{sec:logcal}
+We can rewrite Eq.~\eqref{eq:vis_red} to \citep{Liu2010}
+\begin{eqnarray}
+\ln |d_{pq}| &=& \ln |g_p| + \ln |g_q| + \ln |y_{\phi_{pq}}| + s_{pq} \label{eq:logcal_amp}\\
+\angle d_{pq} &=& \angle g_p - \angle g_q + \angle \phi_{pq} + t_{pq} \label{eq:logcal_phase}
+\end{eqnarray}
+where
+\begin{align}
+s_{pq} &= \mathscr{R} \left \{1 + \frac{n_{pq}}{g_p\conj{g_q}y_{\phi_{pq}}} \right \}, & t_{pq} &= \mathscr{I}\left \{1 + \frac{n_{pq}}{g_p\conj{g_q}y_{\phi_{pq}}} \right \}.
+\end{align}
+Moreover, $\mathscr{R}\{*\}$ and $\mathscr{I}\{*\}$ denote the real and imaginary operators.
+
+With the aid of Eq.~\eqref{eq:logcal_amp} we can construct the following linear system:
+\begin{equation}
+\label{eq:log_system}
+\bdelta = \bJ\bzeta + \bs, 
+\end{equation}
+where
+\begin{align}
+\left [ \bdelta \right ]_{\alpha_{pq}} &= \ln |d_{pq}|, & \left [ \bs \right ]_{\alpha_{pq}} = s_{pq} 
+\end{align}
+and
+\begin{equation}
+\left [ \bzeta \right ]_{\alpha_{pq}} = \begin{cases} \ln |g_j| & \textrm{if}~i\leq N \\ \ln |y_{j-N}| & \textrm{otherwise} \end{cases}. 
+\end{equation}
+Furthermore,
+\begin{equation}
+\bJ = 
+\begin{bmatrix}
+\bN & \bM
+\end{bmatrix}
+\end{equation}
+with 
+\begin{equation}
+[\bN]_{\alpha_{pq},j} = \begin{cases}
+       1 & \textrm{if}~(p=j)~\textrm{or}~(q=j)\\
+       0 & \textrm{otherwise}
+      \end{cases}
+\end{equation}
+and
+\begin{equation}
+[\bM]_{\alpha_{pq},j} = \begin{cases}
+       1 & \textrm{if}~\phi_{pq}=j\\
+       0 & \textrm{otherwise}
+      \end{cases}
+\end{equation}
+
+The least squares solution to Eq.\eqref{eq:log_system}, under the incorrect assumption that $s_{pq}$ is normally distributed, is given by
+\begin{equation}
+\bzeta = (\bA^T\bA)^{-1} \bA^T\bdelta. 
+\end{equation}
+How to correctly deal with the fact that $s_{pq}$ is not normally distributed is discussed in \citep{Liu2010}. A similar linear system can be constructed for Eq.~\eqref{eq:logcal_phase}. 
+From the least squares solutions to the linear systems associated with Eq.~\eqref{eq:logcal_amp} and Eq.~\eqref{eq:logcal_phase} we can compute the antenna gains as well as the true visibilities.
+
 \section{LINCAL}
+\label{sec:lincal}
 In this section we derive LINCAL \citep{Liu2010} from first principles.
@@
  \bmath{\varphi} &= [\varphi_1, \varphi_2,\cdots \varphi_N]^T\\
- \widetilde{\bmath{\eta}} &= [\widetilde{\eta_1}, \widetilde{\eta_2},\cdots \widetilde{\eta_L}]^T\\
- \widetilde{\bmath{\varphi}} &= [\widetilde{\varphi_1}, \widetilde{\varphi_2},\cdots \widetilde{\varphi_L}]^T
+ \widetilde{\bmath{\eta}} &= [\widetilde{\eta}_1, \widetilde{\eta}_2,\cdots \widetilde{\eta}_L]^T\\
+ \widetilde{\bmath{\varphi}} &= [\widetilde{\varphi}_1, \widetilde{\varphi}_2,\cdots \widetilde{\varphi}_L]^T
 \end{aligned}.
@@
 \begin{equation}
+\label{eq:Jac_lin}
 \bJ = 

### 2016-10-19 paper.tex
LOGCAL sort of complete
reformatting the start of the paper

@@
 
- \section{Visibilities}
+\section{Preliminaries}
+
+\subsection{Measurement Equation}
 The observed visibility $d_{pq}$ measured by the baseline formed by antenna $p$ and $q$ can be described as
@@
 
-\subsection{Redundant Array Geometry Function}
+\subsection{Indexing Functions}
 If an array is redundant then its redundant baselines sample the exact same visibilites in the $uv$-plane, i.e. if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$. We can
@@
 
-\subsection{Signal-to-noise Ratio}
+\subsection{Redundant Measurement Equation}
 If the array is redundant and the function $\phi_{ij}$ is known we can rewrite Eq.~\eqref{eq:vis_definition} as
@@
 \begin{equation}
-\left [ \bzeta \right ]_{\alpha_{pq}} = \begin{cases} \ln |g_j| & \textrm{if}~i\leq N \\ \ln |y_{j-N}| & \textrm{otherwise} \end{cases}. 
+\left [ \bzeta \right ]_{j} = \begin{cases} \ln |g_j| & \textrm{if}~j\leq N \\ \ln |y_{j-N}| & \textrm{otherwise} \end{cases}. 
 \end{equation}

### 2016-10-19 paper.tex
cleaning up first section

@@
 
-\subsection{Measurement Equation}
-The observed visibility $d_{pq}$ measured by the baseline formed by antenna $p$ and $q$ can be described as
-
+\subsection{Unpolarized Measurement Equation}
+The observed visibility $d_{pq}$ measured by the baseline formed by antenna $p$ and $q$ is described by
 \begin{equation}
@@
 and $n_{pq}$ is the thermal noise component.
+Eq.~\eqref{eq:vis_definition} is known as the unpolarized measurement equation.
 Conjugation is denoted by $\conj{(*)}$.
-During the course of an actual observation the true value of $g_p$, $g_q$ and $y_{pq}$ are unknown and are the quantities which infact have
-to be estimated.
-
-The real and imaginary components of the thermal noise is normally distributed with a mean of zero and a standard deviation that is equal to [NB - are the real and imaginary parts defined with same sigma?]
-\begin{equation}
-\sigma = \frac{\sqrt{2}k_{B}T_{\textrm{sys}}}{A\eta\sqrt{\Delta \nu \tau}}, 
-\end{equation}
-where $k_B$ is Boltzmann's constant, $T_{\textrm{sys}}$ is equal to the system temperature, $A$ is the effective collecting area of each element in the array, $\eta$ is a dimensionless
-efficiency factor, $\Delta \nu$ is the observational bandwidth and $\tau$ is the integration time per visibility.
+The true value of $g_p$, $g_q$ and $y_{pq}$ are of course unknown and is what we try to estimate when we calibrate
+and image.
+
+The real and imaginary components of the thermal noise is normally distributed with a mean of zero and a standard deviation proportional to   
+\begin{equation}
+\sigma \propto \frac{T_{\textrm{sys}}}{\sqrt{\Delta \nu \tau}}, 
+\end{equation}
+where $T_{\textrm{sys}}$ is equal to the system temperature, $\Delta \nu$ is the observational bandwidth and $\tau$ is the integration time per visibility.
 
 \subsection{Indexing Functions}
+In general there are two ways of indexing the visibilities associated with different baselines.
+We can use the composite index $pq$, which is the strategy that is followed in Eq.~\eqref{eq:vis_definition},
+or we can use a single unique baseline index which we assign in numerical order to each composite index; assuming the composite indexes are placed in standard baseline ordering.
+The composite indexes $pq$ are ordered using standard baseline ordering
+
 If an array is redundant then its redundant baselines sample the exact same visibilites in the $uv$-plane, i.e. if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$. We can

### 2016-10-19 paper.tex
improving first sections

@@
 \subsection{Indexing Functions}
-In general there are two ways of indexing the visibilities associated with different baselines.
+In general there are two ways of indexing the true visibilities associated with different baselines.
 We can use the composite index $pq$, which is the strategy that is followed in Eq.~\eqref{eq:vis_definition},
-or we can use a single unique baseline index which we assign in numerical order to each composite index; assuming the composite indexes are placed in standard baseline ordering.
-The composite indexes $pq$ are ordered using standard baseline ordering
-
-If an array is redundant then its redundant baselines sample the exact same visibilites in the $uv$-plane, i.e. if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$. We can
-therefore also use the following useful indexing strategy: we can use singular redundant baseline indexes instead of composite antenna pairs to label the true observed visibilities.
-
-Let $\phi_{ij}:\mathbb{N}^2\rightarrow\mathbb{N}^+$ be the function that maps composite antenna pairs to the unique redundant baseline indexes of an array. We refer to this function 
-as the redundant array geometry function in this paper. This mapping is not symmetric as 
-$\phi_{ij} = 0$ if $i>j$. We can also define the following symmetric variant of $\phi_{ij}$:
+or we can use a single unique baseline index which we assign in numerical order to each composite index; assuming the composite indexes are arranged in standard baseline order.
+The composite indexes $pq$ are in standard baseline order when they are arranged as follow: $12, 13, 1N, 23\cdots$. We use $N$ to denote the number of antennas in the array. Let 
+\begin{equation}
+\alpha_{pq} =
+\begin{cases}
+(q-p) + (p-1)\left (N-\frac{1}{2}p \right ) & \textrm{if}~p<q\\
+%B + (p-q) + (q-1)\left (N-\frac{1}{2}q \right )) & \textrm{if}~p>q\\
+%B+q + \frac{1}{2}(p-1)(p-2) & \textrm{if}~p>q\\
+0 & \textrm{otherwise}
+\end{cases}.
+\end{equation}
+We can use $\alpha_{pq}$ to map composite indexes to unique baseline indexes, i.e. 
+\begin{equation}
+12,13,\cdots,1N,23,\cdots \xrightarrow{\alpha_{pq}} 1,2,3,\cdots,B.
+\end{equation}
+Moreover, $B$ denotes the number of baselines in the array, i.e. $B = \frac{N^2-N}{2}$.
+
+If an array is redundant then its redundant baselines should sample the exact same visibilities in the $uv$-plane, i.e. if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$. The 
+fact that redundant baselines should measure the exact same visibility implies that every baseline does not require a unique baselines index; instead we can use a separate index for each unique redundant baseline.
+Let $\phi_{pq}$ be the function that maps composite antenna pairs to the redundant baseline indexes of an array.
+Note that $\phi_{pq}$ is not symmetric as
+$\phi_{pq} = 0$ if $p>q$. We can also define the following symmetric variant of $\phi_{pq}$:
 \begin{equation}
@@
 \begin{cases}
-\phi_{ij}~\textrm{if}~i\leq j\\
-\phi_{ji}~\textrm{if}~i>j
+\phi_{pq}~\textrm{if}~p \leq q\\
+\phi_{qp}~\textrm{if}~p>q
 \end{cases}.
 \end{equation} 
-It is possible to construct a simple analytic expression for $\zeta_{ij}$ when our array is in an east-west regular configuration, namely $\zeta_{ij} = |j-i|$. 
-It becomes increasingly more difficult to construct analytic expressions of $\zeta_{ij}$ for other more complicated array layouts. The empirically constructed symmetric geometry
+It is possible to construct a simple analytic expression for $\zeta_{pq}$ when our array is in an east-west regular configuration, namely $\zeta_{pq} = |q-p|$. 
+It becomes increasingly more difficult to construct analytic expressions of $\zeta_{pq}$ for other more complicated array layouts. The empirically constructed symmetric geometry
 functions of three different redundant layouts are depicted in Fig.~\ref{fig:geometry_function}. We denote the range of $\zeta_{ij}$ with $\mathcal{R}(\zeta_{ij})$. The maximal element 

### 2016-10-19 software/simulator.py
new simulator

@@ -302,9 +302,10 @@ class sim():
           plt.ylabel("S-N [m]",fontsize=label_size)
           plt.title(title+"-"+str(self.N),fontsize=label_size)
     
-          plt.ylim(-1.1*m,1.1*m)
-          plt.xlim(-1.1*m,1.1*m)
+          plt.ylim(-1*m,1*m)
+          plt.xlim(-1*m,1*m)
           plt.tight_layout()
+          plt.axis("equal")
           plt.show() 
 
       '''

### 2016-10-19 software/simulator.py
new simulator

@@ -301,11 +301,8 @@ class sim():
           plt.xlabel("W-E [m]",fontsize=label_size)
           plt.ylabel("S-N [m]",fontsize=label_size)
           plt.title(title+"-"+str(self.N),fontsize=label_size)
-    
-          plt.ylim(-1*m,1*m)
-          plt.xlim(-1*m,1*m)
-          plt.tight_layout()
           plt.axis("equal")
+          plt.axis([-1.1*m, 1.1*m,-1.1*m, 1.1*m])
           plt.show() 
 
       '''

### 2016-12-05 software/redundant_selfcal.py
fixed bug

@@ -138,7 +138,7 @@ def do_red_cal_experiment(SNR=5,min_order=1,max_order=2,layout="HEX",exp_number=
         
         method = "R_StEFCal"
         if freq_enabled:
-           dir_name = layout+"_"+method+"_"+str(SNR)+"_"+str(e)+"_"+type_exp+"_f
+           dir_name = layout+"_"+method+"_"+str(SNR)+"_"+str(e)+"_"+type_exp+"_f"
         else:
            dir_name = layout+"_"+method+"_"+str(SNR)+"_"+str(e)+"_"+type_exp
     

### 2016-12-05 paper.tex
started working on abstract, still lots to be done

@@
 
-\title[Unknown]{TBD}
+\title[Unknown]{Redundant calibration as a complex optimization problem}
 \author[A. V. Raveendran and A. N. Other]{A. V. Raveendran$^{1}$\thanks{E-mail:
@@
 \begin{abstract}
-TBD
+Traditionally redundant calibration is implemented via a layered algorithmic pipeline, i.e. a basic firstcal is performed, followed by LOGCAL and LINCAL.  We first show
+that LINCAL is effectively the non-linear least-squares Gauss-Newton algorithm.
+An easy way to reduce the number of layers that are currently used would therefore be to employ the Levenberg-Marquardt algorithm instead.
+We further show that instead of splitting our calibration problem into its phase and amplitude components,
+which is the avenue taken by LINCAL, we can treat the complex unknown parameters and their complex conjugates as separate unknowns instead.
+
+In this paper we use complex least-squares to perform redundant calibration. Traditionally redundant calibration is performed by splitting the problem into its 
+phase and amplitude components, after-which these two components are solved for separately. During complex least-squares, we treat the complex
+parameters and their conjugates as independent quantities instead.     
 \end{abstract}

### 2016-12-05 paper.tex
abstract started

@@
 We further show that instead of splitting our calibration problem into its phase and amplitude components,
-which is the avenue taken by LINCAL, we can treat the complex unknown parameters and their complex conjugates as separate unknowns instead.
+which is the avenue taken by LINCAL, we can treat the complex unknown parameters and their complex conjugates as separate unknowns instead. In the latter approach 
+we use Wirtinger derivatives instead of the classical definition of differentiation. Lastly, we compare two previously proposed methods, to increase the speed of redundant 
+calibration, with each other. The two approaches we investigate is the alternating direction implicit method and the preconditioned conjugate gradient method. We specifically use 
+the Wirtinger framework in this paper, instead of the phase amplitude approach, as it is easy to show that the alternating direction implicit method is simply a special case
 

### 2016-12-05 paper.tex
newest version of paper

--- 2016-10-19 paper.tex
@@
 \begin{abstract}
-TBD
+Traditionally redundant calibration is implemented via a hierarchical algorithmic strategy, i.e. a basic firstcal is performed, followed by LOGCAL and LINCAL.  We first show
+that LINCAL is effectively the non-linear least-squares gradient-based Gauss-Newton algorithm.
+Therefore, an easy way to reduce the number of layers used in performing redundant calibration is to employ the Levenberg-Marquardt algorithm instead (which would improve the convergence properties of stock standard LINCAL).
+We further show that instead of splitting our calibration problem into its phase and amplitude components,
+which is the approach taken by LINCAL, we can treat the complex unknown parameters and their complex conjugates as separate unknowns instead. In the latter approach 
+we make use of the gradient-based complex Levenberg-Marquardt algorithm and not complex Gauss-Newton for the same reason stated earlier. 
+Lastly, we compare two previously proposed methods to speedup redundant 
+calibration, namely the alternating direction implicit method and the preconditioned conjugate gradient method. We found that the alternating 
+direction implicit method outperforms the preconditioned conjugate gradient method, however the preconditioned conjugate gradient method is more robust as it is easier to adapt to other calibration use cases.
+The reason we use the complex optimization framework, instead of the phase-amplitude approach, in this paper is: in this framework
+it becomes demonstrate that the alternating direction implicit method is in fact a special case of the complex Levenberg-Marquardt algorithm.
 \end{abstract}

### 2016-12-05 paper.tex
newest version of paper

@@
 We further show that instead of splitting our calibration problem into its phase and amplitude components,
-which is the approach taken by LINCAL, we can treat the complex unknown parameters and their complex conjugates as separate unknowns instead. In the latter approach 
-we make use of the gradient-based complex Levenberg-Marquardt algorithm and not complex Gauss-Newton for the same reason stated earlier. 
+which is the approach taken by LINCAL, we can treat the complex unknown model parameters and their complex conjugates as separate unknowns instead. In the latter framework
+we make use of the gradient-based complex Levenberg-Marquardt algorithm (not its real valued counterpart) to determine our fundamental update step. 
 Lastly, we compare two previously proposed methods to speedup redundant 
 calibration, namely the alternating direction implicit method and the preconditioned conjugate gradient method. We found that the alternating 
-direction implicit method outperforms the preconditioned conjugate gradient method, however the preconditioned conjugate gradient method is more robust as it is easier to adapt to other calibration use cases.
-The reason we use the complex optimization framework, instead of the phase-amplitude approach, in this paper is: in this framework
-it becomes demonstrate that the alternating direction implicit method is in fact a special case of the complex Levenberg-Marquardt algorithm.
+direction implicit method outperforms the preconditioned conjugate gradient method, however the preconditioned conjugate gradient method is more robust as it would be easier to adapt to other calibration use cases.
+The reason we exclusively use the complex optimization framework in this paper, instead of the phase-amplitude approach, is that in our opinion the complex framework makes it much easier to derive an analytical tractable update step.
+More importantly however, when we work in this complex framework it becomes trivial to show that the alternating direction implicit method is in actual fact just a
+special case of the more general complex Levenberg-Marquardt algorithm.
 \end{abstract}

### 2016-12-06 paper.tex
changed layout of paper

@@
 
-\subsection{Unpolarized Measurement Equation}
+\subsection{Scalar Measurement Equation}
 The observed visibility $d_{pq}$ measured by the baseline formed by antenna $p$ and $q$ is described by
@@
 
+NB:: ADD GN AND LM UPDATE RULES HERE
+
 The gradient based minimization algorithms that are generally used to solve least squares problems generally require the model $\bnu$ to be differentiable
@@
 
-\subsection{Redundant Calibration}
+\section{Redundant Calibration}
+
+\subsection{Historical Overview}
+Layered approach, LOGCAL, LINCAL, Stefan's conference paper, Noordam, other ASTRON paper, newest papers on redundant calibration, marthi paper etc...
+
+\subsection{Real-Imag Framework}
+Here we discuss the fact that LINCAL is just GN
+
+OLD COPIED TEXT
+
+As we mentioned in the beginning of this section, the Wirtinger framework is not the only way of ensuring that $\bv$ is analytic in its argument. We could also 
+have divided the problem in Eq.~\eqref{eq:least_squares_red} into its real and imaginary components after-which we solve for the real and imaginary parts of the parameters separately.
+There is a third way
+of dealing with the issue.
+We can split the problem in Eq.~\eqref{eq:least_squares_complex} into its phase and amplitude components and then solve for the amplitude and phase of the parameters separately.
+This can be achieved by first rewriting Eq.~\eqref{eq:least_squares_complex} to
+\begin{equation}
+\label{eq:least_squares_lincal}
+\min_{\bmath{\varrho}} \|\breve{\br}\| = \min_{\bmath{\varrho}} \|\breve{\bd} - \breve{\bv}(\bmath{\varrho})\|, 
+\end{equation}
+where $\bmath{\varrho}$ is the phase and amplitude parameter vector and is defined in Appendix~\ref{sec:lincal}. The Jacobian associated with 
+Eq.~\eqref{eq:least_squares_lincal} is equal to Eq.~\eqref{eq:Jac_lin}. It should therefore be obvious to the reader that 
+performing one iteration of LINCAL is equivalent to a basic GN update step; the GN update step was just re-derived in Appendix~\ref{sec:lincal} from first principles.
+Therefore, a straightforward way of improving the convergence properties of stock standard LINCAL would be to simply add the LM damping factor to the basic LINCAL update step.
+
+\subsection{Complex Framework}
+
 \citet{Smirnov2015} recently proposed to use this Wirtinger framework instead of the classic real and imaginary approach to perform normal skymodel-based calibration. We will now show
@@
 
-\subsection{Complex GN and LM}
+\subsubsection{Complex GN and LM}
 In this section we discuss how we can use the GN and LM algorithms to solve Eq.~\eqref{eq:least_squares_complex} (i.e. estimate $\breve{\bz}$). The complex Jacobian is defined to be
@@
 
-\subsection{Analytic Expressions}
+\subsubsection{Analytic Expressions}
 If we apply the definition in Eq.~\eqref{eq:Jacobian} to Eq.~\eqref{eq:least_squares_complex} we obtain the following analytic result:
@@
 
-\subsection{Simplifying GN}
+\subsubsection{Simplifying GN}
 
@@
 
-\subsection{GN and LINCAL}
-As we mentioned in the beginning of this section, the Wirtinger framework is not the only way of ensuring that $\bv$ is analytic in its argument. We could also 
-have divided the problem in Eq.~\eqref{eq:least_squares_red} into its real and imaginary components after-which we solve for the real and imaginary parts of the parameters separately.
-There is a third way
-of dealing with the issue.
-We can split the problem in Eq.~\eqref{eq:least_squares_complex} into its phase and amplitude components and then solve for the amplitude and phase of the parameters separately.
-This can be achieved by first rewriting Eq.~\eqref{eq:least_squares_complex} to
-\begin{equation}
-\label{eq:least_squares_lincal}
-\min_{\bmath{\varrho}} \|\breve{\br}\| = \min_{\bmath{\varrho}} \|\breve{\bd} - \breve{\bv}(\bmath{\varrho})\|, 
-\end{equation}
-where $\bmath{\varrho}$ is the phase and amplitude parameter vector and is defined in Appendix~\ref{sec:lincal}. The Jacobian associated with 
-Eq.~\eqref{eq:least_squares_lincal} is equal to Eq.~\eqref{eq:Jac_lin}. It should therefore be obvious to the reader that 
-performing one iteration of LINCAL is equivalent to a basic GN update step; the GN update step was just re-derived in Appendix~\ref{sec:lincal} from first principles.
-Therefore, a straightforward way of improving the convergence properties of stock standard LINCAL would be to simply add the LM damping factor to the basic LINCAL update step.
-
-\subsection{Diagonalization and StEFCal}
+\section{Results}
+
+\subsection{Simulations}
+
+\subsection{Alternating Direction Implicit Method}
+
+%\subsection{Diagonalization and StEFCal}
 StEFCal (Statistically Efficient Calibration) is an alternating direction implicit method, in which the measurement equation (cite the equation earlier in paper) is linearized by assuming that the gains are known, but
@@
 of $2\frac{1}{3}$.
+
+NB:: NEED TO ADD SNR PLOTS HERE
+PLUS IMAGE OF HESSIAN FOR THE DISCUSSION
+
+\subsection{Conjugate Gradient Method}
+
+\subsubsection{Spectral Condition Number}
+kappa + itr plots
+\subsubsection{Sparsity}
+sparsity derivation
+sparsity + computational complexity plots
+
+\subsection{Comparison}
+itr numbers plot and graph of $kP = P^{1.7}$.
 

### 2016-12-06 paper.tex
added GN and LM to LEASTSQR section

@@
 \end{equation}
-where $\Lambda$ is known as the objective function, $\brho$ is the residual vector, $\bdelta$ is the observed data vector, $\bnu$ is the model data vector and $\bzeta$ is the solution which minimizes Eq.~\eqref{eq:least_squares}.
+where $\Lambda$ is known as the objective function, $\brho$ is the residual vector, $\bdelta$ is the observed data vector, $\bnu$ is the model data vector and $\bzeta$ is the model parameter vector that minimizes Eq.~\eqref{eq:least_squares}.
+In Eq.~\eqref{eq:least_squares}, we assume that the optimization problem is real valued.
 The main objective of least squares is to find the model parameters which minimizes the squared difference between the observed data and the model predicted data. Moreover, $\|*\|_F$ denotes the 
@@
 In this paper we will be focusing on the following gradient minimization algorithms: Gauss-Newton and Levenberg-Marquardt \citep{Levenberg1944,Marquardt1963}.
-
-NB:: ADD GN AND LM UPDATE RULES HERE
-
-The gradient based minimization algorithms that are generally used to solve least squares problems generally require the model $\bnu$ to be differentiable
+The GN update step is defined as
+\begin{equation}
+ \Delta \bzeta = (\bJ^T\bJ)^{-1} \bJ^T\brho,
+\end{equation}
+while the LM update takes the following form
+\begin{equation}
+ \Delta \bzeta = (\bJ^T\bJ+\lambda_{\textrm{LM}}\bD)^{-1} \bJ^T\brho,
+\end{equation}
+where $\bJ$ is the Jacobian matrix (i.e. the derivative of the model $\bnu$ with respect to the parameter vector $\bzeta$), $\lambda_{\textrm{LM}}$ denotes the 
+damping factor, while $\bD$ is a diagonal matrix whose diagonal entries are equal to the diagonal entries of $\bJ^T\bJ$. These update rules are used in an iterative manner to 
+determine new estimates of $\bzeta$.
+
+The gradient based minimization algorithms that are generally used to solve least squares problems (i.e. GN and LM) generally require the model $\bnu$ to be differentiable
 towards each model parameter. When the least squares problem is complex it becomes less straightforward to apply these gradient based minimization methods. The reason being,
@@
 
-The only way to circumvent the differentiability conundrum that plaques complex least square problems is to recast our complex optimization problem in such a way that $\bnu$ becomes analytic in its argument.
+The only way to circumvent the differentiability conundrum, associated with complex least square problems, is to recast our complex optimization problem in such a way that $\bnu$ becomes analytic in its argument.
 The standard way of achieving this is to divide the complex optimization problem into its real and imaginary parts and then to solve for the real and imaginary parts of the parameters separately.
@@
 \end{align}
+In this alternative approach we treat the complex parameters and their conjugates as separate parameters.
 

### 2016-12-06 paper.tex
added some historical overview to red calibration paper

@@
 \subsection{Historical Overview}
+
+[Redundant calibration has been in use for a long time. In the early 1980's it was already used to produce high-dynamic range images of 3C84.] RW The first 
+implementations of redundant calibration made use of the logarithm function, because when the logarithm is applied to the scalar measurement equation redundant 
+calibration is transformed into a linear problem cite Wieringa,camps and Lui. This approach is also generally referred to as LOGCAL.
+LOGCAL is reviewed in Appendix ??.
+The next improvement was to perform a firt-order taylornomial expansion (which is equivalent to the GN algorithm Kurien 2016) of the measurement equation an then to solve for the
+perturbations of the model-parameters.
+This approach is commonly referred to as LINCAL.
+LINCAL is derived from first principals in Appendix .
+Two ways of speeding up redundant calibration has also been
+proposed. The first technique makes use of the alternating direction implicit method, in which we alternate between solving the gains, their conjugates and 
+the true observed visibilities (wijnholds and marthi). Using the conjugate gradient method to exploit the sparsity inherent in the redundant calibration problem, 
+and in doing so speedup the algorithm itself was proposed in . A proper comparison of between these two approaches have not been done before.
+
 Layered approach, LOGCAL, LINCAL, Stefan's conference paper, Noordam, other ASTRON paper, newest papers on redundant calibration, marthi paper etc...

### 2016-12-06 paper.tex
started adding some historical overview to red calibration paper

@@
 
-[Redundant calibration has been in use for a long time. In the early 1980's it was already used to produce high-dynamic range images of 3C84.] RW The first 
-implementations of redundant calibration made use of the logarithm function, because when the logarithm is applied to the scalar measurement equation redundant 
+[Redundant calibration has been in use for a long time. It was already being used in the 1980's. Noordam used redundant calibration to produce high-dynamic range images of 3C84 with the 
+WSRT array.] RW The first implementations of redundant calibration made use of the logarithm function, because when the logarithm is applied to the scalar measurement equation redundant 
 calibration is transformed into a linear problem cite Wieringa,camps and Lui. This approach is also generally referred to as LOGCAL.
 LOGCAL is reviewed in Appendix ??.
-The next improvement was to perform a firt-order taylornomial expansion (which is equivalent to the GN algorithm Kurien 2016) of the measurement equation an then to solve for the
+The next improvement was to perform a firt-order taylor expansion (which is equivalent to the GN algorithm Kurien 2016) of the measurement equation an then to solve for the
 perturbations of the model-parameters.
@@
 the true observed visibilities (wijnholds and marthi). Using the conjugate gradient method to exploit the sparsity inherent in the redundant calibration problem, 
-and in doing so speedup the algorithm itself was proposed in . A proper comparison of between these two approaches have not been done before.
+and in doing so speedup the algorithm itself was proposed in . A proper comparison between these two approaches has not been done before.
+
+Newly built instruments like LOFAR, ..... has also sparked a renewed interest in redundant calibration. Redundancy is also advantageous for detecting the EoR,
+which explains why some of the previsously mentioned instruments have redundant layouts.
 

### 2016-12-06 paper.tex
added some text to sparsity section

@@
 
-[Redundant calibration has been in use for a long time. It was already being used in the 1980's. Noordam used redundant calibration to produce high-dynamic range images of 3C84 with the 
+[Redundant calibration has been in use for a long time. Noordam used redundant calibration to produce high-dynamic range images of 3C84 with the 
 WSRT array.] RW The first implementations of redundant calibration made use of the logarithm function, because when the logarithm is applied to the scalar measurement equation redundant 
@@
 \subsubsection{Sparsity}
+
+% \begin{enumerate}
+% \item $\boldsymbol{H}$: is a $(4N-2)\times(4N-2)$ or a $P \times P$ matrix. $N$ denotes the number of antennas and $P$ denotes the number of parameters. This matrix
+% contains $6N^2 - 2N - 2$ non-zero entries. It contains $16N^2 - 16N + 4$ entries. We can construct this matrix with $4N^2 -  4N$ or $\frac{1}{4} P^2 -1$ elementary operations.
+% \item $\boldsymbol{A}$: is a $(2N-1)\times(2N-1)$ matrix. This matrix has $N^2 + N - 1$ non-zero entries. We can construct this matrix with $2\frac{1}{2} (N^2 -  N)$ elementary operations.  
+% \item $\boldsymbol{B}$: is a $(2N-1)\times(2N-1)$ matrix. This matrix has $2N^2 - 2N$ non-zero entries. We can construct this matrix with $1\frac{1}{2} (N^2 -  N)$ elementary operations.
+% \item $\boldsymbol{C}$: is a $N\times N$ matrix. This matrix has $N$ non-zero entries. We can construct this matrix with $N^2-N$ elementary operations.
+% \item $\boldsymbol{D}$: is a $N \times (N-1)$ matrix. This matrix has $\frac{1}{2} (N^2 -  N)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations. 
+% \item $\boldsymbol{E}$: is a $(N-1) \times (N-1)$ matrix. This matrix has $(N-1)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.  
+% \item $\boldsymbol{F}$: is a $N \times N$ matrix. This matrix has $N^2 - N$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.
+% \item $\boldsymbol{G}$: is a $N \times (N-1)$ matrix. This matrix has $\frac{1}{2} (N^2 -  N)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.
+% \item $\boldsymbol{0}$: is a $(N-1) \times (N-1)$ all zero matrix. 
+% \end{enumerate}
+
+%We can construct $\boldsymbol{C}-\boldsymbol{G}$ in $O(N^2)$
+
+As alluded to in section ..., the second factor that determines the computational complexity of the conjugate gradient method when used to perform redundant calibration is the amount of non-zero
+entries in the Hessian matrix $\bH$, which is the matrix we are trying to invert. We have also shown in section that the first factor, the spectral condition number of $\bH$, can be reduced 
+to a constant by employing preconditioning and can therefore be completely eliminated from the computational complexity equation. A useful matrix quantity
+which we will use in this section is the sparsity factor $\gamma$ of a matrix, which is the ratio between the amount of zero entries and the total amount of entries in a matrix. We 
+can therefore define $\gamma$ as follows:
+\begin{equation}
+ \gamma = \left (1 - \frac{m}{P^2} \right ) 
+\end{equation}
+It is trivial to determine an analytic expression for the sparsity ratio of $\bH$ when our redundant array is in a regular east-west geometric configuration
+The sparsity and the asymptotic sparsity ratio of $\boldsymbol{H}$ are equal to
+\begin{align}
+\gamma &= \frac{5N^2-7N+3}{8N^2-8N+2} & \gamma_{\infty} &= \lim_{N\rightarrow \infty}\gamma = \frac{5}{8}. 
+\end{align}
+However when our redundant array makes use of a more complicated geometric layout it becomes much more difficult to construct such expressions as $\bH$ itself 
+becomes much harder to express analytically.
+
+In Fig... we plot the sparsity ratio of three different geometric layouts as a function of the number antennas in the array. Discuss the solid lines 
+and how they relate to the two equations from above.
+
+We can now use the sparsity factor to compute the order of the computational cost of inverting $\bH$ as follows:
+\begin{align}
+P^{c} &= (1 - \gamma)P^2\\
+c &= \log_{P}(1 - \gamma) + 2 & c_{\infty} &= \lim_{N\rightarrow \infty} c = 2
+\end{align}
+
+Asymptotically the computational cost of inverting $\bH$ is $O(P^2)$. However, the computational cost converges 
+very slowly to its asymptotic value, and in general is equal to $O(P^{c})$, where $c \sim 1.7$ if a Hexagonal geometric layout is employed.
+
 sparsity derivation

### 2016-12-06 paper.tex
added some text to sparsity section

@@
 
-We can now use the sparsity factor to compute the order of the computational cost of inverting $\bH$ as follows:
-\begin{align}
-P^{c} &= (1 - \gamma)P^2\\
-c &= \log_{P}(1 - \gamma) + 2 & c_{\infty} &= \lim_{N\rightarrow \infty} c = 2
-\end{align}
-
+We can now use the sparsity factor to compute the order of the computational cost of inverting $\bH$ with
+\begin{equation}
+P^{c} = (1 - \gamma)P^2.
+\end{equation}
+Solving for $c$ we obtain
+\begin{align}
+c &= \log_{P}(1 - \gamma) + 2 & c_{\infty} &= \lim_{N\rightarrow \infty} c = 2.
+\end{align}
 Asymptotically the computational cost of inverting $\bH$ is $O(P^2)$. However, the computational cost converges 

### 2016-12-07 software/plot_func.py
added plot function

@@ -418,7 +418,7 @@ def plot_err_itr(SNR=1000,num=4,e_upper1=5,e_upper2=5):
 def mad(x):
     return np.median(np.absolute(x - np.median(x)))
 
-def plot_precentage_error(SNR=10,k_upper1=5,k_upper2=5):
+def plot_precentage_error(SNR=10,k_upper1=0,k_upper2=3,extra_string="_G_OLD_f"):
     
     N = np.array([7,19,37,61,91,108,169,217])
 
@@ -436,7 +436,7 @@ def plot_precentage_error(SNR=10,k_upper1=5,k_upper2=5):
         precentage_stef_dic[str(N[j])]=np.array([],dtype=float)
 
     for k in xrange(k_upper1):
-        PCG_DIR = "./HEX_PCG_"+str(SNR)+"_False_"+str(k)
+        PCG_DIR = "./HEX_PCG_"+str(SNR)+"_False_"+str(k)+extra_string
         PCG_FILE_LIST = np.array([PCG_DIR+"/1_7_9_"+PCG_DIR[2:]+".p",PCG_DIR+"/2_19_30_"+PCG_DIR[2:]+".p",PCG_DIR+"/3_37_63_"+PCG_DIR[2:]+".p",PCG_DIR+"/4_61_108_"+PCG_DIR[2:]+".p",PCG_DIR+"/5_91_165_"+PCG_DIR[2:]+".p",PCG_DIR+"/6_127_234_"+PCG_DIR[2:]+".p",PCG_DIR+"/7_169_315_"+PCG_DIR[2:]+".p",PCG_DIR+"/8_217_408_"+PCG_DIR[2:]+".p"]) 
         
         for i in xrange(len(PCG_FILE_LIST)):
@@ -476,7 +476,7 @@ def plot_precentage_error(SNR=10,k_upper1=5,k_upper2=5):
             file_p.close()
     
     for k in xrange(k_upper2):
-        STEF_DIR = "./HEX_R_StEFCal_"+str(SNR)+"_"+str(k)
+        STEF_DIR = "./HEX_R_StEFCal_"+str(SNR)+"_"+str(k)+extra_string
[…]

### 2016-12-07 software/plot_func.py
if statements to handle corrupted files

@@ -483,6 +483,15 @@ def plot_precentage_error(SNR=10,k_upper1=0,k_upper2=3,extra_string="_G_OLD_f"):
             print "STEF_FILE_LIST = ",STEF_FILE_LIST[i]
             if not os.path.isfile(STEF_FILE_LIST[i]):
                break
+            if STEF_FILE_LIST[i] == "./HEX_R_StEFCal_5_2_G_OLD_f/8_217_408_HEX_R_StEFCal_5_2_G_OLD_f.p":#corrupted file
+               break            
+            if STEF_FILE_LIST[i] == "./HEX_R_StEFCal_3_2_G_OLD_f/7_169_315_HEX_R_StEFCal_3_2_G_OLD_f.p":#corrupted file
+               break
+            if  STEF_FILE_LIST[i] == "./HEX_R_StEFCal_1_2_G_OLD_f/5_91_165_HEX_R_StEFCal_1_2_G_OLD_f.p":#corrupted file
+               break
+            if STEF_FILE_LIST[i] ==  "./HEX_R_StEFCal_-1_1_G_OLD_f/7_169_315_HEX_R_StEFCal_-1_1_G_OLD_f.p":#corrupted file
+            
+
             file_p = open(STEF_FILE_LIST[i], 'rb')
             order = pickle.load(file_p)
             N_v = pickle.load(file_p)
@@ -900,5 +909,11 @@ if __name__ == "__main__":
    #plot_outer_loop(SNR=5)
    #plot_time()
    #plot_sparsity()
-   plot_precentage_error(SNR=1000,k_upper1=0,k_upper2=3)
+   #plot_precentage_error(SNR=1000,k_upper1=0,k_upper2=3)
+   #plot_precentage_error(SNR=20,k_upper1=0,k_upper2=3)
+   #plot_precentage_error(SNR=10,k_upper1=0,k_upper2=3)
+   #plot_precentage_error(SNR=5,k_upper1=0,k_upper2=3)
+   #plot_precentage_error(SNR=3,k_upper1=0,k_upper2=3)
+   plot_precentage_error(SNR=1,k_upper1=0,k_upper2=3)
+   plot_precentage_error(SNR=-1,k_upper1=0,k_upper2=2)
    #plot_err_itr(num=4)

### 2016-12-07 software/plot_func.py
if statements to handle corrupted files

@@ -490,7 +490,7 @@ def plot_precentage_error(SNR=10,k_upper1=0,k_upper2=3,extra_string="_G_OLD_f"):
             if  STEF_FILE_LIST[i] == "./HEX_R_StEFCal_1_2_G_OLD_f/5_91_165_HEX_R_StEFCal_1_2_G_OLD_f.p":#corrupted file
                break
             if STEF_FILE_LIST[i] ==  "./HEX_R_StEFCal_-1_1_G_OLD_f/7_169_315_HEX_R_StEFCal_-1_1_G_OLD_f.p":#corrupted file
-            
+               break            
 
             file_p = open(STEF_FILE_LIST[i], 'rb')
             order = pickle.load(file_p)

### 2016-12-07 software/plot_func.py
newest plot_func.py

@@ -914,6 +914,6 @@ if __name__ == "__main__":
    #plot_precentage_error(SNR=10,k_upper1=0,k_upper2=3)
    #plot_precentage_error(SNR=5,k_upper1=0,k_upper2=3)
    #plot_precentage_error(SNR=3,k_upper1=0,k_upper2=3)
-   plot_precentage_error(SNR=1,k_upper1=0,k_upper2=3)
+   #plot_precentage_error(SNR=1,k_upper1=0,k_upper2=3)
    plot_precentage_error(SNR=-1,k_upper1=0,k_upper2=2)
    #plot_err_itr(num=4)

### 2016-12-07 software/plot_func.py
added the SNR plot

@@ -555,6 +555,38 @@ def plot_precentage_error(SNR=10,k_upper1=0,k_upper2=3,extra_string="_G_OLD_f"):
     ax.legend(loc=5)
     plt.show()
 
+def plot_prec_err_paper():
+    
+    SNR = np.array([-1,1,3,5,10,20,1000])
+    color = ['r','b','g','y','m','c','k'] 
+    matplotlib.rcParams.update({'font.size': 15})
+    fig = plt.figure()
+    ax = fig.add_subplot(1,1,1)
+
+    for k in xrange(len(SNR)):
+        p_file = "prec_error_"+str(SNR[k])+".p"
+        output = open(p_file, 'rb')
+        N1 = pickle.load(output)
+        precentage_stef_mean1 = pickle.load(output)
+        precentage_stef_std1 = pickle.load(output)
+        N1 = N1[2:] 
+        precentage_stef_mean1 = precentage_stef_mean1[2:]
+        precentage_stef_std1 = precentage_stef_std1[2:]
+        output.close()
+          
+        ax.plot(N1,precentage_stef_mean1*100,color[k],lw=2,label="SNR="+str(SNR[k]))
+        if SNR[k] <> 1000:
+           ax.fill_between(N1, (precentage_stef_mean1-precentage_stef_std1)*100, (precentage_stef_mean1+precentage_stef_std1)*100,alpha=0.2, edgecolor='k', facecolor=color[k])
+
+    ax.legend(prop={'size': 14})
+    ax.set_ylim([-5,30])
+    ax.set_xlim([37,217])
+    ax.set_xlabel("$N$ [antennas]")
+    ax.set_ylabel("% Error")
+    plt.grid('on')
+    plt.show()
+
+
 def plot_prec_err_presentation():
     output = open("prec_error_5.p", 'rb')
     
@@ -915,5 +947,6 @@ if __name__ == "__main__":
    #plot_precentage_error(SNR=5,k_upper1=0,k_upper2=3)
    #plot_precentage_error(SNR=3,k_upper1=0,k_upper2=3)
    #plot_precentage_error(SNR=1,k_upper1=0,k_upper2=3)
-   plot_precentage_error(SNR=-1,k_upper1=0,k_upper2=2)
+   #plot_precentage_error(SNR=-1,k_upper1=0,k_upper2=2)
+   plot_prec_err_paper()
    #plot_err_itr(num=4)

### 2016-12-08 paper.tex
added gain error table

@@
 
+\begin{table*}
+\centering
+\caption{The gain error models used in this paper. We have used the symbol $x$ here as a proxy as it can either refer to $t$ or $\nu$. We either
+performed our simulations over multiple time-slots and one frequency channel or one timeslot and multiple frequencey channels. We use $U$ to represent 
+a uniform distribution. Moreover, $c$ in the left most column denotes the speed of light.}
+\begin{tabular}{|c c c c|} 
+\hline
+Number tag & 1 & 2 & 3\\
+Model & Sinusoidal amplitude and phase & Sinusoidal real and imaginary parts & Phase slope \\ [0.5ex] 
+\hline\hline
+Function & $(A+1)e^{jP}$ & $A\cos(2\pi fx+B)+1.5A+jC\sin(2\pi fx+D)$ & $e^{jP}$ \\ 
+\hline
+Parameters & $A=a\cos(2\pi fx +b)$  & $f=5$ & $P=\tau x$ \\
+ & $P =c \cos(2\pi fx +d)$ & $A,C\sim U[0.5,10]$ & $\tau = \frac{l}{c}$ \\
+ & $f=5$ & $B,D\sim U[0,2\pi]$ &  $l\sim U[5,50]$\\
+ & $a\sim U[0.8,0.9]$ &  & \\ 
+ & $c\sim U[0.5,5]$ &  &  \\ 
+ & $b,d\sim U[0,2\pi]$ &  &  \\ 
+\hline
+\end{tabular}
+\end{table*}
+
 \subsection{Alternating Direction Implicit Method}

### 2016-12-08 software/simulator.py
fixed simulator

@@ -35,7 +35,7 @@ class sim():
       order - Fundamental parameter which determines the layout size
       '''
       
-      def __init__(self,h_min=-1,h_max=1,dec=(-74-39./60-37.481)*(np.pi/180),lat=(-30 - 43.0/60.0 - 17.34/3600)*(np.pi/180),freq=150*10**6,layout="HEX",nsteps=600,bas_len=50,order=1,seed=1):
+      def __init__(self,h_min=-1,h_max=1,dec=(-74-39./60-37.481/3600)*(np.pi/180),lat=(-30 - 43.0/60.0 - 17.34/3600)*(np.pi/180),freq=150*10**6,layout="HEX",nsteps=600,bas_len=50,order=1,seed=1):
           self.h_min = h_min
           self.h_max = h_max
           self.nsteps = nsteps

### 2016-12-08 paper.tex
fixed simulator
added fixed simulation parameters
added simulation details

@@
 \centering
+\caption{The following fixed simulation parameters were used to generate the results of this paper.}
+\begin{tabular}{|c c c c c c c|} 
+\hline
+Declination & Latitude & Minimum baseline & Geometry& Sources & Flux & Spatial \\
+\hline 
+\hline
+ $-74^{\circ}39'37.481''$ & $-30^{\circ}43'17.34''$ & 20 & Hexagonal &100 & Pareto (shape param. of 2) & $U[-3,3]$ (deg.) 
+\end{tabular}
+\end{table*}
+
+\begin{table}
+\centering
+\caption{We generated results using two major simulation setups. We either used one frequency channel and multiple time-slots, or one time-slot and many frequency 
+channels. The most important parameters used in realizing these two major setups are presented here.}
+\begin{tabular}{|c c c|} 
+\hline
+ & Setup 1 & Setup 2\\
+\hline
+\hline
+ Num. channels & 1024 & 1\\
+$\nu$-range & 100-200 MHz & 1.4 GHz\\
+Num. timeslots & 1 & 50\\
+%Hour angle &-1^h & [-1^h,1^h]\\
+Gain error model & 2 & 1\\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{table*}
+\centering
 \caption{The gain error models used in this paper. We have used the symbol $x$ here as a proxy as it can either refer to $t$ or $\nu$. We either
@@
 Number tag & 1 & 2 & 3\\
-Model & Sinusoidal amplitude and phase & Sinusoidal real and imaginary parts & Phase slope \\ [0.5ex] 
+Model & Sinusoidal: amplitude and phase & Sinusoidal: real and imaginary parts & Linear phase slope \\ [0.5ex] 
 \hline\hline
@@
  & $P =c \cos(2\pi fx +d)$ & $A,C\sim U[0.5,10]$ & $\tau = \frac{l}{c}$ \\
- & $f=5$ & $B,D\sim U[0,2\pi]$ &  $l\sim U[5,50]$\\
+ & $f=5$ & $B,D\sim U[0,2\pi]$ &  $l\sim U[5,50]$ (m)\\
  & $a\sim U[0.8,0.9]$ &  & \\ 

### 2016-12-08 paper.tex
added some text to simulation section

@@
 
+The results reported in this paper were generated using two major experimental setups, namely a single time-slot and multiple frequency channels setup or 
+a single frquency channel and multiple time-slots setup. The simulation parameters which where kept fixed in both these setups can be found in 
+Table~\ref{tab:fixed_parm}. The sky-model we used to simulate the visibilities were the same for both setups. We generated 100 sources, each source had random 
+flux and position. The flux of the sources followed a Pareto distribution (with a shape parameter of 2), while their positions were uniformly 
+distributed in a 3 by 3 degree square grid around the field center. We used a hexagonal redundant layout for our antenna positions. We produced 
+results for different
+
 \begin{table*}
@@
 \end{tabular}
+\label{tab:fixed_parm}
 \end{table*}
@@
 \centering
-\caption{We generated results using two major simulation setups. We either used one frequency channel and multiple time-slots, or one time-slot and many frequency 
+\caption{We generated results using two major simulation setups. We either used one frequency channel and multiple time-slots, or one time-slot and multiple frequency 
 channels. The most important parameters used in realizing these two major setups are presented here.}
@@
 %Hour angle &-1^h & [-1^h,1^h]\\
-Gain error model & 2 & 1\\
+Gain error model (Table~\ref{tab:gain_parm}) & 2 & 1\\
 \hline
 \end{tabular}
+\label{tab:ch_parm}
 \end{table}
@@
 \centering
-\caption{The gain error models used in this paper. We have used the symbol $x$ here as a proxy as it can either refer to $t$ or $\nu$. We either
+\caption{The gain error models used in this paper. We have used the symbol $x$ here as a proxy as it can either refer to time-slots or channels. We either
 performed our simulations over multiple time-slots and one frequency channel or one timeslot and multiple frequencey channels. We use $U$ to represent 
@@
 \end{tabular}
+\label{tab:gain_parm}
 \end{table*}

### 2016-12-09 paper.tex
finished simulation section

@@
 \subsection{Simulations}
-
-The results reported in this paper were generated using two major experimental setups, namely a single time-slot and multiple frequency channels setup or 
-a single frquency channel and multiple time-slots setup. The simulation parameters which where kept fixed in both these setups can be found in 
-Table~\ref{tab:fixed_parm}. The sky-model we used to simulate the visibilities were the same for both setups. We generated 100 sources, each source had random 
-flux and position. The flux of the sources followed a Pareto distribution (with a shape parameter of 2), while their positions were uniformly 
-distributed in a 3 by 3 degree square grid around the field center. We used a hexagonal redundant layout for our antenna positions. We produced 
-results for different
+The main purpose of the simulations conducted in this paper is to genrate experimental results with which we can compare the computational complexity of ADI and PCG.
+We used an hexagonal geometric layout with an adjustable number of rings during the
+simulations.
+For the skymodel we generated 100 sources, each source had a random flux value and position.
+The flux of the sources followed a Pareto distribution (with a shape parameter of 2), while their positions were uniformly
+distributed in a 3 by 3 degree square grid around the field center. The remaining fixed simulation parameters can all be found in Table~\ref{tab:fixed_parm}. 
+We also employed three different gain error models which we summarize in Table~\ref{tab:gain_parm}. We found that it did not matter which gain error model we used during
+our simulations, as the model we used did not have an impact on the results.
+Moreover, we had two major experimental setups, we either made use of simulated observations consisting of one frequency channel and multiple time-slots or one time-slot and multiple channels.
+We describe these two experimental setups in more detail in Table~\ref{tab:ch_parm}. As was the case for the gain error models, the experimental setup 
+that was used did not alter the simulation results.
+
+The main flow of a single simulation is as follow. We created corrupted visibilities, using either one of the experimental setups summarized in Table~\ref{tab:ch_parm} and
+the fixed simulation parameters presented in Table~\ref{tab:fixed_parm}. We then used per time-slot and channel redundant calibration, i.e. we either used the ADI or PCG algorithm, to 
+estimate the gains and the true observed visibilities. To obtain statistically significant results we performed, for both the ADI and PCG algorithms, multiple statistical independent simulations.
+
+%We compare ADI with PCG in Section \textbf{??}, by using the results generated from the simulations we performed using 
+%these two methods.
+
+%The results reported in this paper were generated using two major experimental setups, namely a single time-slot and multiple frequency channels setup or 
+%a single frquency channel and multiple time-slots setup. The simulation parameters which where kept fixed in both these setups can be found in 
+%Table~\ref{tab:fixed_parm}. The sky-model we used to simulate the visibilities were the same for both setups.   We used a hexagonal redundant layout for our antenna positions. We produced 
+%results for different
 
@@
 
-\begin{table}
-\centering
-\caption{We generated results using two major simulation setups. We either used one frequency channel and multiple time-slots, or one time-slot and multiple frequency 
-channels. The most important parameters used in realizing these two major setups are presented here.}
-\begin{tabular}{|c c c|} 
-\hline
- & Setup 1 & Setup 2\\
-\hline
-\hline
- Num. channels & 1024 & 1\\
-$\nu$-range & 100-200 MHz & 1.4 GHz\\
-Num. timeslots & 1 & 50\\
-%Hour angle &-1^h & [-1^h,1^h]\\
-Gain error model (Table~\ref{tab:gain_parm}) & 2 & 1\\
-\hline
-\end{tabular}
-\label{tab:ch_parm}
-\end{table}
-
 \begin{table*}
@@
 \caption{The gain error models used in this paper. We have used the symbol $x$ here as a proxy as it can either refer to time-slots or channels. We either
-performed our simulations over multiple time-slots and one frequency channel or one timeslot and multiple frequencey channels. We use $U$ to represent 
+performed our simulations over multiple time-slots and one frequency channel or one timeslot and multiple frequencey channels (see Table~\ref{tab:ch_parm}). We use $U$ to represent 
 a uniform distribution. Moreover, $c$ in the left most column denotes the speed of light.}
@@
 
+\begin{table}
+\centering
+\caption{We generated results using two major experimental setups. We either used one frequency channel and multiple time-slots, or one time-slot and multiple frequency 
+channels. The most important parameters used in realizing these two major setups are presented here. Note that we refer to the gain error models in Table~\ref{tab:gain_parm} in the 
+last row of this table.}
+\begin{tabular}{|c c c|} 
+\hline
+ & Setup 1 & Setup 2\\
+\hline
+\hline
+ Num. channels & 1024 & 1\\
+$\nu$-range & 100-200 MHz & 1.4 GHz\\
+Num. timeslots & 1 & 50\\
+%Hour angle &-1^h & [-1^h,1^h]\\
+Gain error model (Table~\ref{tab:gain_parm}) & 2 & 1\\
+\hline
+\end{tabular}
+\label{tab:ch_parm}
+\end{table}
+
 \subsection{Alternating Direction Implicit Method}

### 2016-12-09 paper.tex
added kappa and itr plots

@@
 kappa + itr plots
+\begin{figure*}
+\centering
+\subfigure[$\kappa$]{\includegraphics[width=0.47\textwidth]{./kappa.pdf}\label{fig:kappa}}
+% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
+\subfigure[Iterations required before and after Preconditioning]{\includegraphics[width=0.47\textwidth]{./cg_itr.pdf}\label{fig:cg_itr}}
+\caption{The graphs presented here were generated using the second experimental setup in Table~\ref{tab:ch_parm}.
+In the left panel we have plotted the spectral condition number $\kappa$ of the Hessian $\bH$, before (magenta and red curves) and after preconditioning (blue curve), at different SNR values as a function of the
+number of antennas $N$ in the array. This plot shows us that we can drastically reduce the condition number of $\bH$ by preconditioning. Moreover, $\kappa$ is also
+independent of the problem size (the number of antennas in the array). The fact that $\kappa$ is small and independent of problem size implies that it simply drops out 
+of Eq.~\eqref{eq:cg_bound}.
+The right hand panel confirms this result.
+In the right hand panel we have plotted the number of major iterations (line 6 of Algorithm \ref{algo:CG})) required by the conjugate
+gradient method to invert $\bH$, before (magenta and red curves) and after preconditioning (blue and green curves), at different SNR values as a function of the number of antennas $N$ in the array.  
+method. The right hand panel shows us that the number of major iterations needed by the preconditioned conjugate gradient method to invert $\bH$ is much smaller than the 
+dimension of $\bH$ and is independent of the problem size (the number of antennas in the array). 
+\label{fig:sup_anti}} 
+\end{figure*}
+
 \subsubsection{Sparsity}

### 2016-12-09 paper.tex
added kappa and itr plots to paper

@@
 number of antennas $N$ in the array. This plot shows us that we can drastically reduce the condition number of $\bH$ by preconditioning. Moreover, $\kappa$ is also
-independent of the problem size (the number of antennas in the array). The fact that $\kappa$ is small and independent of problem size implies that it simply drops out 
+independent of the problem size (the number of antennas in the array). The fact that $\kappa$ is small and independent of problem size (effectively constant) implies that it simply drops out 
 of Eq.~\eqref{eq:cg_bound}.
 The right hand panel confirms this result.
-In the right hand panel we have plotted the number of major iterations (line 6 of Algorithm \ref{algo:CG})) required by the conjugate
+In the right hand panel we have plotted the number of major iterations (line 6 of Algorithm \ref{algo:CG}) required by the conjugate
 gradient method to invert $\bH$, before (magenta and red curves) and after preconditioning (blue and green curves), at different SNR values as a function of the number of antennas $N$ in the array.  
-method. The right hand panel shows us that the number of major iterations needed by the preconditioned conjugate gradient method to invert $\bH$ is much smaller than the 
-dimension of $\bH$ and is independent of the problem size (the number of antennas in the array). 
+The right hand panel therefore shows us that the number of major iterations needed by the preconditioned conjugate gradient method to invert $\bH$ is independent of the number of antennas in the array
+and is much smaller than the dimension of $\bH$. 
 \label{fig:sup_anti}} 

### 2016-12-09 paper.tex
added some of the sparsity plots to the paper

@@
 and is much smaller than the dimension of $\bH$. 
-\label{fig:sup_anti}} 
+\label{fig:kappa_itr}} 
 \end{figure*}
@@
 \begin{align}
-\gamma &= \frac{5N^2-7N+3}{8N^2-8N+2} & \gamma_{\infty} &= \lim_{N\rightarrow \infty}\gamma = \frac{5}{8}. 
+\gamma &= \frac{5N^2-7N+3}{8N^2-8N+2} & \gamma_{\infty} &= \lim_{N\rightarrow \infty}\gamma = \frac{5}{8} \label{eq:gamma}. 
 \end{align}
@@
 \begin{align}
-c &= \log_{P}(1 - \gamma) + 2 & c_{\infty} &= \lim_{N\rightarrow \infty} c = 2.
+c &= \log_{P}(1 - \gamma) + 2 & c_{\infty} &= \lim_{N\rightarrow \infty} c = 2. \label{eq:c}
 \end{align}
@@
 itr numbers plot and graph of $kP = P^{1.7}$.
+
+\begin{figure*}
+\centering
+\subfigure[$\gamma$]{\includegraphics[width=0.47\textwidth]{./sparsity.pdf}\label{fig:kappa}}
+% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
+\subfigure[$c$]{\includegraphics[width=0.46\textwidth]{./comp_order.pdf}\label{fig:cg_itr}}
+\caption{In the left panel we plot the sparsity ratio $\gamma$ (the number of zero entries relative to the total number of entries in a matrix) of the Hessian $\bH$ as a function of the number 
+of antennas $N$ in the array for different geometries, namely an hexagonal grid (blue circles), a square grid (green crosses) and a regular east-west grid (red circles).
+For the regular east-west grid we are able to construct an analytical expression for $\gamma$ (the dashed red line).
+Taking the limit
+of this expression, Eq.~\eqref{eq:gamma}, as $N\rightarrow \infty$ produces the black dotted line.
+In the right panel we plot the order of the computational cost $c$ for inverting $\bH$ as a function of the number
+of antennas in the array for different geometries. We use the same color scheme in the right and left panels. We compute $c$ using Eq.~\eqref{eq:c}. We can again compute 
+an analytic expression (dashed red line) for the east-west grid using Eq.~\eqref{eq:c} and Eq.~\eqref{eq:gamma}.  
+\label{fig:sup_anti}} 
+\end{figure*}
 

### 2016-12-09 paper.tex
added some sparsity results

@@
 In the right panel we plot the order of the computational cost $c$ for inverting $\bH$ as a function of the number
-of antennas in the array for different geometries. We use the same color scheme in the right and left panels. We compute $c$ using Eq.~\eqref{eq:c}. We can again compute 
-an analytic expression (dashed red line) for the east-west grid using Eq.~\eqref{eq:c} and Eq.~\eqref{eq:gamma}.  
+of antennas in the array for different geometries. We use the same color scheme in the right and left panels. We compute $c$ using Eq.~\eqref{eq:c}. The red dashed line 
+in the righ panel is associated with the theoretical curve of $c$ we computed for the east-west regular grid. The limit of the red-dashed curve is two, which we 
+have plotted with a black dotted line.
 \label{fig:sup_anti}} 

### 2016-12-12 software/analytic.py
fixed deepcopy bug

@@ -165,6 +165,32 @@ class term():
 
 
       '''
+      The input term is the one we are copying to, the host term is the one we are copying
+      '''
+      def deepcopy_term(self,new_term):
+          if len(self.g_array) <> 0:
+             new_term.g_array = deepcopy(self.g_array) 
+          if len(self.y_array) <> 0:
+             new_term.y_array = deepcopy(self.y_array) 
+          if len(self.yc_array) <> 0:
+             new_term.yc_array = deepcopy(self.yc_array)
+          if len(self.gc_array) <> 0:
+             new_term.gc_array = deepcopy(self.gc_array)  
+          if len(self.ga_array) <> 0:
+             new_term.ga_array = deepcopy(self.ga_array)
+          if len(self.ya_array) <> 0:
+             new_term.ya_array = deepcopy(self.ya_array)
+          if len(self.constant_array) <> 0:
+             new_term.constant_array = deepcopy(self.constant_array)
+          if len(self.a_array) <> 0:
+             new_term.a_array = deepcopy(self.a_array)	
+          if len(self.b_array) <> 0:
+             new_term.b_array = deepcopy(self.b_array)	 
+          new_term.zero = deepcopy(self.zero)
+          new_term.const = deepcopy(self.const) 
+          return new_term
+                    
+      '''
       Substitutes real values of g and y into term to obtain result
 
       RETURNS:
@@ -329,9 +355,18 @@ class term():
       ''' 
       def conjugate(self):
           #print "self.g_array[0].to_string = ",self.g_array[0].to_string()
-          g_t = deepcopy(self.g_array)
-          self.g_array = deepcopy(self.gc_array)
+          if len(self.g_array) <> 0:
+             g_t = deepcopy(self.g_array)
+          else:
+             g_t = np.array([],dtype=object)
+          
+          if len(self.gc_array) <> 0:
+             self.g_array = deepcopy(self.gc_array)
+          else:
+             self.g_array = np.array([],dtype=object)            
+
           self.gc_array = g_t
+          
           #print "self.g_array[0].to_string = ",self.g_array[0].to_string()
 
           for k in xrange(len(self.g_array)):
@@ -340,8 +375,21 @@ class term():
           for k in xrange(len(self.gc_array)):
               self.gc_array[k].conjugate() 
 
-          y_t = deepcopy(self.y_array)
-          self.y_array = deepcopy(self.yc_array)
+          #OLD CODE
[…]

### 2016-12-12 software/H_gen.py
depricated H_gen.py

@@ -1,0 +1,1319 @@
+#DEPRICATED AND NOT FUNCTIONING DUE TO THE DEEPCOPY BUG
+
+import numpy as np
+import pylab as plt
+import pickle
+from scipy import optimize
+import matplotlib as mpl
+from copy import deepcopy
+import networkx as nx
+from networkx.utils import reverse_cuthill_mckee_ordering, cuthill_mckee_ordering
+
+
+
+
+class factor():
+
+    def __init__(self,
+                type_factor = "g",
+                index = 1, 
+                exponent = 1,
+                conjugate_value = False,
+                mag = False,
+                ant_p = 0,
+                ant_q = 0,
+                print_f = True):
+                self.type_factor = type_factor
+                self.index = index
+                self.exponent = exponent
+                self.conjugate_value = conjugate_value
+                self.mag = mag
+                self.ant_p = ant_p
+                self.ant_q = ant_q
+                self.print_f = print_f 
+             
+
+    def to_string(self):
+        if self.print_f:
+           if self.type_factor == "y":
+              string_out = self.type_factor + "_{" + str(self.ant_p)+str(self.ant_q)+"}"
+           else:
+              string_out = self.type_factor + "_" + str(self.index)
+        else:   
+           string_out = self.type_factor + "_" + str(self.index)
+        if self.conjugate_value:
+           string_out = string_out + "^" + "*"
+           if self.exponent > 1:
+              string_out = "("+string_out+")"+"^"+str(self.exponent)
+        else:
+           if self.mag:
+              string_out = "|"+string_out+"|^2"
+           else:  
+              if self.exponent > 1:
+                 string_out = string_out+"^"+str(self.exponent)
+        return string_out   
+           
+    def conjugate(self):
+        if self.conjugate_value:
+           self.conjugate_value = False
+        else:
+           self.conjugate_value = True
+
+    def equal(self,factor_in):
+        if self.type_factor == factor_in.type_factor:
+           if self.index == factor_in.index:
+              if self.exponent == factor_in.exponent:
+                 if self.conjugate_value == factor_in.conjugate_value:
+                    return true
+        return false
+
+class term():
+      def __init__(self):
+          self.g_array = np.array([],dtype=object)
+          self.y_array = np.array([],dtype=object)
+          self.yc_array = np.array([],dtype=object) 
+          self.gc_array = np.array([],dtype=object)
[…]

### 2016-12-12 software/analytic.py
added code to generate Hessians for paper

@@ -2687,7 +2687,7 @@ EXAMPLE OF HOW TO USE THE CODE TO CREATE ANALYTIC EXPRESSION OF COMPLEX HESSIAN
 def Complex_example():
     r = redundant()
     #r.create_regular()
-    r.create_redundant(layout="HEX",order=1,print_pq=False) 
+    r.create_redundant(layout="REG",order=5,print_pq=False) 
     r.create_J1()
     r.create_J2()
     r.conjugate_J1_J2()
@@ -2767,11 +2767,57 @@ def Simple_example():
  
     print "e1 = ",e1.to_string(simplify_const=True)
 
+'''
+PLOT HESSIANS FOR PAPER
+'''
+def plot_hessian_paper(layout="REG",order=5,print_pq=False,label_size=20,cs="jet",step2=4):
+    r = redundant()
+    #r.create_regular()
+    r.create_redundant(layout=layout,order=order,print_pq=print_pq) 
+    r.create_J1()
+    r.create_J2()
+    r.conjugate_J1_J2()
+    r.create_J()
+    r.hermitian_transpose_J()
+    r.compute_H()          
+    H_int = r.to_int_H()
+    mpl.rcParams['xtick.labelsize'] = label_size 
+    mpl.rcParams['ytick.labelsize'] = label_size 
+    data = H_int
+    # get discrete colormap
+    cmap = plt.get_cmap(cs, np.max(data)-np.min(data)+1)
+    # set limits .5 outside true range
+    mat = plt.matshow(data,cmap=cmap,vmin = np.min(data)-.5, vmax = np.max(data)+.5)
+    #tell the colorbar to tick at integers
+    ticks = ticks=np.arange(np.min(data),np.max(data)+1,1)
+    cax = plt.colorbar(mat, ticks=ticks)
+    x = np.arange(data.shape[0],step=step2,dtype=int)
+    if x[-1] <> data.shape[0]-1:
+       x = np.append(x,np.array([data.shape[0]-1]))
+    plt.xticks(x, x+1)
+    y = np.arange(data.shape[0],step=step2,dtype=int)
+    if y[-1] <> data.shape[0]-1:
+       y = np.append(y,np.array([data.shape[0]-1]))
+    plt.yticks(y, y+1)
+    plt.xlabel("$j$",fontsize=label_size+5)
+    plt.ylabel("$i$",fontsize=label_size+5)
+    #plt.title("$\zeta_{pq}$",fontsize=label_size+5)
[…]

### 2016-12-12 paper.tex
added more plots to the paper

@@
 
+\begin{figure*}
+\centering
+\subfigure[Regular layout]{\includegraphics[width=0.47\textwidth]{./reg_hessian.pdf}\label{fig:hessian_reg}}
+% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
+\subfigure[Hexagonal layout]{\includegraphics[width=0.47\textwidth]{./hex_hessian.pdf}\label{fig:hessian_hex}}
+\caption{Hallo
+\label{fig:hessian}} 
+\end{figure*}
+
 %\subsection{Diagonalization and StEFCal}
@@
 sparsity + computational complexity plots
-
-\subsection{Comparison}
-itr numbers plot and graph of $kP = P^{1.7}$.
-
 \begin{figure*}
@@
 
+\subsection{Comparison}
+itr numbers plot and graph of $kP = P^{1.7}$.
+
+\begin{figure*}
+\centering
+\subfigure[Regular layout]{\includegraphics[width=0.47\textwidth]{./outerloop.pdf}\label{fig:outerloop}}
+% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
+\subfigure[Hexagonal layout]{\includegraphics[width=0.47\textwidth]{./diff.pdf}\label{fig:diff}}
+\caption{Hallo
+\label{fig:out_diff}} 
+\end{figure*}
+
 % fiducial initial guess 

### 2016-12-12 software/plot_func.py

@@ -559,7 +559,7 @@ def plot_prec_err_paper():
     
     SNR = np.array([-1,1,3,5,10,20,1000])
     color = ['r','b','g','y','m','c','k'] 
-    matplotlib.rcParams.update({'font.size': 15})
+    matplotlib.rcParams.update({'font.size': 22})
     fig = plt.figure()
     ax = fig.add_subplot(1,1,1)
 
@@ -578,7 +578,7 @@ def plot_prec_err_paper():
         if SNR[k] <> 1000:
            ax.fill_between(N1, (precentage_stef_mean1-precentage_stef_std1)*100, (precentage_stef_mean1+precentage_stef_std1)*100,alpha=0.2, edgecolor='k', facecolor=color[k])
 
-    ax.legend(prop={'size': 14})
+    ax.legend(prop={'size': 18})
     ax.set_ylim([-5,30])
     ax.set_xlim([37,217])
     ax.set_xlabel("$N$ [antennas]")
@@ -653,11 +653,11 @@ def plot_outerloop_pres():
     ax = fig.add_subplot(1,1,1)
     ax.plot(N1,outerloop_pcg_mean1,"r",lw=2,label="PCG (SNR=1000)")
     ax.fill_between(N1,outerloop_pcg_mean1-outerloop_pcg_std1, outerloop_pcg_mean1+outerloop_pcg_std1,alpha=0.2, edgecolor='k', facecolor='r')
-    ax.plot(N1,outerloop_stef_mean1,"b",lw=2,label="StEf (SNR=1000)")
+    ax.plot(N1,outerloop_stef_mean1,"b",lw=2,label="ADI (SNR=1000)")
     ax.fill_between(N1, outerloop_stef_mean1-outerloop_stef_std1, outerloop_stef_mean1+outerloop_stef_std1,alpha=0.2, edgecolor='k', facecolor='b')
     ax.plot(N1,outerloop_pcg_mean2,"m",lw=2,label="PCG (SNR=5)")
     ax.fill_between(N2,outerloop_pcg_mean2-outerloop_pcg_std2, outerloop_pcg_mean2+outerloop_pcg_std2,alpha=0.2, edgecolor='k', facecolor='m')
-    ax.plot(N1,outerloop_stef_mean2,"g",lw=2,label="StEf (SNR=5)")
[…]

### 2016-12-12 paper.tex
added pl graphs

@@
 \subsection{Indexing Functions}
+\begin{figure*}
+\centering
+\subfigure[Regular layout]{\includegraphics[width=0.47\textwidth]{./ln.pdf}\label{fig:ln}}
+% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
+\subfigure[Hexagonal layout]{\includegraphics[width=0.47\textwidth]{./pn.pdf}\label{fig:pn}}
+\caption{Hallo
+\label{fig:pl}} 
+\end{figure*}
+
 In general there are two ways of indexing the true visibilities associated with different baselines.

### 2016-12-12 software/plot_func.py

@@ -782,6 +782,40 @@ def plot_k():
     
     plt.show() 
 
+def plot_P():
+    matplotlib.rcParams.update({'font.size': 22})
+    N = np.arange(7,218)
+
+    P_reg = 4*N-2
+    P_hex = 6*N - np.sqrt(12*N-3) - 1
+    P_sqr = 6*N - 4*np.sqrt(N)
+
+    plt.plot(N,P_reg,label="REG",lw=2)
+    plt.plot(N,P_hex,label="HEX",lw=2)
+
+    plt.xlabel(r'$N$ [antennas]')
+    plt.ylabel(r'$P$')
+    plt.legend(loc=5,prop={'size': 18})
+    plt.xlim([37,217])
+    
+    plt.grid('on')
+
+    plt.show()
+
+    L_reg = N-1
+    L_hex = 2*N - 0.5*np.sqrt(12*N-3) - 0.5 
+    plt.plot(N,L_reg,label="REG",lw=2)
+    plt.plot(N,L_hex,label="HEX",lw=2)
+    #plt.plot(N,P_sqr)
+    plt.xlabel(r'$N$ [antennas]')
+    plt.ylabel(r'$L$')
+    plt.legend(loc=5,prop={'size': 18})
+    plt.xlim([7,217])
+    
+    plt.grid('on')
+
+    plt.show()   
+
 
 def plot_time(SNR=1000,k_upper1=5,k_upper2=5,k_upper3=5):
     
@@ -995,7 +1029,8 @@ if __name__ == "__main__":
    #plot_kappa_itr_pres()
    #plot_prec_err_paper()
    #plot_sparsity()
-   plot_k()
+   #plot_k()
+   plot_P()
 
    #plot_kappa_itr(SNR=5)
    #plot_outer_loop(SNR=5)

### 2016-12-14 paper.tex
added prec error plot
started red calibration section properly

@@
 \section{Introduction}
+
+\subsection{Historical Overview}
+
+[Redundant calibration has been in use for a long time. Noordam used redundant calibration to produce high-dynamic range images of 3C84 with the 
+WSRT array.] RW The first implementations of redundant calibration made use of the logarithm function, because when the logarithm is applied to the scalar measurement equation redundant 
+calibration is transformed into a linear problem cite Wieringa,camps and Lui. This approach is also generally referred to as LOGCAL.
+LOGCAL is reviewed in Appendix ??.
+The next improvement was to perform a firt-order taylor expansion (which is equivalent to the GN algorithm Kurien 2016) of the measurement equation an then to solve for the
+perturbations of the model-parameters.
+This approach is commonly referred to as LINCAL.
+LINCAL is derived from first principals in Appendix .
+Two ways of speeding up redundant calibration has also been
+proposed. The first technique makes use of the alternating direction implicit method, in which we alternate between solving the gains, their conjugates and 
+the true observed visibilities (wijnholds and marthi). Using the conjugate gradient method to exploit the sparsity inherent in the redundant calibration problem, 
+and in doing so speedup the algorithm itself was proposed in . A proper comparison between these two approaches has not been done before.
+
+Newly built instruments like LOFAR, ..... has also sparked a renewed interest in redundant calibration. Redundancy is also advantageous for detecting the EoR,
+which explains why some of the previsously mentioned instruments have redundant layouts.
+
+Layered approach, LOGCAL, LINCAL, Stefan's conference paper, Noordam, other ASTRON paper, newest papers on redundant calibration, marthi paper etc...
 
@@
 \section{Redundant Calibration}
-
-\subsection{Historical Overview}
-
-[Redundant calibration has been in use for a long time. Noordam used redundant calibration to produce high-dynamic range images of 3C84 with the 
-WSRT array.] RW The first implementations of redundant calibration made use of the logarithm function, because when the logarithm is applied to the scalar measurement equation redundant 
-calibration is transformed into a linear problem cite Wieringa,camps and Lui. This approach is also generally referred to as LOGCAL.
-LOGCAL is reviewed in Appendix ??.
-The next improvement was to perform a firt-order taylor expansion (which is equivalent to the GN algorithm Kurien 2016) of the measurement equation an then to solve for the
-perturbations of the model-parameters.
-This approach is commonly referred to as LINCAL.
-LINCAL is derived from first principals in Appendix .
-Two ways of speeding up redundant calibration has also been
-proposed. The first technique makes use of the alternating direction implicit method, in which we alternate between solving the gains, their conjugates and 
-the true observed visibilities (wijnholds and marthi). Using the conjugate gradient method to exploit the sparsity inherent in the redundant calibration problem, 
-and in doing so speedup the algorithm itself was proposed in . A proper comparison between these two approaches has not been done before.
-
-Newly built instruments like LOFAR, ..... has also sparked a renewed interest in redundant calibration. Redundancy is also advantageous for detecting the EoR,
-which explains why some of the previsously mentioned instruments have redundant layouts.
-
-Layered approach, LOGCAL, LINCAL, Stefan's conference paper, Noordam, other ASTRON paper, newest papers on redundant calibration, marthi paper etc...
+Redundant calibration can be formulated as a least-squares problem:
+\begin{equation}
+\label{eq:least_squares_red}
+\min_{\bz} \|\br\| = \min_{\bz} \|\bd - \bv(\bz)\|, 
+\end{equation}
+where $\bz = [\bg^T,\by^T]^T$. As discussed in Section~\ref{}, we can solve the above minimization problem in one of two ways. We can split the problem into its real and
+imaginary parts and then solve for the real and imaginary parts of the parameters seperately or we can employ the notion of Wirtinger derivatives. The former 
+approach is discussed in detail in Section~\ref{}, while the latter approach is described in Section~\ref{}.
 
@@
 
+\begin{figure}
+\includegraphics[width=0.47\textwidth]{./prec_error.pdf} 
+\caption{Precentage Error}
+\label{fig:prec_error}
+\end{figure}
+
 %\subsection{Diagonalization and StEFCal}

### 2016-12-14 paper.tex
added some text to intro

@@
 \section{Introduction}
+Calibration is the procedure by which we remove the instrumental and environmental iduced errors from our intereferometric data. We normally model 
+the propogation effects This is normally accomplished using a 
+least-squares solver, i.e.
+we find the propogation effects (antenna gains) that minimizes the difference betweem our observed and predicted visibilities \citep{RIME1}.
+There are two recent developments
+in Calibration approaches that are worth
 

### 2016-12-14 paper.tex
started writing a brief introduction

--- 2016-12-14 paper.tex
@@
 \section{Introduction}
+The propogation effects (i.e.
+the ionosphere, antenna gains, etc.) a celestial signal experiences along its propogation path is modelled with the radio interforemetric measurement equation (RIME).
+%\citep{ME1,RRIME1}.
+Calibration is the procedure by which we remove the errors that these propogation effects induce onto our intereferometric data and is normally accomplished using a 
+least-squares solver.
+In the case where we only model the antenna gains, which is the most basic form of the RIME, we can define calibration as finding the antenna gains that minimizes the difference betweem our observed and predicted visibilities.
+There are two recent developments
+in calibration which should be explicitly mentioned, since they form the foundation this paper builds on. The first is the discovery of StEfCal \citep{salvini2014}, which 
+is an alternating direction implicit (ADI) skymodel-based calibration algorithm. It works by assuming the gains are known while solving for their conjugates. Now we do the reverse,
+assume the conugates are known and solve for the gains. This procedure results in the linearization of the calibration algorithm and reduces the computational complexity 
+of calibration from $\mathbb{O}(N^3)$ to $\mathbb{O}(N^3)$, where $N$ denotes the number of antennas in the array. The second is to not break the calibration problem 
+into its real and imaginary parts and then to solve for the real and imaginary parts of the parameters seperately, but to rather treat the antenna gains and their 
+conjugates as seperate parameters to estimate \citep{smirnov2015}. \citet{smirnov2015} also showed that the StEfCal algorithm itself is merely a special case of this
+gain-and-conjugate framework.
+
+Redundant calibration differs from normal calibration in that it does not require a skymodel. With redundant calibration we do not only solve for the antenna gains, but also 
+for the true observed visibilities. We are able to do so, because we have less unkonws than baseline equations. Since we
 

### 2016-12-14 paper.tex
added new text to intro

@@
 assume the conugates are known and solve for the gains. This procedure results in the linearization of the calibration algorithm and reduces the computational complexity 
-of calibration from $\mathbb{O}(N^3)$ to $\mathbb{O}(N^3)$, where $N$ denotes the number of antennas in the array. The second is to not break the calibration problem 
+of calibration from $\mathbb{O}(N^3)$ to $\mathbb{O}(N^3)$, where $N$ denotes the number of antennas in the array. The second development is the idea; to not break the calibration problem 
 into its real and imaginary parts and then to solve for the real and imaginary parts of the parameters seperately, but to rather treat the antenna gains and their 
 conjugates as seperate parameters to estimate \citep{smirnov2015}. \citet{smirnov2015} also showed that the StEfCal algorithm itself is merely a special case of this
-gain-and-conjugate framework.
-
+gain-and-conjugate approach.
+
+Two baselines are redundant if their beaseline difference vector have the same lenght and orientation, which imply that they measure the same $uv$-modes.
+When we use a redundant array layout we dramatically
+reduce the number of unkowns. Redundant calibration is the procedure by which we leverage the fact that we have 
 Redundant calibration differs from normal calibration in that it does not require a skymodel. With redundant calibration we do not only solve for the antenna gains, but also 
 for the true observed visibilities. We are able to do so, because we have less unkonws than baseline equations. Since we
-
-\subsection{Historical Overview}
 
@@
 
-Main contributions of the paper:
-
-\begin{enumerate}
- \item Present a general framework which unifies all the techniques developed so far showing they are all related (LINCAL, non-linear estimator, etc ...). They are all non-linear
- least-squares techniques, i.e. they either employ Gauss-Newton or the Levenberg-Marquardt algorithms. We have to start with LINCAL showing that it is GN, then we have to motivate 
- why we want to use Oleg's complex optimization framework.
- \item Use Oleg's complex optimization framework to re-derive the non-linear technique proposed by Marthi and Chengular. The novelty lies, in the fact that by using Oleg's
- framework we can find analytic expressions for the Jacobian, the Hessian and the Jacobian-residual product (which is not even the case for LINCAL). State that the algorithm is 
- effectively related to SteFCal and is eff an independent rediscovery of SteFCal and an extension of it into the redundant domain.
- \item Also we present the array geometry function to help us make the derivations from Marthi and Chengular easier to read and understand for a general layout.
- \item We also mention at this point that in Oleg's paper the question is raised is there a fast way of computing the exact inverse, we then present this new method, which is called
- conjugate gradient method. We discuss the algorithm and the two things which bound its execution time. Which is $kappa$ and $m$ (spectral condition number and its sparsity). Then
- we explore both of these parameters.
- \end{enumerate}
-
- **************
- FLOW OF PAPER
- **************
-
- \begin{enumerate}
- \item We need the definition of visibilities as in Liu. DEFINE SNR HERE ALREADY together maybe with the sigma value of the noise.
- \item Introduce the array geometery function.
- \item Write down logcal and lincal in matrix form... ?
- \item Short discussion of least squares and jacobian and hessian's. General GN and LM update rules.
- \item Introduce redundant calibration as a least squares problem. We will use this general fact to derive both popular methods.
- \item Propose a possible solution witch leads to lincal. Mention that this approach works as in this form the function is differentiable (a taylor expansion in the 
- parameters exists). Maybe mention that we can also divide the problem into real and imaginary etc...
- \item Show how this relates to lincal for example ---> show lincal is GN.
- \item Now introduce complex optimization ---> Main motivation for switching to the alternative framework is that the the differentials are very simplistic. We wish to show that
- we can derive the method of Chengular.
- \item Do the derivation of Chengular. 
- \item Brief discussion abouth Chengular and SteFCal and Complex Optimization. Here I show that in Oleg's paper he re-derives an algorithm called SteFCal. Stefcal works
- on the basis of alternating direction implicit method. The first signs of achieving a similar algorithm already appeared in Stefan's conference paper in which the alternating
- idea was first proposed. Then I mention Chegular re-derives the expression by using partial derivatives and extends it to redundant. One could also have used the linear alternating
- approach. In an attempt to merge the terminology that has independently develop in the general calibration and redundant calibration literature and to emphasize the close
- relationship between Stefcal and the approach derived by I will use the label Redunandat SteFCal (R-Stefcal) to refer to the NLS method proposed by ... Lastly we mention that similarly
- to how Oleg re-derived stefcal in, we have achieved the same approach.
- \item Faster Exact inverse. A question that Oleg poses in his paper is, does there exist a faster way to take the exact inverse of JhJ? One that is almost linear, and
- implies that we can therefore implement the full LM algorithm. The aim here is of course to reduce the number of iterations that are required to converge by using the 
- full inverse instead which would hopefully provide enough of a speedup to compensate for the more expensive full-inverse. The algorithm we propose is the conjugate 
- gradient method.
- \item We give the images of the HESSIAN of both lincal and the complex method (number of terms). So we can mention that both are sparse and contain diagonal entries that 
- are more significant than the off-diagonal entries. What linear inversion approach can take advantage of both these phenomenon. One such technique is 
- cg. 
- \item Briefly discuss CG and how its computationally bounded by its condition number and its sparsity (how does the diagonal play a role).
- \item Simulation description
- \item Will CG improve things?
- \item Now first discuss the condition
- number of the Hessian before and after pre-conditioning (pre-codnitioning can only be applied if a good inverse of a matrix is known, if it is known then it can improve 
- the spectral condition number of a matrix. We present here the kappa and iteration number graphs for the HEXAGONAL layout. Although the
- \item Now we discuss the sparsity results. 
- \item Provide a table that theoretical compares R-StEFCal and SPARC.
- \item Number of outerloop iterations. 
- \item Timing results.
- \item Accuracy results.
- \item Maybe some freq simulations.
- \end{enumerate}
-
- We
+EOR SECTION SOMEWHERE HERE
+
+MAIN CONTRIBUTION OF THIS PAPER COMES HERE
+
+% Main contributions of the paper:
+% 
+% \begin{enumerate}
+%  \item Present a general framework which unifies all the techniques developed so far showing they are all related (LINCAL, non-linear estimator, etc ...). They are all non-linear
+%  least-squares techniques, i.e. they either employ Gauss-Newton or the Levenberg-Marquardt algorithms. We have to start with LINCAL showing that it is GN, then we have to motivate 
+%  why we want to use Oleg's complex optimization framework.
+%  \item Use Oleg's complex optimization framework to re-derive the non-linear technique proposed by Marthi and Chengular. The novelty lies, in the fact that by using Oleg's
+%  framework we can find analytic expressions for the Jacobian, the Hessian and the Jacobian-residual product (which is not even the case for LINCAL). State that the algorithm is 
+%  effectively related to SteFCal and is eff an independent rediscovery of SteFCal and an extension of it into the redundant domain.
+%  \item Also we present the array geometry function to help us make the derivations from Marthi and Chengular easier to read and understand for a general layout.
+%  \item We also mention at this point that in Oleg's paper the question is raised is there a fast way of computing the exact inverse, we then present this new method, which is called
+%  conjugate gradient method. We discuss the algorithm and the two things which bound its execution time. Which is $kappa$ and $m$ (spectral condition number and its sparsity). Then
+%  we explore both of these parameters.
+%  \end{enumerate}
+%  
+%  **************
+%  FLOW OF PAPER
+%  **************
+%  
+%  \begin{enumerate}
+%  \item We need the definition of visibilities as in Liu. DEFINE SNR HERE ALREADY together maybe with the sigma value of the noise.
+%  \item Introduce the array geometery function.
+%  \item Write down logcal and lincal in matrix form... ?
+%  \item Short discussion of least squares and jacobian and hessian's. General GN and LM update rules.
+%  \item Introduce redundant calibration as a least squares problem. We will use this general fact to derive both popular methods.
+%  \item Propose a possible solution witch leads to lincal. Mention that this approach works as in this form the function is differentiable (a taylor expansion in the 
+%  parameters exists). Maybe mention that we can also divide the problem into real and imaginary etc...
+%  \item Show how this relates to lincal for example ---> show lincal is GN.
+%  \item Now introduce complex optimization ---> Main motivation for switching to the alternative framework is that the the differentials are very simplistic. We wish to show that
+%  we can derive the method of Chengular.
+%  \item Do the derivation of Chengular. 
+%  \item Brief discussion abouth Chengular and SteFCal and Complex Optimization. Here I show that in Oleg's paper he re-derives an algorithm called SteFCal. Stefcal works
+%  on the basis of alternating direction implicit method. The first signs of achieving a similar algorithm already appeared in Stefan's conference paper in which the alternating
+%  idea was first proposed. Then I mention Chegular re-derives the expression by using partial derivatives and extends it to redundant. One could also have used the linear alternating
+%  approach. In an attempt to merge the terminology that has independently develop in the general calibration and redundant calibration literature and to emphasize the close
+%  relationship between Stefcal and the approach derived by I will use the label Redunandat SteFCal (R-Stefcal) to refer to the NLS method proposed by ... Lastly we mention that similarly
+%  to how Oleg re-derived stefcal in, we have achieved the same approach.
+%  \item Faster Exact inverse. A question that Oleg poses in his paper is, does there exist a faster way to take the exact inverse of JhJ? One that is almost linear, and
+%  implies that we can therefore implement the full LM algorithm. The aim here is of course to reduce the number of iterations that are required to converge by using the 
+%  full inverse instead which would hopefully provide enough of a speedup to compensate for the more expensive full-inverse. The algorithm we propose is the conjugate 
+%  gradient method.
+%  \item We give the images of the HESSIAN of both lincal and the complex method (number of terms). So we can mention that both are sparse and contain diagonal entries that 
+%  are more significant than the off-diagonal entries. What linear inversion approach can take advantage of both these phenomenon. One such technique is 
+%  cg. 
+%  \item Briefly discuss CG and how its computationally bounded by its condition number and its sparsity (how does the diagonal play a role).
+%  \item Simulation description
+%  \item Will CG improve things?
+%  \item Now first discuss the condition
+%  number of the Hessian before and after pre-conditioning (pre-codnitioning can only be applied if a good inverse of a matrix is known, if it is known then it can improve 
+%  the spectral condition number of a matrix. We present here the kappa and iteration number graphs for the HEXAGONAL layout. Although the
+%  \item Now we discuss the sparsity results. 
+%  \item Provide a table that theoretical compares R-StEFCal and SPARC.
+%  \item Number of outerloop iterations. 
+%  \item Timing results.
+%  \item Accuracy results.
+%  \item Maybe some freq simulations.
+%  \end{enumerate}
+%  
+%  We
 

### 2017-01-03 paper.tex
busy fixing section 5

@@
 There are two recent developments
-in calibration which should be explicitly mentioned, since they form the foundation this paper builds on. The first is the discovery of StEfCal \citep{salvini2014}, which 
+in calibration which should be explicitly mentioned, since they form the foundation this paper builds on. The first is the discovery of StEfCal \citep{Salvini2014}, which 
 is an alternating direction implicit (ADI) skymodel-based calibration algorithm. It works by assuming the gains are known while solving for their conjugates. Now we do the reverse,
@@
 into its real and imaginary parts and then to solve for the real and imaginary parts of the parameters seperately, but to rather treat the antenna gains and their 
-conjugates as seperate parameters to estimate \citep{smirnov2015}. \citet{smirnov2015} also showed that the StEfCal algorithm itself is merely a special case of this
+conjugates as seperate parameters to estimate \citep{Smirnov2015}. \citet{smirnov2015} also showed that the StEfCal algorithm itself is merely a special case of this
 gain-and-conjugate approach.
@@
 \end{align}
+
+NB COLUMN VECTORS NEED TO ADD THIS...
+
 In Eq.~\eqref{eq:vec_definitions}, the size of the three vectors are equal to $B$ (i.e. $p<q$), which denotes the number of baselines in the array. Moreover,
@@
 \section{Least Squares}
+\label{sec:ls}
 The method of nonlinear least squares is tanatamount to solving the following optimization problem:  
@@
 
-Redundant calibration is achieved by solving the following optimization problem:
-
 The only way to circumvent the differentiability conundrum, associated with complex least square problems, is to recast our complex optimization problem in such a way that $\bnu$ becomes analytic in its argument.
@@
 \label{eq:least_squares_red}
-\min_{\bz} \|\br\| = \min_{\bz} \|\bd - \bv(\bz)\|, 
-\end{equation}
-where $\bz = [\bg^T,\by^T]^T$. As discussed in Section~\ref{}, we can solve the above minimization problem in one of two ways. We can split the problem into its real and
+\min_{\bz} \Lambda(\bz) = \min_{\bz} \|\br\|_F^2 = \min_{\bz} \|\bd - \bv(\bz)\|_F^2, 
+\end{equation}
+where $\bz = [\bg^T,\by^T]^T$\footnote{In this paper we use the same symbol to denote corrupted predicted visibilities and true observed corrupted visibilities as it
+is trivial to distinguish which one is implied from the context.}.
+As discussed in Section~\ref{sec:ls}, we can solve the above minimization problem in one of two ways.
+We can split the problem into its real and
 imaginary parts and then solve for the real and imaginary parts of the parameters seperately or we can employ the notion of Wirtinger derivatives. The former 
-approach is discussed in detail in Section~\ref{}, while the latter approach is described in Section~\ref{}.
+approach is discussed in detail in Section~\ref{sec:ri}, while the latter approach is described in Section~\ref{sec:w}.
 
 \subsection{Real-Imag Framework}
-Here we discuss the fact that LINCAL is just GN
-
-OLD COPIED TEXT
-
-As we mentioned in the beginning of this section, the Wirtinger framework is not the only way of ensuring that $\bv$ is analytic in its argument. We could also 
-have divided the problem in Eq.~\eqref{eq:least_squares_red} into its real and imaginary components after-which we solve for the real and imaginary parts of the parameters separately.
-There is a third way
-of dealing with the issue.
-We can split the problem in Eq.~\eqref{eq:least_squares_complex} into its phase and amplitude components and then solve for the amplitude and phase of the parameters separately.
-This can be achieved by first rewriting Eq.~\eqref{eq:least_squares_complex} to
+\label{sec:ri}
+As we mentioned in Section~\ref{sec:ls} using least-squares to solve complex problems can be problematic as $\bv$ needs to be analytic in its argument. In the case 
+of Eq.~\eqref{eq:least_squares_red}, \citet{Liu2010} circumvents this difficulty by 
+splitting the problem into its phase and amplitude components.
+
+The aforementioned is achieved by first rewriting Eq.~\eqref{eq:least_squares_red} to
 \begin{equation}
@@
 \end{equation}
-where $\bmath{\varrho}$ is the phase and amplitude parameter vector and is defined in Appendix~\ref{sec:lincal}. The Jacobian associated with 
-Eq.~\eqref{eq:least_squares_lincal} is equal to Eq.~\eqref{eq:Jac_lin}. It should therefore be obvious to the reader that 
-performing one iteration of LINCAL is equivalent to a basic GN update step; the GN update step was just re-derived in Appendix~\ref{sec:lincal} from first principles.
+where $\breve{\bd} = [\bd^T,\conj{\bd}^T]^T$, $\breve{\bv} = [\bv^T,\conj{\bv}^T]$ and $\bmath{\varrho}$ is the phase and amplitude parameter vector which we define in Appendix~\ref{sec:lincal}. 
+With the aid of Eq.~\eqref{eq:least_squares_lincal} we can derive the LINCAL algorithm (see Appendix~\ref{sec:lincal}).
+The Jacobian associated with Eq.~\eqref{eq:least_squares_lincal} is equal to Eq.~\eqref{eq:Jac_lin}. 
+At the hand of Appendix~\ref{sec:lincal} it becomes evident that LINCAL is equivalent to a basic GN update step \citep{Kurien2016}. 
 Therefore, a straightforward way of improving the convergence properties of stock standard LINCAL would be to simply add the LM damping factor to the basic LINCAL update step.
@@
 \subsection{Complex Framework}
-
-\citet{Smirnov2015} recently proposed to use this Wirtinger framework instead of the classic real and imaginary approach to perform normal skymodel-based calibration. We will now show
-that the Wirtinger framework can also be applied to redundant calibration.
-
-We can perform redundant calibration by solving the following optimization problem: 
-\begin{equation}
-\label{eq:least_squares_red}
-\min_{\bz} \Lambda(\bz) = \min_{\bz} \|\br\|_F^2 = \min_{\bz} \|\bd - \bv(\bz)\|_F^2, 
-\end{equation}
-where $\bz = [\bg,\by]^T$.
+\label{sec:w}
+
+\citet{Smirnov2015} recently proposed to use the Wirtinger derivative instead of the classic real and imaginary approach to perform normal skymodel-based calibration. We will now show
+that the Wirtinger derivative can also be applied to redundant calibration.
+
+%We can perform redundant calibration by solving the following optimization problem: 
+%\begin{equation}
+%\label{eq:least_squares_red}
+%\min_{\bz} \Lambda(\bz) = \min_{\bz} \|\br\|_F^2 = \min_{\bz} \|\bd - \bv(\bz)\|_F^2, 
+%\end{equation}
+%where $\bz = [\bg,\by]^T$.
 The other vectors in Eq.~\eqref{eq:least_squares_red} were defined in Section~\ref{} \footnote{In this paper we use the same symbol to denote corrupted predicted visibilities and true observed corrupted visibilities as it
-is trivial to distinguish which one is implied from the context.
+%is trivial to distinguish which one is implied from the context.
 Using the same symbol for these two quantities improves the readability of the paper.}.
 Note the similarity between Eq.~\eqref{eq:least_squares} and Eq.~\eqref{eq:least_squares_red},
-i.e. redundant calibration can be cast as a non-linear least squares problem.
+%i.e. redundant calibration can be cast as a non-linear least squares problem.
 
@@
 \end{equation}
-where $\breve{\bz} = [\bz,\conj{\bz}]^T$, $\breve{\bd} = [\bd,\conj{\bd}]^T$, $\breve{\bv} = [\bv,\conj{\bv}]$.
-Eq.~\eqref{eq:least_squares_complex} is the augmented counterpart of Eq.~\eqref{eq:least_squares_red}.
+where $\breve{\bz} = [\bz^T,\conj{\bz}^T]^T$. Eq.~\eqref{eq:least_squares_complex} is the complex conjugate augmented counterpart of Eq.~\eqref{eq:least_squares_red}.
 If we use the Wirtinger derivatives, Eq.~\eqref{eq:wir}, instead of the classic notion of a derivative to define the gradient operator then 
@@
 \end{equation}
-The LM update rule is very similar, the major difference being a damping factor $\lambda$ is introduced:
+The LM update rule is very similar, the major difference being a damping factor $\lambda_{\textrm{LM}}$ is introduced:
 \begin{equation}
 \label{eq:LM_update}
-\Delta \breve{\bz} = (\bJ^H\bJ + \lambda\bD)^{-1}\bJ^H\breve{\br},
+\Delta \breve{\bz} = (\bJ^H\bJ + \lambda_{\textrm{LM}}\bD)^{-1}\bJ^H\breve{\br},
 \end{equation}

### 2017-01-03 paper.tex
fixing section 6

@@
 \section{Results}
+In this section we compare the computational complexity of the ADI \citep{Marthi2014} and PCG \citep{Liu2010} methods. In our comparison we focus on the most 
+expensive operation of the two algorithms, which is the inverse of the Hessian (or modified Hessian in the case of LM).
+We restrict ourselves to comparing the most expensive operation of the two methods, namely the computing the inverse which is the inverse of the Hessian (or modified Hessian in the case of LM), as
+it becomes trivial to rate these two algorithms relative to each other.
+
+outperforms the PCG method in this regard it will certainly outperform it on other
 

### 2017-01-03 paper.tex
still fixing section 6

--- 2017-01-03 paper.tex
@@
 \section{Results}
+In this section we compare the computational complexity of the ADI \citep{Marthi2014} and PCG \citep{Liu2010} methods. We restrict ourselves to comparing the most 
+expensive operation of the two methods namely, computing the inverse of the Hessian (or modified Hessian in the case of LM). It becomes trivial to rate these two 
+algorithms relative to each other once the computational complexity of the inverse operation is known.
+
+In Section~\ref{sec:adi} we re-derive the ADI method presented in \citet{Marthi2014}. In \citet{Marthi2014} the ADI method is derived by taking the derivative of the 
+objective function $\Lambda$ relative to $\bg$ and $\by$ and setting the result to naught. When derived in this way, its relation to the recent works \citep{Salvini2014,Smirnov2015} are unclear.
+We use a similar approach to the diagonal approximate method proposed in \citet{Smirnov2015} to show that the ADI algorithm presented in \citet{Marthi2014} is in fact redundant 
+StEfCal.
 
@@
 \subsection{Alternating Direction Implicit Method}
+\label{sec:adi}
 

### 2017-01-03 paper.tex
fixing section 6

@@
 expensive operation of the two methods namely, computing the inverse of the Hessian (or modified Hessian in the case of LM). It becomes trivial to rate these two 
-algorithms relative to each other once the computational complexity of the inverse operation is known.
+algorithms relative to each other once the computational complexity of the inverse operation is known. The simulations we performed to compare these two 
+methods are summarized in Section~\ref{sec:simulations}.
 
@@
 objective function $\Lambda$ relative to $\bg$ and $\by$ and setting the result to naught. When derived in this way, its relation to the recent works \citep{Salvini2014,Smirnov2015} are unclear.
-We use a similar approach to the diagonal approximate method proposed in \citet{Smirnov2015} to show that the ADI algorithm presented in \citet{Marthi2014} is in fact redundant 
-StEfCal.
+We use the diagonal approximate method proposed in \citet{Smirnov2015} to show that the ADI algorithm presented in \citet{Marthi2014} is in actual fact redundant 
+StEfCal \citep{Salvini2014}. Since we approximate the Hessian in the ADI method with its diagonal the computational complexity of its inverse becomes $\mathbb{O}(P)$\footnote{The cost of constructing 
+the diagonal is not included in this estimate.}.
+
+\citet{Liu2010} proposed the PCG method to compute the inverse of the Hessian, however the computational complexity of this approach was not analyzed therein. In 
+Section~\ref{sec:pcg}, we do a proper computational cost analysis of PCG. We find that the computational complexity of PCG is bounded by $\mathbb{O}(P^2)$, which is 
+a lot better than straightforward inversion which takes $\mathbb{O}(P^3)$.
+
+Even though the PCG method is computationally more expensive than the ADI method we can not arrive at a final conclusion at this point. With the ADI method 
+we only approximate the inverse of the Hessian, while in the case of the PCG method we compute the exact inverse. So the question arises whether the PCG method 
+will converge much faster than the ADI method even though each iteration (which includes an inversion) is more expensive. We investigate this question in Section~\ref{sec:comparison}.
 
 \subsection{Simulations}
+\label{sec:simulations}
 The main purpose of the simulations conducted in this paper is to genrate experimental results with which we can compare the computational complexity of ADI and PCG.
@@
 \subsection{Conjugate Gradient Method}
+\label{sec:pcg}
 
@@
 \subsection{Comparison}
+\label{sec:comparison}
 itr numbers plot and graph of $kP = P^{1.7}$.

### 2017-01-04 paper.tex
worked on section 6

@@
 
+In Section~\ref{sec:sbc} we briefly review the ADI method that is used for skymodel-based calibration. In Section~\ref{sec:red} we show how the ADI method used for skymodel-based
+calibration relates to the redundant ADI method presented in \citet{Marthi2014}.
+We also show that the computational complexity of the ADI method to invert the Hessian (assuming it has alredy been constructed)
+is equal to $\mathbb{O}(P)$.
+
 \begin{figure*}
@@
 %\subsection{Diagonalization and StEFCal}
+\subsubsection{Skymodel-based Calibration}
+\label{sec:sbc}
 StEFCal (Statistically Efficient Calibration) is an alternating direction implicit method, in which the measurement equation (cite the equation earlier in paper) is linearized by assuming that the gains are known, but
@@
 is reached.
-By correctly exploiting the symmetry inherent in the calibration problem this alternating approach can be replaced by a single iterative update step [MUST IMPROVE HERE].
+By correctly exploiting the symmetry inherent in the calibration problem this alternating approach can be replaced by a single iterative new parameter estimation step [MUST IMPROVE HERE].
 \cite{Salvini2014} also show that
-the convergence of the algorithm can be improved if for every even iteration the basic update step is modified as follows: the current gain estimate is replaced by the average of the current gain solution and the gain solution of the previous odd iteration.
+the convergence of the algorithm can be improved if for every even iteration the basic new parameter estimation step is modified as follows: the current gain estimate is replaced by the average of the current gain solution and the gain solution of the previous odd iteration.
 StEFCal lowers the computational complexity of normal skymodel-based calibration from O$(P^3)$ to O$(P^2)$.
@@
 $\bH$ and $\bJ^H\breve{\br}$. While inspecting the skymodel-based $\bH$ it becomes apparent that its diagonal entries are much more significant than its off-diagonal entries.
-By approximating $\bH$ by its diagonal and substituting this approximate Hessian back into the GN update step they are able to re-derive the fundamental StEFCal update step. They then show that the 
-averaging update step used on even iterations can be derived by substituting the approximate $H$ into the LM update step, whilst setting $\lambda$ to one.
-
-The natural question now arises: is it possible to derive a redundant calibration variant of the SteFCal algorithm if we follow a similar approach to what \citet{Smirnov2015}
+By approximating $\bH$ by its diagonal and substituting this approximate Hessian back into the GN update step they are able to re-derive the fundamental iterative StEFCal parameter estimation step.
+They then show that the
+averaging parmeter step used on even iterations can be derived by substituting the approximate $H$ into the LM update step, whilst setting $\lambda$ to one.
+
+The StEFCal equivalent of Eq.~\eqref{eq:two_thrids}
+is equal to \citep{Smirnov2015}
+\begin{equation}
+\label{eq:one_half}
+\breve{\bg}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{1}{2}\breve{\bg}_{k}. 
+\end{equation}
+Note that in Eq.~\eqref{eq:one_half} the influence that the estimate of the parameters from the previous iteration has on the current iteration is scaled by a half, while
+in Eq.~\eqref{eq:two_thrids} it is scaled by two thirds. Moreover, the diagonal-approximate iterative StEFCal estimation step is \citep{Smirnov2015}
+\begin{align}
+\breve{\bg}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bg}_k,\label{eq:stef_lambda}\\
+ &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bg}_k, \label{eq:stef_alpha}  
+\end{align}
+where $\widetilde{\bH}$ denotes the diagonal approximation of $\bJ^H\bJ$, $\alpha = \frac{1}{1+\lambda}$. Moreover, $\lambda = \frac{1-\alpha}{\alpha}$.
+
+On the odd iterations $\lambda$ is chosen to be zero, which corresponds to an $\alpha$ equal to one. On the even iteration $\lambda$ is chosen to be equal to one, which corresponds
+to an $\alpha$ equal to a half. Note that, on the odd iteration the influence that the estimate of the parameters of the previous iteration has on the current estimate is scaled
+by a half. Note the similarity between the last terms of Eq.~\eqref{eq:stef_alpha} and Eq.~\eqref{eq:one_half} when $\alpha$ is an half.
+
+\subsubsection{Redundant Calibration}
+\label{sec:red}
+The natural question now arises: is it possible to derive a redundant calibration variant of the StEfCal algorithm if we follow a similar approach to what \citet{Smirnov2015}
 proposed?
@@
 The question now arises what value should we use for $\alpha$? Although an exact mathematical proof for the optimal choice of $\alpha$ still eludes us, it is still possible to
-venture a clever guess by studying what the choice of $\alpha$ is in the case of StEFCal. The StEFCal equivalent of Eq.~\eqref{eq:two_thrids}
-is equal to
-\begin{equation}
-\label{eq:one_half}
-\breve{\bg}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{1}{2}\breve{\bg}_{k}. 
-\end{equation}
-Note that in Eq.~\eqref{eq:one_half} the influence that the estimate of the parameters from the previous iteration has on the current iteration is scaled by a half, while
-in Eq.~\eqref{eq:two_thrids} it is scaled by two thirds. Moreover, the StEFCal equivalent of Eq.~\eqref{eq:lambda} and Eq.~\eqref{eq:alpha} are
-\begin{align}
-\breve{\bg}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bg}_k,\label{eq:stef_lambda}\\
- &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bg}_k, \label{eq:stef_alpha}  
-\end{align}
-
-On the odd iterations $\lambda$ is chosen to be zero, which corresponds to an $\alpha$ equal to one. On the even iteration $\lambda$ is chosen to be equal to one, which corresponds
-to an $\alpha$ equal to a half. Note that, on the odd iteration the influence that the estimate of the parameters of the previous iteration has on the current estimate is scaled
-by a half. Which is exactly the same as in, Eq.~\eqref{eq:one_half}.
+venture a clever guess by studying what the choice of $\alpha$ is in the case of StEFCal.
 

### 2017-01-04 paper.tex
added symbol for modified hessian

@@
 \newcommand{\bmJ}{\bmath{\mathcal{J}}}
+\newcommand{\bmH}{\bmath{\mathcal{H}}}
 \newcommand{\bzero}{\bmath{0}}
@@
 \section{Conjugate Gradient Method}
+\label{sec:conj_grad}
 The Conjugate Gradient (CG) method was independently discovered by \citet{Hestenes1973} and \citet{Stiefel1952}. They later published a joint paper, which is now considered as the seminal
@@
 \subsection{Normal Equation}
+\label{sec:normal}
 Lets assume that instead of Eq.~\eqref{eq:linear_system} we have a more general linear system, i.e.
@@
 \end{equation}
-The LM update rule is very similar, the major difference being a damping factor $\lambda_{\textrm{LM}}$ is introduced:
+The LM update rule is very similar, the major difference being a damping factor $\lambda$ is introduced:
 \begin{equation}
 \label{eq:LM_update}
-\Delta \breve{\bz} = (\bJ^H\bJ + \lambda_{\textrm{LM}}\bD)^{-1}\bJ^H\breve{\br},
-\end{equation}
-where $\bD=\bI\odot\bJ^H\bJ$.
+\Delta \breve{\bz} = (\bJ^H\bJ + \lambda\bD)^{-1}\bJ^H\breve{\br},
+\end{equation}
+where $\bD=\bI\odot\bJ^H\bJ$. We refer to $\bJ^H\bJ$ as the Hessian matrix $\bH$ and to $\bJ^H\bJ + \lambda\bD$ as the modified Hessian matrix $\bmH$.
 
@@
 \end{equation}
-Eq.~\eqref{eq:two_thrids} shows us that in the case of redundant calibration we can calculate the GN update step without calculating the residual. The 
-normal skymodel-based equivalent of Eq.~\eqref{eq:two_thrids} is equal to \citep{Smirnov2015}:
-\begin{equation}
-\label{eq:one_half}
-\breve{\bg}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{1}{2}\breve{\bg}_{k}, 
-\end{equation}
-where $\breve{\bg} = [\bg^T,\conj{\bg}^T]^T$
-
-FIX TRANSPOSE PROBLEM
+Eq.~\eqref{eq:two_thrids} shows us that in the case of redundant calibration we can calculate the GN update step without calculating the residual. 
+% The 
+% normal skymodel-based equivalent of Eq.~\eqref{eq:two_thrids} is equal to \citep{Smirnov2015}:
+% \begin{equation}
+% \label{eq:one_half}
+% \breve{\bg}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{1}{2}\breve{\bg}_{k}, 
+% \end{equation}
+% where $\breve{\bg} = [\bg^T,\conj{\bg}^T]^T$
+% 
+% FIX TRANSPOSE PROBLEM
 
@@
 to an $\alpha$ equal to a half. Note that, on the odd iteration the influence that the estimate of the parameters of the previous iteration has on the current estimate is scaled
-by a half. Note the similarity between the last terms of Eq.~\eqref{eq:stef_alpha} and Eq.~\eqref{eq:one_half} when $\alpha$ is an half.
+by a half. Note the similarity between the last terms of Eq.~\eqref{eq:stef_alpha} and Eq.~\eqref{eq:one_half} when $\alpha$ is a half.
 
@@
 \breve{\bz}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bz}_k,\label{eq:lambda}\\
- &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bz}_k, \label{eq:alpha}  
-\end{align}
-where $\alpha = \frac{1}{1+\lambda}$. Moreover, $\lambda = \frac{1-\alpha}{\alpha}$. The analytic expression of $\bJ^H\breve{\bd}$ will be very similar to the analytic 
+ &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bz}_k. \label{eq:alpha}  
+\end{align}
+The analytic expression of $\bJ^H\breve{\bd}$ will be very similar to the analytic 
 expression of $\bJ^H\breve{\br}$, the only difference being that in Eq.~\eqref{eq:ab} the letter $r$ would be replaced by a $d$. If we substitute the analytic expression
@@
 \end{equation}
-
-The question now arises what value should we use for $\alpha$? Although an exact mathematical proof for the optimal choice of $\alpha$ still eludes us, it is still possible to
-venture a clever guess by studying what the choice of $\alpha$ is in the case of StEFCal.
-
-If we wish to scale $\breve{\bz}$ by two thirds in Eq.~\eqref{eq:alpha} we should choose $\alpha$ equal to a third, which corresponds to a $\lambda$ equal to two.
-
-At this point there is no way of knowing if this is mere coincidence. DEF REWRITE THIS.
-
-REWRTIE THIS IN A DIFFERENT WAY. START WITH CHENGULAR AND JUST MENTION THE LINK TO STEFCAL'S CHOICE OF ALPHA.
-
-The update rules in Eq.~\eqref{eq:lambda} and Eq.~\eqref{eq:alpha} have been derived before, by Chengular... The approach they followed is discussed in the Appendix.
-
-Interestingly enough, if we substitute $\alpha = 1$ into Eq.~\ref{eq:g_update} then it reduces to the skymodel-based StEFCal update rule. In Chengular they propose 
-an $\alpha = 0.3$.
-
-In an attempt to merge the terminology that has independently develop in the general calibration and redundant calibration literature and to emphasize the close
-relationship between Stefcal and the approach derived by I will use the label Redunandat SteFCal (R-Stefcal) to refer to the NLS method proposed by ...
-
-****\\
-
-with $\alpha = \frac{1}{1+\lambda}$. Moreover $\lambda = \frac{1-\alpha}{\alpha}$. So in Marthi and Chengular (2014) they use an $\alpha$ of 0.3, which corresponds to an LM damping factor $\lambda$
-of $2\frac{1}{3}$.
-
-NB:: NEED TO ADD SNR PLOTS HERE
-PLUS IMAGE OF HESSIAN FOR THE DISCUSSION
-
-\subsection{Conjugate Gradient Method}
+The computational complexity of inverting $\widetilde{\bH}$ is $\mathbb{O}(P)$.
+
+Eq.~\eqref{eq:g_update} and Eq.~\eqref{eq:y_update} are the exact same iterative new parameter estimators derived by \citet{Marthi2014}. Als note that Eq.~\eqref{eq:g_update} is the fundamental
+iterative StEfCal new parameter estimator. We have therefore successfully used the diagonal approximation approach suggested by \citet{Smirnov2015} to re-derive the ADI method presented in 
+\citet{Marthi2014} and in doing so we have shown that the algorithm in \citet{Marthi2014} is actually redundant StEfCal. We could also have shown this by using a similar approach 
+to the one presented in \citet{Salvini2014}.
+
+The question now arises what value should we use for $\alpha$? In \citet{Marthi2014} they propose 
+$\alpha = 0.3$. Although an exact mathematical proof for the optimal choice of $\alpha$ still eludes us, it is still possible to
+venture a clever guess by studying what the choice of $\alpha$ is in the case of skymodel-based StEFCal. Recall from Section~\ref{sec:sbc} that $\alpha$ is equal to a half 
+in the case of StEFCal.
+If we wish to mimic this behaviour for the redundant calibration use case we need to scale $\breve{\bz}$ by two thirds in Eq.~\eqref{eq:alpha} (see Eq.~\eqref{eq:two_thrids}) which implies that we should choose $\alpha$ equal to a third, which corresponds to a $\lambda$ equal to two,
+which is the value we used in this paper.
+
+% At this point there is no way of knowing if this is mere coincidence. DEF REWRITE THIS.
+% 
+% REWRTIE THIS IN A DIFFERENT WAY. START WITH CHENGULAR AND JUST MENTION THE LINK TO STEFCAL'S CHOICE OF ALPHA.
+% 
+% The update rules in Eq.~\eqref{eq:lambda} and Eq.~\eqref{eq:alpha} have been derived before, by Chengular... The approach they followed is discussed in the Appendix.
+% 
+% Interestingly enough, if we substitute $\alpha = 1$ into Eq.~\ref{eq:g_update} then it reduces to the skymodel-based StEFCal update rule. In Chengular they propose 
+% an $\alpha = 0.3$.
+% 
+% In an attempt to merge the terminology that has independently develop in the general calibration and redundant calibration literature and to emphasize the close
+% relationship between Stefcal and the approach derived by I will use the label Redunandat SteFCal (R-Stefcal) to refer to the NLS method proposed by ... 
+% 
+% ****\\
+% 
+% with $\alpha = \frac{1}{1+\lambda}$. Moreover $\lambda = \frac{1-\alpha}{\alpha}$. So in Marthi and Chengular (2014) they use an $\alpha$ of 0.3, which corresponds to an LM damping factor $\lambda$
+% of $2\frac{1}{3}$.
+% 
+% NB:: NEED TO ADD SNR PLOTS HERE
+% PLUS IMAGE OF HESSIAN FOR THE DISCUSSION
+
+\subsection{Preconditioned Conjugate Gradient Method}
 \label{sec:pcg}
+In this section we investigate the computational complexity of the PCG method when applied to the redundant calibration problem. 
+The PCG method was presented in Section~\ref{sec:conj_grad}.
+The PCG method differs from the ADI method presented in Section~\ref{sec:adi} in that in the case of PCG we compute the exact inverse of the modified Hessian instead 
+of using some sort of approximation as is the case for ADI. As we mentioned in Section~\ref{sec:pcg} there
+are two factors which influence the computational complexity of the PCG method, namely the spectral condition number and the sparsity of the matrix we wish to invert. 
+We used a $\lambda$ value of two, implying we used Eq.~\eqref{eq:LM_update} instead of Eq.~\eqref{eq:GN_update} throughout this section.
+We respectively discuss the spectral condition number and the sparsity of the modified Hessian, which is a non-singular matrix in Section~\ref{sec:scn} and Section~\ref{sec:sparsity}.
+It is interesting to note, that the Hessian itself is singular which means that if we are to use Eq.~\eqref{eq:GN_update} instead we would need to be able invert a singular matrix.
+In Section~\ref{sec:normal},
+we showed that the PCG method can in fact be used to invert a singular Hessian matrix.
 
 \subsubsection{Spectral Condition Number}
+\label{sec:scn}
 kappa + itr plots
@@
 \caption{The graphs presented here were generated using the second experimental setup in Table~\ref{tab:ch_parm}.
-In the left panel we have plotted the spectral condition number $\kappa$ of the Hessian $\bH$, before (magenta and red curves) and after preconditioning (blue curve), at different SNR values as a function of the
+In the left panel we have plotted the spectral condition number $\kappa$ of the modified Hessian $\bH$, before (magenta and red curves) and after preconditioning (blue curve), at different SNR values as a function of the
 number of antennas $N$ in the array. This plot shows us that we can drastically reduce the condition number of $\bH$ by preconditioning. Moreover, $\kappa$ is also
@@
 \subsubsection{Sparsity}
-
+\label{sec:sparsity}
 % \begin{enumerate}

### 2017-01-04 paper.tex
newest version

@@
 \subsection{Preconditioning}
+\label{sec:precon}
 Preconditioning is a technique which allows us to improve the condition number of a matrix. Let $\bM$ be a positive-definite Hermitian matrix.
@@
 The PCG method was presented in Section~\ref{sec:conj_grad}.
-The PCG method differs from the ADI method presented in Section~\ref{sec:adi} in that in the case of PCG we compute the exact inverse of the modified Hessian instead 
+The PCG method differs from the ADI method presented in Section~\ref{sec:adi} in that in the case of PCG we compute the exact inverse of the modified Hessian $\bmH$ instead 
 of using some sort of approximation as is the case for ADI. As we mentioned in Section~\ref{sec:pcg} there
-are two factors which influence the computational complexity of the PCG method, namely the spectral condition number and the sparsity of the matrix we wish to invert. 
+are two factors which influence the computational complexity of the PCG method, namely the spectral condition number and the sparsity of the matrix we wish to invert (see Eq.~\eqref{eq:cg_bound}). 
 We used a $\lambda$ value of two, implying we used Eq.~\eqref{eq:LM_update} instead of Eq.~\eqref{eq:GN_update} throughout this section.
-We respectively discuss the spectral condition number and the sparsity of the modified Hessian, which is a non-singular matrix in Section~\ref{sec:scn} and Section~\ref{sec:sparsity}.
-It is interesting to note, that the Hessian itself is singular which means that if we are to use Eq.~\eqref{eq:GN_update} instead we would need to be able invert a singular matrix.
+We respectively discuss the spectral condition number and the sparsity of the modified Hessian $\bmH$, which is a non-singular matrix, in Section~\ref{sec:scn} and Section~\ref{sec:sparsity}.
+It is interesting to note, that the Hessian $\bH$ itself is singular which means that if we are to use Eq.~\eqref{eq:GN_update} instead we would need to be able to invert a singular matrix.
 In Section~\ref{sec:normal},
-we showed that the PCG method can in fact be used to invert a singular Hessian matrix.
+we showed that the PCG method can in fact be used to invert a singular Hessian $\bH$.
 
@@
 \label{sec:scn}
-kappa + itr plots
+In Section~\ref{sec:precon} we mentioned that the convergence properties of CG can be dramatically improved through preconditioning.
+Inspecting Fig.~\ref{fig:hessian} reveals that the Jacobian preconditioner
+is a good candidate for the redundant use case. In Fig.~\ref{fig:kappa} we see that this choice of preconditioner performs very well, the value of the spectral condition number $\kappa$
+becomes constant and independent of problem size, i.e. the number of antennas in the array. Preconditioning effectively eliminates it from Eq.~\eqref{eq:cg_bound}. This result
+is confirmed by Fig.~\ref{fig:cg_itr}.
+
+%kappa + itr plots
 \begin{figure*}
@@
 \caption{The graphs presented here were generated using the second experimental setup in Table~\ref{tab:ch_parm}.
-In the left panel we have plotted the spectral condition number $\kappa$ of the modified Hessian $\bH$, before (magenta and red curves) and after preconditioning (blue curve), at different SNR values as a function of the
-number of antennas $N$ in the array. This plot shows us that we can drastically reduce the condition number of $\bH$ by preconditioning. Moreover, $\kappa$ is also
+In the left panel we have plotted the spectral condition number $\kappa$ of the modified Hessian $\bmH$, before (magenta and red curves) and after preconditioning (blue curve), at different SNR values as a function of the
+number of antennas $N$ in the array. This plot shows us that we can drastically reduce the condition number of $\bmH$ by preconditioning. Moreover, $\kappa$ is also
 independent of the problem size (the number of antennas in the array). The fact that $\kappa$ is small and independent of problem size (effectively constant) implies that it simply drops out 
@@
 In the right hand panel we have plotted the number of major iterations (line 6 of Algorithm \ref{algo:CG}) required by the conjugate
-gradient method to invert $\bH$, before (magenta and red curves) and after preconditioning (blue and green curves), at different SNR values as a function of the number of antennas $N$ in the array.  
-The right hand panel therefore shows us that the number of major iterations needed by the preconditioned conjugate gradient method to invert $\bH$ is independent of the number of antennas in the array
-and is much smaller than the dimension of $\bH$. 
+gradient method to invert $\bmH$, before (magenta and red curves) and after preconditioning (blue and green curves), at different SNR values as a function of the number of antennas $N$ in the array.  
+The right hand panel therefore shows us that the number of major iterations needed by the preconditioned conjugate gradient method to invert $\bmH$ is independent of the number of antennas in the array.
+The number of major iterations are also much smaller than the dimension of $\bmH$. 
 \label{fig:kappa_itr}} 
@@
 
-As alluded to in section ..., the second factor that determines the computational complexity of the conjugate gradient method when used to perform redundant calibration is the amount of non-zero
-entries in the Hessian matrix $\bH$, which is the matrix we are trying to invert. We have also shown in section that the first factor, the spectral condition number of $\bH$, can be reduced 
-to a constant by employing preconditioning and can therefore be completely eliminated from the computational complexity equation. A useful matrix quantity
-which we will use in this section is the sparsity factor $\gamma$ of a matrix, which is the ratio between the amount of zero entries and the total amount of entries in a matrix. We 
+The second factor that determines the computational complexity of the conjugate gradient method when used to perform redundant calibration is the amount of non-zero
+entries in the modiefied Hessian matrix $\bmH$. 
+%We have also shown in section that the first factor, the spectral condition number of $\bH$, can be reduced 
+%to a constant by employing preconditioning and can therefore be completely eliminated from the computational complexity equation. 
+A useful matrix quantity which we will use in this section is the sparsity factor $\gamma$ of a matrix, which is the ratio between the amount of zero entries and the total amount of entries in a matrix.
+We
 can therefore define $\gamma$ as follows:
@@
 \end{equation}
-It is trivial to determine an analytic expression for the sparsity ratio of $\bH$ when our redundant array is in a regular east-west geometric configuration
-The sparsity and the asymptotic sparsity ratio of $\boldsymbol{H}$ are equal to
+It is trivial to determine an analytic expression for the sparsity ratio of $\bmH$ when our redundant array is in a regular east-west geometric configuration
+The sparsity and the asymptotic sparsity ratio of $\bmH$ are equal to
 \begin{align}
@@
 \end{align}
-However when our redundant array makes use of a more complicated geometric layout it becomes much more difficult to construct such expressions as $\bH$ itself 
+However when our redundant array makes use of a more complicated geometric layout it becomes much more difficult to construct such expressions as $\bmH$ itself 
 becomes much harder to express analytically.

### 2017-01-04 paper.tex
added comparison section

@@
 
-In Fig... we plot the sparsity ratio of three different geometric layouts as a function of the number antennas in the array. Discuss the solid lines 
-and how they relate to the two equations from above.
-
-We can now use the sparsity factor to compute the order of the computational cost of inverting $\bH$ with
+In Fig.~\ref{fig:gamma} we plot the sparsity ratio for three different geometric layouts as a function of the number antennas in the array. Although, we can not explicitly calculate 
+the limit of the hexagonal and square layouts, Fig.~\ref{fig:gamma} seems to indicate that these limits do exist.
+
+We can now use the sparsity factor to compute the order of the computational complexity of inverting $\bmH$ with
 \begin{equation}
@@
 \end{align}
-Asymptotically the computational cost of inverting $\bH$ is $O(P^2)$. However, the computational cost converges 
-very slowly to its asymptotic value, and in general is equal to $O(P^{c})$, where $c \sim 1.7$ if a Hexagonal geometric layout is employed.
-
-sparsity derivation
-sparsity + computational complexity plots
+In Fig.~\ref{fig:c} we plot the value of $c$ for three different geometric layouts as a function of the number antennas in the array.
+It is clear from Eq.~\eqref{eq:c} and Fig.~\ref{fig:c} that the computational complexity, under the assumption that $\lim_{N\rightarrow \infty} \gamma$ exists, is asymptotically bounded by $O(P^2)$.
+However, the computational complexity converges
+very slowly to its asymptotic value, and in general is equal to $O(P^{c})$, where $c \sim 1.7$ if a Hexagonal geometric layout containing less than 200 antennas is employed.
+
 \begin{figure*}
 \centering
-\subfigure[$\gamma$]{\includegraphics[width=0.47\textwidth]{./sparsity.pdf}\label{fig:kappa}}
+\subfigure[$\gamma$]{\includegraphics[width=0.47\textwidth]{./sparsity.pdf}\label{fig:gamma}}
 % V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
-\subfigure[$c$]{\includegraphics[width=0.46\textwidth]{./comp_order.pdf}\label{fig:cg_itr}}
-\caption{In the left panel we plot the sparsity ratio $\gamma$ (the number of zero entries relative to the total number of entries in a matrix) of the Hessian $\bH$ as a function of the number 
+\subfigure[$c$]{\includegraphics[width=0.46\textwidth]{./comp_order.pdf}\label{fig:c}}
+\caption{In the left panel we plot the sparsity ratio $\gamma$ (the number of zero entries relative to the total number of entries in a matrix) of the modified Hessian $\bmH$ as a function of the number
 of antennas $N$ in the array for different geometries, namely an hexagonal grid (blue circles), a square grid (green crosses) and a regular east-west grid (red circles).
@@
 of this expression, Eq.~\eqref{eq:gamma}, as $N\rightarrow \infty$ produces the black dotted line.
-In the right panel we plot the order of the computational cost $c$ for inverting $\bH$ as a function of the number
+In the right panel we plot the order of the computational cost $c$ for inverting $\bmH$ as a function of the number
 of antennas in the array for different geometries. We use the same color scheme in the right and left panels. We compute $c$ using Eq.~\eqref{eq:c}. The red dashed line 
-in the righ panel is associated with the theoretical curve of $c$ we computed for the east-west regular grid. The limit of the red-dashed curve is two, which we 
+in the right panel is associated with the theoretical curve of $c$ we computed for the east-west regular grid. The limit of the red-dashed curve is two, which we 
 have plotted with a black dotted line.
-\label{fig:sup_anti}} 
+\label{fig:sparsity}} 
 \end{figure*}
@@
 \label{sec:comparison}
-itr numbers plot and graph of $kP = P^{1.7}$.
+We are now in a position to make a final comparison between the ADI and PCG methods.
+One inversion using the ADI method is very cheap, while one inverse using PCG is more expensive.
+On the other side of the coin,  one inversion with PCG is much more
+accurate than one inversion with ADI.
+The question is now, if the inverse using the PCG method is much more accurate will the LM implementation using it require much less LM-iterations than its ADI counterpart.
+Moreover, will the number
+of LM iterations PCG require be so few that it manages to beat the ADI approach in any case. To answer this we first need to compute the number of iterations $k$; ADI-LM can make in the 
+same time it takes one single PCG-LM iteration to execute. We can compute this with
+\begin{align}
+\label{eq:k}
+ kP &\approx P^{1.7}, & k &\approx P^{0.7}.
+\end{align}
+Next we need to measure the amount of LM iterations we save by using the PCG method and compare it to Eq.\eqref{eq:k}. We make this comparison 
+in Fig.~\ref{fig:out_diff}. Fig.~\ref{fig:out_diff} tells us that the ADI method will outperform the PCG method as the number of LM iterations we save by taking the 
+full inverse is not enough to exceed $k$.
 
@@
 \centering
-\subfigure[Regular layout]{\includegraphics[width=0.47\textwidth]{./outerloop.pdf}\label{fig:outerloop}}
+\subfigure[Number of LM iterations required by ADI and PCG.]{\includegraphics[width=0.47\textwidth]{./outerloop.pdf}\label{fig:outerloop}}
 % V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
-\subfigure[Hexagonal layout]{\includegraphics[width=0.47\textwidth]{./diff.pdf}\label{fig:diff}}
-\caption{Hallo
+\subfigure[Difference between ADI and PCG.]{\includegraphics[width=0.47\textwidth]{./diff.pdf}\label{fig:diff}}
+\caption{The graphs presented here were generated using the second experimental setup in Table~\ref{tab:ch_parm}. In the left panel we have the number of LM iterations 
+that is required for the LM algorithm using the ADI (green and blue curve) and the PCG (magenta and red curve) methods to converge to a parameter convergence criteria of 1e-6 whilst using different 
+SNR values as a function of number of antennas $N$ in the array.
+In the right panel we have the average amount of LM iterations (difference between the ADI and PCG curves in the left panel) we save (green and blue curve) by computing the full-inverse with the PCG method whilst using different SNR values.
+The black curve is associated with $k$ (see Eq.~\eqref{eq:k}). Clearly we do not save enough iterations using the full-inverse of $\bmH$ for the PCG method to outperform the ADI method.
 \label{fig:out_diff}} 

### 2017-01-05 paper.tex
worked a bit on the introduction

@@
 
+\newcommand{\aaps}{A\&AS}
+\newcommand{\aap}{A\&A}
+\newcommand{\mnras}{MNRAS}
+\newcommand{\nat}{Nature}
+\newcommand{\physrep}{Phys. Rep.}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{abstract}
-Traditionally redundant calibration is implemented via a hierarchical algorithmic strategy, i.e. a basic firstcal is performed, followed by LOGCAL and LINCAL.  We first show
-that LINCAL is effectively the non-linear least-squares gradient-based Gauss-Newton algorithm.
-Therefore, an easy way to reduce the number of layers used in performing redundant calibration is to employ the Levenberg-Marquardt algorithm instead (which would improve the convergence properties of stock standard LINCAL).
-We further show that instead of splitting our calibration problem into its phase and amplitude components,
-which is the approach taken by LINCAL, we can treat the complex unknown model parameters and their complex conjugates as separate unknowns instead. In the latter framework
-we make use of the gradient-based complex Levenberg-Marquardt algorithm (not its real valued counterpart) to determine our fundamental update step. 
-Lastly, we compare two previously proposed methods to speedup redundant 
-calibration, namely the alternating direction implicit method and the preconditioned conjugate gradient method. We found that the alternating 
-direction implicit method outperforms the preconditioned conjugate gradient method, however the preconditioned conjugate gradient method is more robust as it would be easier to adapt to other calibration use cases.
-The reason we exclusively use the complex optimization framework in this paper, instead of the phase-amplitude approach, is that in our opinion the complex framework makes it much easier to derive an analytical tractable update step.
-More importantly however, when we work in this complex framework it becomes trivial to show that the alternating direction implicit method is in actual fact just a
-special case of the more general complex Levenberg-Marquardt algorithm.
+Traditionally redundant calibration is implemented via a hierarchical algorithmic strategy, i.e. a basic \textsc{firstcal} is performed, followed by \textsc{logcal} and \textsc{lincal}.  
+\textsc{logcal} requires the use of the logartihm, while \textsc{lincal} makes use of the gradient-based Gauss-Newton algorithm.
+We show that instead of splitting our calibration problem into its phase and amplitude components,
+which is the approach taken by \textsc{lincal}, we can treat the complex unknown model parameters and their complex conjugates as separate unknowns instead. 
+We further propose the use of the Levenberg-Marquardt algorithm, instead of the Gauss-Newton algorithm, to improve the convergence properties of the proposed complex
+framework, which in turn reduces the number of layers that redundant calibration pipelines require.
+We also compare two previously proposed methods to speedup redundant calibration, namely the alternating direction implicit method and the preconditioned conjugate gradient method.
+We found that the alternating
+direction implicit method outperforms the preconditioned conjugate gradient method, however the preconditioned conjugate gradient method is more robust as it is more straightforward to adapt to other calibration use cases.
+We mainly focus on the complex framework in this paper rather than the amplitude and phase approach as it makes it trivial to relate the alternating direction implicit method 
+and the Levenberg-Marquardt algorithm with each other.
 \end{abstract}
+
+%that \textsc{lincal} is effectively the non-linear least-squares gradient-based Gauss-Newton algorithm.
+Therefore, an easy way to reduce the number of layers used in performing redundant calibration is to employ the Levenberg-Marquardt algorithm instead (which would improve the convergence properties of stock standard \textsc{lincal})
 
@@
 The propogation effects (i.e.
-the ionosphere, antenna gains, etc.) a celestial signal experiences along its propogation path is modelled with the radio interforemetric measurement equation (RIME).
-%\citep{ME1,RRIME1}.
+the ionosphere, antenna gains, etc.) a celestial signal experiences along its propogation path is modelled with the radio interforemetric measurement equation (RIME) \citep{ME1,RRIME1}.
 Calibration is the procedure by which we remove the errors that these propogation effects induce onto our intereferometric data and is normally accomplished using a 
@@
 In the case where we only model the antenna gains, which is the most basic form of the RIME, we can define calibration as finding the antenna gains that minimizes the difference betweem our observed and predicted visibilities.
-There are two recent developments
-in calibration which should be explicitly mentioned, since they form the foundation this paper builds on. The first is the discovery of StEfCal \citep{Salvini2014}, which 
-is an alternating direction implicit (ADI) skymodel-based calibration algorithm. It works by assuming the gains are known while solving for their conjugates. Now we do the reverse,
-assume the conugates are known and solve for the gains. This procedure results in the linearization of the calibration algorithm and reduces the computational complexity 
-of calibration from $\mathbb{O}(N^3)$ to $\mathbb{O}(N^3)$, where $N$ denotes the number of antennas in the array. The second development is the idea; to not break the calibration problem 
-into its real and imaginary parts and then to solve for the real and imaginary parts of the parameters seperately, but to rather treat the antenna gains and their 
-conjugates as seperate parameters to estimate \citep{Smirnov2015}. \citet{smirnov2015} also showed that the StEfCal algorithm itself is merely a special case of this
+
+There are two recent developments in skymodel-based calibration which should be explicitly mentioned at this point, since they form the foundation this paper builds on.
+The first is the discovery of \textsc{StEfCal} \citep{Mitchell:MWA-cal,Salvini2014}, which
+is an alternating direction implicit (ADI) skymodel-based calibration algorithm. It works by assuming the gains are known while solving for their conjugates and then performing the reverse. 
+This procedure results in the linearization of the calibration algorithm and reduces the computational complexity of calibration from $\mathbb{O}(N^3)$ to $\mathbb{O}(N^2)$, where $N$ denotes the number of antennas in the array.
+assume the conugates are known and solve for the gains. The second development is the idea: to not break the calibration problem 
+into its real and imaginary parts and then to solve for the real and imaginary parts of the unknown parameters seperately, but to rather treat the antenna gains and their 
+conjugates as seperate parameters to estimate \citep{Smirnov2015}. \citet{Smirnov2015} also showed that the \textsc{StEfCal} algorithm itself is merely a special case of this
 gain-and-conjugate approach.
 
-Two baselines are redundant if their beaseline difference vector have the same lenght and orientation, which imply that they measure the same $uv$-modes.
+Two baselines are redundant if their beaseline difference vectors have the same lenght and orientation, which imply that they measure the same $uv$-modes.
 When we use a redundant array layout we dramatically
-reduce the number of unkowns. Redundant calibration is the procedure by which we leverage the fact that we have 
-Redundant calibration differs from normal calibration in that it does not require a skymodel. With redundant calibration we do not only solve for the antenna gains, but also 
-for the true observed visibilities. We are able to do so, because we have less unkonws than baseline equations. Since we
-
-[Redundant calibration has been in use for a long time. Noordam used redundant calibration to produce high-dynamic range images of 3C84 with the 
-WSRT array.] RW The first implementations of redundant calibration made use of the logarithm function, because when the logarithm is applied to the scalar measurement equation redundant 
-calibration is transformed into a linear problem cite Wieringa,camps and Lui. This approach is also generally referred to as LOGCAL.
-LOGCAL is reviewed in Appendix ??.
-The next improvement was to perform a firt-order taylor expansion (which is equivalent to the GN algorithm Kurien 2016) of the measurement equation an then to solve for the
-perturbations of the model-parameters.
-This approach is commonly referred to as LINCAL.
-LINCAL is derived from first principals in Appendix .
-Two ways of speeding up redundant calibration has also been
-proposed. The first technique makes use of the alternating direction implicit method, in which we alternate between solving the gains, their conjugates and 
-the true observed visibilities (wijnholds and marthi). Using the conjugate gradient method to exploit the sparsity inherent in the redundant calibration problem, 
-and in doing so speedup the algorithm itself was proposed in . A proper comparison between these two approaches has not been done before.
+reduce the number of unkowns. If our array is redundant enough we reduce the number of unknowns to such an extend that we are able to solve for the unknown 
+observed visibilities themselves in addition to the antenna gains. Redundant calibration is the procedure by which we estimate the unknown observed visibilities in addition
+to the antenna gains.
+
+Redundant calibration was already in use in the 1980's. \citet{Noordam1982} used redundant calibration and the WSRT (Westerbork Synthesis Radio Telescope) to produce high-dynamic range images of 3C84.
+The first implementations of redundant calibration made use of the logarithm function, because when the logarithm is applied to the scalar measurement equation, redundant 
+calibration is transformed into a linear problem \citep{Wieringa1992,Camps2003,Liu2010}. This approach is commenly referred to as \textsc{logcal}.
+We briefly discuss \textsc{logcal} in Appendix~\ref{sec:logcal}.
+The next improvement in redundant calibration was to perform a firt-order taylor expansion, which is equivalent to the Gauss-Newton (GN) algorithm \citep{Kurien2016}, of the measurement equation an then to solve for the
+perturbations of the model-parameters \citep{Liu2010}. The implementation of  this approach is called \textsc{lincal}. \textsc{lincal} is derived from first principals in Appendix~\ref{sec:lincal}. 
+\textsc{logcal} and \textsc{lincal} are both implemented in \textsc{omnical}\footnote{https://github.com/jeffzhen/omnical}, which is a layered redundant calibration pipeline \citep{Zheng2014,Ali2015}.
+The solutions of \textsc{logcal} are used as initial input for \textsc{lincal}, since the GN algorithm requires that its starting parameter guesses be in the vacinity 
+of the solution.
+
+Two ways of speeding up redundant calibration have also been proposed recently. The first technique makes use of the ADI method, in which we alternate between solving the gains, their conjugates and 
+the true observed visibilities \citep{Wijnholds2012,Marthi2014}.
+The second is to use the preconditioned conjugate gradthient (PCG) method.
+PCG allows us to exploit the sparsity inherent in the redundant calibration problem,
+to speed it up \citep{Liu2010}. A proper comparison between these two approaches has not been done before.
 
@@
 
-Layered approach, LOGCAL, LINCAL, Stefan's conference paper, Noordam, other ASTRON paper, newest papers on redundant calibration, marthi paper etc...
+Layered approach, LOGCAL, \textsc{lincal}, Stefan's conference paper, Noordam, other ASTRON paper, newest papers on redundant calibration, marthi paper etc...
 
@@
 % \begin{enumerate}
-%  \item Present a general framework which unifies all the techniques developed so far showing they are all related (LINCAL, non-linear estimator, etc ...). They are all non-linear
-%  least-squares techniques, i.e. they either employ Gauss-Newton or the Levenberg-Marquardt algorithms. We have to start with LINCAL showing that it is GN, then we have to motivate 
+%  \item Present a general framework which unifies all the techniques developed so far showing they are all related (\textsc{lincal}, non-linear estimator, etc ...). They are all non-linear
+%  least-squares techniques, i.e. they either employ Gauss-Newton or the Levenberg-Marquardt algorithms. We have to start with \textsc{lincal} showing that it is GN, then we have to motivate 
 %  why we want to use Oleg's complex optimization framework.
 %  \item Use Oleg's complex optimization framework to re-derive the non-linear technique proposed by Marthi and Chengular. The novelty lies, in the fact that by using Oleg's
-%  framework we can find analytic expressions for the Jacobian, the Hessian and the Jacobian-residual product (which is not even the case for LINCAL). State that the algorithm is 
+%  framework we can find analytic expressions for the Jacobian, the Hessian and the Jacobian-residual product (which is not even the case for \textsc{lincal}). State that the algorithm is 
 %  effectively related to SteFCal and is eff an independent rediscovery of SteFCal and an extension of it into the redundant domain.
@@
 where $\breve{\bd} = [\bd^T,\conj{\bd}^T]^T$, $\breve{\bv} = [\bv^T,\conj{\bv}^T]$ and $\bmath{\varrho}$ is the phase and amplitude parameter vector which we define in Appendix~\ref{sec:lincal}. 
-With the aid of Eq.~\eqref{eq:least_squares_lincal} we can derive the LINCAL algorithm (see Appendix~\ref{sec:lincal}).
+With the aid of Eq.~\eqref{eq:least_squares_lincal} we can derive the \textsc{lincal} algorithm (see Appendix~\ref{sec:lincal}).
 The Jacobian associated with Eq.~\eqref{eq:least_squares_lincal} is equal to Eq.~\eqref{eq:Jac_lin}. 
-At the hand of Appendix~\ref{sec:lincal} it becomes evident that LINCAL is equivalent to a basic GN update step \citep{Kurien2016}. 
-Therefore, a straightforward way of improving the convergence properties of stock standard LINCAL would be to simply add the LM damping factor to the basic LINCAL update step.
+At the hand of Appendix~\ref{sec:lincal} it becomes evident that \textsc{lincal} is equivalent to a basic GN update step \citep{Kurien2016}. 
+Therefore, a straightforward way of improving the convergence properties of stock standard \textsc{lincal} would be to simply add the LM damping factor to the basic \textsc{lincal} update step.
 
@@
 
-\section{LOGCAL}
+\section{\textsc{LOGCAL}}
 \label{sec:logcal}
@@
 
-\section{LINCAL}
+\section{\textsc{lincal}}
 \label{sec:lincal}
-In this section we derive LINCAL \citep{Liu2010} from first principles.
+In this section we derive \textsc{lincal} \citep{Liu2010} from first principles.
 %Assume that $\eta_p,\eta_q,\widetilde{\eta}_{\phi_{pq}},\varphi_p,\varphi_q$ and $\widetilde{\varphi}_{\phi_{pq}}$ are real valued variables.

### 2017-01-05 paper.tex
worked a bit on the introduction
added introduction to paper

@@
 We further propose the use of the Levenberg-Marquardt algorithm, instead of the Gauss-Newton algorithm, to improve the convergence properties of the proposed complex
-framework, which in turn reduces the number of layers that redundant calibration pipelines require.
+framework, which in turn reduces the number of layers that redundant calibration pipelines would require.
 We also compare two previously proposed methods to speedup redundant calibration, namely the alternating direction implicit method and the preconditioned conjugate gradient method.
@@
 
-Redundant calibration was already in use in the 1980's. \citet{Noordam1982} used redundant calibration and the WSRT (Westerbork Synthesis Radio Telescope) to produce high-dynamic range images of 3C84.
+Redundant calibration was already quite popular in the 1980's.
+\citet{Noordam1982} used redundant calibration and the WSRT (Westerbork Synthesis Radio Telescope) to produce high-dynamic range images of 3C84.
 The first implementations of redundant calibration made use of the logarithm function, because when the logarithm is applied to the scalar measurement equation, redundant 
@@
 the true observed visibilities \citep{Wijnholds2012,Marthi2014}.
-The second is to use the preconditioned conjugate gradthient (PCG) method.
+The second is to use the preconditioned conjugate gradient (PCG) method.
 PCG allows us to exploit the sparsity inherent in the redundant calibration problem,
@@
 
-Newly built instruments like LOFAR, ..... has also sparked a renewed interest in redundant calibration. Redundancy is also advantageous for detecting the EoR,
-which explains why some of the previsously mentioned instruments have redundant layouts.
-
-Layered approach, LOGCAL, \textsc{lincal}, Stefan's conference paper, Noordam, other ASTRON paper, newest papers on redundant calibration, marthi paper etc...
-
-EOR SECTION SOMEWHERE HERE
-
-MAIN CONTRIBUTION OF THIS PAPER COMES HERE
+New instruments like LOFAR (Low Frequency Array) \citep{Noorishad2012}, PAPER (Precision Array for Probing the Epoch of Reionization) \citep{Ali2015} and HERA (Hydrogen Epoch of Reionization Array) \citep{deboer2015} have recently sparked a renewed interest in redundant calibration.
+Using redundancy is especially advantageous when we are trying to detect the EoR (Epoch of Reionization) 21 cm signal,
+which is inherently quite faint \citep{Parsons2012}. Since redundant baselines make statistically independent measurements of the same $uv$-modes redundancy can greatly 
+improve the SNR (signal-to-noise ratio) of EoR detection experiments. This explains why PAPER and HERA, are highly redundant arrays, as they are both primarily used to detect the EoR 21 cm signal.
+[GIANNI TO REWRITE AND EXTEND EOR SECTION]
+
+In this paper we propose to use the complex framework introduced by \citet{Smirnov2015} instead of the traditional amplitude and phase approach used by \textsc{lincal}.
+We also propose to use the Levenberg-Marquardt gradient-based algorithm instead of the Gauss-Newton algorithm as it exhibits better convergence properties.
+In doing so the amount of layers that redundant calibration pipelines need are effectively reduced. We also compare two recent redundant calibration speedup algorithms
+with each other, namely ADI \citep{Marthi2014} and PCG \citep{Liu2010}.
+In this paper, we decided to mainly focus on the complex framework proposed by \citet{Smirnov2015} rather than the traditional phase and amplitude approach as it
+allows us to trivially prove that the ADI method proposed by \citep{Marthi2014} is merely a special case of the more general LM algorithm. Moreover, this framework also 
+allows us to show that the ADI method proposed by \citep{Marthi2014} can for all intents and purposes be renamed to redundant \textsc{StEfCal} \citep{Mitchell:MWA-cal,Salvini2014}.
+
+ADD SUMMARY PARAGRAPH WHICH DEPENDS ON PAPER LAYOUT
 

### 2017-01-06 paper.tex
moved paper content around

@@
 
-\subsection{Scalar Measurement Equation}
-The observed visibility $d_{pq}$ measured by the baseline formed by antenna $p$ and $q$ is described by
-\begin{equation}
-\label{eq:vis_definition}
-d_{pq} = g_{p}\conj{g_q}y_{pq} + n_{pq},
-\end{equation}
-where $g_{p}$ and $g_{q}$ are the direction independent gains associated with antenna $p$ and $q$, $y_{pq}$ denotes the true visibility that baseline $pq$ measured
-and $n_{pq}$ is the thermal noise component.
-Eq.~\eqref{eq:vis_definition} is known as the unpolarized measurement equation.
-Conjugation is denoted by $\conj{(*)}$.
-The true value of $g_p$, $g_q$ and $y_{pq}$ are of course unknown and is what we try to estimate when we calibrate
-and image.
-
-The real and imaginary components of the thermal noise is normally distributed with a mean of zero and a standard deviation proportional to   
-\begin{equation}
-\sigma \propto \frac{T_{\textrm{sys}}}{\sqrt{\Delta \nu \tau}}, 
-\end{equation}
-where $T_{\textrm{sys}}$ is equal to the system temperature, $\Delta \nu$ is the observational bandwidth and $\tau$ is the integration time per visibility.
-
-\subsection{Indexing Functions}
-\begin{figure*}
-\centering
-\subfigure[Regular layout]{\includegraphics[width=0.47\textwidth]{./ln.pdf}\label{fig:ln}}
-% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
-\subfigure[Hexagonal layout]{\includegraphics[width=0.47\textwidth]{./pn.pdf}\label{fig:pn}}
-\caption{Hallo
-\label{fig:pl}} 
-\end{figure*}
-
-In general there are two ways of indexing the true visibilities associated with different baselines.
-We can use the composite index $pq$, which is the strategy that is followed in Eq.~\eqref{eq:vis_definition},
-or we can use a single unique baseline index which we assign in numerical order to each composite index; assuming the composite indexes are arranged in standard baseline order.
-The composite indexes $pq$ are in standard baseline order when they are arranged as follow: $12, 13, 1N, 23\cdots$. We use $N$ to denote the number of antennas in the array. Let 
-\begin{equation}
-\alpha_{pq} =
-\begin{cases}
-(q-p) + (p-1)\left (N-\frac{1}{2}p \right ) & \textrm{if}~p<q\\
-%B + (p-q) + (q-1)\left (N-\frac{1}{2}q \right )) & \textrm{if}~p>q\\
-%B+q + \frac{1}{2}(p-1)(p-2) & \textrm{if}~p>q\\
-0 & \textrm{otherwise}
-\end{cases}.
-\end{equation}
-We can use $\alpha_{pq}$ to map composite indexes to unique baseline indexes, i.e. 
-\begin{equation}
-12,13,\cdots,1N,23,\cdots \xrightarrow{\alpha_{pq}} 1,2,3,\cdots,B.
-\end{equation}
-Moreover, $B$ denotes the number of baselines in the array, i.e. $B = \frac{N^2-N}{2}$.
-
-If an array is redundant then its redundant baselines should sample the exact same visibilities in the $uv$-plane, i.e. if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$. The 
-fact that redundant baselines should measure the exact same visibility implies that every baseline does not require a unique baselines index; instead we can use a separate index for each unique redundant baseline.
-Let $\phi_{pq}$ be the function that maps composite antenna pairs to the redundant baseline indexes of an array.
-Note that $\phi_{pq}$ is not symmetric as
-$\phi_{pq} = 0$ if $p>q$. We can also define the following symmetric variant of $\phi_{pq}$:
-\begin{equation}
-\zeta_{ij} = 
-\begin{cases}
-\phi_{pq}~\textrm{if}~p \leq q\\
-\phi_{qp}~\textrm{if}~p>q
-\end{cases}.
-\end{equation} 
-It is possible to construct a simple analytic expression for $\zeta_{pq}$ when our array is in an east-west regular configuration, namely $\zeta_{pq} = |q-p|$. 
-It becomes increasingly more difficult to construct analytic expressions of $\zeta_{pq}$ for other more complicated array layouts. The empirically constructed symmetric geometry
-functions of three different redundant layouts are depicted in Fig.~\ref{fig:geometry_function}. We denote the range of $\zeta_{ij}$ with $\mathcal{R}(\zeta_{ij})$. The maximal element 
-that $\zeta_{ij}$ can ascertain is denoted by $L$ and can be construed as the maximal number of unique redundant baselines which can be formed for a given 
-array layout. Alternatively, we can interpret the symmetric geometry functions 
-in Fig.~\ref{fig:geometry_function} as matrices. In this alternative paradigm, $\zeta_{ij}$ denotes a matrix entry instead. The dimension of these so called geometry 
-matrices are $N\times N$, where $N$ denotes the number of antennas in the array.
-
-\begin{figure*}
-\centering
-\subfigure[Hexagonal layout]
-{\includegraphics[width=0.33\textwidth]{./HEX_lay.pdf}\label{fig:HEX_lay}}
-\subfigure[Square layout]
-{\includegraphics[width=0.33\textwidth]{./SQR_lay.pdf}\label{fig:SQR_lay}}
-\subfigure[Regular east-west layout]
-{\includegraphics[width=0.33\textwidth]{./REG_lay.pdf}\label{fig:REG_lay}}
-
-\subfigure[Hexagonal: $\zeta_{pq}$]
-{\includegraphics[width=0.33\textwidth]{./HEX_phi.pdf}\label{fig:HEX_phi}}
-\subfigure[Square: $\zeta_{pq}$]
-{\includegraphics[width=0.33\textwidth]{./SQR_phi.pdf}\label{fig:SQR_phi}}
-\subfigure[Regular east-west: $\zeta_{pq}$]
-{\includegraphics[width=0.33\textwidth]{./REG_phi.pdf}\label{fig:REG_phi}}
-\caption{In the top panel we have three different redundant antenna layouts, namely a hexagonal (left), square (middle) and a regular east-west (right) layout. 
-In the bottom panel we have the respective symmetric redundancy geometry functions for the layouts on the top panel. We used 91 antennas
-to construct the the hexagonal layout, while a 100 antennas where used in the square and east-west layouts. The maximal amount of redundant baselines that can be formed for 
-the hexagonal, square and east-west layouts in the top panel are 165, 180 and 99 baselines respectively.\label{fig:geometry_function}}
-\end{figure*}
-
-If the function $\phi_{ij}$ is known and one is given one of the two antenna indexes that together form a redundant baseline as well as the redundant baseline index itself then it is possible 
-to calculate the unknown antenna index. This can be computed via the following two expressions:
-\begin{equation}
-\xi_{ij} = 
-\begin{cases}
-p~\textrm{if}~\exists! ~ p \in \mathbb{N} ~ s.t. ~(\phi_{pi} = j)\\
-0~\textrm{otherwise}
-\end{cases},
-\end{equation}
-and
-\begin{equation}
-\psi_{ij} = 
-\begin{cases}
-q~\textrm{if}~\exists! ~ q \in \mathbb{N} ~ s.t. ~(\phi_{iq} = j)\\
-0~\textrm{otherwise}
-\end{cases}
-\end{equation}
-We use $\xi_{ij}$ to retrieve the first antenna index of the composite antenna pair associated with a particular baseline, if the index of the first antenna in the composite antenna pair and the unique redundant baseline index are known, while we use $\psi_{ij}$ to obtain
-a similar result; the second antenna index of the composite antenna pair. \textbf{NB::} I still need to plot $\zeta_{ij}$ for different antenna layouts.
-
-\subsection{Redundant Measurement Equation}
-If the array is redundant and the function $\phi_{ij}$ is known we can rewrite Eq.~\eqref{eq:vis_definition} as
-\begin{equation}
-\label{eq:vis_red}
-d_{pq} = g_{p}\conj{g_q}y_{\phi_{pq}} + n_{pq}.
-\end{equation}
-Eq.~\eqref{eq:vis_red} can also be expressed in the following vector form 
-\begin{equation}
-\label{eq:vis_red_vec}
-\bd = \bv + \bn, 
-\end{equation}
-where 
-\begin{align}
- \left [ \bd \right]_{\alpha_{pq}} &= d_{pq}, & \left [ \bv \right ]_{\alpha_{pq}} &= g_p y_{\phi_{pq}} \conj{g_q}\nonumber\\
- \left [ \bn \right ]_{\alpha_{pq}} &= n_{pq}. &  &\label{eq:vec_definitions}
-\end{align}
-
-NB COLUMN VECTORS NEED TO ADD THIS...
-
-In Eq.~\eqref{eq:vec_definitions}, the size of the three vectors are equal to $B$ (i.e. $p<q$), which denotes the number of baselines in the array. Moreover,
-\begin{equation}
-\alpha_{pq} =
-\begin{cases}
-(q-p) + (p-1)\left (N-\frac{1}{2}p \right ) & \textrm{if}~p<q\\
-B + (p-q) + (q-1)\left (N-\frac{1}{2}q \right )) & \textrm{if}~p>q\\
-%B+q + \frac{1}{2}(p-1)(p-2) & \textrm{if}~p>q\\
-0 & \textrm{otherwise}
-\end{cases}.
-\end{equation}
-The easiest way to construct the vectors in Eq.~\ref{eq:vec_definitions} is to enumerate through the composite indexes $pq$ in the following order $12, 13,\cdots,1N,23,\cdots$.
-For each composite index we then compute the right hand side of the appropriate vector element definition which we then need to assign to the 
-correct vector position.
-The correct position is calculated with $\alpha_{pq}$.
-Using the aforementioned order, however, allows you to fill the vector in numerical order, i.e.
-$12,13,\cdots,1N,23,\cdots \xrightarrow{\alpha_{pq}} 1,2,3,\cdots,B$.
-
-We will be using the following SNR (signal-to-noise ratio) definition in this paper  
-\begin{equation}
-\textrm{SNR} = \frac{<\bv\odot\conj{\bv}>_{t,pq}}{<\bn\odot\conj{\bn}>_{t,pq}}, 
-\end{equation}
-where $<*>_{t,pq}$ denotes averaging over time and baselines and $\odot$ denotes the Hadamard product.
-The definition we use here is based on the SNR definitions used in \citet{Liu2010} and \citet{Marthi2014}.
-
-\section{Conjugate Gradient Method}
+\subsection{Conjugate Gradient Method}
 \label{sec:conj_grad}
@@
 
-\subsection{Computational Complexity}
+\subsubsection{Computational Complexity}
 \citet{Kaniel1966} was able to determine convergence bounds for CG using Chebyshev polynomials. A much more in depth analysis of CG convergence can be found in
@@
 
-\subsection{Preconditioning}
+\subsubsection{Preconditioning}
 \label{sec:precon}
@@
 
-\subsection{Normal Equation}
+\subsubsection{Normal Equation}
 \label{sec:normal}
@@
 
-\section{Least Squares}
+\subsection{Least Squares}
 \label{sec:ls}
@@
 
-\section{Redundant Calibration}
-Redundant calibration can be formulated as a least-squares problem:
+\section{Problem Statement}
+
+\subsection{Scalar Measurement Equation}
+The observed visibility $d_{pq}$ measured by the baseline formed by antenna $p$ and $q$ is described by
+\begin{equation}
+\label{eq:vis_definition}
+d_{pq} = g_{p}\conj{g_q}y_{pq} + n_{pq},
+\end{equation}
+where $g_{p}$ and $g_{q}$ are the direction independent gains associated with antenna $p$ and $q$, $y_{pq}$ denotes the true visibility that baseline $pq$ measured
+and $n_{pq}$ is the thermal noise component.
+Eq.~\eqref{eq:vis_definition} is known as the scalar measurement equation.
+Conjugation is denoted by $\conj{(*)}$.
+The true value of $g_p$, $g_q$ and $y_{pq}$ are unknown.
+We try to estimate the value of
+$g_p$, $g_q$ and $y_{pq}$ by calibrating our data and then imaging the result.
+
+The real and imaginary components of the thermal noise is normally distributed with a mean of zero and a standard deviation proportional to   
+\begin{equation}
+\sigma \propto \frac{T_{\textrm{sys}}}{\sqrt{\Delta \nu \tau}}, 
+\end{equation}
+where $T_{\textrm{sys}}$ is equal to the system temperature, $\Delta \nu$ is the observational bandwidth and $\tau$ is the integration time per visibility.
+
+\subsection{Indexing Functions}
+
+In general there are two ways of indexing the true visibilities associated with different baselines.
+We can use the composite index $pq$, which is the strategy that is followed in Eq.~\eqref{eq:vis_definition},
+or we can use a single unique baseline index which we assign in numerical order to each composite index; assuming the composite indexes are arranged in standard baseline order.
+The composite indexes $pq$ are in standard baseline order when they are arranged as follow: $12, 13, 1N, 23\cdots$. Let 
+\begin{equation}
+\alpha_{pq} =
+\begin{cases}
+(q-p) + (p-1)\left (N-\frac{1}{2}p \right ) & \textrm{if}~p<q\\
+%B + (p-q) + (q-1)\left (N-\frac{1}{2}q \right )) & \textrm{if}~p>q\\
+%B+q + \frac{1}{2}(p-1)(p-2) & \textrm{if}~p>q\\
+0 & \textrm{otherwise}
+\end{cases}.
+\end{equation}
+We can use $\alpha_{pq}$ to map composite indexes to unique baseline indexes, i.e. 
+\begin{equation}
+12,13,\cdots,1N,23,\cdots \xrightarrow{\alpha_{pq}} 1,2,3,\cdots,B.
+\end{equation}
+We denote the number of baselines in the array with $B$, i.e. $B = \frac{N^2-N}{2}$.
+
+Recall that redundant baselines sample the exact same visibilities in the $uv$-plane, i.e. if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$. 
+The aforementioned fact allows us to group all the redundant baselines in a redundant array into the same redundant baseline group. When our array 
+is a redundant array we therefore only require single index values for each redundant baseline group, instead of having a unique single index label for 
+each and every baseline. Let $\phi_{pq}$ be the function that maps composite antenna pairs to the appropriate single redundant baseline group index. Note that $\phi_{pq}$ is not symmetric as 
+$\phi_{pq} = 0$ if $p>q$. We can also define the following symmetric variant of $\phi_{pq}$:
+\begin{equation}
+\zeta_{pq} = 
+\begin{cases}
+\phi_{pq}~\textrm{if}~p \leq q\\
+\phi_{qp}~\textrm{if}~p>q
+\end{cases}.
+\end{equation} 
+It is possible to construct a simple analytic expression for $\zeta_{pq}$ when our array is in an east-west regular configuration, namely $\zeta_{pq} = |q-p|$. 
+It becomes increasingly more difficult to construct analytic expressions of $\zeta_{pq}$ for other more complicated array layouts. 
+We will refer to $\zeta_{pq}$ as the symmetric geometric function. The empirically constructed symmetric geometry
+functions of three different redundant layouts are depicted in Fig.~\ref{fig:geometry_function}. We denote the range of $\zeta_{pq}$ with $\mathcal{R}(\zeta_{pq})$. The maximal element 
+that $\zeta_{pq}$ can ascertain is denoted by $L$ and can be construed as the maximal number of unique redundant baseline groups which can be formed for a given 
+array layout.
+Alternatively, we can interpret the symmetric geometry functions in Fig.~\ref{fig:geometry_function} as geometry matrices.
+In this alternative paradigm, $\zeta_{pq}$ denotes a matrix entry instead.
+The dimension of these geometry
+matrices are $N\times N$.
+
+\begin{figure*}
+\centering
+\subfigure[Hexagonal layout]
+{\includegraphics[width=0.33\textwidth]{./HEX_lay.pdf}\label{fig:HEX_lay}}
+\subfigure[Square layout]
+{\includegraphics[width=0.33\textwidth]{./SQR_lay.pdf}\label{fig:SQR_lay}}
+\subfigure[Regular east-west layout]
+{\includegraphics[width=0.33\textwidth]{./REG_lay.pdf}\label{fig:REG_lay}}
+
+\subfigure[Hexagonal: $\zeta_{pq}$]
+{\includegraphics[width=0.33\textwidth]{./HEX_phi.pdf}\label{fig:HEX_phi}}
+\subfigure[Square: $\zeta_{pq}$]
+{\includegraphics[width=0.33\textwidth]{./SQR_phi.pdf}\label{fig:SQR_phi}}
+\subfigure[Regular east-west: $\zeta_{pq}$]
+{\includegraphics[width=0.33\textwidth]{./REG_phi.pdf}\label{fig:REG_phi}}
+\caption{In the top panel we have three different redundant antenna layouts, namely a hexagonal (left), square (middle) and a regular east-west (right) layout. 
+In the bottom panel we have the respective symmetric redundancy geometry functions for the layouts on the top panel. We used 91 antennas
+to construct the the hexagonal layout, while a 100 antennas where used in the square and east-west layouts. The maximal amount of redundant baselines that can be formed for 
+the hexagonal, square and east-west layouts in the top panel are 165, 180 and 99 baselines respectively.\label{fig:geometry_function}}
+\end{figure*}
+
+If the function $\phi_{pq}$ is known and one is given one of the two antenna indexes that together form a redundant baseline as well as the redundant baseline group index itself then it is possible 
+to calculate the unknown antenna index. This can be computed via the following two expressions:
+\begin{equation}
+\xi_{ij} = 
+\begin{cases}
+p~\textrm{if}~\exists! ~ p \in \mathbb{N} ~ s.t. ~(\phi_{pi} = j)\\
+0~\textrm{otherwise}
+\end{cases},
+\end{equation}
+and
+\begin{equation}
+\psi_{ij} = 
+\begin{cases}
+q~\textrm{if}~\exists! ~ q \in \mathbb{N} ~ s.t. ~(\phi_{iq} = j)\\
+0~\textrm{otherwise}
+\end{cases}
+\end{equation}
+We use $\xi_{ij}$ to retrieve the first antenna index of the composite antenna pair associated with a particular baseline, if the index of the first antenna in the composite antenna pair and the unique redundant baseline group index are known, while we use $\psi_{ij}$ to obtain
+a similar result: the second antenna index of the composite antenna pair. %\textbf{NB::} I still need to plot $\zeta_{ij}$ for different antenna layouts.
+
+\subsection{Redundant Measurement Equation}
+If the array is redundant and the function $\phi_{pq}$ is known we can rewrite Eq.~\eqref{eq:vis_definition} as
+\begin{equation}
+\label{eq:vis_red}
+d_{pq} = g_{p}\conj{g_q}y_{\phi_{pq}} + n_{pq}.
+\end{equation}
+Eq.~\eqref{eq:vis_red} can also be expressed in the following vector form 
+\begin{equation}
+\label{eq:vis_red_vec}
+\bd = \bv + \bn, 
+\end{equation}
+where 
+\begin{align}
+ \left [ \bd \right]_{\alpha_{pq}} &= d_{pq}, & \left [ \bv \right ]_{\alpha_{pq}} &= g_p y_{\phi_{pq}} \conj{g_q},\nonumber\\
+ \left [ \bn \right ]_{\alpha_{pq}} &= n_{pq}. &  &\label{eq:vec_definitions}
+\end{align}
+The vectors in Eq.~\eqref{eq:vec_definitions} are column vectors of size $B$ (i.e. $p<q$).
+
+We have now defined enough notation to enable us to formulate redundant calibration as a least-squares problem:
 \begin{equation}
@@
 \end{equation}
-where $\bz = [\bg^T,\by^T]^T$\footnote{In this paper we use the same symbol to denote corrupted predicted visibilities and true observed corrupted visibilities as it
-is trivial to distinguish which one is implied from the context.}.
-As discussed in Section~\ref{sec:ls}, we can solve the above minimization problem in one of two ways.
-We can split the problem into its real and
+where
+\begin{align}
+ \by &= [y_1,\cdots,y_L]^T, & \bg &=[g_1,\cdots,g_N]^T,\nonumber\\
+ \bz &= [\bg^T,\by^T]^T. &  &\label{eq:parm_definitions}
+ \end{align}
+The number of model parameters that we need to solve for in Eq.~\eqref{eq:least_squares_red} is equal to $P = 2(N+L)$, since redundant calibration is a complex problem.
+We plot the number of unique baseline groups $L$ and the number of parameters we need to estimate $P$ for two different geometric layouts in Fig~\ref{fig:pl}.
+The analytic expressions for the red and the blue curve in Fig.~\ref{fig:ln} are respectively equal to
+\begin{equation}
+L = 2N-\frac{1}{2}\sqrt{12N-3}-\frac{1}{2}, 
+\end{equation}
+and
+\begin{equation}
+L = N - 1. 
+\end{equation}
+
+\begin{figure*}
+\centering
+\subfigure[$L$.]{\includegraphics[width=0.47\textwidth]{./ln.pdf}\label{fig:ln}}
+% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
+\subfigure[$P$.]{\includegraphics[width=0.47\textwidth]{./pn.pdf}\label{fig:pn}}
+\caption{In the left panel we plot the number of unique baseline groups $L$ one can form when using a hexagonal (red curve) and regular east-west layout (blue curve) as 
+a function of the number of antennas $N$ in the array.
+In the right panel we plot the total number of parameters we need to estimate when employing redundant calibration for a hexagonal (red curve) and regular east-west layout (blue curve) as
+a function of the number of antennas $N$ in the array. 
+\label{fig:pl}} 
+\end{figure*}
+
+\subsection{Signal-to-noise Ratio}
+The following SNR (signal-to-noise ratio) definition is used in this paper  
+\begin{equation}
+\label{}
+\textrm{SNR} = \frac{<\bv\odot\conj{\bv}>_{\nu,t,pq}}{<\bn\odot\conj{\bn}>_{\nu,t,pq}}, 
+\end{equation}
+where $<*>_{\nu,t,pq}$ denotes averaging over frequency, time and baselines and $\odot$ denotes the Hadamard product.
+The definition we use here is based on the SNR definitions used in \citet{Liu2010} and \citet{Marthi2014}.
+
+HERE I CAN WRITE SOMETHING ABOUT THE LIMITATIONS OF THE PAPER.
+
+\section{Redundant Calibration}
+As discussed in Section~\ref{sec:ls}, we can solve Eq.~\eqref{eq:least_squares_red} in one of two ways. We can split the problem into its real and
 imaginary parts and then solve for the real and imaginary parts of the parameters seperately or we can employ the notion of Wirtinger derivatives. The former 
-approach is discussed in detail in Section~\ref{sec:ri}, while the latter approach is described in Section~\ref{sec:w}.
+approach is discussed in Section~\ref{sec:ri}, while the latter approach is described in detail in Section~\ref{sec:w}.
 

### 2017-01-06 paper.tex
added some more structure to paper

@@
 \section{Preliminaries}
+There are two major pre-existing algorithms we use throughout this paper, namely the Conjugate Gradient (CG) method and least-squares parameter estimation.
+We respectively discuss the CG method and least-squares in Section~\ref{sec:conj_grad} and Section~\ref{sec:ls}.
 
@@
 \label{sec:conj_grad}
-The Conjugate Gradient (CG) method was independently discovered by \citet{Hestenes1973} and \citet{Stiefel1952}. They later published a joint paper, which is now considered as the seminal
+The CG method was independently discovered by \citet{Hestenes1973} and \citet{Stiefel1952}. They later published a joint paper, which is now considered as the seminal
 reference on CG \citep{Hestenes1952}. The iterative CG method is used to solve a particular class of linear system. The CG method is generally used to solve
@@
 \caption{Conjugate Gradient Method.
-Inputs: $\bA$, $\bb$, a starting value $x$, maximum number of iterations $k_{\textrm{max}}$ and an error tolerance $\epsilon<1$ Output: $\bx$ the solution to Eq.~\eqref{eq:linear_system}.
+Inputs: $\bA$, $\bb$, a starting value for $\bx$, maximum number of iterations $k_{\textrm{max}}$ and an error tolerance $\epsilon<1$ Output: $\bx$ the solution to Eq.~\eqref{eq:linear_system}.
 \citep{Shewchuk1994}.}\label{algo:CG}
@@
 \end{equation}
-It should now also be clear to the reader why the CG method is especially applicable to sparse linear systems, since in those cases $m < P^2$.
+It should now also be clear to the reader why the CG method is especially applicable to sparse linear systems. In sparse linear systems $m < P^2$.
 
@@
 of CG (see Eq.~\eqref{eq:cg_bound}).
-The best possible choice of $\bM$ would therefore be $\bA$, which would lead to $\kappa$ (as the largest and smallest eigenvalue of the Identity matrix is equal to one) becoming unity.
+The best possible choice of $\bM$ would therefore be $\bA$, which would lead to $\kappa$ becoming unity (as the largest and smallest eigenvalue of the Identity matrix is equal to one).
 If $\kappa$ is one then Eq.~\eqref{eq:cg_bound} simplifies and becomes $\mathbb{O}(m)$.
@@
 \end{equation}
-where $\bI$ is the identity matrix. Since the Jacobian preconditioner is a diagonal matrix its inverse can be obtained trivially.
+where $\bI$ is the identity matrix and $\odot$ denotes the Hadamard product. Since the Jacobian preconditioner is a diagonal matrix its inverse can be obtained trivially.
 
@@
 \label{eq:normal_equation}
-\bB^H\bB\bx = \bB^H\bb 
+\bB^H\bB\bx = \bB^H\bb, 
 \end{equation}
@@
 \end{equation}
+In this paper, we denote the Hermitian transpose with $(*)^H$ and the Frobenius norm with $\|*\|_F$.
+
 There are a few important observations that we can make while inspecting Eq.~\eqref{eq:normal_equation}:
@@
 \end{enumerate}
+In this paper, we denote the column range of a matrix $\bA$ with $\mathcal{R}(\bA)$.
 
@@
 square matrix as long as $\bb$ is in the column range of $\bA$ \citep{Lu2015}.
-
-In light of \citet{Lu2015}, the two observations in the numbered list we presented earlier in this section imply that Eq.~\eqref{eq:normal_equation}
+In light of \citet{Lu2015}, the two observations in the numbered list we presented in the previous paragraph imply that Eq.~\eqref{eq:normal_equation}
 can be solved by employing the CG method. The solution returned by CG is however only unique if $\bB^H\bB$ is positive-definite (invertible). The solution
@@
 \end{equation}
-where $\Lambda$ is known as the objective function, $\brho$ is the residual vector, $\bdelta$ is the observed data vector, $\bnu$ is the model data vector and $\bzeta$ is the model parameter vector that minimizes Eq.~\eqref{eq:least_squares}.
+where $\Lambda$ is known as the objective function, $\brho$ is the residual vector, $\bdelta$ is the observed data vector, $\bnu$ is the model-predicted data vector and $\bzeta$ is the model-parameter vector that minimizes Eq.~\eqref{eq:least_squares}.
 In Eq.~\eqref{eq:least_squares}, we assume that the optimization problem is real valued.
-The main objective of least squares is to find the model parameters which minimizes the squared difference between the observed data and the model predicted data. Moreover, $\|*\|_F$ denotes the 
-Frobenius norm.
+The main objective of least squares is to find the model-parameters which minimizes the squared difference between the observed data and the model-predicted data.
 
@@
 \begin{equation}
- \Delta \bzeta = (\bJ^T\bJ+\lambda_{\textrm{LM}}\bD)^{-1} \bJ^T\brho,
-\end{equation}
-where $\bJ$ is the Jacobian matrix (i.e. the derivative of the model $\bnu$ with respect to the parameter vector $\bzeta$), $\lambda_{\textrm{LM}}$ denotes the 
-damping factor, while $\bD$ is a diagonal matrix whose diagonal entries are equal to the diagonal entries of $\bJ^T\bJ$. These update rules are used in an iterative manner to 
-determine new estimates of $\bzeta$.
+ \Delta \bzeta = (\bJ^T\bJ+\lambda\bD)^{-1} \bJ^T\brho,
+\end{equation}
+where $\bJ$ is the Jacobian matrix (i.e. the derivative of the model $\bnu$ with respect to the parameter vector $\bzeta$), $\lambda$ denotes the 
+damping factor, while $\bD$ is a diagonal matrix whose diagonal entries are equal to the diagonal entries of $\bJ^T\bJ$. In this paper we denote 
+matrix transposition with $(*)^T$.
+These update steps can be used in an iterative manner to determine new estimates of $\bzeta$.
 
@@
 many complex functions are not analytic (no Taylor expansion around a fiducial point exists) and therefore not differentiable if the classic notion of differentiation is used. For instance if the 
-classical definition of differentiation is used then $\frac{\partial z}{\partial \conj{z}}$, where $z \in \mathbb{C}$, does not exist.
+classical definition of differentiation is used then $\frac{\partial \conj{z}}{\partial z}$, where $z \in \mathbb{C}$, does not exist.
 
@@
 \section{Problem Statement}
+In Section~\ref{sec:s_mes}, we present the scalar measurement equation. Next, we define some indexing functions in Section~\ref{sec:indexing}. The measurement 
+equation and the indexing functions from Section~\ref{sec:indexing} are then used in Section~\ref{sec:r_mes} to phrase redundant calibration as a least-squares problem, i.e.
+Eq.~\eqref{eq:least_squares_red}.
+We present the Signal-to-noise ratio definition we used in this paper in Section\ref{sec:snr}. The general scope and limitations of this paper is discussed in Section~\ref{sec:scope}.
 
 \subsection{Scalar Measurement Equation}
+\label{sec:s_mes}
 The observed visibility $d_{pq}$ measured by the baseline formed by antenna $p$ and $q$ is described by
@@
 \subsection{Indexing Functions}
-
+\label{sec:indexing}
 In general there are two ways of indexing the true visibilities associated with different baselines.
@@
 \subsection{Redundant Measurement Equation}
+\label{sec:r_mes}
 If the array is redundant and the function $\phi_{pq}$ is known we can rewrite Eq.~\eqref{eq:vis_definition} as
@@
 The number of model parameters that we need to solve for in Eq.~\eqref{eq:least_squares_red} is equal to $P = 2(N+L)$, since redundant calibration is a complex problem.
-We plot the number of unique baseline groups $L$ and the number of parameters we need to estimate $P$ for two different geometric layouts in Fig~\ref{fig:pl}.
+We plot $L$ and $P$ for two different geometric layouts in Fig~\ref{fig:pl}.
 The analytic expressions for the red and the blue curve in Fig.~\ref{fig:ln} are respectively equal to
@@
 \centering
-\subfigure[$L$.]{\includegraphics[width=0.47\textwidth]{./ln.pdf}\label{fig:ln}}
+\subfigure[$L$]{\includegraphics[width=0.47\textwidth]{./ln.pdf}\label{fig:ln}}
 % V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
-\subfigure[$P$.]{\includegraphics[width=0.47\textwidth]{./pn.pdf}\label{fig:pn}}
+\subfigure[$P$]{\includegraphics[width=0.47\textwidth]{./pn.pdf}\label{fig:pn}}
 \caption{In the left panel we plot the number of unique baseline groups $L$ one can form when using a hexagonal (red curve) and regular east-west layout (blue curve) as 
 a function of the number of antennas $N$ in the array.
-In the right panel we plot the total number of parameters we need to estimate when employing redundant calibration for a hexagonal (red curve) and regular east-west layout (blue curve) as
+In the right panel we plot the total number of parameters $P$ we need to estimate when employing redundant calibration for a hexagonal (red curve) and regular east-west layout (blue curve) as
 a function of the number of antennas $N$ in the array. 
@@
 \subsection{Signal-to-noise Ratio}
+\label{sec:snr}
 The following SNR (signal-to-noise ratio) definition is used in this paper  
@@
 
-HERE I CAN WRITE SOMETHING ABOUT THE LIMITATIONS OF THE PAPER.
+\subsection{Paper Scope}
+\label{sec:scope}
+We only deal with per-timeslot and frequency channel calibration. No solution smoothing  
+is applied between the different timeslots and channels. The reader interested in this is referred to \citep{Zheng2014}. 
+Moreover, we do not explicitly deal with any of the degeneracies normally associated with redundant calibration \citep{Zheng2014,Kurien2016}.
+Similar to the approach taken in \citep{Marthi2014}, we assume that the degeneracies present in the redundant calibration solutions are removed with a final round of skymodel-based calibration.
 
@@
 
+\section{Conclusion}
+Hallo
+
 % fiducial initial guess 

### 2017-01-06 paper.tex
structure improved even more

@@
 where $\bJ$ is the Jacobian matrix (i.e. the derivative of the model $\bnu$ with respect to the parameter vector $\bzeta$), $\lambda$ denotes the 
-damping factor, while $\bD$ is a diagonal matrix whose diagonal entries are equal to the diagonal entries of $\bJ^T\bJ$. In this paper we denote 
-matrix transposition with $(*)^T$.
+damping factor, while $\bD$ is a diagonal matrix whose diagonal entries are equal to the diagonal entries of $\bJ^T\bJ$. In this paper we respectively denote 
+matrix transposition and matrix inversion with $(*)^T$ and $(*)^{-1}$.
 These update steps can be used in an iterative manner to determine new estimates of $\bzeta$.
@@
 Eq.~\eqref{eq:least_squares_red}.
-We present the Signal-to-noise ratio definition we used in this paper in Section\ref{sec:snr}. The general scope and limitations of this paper is discussed in Section~\ref{sec:scope}.
+We present the Signal-to-noise ratio (SNR) definition we used in this paper in Section\ref{sec:snr}. The general scope and limitations of this paper is discussed in Section~\ref{sec:scope}.
 
@@
 The true value of $g_p$, $g_q$ and $y_{pq}$ are unknown.
-We try to estimate the value of
+We normally try to estimate the true value of
 $g_p$, $g_q$ and $y_{pq}$ by calibrating our data and then imaging the result.
@@
 The aforementioned fact allows us to group all the redundant baselines in a redundant array into the same redundant baseline group. When our array 
-is a redundant array we therefore only require single index values for each redundant baseline group, instead of having a unique single index label for 
-each and every baseline. Let $\phi_{pq}$ be the function that maps composite antenna pairs to the appropriate single redundant baseline group index. Note that $\phi_{pq}$ is not symmetric as 
+is a redundant array we therefore only require single index values for each redundant baseline group, instead of a unique single index label for 
+each baseline. Let $\phi_{pq}$ be the function that maps composite antenna pairs to the appropriate single redundant baseline group index. Note that $\phi_{pq}$ is not symmetric as 
 $\phi_{pq} = 0$ if $p>q$. We can also define the following symmetric variant of $\phi_{pq}$:
@@
 \begin{align}
- \by &= [y_1,\cdots,y_L]^T, & \bg &=[g_1,\cdots,g_N]^T,\nonumber\\
+ \bg &=[g_1,\cdots,g_N]^T, & \by &= [y_1,\cdots,y_L]^T,\nonumber\\
  \bz &= [\bg^T,\by^T]^T. &  &\label{eq:parm_definitions}
@@
 \label{sec:snr}
-The following SNR (signal-to-noise ratio) definition is used in this paper  
+The following SNR definition is used in this paper  
 \begin{equation}

### 2017-01-09 paper.tex
Improved new section 5

@@
 \section{Results}
-In this section we compare the computational complexity of the ADI \citep{Marthi2014} and PCG \citep{Liu2010} methods. We restrict ourselves to comparing the most 
-expensive operation of the two methods namely, computing the inverse of the Hessian (or modified Hessian in the case of LM). It becomes trivial to rate these two 
+In this section we compare the computational complexity of two newly proposed speedup algorithms for redundant calibration, namely the ADI \citep{Marthi2014} and PCG \citep{Liu2010} methods.
+We restrict ourselves to comparing the most
+expensive operation of the two methods namely, computing the inverse of the Hessian $\bH$ (or modified Hessian $\bmH$ in the case of LM). It becomes trivial to rate these two 
 algorithms relative to each other once the computational complexity of the inverse operation is known. The simulations we performed to compare these two 
@@
 
-In Section~\ref{sec:adi} we re-derive the ADI method presented in \citet{Marthi2014}. In \citet{Marthi2014} the ADI method is derived by taking the derivative of the 
-objective function $\Lambda$ relative to $\bg$ and $\by$ and setting the result to naught. When derived in this way, its relation to the recent works \citep{Salvini2014,Smirnov2015} are unclear.
-We use the diagonal approximate method proposed in \citet{Smirnov2015} to show that the ADI algorithm presented in \citet{Marthi2014} is in actual fact redundant 
-StEfCal \citep{Salvini2014}. Since we approximate the Hessian in the ADI method with its diagonal the computational complexity of its inverse becomes $\mathbb{O}(P)$\footnote{The cost of constructing 
-the diagonal is not included in this estimate.}.
-
-\citet{Liu2010} proposed the PCG method to compute the inverse of the Hessian, however the computational complexity of this approach was not analyzed therein. In 
-Section~\ref{sec:pcg}, we do a proper computational cost analysis of PCG. We find that the computational complexity of PCG is bounded by $\mathbb{O}(P^2)$, which is 
+The basic difference between these two algorithms can be understood by inspecting Fig.~\ref{fig:hessian}. The first observation we can make by inspecting 
+Fig.~\ref{fig:hessian_reg} and Fig.~\ref{fig:hessian_hex} is that the diagonal entries of the Hessians are more significant than the off-diagonal entries.
+It therefore makes sense to approximate the Hessian (or the modified Hessian) with its diagonal, which greatly simplifies the computational complexity of inverting 
+the Hessian, i.e. it becomes $\mathbb{O}(P)$\footnote{The cost of constructing 
+the diagonal is not included in this estimate.} \citep{Smirnov2015}. This diagonal-approximation approach is effectively ADI \citep{Marthi2014}. The exact details of this 
+approach is discussed in Section~\ref{sec:adi}. We will also show how \citet{Marthi2014}, \citet{Smirnov2015} and \citet{Salvini2014} relate to each other in that same section.
+
+The second observation we can make from Fig.~\ref{fig:hessian_reg} and Fig.~\ref{fig:hessian_hex} is that the Hessians contain many zero entries, i.e.
+they are sparse matrices.
+This, in conjunction with the more significant diagonal entries observation, implies that the PCG method (presented in Section~\ref{sec:conj_grad})
+can be applied to invert them in a computationally efficient manner.
+\citet{Liu2010} proposed the PCG method to compute the inverse of the Hessian, however the computational complexity of this approach was not analyzed therein.
+In Section~\ref{sec:pcg}, we do a proper computational complexity analysis of PCG. We find that the computational complexity of PCG is asymptotically bounded by $\mathbb{O}(P^2)$, which is 
 a lot better than straightforward inversion which takes $\mathbb{O}(P^3)$.
+
+\begin{figure*}
+\centering
+\subfigure[Regular layout]{\includegraphics[width=0.47\textwidth]{./reg_hessian.pdf}\label{fig:hessian_reg}}
+% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
+\subfigure[Hexagonal layout]{\includegraphics[width=0.47\textwidth]{./hex_hessian.pdf}\label{fig:hessian_hex}}
+\caption{The number of analytic terms out of which the entries of the Hessian $\bH$ consists for two different geometric layouts, namely a regular east-west grid with $N = 5$ (left panel) and 
+a hexagonal grid with $N = 7$. The diagonal entries of these two Hessians are clearly more significant than their off-diagonal entries. Moreover, these two Hessians, also 
+contain many zero-entries. Note that the location of the zero-entries are dependent on the geometry of the array layout.
+\label{fig:hessian}} 
+\end{figure*}
+
+%Since we approximate the Hessian in the ADI method with its diagonal the computational complexity of its inverse becomes $\mathbb{O}(P)$\footnote{The cost of constructing 
+%the diagonal is not included in this estimate.}.
 
@@
 \label{sec:simulations}
-The main purpose of the simulations conducted in this paper is to genrate experimental results with which we can compare the computational complexity of ADI and PCG.
+%The main purpose of the simulations conducted in this paper is to genrate experimental results with which we can compare the computational complexity of ADI and PCG.
 We used an hexagonal geometric layout with an adjustable number of rings during the
-simulations.
+%simulations. 
 For the skymodel we generated 100 sources, each source had a random flux value and position.
 The flux of the sources followed a Pareto distribution (with a shape parameter of 2), while their positions were uniformly
-distributed in a 3 by 3 degree square grid around the field center. The remaining fixed simulation parameters can all be found in Table~\ref{tab:fixed_parm}. 
+distributed in a 3$^{\circ}$ by 3$^{\circ}$ square grid around the field center. The remaining fixed simulation parameters can all be found in Table~\ref{tab:fixed_parm}. 
 We also employed three different gain error models which we summarize in Table~\ref{tab:gain_parm}. We found that it did not matter which gain error model we used during
@@
 
-In Section~\ref{sec:sbc} we briefly review the ADI method that is used for skymodel-based calibration. In Section~\ref{sec:red} we show how the ADI method used for skymodel-based
-calibration relates to the redundant ADI method presented in \citet{Marthi2014}.
-We also show that the computational complexity of the ADI method to invert the Hessian (assuming it has alredy been constructed)
-is equal to $\mathbb{O}(P)$.
-
-\begin{figure*}
-\centering
-\subfigure[Regular layout]{\includegraphics[width=0.47\textwidth]{./reg_hessian.pdf}\label{fig:hessian_reg}}
-% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
-\subfigure[Hexagonal layout]{\includegraphics[width=0.47\textwidth]{./hex_hessian.pdf}\label{fig:hessian_hex}}
-\caption{Hallo
-\label{fig:hessian}} 
-\end{figure*}
+In Section~\ref{sec:sbc} we briefly review the ADI method that is used for skymodel-based calibration. 
+%In Section~\ref{sec:red} we show how the ADI method used for skymodel-based calibration relates to the redundant ADI method presented in \citet{Marthi2014}. 
+In Section~\ref{sec:red}, we re-derive the ADI method presented in \citet{Marthi2014}. In \citet{Marthi2014} the ADI method is derived by taking the derivative of the 
+objective function $\Lambda$ relative to $\bg$ and $\by$ and setting the result to naught. When derived in this way, its relation to the recent works \citep{Salvini2014,Smirnov2015} are unclear.
+In Section~\ref{sec:red}, we use the diagonal-approximate approach proposed in \citet{Smirnov2015} to show that the ADI algorithm presented in \citet{Marthi2014} is in actual fact redundant 
+\textsc{StEfCal} \citep{Salvini2014}.
 
@@
 \label{sec:sbc}
-StEFCal (Statistically Efficient Calibration) is an alternating direction implicit method, in which the measurement equation (cite the equation earlier in paper) is linearized by assuming that the gains are known, but
-that their conjugates are unknown (REMEMBER TO CITE ALL THE STEFCAL PAPERS MITCHELL ETC...).
+\textsc{StEFCal} is an alternating direction implicit method, in which the measurement equation, Eq.~\eqref{eq:vis_definition}, is linearized by assuming that the gains are known, but
+that their conjugates are unknown \citep{Mitchell:MWA-cal,Salvini2014}.
 Under this assumption, the conjugates of the gains can be obtained trivially.
@@
 is reached.
-By correctly exploiting the symmetry inherent in the calibration problem this alternating approach can be replaced by a single iterative new parameter estimation step [MUST IMPROVE HERE].
+By correctly exploiting the symmetry inherent in the calibration problem this alternating approach can be replaced by a single iterative new parameter estimation step.
 \cite{Salvini2014} also show that
-the convergence of the algorithm can be improved if for every even iteration the basic new parameter estimation step is modified as follows: the current gain estimate is replaced by the average of the current gain solution and the gain solution of the previous odd iteration.
-StEFCal lowers the computational complexity of normal skymodel-based calibration from O$(P^3)$ to O$(P^2)$.
-
-\citet{Smirnov2015} recently showed that one can also derive StEFCal by viewing normal skymodel-based calibration as a complex optimization problem.
+the convergence of the algorithm can be improved if for every even iteration the basic new parameter estimation step is modified as follows: the current gain estimate is replaced by the average of the current gain estimate and the gain estimate of the previous odd iteration.
+\textsc{StEFCal} lowers the computational complexity of normal skymodel-based calibration from $\mathbb{O}(N^3)$ to $\mathbb{O}(N^2)$.
+
+\citet{Smirnov2015} recently showed that one can also derive \textsc{StEFCal} by viewing normal skymodel-based calibration as a complex optimization problem.
 They first compute the skymodel-based calibration equivalent analytic expressions of $\bJ$,

### 2017-01-09 paper.tex
added new info to the paper

@@
 \section{Results}
+\label{sec:results}
 In this section we compare the computational complexity of two newly proposed speedup algorithms for redundant calibration, namely the ADI \citep{Marthi2014} and PCG \citep{Liu2010} methods.
@@
 the diagonal is not included in this estimate.} \citep{Smirnov2015}. This diagonal-approximation approach is effectively ADI \citep{Marthi2014}. The exact details of this 
-approach is discussed in Section~\ref{sec:adi}. We will also show how \citet{Marthi2014}, \citet{Smirnov2015} and \citet{Salvini2014} relate to each other in that same section.
+approach is discussed in Section~\ref{sec:adi}.
+We will also show how \citet{Marthi2014}, \citet{Smirnov2015} and \citet{Salvini2014} relate to each other in that same section.
+In Appendix~\ref{sec:sbc} we briefly summarize the ADI method that is used for skymodel-based calibration.
 
@@
 
-In Section~\ref{sec:sbc} we briefly review the ADI method that is used for skymodel-based calibration. 
-%In Section~\ref{sec:red} we show how the ADI method used for skymodel-based calibration relates to the redundant ADI method presented in \citet{Marthi2014}. 
-In Section~\ref{sec:red}, we re-derive the ADI method presented in \citet{Marthi2014}. In \citet{Marthi2014} the ADI method is derived by taking the derivative of the 
+%In Section~\ref{sec:adi} we show how the ADI method used for skymodel-based calibration relates to the redundant ADI method presented in \citet{Marthi2014}. 
+In this section we re-derive the ADI method presented in \citet{Marthi2014}. In \citet{Marthi2014} the ADI method is derived by taking the derivative of the 
 objective function $\Lambda$ relative to $\bg$ and $\by$ and setting the result to naught. When derived in this way, its relation to the recent works \citep{Salvini2014,Smirnov2015} are unclear.
-In Section~\ref{sec:red}, we use the diagonal-approximate approach proposed in \citet{Smirnov2015} to show that the ADI algorithm presented in \citet{Marthi2014} is in actual fact redundant 
-\textsc{StEfCal} \citep{Salvini2014}.
+We will use the diagonal-approximate approach proposed in \citet{Smirnov2015} to show that the ADI algorithm presented in \citet{Marthi2014} is in actual fact redundant 
+\textsc{StEfCal} \citep{Salvini2014}. Perusing through Appendix~\ref{sec:sbc} is advised before continuing this section.
+
+As we mention in Section~\ref{sec:results}, inspecting Fig.~\ref{fig:hessian} (also see Eq.~\eqref{eq:red_H}) reveals that in the case of redundant calibration the Hessian's diagonal entries are more significant than its off-diagonal entries.
+If we substitute $\bJ^H\bJ$ with $\widetilde{\bH}$ (by its diagonal approximation) and replace $\breve{\br}$ with $\breve{\bd} - \breve{\bv}$ in Eq.~\eqref{eq:GN_update} we obtain
+\begin{equation}
+\label{eq:appr_GN}
+ \Delta \breve{\bz} \approx \widetilde{\bH}^{-1}\bJ^H(\breve{\bd}-\breve{\bv})
+\end{equation}
+Utilizing the second identity in Eq.~\eqref{eq:identities} allows us to simplify Eq.~\eqref{eq:appr_GN} to
+\begin{equation}
+  \Delta \breve{\bz} \approx \widetilde{\bH}^{-1}\bJ^H\breve{\bd}-\breve{\bz}.
+\end{equation}
+It now trivially follows that
+\begin{equation}
+ \breve{\bz}_{k+1} \approx \widetilde{\bH}^{-1}\bJ^H\breve{\bd}.
+\end{equation}
+We obtain a similar result if we repeat the above procedure, only in this case we use Eq.~\eqref{eq:LM_update} instead,
+namely
+\begin{align}
+\breve{\bz}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bz}_k,\label{eq:lambda}\\
+ &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bz}_k, \label{eq:alpha}  
+\end{align}
+where $\alpha = \frac{1}{1+\lambda}$. Moreover, $\lambda = \frac{1-\alpha}{\alpha}$.
+The analytic expression of $\bJ^H\breve{\bd}$ will be very similar to the analytic 
+expression of $\bJ^H\breve{\br}$, the only difference being that in Eq.~\eqref{eq:ab} the letter $r$ would be replaced by a $d$. If we substitute the analytic expression
+of $\bJ^H\breve{\bd}$ and $\widetilde{\bH}^{-1}$ (which is easily computed as the matrix is a diagonal matrix) into Eq.~\eqref{eq:alpha} we obtain the following two update rules:
+\begin{equation}
+\label{eq:g_update}
+g_{i}^{k+1} = \alpha \frac{\sum_{j\neq i} g_j^k \widetilde{y}_{ij}^{~\!\!k} d_{ij}}{\sum_{j\neq i} |g_j^k|^2|y_{\zeta_{ij}}^k|^2} + (1-\alpha) g_i^k, 
+\end{equation}
+and
+\begin{equation}
+\label{eq:y_update}
+y_{i}^{k+1} = \alpha \frac{\sum_{pq \in \mathcal{PQ}_i} \conj{g}_p^k g_q^k d_{pq}}{\sum_{pq \in \mathcal{PQ}_i}|g_p^k|^2|g_q^k|^2} + (1-\alpha) y_i^k. 
+\end{equation}
+The computational complexity of inverting $\widetilde{\bH}$ is $\mathbb{O}(P)$.
+
+Eq.~\eqref{eq:g_update} and Eq.~\eqref{eq:y_update} are the exact same iterative new parameter estimators derived by \citet{Marthi2014}. Als note that Eq.~\eqref{eq:g_update} is the fundamental
+gain estimator associated with StEfCal. We have therefore successfully used the diagonal approximation approach suggested by \citet{Smirnov2015} to re-derive the ADI method presented in 
+\citet{Marthi2014} and in doing so we have shown that the algorithm in \citet{Marthi2014} is actually redundant StEfCal. We could also have shown this by using a similar approach 
+to the one presented in \citet{Salvini2014}.
+
+The question now arises what value should we use for $\alpha$? In \citet{Marthi2014} they propose 
+$\alpha = 0.3$. Although an exact mathematical proof for the optimal choice of $\alpha$ still eludes us, it is still possible to
+venture a clever guess by studying what the choice of $\alpha$ is in the case of skymodel-based StEFCal. Recall from Section~\ref{sec:sbc} that $\alpha$ is equal to a half 
+in the case of StEFCal (for even iterations).
+If we wish to mimic this behaviour for the redundant calibration use case we need to scale $\breve{\bz}$ by two thirds in Eq.~\eqref{eq:alpha} (see Eq.~\eqref{eq:two_thrids}) which implies that we should choose $\alpha$ equal to a third, which corresponds to a $\lambda$ equal to two.
+We used $\lambda$ equal to two in this paper.
 
@@
 \end{figure}
-
-%\subsection{Diagonalization and StEFCal}
-\subsubsection{Skymodel-based Calibration}
-\label{sec:sbc}
-\textsc{StEFCal} is an alternating direction implicit method, in which the measurement equation, Eq.~\eqref{eq:vis_definition}, is linearized by assuming that the gains are known, but
-that their conjugates are unknown \citep{Mitchell:MWA-cal,Salvini2014}.
-Under this assumption, the conjugates of the gains can be obtained trivially.
-Once the conjugates of the gains have been computed, the gains can be solved for.
-One can then alternate between computing the gains and their conjugates until convergence
-is reached.
-By correctly exploiting the symmetry inherent in the calibration problem this alternating approach can be replaced by a single iterative new parameter estimation step.
-\cite{Salvini2014} also show that
-the convergence of the algorithm can be improved if for every even iteration the basic new parameter estimation step is modified as follows: the current gain estimate is replaced by the average of the current gain estimate and the gain estimate of the previous odd iteration.
-\textsc{StEFCal} lowers the computational complexity of normal skymodel-based calibration from $\mathbb{O}(N^3)$ to $\mathbb{O}(N^2)$.
-
-\citet{Smirnov2015} recently showed that one can also derive \textsc{StEFCal} by viewing normal skymodel-based calibration as a complex optimization problem.
-They first compute the skymodel-based calibration equivalent analytic expressions of $\bJ$,
-$\bH$ and $\bJ^H\breve{\br}$. While inspecting the skymodel-based $\bH$ it becomes apparent that its diagonal entries are much more significant than its off-diagonal entries.
-By approximating $\bH$ by its diagonal and substituting this approximate Hessian back into the GN update step they are able to re-derive the fundamental iterative StEFCal parameter estimation step.
-They then show that the
-averaging parmeter step used on even iterations can be derived by substituting the approximate $H$ into the LM update step, whilst setting $\lambda$ to one.
-
-The StEFCal equivalent of Eq.~\eqref{eq:two_thrids}
-is equal to \citep{Smirnov2015}
-\begin{equation}
-\label{eq:one_half}
-\breve{\bg}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{1}{2}\breve{\bg}_{k}. 
-\end{equation}
-Note that in Eq.~\eqref{eq:one_half} the influence that the estimate of the parameters from the previous iteration has on the current iteration is scaled by a half, while
-in Eq.~\eqref{eq:two_thrids} it is scaled by two thirds. Moreover, the diagonal-approximate iterative StEFCal estimation step is \citep{Smirnov2015}
-\begin{align}
-\breve{\bg}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bg}_k,\label{eq:stef_lambda}\\
- &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bg}_k, \label{eq:stef_alpha}  
-\end{align}
-where $\widetilde{\bH}$ denotes the diagonal approximation of $\bJ^H\bJ$, $\alpha = \frac{1}{1+\lambda}$. Moreover, $\lambda = \frac{1-\alpha}{\alpha}$.
-
-On the odd iterations $\lambda$ is chosen to be zero, which corresponds to an $\alpha$ equal to one. On the even iteration $\lambda$ is chosen to be equal to one, which corresponds
-to an $\alpha$ equal to a half. Note that, on the odd iteration the influence that the estimate of the parameters of the previous iteration has on the current estimate is scaled
-by a half. Note the similarity between the last terms of Eq.~\eqref{eq:stef_alpha} and Eq.~\eqref{eq:one_half} when $\alpha$ is a half.
-
-\subsubsection{Redundant Calibration}
-\label{sec:red}
-The natural question now arises: is it possible to derive a redundant calibration variant of the StEfCal algorithm if we follow a similar approach to what \citet{Smirnov2015}
-proposed?
-
-Inspecting Eq.~\eqref{eq:red_H} reveals that in the case of redundant calibration the Hessian's diagonal entries are also more significant than its off-diagonal entries.
-If we substitute $\bJ^H\bJ$ with $\widetilde{\bH}$ (by its diagonal approximation) and replace $\breve{\br}$ with $\breve{\bd} - \breve{\bv}$ in Eq.~\eqref{eq:GN_update} we obtain
-\begin{equation}
-\label{eq:appr_GN}
- \Delta \breve{\bz} \approx \widetilde{\bH}^{-1}\bJ^H(\breve{\bd}-\breve{\bv})
-\end{equation}
-Utilizing the second identity in Eq.~\eqref{eq:identities} allows us to simplify Eq.~\eqref{eq:appr_GN} to
-\begin{equation}
-  \Delta \breve{\bz} \approx \widetilde{\bH}^{-1}\bJ^H\breve{\bd}-\breve{\bz}.
-\end{equation}
-It now trivially follows that
-\begin{equation}
- \breve{\bz}_{k+1} \approx \widetilde{\bH}^{-1}\bJ^H\breve{\bd}.
-\end{equation}
-We obtain a similar result if we repeat the above procedure, only in this case we use Eq.~\eqref{eq:LM_update} instead,
-namely
-\begin{align}
-\breve{\bz}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bz}_k,\label{eq:lambda}\\
- &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bz}_k. \label{eq:alpha}  
-\end{align}
-The analytic expression of $\bJ^H\breve{\bd}$ will be very similar to the analytic 
-expression of $\bJ^H\breve{\br}$, the only difference being that in Eq.~\eqref{eq:ab} the letter $r$ would be replaced by a $d$. If we substitute the analytic expression
-of $\bJ^H\breve{\bd}$ and $\widetilde{\bH}^{-1}$ (which is easily computed as the matrix is a diagonal matrix) into Eq.~\eqref{eq:alpha} we obtain the following two update rules:
-\begin{equation}
-\label{eq:g_update}
-g_{i}^{k+1} = \alpha \frac{\sum_{j\neq i} g_j^k \widetilde{y}_{ij}^{~\!\!k} d_{ij}}{\sum_{j\neq i} |g_j^k|^2|y_{\zeta_{ij}}^k|^2} + (1-\alpha) g_i^k, 
-\end{equation}
-and
-\begin{equation}
-\label{eq:y_update}
-y_{i}^{k+1} = \alpha \frac{\sum_{pq \in \mathcal{PQ}_i} \conj{g}_p^k g_q^k d_{pq}}{\sum_{pq \in \mathcal{PQ}_i}|g_p^k|^2|g_q^k|^2} + (1-\alpha) y_i^k. 
-\end{equation}
-The computational complexity of inverting $\widetilde{\bH}$ is $\mathbb{O}(P)$.
-
-Eq.~\eqref{eq:g_update} and Eq.~\eqref{eq:y_update} are the exact same iterative new parameter estimators derived by \citet{Marthi2014}. Als note that Eq.~\eqref{eq:g_update} is the fundamental
-iterative StEfCal new parameter estimator. We have therefore successfully used the diagonal approximation approach suggested by \citet{Smirnov2015} to re-derive the ADI method presented in 
-\citet{Marthi2014} and in doing so we have shown that the algorithm in \citet{Marthi2014} is actually redundant StEfCal. We could also have shown this by using a similar approach 
-to the one presented in \citet{Salvini2014}.
-
-The question now arises what value should we use for $\alpha$? In \citet{Marthi2014} they propose 
-$\alpha = 0.3$. Although an exact mathematical proof for the optimal choice of $\alpha$ still eludes us, it is still possible to
-venture a clever guess by studying what the choice of $\alpha$ is in the case of skymodel-based StEFCal. Recall from Section~\ref{sec:sbc} that $\alpha$ is equal to a half 
-in the case of StEFCal.
-If we wish to mimic this behaviour for the redundant calibration use case we need to scale $\breve{\bz}$ by two thirds in Eq.~\eqref{eq:alpha} (see Eq.~\eqref{eq:two_thrids}) which implies that we should choose $\alpha$ equal to a third, which corresponds to a $\lambda$ equal to two,
-which is the value we used in this paper.
 
@@
 Note that we do not explicitly denote the $k$ index in Eq.~\eqref{eq:LINCAL_update}. We update our parameter vector until convergence is reached.
+
+\section{Skymodel-based Calibration}
+\label{sec:sbc}
+\textsc{StEFCal} is an alternating direction implicit method, in which the measurement equation, Eq.~\eqref{eq:vis_definition}, is linearized by assuming that the gains are known, but
+that their conjugates are unknown \citep{Mitchell:MWA-cal,Salvini2014}.
+Under this assumption, the conjugates of the gains can be obtained trivially.
+Once the conjugates of the gains have been computed, the gains can be solved for.
+One can then alternate between computing the gains and their conjugates until convergence
+is reached.
+By correctly exploiting the symmetry inherent in the calibration problem this alternating approach can be replaced by a single iterative new parameter estimation step.
+\cite{Salvini2014} also show that
+the convergence of the algorithm can be improved if for every even iteration the basic new parameter estimation step is modified as follows: the current gain estimate is replaced by the average of the current gain estimate and the gain estimate of the previous odd iteration.
+\textsc{StEFCal} lowers the computational complexity of normal skymodel-based calibration from $\mathbb{O}(N^3)$ to $\mathbb{O}(N^2)$.
+
+\citet{Smirnov2015} recently showed that one can also derive \textsc{StEFCal} by viewing normal skymodel-based calibration as a complex optimization problem.
+As in the redundant use case depicted in Fig.~\ref{fig:hessian}, when we inspect the skymodel-based $\bH$ it becomes apparent that its diagonal entries are much more significant than its off-diagonal entries.
+By approximating $\bH$ by its diagonal and substituting this approximate Hessian $\widetilde{\bH}$ back into the fundamental skymodel-based LM update step we obtain \citep{Smirnov2015} 
+\begin{align}
+\breve{\bg}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bg}_k,\label{eq:stef_lambda}\\
+ &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bg}_k, \label{eq:stef_alpha}  
+\end{align}
+
+If we choose $\lambda$ to be zero, which corresponds to an $\alpha$ equal to one, we obtain the new parameter estimation step of \textsc{StEfCal}; used during odd iterations.
+On the other hand if we choose $\lambda$ equal to one, which corresponds
+to an $\alpha$ equal to a half, we obtain the new parameter estimation step of \textsc{StEfCal}; used during even iterations.
+
+Furthermore, the simplified skymodel-based GN equivalent of Eq.~\eqref{eq:two_thrids} is equal to
+\citep{Smirnov2015}
+\begin{equation}
+\label{eq:one_half}
+\breve{\bg}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{1}{2}\breve{\bg}_{k}. 
+\end{equation}
+Note that in Eq.~\eqref{eq:one_half} the influence that the estimate of the parameters from the previous iteration has on the current iteration is scaled by a half, while
+in Eq.~\eqref{eq:two_thrids} it is scaled by two thirds. Note the similarity between the last terms of Eq.~\eqref{eq:stef_alpha} and Eq.~\eqref{eq:one_half} when $\alpha$ is a half.
+
 \label{lastpage}

### 2017-01-09 paper.tex
added prec error results

@@
 \begin{align}
- \left [ \bd \right]_{\alpha_{pq}} &= d_{pq}, & \left [ \bv \right ]_{\alpha_{pq}} &= g_p y_{\phi_{pq}} \conj{g_q},\nonumber\\
+ \left [ \bd \right]_{\alpha_{pq}} &= d_{pq}, & \left [ \bv \right ]_{\alpha_{pq}} &= v_{pq}=g_p y_{\phi_{pq}} \conj{g_q},\nonumber\\
  \left [ \bn \right ]_{\alpha_{pq}} &= n_{pq}. &  &\label{eq:vec_definitions}
@@
 Eq.~\eqref{eq:g_update} and Eq.~\eqref{eq:y_update} are the exact same iterative new parameter estimators derived by \citet{Marthi2014}. Als note that Eq.~\eqref{eq:g_update} is the fundamental
-gain estimator associated with StEfCal. We have therefore successfully used the diagonal approximation approach suggested by \citet{Smirnov2015} to re-derive the ADI method presented in 
+gain estimator associated with \textsc{StEfCal}. We have therefore successfully used the diagonal approximation approach proposed by \citet{Smirnov2015} to re-derive the ADI method presented in 
 \citet{Marthi2014} and in doing so we have shown that the algorithm in \citet{Marthi2014} is actually redundant StEfCal. We could also have shown this by using a similar approach 
@@
 
-The question now arises what value should we use for $\alpha$? In \citet{Marthi2014} they propose 
+The question now arises what value should we use for $\alpha$? In \citet{Marthi2014}, they propose 
 $\alpha = 0.3$. Although an exact mathematical proof for the optimal choice of $\alpha$ still eludes us, it is still possible to
-venture a clever guess by studying what the choice of $\alpha$ is in the case of skymodel-based StEFCal. Recall from Section~\ref{sec:sbc} that $\alpha$ is equal to a half 
-in the case of StEFCal (for even iterations).
+venture a clever guess by studying what the choice of $\alpha$ is in the case of skymodel-based \textsc{StEFCal}. The value of $\alpha$ is equal to a half 
+in the case of \textsc{StEFCal} (for even iterations); see Appendix~\ref{sec:sbc}.
 If we wish to mimic this behaviour for the redundant calibration use case we need to scale $\breve{\bz}$ by two thirds in Eq.~\eqref{eq:alpha} (see Eq.~\eqref{eq:two_thrids}) which implies that we should choose $\alpha$ equal to a third, which corresponds to a $\lambda$ equal to two.
@@
 
+Optimality results of ADI were presented in \citet{Marthi2014}. We do not repeat their analysis here. We however, do present our own ADI results in Fig.~\ref{fig:prec_error} that confirm that the
+ADI method can accurately estimate the antenna gains and the observed visibilities at different SNR values. To quantity the exact error in our estimates we used the precentage error $\beta$ defined 
+as 
+\begin{equation}
+\label{eq:beta}
+\beta = \frac{\sum_{pq} |v_{pq} - \widehat{v}_{pq} |^2}{v_{pq}}.
+\end{equation}
+We have already defined $v_{pq}$ in Eq.~\eqref{eq:vec_definitions}. Moreover, in Eq.~\eqref{eq:beta}, $\widehat{v}_{pq}$ denotes the ADI estimate of $v_pq$.
+
 \begin{figure}
 \includegraphics[width=0.47\textwidth]{./prec_error.pdf} 
-\caption{Precentage Error}
+\caption{We plot the precentage error $\beta$ between the estimated 
+predicted visibilities and the true predicted visibilities for different SNR values as a function of the number of antennas in the array $N$.
+The graph presented here was generated using the first experimental setup in Table~\ref{tab:ch_parm}.}
 \label{fig:prec_error}

### 2017-01-09 paper.tex
newest version

@@
 \label{sec:adi}
-
 %In Section~\ref{sec:adi} we show how the ADI method used for skymodel-based calibration relates to the redundant ADI method presented in \citet{Marthi2014}. 
@@
 \label{eq:beta}
-\beta = \frac{\sum_{pq} |v_{pq} - \widehat{v}_{pq} |^2}{v_{pq}}.
-\end{equation}
-We have already defined $v_{pq}$ in Eq.~\eqref{eq:vec_definitions}. Moreover, in Eq.~\eqref{eq:beta}, $\widehat{v}_{pq}$ denotes the ADI estimate of $v_pq$.
+\beta = \frac{\sum_{pq} |v_{pq} - \widehat{v}_{pq} |^2}{\sum_{pq} |v_{pq}|^2}.
+\end{equation}
+We have already defined $v_{pq}$ in Eq.~\eqref{eq:vec_definitions}. Moreover, in Eq.~\eqref{eq:beta}, $\widehat{v}_{pq}$ denotes the ADI estimate of $v_{pq}$.
 
@@
 In this section we investigate the computational complexity of the PCG method when applied to the redundant calibration problem. 
-The PCG method was presented in Section~\ref{sec:conj_grad}.
-The PCG method differs from the ADI method presented in Section~\ref{sec:adi} in that in the case of PCG we compute the exact inverse of the modified Hessian $\bmH$ instead 
-of using some sort of approximation as is the case for ADI. As we mentioned in Section~\ref{sec:pcg} there
-are two factors which influence the computational complexity of the PCG method, namely the spectral condition number and the sparsity of the matrix we wish to invert (see Eq.~\eqref{eq:cg_bound}). 
+As we mentioned in Section~\ref{sec:pcg} there are two factors which influence the computational complexity of the PCG method, namely the spectral condition number and the sparsity of the matrix we wish to invert (see Eq.~\eqref{eq:cg_bound}).
 We used a $\lambda$ value of two, implying we used Eq.~\eqref{eq:LM_update} instead of Eq.~\eqref{eq:GN_update} throughout this section.
@@
 In Section~\ref{sec:normal},
-we showed that the PCG method can in fact be used to invert a singular Hessian $\bH$.
+we showed that the PCG method would in fact be able to handle such a scenario.
 
@@
 is a good candidate for the redundant use case. In Fig.~\ref{fig:kappa} we see that this choice of preconditioner performs very well, the value of the spectral condition number $\kappa$
-becomes constant and independent of problem size, i.e. the number of antennas in the array. Preconditioning effectively eliminates it from Eq.~\eqref{eq:cg_bound}. This result
+becomes constant and independent of problem size, i.e. the number of antennas in the array. Preconditioning therefore effectively eliminates $\kappa$ from Eq.~\eqref{eq:cg_bound}. This result
 is confirmed by Fig.~\ref{fig:cg_itr}.
+Fig.~\ref{fig:cg_itr} shows us that the number of major iterations needed by the preconditioned conjugate gradient method to invert $\bmH$ is independent of the number of antennas in the array.
+Moreover, the number of major iterations required are also much smaller than the dimension of $\bmH$.
 
@@
 In the left panel we have plotted the spectral condition number $\kappa$ of the modified Hessian $\bmH$, before (magenta and red curves) and after preconditioning (blue curve), at different SNR values as a function of the
-number of antennas $N$ in the array. This plot shows us that we can drastically reduce the condition number of $\bmH$ by preconditioning. Moreover, $\kappa$ is also
-independent of the problem size (the number of antennas in the array). The fact that $\kappa$ is small and independent of problem size (effectively constant) implies that it simply drops out 
-of Eq.~\eqref{eq:cg_bound}.
+number of antennas $N$ in the array.
 The right hand panel confirms this result.
@@
 gradient method to invert $\bmH$, before (magenta and red curves) and after preconditioning (blue and green curves), at different SNR values as a function of the number of antennas $N$ in the array.
-The right hand panel therefore shows us that the number of major iterations needed by the preconditioned conjugate gradient method to invert $\bmH$ is independent of the number of antennas in the array.
-The number of major iterations are also much smaller than the dimension of $\bmH$. 
 \label{fig:kappa_itr}} 
@@
 The second factor that determines the computational complexity of the conjugate gradient method when used to perform redundant calibration is the amount of non-zero
-entries in the modiefied Hessian matrix $\bmH$. 
+entries in the modified Hessian matrix $\bmH$. 
 %We have also shown in section that the first factor, the spectral condition number of $\bH$, can be reduced 
@@
 
-In Fig.~\ref{fig:gamma} we plot the sparsity ratio for three different geometric layouts as a function of the number antennas in the array. Although, we can not explicitly calculate 
-the limit of the hexagonal and square layouts, Fig.~\ref{fig:gamma} seems to indicate that these limits do exist.
-
-We can now use the sparsity factor to compute the order of the computational complexity of inverting $\bmH$ with
+In Fig.~\ref{fig:gamma} we plot the empirically determined sparsity ratios for three different geometric layouts as a function of the number antennas in the array.
+Although, we can not explicitly calculate
+the limit of the hexagonal and square layout curves, Fig.~\ref{fig:gamma} seems to indicate that they do exist. Moreover, we also plot the left most equation
+of Eq.~\eqref{eq:gamma} (red dashed line) and the right most equation of Eq.~\eqref{eq:gamma} (black dotted curve) in Fig.~\ref{fig:gamma}.
+
+We can now use the sparsity factor to compute the order of the computational complexity of inverting $\bmH$ (assuming the effect of $\kappa$ is negliable) with
 \begin{equation}
@@
 \end{align}
-In Fig.~\ref{fig:c} we plot the value of $c$ for three different geometric layouts as a function of the number antennas in the array.
-It is clear from Eq.~\eqref{eq:c} and Fig.~\ref{fig:c} that the computational complexity, under the assumption that $\lim_{N\rightarrow \infty} \gamma$ exists, is asymptotically bounded by $O(P^2)$.
-However, the computational complexity converges
-very slowly to its asymptotic value, and in general is equal to $O(P^{c})$, where $c \sim 1.7$ if a Hexagonal geometric layout containing less than 200 antennas is employed.
+Note, that we can only compute $c_{\infty}$ in the rightmost equation if $\lim_{N\rightarrow \infty} \gamma$ exists. 
+In Fig.~\ref{fig:c}, we plot the empiracally determined value of $c$ for three different geometric layouts as a function of the number antennas in the array.
+Furhtermore, in Fig.~\ref{fig:c}, the analytic value of $c$ for the regular east-west grid is plotted as a dashed red curve, while the limit of this analytic curve is plotted with a dotted black curve.
+
+Eq.~\eqref{eq:gamma}, Eq.~\eqref{eq:c} and Fig.~\ref{fig:sparsity} indicate that the computational complexity, under the assumption that $\lim_{N\rightarrow \infty} \gamma$ exists, is asymptotically bounded by $O(P^2)$.
+However, the computational complexity converges very slowly to its asymptotic value, and in general is equal to $O(P^{c})$, where $c \sim 1.7$ if a Hexagonal geometric layout containing less than 200 antennas is employed.
 
@@
 Taking the limit
-of this expression, Eq.~\eqref{eq:gamma}, as $N\rightarrow \infty$ produces the black dotted line.
-In the right panel we plot the order of the computational cost $c$ for inverting $\bmH$ as a function of the number
-of antennas in the array for different geometries. We use the same color scheme in the right and left panels. We compute $c$ using Eq.~\eqref{eq:c}. The red dashed line 
+of this analytic expression produces the black dotted line. In the right panel we plot the order of the computational cost $c$ for inverting $\bmH$ as a function of the number 
+of antennas in the array for different geometries. We use the same color scheme in the right and left panels. The red dashed line 
 in the right panel is associated with the theoretical curve of $c$ we computed for the east-west regular grid. The limit of the red-dashed curve is two, which we 

### 2017-01-11 paper.tex
started working on section 5 again

@@
 \label{sec:results}
-In this section we compare the computational complexity of two newly proposed speedup algorithms for redundant calibration, namely the ADI \citep{Marthi2014} and PCG \citep{Liu2010} methods.
+In the previous section we discussed the complex GN and LM algorithms and how they relate to redundant calibration. If we inspect Eq.~\eqref{eq:GN_update} and 
+Eq.~\eqref{eq:GN_update} we see that the most expensive operation used by the GN and LM algorithms is the inversion of the Hessian $\bH$. The most logical 
+way of speeding up the algorithm is to investigate whether we can do this inversion in some efficient way.
+
+Two recent algorithms have been proposed that are able to invert $\bH$ in an efficient manner, namely the ADI \citep{Marthi2014} and PCG \citep{Liu2010} methods. 
 We restrict ourselves to comparing the most 
@@
 
+\subsubsection{Skymodel-based Calibration}
+\label{sec:sbc}
+\textsc{StEFCal} is an alternating direction implicit method, in which the measurement equation, Eq.~\eqref{eq:vis_definition}, is linearized by assuming that the gains are known, but
+that their conjugates are unknown \citep{Mitchell:MWA-cal,Salvini2014}.
+Under this assumption, the conjugates of the gains can be obtained trivially.
+Once the conjugates of the gains have been computed, the gains can be solved for.
+One can then alternate between computing the gains and their conjugates until convergence
+is reached.
+By correctly exploiting the symmetry inherent in the calibration problem this alternating approach can be replaced by a single iterative new parameter estimation step.
+\cite{Salvini2014} also show that
+the convergence of the algorithm can be improved if for every even iteration the basic new parameter estimation step is modified as follows: the current gain estimate is replaced by the average of the current gain estimate and the gain estimate of the previous odd iteration.
+\textsc{StEFCal} lowers the computational complexity of normal skymodel-based calibration from $\mathbb{O}(N^3)$ to $\mathbb{O}(N^2)$.
+
+\citet{Smirnov2015} recently showed that one can also derive \textsc{StEFCal} by viewing normal skymodel-based calibration as a complex optimization problem.
+As in the redundant use case depicted in Fig.~\ref{fig:hessian}, when we inspect the skymodel-based $\bH$ it becomes apparent that its diagonal entries are much more significant than its off-diagonal entries.
+By approximating $\bH$ by its diagonal and substituting this approximate Hessian $\widetilde{\bH}$ back into the fundamental skymodel-based LM update step we obtain \citep{Smirnov2015} 
+\begin{align}
+\breve{\bg}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bg}_k,\label{eq:stef_lambda}\\
+ &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bg}_k, \label{eq:stef_alpha}  
+\end{align}
+
+If we choose $\lambda$ to be zero, which corresponds to an $\alpha$ equal to one, we obtain the new parameter estimation step of \textsc{StEfCal}; used during odd iterations.
+On the other hand if we choose $\lambda$ equal to one, which corresponds
+to an $\alpha$ equal to a half, we obtain the new parameter estimation step of \textsc{StEfCal}; used during even iterations.
+
+Furthermore, the simplified skymodel-based GN equivalent of Eq.~\eqref{eq:two_thrids} is equal to
+\citep{Smirnov2015}
+\begin{equation}
+\label{eq:one_half}
+\breve{\bg}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{1}{2}\breve{\bg}_{k}. 
+\end{equation}
+Note that in Eq.~\eqref{eq:one_half} the influence that the estimate of the parameters from the previous iteration has on the current iteration is scaled by a half, while
+in Eq.~\eqref{eq:two_thrids} it is scaled by two thirds. Note the similarity between the last terms of Eq.~\eqref{eq:stef_alpha} and Eq.~\eqref{eq:one_half} when $\alpha$ is a half.
+
+\subsubsection{Redundant Calibration}
 As we mention in Section~\ref{sec:results}, inspecting Fig.~\ref{fig:hessian} (also see Eq.~\eqref{eq:red_H}) reveals that in the case of redundant calibration the Hessian's diagonal entries are more significant than its off-diagonal entries.
@@
 
-\section{Skymodel-based Calibration}
-\label{sec:sbc}
-\textsc{StEFCal} is an alternating direction implicit method, in which the measurement equation, Eq.~\eqref{eq:vis_definition}, is linearized by assuming that the gains are known, but
-that their conjugates are unknown \citep{Mitchell:MWA-cal,Salvini2014}.
-Under this assumption, the conjugates of the gains can be obtained trivially.
-Once the conjugates of the gains have been computed, the gains can be solved for.
-One can then alternate between computing the gains and their conjugates until convergence
-is reached.
-By correctly exploiting the symmetry inherent in the calibration problem this alternating approach can be replaced by a single iterative new parameter estimation step.
-\cite{Salvini2014} also show that
-the convergence of the algorithm can be improved if for every even iteration the basic new parameter estimation step is modified as follows: the current gain estimate is replaced by the average of the current gain estimate and the gain estimate of the previous odd iteration.
-\textsc{StEFCal} lowers the computational complexity of normal skymodel-based calibration from $\mathbb{O}(N^3)$ to $\mathbb{O}(N^2)$.
-
-\citet{Smirnov2015} recently showed that one can also derive \textsc{StEFCal} by viewing normal skymodel-based calibration as a complex optimization problem.
-As in the redundant use case depicted in Fig.~\ref{fig:hessian}, when we inspect the skymodel-based $\bH$ it becomes apparent that its diagonal entries are much more significant than its off-diagonal entries.
-By approximating $\bH$ by its diagonal and substituting this approximate Hessian $\widetilde{\bH}$ back into the fundamental skymodel-based LM update step we obtain \citep{Smirnov2015} 
-\begin{align}
-\breve{\bg}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bg}_k,\label{eq:stef_lambda}\\
- &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bg}_k, \label{eq:stef_alpha}  
-\end{align}
-
-If we choose $\lambda$ to be zero, which corresponds to an $\alpha$ equal to one, we obtain the new parameter estimation step of \textsc{StEfCal}; used during odd iterations.
-On the other hand if we choose $\lambda$ equal to one, which corresponds
-to an $\alpha$ equal to a half, we obtain the new parameter estimation step of \textsc{StEfCal}; used during even iterations.
-
-Furthermore, the simplified skymodel-based GN equivalent of Eq.~\eqref{eq:two_thrids} is equal to
-\citep{Smirnov2015}
-\begin{equation}
-\label{eq:one_half}
-\breve{\bg}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{1}{2}\breve{\bg}_{k}. 
-\end{equation}
-Note that in Eq.~\eqref{eq:one_half} the influence that the estimate of the parameters from the previous iteration has on the current iteration is scaled by a half, while
-in Eq.~\eqref{eq:two_thrids} it is scaled by two thirds. Note the similarity between the last terms of Eq.~\eqref{eq:stef_alpha} and Eq.~\eqref{eq:one_half} when $\alpha$ is a half.
-
 \label{lastpage}

### 2017-01-11 paper.tex
added section 5 to the paper

@@
 Two recent algorithms have been proposed that are able to invert $\bH$ in an efficient manner, namely the ADI \citep{Marthi2014} and PCG \citep{Liu2010} methods. 
-We restrict ourselves to comparing the most 
-expensive operation of the two methods namely, computing the inverse of the Hessian $\bH$ (or modified Hessian $\bmH$ in the case of LM). It becomes trivial to rate these two 
-algorithms relative to each other once the computational complexity of the inverse operation is known. The simulations we performed to compare these two 
+Note that here we focus solely on comparing the computational complexity of inverting $\bH$ using either one of these two methods. Rating the LM or GN algorithms
+using one of these approaches relative to each other once the computation complexity of inverting $\bH$ is known is trivial. The simulations we performed to compare these two 
 methods are summarized in Section~\ref{sec:simulations}.
 
-The basic difference between these two algorithms can be understood by inspecting Fig.~\ref{fig:hessian}. The first observation we can make by inspecting 
+%We restrict ourselves to comparing the most 
+%expensive operation of the two methods namely, computing the inverse of the Hessian $\bH$ (or modified Hessian $\bmH$ in the case of LM). It becomes trivial to rate these two 
+%algorithms relative to each other once the computational complexity of the inverse operation is known.
+
+The basic difference between these two speedup algorithms can be understood by inspecting Fig.~\ref{fig:hessian}. The first observation we can make by inspecting 
 Fig.~\ref{fig:hessian_reg} and Fig.~\ref{fig:hessian_hex} is that the diagonal entries of the Hessians are more significant than the off-diagonal entries.
-It therefore makes sense to approximate the Hessian (or the modified Hessian) with its diagonal, which greatly simplifies the computational complexity of inverting 
-the Hessian, i.e. it becomes $\mathbb{O}(P)$\footnote{The cost of constructing 
+It therefore makes sense to approximate the $\bH$ (or the modified Hessian $\bmH$) with its diagonal, which greatly simplifies the computational complexity of inverting 
+$\bH$, i.e. it becomes $\mathbb{O}(P)$\footnote{The cost of constructing 
 the diagonal is not included in this estimate.} \citep{Smirnov2015}. This diagonal-approximation approach is effectively ADI \citep{Marthi2014}. The exact details of this 
@@
 We will also show how \citet{Marthi2014}, \citet{Smirnov2015} and \citet{Salvini2014} relate to each other in that same section.
-In Appendix~\ref{sec:sbc} we briefly summarize the ADI method that is used for skymodel-based calibration.
+It should be noted that the ADI method actually refers to the complete LM update step using the approximate $\bH$ and not just
+to the inversion of the approximate $\bH$. Furhtermore, the name ADI also implies that with this algorithm we alternate between computing the 
+gains, their conjugates and the true observed visibilities. %In Appendix~\ref{sec:sbc} we briefly summarize the ADI method that is used for skymodel-based calibration.
 
@@
 can be applied to invert them in a computationally efficient manner.
-\citet{Liu2010} proposed the PCG method to compute the inverse of the Hessian, however the computational complexity of this approach was not analyzed therein.
+\citet{Liu2010} proposed the PCG method to compute the inverse of $\bH$, however the computational complexity of this approach was not analyzed therein.
 In Section~\ref{sec:pcg}, we do a proper computational complexity analysis of PCG. We find that the computational complexity of PCG is asymptotically bounded by $\mathbb{O}(P^2)$, which is 
@@
 %In Section~\ref{sec:adi} we show how the ADI method used for skymodel-based calibration relates to the redundant ADI method presented in \citet{Marthi2014}.
-In this section we re-derive the ADI method presented in \citet{Marthi2014}. In \citet{Marthi2014} the ADI method is derived by taking the derivative of the 
+
+We start this section by briefly reviewing the ADI method as it pertains to skymodel-based calibration in Section~\ref{sec:sbc} \citep{Salvini2014,Smirnov2015}.
+In Section~\ref{sec:red_c} we re-derive the redundant calibration ADI method presented in \citet{Marthi2014}. In \citet{Marthi2014} the ADI method is derived by taking the derivative of the 
 objective function $\Lambda$ relative to $\bg$ and $\by$ and setting the result to naught. When derived in this way, its relation to the recent works \citep{Salvini2014,Smirnov2015} are unclear.
 We will use the diagonal-approximate approach proposed in \citet{Smirnov2015} to show that the ADI algorithm presented in \citet{Marthi2014} is in actual fact redundant 
-\textsc{StEfCal} \citep{Salvini2014}. Perusing through Appendix~\ref{sec:sbc} is advised before continuing this section.
+\textsc{StEfCal} \citep{Salvini2014}.
 
@@
 \citet{Smirnov2015} recently showed that one can also derive \textsc{StEFCal} by viewing normal skymodel-based calibration as a complex optimization problem.
-As in the redundant use case depicted in Fig.~\ref{fig:hessian}, when we inspect the skymodel-based $\bH$ it becomes apparent that its diagonal entries are much more significant than its off-diagonal entries.
-By approximating $\bH$ by its diagonal and substituting this approximate Hessian $\widetilde{\bH}$ back into the fundamental skymodel-based LM update step we obtain \citep{Smirnov2015} 
+They realized while inspecting the skymodel-based $\bH$ that its diagonal entries are much more significant than its off-diagonal entries.
+As we mention in Section~\ref{sec:results}, the exact same conclusion can be reached if we inspect the Hessian $\bH$ that is associated with redundant calibration (see Fig.~\ref{fig:hessian}). 
+By approximating the skymodel-based $\bH$ by its diagonal and substituting this approximate Hessian $\widetilde{\bH}$ back into the fundamental skymodel-based LM update step we obtain \citep{Smirnov2015}
 \begin{align}
@@
 \subsubsection{Redundant Calibration}
-As we mention in Section~\ref{sec:results}, inspecting Fig.~\ref{fig:hessian} (also see Eq.~\eqref{eq:red_H}) reveals that in the case of redundant calibration the Hessian's diagonal entries are more significant than its off-diagonal entries.
+\label{sec:red_c}
+Since the diagonal entries of the $\bH$ associated with redundant calibration are more significant than its off-diagonal entries it makes 
+sense to try and approximate it with its diagonal (see Fig.~\ref{fig:hessian} and Eq.~\eqref{eq:red_H}).
 If we substitute $\bJ^H\bJ$ with $\widetilde{\bH}$ (by its diagonal approximation) and replace $\breve{\br}$ with $\breve{\bd} - \breve{\bv}$ in Eq.~\eqref{eq:GN_update} we obtain
@@
 
-Eq.~\eqref{eq:g_update} and Eq.~\eqref{eq:y_update} are the exact same iterative new parameter estimators derived by \citet{Marthi2014}. Als note that Eq.~\eqref{eq:g_update} is the fundamental
+Eq.~\eqref{eq:g_update} and Eq.~\eqref{eq:y_update} are the exact same iterative new parameter estimators derived by \citet{Marthi2014}. Also note that Eq.~\eqref{eq:g_update} is the fundamental
 gain estimator associated with \textsc{StEfCal}. We have therefore successfully used the diagonal approximation approach proposed by \citet{Smirnov2015} to re-derive the ADI method presented in 
-\citet{Marthi2014} and in doing so we have shown that the algorithm in \citet{Marthi2014} is actually redundant StEfCal. We could also have shown this by using a similar approach 
-to the one presented in \citet{Salvini2014}.
+\citet{Marthi2014} and in doing so we have shown that the algorithm in \citet{Marthi2014} is actually redundant StEfCal. We could also have shown this by using the alternating approach 
+presented in \citet{Salvini2014}; which explains why, in this paper, we have been referring to the algorithm presented in \citet{Marthi2014} as the ADI method.
 
@@
 venture a clever guess by studying what the choice of $\alpha$ is in the case of skymodel-based \textsc{StEFCal}. The value of $\alpha$ is equal to a half 
-in the case of \textsc{StEFCal} (for even iterations); see Appendix~\ref{sec:sbc}.
+in the case of \textsc{StEFCal} (for even iterations); see Section~\ref{sec:sbc}.
 If we wish to mimic this behaviour for the redundant calibration use case we need to scale $\breve{\bz}$ by two thirds in Eq.~\eqref{eq:alpha} (see Eq.~\eqref{eq:two_thrids}) which implies that we should choose $\alpha$ equal to a third, which corresponds to a $\lambda$ equal to two.

### 2017-01-11 paper.tex
started writing the conclusion

@@
 We are now in a position to make a final comparison between the ADI and PCG methods.
-One inversion using the ADI method is very cheap, while one inverse using PCG is more expensive.
+Inverting $\bH$ by assuming it is a diagonal matrix (i.e.
+using the ADI method) is very cheap, while inverting $\bH$ using PCG is more expensive.
 On the other side of the coin,  one inversion with PCG is much more
-accurate than one inversion with ADI.
+accurate than using the approximate inverse which is associated with the ADI method.
 The question is now, if the inverse using the PCG method is much more accurate will the LM implementation using it require much less LM-iterations than its ADI counterpart.
 Moreover, will the number
-of LM iterations PCG require be so few that it manages to beat the ADI approach in any case. To answer this we first need to compute the number of iterations $k$; ADI-LM can make in the 
-same time it takes one single PCG-LM iteration to execute. We can compute this with
+of LM iterations PCG require be so few that it manages to beat the ADI method after all.
+To answer this we first need to compute the number of iterations, $k$, the LM implementation associated with the ADI method can make in the
+same time it takes the LM implementation associated with the PCG method to perform a single iteration. We can compute this with
 \begin{align}
@@
 \caption{The graphs presented here were generated using the second experimental setup in Table~\ref{tab:ch_parm}. In the left panel we have the number of LM iterations 
-that is required for the LM algorithm using the ADI (green and blue curve) and the PCG (magenta and red curve) methods to converge to a parameter convergence criteria of 1e-6 whilst using different 
-SNR values as a function of number of antennas $N$ in the array.
+that is required for the LM algorithm using the ADI (green and blue curve) and the PCG (magenta and red curve) methods to converge to a parameter error tolerance of 1e-6 whilst using different 
+SNR values as a function of the number of antennas $N$ in the array.
 In the right panel we have the average amount of LM iterations (difference between the ADI and PCG curves in the left panel) we save (green and blue curve) by computing the full-inverse with the PCG method whilst using different SNR values.
-The black curve is associated with $k$ (see Eq.~\eqref{eq:k}). Clearly we do not save enough iterations using the full-inverse of $\bmH$ for the PCG method to outperform the ADI method.
+The black curve is associated with $k$, the number of iterations the LM implementation using the ADI method can perform in the same time it takes the LM implementation using 
+the PCG method to perform one iteration.
 \label{fig:out_diff}} 
@@
 \section{Conclusion}
-Hallo
+In Section~\ref{sec:} we showed that instead of using the traditional phase-and-amplitude approach to perform redundant calibration we can actually use a purely complex
+framework. In this framework the unknown variables and their conjugates are treated as independent variables that have to be estimated. We have also 
+proposed to use the LM algorithm, instead of the GN algorithm, to perform the actual parameter estimation within this complex framework.
+The LM algorithm exhibits better convergence properties than the
+GN algorithm. This has the immediate consequence that the amount of algorithmic layers that are needed to perform redundant calibration are reduced.
+Them most expensive operation associated with the LM algorithm (as well as the GN algorithm) is the inversion of $\bH$.
+We also compared the computational complexity of two previously proposed algorithms to efficiently invert $\bH$ for the redundant calibration use case, namely 
+ADI and PCG. We showed that the computational complexity of the ADI method is $\mathbb{O}(P)$, while the computational complexity of the PCG method is asymptotically bounded
+by $\mathbb{O}(P^2)$. Both of these algorithms perform much better than stock standard inversion, which has a computational complexity of $\mathbb{O}(P^3)$.
+Although the ADI method does execute faster than the PCG method it is less robust as it
 

### 2017-01-11 paper.tex
more added to the conclusion

@@
 by $\mathbb{O}(P^2)$. Both of these algorithms perform much better than stock standard inversion, which has a computational complexity of $\mathbb{O}(P^3)$.
-Although the ADI method does execute faster than the PCG method it is less robust as it
+Although the ADI method does execute faster than the PCG method it is far less robust as it would be less straightforward to apply to other calibration uses cases. The reason being,
+the diagonal approximation we make with might not be accurate
 

### 2017-01-12 paper.tex
added conclusion

@@
 \section{Redundant Calibration}
+\label{sec:main_sec}
 As discussed in Section~\ref{sec:ls}, we can solve Eq.~\eqref{eq:least_squares_red} in one of two ways. We can split the problem into its real and
@@
 \subsubsection{Complex GN and LM}
+\label{sec:complex_GN_LM}
 In this section we discuss how we can use the GN and LM algorithms to solve Eq.~\eqref{eq:least_squares_complex} (i.e. estimate $\breve{\bz}$). The complex Jacobian is defined to be
@@
 \section{Conclusion}
-In Section~\ref{sec:} we showed that instead of using the traditional phase-and-amplitude approach to perform redundant calibration we can actually use a purely complex
-framework. In this framework the unknown variables and their conjugates are treated as independent variables that have to be estimated. We have also 
-proposed to use the LM algorithm, instead of the GN algorithm, to perform the actual parameter estimation within this complex framework.
+In Section~\ref{sec:main_sec} we showed that instead of using the traditional phase-and-amplitude approach to perform redundant calibration we can actually use a purely complex
+framework. In this framework the unknown variables and their conjugates are treated as independent variables that have to be estimated. 
+We then derived the GN and LM update steps associated with this complex framework (see Section~\ref{sec:complex_GN_LM}).
+In this paper we focused more on the LM algorithm than on the GN algorithm.
 The LM algorithm exhibits better convergence properties than the
 GN algorithm. This has the immediate consequence that the amount of algorithmic layers that are needed to perform redundant calibration are reduced.
-Them most expensive operation associated with the LM algorithm (as well as the GN algorithm) is the inversion of $\bH$.
-We also compared the computational complexity of two previously proposed algorithms to efficiently invert $\bH$ for the redundant calibration use case, namely 
-ADI and PCG. We showed that the computational complexity of the ADI method is $\mathbb{O}(P)$, while the computational complexity of the PCG method is asymptotically bounded
+
+The most expensive operation associated with the LM algorithm (as well as the GN algorithm) is the inversion of $\bH$.
+We compared the computational complexity of two previously proposed algorithms to efficiently invert $\bH$ for the redundant calibration use case in Section~\ref{sec:results}, namely 
+ADI \citep{Marthi2014} and PCG \citep{Liu2010}.
+We showed that the computational complexity of the ADI method is $\mathbb{O}(P)$, while the computational complexity of the PCG method is asymptotically bounded
 by $\mathbb{O}(P^2)$. Both of these algorithms perform much better than stock standard inversion, which has a computational complexity of $\mathbb{O}(P^3)$.
 Although the ADI method does execute faster than the PCG method it is far less robust as it would be less straightforward to apply to other calibration uses cases. The reason being,
-the diagonal approximation we make with might not be accurate
+the diagonal approximation we make with ADI might not be accurate enough in general.
+
+Lastly, we showed how \citet{Marthi2014}, \citet{Salvini2014} and \citet{Smirnov2015} relate to each other (see Section~\ref{sec:red_c}).
+The ADI method associated with skymodel-based calibration (i.e.
+\textsc{StEfCal}) is presented
+in \citet{Salvini2014}. \citet{Smirnov2015} showed that it is possible to rederive this skymodel-based ADI method through complex optimization. In particular, they showed that if 
+we approximate the $\bH$ associated with the LM algorithm with its diagonal then we can rederive the ADI update rules presented in \citet{Salvini2014}. 
+\citet{Marthi2014} applied the ADI method to the redundant use case.
+This fact is not immediatly clear when inspecting \citet{Marthi2014}.
+The ADI update rules presented \citet{Marthi2014} were derived by taking the partial derivatives of
+the objective function towards the unknown parameters, setting the intermediate results to zero and then solving for the unknowns. In this paper we show that 
+these same ADI update rules can be derived using the diagonal approximation technique proposed in \citet{Smirnov2015}. The fact that we can use this diagonal approximation
+technique to derive these update rules imply that the algorithm presented in \citet{Marthi2014} is in fact the ADI method (i.e. redundant \textsc{StEfCal}). This same
+fact can be established by directly applying the skymodel-based ADI derivation presented in \citet{Salvini2014} to the redundant use case.
 

### 2017-01-12 paper.tex
fixed abstract up a bit

@@
 \begin{abstract}
-Traditionally redundant calibration is implemented via a hierarchical algorithmic strategy, i.e. a basic \textsc{firstcal} is performed, followed by \textsc{logcal} and \textsc{lincal}.  
-\textsc{logcal} requires the use of the logartihm, while \textsc{lincal} makes use of the gradient-based Gauss-Newton algorithm.
-We show that instead of splitting our calibration problem into its phase and amplitude components,
-which is the approach taken by \textsc{lincal}, we can treat the complex unknown model parameters and their complex conjugates as separate unknowns instead. 
-We further propose the use of the Levenberg-Marquardt algorithm, instead of the Gauss-Newton algorithm, to improve the convergence properties of the proposed complex
-framework, which in turn reduces the number of layers that redundant calibration pipelines would require.
+Traditionally redundant calibration is implemented via a hierarchical algorithmic strategy.  
+We show that instead of splitting our redundant calibration problem into its phase-and-amplitude components,
+which is the more traditional approach, we can treat the complex unknown model parameters and their complex conjugates as separate unknowns instead. 
+We also derive the Levenberg-Maquardt and Gauss-Newton update steps associated with this alternative parameter-and-conjugate framework.
+We focus more on the Levenberg-Maquardt algorithm in this paper as it exhibits better convergence properties when compared to Gauss-Newton.
+Using the Levenberg-Maquardt algorithm, therefore, reduces the number of algorithmic layers that are needed to perform redundant calibration.
+The Gauss-Newton algorithm is, however, still the de facto standard that is currently being used to perform redundant calibration.
 We also compare two previously proposed methods to speedup redundant calibration, namely the alternating direction implicit method and the preconditioned conjugate gradient method.
@@
 direction implicit method outperforms the preconditioned conjugate gradient method, however the preconditioned conjugate gradient method is more robust as it is more straightforward to adapt to other calibration use cases.
-We mainly focus on the complex framework in this paper rather than the amplitude and phase approach as it makes it trivial to relate the alternating direction implicit method 
-and the Levenberg-Marquardt algorithm with each other.
+We make use of the parameter-and-conjugate framework in this paper, rather than the more traditional amplitude-and-phase approach as it makes it trivial to relate the alternating direction implicit method
+and the Levenberg-Marquardt algorithm with one another.
 \end{abstract}
@@
 The propogation effects (i.e.
-the ionosphere, antenna gains, etc.) a celestial signal experiences along its propogation path is modelled with the radio interforemetric measurement equation (RIME) \citep{ME1,RRIME1}.
-Calibration is the procedure by which we remove the errors that these propogation effects induce onto our intereferometric data and is normally accomplished using a 
+the ionosphere, antenna gains, etc.) a celestial signal experiences along its propagation path is modelled with the radio interferometric measurement equation (RIME) \citep{ME1,RRIME1}.
+Calibration is the procedure by which we remove the errors that these propagation effects induce onto our interferometric data and is normally accomplished using a 
 least-squares solver.
@@
 There are two recent developments in skymodel-based calibration which should be explicitly mentioned at this point, since they form the foundation this paper builds on.
-The first is the discovery of \textsc{StEfCal} \citep{Mitchell:MWA-cal,Salvini2014}, which
+The first is the discovery of \textsc{StEfCal} (Statistically Efficient Calibration) \citep{Mitchell:MWA-cal,Salvini2014}, which
 is an alternating direction implicit (ADI) skymodel-based calibration algorithm. It works by assuming the gains are known while solving for their conjugates and then performing the reverse. 
 This procedure results in the linearization of the calibration algorithm and reduces the computational complexity of calibration from $\mathbb{O}(N^3)$ to $\mathbb{O}(N^2)$, where $N$ denotes the number of antennas in the array.
-assume the conugates are known and solve for the gains. The second development is the idea: to not break the calibration problem 
-into its real and imaginary parts and then to solve for the real and imaginary parts of the unknown parameters seperately, but to rather treat the antenna gains and their 
+The second development is the idea: to not break the calibration problem 
+into its real and imaginary parts and then to solve for the real and imaginary parts of the unknown parameters separately, but to rather treat the antenna gains and their 
 conjugates as seperate parameters to estimate \citep{Smirnov2015}. \citet{Smirnov2015} also showed that the \textsc{StEfCal} algorithm itself is merely a special case of this
-gain-and-conjugate approach.
-
-Two baselines are redundant if their beaseline difference vectors have the same lenght and orientation, which imply that they measure the same $uv$-modes.
+gain-and-conjugate framework.
+
+Two baselines are redundant if their baseline difference vectors have the same length and orientation, which imply that they measure the same $uv$-modes.
 When we use a redundant array layout we dramatically
-reduce the number of unkowns. If our array is redundant enough we reduce the number of unknowns to such an extend that we are able to solve for the unknown 
+reduce the number of unknowns. If our array is redundant enough we reduce the number of unknowns to such an extend that we are able to solve for the unknown 
 observed visibilities themselves in addition to the antenna gains. Redundant calibration is the procedure by which we estimate the unknown observed visibilities in addition
@@
 
-Redundant calibration was already quite popular in the 1980's.
-\citet{Noordam1982} used redundant calibration and the WSRT (Westerbork Synthesis Radio Telescope) to produce high-dynamic range images of 3C84.
+Redundant calibration was popularized in the 1980's. \citet{Noordam1982} used redundant calibration and the WSRT (Westerbork Synthesis Radio Telescope) to produce high-dynamic range images of 3C84.
 The first implementations of redundant calibration made use of the logarithm function, because when the logarithm is applied to the scalar measurement equation, redundant 
-calibration is transformed into a linear problem \citep{Wieringa1992,Camps2003,Liu2010}. This approach is commenly referred to as \textsc{logcal}.
+calibration is transformed into a linear problem \citep{Wieringa1992,Camps2003,Liu2010}. The implementation associated with this approach is commonly referred to as \textsc{logcal}.
 We briefly discuss \textsc{logcal} in Appendix~\ref{sec:logcal}.
-The next improvement in redundant calibration was to perform a firt-order taylor expansion, which is equivalent to the Gauss-Newton (GN) algorithm \citep{Kurien2016}, of the measurement equation an then to solve for the
-perturbations of the model-parameters \citep{Liu2010}. The implementation of  this approach is called \textsc{lincal}. \textsc{lincal} is derived from first principals in Appendix~\ref{sec:lincal}. 
+The next improvement in redundant calibration was to perform a firt-order taylor expansion, which is equivalent to the Gauss-Newton (GN) algorithm \citep{Kurien2016}, of the measurement equation and then to solve for the
+perturbations of the model-parameters \citep{Liu2010}.
+\citet{Liu2010} perform the first order Taylor expansion of the measurement equation by splitting the unknown parameters into their phase-and-amplitude
+components. The implementation of  this approach is called \textsc{lincal}. \textsc{lincal} is derived from first principals in Appendix~\ref{sec:lincal}. 
+Currently, redundant calibration is usually implemented using a layered algorithmic strategy.
 \textsc{logcal} and \textsc{lincal} are both implemented in \textsc{omnical}\footnote{https://github.com/jeffzhen/omnical}, which is a layered redundant calibration pipeline \citep{Zheng2014,Ali2015}.
-The solutions of \textsc{logcal} are used as initial input for \textsc{lincal}, since the GN algorithm requires that its starting parameter guesses be in the vacinity 
+The solutions of \textsc{logcal} are used as initial input for \textsc{lincal}, since the GN algorithm requires that its starting parameter guesses be in the vicinity 
 of the solution.
+
+New instruments like LOFAR (Low Frequency Array) \citep{Noorishad2012}, PAPER (Precision Array for Probing the Epoch of Reionization) \citep{Ali2015} and HERA (Hydrogen Epoch of Reionization Array) \citep{deboer2015} have also recently sparked a renewed interest in redundant calibration.
+Using redundancy is especially advantageous when we are trying to detect the EoR (Epoch of Reionization) 21 cm signal,
+which is inherently quite faint \citep{Parsons2012}. Since redundant baselines make statistically independent measurements of the same $uv$-modes redundancy can greatly 
+improve the SNR (signal-to-noise ratio) of EoR detection experiments. This explains why PAPER and HERA, are highly redundant arrays, as they are both primarily used to detect the EoR 21 cm signal.
+
+\noindent
+\textbf{[GIANNI TO REWRITE AND EXTEND EOR SECTION]}
 
@@
 
-New instruments like LOFAR (Low Frequency Array) \citep{Noorishad2012}, PAPER (Precision Array for Probing the Epoch of Reionization) \citep{Ali2015} and HERA (Hydrogen Epoch of Reionization Array) \citep{deboer2015} have recently sparked a renewed interest in redundant calibration.
-Using redundancy is especially advantageous when we are trying to detect the EoR (Epoch of Reionization) 21 cm signal,
-which is inherently quite faint \citep{Parsons2012}. Since redundant baselines make statistically independent measurements of the same $uv$-modes redundancy can greatly 
-improve the SNR (signal-to-noise ratio) of EoR detection experiments. This explains why PAPER and HERA, are highly redundant arrays, as they are both primarily used to detect the EoR 21 cm signal.
-[GIANNI TO REWRITE AND EXTEND EOR SECTION]
-
-In this paper we propose to use the complex framework introduced by \citet{Smirnov2015} instead of the traditional amplitude and phase approach used by \textsc{lincal}.
-We also propose to use the Levenberg-Marquardt gradient-based algorithm instead of the Gauss-Newton algorithm as it exhibits better convergence properties.
-In doing so the amount of layers that redundant calibration pipelines need are effectively reduced. We also compare two recent redundant calibration speedup algorithms
-with each other, namely ADI \citep{Marthi2014} and PCG \citep{Liu2010}.
-In this paper, we decided to mainly focus on the complex framework proposed by \citet{Smirnov2015} rather than the traditional phase and amplitude approach as it
+In this paper we propose to use the complex framework introduced by \citet{Smirnov2015} instead of the traditional amplitude-and-phase approach used by \textsc{lincal}.
+We also propose to use the Levenberg-Marquardt (LM) gradient-based algorithm instead of the Gauss-Newton (GN) algorithm as it exhibits better convergence properties.
+Using the LM algorithm, reduces the number of algorithmic layers that are needed to perform redundant calibration. 
+We also compare two recent redundant calibration speedup algorithms with each other, namely the ADI \citep{Marthi2014} and PCG \citep{Liu2010} methods.
+In this paper, we decided to mainly focus on the complex framework proposed by \citet{Smirnov2015} rather than the traditional phase-and-amplitude approach as it
 allows us to trivially prove that the ADI method proposed by \citep{Marthi2014} is merely a special case of the more general LM algorithm. Moreover, this framework also 

### 2017-01-12 paper.tex
added dimension table

@@
 which is inherently quite faint \citep{Parsons2012}. Since redundant baselines make statistically independent measurements of the same $uv$-modes redundancy can greatly 
-improve the SNR (signal-to-noise ratio) of EoR detection experiments. This explains why PAPER and HERA, are highly redundant arrays, as they are both primarily used to detect the EoR 21 cm signal.
+improve the SNR (signal-to-noise ratio) of EoR detection experiments.
+This explains why PAPER and HERA, are highly redundant arrays, as they are both primarily used for EoR 21 cm detection experiments.
 
@@
 Two ways of speeding up redundant calibration have also been proposed recently. The first technique makes use of the ADI method, in which we alternate between solving the gains, their conjugates and 
-the true observed visibilities \citep{Wijnholds2012,Marthi2014}.
+the true observed visibilities \citep{Wijnholds2012,Marthi2014}. Note that the fact that \citet{Marthi2014} is making use of the ADI method is not immediately clear
+from the paper itself.
+We will elaborate on this later on in the paper (see Section~\ref{sec:results}).
 The second is to use the preconditioned conjugate gradient (PCG) method.
@@
 In this paper, we decided to mainly focus on the complex framework proposed by \citet{Smirnov2015} rather than the traditional phase-and-amplitude approach as it
-allows us to trivially prove that the ADI method proposed by \citep{Marthi2014} is merely a special case of the more general LM algorithm. Moreover, this framework also 
-allows us to show that the ADI method proposed by \citep{Marthi2014} can for all intents and purposes be renamed to redundant \textsc{StEfCal} \citep{Mitchell:MWA-cal,Salvini2014}.
-
-ADD SUMMARY PARAGRAPH WHICH DEPENDS ON PAPER LAYOUT
+allows us to trivially prove that the ADI method proposed by \citep{Marthi2014} is merely a special case of the more general LM algorithm.
+
+%Moreover, this framework also 
+%allows us to show that the ADI method proposed by \citep{Marthi2014} can for all intents and purposes be renamed to redundant \textsc{StEfCal} \citep{Mitchell:MWA-cal,Salvini2014}.
+
+We start the paper by presenting some useful preliminaries in Section~\ref{sec:preliminaries}. In Section~\ref{sec:p_statement}, we formulate redundant calibration 
+as a least squares problem, i.e. Eq~\eqref{eq:least_squares_red}. In Section~\ref{sec:main_sec}, we use the complex framework proposed by \citet{Smirnov2015} to derive a numerical
+algorithm (see Eq.~\eqref{eq:GN_update} to Eq.~\eqref{eq:update}), which we can use to solve Eq~\eqref{eq:least_squares_red}.
+In Section~\ref{sec:results}, we compare the computational complexity of two recently proposed redundant calibration speedup algorithms with one another.
+We end the paper with some final conclusions.
 
@@
 \section{Preliminaries}
-There are two major pre-existing algorithms we use throughout this paper, namely the Conjugate Gradient (CG) method and least-squares parameter estimation.
+\label{sec:preliminaries}
+There are two major pre-existing algorithms we use throughout this paper, namely the Conjugate Gradient (CG) method and the least-squares solver.
 We respectively discuss the CG method and least-squares in Section~\ref{sec:conj_grad} and Section~\ref{sec:ls}.
@@
 \While{$k\leq k_{\textrm{max}}$ and $\delta_{\textrm{new}} > \epsilon^2\delta_0$}
-\State $\alpha_k \gets \frac{\delta_{\textrm{new}}}{\bp^H\bA\bp}$
-\State $\bx \gets \bx + \alpha_k\bp$
+\State $\alpha \gets \frac{\delta_{\textrm{new}}}{\bp^H\bA\bp}$
+\State $\bx \gets \bx + \alpha\bp$
 \State $\br \gets \br - \alpha\bA\bp$
@@
 \State $\delta_{\textrm{new}} \gets \br^H\br$
-\State $\beta_{k} \gets \frac{\delta_{\textrm{new}}}{\delta_{\textrm{old}}}$
+\State $\beta \gets \frac{\delta_{\textrm{new}}}{\delta_{\textrm{old}}}$
 \State $\bp \gets \br + \beta\bp$
@@
 of CG (see Eq.~\eqref{eq:cg_bound}).
+In this paper, we denote matrix inversion with $(*)^{-1}$.
 The best possible choice of $\bM$ would therefore be $\bA$, which would lead to $\kappa$ becoming unity (as the largest and smallest eigenvalue of the Identity matrix is equal to one).
@@
 The most standard way of solving Eq.~\eqref{eq:least_squares} is to use a gradient type of minimization algorithm.
-In this paper we will be focusing on the following gradient minimization algorithms: Gauss-Newton and Levenberg-Marquardt \citep{Levenberg1944,Marquardt1963}.
+In this paper we will be focusing on the following gradient minimization algorithms: GN and LM \citep{Levenberg1944,Marquardt1963}.
 The GN update step is defined as
@@
 where $\bJ$ is the Jacobian matrix (i.e. the derivative of the model $\bnu$ with respect to the parameter vector $\bzeta$), $\lambda$ denotes the 
-damping factor, while $\bD$ is a diagonal matrix whose diagonal entries are equal to the diagonal entries of $\bJ^T\bJ$. In this paper we respectively denote 
-matrix transposition and matrix inversion with $(*)^T$ and $(*)^{-1}$.
+damping factor, while $\bD$ is a diagonal matrix whose diagonal entries are equal to the diagonal entries of $\bJ^T\bJ$. In this paper, we denote 
+matrix transposition with $(*)^T$.
 These update steps can be used in an iterative manner to determine new estimates of $\bzeta$.
@@
 \section{Problem Statement}
+\label{sec:p_statement}
 In Section~\ref{sec:s_mes}, we present the scalar measurement equation. Next, we define some indexing functions in Section~\ref{sec:indexing}. The measurement 
@@
 Eq.~\eqref{eq:least_squares_red}.
-We present the Signal-to-noise ratio (SNR) definition we used in this paper in Section\ref{sec:snr}. The general scope and limitations of this paper is discussed in Section~\ref{sec:scope}.
+We present the Signal-to-noise ratio (SNR) definition we used in this paper in Section~\ref{sec:snr}. The limitations of this paper are discussed in Section~\ref{sec:scope}.
 
@@
 Recall that redundant baselines sample the exact same visibilities in the $uv$-plane, i.e. if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$. 
-The aforementioned fact allows us to group all the redundant baselines in a redundant array into the same redundant baseline group. When our array 
-is a redundant array we therefore only require single index values for each redundant baseline group, instead of a unique single index label for 
+Let us now group all the baselines that measure the same $uv$-modes in a redundant array into the same redundant baseline group.
+Assigning the baselines of a redundant array to redundant baseline groups has the advantage that it reduces the number of single index values we require.
+We now only require single index values for each redundant baseline group, instead of a unique single index label for
 each baseline. Let $\phi_{pq}$ be the function that maps composite antenna pairs to the appropriate single redundant baseline group index. Note that $\phi_{pq}$ is not symmetric as 
@@
 0~\textrm{otherwise}
-\end{cases}
+\end{cases}.
 \end{equation}
@@
 \end{equation}
-where $<*>_{\nu,t,pq}$ denotes averaging over frequency, time and baselines and $\odot$ denotes the Hadamard product.
-The definition we use here is based on the SNR definitions used in \citet{Liu2010} and \citet{Marthi2014}.
-
-\subsection{Paper Scope}
+where $<*>_{\nu,t,pq}$ denotes averaging over frequency, time and baselines. The definition we use here is based on the SNR definitions used in \citet{Liu2010} and \citet{Marthi2014}.
+
+\subsection{Limitations}
 \label{sec:scope}
@@
 \end{equation}
-where $\breve{\bd} = [\bd^T,\conj{\bd}^T]^T$, $\breve{\bv} = [\bv^T,\conj{\bv}^T]$ and $\bmath{\varrho}$ is the phase and amplitude parameter vector which we define in Appendix~\ref{sec:lincal}. 
-With the aid of Eq.~\eqref{eq:least_squares_lincal} we can derive the \textsc{lincal} algorithm (see Appendix~\ref{sec:lincal}).
+where $\breve{\bd} = [\bd^T,\conj{\bd}^T]^T$, $\breve{\bv} = [\bv^T,\conj{\bv}^T]^T$ and $\bmath{\varrho}$ is the phase and amplitude parameter vector which we define in Appendix~\ref{sec:lincal}. 
+With the aid of Eq.~\eqref{eq:least_squares_lincal} we can derive the algorithm implemented in \textsc{lincal} (see Appendix~\ref{sec:lincal}).
 The Jacobian associated with Eq.~\eqref{eq:least_squares_lincal} is equal to Eq.~\eqref{eq:Jac_lin}. 
@@
 Therefore, a straightforward way of improving the convergence properties of stock standard \textsc{lincal} would be to simply add the LM damping factor to the basic \textsc{lincal} update step.
+In doing so the need to use $\textsc{logcal}$ before $\textsc{lincal}$ becomes superfluous. The use of $\textsc{logcal}$ may, however, still offer 
+a computational advantage, as it is a very fast method to obtain a good starting solution, and would reduce the number of iterations needed by the LM algorithm.
+Proper experiments would need to be conducted to determine whether it is worth keeping $\textsc{logcal}$ in the algorithmic hierarchy.
 
@@
 \label{sec:complex_GN_LM}
-In this section we discuss how we can use the GN and LM algorithms to solve Eq.~\eqref{eq:least_squares_complex} (i.e. estimate $\breve{\bz}$). The complex Jacobian is defined to be
+The complex Jacobian is defined to be
 \begin{equation}
@@
 \subsubsection{Analytic Expressions}
+\label{sec:analytic}
 If we apply the definition in Eq.~\eqref{eq:Jacobian} to Eq.~\eqref{eq:least_squares_complex} we obtain the following analytic result:
@@
 \begin{cases}
- \sum_{pq \in \mathcal{PQ}_i} \left | g_p \right |^2 \left | g_q \right |^2  & \textrm{if} ~ i=j\\
+ \sum_{rs \in \mathcal{RS}_i} \left | g_r \right |^2 \left | g_s \right |^2  & \textrm{if} ~ i=j\\
  0 & \textrm{otherwise}
@@
 \end{equation}
-Moreover, $(*)^T$ denotes matrix transposition and
-\begin{equation}
-\mathcal{PQ}_i = \left\{pq\in\mathbb{N}^2|(\phi_{pq} = i) \right\}.
+Moreover, 
+\begin{equation}
+\mathcal{RS}_i = \left\{rs\in\mathbb{N}^2|(\phi_{rs} = i) \right\}.
 \end{equation}
@@
 \label{eq:ab}
-\left [ \ba \right ]_i &= \sum_{k\neq i} g_k \widetilde{y}_{ik}r_{ik},  & \left [ \bb \right ]_i &= \sum_{pq\in\mathcal{PQ}_i}\conj{g}_p g_q r_{pq},
+\left [ \ba \right ]_i &= \sum_{k\neq i} g_k \widetilde{y}_{ik}r_{ik},  & \left [ \bb \right ]_i &= \sum_{rs\in\mathcal{RS}_i}\conj{g}_r g_s r_{rs},
 \end{align}
@@
 \end{equation}
-Moreover $\ba$ and $\bb$ are both column vectors.
-
-HAVE TO RECHECK THE ABOVE EQUATIONS DIRECTLY FROM THE CODE.
-DEFINED ALL OPERATORS AND SYMBOLS?
+Moreover $\ba$ and $\bb$ are both column vectors. The length of $\ba$ and $\bb$ is $N$ and $L$, respectively.
+The dimensions of the matrices we defined in this section are presented in Table~\ref{tab:matrix_dimensions}.
 
@@
 % \end{equation}
+
+\begin{table}
+\centering
+\caption{The dimensions of the matrices defined in Section~\ref{sec:analytic}.}
+\begin{tabular}{|c c|} 
+\hline
+Matrix & Dimension\\
+\hline
+\hline
+$\bJ$ & $2B \times P$ \\
+$\bM$ & $B \times (N+L)$ \\
+$\bN$ & $B \times (N+L)$ \\
+$\bO$ & $B \times N$ \\
+$\bP$ & $B \times L$ \\
+$\bQ$ & $B \times N$ \\
+$\bzero$ & $B \times L$ \\
+\hline
+\hline
+$\bH$ & $P\times P$\\
+$\bA$ & $(N+L)\times (N+L)$\\
+$\bB$ & $(N+L)\times (N+L)$\\
+$\bC$ & $N \times N$\\
+$\bD$ & $N \times L$\\
+$\bE$ & $L \times L$\\
+$\bF$ & $N \times N$\\
+$\bG$ & $N \times L$\\
+$\bzero$ & $L \times L$\\
+\hline
+\end{tabular}
+\label{tab:matrix_dimensions}
+\end{table}
 

### 2017-01-12 paper.tex
newest version

@@
 \newcommand{\bzero}{\bmath{0}}
+\newcommand{\bone}{\bmath{1}}
 %\newcommand{\bvarrho}{\bmath{\varrho}}
@@
 from the paper itself.
-We will elaborate on this later on in the paper (see Section~\ref{sec:results}).
+We will elaborate on this later on in Section~\ref{sec:adi}.
 The second is to use the preconditioned conjugate gradient (PCG) method.
@@
 \label{sec:results}
-In the previous section we discussed the complex GN and LM algorithms and how they relate to redundant calibration. If we inspect Eq.~\eqref{eq:GN_update} and 
-Eq.~\eqref{eq:GN_update} we see that the most expensive operation used by the GN and LM algorithms is the inversion of the Hessian $\bH$. The most logical 
+In Section~\ref{sec:main_sec} we discussed the complex GN and LM algorithms and how they relate to redundant calibration. If we inspect Eq.~\eqref{eq:GN_update} and 
+Eq.~\eqref{eq:LM_update} we see that the most expensive operation used by the GN and LM algorithms is the inversion of the Hessian $\bH$. The most logical 
 way of speeding up the algorithm is to investigate whether we can do this inversion in some efficient way.
@@
 Note that here we focus solely on comparing the computational complexity of inverting $\bH$ using either one of these two methods. Rating the LM or GN algorithms
-using one of these approaches relative to each other once the computation complexity of inverting $\bH$ is known is trivial. The simulations we performed to compare these two 
+using one of these approaches relative to each other once the computational complexity of inverting $\bH$ is known is trivial. The simulations we performed to compare these two 
 methods are summarized in Section~\ref{sec:simulations}.
@@
 $\bH$, i.e. it becomes $\mathbb{O}(P)$\footnote{The cost of constructing 
-the diagonal is not included in this estimate.} \citep{Smirnov2015}. This diagonal-approximation approach is effectively ADI \citep{Marthi2014}. The exact details of this 
-approach is discussed in Section~\ref{sec:adi}.
+the diagonal is not included in this estimate.} \citep{Smirnov2015}. This diagonal-approximation approach is effectively ADI \citep{Marthi2014}. The exact details of why this 
+is the case is discussed in Section~\ref{sec:adi}.
 We will also show how \citet{Marthi2014}, \citet{Smirnov2015} and \citet{Salvini2014} relate to each other in that same section.
-It should be noted that the ADI method actually refers to the complete LM update step using the approximate $\bH$ and not just
-to the inversion of the approximate $\bH$. Furhtermore, the name ADI also implies that with this algorithm we alternate between computing the 
-gains, their conjugates and the true observed visibilities. %In Appendix~\ref{sec:sbc} we briefly summarize the ADI method that is used for skymodel-based calibration.
+It should be noted here that strictly speaking the ADI method actually refers to the complete LM update step using the approximate $\bH$ and not just
+to the inversion of the approximate $\bH$.
+To improve the readability of the paper, we use the acronym ADI to refer to both the faster approximate inverse approach and the complete LM algorithm that uses it.
+Which one of these two concepts we are actually referring to, however, is always perfectly clear from the context.
+%Furhtermore, the name ADI also implies that with this algorithm we alternate between computing the 
+%gains, their conjugates and the true observed visibilities. %In Appendix~\ref{sec:sbc} we briefly summarize the ADI method that is used for skymodel-based calibration.
 
@@
 Even though the PCG method is computationally more expensive than the ADI method we can not arrive at a final conclusion at this point. With the ADI method 
-we only approximate the inverse of the Hessian, while in the case of the PCG method we compute the exact inverse. So the question arises whether the PCG method 
+we only approximate the inverse of the Hessian, while in the case of the PCG method we compute the exact inverse. So the question arises whether the LM algorithm using the PCG method 
 will converge much faster than the ADI method even though each iteration (which includes an inversion) is more expensive. We investigate this question in Section~\ref{sec:comparison}.
@@
 
-The main flow of a single simulation is as follow. We created corrupted visibilities, using either one of the experimental setups summarized in Table~\ref{tab:ch_parm} and
-the fixed simulation parameters presented in Table~\ref{tab:fixed_parm}. We then used per time-slot and channel redundant calibration, i.e. we either used the ADI or PCG algorithm, to 
-estimate the gains and the true observed visibilities. To obtain statistically significant results we performed, for both the ADI and PCG algorithms, multiple statistical independent simulations.
+The main flow of a single simulation is as follow.
+We created corrupted visibilities, using the fixed simulation parameters presented in Table~\ref{tab:fixed_parm} and either one of the experimental setups summarized in Table~\ref{tab:ch_parm}.
+We then used per time-slot and channel redundant calibration, i.e.
+we use the LM algorithm which in turn makes use of either the ADI or PCG method, to
+estimate the gains and the true observed visibilities. To obtain statistically significant results we performed, for both the ADI and PCG methods, multiple statistical independent simulations.
 
@@
 Gain error model (Table~\ref{tab:gain_parm}) & 2 & 1\\
+Initial guess: $\bz_0$ & $\bone$ & $\bone$\\
 \hline
@@
 %In Section~\ref{sec:adi} we show how the ADI method used for skymodel-based calibration relates to the redundant ADI method presented in \citet{Marthi2014}. 
-
 We start this section by briefly reviewing the ADI method as it pertains to skymodel-based calibration in Section~\ref{sec:sbc} \citep{Salvini2014,Smirnov2015}.
-In Section~\ref{sec:red_c} we re-derive the redundant calibration ADI method presented in \citet{Marthi2014}. In \citet{Marthi2014} the ADI method is derived by taking the derivative of the 
+
+In Section~\ref{sec:red_c}, we re-derive the fast redundant calibration algorithm presented in \citet{Marthi2014} using the complex framework proposed by \citet{Smirnov2015}.
+In \citet{Marthi2014} it is derived by taking the derivative of the
 objective function $\Lambda$ relative to $\bg$ and $\by$ and setting the result to naught. When derived in this way, its relation to the recent works \citep{Salvini2014,Smirnov2015} are unclear.
-We will use the diagonal-approximate approach proposed in \citet{Smirnov2015} to show that the ADI algorithm presented in \citet{Marthi2014} is in actual fact redundant 
-\textsc{StEfCal} \citep{Salvini2014}.
+To elaborate a little, the conclusion that the algorithm presented in \citet{Marthi2014} is nothing more than a redundant extension of the skymodel-based ADI method (i.e.
+redundant \textsc{StEfCal}) is in actual fact not obvious either (something we only briefly mentioned in the introduction and then for the sake of simplicity and convenience chose to ignore until now) if the aforementioned
+partial derivative strategy is used to derive it. In Section~\ref{sec:red_c}, we will, therefore, also explicitly show that 
+the redundant calibration algorithm presented in \citet{Marthi2014} is in actual fact nothing more than an ADI method and in doing so we will have justified the naming 
+convention we have used throughout the paper.
 
@@
 \label{sec:sbc}
-\textsc{StEFCal} is an alternating direction implicit method, in which the measurement equation, Eq.~\eqref{eq:vis_definition}, is linearized by assuming that the gains are known, but
+\textsc{StEFCal} is an ADI method, in which the measurement equation, Eq.~\eqref{eq:vis_definition}, is linearized by assuming that the gains are known, but
 that their conjugates are unknown \citep{Mitchell:MWA-cal,Salvini2014}.
@@
 \end{align}
+where $\alpha = \frac{1}{1+\lambda}$. Moreover, $\lambda = \frac{1-\alpha}{\alpha}$.
 
@@
 sense to try and approximate it with its diagonal (see Fig.~\ref{fig:hessian} and Eq.~\eqref{eq:red_H}).
-If we substitute $\bJ^H\bJ$ with $\widetilde{\bH}$ (by its diagonal approximation) and replace $\breve{\br}$ with $\breve{\bd} - \breve{\bv}$ in Eq.~\eqref{eq:GN_update} we obtain
+If we substitute $\bJ^H\bJ$ with $\widetilde{\bH}$ and replace $\breve{\br}$ with $\breve{\bd} - \breve{\bv}$ in Eq.~\eqref{eq:GN_update} we obtain
 \begin{equation}
@@
 \breve{\bz}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bz}_k,\label{eq:lambda}\\
- &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bz}_k, \label{eq:alpha}  
-\end{align}
-where $\alpha = \frac{1}{1+\lambda}$. Moreover, $\lambda = \frac{1-\alpha}{\alpha}$.
+ &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bz}_k. \label{eq:alpha}  
+\end{align}
 The analytic expression of $\bJ^H\breve{\bd}$ will be very similar to the analytic 
@@
 Eq.~\eqref{eq:g_update} and Eq.~\eqref{eq:y_update} are the exact same iterative new parameter estimators derived by \citet{Marthi2014}. Also note that Eq.~\eqref{eq:g_update} is the fundamental
-gain estimator associated with \textsc{StEfCal}. We have therefore successfully used the diagonal approximation approach proposed by \citet{Smirnov2015} to re-derive the ADI method presented in 
-\citet{Marthi2014} and in doing so we have shown that the algorithm in \citet{Marthi2014} is actually redundant StEfCal. We could also have shown this by using the alternating approach 
-presented in \citet{Salvini2014}; which explains why, in this paper, we have been referring to the algorithm presented in \citet{Marthi2014} as the ADI method.
+gain estimator associated with \textsc{StEfCal}.
+We have therefore successfully used the diagonal approximation approach proposed by \citet{Smirnov2015} to re-derive the fast redundant calibration method presented in
+\citet{Marthi2014} and in doing so we have shown that this algorithm is in actual fact nothing more than the ADI method (i.e.
+redundant \textsc{StEfCal}), which justifies the naming convention we have been using for this method.
+We could also have shown this same result by using
+a more direct approach. We could simply have used the alternating strategy presented in the skymodel-based ADI paper itself \citep{Salvini2014}.
 
@@
 Optimality results of ADI were presented in \citet{Marthi2014}. We do not repeat their analysis here. We however, do present our own ADI results in Fig.~\ref{fig:prec_error} that confirm that the
-ADI method can accurately estimate the antenna gains and the observed visibilities at different SNR values. To quantity the exact error in our estimates we used the precentage error $\beta$ defined 
+ADI method can accurately estimate the antenna gains and the observed visibilities at different SNR values. To quantity the exact error in our estimates we used the percentage error $\beta$ defined 
 as 
@@
 \includegraphics[width=0.47\textwidth]{./prec_error.pdf} 
-\caption{We plot the precentage error $\beta$ between the estimated 
+\caption{We plot the percentage error $\beta$ between the estimated 
 predicted visibilities and the true predicted visibilities for different SNR values as a function of the number of antennas in the array $N$.

### 2017-01-13 paper.tex
tried fixing the chicken egg probelm

@@
 
-In Section~\ref{sec:red_c}, we re-derive the fast redundant calibration algorithm presented in \citet{Marthi2014} using the complex framework proposed by \citet{Smirnov2015}.
-In \citet{Marthi2014} it is derived by taking the derivative of the
-objective function $\Lambda$ relative to $\bg$ and $\by$ and setting the result to naught. When derived in this way, its relation to the recent works \citep{Salvini2014,Smirnov2015} are unclear.
-To elaborate a little, the conclusion that the algorithm presented in \citet{Marthi2014} is nothing more than a redundant extension of the skymodel-based ADI method (i.e.
-redundant \textsc{StEfCal}) is in actual fact not obvious either (something we only briefly mentioned in the introduction and then for the sake of simplicity and convenience chose to ignore until now) if the aforementioned
-partial derivative strategy is used to derive it. In Section~\ref{sec:red_c}, we will, therefore, also explicitly show that 
-the redundant calibration algorithm presented in \citet{Marthi2014} is in actual fact nothing more than an ADI method and in doing so we will have justified the naming 
-convention we have used throughout the paper.
+In Section~\ref{sec:red_c}, we will show that the fast redundant calibration algorithm, that is presented in \citet{Marthi2014},
+is in fact nothing more than an extension of the skymodel-based ADI method, presented in \citet{Salvini2014}, to the redundant calibration use case (i.e. redundant \textsc{StEfCal}).
+In doing so we will justify the naming convention we have been using up until now.
+
+In Section~\ref{sec:adi_results}, we present some ADI simulation results. 
+%We will mainly accomplish this by using the complex framework proposed by \citet{Smirnov2015}.
 
@@
 \label{sec:red_c}
-Since the diagonal entries of the $\bH$ associated with redundant calibration are more significant than its off-diagonal entries it makes 
+In this section, we re-derive the fast redundant calibration algorithm presented in \citet{Marthi2014} using a different approach to the one presented in the paper itself.
+In \citet{Marthi2014}, the algorithm is derived by taking the derivative of the
+objective function $\Lambda$ relative to $\bg$ and $\by$ and setting the result to naught.
+
+It is in fact trivial to extend the ADI method presented in \citet{Salvini2014} to the redundant use case, a.k.a.
+we now alternate between solving for the antenna gains their conjugates and the true observed visibilities.
+If we do so we are able to re-derive the redundant calibration algorithm presented in \citet{Marthi2014}.
+This implies that the fast redundant calibration algorithm presented in \citet{Marthi2014} is effectively redundant \textsc{StEfCal} (i.e.
+some kind of ADI method).
+
+Recall, that \citet{Smirnov2015} showed that \textsc{StEfCal} can also be derived by approximating the $\bH$, associated with skymodel-based calibration, with 
+its diagonal. We will now use the diagonal approximation approach presented in \citet{Smirnov2015} to yet again re-derive the redundant calibration algorithm presented in \citet{Marthi2014}.
+
+Since the diagonal entries of the $\bH$, associated with redundant calibration, are more significant than its off-diagonal entries it makes 
 sense to try and approximate it with its diagonal (see Fig.~\ref{fig:hessian} and Eq.~\eqref{eq:red_H}).
@@
 Eq.~\eqref{eq:g_update} and Eq.~\eqref{eq:y_update} are the exact same iterative new parameter estimators derived by \citet{Marthi2014}. Also note that Eq.~\eqref{eq:g_update} is the fundamental
-gain estimator associated with \textsc{StEfCal}.
-We have therefore successfully used the diagonal approximation approach proposed by \citet{Smirnov2015} to re-derive the fast redundant calibration method presented in
-\citet{Marthi2014} and in doing so we have shown that this algorithm is in actual fact nothing more than the ADI method (i.e.
-redundant \textsc{StEfCal}), which justifies the naming convention we have been using for this method.
-We could also have shown this same result by using
-a more direct approach. We could simply have used the alternating strategy presented in the skymodel-based ADI paper itself \citep{Salvini2014}.
+gain estimator associated with skymodel-based \textsc{StEfCal}.
+We have, therefore, successfully used the diagonal approximation approach proposed by \citet{Smirnov2015} to re-derive the fast redundant calibration method presented in
+\citet{Marthi2014}.
+%and in doing so we have confirmed shown that this algorithm is in actual fact nothing more than redundant \textsc{StEfCal}, i.e.
+some form of ADI method, justifying the naming convention we have been using for it.
+Re-deriving the algorithm presented in \citet{Marthi2014} in this way serves as a further confirmation that it is indeed nothing more than redundant \textsc{StEfCal} (i.e. some kind of ADI method).
 
@@
 
-Optimality results of ADI were presented in \citet{Marthi2014}. We do not repeat their analysis here. We however, do present our own ADI results in Fig.~\ref{fig:prec_error} that confirm that the
+\subsubsection{ADI simulation results}
+\label{sec:adi_results}
+Optimality results of ADI were presented in \citet{Marthi2014}. We do not repeat their analysis here. We however, do present our own ADI results, in Fig.~\ref{fig:prec_error}, that confirm that the
 ADI method can accurately estimate the antenna gains and the observed visibilities at different SNR values. To quantity the exact error in our estimates we used the percentage error $\beta$ defined 
@@
 \label{eq:beta}
-\beta = \frac{\sum_{pq} |v_{pq} - \widehat{v}_{pq} |^2}{\sum_{pq} |v_{pq}|^2}.
-\end{equation}
-We have already defined $v_{pq}$ in Eq.~\eqref{eq:vec_definitions}. Moreover, in Eq.~\eqref{eq:beta}, $\widehat{v}_{pq}$ denotes the ADI estimate of $v_{pq}$.
+\beta = \frac{\|\bv - \widehat{\bv}\|_F^2}{\|\bv\|_F^2}.
+\end{equation}
+We have already defined $\bv$ in Eq.~\eqref{eq:vec_definitions}. Moreover, in Eq.~\eqref{eq:beta}, $\widehat{\bv}$ denotes the ADI estimate of $\bv$.
 

### 2017-01-13 paper.tex
more chicken egg corrections

@@
 Recall, that \citet{Smirnov2015} showed that \textsc{StEfCal} can also be derived by approximating the $\bH$, associated with skymodel-based calibration, with 
-its diagonal. We will now use the diagonal approximation approach presented in \citet{Smirnov2015} to yet again re-derive the redundant calibration algorithm presented in \citet{Marthi2014}.
+its diagonal.
+To double check if our statement that the algorithm in \citet{Marthi2014} is in actual fact redundant \textsc{StEfCal} (i.e.
+some kind of ADI method) we will now use the diagonal approximation approach presented in \citet{Smirnov2015} to yet again re-derive the redundant calibration algorithm presented in \citet{Marthi2014}.
 
@@
 \textsc{StEfCal}) is presented
-in \citet{Salvini2014}. \citet{Smirnov2015} showed that it is possible to rederive this skymodel-based ADI method through complex optimization. In particular, they showed that if 
+in \citet{Salvini2014}. \citet{Smirnov2015} showed that it is possible to re-derive this skymodel-based ADI method through complex optimization. In particular, they showed that if 
 we approximate the $\bH$ associated with the LM algorithm with its diagonal then we can rederive the ADI update rules presented in \citet{Salvini2014}. 
-\citet{Marthi2014} applied the ADI method to the redundant use case.
-This fact is not immediatly clear when inspecting \citet{Marthi2014}.
-The ADI update rules presented \citet{Marthi2014} were derived by taking the partial derivatives of
-the objective function towards the unknown parameters, setting the intermediate results to zero and then solving for the unknowns. In this paper we show that 
-these same ADI update rules can be derived using the diagonal approximation technique proposed in \citet{Smirnov2015}. The fact that we can use this diagonal approximation
-technique to derive these update rules imply that the algorithm presented in \citet{Marthi2014} is in fact the ADI method (i.e. redundant \textsc{StEfCal}). This same
-fact can be established by directly applying the skymodel-based ADI derivation presented in \citet{Salvini2014} to the redundant use case.
+\citet{Marthi2014} actually proposes to use the ADI method to perform redundant calibration.
+This fact is not immediately clear when superficially inspecting \citet{Marthi2014}.
+The redundant calibration algorithm presented in \citet{Marthi2014} was derived by taking the partial derivatives of
+the objective function towards the unknown parameters, setting the intermediate results to zero and then solving for the unknowns. In Section~\ref{sec:red_c}, we showed that 
+this algorithm is in actual fact an extension of the skymodel-based ADI method presented in \citet{Salvini2014} to the redundant calibration use case (i.e.
+we showed that this algorithm is in actual fact redundant \textsc{StEfCal}).
+We also re-derived the aforementioned redundant calibration algorithm by applying the diagonal approximation approach presented in \citet{Smirnov2015}, which serves as yet another additional confirmation
+that this algorithm is in fact redundant \textsc{StEfCal} (i.e. that it is some kind of ADI method).
 

### 2017-01-24 paper.tex
fixed Jonathan's comments

@@
 \begin{abstract}
-Traditionally redundant calibration is implemented via a hierarchical algorithmic strategy.  
-We show that instead of splitting our redundant calibration problem into its phase-and-amplitude components,
-which is the more traditional approach, we can treat the complex unknown model parameters and their complex conjugates as separate unknowns instead. 
+In this paper we show that instead of splitting our redundant calibration problem into its phase-and-amplitude components (the more traditional approach) we can treat the complex unknown model parameters and their complex conjugates as separate unknowns.
 We also derive the Levenberg-Maquardt and Gauss-Newton update steps associated with this alternative parameter-and-conjugate framework.
 We focus more on the Levenberg-Maquardt algorithm in this paper as it exhibits better convergence properties when compared to Gauss-Newton.
+Traditionally redundant calibration is implemented via a hierarchical algorithmic strategy.
 Using the Levenberg-Maquardt algorithm, therefore, reduces the number of algorithmic layers that are needed to perform redundant calibration.
 The Gauss-Newton algorithm is, however, still the de facto standard that is currently being used to perform redundant calibration.
-We also compare two previously proposed methods to speedup redundant calibration, namely the alternating direction implicit method and the preconditioned conjugate gradient method.
+We also compare two previously proposed methods to accelerate redundant calibration, namely the alternating direction implicit method and the preconditioned conjugate gradient method.
 We found that the alternating
-direction implicit method outperforms the preconditioned conjugate gradient method, however the preconditioned conjugate gradient method is more robust as it is more straightforward to adapt to other calibration use cases.
+direction implicit method outperforms the preconditioned conjugate gradient method.
+However, the preconditioned conjugate gradient method is more robust as it is more straightforward to adapt to other calibration use cases.
 We make use of the parameter-and-conjugate framework in this paper, rather than the more traditional amplitude-and-phase approach as it makes it trivial to relate the alternating direction implicit method
@@
 \section{Introduction}
-The propogation effects (i.e.
-the ionosphere, antenna gains, etc.) a celestial signal experiences along its propagation path is modelled with the radio interferometric measurement equation (RIME) \citep{ME1,RRIME1}.
-Calibration is the procedure by which we remove the errors that these propagation effects induce onto our interferometric data and is normally accomplished using a 
+The effects (i.e.
+the ionosphere, antenna gains, etc.) a celestial signal experiences along its propagation path are modelled with the radio interferometric measurement equation (RIME) \citep{ME1,RRIME1}.
+Calibration is the procedure by which we remove the errors that these propagation effects introduce into interferometric data and is normally accomplished using a 
 least-squares solver.
-In the case where we only model the antenna gains, which is the most basic form of the RIME, we can define calibration as finding the antenna gains that minimizes the difference betweem our observed and predicted visibilities.
+In the case where we only model the antenna gains, which is the most basic form of the RIME, we can define calibration as finding the antenna gains that minimize the difference between our observed and predicted visibilities.
 
@@
 This procedure results in the linearization of the calibration algorithm and reduces the computational complexity of calibration from $\mathbb{O}(N^3)$ to $\mathbb{O}(N^2)$, where $N$ denotes the number of antennas in the array.
-The second development is the idea: to not break the calibration problem 
-into its real and imaginary parts and then to solve for the real and imaginary parts of the unknown parameters separately, but to rather treat the antenna gains and their 
-conjugates as seperate parameters to estimate \citep{Smirnov2015}. \citet{Smirnov2015} also showed that the \textsc{StEfCal} algorithm itself is merely a special case of this
+The second development is the idea: instead of breaking the calibration problem 
+into its real and imaginary parts and then solving for the real and imaginary parts of the unknown parameters separately we can treat the antenna gains and their 
+conjugates as separate parameters to estimate \citep{Smirnov2015}. \citet{Smirnov2015} also showed that the \textsc{StEfCal} algorithm itself is merely a special case of this
 gain-and-conjugate framework.
@@
 Two baselines are redundant if their baseline difference vectors have the same length and orientation, which imply that they measure the same $uv$-modes.
-When we use a redundant array layout we dramatically
-reduce the number of unknowns. If our array is redundant enough we reduce the number of unknowns to such an extend that we are able to solve for the unknown 
-observed visibilities themselves in addition to the antenna gains. Redundant calibration is the procedure by which we estimate the unknown observed visibilities in addition
-to the antenna gains.
+When we use a redundant array layout, we dramatically
+reduce the number of unknowns. If our array is redundant enough we reduce the number of unknowns to such an extent that we are able to solve for the unknown 
+observed visibilities themselves in addition to the antenna gains. Redundant calibration is the procedure by which we estimate both the unknown observed visibilities and the antenna gains.
 
@@
 We briefly discuss \textsc{logcal} in Appendix~\ref{sec:logcal}.
-The next improvement in redundant calibration was to perform a firt-order taylor expansion, which is equivalent to the Gauss-Newton (GN) algorithm \citep{Kurien2016}, of the measurement equation and then to solve for the
+The next improvement in redundant calibration was to perform a first order Taylor expansion, which is equivalent to the Gauss-Newton (GN) algorithm \citep{Kurien2016}, of the measurement equation and then to solve for the
 perturbations of the model-parameters \citep{Liu2010}.
@@
 The solutions of \textsc{logcal} are used as initial input for \textsc{lincal}, since the GN algorithm requires that its starting parameter guesses be in the vicinity 
-of the solution.
-
-New instruments like LOFAR (Low Frequency Array) \citep{Noorishad2012}, PAPER (Precision Array for Probing the Epoch of Reionization) \citep{Ali2015} and HERA (Hydrogen Epoch of Reionization Array) \citep{deboer2015} have also recently sparked a renewed interest in redundant calibration.
+of the true solution.
+
+New instruments like LOFAR (Low Frequency Array) \citep{Noorishad2012}, PAPER (Precision Array for Probing the Epoch of Reionization) \citep{Ali2015} and HERA (Hydrogen Epoch of Reionization Array) \citep{deboer2015} have also sparked a renewed interest in redundant calibration.
 Using redundancy is especially advantageous when we are trying to detect the EoR (Epoch of Reionization) 21 cm signal,
 which is inherently quite faint \citep{Parsons2012}. Since redundant baselines make statistically independent measurements of the same $uv$-modes redundancy can greatly 
-improve the SNR (signal-to-noise ratio) of EoR detection experiments.
-This explains why PAPER and HERA, are highly redundant arrays, as they are both primarily used for EoR 21 cm detection experiments.
+improve the SNR (signal-to-noise ratio) of EoR detection experiments. This explains why PAPER and HERA are highly redundant arrays, as they are both primarily used for EoR 21 cm detection experiments.
 
@@
 
-Two ways of speeding up redundant calibration have also been proposed recently. The first technique makes use of the ADI method, in which we alternate between solving the gains, their conjugates and 
-the true observed visibilities \citep{Wijnholds2012,Marthi2014}. Note that the fact that \citet{Marthi2014} is making use of the ADI method is not immediately clear
+Two ways of speeding up redundant calibration have recently been proposed. The first technique makes use of the ADI method, in which we alternate between solving the gains, their conjugates and 
+the true observed visibilities \citep{Wijnholds2012,Marthi2014}. Note that the fact that \citet{Marthi2014} make use of the ADI method is not immediately clear
 from the paper itself.
-We will elaborate on this later on in Section~\ref{sec:adi}.
+We will elaborate on this in Section~\ref{sec:adi}.
 The second is to use the preconditioned conjugate gradient (PCG) method.
-PCG allows us to exploit the sparsity inherent in the redundant calibration problem,
+PCG allows us to exploit the sparsity inherent in the redundant calibration problem
 to speed it up \citep{Liu2010}. A proper comparison between these two approaches has not been done before.
+
+It is also interesting to mention here that a completely different calibration strategy can also be used to calibrate redundant arrays, \textsc{CorrCal} \citep{Sievers2017}\footnote{https://github.com/sievers/corrcal2}.
+In \textsc{CorrCal} the unknown
+visibilities are no longer solved for. The sky is actually modelled, via two covariance matrices. 
+The first covariance matrix is associated with known point sources. The second covariance matrix models the remaining sky in a statistical sense. The unknown gains, however, are still 
+estimated. We do not elaborate any further on this approach in this paper.
 
@@
 We also propose to use the Levenberg-Marquardt (LM) gradient-based algorithm instead of the Gauss-Newton (GN) algorithm as it exhibits better convergence properties.
-Using the LM algorithm, reduces the number of algorithmic layers that are needed to perform redundant calibration. 
+Using the LM algorithm reduces the number of algorithmic layers that are needed to perform redundant calibration. 
 We also compare two recent redundant calibration speedup algorithms with each other, namely the ADI \citep{Marthi2014} and PCG \citep{Liu2010} methods.
-In this paper, we decided to mainly focus on the complex framework proposed by \citet{Smirnov2015} rather than the traditional phase-and-amplitude approach as it
+In this paper, we decided to focus on the complex framework proposed by \citet{Smirnov2015} rather than the traditional phase-and-amplitude approach as it
 allows us to trivially prove that the ADI method proposed by \citep{Marthi2014} is merely a special case of the more general LM algorithm.
@@
 We start the paper by presenting some useful preliminaries in Section~\ref{sec:preliminaries}. In Section~\ref{sec:p_statement}, we formulate redundant calibration 
-as a least squares problem, i.e. Eq~\eqref{eq:least_squares_red}. In Section~\ref{sec:main_sec}, we use the complex framework proposed by \citet{Smirnov2015} to derive a numerical
+as a least squares problem, i.e. Eq~\eqref{eq:least_squares_red}. In Section~\ref{sec:main_sec} we use the complex framework proposed by \citet{Smirnov2015} to derive a numerical
 algorithm (see Eq.~\eqref{eq:GN_update} to Eq.~\eqref{eq:update}), which we can use to solve Eq~\eqref{eq:least_squares_red}.
@@
 In this paper, we denote matrix inversion with $(*)^{-1}$.
-The best possible choice of $\bM$ would therefore be $\bA$, which would lead to $\kappa$ becoming unity (as the largest and smallest eigenvalue of the Identity matrix is equal to one).
+The best possible choice of $\bM$ would therefore be $\bA$, which would lead to $\kappa$ becoming unity (as the largest and smallest eigenvalue of the Identity matrix are equal to one).
 If $\kappa$ is one then Eq.~\eqref{eq:cg_bound} simplifies and becomes $\mathbb{O}(m)$.
@@
 
-If the diagonal entries of $\bA$ is much more significant than its off-diagonal entries then the 
+If the diagonal entries of $\bA$ are much more significant than its off-diagonal entries then the 
 Jacobian preconditioner is a good choice of $\bM$. The Jacobian preconditioner is computed as follows:
@@
 Recall that we mentioned that in general we can only solve Eq.~\eqref{eq:linear_system} with CG if $\bA$ is a square positive-definite Hermitian matrix.
-The CG method actually has an even broader applicability, it can be used to solve Eq.~\eqref{eq:linear_system} even if $\bA$ is an Hermitian positive semi-definite
+The CG method actually has an even broader applicability: it can be used to solve Eq.~\eqref{eq:linear_system} even if $\bA$ is a Hermitian positive semi-definite
 square matrix as long as $\bb$ is in the column range of $\bA$ \citep{Lu2015}.
 In light of \citet{Lu2015}, the two observations in the numbered list we presented in the previous paragraph imply that Eq.~\eqref{eq:normal_equation}
-can be solved by employing the CG method. The solution returned by CG is however only unique if $\bB^H\bB$ is positive-definite (invertible). The solution
-obtained by CG, whether it be unique or not, does however minimize Eq.~\eqref{eq:norm_ls} in a least squares sense.
-
-Note that if $\bA$ is positive semi-definite its smallest eigenvalue is zero, which implies that Eq.~\eqref{eq:kappa} is not defined. \citet{Lu2015} show that when $\bA$
+can be solved by employing the CG method. The solution returned by CG is, however, only unique if $\bB^H\bB$ is positive-definite (invertible). The solution
+obtained by CG, whether it is unique or not, does minimize Eq.~\eqref{eq:norm_ls} in a least squares sense.
+
+Note that if $\bA$ is positive semi-definite, its smallest eigenvalue is zero. This implies that Eq.~\eqref{eq:kappa} is not defined. \citet{Lu2015} show that when $\bA$
 is a square positive semi-definite Hermitian matrix we do not use the spectral condition number to bound the convergence complexity of CG, but the general spectral condition
@@
 In Eq.~\eqref{eq:least_squares}, we assume that the optimization problem is real valued.
-The main objective of least squares is to find the model-parameters which minimizes the squared difference between the observed data and the model-predicted data.
-
-The most standard way of solving Eq.~\eqref{eq:least_squares} is to use a gradient type of minimization algorithm.
+The main objective of least squares is to find the model-parameters which minimize the squared difference between the observed data and the model-predicted data.
+
+The most standard way of solving Eq.~\eqref{eq:least_squares} is to use a gradient-type minimization algorithm.
 In this paper we will be focusing on the following gradient minimization algorithms: GN and LM \citep{Levenberg1944,Marquardt1963}.
@@
 where $\bJ$ is the Jacobian matrix (i.e. the derivative of the model $\bnu$ with respect to the parameter vector $\bzeta$), $\lambda$ denotes the 
-damping factor, while $\bD$ is a diagonal matrix whose diagonal entries are equal to the diagonal entries of $\bJ^T\bJ$. In this paper, we denote 
+damping factor and $\bD$ is a diagonal matrix whose diagonal entries are equal to the diagonal entries of $\bJ^T\bJ$. In this paper, we denote 
 matrix transposition with $(*)^T$.
@@
 The gradient based minimization algorithms that are generally used to solve least squares problems (i.e. GN and LM) generally require the model $\bnu$ to be differentiable
-towards each model parameter. When the least squares problem is complex it becomes less straightforward to apply these gradient based minimization methods. The reason being,
-many complex functions are not analytic (no Taylor expansion around a fiducial point exists) and therefore not differentiable if the classic notion of differentiation is used. For instance if the 
+towards each model parameter. When the least squares problem is complex it becomes less straightforward to apply these gradient based minimization methods, as
+many complex functions are not analytic (no Taylor expansion around a fiducial point exists) and therefore not differentiable if the classic notion of differentiation is used. For instance, if the 
 classical definition of differentiation is used then $\frac{\partial \conj{z}}{\partial z}$, where $z \in \mathbb{C}$, does not exist.
 
-The only way to circumvent the differentiability conundrum, associated with complex least square problems, is to recast our complex optimization problem in such a way that $\bnu$ becomes analytic in its argument.
+The only way to circumvent the differentiability conundrum associated with complex least square problems is to recast our complex optimization problem in such a way that $\bnu$ becomes analytic in its argument.
 The standard way of achieving this is to divide the complex optimization problem into its real and imaginary parts and then to solve for the real and imaginary parts of the parameters separately.
@@
 \end{align}
-In this alternative approach we treat the complex parameters and their conjugates as separate parameters.
+In this alternative approach we treat the complex parameters and their conjugates as separate variables.
 
@@
 Eq.~\eqref{eq:least_squares_red}.
-We present the Signal-to-noise ratio (SNR) definition we used in this paper in Section~\ref{sec:snr}. The limitations of this paper are discussed in Section~\ref{sec:scope}.
+We present the signal-to-noise ratio (SNR) definition we used in this paper in Section~\ref{sec:snr}. The limitations of this paper are discussed in Section~\ref{sec:scope}.
 
@@
 
-The real and imaginary components of the thermal noise is normally distributed with a mean of zero and a standard deviation proportional to   
-\begin{equation}
-\sigma \propto \frac{T_{\textrm{sys}}}{\sqrt{\Delta \nu \tau}}, 
-\end{equation}
-where $T_{\textrm{sys}}$ is equal to the system temperature, $\Delta \nu$ is the observational bandwidth and $\tau$ is the integration time per visibility.
+The real and imaginary components of the thermal noise are normally distributed with a mean of zero and a standard deviation $\sigma$, where   
+\begin{equation}
+\sigma \propto \frac{T_{\textrm{sys}}}{\sqrt{\Delta \nu \tau}}. 
+\end{equation}
+Here $T_{\textrm{sys}}$ is equal to the system temperature, $\Delta \nu$ is the observational bandwidth and $\tau$ is the integration time per visibility.
 
@@
 We can use the composite index $pq$, which is the strategy that is followed in Eq.~\eqref{eq:vis_definition},
-or we can use a single unique baseline index which we assign in numerical order to each composite index; assuming the composite indexes are arranged in standard baseline order.
+or we can use a single unique baseline index which we assign in numerical order to each composite index, assuming the composite indexes are arranged in standard baseline order.
 The composite indexes $pq$ are in standard baseline order when they are arranged as follow: $12, 13, 1N, 23\cdots$. Let 
@@
 Let us now group all the baselines that measure the same $uv$-modes in a redundant array into the same redundant baseline group.
-Assigning the baselines of a redundant array to redundant baseline groups has the advantage that it reduces the number of single index values we require.
-We now only require single index values for each redundant baseline group, instead of a unique single index label for
+Assigning the baselines of a redundant array to different redundant baseline groups has the advantage that it reduces the number of single index values we require.
+We now only require single index values for each redundant baseline group instead of a unique single index label for
 each baseline. Let $\phi_{pq}$ be the function that maps composite antenna pairs to the appropriate single redundant baseline group index. Note that $\phi_{pq}$ is not symmetric as 
@@
 It is possible to construct a simple analytic expression for $\zeta_{pq}$ when our array is in an east-west regular configuration, namely $\zeta_{pq} = |q-p|$. 
-It becomes increasingly more difficult to construct analytic expressions of $\zeta_{pq}$ for other more complicated array layouts. 
+It becomes increasingly difficult to construct analytic expressions of $\zeta_{pq}$ for more complicated array layouts. 
 We will refer to $\zeta_{pq}$ as the symmetric geometric function. The empirically constructed symmetric geometry
 functions of three different redundant layouts are depicted in Fig.~\ref{fig:geometry_function}. We denote the range of $\zeta_{pq}$ with $\mathcal{R}(\zeta_{pq})$. The maximal element 
-that $\zeta_{pq}$ can ascertain is denoted by $L$ and can be construed as the maximal number of unique redundant baseline groups which can be formed for a given 
+that $\zeta_{pq}$ can ascertain is denoted by $L$ and can be interpreted as the maximal number of unique redundant baseline groups which can be formed for a given 
 array layout.
 Alternatively, we can interpret the symmetric geometry functions in Fig.~\ref{fig:geometry_function} as geometry matrices.
-In this alternative paradigm, $\zeta_{pq}$ denotes a matrix entry instead.
+In this alternative paradigm, $\zeta_{pq}$ denotes a matrix entry.
 The dimension of these geometry
@@
  \end{align}
-The number of model parameters that we need to solve for in Eq.~\eqref{eq:least_squares_red} is equal to $P = 2(N+L)$, since redundant calibration is a complex problem.
+The number of model parameters that we need to solve for in Eq.~\eqref{eq:least_squares_red} is $P = 2(N+L)$, since redundant calibration is a complex problem.
 We plot $L$ and $P$ for two different geometric layouts in Fig~\ref{fig:pl}.
-The analytic expressions for the red and the blue curve in Fig.~\ref{fig:ln} are respectively equal to
+The analytic expressions for the red and the blue curve in Fig.~\ref{fig:ln} are
 \begin{equation}
@@
 \begin{equation}
-L = N - 1. 
-\end{equation}
+L = N - 1, 
+\end{equation}
+respectively.
 
@@
 \end{equation}
-where $<*>_{\nu,t,pq}$ denotes averaging over frequency, time and baselines. The definition we use here is based on the SNR definitions used in \citet{Liu2010} and \citet{Marthi2014}.
+where $<*>_{\nu,t,pq}$ denotes averaging over frequency, time and baseline. The definition we use here is based on the SNR definitions used in \citet{Liu2010} and \citet{Marthi2014}.
 
@@
 As discussed in Section~\ref{sec:ls}, we can solve Eq.~\eqref{eq:least_squares_red} in one of two ways. We can split the problem into its real and
-imaginary parts and then solve for the real and imaginary parts of the parameters seperately or we can employ the notion of Wirtinger derivatives. The former 
-approach is discussed in Section~\ref{sec:ri}, while the latter approach is described in detail in Section~\ref{sec:w}.
-
-\subsection{Real-Imag Framework}
+imaginary parts and then solve for each part of the parameters seperately or we can employ the notion of Wirtinger derivatives. The former 
+approach is discussed in Section~\ref{sec:ri}, while the latter approach is described in Section~\ref{sec:w}.
+
+\subsection{Real-and-imaginary Framework}
 \label{sec:ri}
@@
 of Eq.~\eqref{eq:least_squares_red}, \citet{Liu2010} circumvents this difficulty by 
-splitting the problem into its phase and amplitude components.
-
-The aforementioned is achieved by first rewriting Eq.~\eqref{eq:least_squares_red} to
+splitting the problem into its phase-and-amplitude components.
+
+The aforementioned is achieved by first rewriting Eq.~\eqref{eq:least_squares_red} as
 \begin{equation}
@@
 The Jacobian associated with Eq.~\eqref{eq:least_squares_lincal} is equal to Eq.~\eqref{eq:Jac_lin}. 
-At the hand of Appendix~\ref{sec:lincal} it becomes evident that \textsc{lincal} is equivalent to a basic GN update step \citep{Kurien2016}. 
+Appendix~\ref{sec:lincal} makes it evident that \textsc{lincal} is equivalent to a basic GN update step \citep{Kurien2016}. 
 Therefore, a straightforward way of improving the convergence properties of stock standard \textsc{lincal} would be to simply add the LM damping factor to the basic \textsc{lincal} update step.
@@
 
-\citet{Smirnov2015} recently proposed to use the Wirtinger derivative instead of the classic real and imaginary approach to perform normal skymodel-based calibration. We will now show
+\citet{Smirnov2015} recently proposed to use the Wirtinger derivative instead of the classic real-and-imaginary approach to perform normal skymodel-based calibration. We will now show
 that the Wirtinger derivative can also be applied to redundant calibration.
@@
 %\end{equation}
-
-It is now trivial to compute the Hessian $\bH$ by using Eq.~\eqref{eq:Jacobian_red}. If we substitute Eq.~\eqref{eq:Jacobian_red} into $\bJ^H\bJ$
+We use $\bzero$ to denote an all zero matrix. It is now trivial to compute the Hessian $\bH$ by using Eq.~\eqref{eq:Jacobian_red}. If we substitute Eq.~\eqref{eq:Jacobian_red} into $\bJ^H\bJ$
 we obtain 
@@
 \end{equation}
-Moreover $\ba$ and $\bb$ are both column vectors. The length of $\ba$ and $\bb$ is $N$ and $L$, respectively.
+Additionally, $\ba$ and $\bb$ are both column vectors. The lengths of $\ba$ and $\bb$ are $N$ and $L$ respectively.
 The dimensions of the matrices we defined in this section are presented in Table~\ref{tab:matrix_dimensions}.
@@
 In Section~\ref{sec:main_sec} we discussed the complex GN and LM algorithms and how they relate to redundant calibration. If we inspect Eq.~\eqref{eq:GN_update} and 
-Eq.~\eqref{eq:LM_update} we see that the most expensive operation used by the GN and LM algorithms is the inversion of the Hessian $\bH$. The most logical 
+Eq.~\eqref{eq:LM_update} we see that the most expensive operation used by the GN and LM algorithms is the inversion of the Hessian, $\bH$. The most logical 
 way of speeding up the algorithm is to investigate whether we can do this inversion in some efficient way.
@@
 The basic difference between these two speedup algorithms can be understood by inspecting Fig.~\ref{fig:hessian}. The first observation we can make by inspecting 
-Fig.~\ref{fig:hessian_reg} and Fig.~\ref{fig:hessian_hex} is that the diagonal entries of the Hessians are more significant than the off-diagonal entries.
-It therefore makes sense to approximate the $\bH$ (or the modified Hessian $\bmH$) with its diagonal, which greatly simplifies the computational complexity of inverting 
+Fig.~\ref{fig:hessian_reg} and Fig.~\ref{fig:hessian_hex} is that the diagonal entries of $\bH$ are more significant than the off-diagonal entries in both cases.
+It therefore makes sense to approximate the $\bH$ with its diagonal, which greatly simplifies the computational complexity of inverting 
 $\bH$, i.e. it becomes $\mathbb{O}(P)$\footnote{The cost of constructing 
@@
 
-The second observation we can make from Fig.~\ref{fig:hessian_reg} and Fig.~\ref{fig:hessian_hex} is that the Hessians contain many zero entries, i.e.
+The second observation we can make from Fig.~\ref{fig:hessian_reg} and Fig.~\ref{fig:hessian_hex} is that both Hessians contain many zero entries, i.e.
 they are sparse matrices.
-This, in conjunction with the more significant diagonal entries observation, implies that the PCG method (presented in Section~\ref{sec:conj_grad})
+This, in conjunction with the more significant diagonal entries, implies that the PCG method (presented in Section~\ref{sec:conj_grad})
 can be applied to invert them in a computationally efficient manner.
@@
 \subfigure[Hexagonal layout]{\includegraphics[width=0.47\textwidth]{./hex_hessian.pdf}\label{fig:hessian_hex}}
-\caption{The number of analytic terms out of which the entries of the Hessian $\bH$ consists for two different geometric layouts, namely a regular east-west grid with $N = 5$ (left panel) and 
-a hexagonal grid with $N = 7$. The diagonal entries of these two Hessians are clearly more significant than their off-diagonal entries. Moreover, these two Hessians, also 
-contain many zero-entries. Note that the location of the zero-entries are dependent on the geometry of the array layout.
+\caption{The number of analytic terms out of which the entries of the Hessian $\bH$ consist for two different geometric layouts, namely a regular east-west grid with $N = 5$ (left panel) and 
+a hexagonal grid with $N = 7$. The diagonal entries of these two Hessians are clearly more significant than their off-diagonal entries. Moreover, these two Hessians also 
+contain many zero-entries. Note that the locations of the zero-entries are dependent on the geometry of the array layout.
 \label{fig:hessian}} 
@@
 
-Even though the PCG method is computationally more expensive than the ADI method we can not arrive at a final conclusion at this point. With the ADI method 
+Even though the PCG method is computationally more expensive than the ADI method we cannot arrive at a final conclusion at this point. With the ADI method 
 we only approximate the inverse of the Hessian, while in the case of the PCG method we compute the exact inverse. So the question arises whether the LM algorithm using the PCG method 
@@
 %simulations. 
-For the skymodel we generated 100 sources, each source had a random flux value and position.
+For the skymodel we generated 100 flat spectrum sources.
+Each source had a random flux value and position.
 The flux of the sources followed a Pareto distribution (with a shape parameter of 2), while their positions were uniformly
@@
 
-The main flow of a single simulation is as follow.
-We created corrupted visibilities, using the fixed simulation parameters presented in Table~\ref{tab:fixed_parm} and either one of the experimental setups summarized in Table~\ref{tab:ch_parm}.
+The main flow of a single simulation is as follows.
+We created corrupted visibilities using the fixed simulation parameters presented in Table~\ref{tab:fixed_parm} and either one of the experimental setups summarized in Table~\ref{tab:ch_parm}.
 We then used per time-slot and channel redundant calibration, i.e.
-we use the LM algorithm which in turn makes use of either the ADI or PCG method, to
+we use the LM algorithm, which in turn makes use of either the ADI or PCG method, to
 estimate the gains and the true observed visibilities. To obtain statistically significant results we performed, for both the ADI and PCG methods, multiple statistical independent simulations.
@@
 \label{sec:sbc}
-\textsc{StEFCal} is an ADI method, in which the measurement equation, Eq.~\eqref{eq:vis_definition}, is linearized by assuming that the gains are known, but
+\textsc{StEFCal} is an ADI method in which the measurement equation, Eq.~\eqref{eq:vis_definition}, is linearized by assuming that the gains are known, but
 that their conjugates are unknown \citep{Mitchell:MWA-cal,Salvini2014}.
 Under this assumption, the conjugates of the gains can be obtained trivially.
-Once the conjugates of the gains have been computed, the gains can be solved for.
+Once the conjugates of the gains have been computed, it is possible to solve for the gains.
 One can then alternate between computing the gains and their conjugates until convergence
 is reached.
-By correctly exploiting the symmetry inherent in the calibration problem this alternating approach can be replaced by a single iterative new parameter estimation step.
+By correctly exploiting the symmetry inherent in the calibration problem, this alternating approach can be replaced by a single iterative new parameter estimation step.
 \cite{Salvini2014} also show that
-the convergence of the algorithm can be improved if for every even iteration the basic new parameter estimation step is modified as follows: the current gain estimate is replaced by the average of the current gain estimate and the gain estimate of the previous odd iteration.
+the convergence of the algorithm can be improved, if for every even iteration, the basic new parameter estimation step is modified as follows: the current gain estimate is replaced by the average of the current gain estimate and the gain estimate of the previous odd iteration.
 \textsc{StEFCal} lowers the computational complexity of normal skymodel-based calibration from $\mathbb{O}(N^3)$ to $\mathbb{O}(N^2)$.
@@
 
-If we choose $\lambda$ to be zero, which corresponds to an $\alpha$ equal to one, we obtain the new parameter estimation step of \textsc{StEfCal}; used during odd iterations.
-On the other hand if we choose $\lambda$ equal to one, which corresponds
-to an $\alpha$ equal to a half, we obtain the new parameter estimation step of \textsc{StEfCal}; used during even iterations.
+If we choose $\lambda$ to be zero, which corresponds to an $\alpha$ equal to one, we obtain the new parameter estimation step of \textsc{StEfCal} used during odd iterations.
+On the other hand, if we choose $\lambda$ equal to one (which corresponds
+to $\alpha=\frac{1}{2}$) we obtain the new parameter estimation step of \textsc{StEfCal} used during even iterations.
 
@@
 \label{sec:red_c}
-In this section, we re-derive the fast redundant calibration algorithm presented in \citet{Marthi2014} using a different approach to the one presented in the paper itself.
+In this section, we re-derive the fast redundant calibration algorithm presented in \citet{Marthi2014} using two different approaches.
+Both these approaches differ from the one presented in \citet{Marthi2014}.
 In \citet{Marthi2014}, the algorithm is derived by taking the derivative of the
-objective function $\Lambda$ relative to $\bg$ and $\by$ and setting the result to naught.
+objective function $\Lambda$ relative to $\bg$ and $\by$, setting the intermediate result to naught and then solving for the unknown parameters.
 
@@
 we now alternate between solving for the antenna gains their conjugates and the true observed visibilities.
-If we do so we are able to re-derive the redundant calibration algorithm presented in \citet{Marthi2014}.
-This implies that the fast redundant calibration algorithm presented in \citet{Marthi2014} is effectively redundant \textsc{StEfCal} (i.e.
+See Appendix~\ref{sec:red_stef_ADI} for the details.
+The result of this extension is that we obtain the redundant calibration algorithm presented in \citet{Marthi2014}.
+This implies that the fast redundant calibration algorithm presented in \citet{Marthi2014} can for all intents and purposes be referred to as redundant \textsc{StEfCal} (i.e.
 some kind of ADI method).
@@
 some kind of ADI method) we will now use the diagonal approximation approach presented in \citet{Smirnov2015} to yet again re-derive the redundant calibration algorithm presented in \citet{Marthi2014}.
+The benefit of doing this second additional derivation is that it also enables us to show that the ADI method is in actual fact just a special case of the LM algorithm.
 
@@
 \end{equation}
-We obtain a similar result if we repeat the above procedure, only in this case we use Eq.~\eqref{eq:LM_update} instead,
-namely
+We obtain a similar result if we repeat the above procedure, only in this case we use Eq.~\eqref{eq:LM_update} instead. Thus,
 \begin{align}
@@
 expression of $\bJ^H\breve{\br}$, the only difference being that in Eq.~\eqref{eq:ab} the letter $r$ would be replaced by a $d$. If we substitute the analytic expression
-of $\bJ^H\breve{\bd}$ and $\widetilde{\bH}^{-1}$ (which is easily computed as the matrix is a diagonal matrix) into Eq.~\eqref{eq:alpha} we obtain the following two update rules:
+of $\bJ^H\breve{\bd}$ and $\widetilde{\bH}^{-1}$ (which is easily computed as the matrix is diagonal) into Eq.~\eqref{eq:alpha} we obtain the following two update rules:
 \begin{equation}
@@
 some form of ADI method, justifying the naming convention we have been using for it.
-Re-deriving the algorithm presented in \citet{Marthi2014} in this way serves as a further confirmation that it is indeed nothing more than redundant \textsc{StEfCal} (i.e. some kind of ADI method).
+%Re-deriving the algorithm presented in \citet{Marthi2014} in this way serves as a further confirmation that it is indeed nothing more than redundant \textsc{StEfCal} (i.e. some kind of ADI method).
 
@@
 If we wish to mimic this behaviour for the redundant calibration use case we need to scale $\breve{\bz}$ by two thirds in Eq.~\eqref{eq:alpha} (see Eq.~\eqref{eq:two_thrids}) which implies that we should choose $\alpha$ equal to a third, which corresponds to a $\lambda$ equal to two.
-We used $\lambda$ equal to two in this paper.
+We used this $\lambda$ value in this paper.
 
@@
 \label{sec:adi_results}
-Optimality results of ADI were presented in \citet{Marthi2014}. We do not repeat their analysis here. We however, do present our own ADI results, in Fig.~\ref{fig:prec_error}, that confirm that the
-ADI method can accurately estimate the antenna gains and the observed visibilities at different SNR values. To quantity the exact error in our estimates we used the percentage error $\beta$ defined 
+Optimality results of ADI were presented in \citet{Marthi2014}. We do not repeat their analysis here. We do, however, present our own ADI results, in Fig.~\ref{fig:prec_error}, that confirm that the
+ADI method can accurately estimate the antenna gains and the observed visibilities at different SNR values. To quantify the exact error in our estimates we used the percentage error $\beta$ defined 
 as 
@@
 We used a $\lambda$ value of two, implying we used Eq.~\eqref{eq:LM_update} instead of Eq.~\eqref{eq:GN_update} throughout this section.
-We respectively discuss the spectral condition number and the sparsity of the modified Hessian $\bmH$, which is a non-singular matrix, in Section~\ref{sec:scn} and Section~\ref{sec:sparsity}.
+We discuss the spectral condition number and the sparsity of the modified Hessian $\bmH$, which is a non-singular matrix, in Section~\ref{sec:scn} and Section~\ref{sec:sparsity} respectively.
 It is interesting to note, that the Hessian $\bH$ itself is singular which means that if we are to use Eq.~\eqref{eq:GN_update} instead we would need to be able to invert a singular matrix.
@@
 Fig.~\ref{fig:cg_itr} shows us that the number of major iterations needed by the preconditioned conjugate gradient method to invert $\bmH$ is independent of the number of antennas in the array.
-Moreover, the number of major iterations required are also much smaller than the dimension of $\bmH$.
+Moreover, the number of major iterations required is also much smaller than the dimension of $\bmH$.
 
@@
 
-The second factor that determines the computational complexity of the conjugate gradient method when used to perform redundant calibration is the amount of non-zero
+The second factor that determines the computational complexity of the conjugate gradient method when used to perform redundant calibration is the number of non-zero
 entries in the modified Hessian matrix $\bmH$. 
@@
 %to a constant by employing preconditioning and can therefore be completely eliminated from the computational complexity equation. 
-A useful matrix quantity which we will use in this section is the sparsity factor $\gamma$ of a matrix, which is the ratio between the amount of zero entries and the total amount of entries in a matrix.
+A useful matrix quantity which we will use in this section is the sparsity factor $\gamma$ of a matrix, which is the ratio between the number of zero entries and the total number of entries in a matrix.
 We
@@
 \end{align}
-However when our redundant array makes use of a more complicated geometric layout it becomes much more difficult to construct such expressions as $\bmH$ itself 
+However when our redundant array makes use of a more complicated geometric layout it becomes much more difficult to construct such expressions, as $\bmH$ itself 
 becomes much harder to express analytically.
 
-In Fig.~\ref{fig:gamma} we plot the empirically determined sparsity ratios for three different geometric layouts as a function of the number antennas in the array.
-Although, we can not explicitly calculate
+In Fig.~\ref{fig:gamma}, we plot the empirically determined sparsity ratios for three different geometric layouts as a function of the number antennas in the array.
+Although, we cannot explicitly calculate
 the limit of the hexagonal and square layout curves, Fig.~\ref{fig:gamma} seems to indicate that they do exist. Moreover, we also plot the left most equation
@@
 \end{align}
-Note, that we can only compute $c_{\infty}$ in the rightmost equation if $\lim_{N\rightarrow \infty} \gamma$ exists. 
-In Fig.~\ref{fig:c}, we plot the empiracally determined value of $c$ for three different geometric layouts as a function of the number antennas in the array.
-Furhtermore, in Fig.~\ref{fig:c}, the analytic value of $c$ for the regular east-west grid is plotted as a dashed red curve, while the limit of this analytic curve is plotted with a dotted black curve.
+
+The evaluation of the limit in the rightmost term of Eq.~\eqref{eq:c} is evaluated as follows: 
+Note that,
+\begin{equation}
+\label{eq:int_limit}
+\lim_{N\rightarrow\infty} P^{\log_{P}(1-\gamma)} = 1-\gamma_{\infty},
+\end{equation}
+which is only possible if 
+\begin{equation}
+\lim_{N\rightarrow\infty} \log_{P}(1-\gamma) = 0,
+\end{equation}
+implying that $c_{\infty}$ is equal to two.
+
+Eq.~\eqref{eq:int_limit}, implies that we can only compute $c_{\infty}$ in the rightmost equation if $\lim_{N\rightarrow \infty} \gamma$ exists. 
+In Fig.~\ref{fig:c}, we plot the empirically determined value of $c$ for three different geometric layouts as a function of the number antennas in the array.
+Furthermore, in Fig.~\ref{fig:c}, the analytic value of $c$ for the regular east-west grid is plotted as a dashed red curve, while the limit of this analytic curve is plotted with a dotted black curve.
 
@@
 Moreover, will the number
-of LM iterations PCG require be so few that it manages to beat the ADI method after all.
+of LM iterations PCG requires be so few that it manages to beat the ADI method after all.
 To answer this we first need to compute the number of iterations, $k$, the LM implementation associated with the ADI method can make in the
@@
 \end{align}
-Next we need to measure the amount of LM iterations we save by using the PCG method and compare it to Eq.\eqref{eq:k}. We make this comparison 
+Next we need to measure the number of LM iterations we save by using the PCG method and compare it to Eq.\eqref{eq:k}. We make this comparison 
 in Fig.~\ref{fig:out_diff}. Fig.~\ref{fig:out_diff} tells us that the ADI method will outperform the PCG method as the number of LM iterations we save by taking the 
@@
 In Section~\ref{sec:main_sec} we showed that instead of using the traditional phase-and-amplitude approach to perform redundant calibration we can actually use a purely complex 
-framework. In this framework the unknown variables and their conjugates are treated as independent variables that have to be estimated. 
-We then derived the GN and LM update steps associated with this complex framework (see Section~\ref{sec:complex_GN_LM}).
+framework. In this framework, the unknown variables and their conjugates are treated as independent variables that have to be estimated. 
+We then derived the GN and LM update steps using this complex framework (see Section~\ref{sec:complex_GN_LM}).
 In this paper we focused more on the LM algorithm than on the GN algorithm.
@@
 by $\mathbb{O}(P^2)$. Both of these algorithms perform much better than stock standard inversion, which has a computational complexity of $\mathbb{O}(P^3)$.
-Although the ADI method does execute faster than the PCG method it is far less robust as it would be less straightforward to apply to other calibration uses cases. The reason being,
+Although the ADI method does execute faster than the PCG method it is far less robust as it would be less straightforward to apply to other calibration uses cases. This is due to the fact that,
 the diagonal approximation we make with ADI might not be accurate enough in general.
@@
 
+\section{Redundant \textsc{StEfCal} and the ADI method}
+\label{sec:red_stef_ADI}
+The basic skymodel-based \textsc{StEfCal} update step is equal to  the leftmost term in Eq.~\eqref{eq:g_update} (barring $\alpha$) \citep{Salvini2014}.
+Assume without any loss of generality that the array is in an east-west regular grid. Furthermore, assume that $\boldsymbol{d}$ has been re-ordered to
+\begin{equation}
+\widetilde{\boldsymbol{d}} = \left[d{12},\cdots,d_{N-1,N},d_{13},\cdots,d_{N-2,N},\cdots,d_{1N}\right]^T .
+\end{equation}
+
+Eq.~\eqref{eq:vis_red_vec} (after dropping the noise term) can now be rewritten as 
+\begin{equation}
+\label{eq:linear_system}
+\widetilde{\boldsymbol{d}} = \boldsymbol{R}\boldsymbol{y}, 
+\end{equation}
+if we assume that $\boldsymbol{g}$ and its conjugate are known vectors. In Eq.~\eqref{eq:linear_system},
+\begin{equation}
+\boldsymbol{R} = 
+\begin{bmatrix}
+g_1\conj{g}_2 & 0 & \cdots & 0\\
+g_2\conj{g}_3 & 0 & \cdots & 0\\
+\vdots & 0 & \cdots & 0\\
+g_{N-1}\conj{g}_N & 0 & \cdots & 0\\
+0 & g_1\conj{g}_3 & \cdots & 0\\
+0 & \vdots & \cdots & 0\\
+0 & g_{N-2}\conj{g}_N & \cdots & 0\\
+0 & 0 & \cdots & 0\\
+  & \vdots & \\
+0 & 0 & \cdots & g_1\conj{g}_N\\  
+\end{bmatrix}.
+\end{equation}
+We can now determine $\boldsymbol{y}$ with
+\begin{equation}
+\label{eq:y_final}
+\boldsymbol{y} = (\boldsymbol{R}^H\boldsymbol{R})^{-1}\boldsymbol{R}^H\widetilde{\boldsymbol{d}}, 
+\end{equation}
+where 
+\begin{equation}
+\label{eq:RHR}
+[\boldsymbol{R}^H\boldsymbol{R}]_{ij} = 
+\begin{cases}
+\sum_{rs\in\mathcal{RS}_i} |g_r|^2|g_s|^2 &\textrm{if}~i=j\\
+0&\textrm{otherwise}
+\end{cases},
+\end{equation}
+and
+\begin{equation}
+\label{eq:RHd}
+[\boldsymbol{R}^H\widetilde{\boldsymbol{d}}]_i = \sum_{rs\in\mathcal{RS}_i} \conj{g}_r g_s d_{rs}. 
+\end{equation}
+Substituting Eq.~\eqref{eq:RHR} and Eq.~\eqref{eq:RHd} into Eq.~\eqref{eq:y_final}, followed by writing out the analytic expression of $y_i$, we obtain
+the leftmost term in Eq.~\eqref{eq:y_update} (barring $\alpha$).
 \label{lastpage}

### 2017-01-24 software/simulator.py
sim changes

@@ -1107,7 +1107,8 @@ def plot_paper_layouts():
     print "s.L = ",s.L
     print "s.N = ",s.N
 
-    s = sim(layout="REG",order=100) #INSTANTIATE OBJECT
+    # Make a 100 for phi paper plot...
+    s = sim(layout="REG",order=10) #INSTANTIATE OBJECT
     s.generate_antenna_layout()
     s.plot_ant(title="REG")
     phi,zeta = s.calculate_phi(s.ant[:,0],s.ant[:,1])
@@ -1130,7 +1131,7 @@ def Gianni_freq_exp():
 
 if __name__ == "__main__":
    #example_usage()
-   #plot_paper_layouts()
-   Gianni_freq_exp()
+   plot_paper_layouts()
+   #Gianni_freq_exp()
    
 

### 2017-01-24 paper.tex
added new layout figures

@@
 We focus more on the Levenberg-Maquardt algorithm in this paper as it exhibits better convergence properties when compared to Gauss-Newton.
-Traditionally redundant calibration is implemented via a hierarchical algorithmic strategy.
+The Gauss-Newton algorithm is currently the de facto standard that is being used to perform redundant calibration.
+It is common practice to implement redundant calibration via a hierarchical algorithmic strategy.
 Using the Levenberg-Maquardt algorithm, therefore, reduces the number of algorithmic layers that are needed to perform redundant calibration.
-The Gauss-Newton algorithm is, however, still the de facto standard that is currently being used to perform redundant calibration.
 We also compare two previously proposed methods to accelerate redundant calibration, namely the alternating direction implicit method and the preconditioned conjugate gradient method.
@@
 
+It is also interesting to mention here that a completely different calibration strategy can also be used to calibrate redundant arrays, \textsc{CorrCal} \citep{Sievers2017}\footnote{https://github.com/sievers/corrcal2}.
+In \textsc{CorrCal} the unknown
+visibilities are no longer solved for. The sky is actually modelled, via two covariance matrices. 
+The first covariance matrix is associated with known point sources. The second covariance matrix models the remaining sky in a statistical sense. The unknown gains, however, are still 
+estimated. We do not elaborate any further on this approach in this paper.
+
 New instruments like LOFAR (Low Frequency Array) \citep{Noorishad2012}, PAPER (Precision Array for Probing the Epoch of Reionization) \citep{Ali2015} and HERA (Hydrogen Epoch of Reionization Array) \citep{deboer2015} have also sparked a renewed interest in redundant calibration.
@@
 \textbf{[GIANNI TO REWRITE AND EXTEND EOR SECTION]}
+
+In this paper we propose to use the complex framework introduced by \citet{Smirnov2015} instead of the traditional amplitude-and-phase approach used by \textsc{lincal}.
+We also propose to use the Levenberg-Marquardt (LM) gradient-based algorithm instead of the Gauss-Newton (GN) algorithm as it exhibits better convergence properties.
+Using the LM algorithm reduces the number of algorithmic layers that are needed to perform redundant calibration.
 
@@
 to speed it up \citep{Liu2010}. A proper comparison between these two approaches has not been done before.
-
-It is also interesting to mention here that a completely different calibration strategy can also be used to calibrate redundant arrays, \textsc{CorrCal} \citep{Sievers2017}\footnote{https://github.com/sievers/corrcal2}.
-In \textsc{CorrCal} the unknown
-visibilities are no longer solved for. The sky is actually modelled, via two covariance matrices. 
-The first covariance matrix is associated with known point sources. The second covariance matrix models the remaining sky in a statistical sense. The unknown gains, however, are still 
-estimated. We do not elaborate any further on this approach in this paper.
-
-In this paper we propose to use the complex framework introduced by \citet{Smirnov2015} instead of the traditional amplitude-and-phase approach used by \textsc{lincal}.
-We also propose to use the Levenberg-Marquardt (LM) gradient-based algorithm instead of the Gauss-Newton (GN) algorithm as it exhibits better convergence properties.
-Using the LM algorithm reduces the number of algorithmic layers that are needed to perform redundant calibration. 
-We also compare two recent redundant calibration speedup algorithms with each other, namely the ADI \citep{Marthi2014} and PCG \citep{Liu2010} methods.
-In this paper, we decided to focus on the complex framework proposed by \citet{Smirnov2015} rather than the traditional phase-and-amplitude approach as it
+Therefore, we also compare the ADI \citep{Marthi2014} and PCG \citep{Liu2010} methods in this paper.
+We decided to focus on the complex framework proposed by \citet{Smirnov2015} rather than the traditional phase-and-amplitude approach as it
 allows us to trivially prove that the ADI method proposed by \citep{Marthi2014} is merely a special case of the more general LM algorithm.
@@
 We start the paper by presenting some useful preliminaries in Section~\ref{sec:preliminaries}. In Section~\ref{sec:p_statement}, we formulate redundant calibration 
-as a least squares problem, i.e. Eq~\eqref{eq:least_squares_red}. In Section~\ref{sec:main_sec} we use the complex framework proposed by \citet{Smirnov2015} to derive a numerical
-algorithm (see Eq.~\eqref{eq:GN_update} to Eq.~\eqref{eq:update}), which we can use to solve Eq~\eqref{eq:least_squares_red}.
+as a least squares problem, i.e. Eq~\eqref{eq:least_squares_red}. In Section~\ref{sec:main_sec} we use the complex framework proposed by \citet{Smirnov2015} to derive the GN and LM
+update rules (see Eq.~\eqref{eq:GN_update} to Eq.~\eqref{eq:update}), which we can use to solve Eq~\eqref{eq:least_squares_red}.
 In Section~\ref{sec:results}, we compare the computational complexity of two recently proposed redundant calibration speedup algorithms with one another.
@@
 In this paper, we denote matrix inversion with $(*)^{-1}$.
-The best possible choice of $\bM$ would therefore be $\bA$, which would lead to $\kappa$ becoming unity (as the largest and smallest eigenvalue of the Identity matrix are equal to one).
+The best possible choice of $\bM$ would therefore be $\bA$, which would lead to $\kappa$ becoming unity (as the largest and smallest eigenvalue of the identity matrix are equal to one).
 If $\kappa$ is one then Eq.~\eqref{eq:cg_bound} simplifies and becomes $\mathbb{O}(m)$.
@@
 \label{sec:ls}
-The method of nonlinear least squares is tanatamount to solving the following optimization problem:  
+The method of nonlinear least squares is tantamount to solving the following optimization problem:  
 \begin{equation}
@@
 \end{equation}
-Here $T_{\textrm{sys}}$ is equal to the system temperature, $\Delta \nu$ is the observational bandwidth and $\tau$ is the integration time per visibility.
+Here, $T_{\textrm{sys}}$ is equal to the system temperature, $\Delta \nu$ is the observational bandwidth and $\tau$ is the integration time per visibility.
 
@@
 each baseline. Let $\phi_{pq}$ be the function that maps composite antenna pairs to the appropriate single redundant baseline group index. Note that $\phi_{pq}$ is not symmetric as 
-$\phi_{pq} = 0$ if $p>q$. We can also define the following symmetric variant of $\phi_{pq}$:
+$\phi_{pq} = 0$ if $p\geq q$. We can also define the following symmetric variant of $\phi_{pq}$:
 \begin{equation}
@@
 In the bottom panel we have the respective symmetric redundancy geometry functions for the layouts on the top panel. We used 91 antennas
-to construct the the hexagonal layout, while a 100 antennas where used in the square and east-west layouts. The maximal amount of redundant baselines that can be formed for 
-the hexagonal, square and east-west layouts in the top panel are 165, 180 and 99 baselines respectively.\label{fig:geometry_function}}
+to construct the hexagonal layout, while 100 antennas where used in the square and east-west layouts.
+In the case of the east-west layout we only plot the positions of the first ten antennas.
+The maximal amount of redundant baseline groups that can be formed for
+the hexagonal, square and east-west layouts are 165, 180 and 99 respectively.\label{fig:geometry_function}}
 \end{figure*}
@@
 We plot $L$ and $P$ for two different geometric layouts in Fig~\ref{fig:pl}.
-The analytic expressions for the red and the blue curve in Fig.~\ref{fig:ln} are
-\begin{equation}
-L = 2N-\frac{1}{2}\sqrt{12N-3}-\frac{1}{2}, 
+The analytic expressions for the red and the blue curve in Fig.~\ref{fig:ln} are \citep{Camps2003}
+\begin{equation}
+L = 2N-\frac{1}{2}\sqrt{12N-3}-\frac{1}{2} 
 \end{equation}
@@
 \begin{equation}
-L = N - 1, 
+L = N - 1 
 \end{equation}
@@
 
-\subsection{Real-and-imaginary Framework}
+\subsection{Real and imaginary Framework}
 \label{sec:ri}
@@
 
-\citet{Smirnov2015} recently proposed to use the Wirtinger derivative instead of the classic real-and-imaginary approach to perform normal skymodel-based calibration. We will now show
+\citet{Smirnov2015} recently proposed to use the Wirtinger derivative instead of the classic real and imaginary approach to perform normal skymodel-based calibration. We will now show
 that the Wirtinger derivative can also be applied to redundant calibration.
@@
 we use the LM algorithm, which in turn makes use of either the ADI or PCG method, to
-estimate the gains and the true observed visibilities. To obtain statistically significant results we performed, for both the ADI and PCG methods, multiple statistical independent simulations.
+estimate the gains and the true observed visibilities.
+Our initial parameter guess $\bz_0$ was chosen to be an all one vector.
+To obtain statistically significant results we performed, for both the ADI and PCG methods, multiple statistical independent simulations.
 
@@
 \centering
-\caption{The following fixed simulation parameters were used to generate the results of this paper.}
+\caption{The following fixed simulation parameters were used to generate the results of this paper. We use $U$ to represent 
+a uniform distribution.}
 \begin{tabular}{|c c c c c c c|} 
@@
 \caption{The gain error models used in this paper. We have used the symbol $x$ here as a proxy as it can either refer to time-slots or channels. We either
-performed our simulations over multiple time-slots and one frequency channel or one timeslot and multiple frequencey channels (see Table~\ref{tab:ch_parm}). We use $U$ to represent 
-a uniform distribution. Moreover, $c$ in the left most column denotes the speed of light.}
+performed our simulations over multiple time-slots and one frequency channel or one timeslot and multiple frequencey channels (see Table~\ref{tab:ch_parm}).
+Moreover, $c$ in the left most column denotes the speed of light.}
 \begin{tabular}{|c c c c|} 
@@
 Gain error model (Table~\ref{tab:gain_parm}) & 2 & 1\\
-Initial guess: $\bz_0$ & $\bone$ & $\bone$\\
+%Initial guess: $\bz_0$ & $\bone$ & $\bone$\\
 \hline
@@
 is in fact nothing more than an extension of the skymodel-based ADI method, presented in \citet{Salvini2014}, to the redundant calibration use case (i.e. redundant \textsc{StEfCal}).
-In doing so we will justify the naming convention we have been using up until now.
+In doing so we will justify the naming convention we have been using for it.
 
@@
 \label{sec:sbc}
-\textsc{StEFCal} is an ADI method in which the measurement equation, Eq.~\eqref{eq:vis_definition}, is linearized by assuming that the gains are known, but
+\textsc{StEfCal} is an ADI method in which the measurement equation, Eq.~\eqref{eq:vis_definition}, is linearized by assuming that the gains are known, but
 that their conjugates are unknown \citep{Mitchell:MWA-cal,Salvini2014}.
@@
 In \citet{Marthi2014}, the algorithm is derived by taking the derivative of the
-objective function $\Lambda$ relative to $\bg$ and $\by$, setting the intermediate result to naught and then solving for the unknown parameters.
+objective function $\Lambda$ relative to $\bg$ and $\by$, setting the intermediate results to naught and then solving for the unknown parameters.
 
@@
 In this section we investigate the computational complexity of the PCG method when applied to the redundant calibration problem. 
-As we mentioned in Section~\ref{sec:pcg} there are two factors which influence the computational complexity of the PCG method, namely the spectral condition number and the sparsity of the matrix we wish to invert (see Eq.~\eqref{eq:cg_bound}).
+As we mentioned in Section~\ref{sec:conj_grad} there are two factors which influence the computational complexity of the PCG method, namely the spectral condition number and the sparsity of the matrix we wish to invert (see Eq.~\eqref{eq:cg_bound}).
 We used a $\lambda$ value of two, implying we used Eq.~\eqref{eq:LM_update} instead of Eq.~\eqref{eq:GN_update} throughout this section.
@@
 
-The evaluation of the limit in the rightmost term of Eq.~\eqref{eq:c} is evaluated as follows: 
-Note that,
+The evaluation of the limit in the rightmost equation of Eq.~\eqref{eq:c} is evaluated as follows: 
+note that
 \begin{equation}
@@
 in Fig.~\ref{fig:out_diff}. Fig.~\ref{fig:out_diff} tells us that the ADI method will outperform the PCG method as the number of LM iterations we save by taking the 
-full inverse is not enough to exceed $k$.
+full inverse is not enough to exceed $k$. In this comparison we have only taken into account the cost of inverting $\bH$. The PCG method might, however, still 
+turn out to be quite useful in practice as it would be more straightforward to apply it to other calibration use cases.
 
@@
 \label{sec:lincal}
-In this section we derive \textsc{lincal} \citep{Liu2010} from first principles.
 %Assume that $\eta_p,\eta_q,\widetilde{\eta}_{\phi_{pq}},\varphi_p,\varphi_q$ and $\widetilde{\varphi}_{\phi_{pq}}$ are real valued variables.
-We can rewrite Eq.~\eqref{eq:vis_red_vec} to
+We can rewrite Eq.~\eqref{eq:vis_red_vec} to \citep{Liu2010}
 \begin{equation}
@@
 The basic skymodel-based \textsc{StEfCal} update step is equal to  the leftmost term in Eq.~\eqref{eq:g_update} (barring $\alpha$) \citep{Salvini2014}.
-Assume without any loss of generality that the array is in an east-west regular grid. Furthermore, assume that $\boldsymbol{d}$ has been re-ordered to
+Assume without any loss of generality that the array is in an east-west regular grid. Furthermore, assume that $\boldsymbol{d}$ (see Eq.~\eqref{eq:vis_red_vec}) has been re-ordered to
 \begin{equation}
@@
 \end{equation}
-
-Eq.~\eqref{eq:vis_red_vec} (after dropping the noise term) can now be rewritten as 
+The vector $\widetilde{\bn}$ should be interpreted in a similar manner.
+
+Eq.~\eqref{eq:vis_red_vec} can now be rewritten as 
 \begin{equation}
 \label{eq:linear_system}
-\widetilde{\boldsymbol{d}} = \boldsymbol{R}\boldsymbol{y}, 
+\widetilde{\boldsymbol{d}} = \boldsymbol{J}\boldsymbol{y} + \widetilde{\bn}, 
 \end{equation}
@@
 \begin{equation}
-\boldsymbol{R} = 
+\boldsymbol{J} = 
 \begin{bmatrix}
@@
 \end{equation}
-We can now determine $\boldsymbol{y}$ with
+We can now estimate $\boldsymbol{y}$ with
 \begin{equation}
 \label{eq:y_final}
-\boldsymbol{y} = (\boldsymbol{R}^H\boldsymbol{R})^{-1}\boldsymbol{R}^H\widetilde{\boldsymbol{d}}, 
+\boldsymbol{y} = (\boldsymbol{J}^H\boldsymbol{J})^{-1}\boldsymbol{J}^H\widetilde{\boldsymbol{d}}, 
 \end{equation}
@@
 \label{eq:RHR}
-[\boldsymbol{R}^H\boldsymbol{R}]_{ij} = 
+[\boldsymbol{J}^H\boldsymbol{J}]_{ij} = 
 \begin{cases}
@@
 \label{eq:RHd}
-[\boldsymbol{R}^H\widetilde{\boldsymbol{d}}]_i = \sum_{rs\in\mathcal{RS}_i} \conj{g}_r g_s d_{rs}. 
-\end{equation}
-Substituting Eq.~\eqref{eq:RHR} and Eq.~\eqref{eq:RHd} into Eq.~\eqref{eq:y_final}, followed by writing out the analytic expression of $y_i$, we obtain
+[\boldsymbol{J}^H\widetilde{\boldsymbol{d}}]_i = \sum_{rs\in\mathcal{RS}_i} \conj{g}_r g_s d_{rs}. 
+\end{equation}
+Substituting Eq.~\eqref{eq:RHR} and Eq.~\eqref{eq:RHd} into Eq.~\eqref{eq:y_final} and then write out the analytic expression of $y_i$ we obtain
 the leftmost term in Eq.~\eqref{eq:y_update} (barring $\alpha$).

### 2017-01-24 software/simulator.py

@@ -305,7 +305,8 @@ class sim():
               #plt.annotate(label, xy = (x, y), xytext = (-15, 15), textcoords = 'offset points', fontsize = 12, ha = 'right', va = 'bottom', bbox = dict(boxstyle = 'round,pad=0.5', fc = 'white', alpha = 0.2), arrowprops = dict(arrowstyle = '->', connectionstyle = 'arc3,rad=0')) 
           plt.xlabel("W-E [m]",fontsize=label_size)
           plt.ylabel("S-N [m]",fontsize=label_size)
-          plt.title(title+"-"+str(self.N),fontsize=label_size)
+          #plt.title(title+"-"+str(self.N),fontsize=label_size)
+          plt.title(title+"-100",fontsize=label_size)
           plt.axis("equal")
           plt.axis([-1.1*m, 1.1*m,-1.1*m, 1.1*m])
           plt.show() 

### 2017-01-24 paper.tex
improved figures

@@
 \subfigure[Hexagonal layout]
-{\includegraphics[width=0.33\textwidth]{./HEX_lay.pdf}\label{fig:HEX_lay}}
+{\includegraphics[width=0.32\textwidth]{./HEX_lay.pdf}\label{fig:HEX_lay}}
 \subfigure[Square layout]
-{\includegraphics[width=0.33\textwidth]{./SQR_lay.pdf}\label{fig:SQR_lay}}
+{\includegraphics[width=0.32\textwidth]{./SQR_lay.pdf}\label{fig:SQR_lay}}
 \subfigure[Regular east-west layout]
-{\includegraphics[width=0.33\textwidth]{./REG_lay.pdf}\label{fig:REG_lay}}
+{\includegraphics[width=0.32\textwidth]{./REG_lay.pdf}\label{fig:REG_lay}}
 

### 2017-01-24 software/plot_func.py

@@ -796,7 +796,7 @@ def plot_P():
     plt.xlabel(r'$N$ [antennas]')
     plt.ylabel(r'$P$')
     plt.legend(loc=5,prop={'size': 18})
-    plt.xlim([37,217])
+    plt.xlim([7,217])
     
     plt.grid('on')
 

### 2017-02-24 paper.tex
started with Gianni's edits

@@
 \usepackage{graphicx}
-\usepackage{algorithm}
-\usepackage[noend]{algpseudocode}
+%\usepackage{algorithm}
+%\usepackage[noend]{algpseudocode}
 \usepackage{mathrsfs}
@@
 \newcommand{\bv}{\bmath{v}}
+\newcommand{\bu}{\bmath{u}}
 \newcommand{\bn}{\bmath{n}}
@@
 \title[Unknown]{Redundant calibration as a complex optimization problem}
-\author[A. V. Raveendran and A. N. Other]{A. V. Raveendran$^{1}$\thanks{E-mail:
-email@address (AVR); otheremail@otheraddress (ANO)} and A. N.
-Other$^{2}$\footnotemark[1]\thanks{This file has been amended to
-highlight the proper use of \LaTeXe\ code with the class file.
-These changes are for illustrative purposes and do not reflect the
-original paper by A. V. Raveendran.}\\
-$^{1}$Indian Institute of Astrophysics, Bangalore 560034, India\\
-$^{2}$Building, Institute, Street Address, City, Code, Country}
+\author[T.L.~Grobler et al.]{T.L.~Grobler$^{1}$\thanks{E-mail: email@address (AVR)}, G.~Bernardi$^{1,2}$, Z.S.~Ali$^{3}$?, C.L.~Carilli$^{4,5}$?, J.S.~Dillon$^{3}$?, A.~Liu$^{3}$?, \newauthor
+A.R.~Parsons$^{3,6}$? and O.M.~Smirnov$^{1,2}$?\\
+$^{1}$Department of Physics and Electronics, Rhodes University, PO Box 94, Grahamstown, 6140, South Africa\\
+$^{2}$SKA SA, 3rd Floor, The Park, Park Road, Pinelands, 7405, South Africa\\
+$^{3}$Dept. of Astronomy, University of California, Berkeley, CA 94720, USA\\
+$^{4}$National Radio Astronomy Obs., Socorro NM\\
+$^{5}$Cavendish Lab., Cambridge UK\\
+$^{6}$Radio Astronomy Lab., U. California, Berkeley CA 94720, USA}
 \begin{document}
 
-\date{Accepted 1988 December 15. Received 1988 December 14; in original form 1988 October 11}
+%\date{Accepted 1988 December 15. Received 1988 December 14; in original form 1988 October 11}
 
@@
 \begin{abstract}
+%In this paper we show that instead of splitting our redundant calibration problem into its phase-and-amplitude components (the more traditional approach) we can treat the complex unknown model parameters and their complex conjugates as separate unknowns.
+We also derive the Levenberg-Maquardt and Gauss-Newton update steps associated with this alternative parameter-and-conjugate framework.
+We focus more on the Levenberg-Maquardt algorithm in this paper as it exhibits better convergence properties when compared to Gauss-Newton.
+The Gauss-Newton algorithm is currently the de facto standard that is being used to perform redundant calibration.
+It is common practice to implement redundant calibration via a hierarchical algorithmic strategy.
+Using the Levenberg-Maquardt algorithm, therefore, reduces the number of algorithmic layers that are needed to perform redundant calibration.
+We also compare two previously proposed methods to accelerate redundant calibration, namely the alternating direction implicit method and the preconditioned conjugate gradient method.
+We found that the alternating direction implicit method outperforms the preconditioned conjugate gradient method.
+However, the preconditioned conjugate gradient method is more robust as it is more straightforward to adapt to other calibration use cases.
+We make use of the parameter-and-conjugate framework in this paper, rather than the more traditional amplitude-and-phase approach as it makes it trivial to relate the alternating direction implicit method and the Levenberg-Marquardt algorithm with one another.
+Observations of the redshifted 21-cm line from the epoch of reionization have recently motivated the construction of low frequency radio arrays with highly redundant configurations that provide an alternative calibration strategy - ''redundant calibration" - and boosts sensitivity on specific scales.
+We show that redundant calibration can be formulated in the domain of complex functions of a complex variable.
+This framework reduces the
+
+has driven advancements in the calibration of radio interferometers and 
+Calibration of the new generation of low frequency radio interferometers faces new challenges due to their very wide field of views 
+The construction of new radio interferometers, particularly operating at low frequencies and 
 In this paper we show that instead of splitting our redundant calibration problem into its phase-and-amplitude components (the more traditional approach) we can treat the complex unknown model parameters and their complex conjugates as separate unknowns.
@@
 We also compare two previously proposed methods to accelerate redundant calibration, namely the alternating direction implicit method and the preconditioned conjugate gradient method.
-We found that the alternating
-direction implicit method outperforms the preconditioned conjugate gradient method.
+We found that the alternating direction implicit method outperforms the preconditioned conjugate gradient method.
 However, the preconditioned conjugate gradient method is more robust as it is more straightforward to adapt to other calibration use cases.
-We make use of the parameter-and-conjugate framework in this paper, rather than the more traditional amplitude-and-phase approach as it makes it trivial to relate the alternating direction implicit method
-and the Levenberg-Marquardt algorithm with one another.
+We make use of the parameter-and-conjugate framework in this paper, rather than the more traditional amplitude-and-phase approach as it makes it trivial to relate the alternating direction implicit method and the Levenberg-Marquardt algorithm with one another.
 \end{abstract}
@@
 
-\section{Preliminaries}
-\label{sec:preliminaries}
-There are two major pre-existing algorithms we use throughout this paper, namely the Conjugate Gradient (CG) method and the least-squares solver.
-We respectively discuss the CG method and least-squares in Section~\ref{sec:conj_grad} and Section~\ref{sec:ls}.
-
-\subsection{Conjugate Gradient Method}
-\label{sec:conj_grad}
-The CG method was independently discovered by \citet{Hestenes1973} and \citet{Stiefel1952}. They later published a joint paper, which is now considered as the seminal
-reference on CG \citep{Hestenes1952}. The iterative CG method is used to solve a particular class of linear system. The CG method is generally used to solve
-\begin{equation}
-\label{eq:linear_system}
-\bb = \bA\bx,
-\end{equation}
-where $\bb\in\mathbb{C}^P$ is a known complex column vector of size $P$, $\bx\in\mathbb{C}^P$ is an unknown complex column vector with the same length as $\bb$, and $\bA\in\mathbb{C}^{P\times P}$ is a square positive-definite Hermitian matrix with the appropriate dimensions.
-Moreover, using the CG method to solve Eq.~\eqref{eq:linear_system} will only be computationally advantageous if $\bA$ is also sparse (contain many zero entries).
-Using the iterative CG method to solve large sparse linear sytems was popularized by \citet{Reid1971}. A good tutorial on the CG method can be found in \citep{Shewchuk1994}.
-The iterative CG method is presented in Algorithm \ref{algo:CG}.
-
-%If $\bA$ is a square positive semi-definite Hermitian matrix, then the conjugate gradient method only converges if $\bb$ is in the column range of $\bA$ (cite Lu2016).
-
-\begin{algorithm}
-\caption{Conjugate Gradient Method.
-Inputs: $\bA$, $\bb$, a starting value for $\bx$, maximum number of iterations $k_{\textrm{max}}$ and an error tolerance $\epsilon<1$ Output: $\bx$ the solution to Eq.~\eqref{eq:linear_system}.
-\citep{Shewchuk1994}.}\label{algo:CG}
-\begin{algorithmic}[1]
-\State $k \gets 0$
-\State $\br \gets \bb - \bA\bx$
-\State $\bp \gets \br$
-\State $\delta_{\textrm{new}} \gets \br^H\br$
-\State $\delta_0 \gets \delta_{\textrm{new}}$
-\While{$k\leq k_{\textrm{max}}$ and $\delta_{\textrm{new}} > \epsilon^2\delta_0$}
-\State $\alpha \gets \frac{\delta_{\textrm{new}}}{\bp^H\bA\bp}$
-\State $\bx \gets \bx + \alpha\bp$
-\State $\br \gets \br - \alpha\bA\bp$
-\State $\delta_{\textrm{old}} \gets \delta_{\textrm{new}}$ 
-\State $\delta_{\textrm{new}} \gets \br^H\br$
-\State $\beta \gets \frac{\delta_{\textrm{new}}}{\delta_{\textrm{old}}}$
-\State $\bp \gets \br + \beta\bp$
-\State $k \gets k + 1$
-\EndWhile
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{Computational Complexity}
-\citet{Kaniel1966} was able to determine convergence bounds for CG using Chebyshev polynomials. A much more in depth analysis of CG convergence can be found in
-\citep{Sluis1986}. As can be seen from Algorithm \ref{algo:CG}, the CG method makes use of a major outer loop (line 6 of Algorithm \ref{algo:CG}). It turns out that at worst 
-the number of major iterations that CG requires to converge\footnote{If the $\bA$-norm is used to calculate the error.} is proportional to $\sqrt{\kappa}$, where $\kappa$ is the spectral condition number of the matrix $\bA$.
-The mathematical definition of $\kappa(\bA)$ is
-\begin{equation}
-\label{eq:kappa}
-\kappa(\bA) = \frac{\lambda_{\textrm{max}}}{\lambda_{\textrm{min}}}, 
-\end{equation}
-where $\lambda_{\textrm{max}}$ and $\lambda_{\textrm{min}}$ respectively denote the largest and the smallest eigenvalue of $\bA$.
-Moreover, the most expensive operation that is conducted within each major loop of Algorithm~\ref{algo:CG} is the vector-matrix product $\bA\bp$, which has a complexity of $\mathbb{O}(m)$, where $m$ is the number of non-zero entries in $\bA$.
-The computational complexity of CG is therefore 
-\begin{equation}
-\label{eq:cg_bound}
-\mathbb{O}(\sqrt{\kappa}m). 
-\end{equation}
-It should now also be clear to the reader why the CG method is especially applicable to sparse linear systems. In sparse linear systems $m < P^2$.
-
-\subsubsection{Preconditioning}
-\label{sec:precon}
-Preconditioning is a technique which allows us to improve the condition number of a matrix. Let $\bM$ be a positive-definite Hermitian matrix.
-We can now indirectly solve for $\bx$ by using 
-\begin{equation}
-\label{eq:prec}
-\bM^{-1}\bA\bx = \bM^{-1}\bb.
-\end{equation}
-If $\bM$ is a good preconditioner then $\kappa(\bM^{-1}\bA) \ll \kappa(\bA)$, i.e. a good preconditioner lowers the condition number of a matrix which in turn increases the convergence speed 
-of CG (see Eq.~\eqref{eq:cg_bound}).
-In this paper, we denote matrix inversion with $(*)^{-1}$.
-The best possible choice of $\bM$ would therefore be $\bA$, which would lead to $\kappa$ becoming unity (as the largest and smallest eigenvalue of the identity matrix are equal to one).
-If $\kappa$ is one then Eq.~\eqref{eq:cg_bound} simplifies and becomes $\mathbb{O}(m)$.
-However, if we knew the inverse of $\bA$ then obtaining $\bx$ would be trivial and there would be no need for the CG method in the first place.
-In short, a good preconditioner $\bM$ must approximate $\bA$ fairly accurately. However, if it is not much less expensive to 
-compute the inverse of $\bM$ than that of $\bA$ we actually gain nothing (see Eq.~\eqref{eq:prec}).
-
-If the diagonal entries of $\bA$ are much more significant than its off-diagonal entries then the 
-Jacobian preconditioner is a good choice of $\bM$. The Jacobian preconditioner is computed as follows:
-\begin{equation}
-\bM = \bA\odot\bI, 
-\end{equation}
-where $\bI$ is the identity matrix and $\odot$ denotes the Hadamard product. Since the Jacobian preconditioner is a diagonal matrix its inverse can be obtained trivially.
-
-\subsubsection{Normal Equation}
-\label{sec:normal}
-Lets assume that instead of Eq.~\eqref{eq:linear_system} we have a more general linear system, i.e.
-\begin{equation}
- \bb = \bB\bx,
-\end{equation}
-where $\bb\in\mathbb{C}^K$, $\bx\in\mathbb{C}^P$  and $\bB\in\mathbb{C}^{K \times P}$, with $K > P$. We can use the normal equation 
-\begin{equation}
-\label{eq:normal_equation}
-\bB^H\bB\bx = \bB^H\bb, 
-\end{equation}
-to estimate the $\bx$ that minimizes
-\begin{equation}
-\label{eq:norm_ls}
-\|\bb-\bB\bx\|_F^2. 
-\end{equation}
-In this paper, we denote the Hermitian transpose with $(*)^H$ and the Frobenius norm with $\|*\|_F$.
-
-There are a few important observations that we can make while inspecting Eq.~\eqref{eq:normal_equation}:
-\begin{enumerate}
-\item $\bB^H\bB$ is a square positive semi-definite Hermitian matrix by construction.
-\item $\bB^H\bb$ is in the column range of $\bB^H\bB$; this follows trivially from 
-\begin{equation}
-\bB^H\bb \in \mathcal{R}(\bB^H) \implies \bB^H\bb \in \mathcal{R}(\bB^H\bB).   
-\end{equation}
-\end{enumerate}
-In this paper, we denote the column range of a matrix $\bA$ with $\mathcal{R}(\bA)$.
-
-Recall that we mentioned that in general we can only solve Eq.~\eqref{eq:linear_system} with CG if $\bA$ is a square positive-definite Hermitian matrix.
-The CG method actually has an even broader applicability: it can be used to solve Eq.~\eqref{eq:linear_system} even if $\bA$ is a Hermitian positive semi-definite
-square matrix as long as $\bb$ is in the column range of $\bA$ \citep{Lu2015}.
-In light of \citet{Lu2015}, the two observations in the numbered list we presented in the previous paragraph imply that Eq.~\eqref{eq:normal_equation}
-can be solved by employing the CG method. The solution returned by CG is, however, only unique if $\bB^H\bB$ is positive-definite (invertible). The solution
-obtained by CG, whether it is unique or not, does minimize Eq.~\eqref{eq:norm_ls} in a least squares sense.
-
-Note that if $\bA$ is positive semi-definite, its smallest eigenvalue is zero. This implies that Eq.~\eqref{eq:kappa} is not defined. \citet{Lu2015} show that when $\bA$
-is a square positive semi-definite Hermitian matrix we do not use the spectral condition number to bound the convergence complexity of CG, but the general spectral condition
-number instead. The general spectral condition number is very similar to the spectral condition number; the only difference being that in Eq.~\eqref{eq:kappa} we replace  
-$\lambda_{\textrm{min}}$ with $\lambda_{\textrm{min}}^+$, which denotes the smallest positive eigenvalue.
-
-\subsection{Least Squares}
-\label{sec:ls}
-The method of nonlinear least squares is tantamount to solving the following optimization problem:  
-\begin{equation}
-\label{eq:least_squares}
-\min_{\bzeta} \Lambda(\bzeta) = \min_{\bzeta} \|\brho\|_F^2 = \min_{\bzeta} \|\bdelta - \bnu(\bzeta)\|_F^2, 
-\end{equation}
-where $\Lambda$ is known as the objective function, $\brho$ is the residual vector, $\bdelta$ is the observed data vector, $\bnu$ is the model-predicted data vector and $\bzeta$ is the model-parameter vector that minimizes Eq.~\eqref{eq:least_squares}.
-In Eq.~\eqref{eq:least_squares}, we assume that the optimization problem is real valued.
-The main objective of least squares is to find the model-parameters which minimize the squared difference between the observed data and the model-predicted data.
-
-The most standard way of solving Eq.~\eqref{eq:least_squares} is to use a gradient-type minimization algorithm.
+{\bf GB: ok here are my comments/questions.
+\begin{itemize}
+\item I think we wanna start with Section 2 being a short review of standard calibration but in the complex optimization framework.
+Therefore I would begin with section 3.1, following by section 2.2.
+In particular I think that the discussion of least squares should be connected to (current) eq.
+15;
+\item we need more context for section 3.2.
+Why do we need to introduce a different indexing?
+This should be motivated.
+As it stands, I think it should remain where it is if we can make a connection to standard calibration, otherwise it should go in section 3.3;
+\item section 3.3 should merge with 4 where redundant calibration is introduced;
+\item section 3.5 goes together with the description of the simulations;
+\item section 4.1: I do not understand the definition in eq.
+32.
+How is the ratio of two matrices defined?
+I also suggest to use the symbol $\equiv$ rather than $=$ for definitions throughout the paper;
+\item can section 4.2.2 be moved to the appendix? Or how much of it?
+\item section 5 should be called something like "Implementation"...
+In Oleg's paper he writes "Inverting $J^H J$ and separability"...
+I find it a little too direct, but we need to find something else as these are not yet "Results".
+I also suggest to revert the order in the way this is presented: we first show that the Hessian is almost diagonal (is this only true in the complex represntation?), then we derive ADI and show that is the same as Marthi \& Chengalur;
+\item I have not seen the simulations where you inserted a phase wrap and you calibrate it. Is there a reason why it is not included?
+\end{itemize}
+}
+
+% \subsection{Least Squares}
+% \label{sec:ls}
+% The method of nonlinear least squares is tantamount to solving the following optimization problem:  
+% \begin{equation}
+% \label{eq:least_squares}
+% \min_{\bzeta} \Lambda(\bzeta) = \min_{\bzeta} \|\brho\|_F^2 = \min_{\bzeta} \|\bdelta - \bnu(\bzeta)\|_F^2, 
+% \end{equation}
+% where $\Lambda$ is known as the objective function, $\brho$ is the residual vector, $\bdelta$ is the observed data vector, $\bnu$ is the model-predicted data vector and $\bzeta$ is the model-parameter vector that minimizes Eq.~\eqref{eq:least_squares}.
+% In Eq.~\eqref{eq:least_squares}, we assume that the optimization problem is real valued.
+% The main objective of least squares is to find the model-parameters which minimize the squared difference between the observed data and the model-predicted data.
+% 
+% The most standard way of solving Eq.~\eqref{eq:least_squares} is to use a gradient-type minimization algorithm.
 In this paper we will be focusing on the following gradient minimization algorithms: GN and LM \citep{Levenberg1944,Marquardt1963}.
-The GN update step is defined as
-\begin{equation}
- \Delta \bzeta = (\bJ^T\bJ)^{-1} \bJ^T\brho,
-\end{equation}
-while the LM update takes the following form
-\begin{equation}
- \Delta \bzeta = (\bJ^T\bJ+\lambda\bD)^{-1} \bJ^T\brho,
-\end{equation}
-where $\bJ$ is the Jacobian matrix (i.e. the derivative of the model $\bnu$ with respect to the parameter vector $\bzeta$), $\lambda$ denotes the 
-damping factor and $\bD$ is a diagonal matrix whose diagonal entries are equal to the diagonal entries of $\bJ^T\bJ$. In this paper, we denote 
-matrix transposition with $(*)^T$.
-These update steps can be used in an iterative manner to determine new estimates of $\bzeta$.
-
-The gradient based minimization algorithms that are generally used to solve least squares problems (i.e. GN and LM) generally require the model $\bnu$ to be differentiable
-towards each model parameter. When the least squares problem is complex it becomes less straightforward to apply these gradient based minimization methods, as
-many complex functions are not analytic (no Taylor expansion around a fiducial point exists) and therefore not differentiable if the classic notion of differentiation is used. For instance, if the 
-classical definition of differentiation is used then $\frac{\partial \conj{z}}{\partial z}$, where $z \in \mathbb{C}$, does not exist.
-
-The only way to circumvent the differentiability conundrum associated with complex least square problems is to recast our complex optimization problem in such a way that $\bnu$ becomes analytic in its argument.
-The standard way of achieving this is to divide the complex optimization problem into its real and imaginary parts and then to solve for the real and imaginary parts of the parameters separately.
-
-Recently, \citet{Sorber2012} proposed an alternative strategy, which involves the use of Wirtinger calculus \citep{Wirtinger1927}. The Wirtinger derivatives 
-are defined to be 
-\begin{align}
-\label{eq:wir}
-\frac{\partial}{\partial z} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} -  i \frac{\partial}{\partial y} \right ),&\frac{\partial}{\partial \conj{z}} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} +  i \frac{\partial}{\partial y} \right ).
-\end{align}
-Using the above definitions we can now easily compute the following:
-\begin{align}
-\frac{\partial z}{\partial z} & = 1, & \frac{\partial \conj{z}}{\partial z}&=0, & \frac{\partial z}{\partial \conj{z}} & = 0, & \frac{\partial \conj{z}}{\partial \conj{z}}&=1.
-\end{align}
-In this alternative approach we treat the complex parameters and their conjugates as separate variables.
-
-\section{Problem Statement}
+% The GN update step is defined as
+% \begin{equation}
+%  \Delta \bzeta = (\bJ^T\bJ)^{-1} \bJ^T\brho,
+% \end{equation}
+% while the LM update takes the following form
+% \begin{equation}
+%  \Delta \bzeta = (\bJ^T\bJ+\lambda\bD)^{-1} \bJ^T\brho,
+% \end{equation}
+% where $\bJ$ is the Jacobian matrix (i.e. the derivative of the model $\bnu$ with respect to the parameter vector $\bzeta$), $\lambda$ denotes the 
+% damping factor and $\bD$ is a diagonal matrix whose diagonal entries are equal to the diagonal entries of $\bJ^T\bJ$. In this paper, we denote 
+% matrix transposition with $(*)^T$.
+% These update steps can be used in an iterative manner to determine new estimates of $\bzeta$.
+% 
+% The gradient based minimization algorithms that are generally used to solve least squares problems (i.e. GN and LM) generally require the model $\bnu$ to be differentiable
+% towards each model parameter. When the least squares problem is complex it becomes less straightforward to apply these gradient based minimization methods, as
+% many complex functions are not analytic (no Taylor expansion around a fiducial point exists) and therefore not differentiable if the classic notion of differentiation is used. For instance, if the 
+% classical definition of differentiation is used then $\frac{\partial \conj{z}}{\partial z}$, where $z \in \mathbb{C}$, does not exist.
+% 
+% The only way to circumvent the differentiability conundrum associated with complex least square problems is to recast our complex optimization problem in such a way that $\bnu$ becomes analytic in its argument.
+% The standard way of achieving this is to divide the complex optimization problem into its real and imaginary parts and then to solve for the real and imaginary parts of the parameters separately.
+% 
+% Recently, \citet{Sorber2012} proposed an alternative strategy, which involves the use of Wirtinger calculus \citep{Wirtinger1927}. The Wirtinger derivatives 
+% are defined to be 
+% \begin{align}
+% \label{eq:wir}
+% \frac{\partial}{\partial z} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} -  i \frac{\partial}{\partial y} \right ),&\frac{\partial}{\partial \conj{z}} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} +  i \frac{\partial}{\partial y} \right ).
+% \end{align}
+% Using the above definitions we can now easily compute the following:
+% \begin{align}
+% \frac{\partial z}{\partial z} & = 1, & \frac{\partial \conj{z}}{\partial z}&=0, & \frac{\partial z}{\partial \conj{z}} & = 0, & \frac{\partial \conj{z}}{\partial \conj{z}}&=1.
+% \end{align}
+% In this alternative approach we treat the complex parameters and their conjugates as separate variables.
+
+\section{Wirtinger Calibration}
 \label{sec:p_statement}
-In Section~\ref{sec:s_mes}, we present the scalar measurement equation. Next, we define some indexing functions in Section~\ref{sec:indexing}. The measurement 
-equation and the indexing functions from Section~\ref{sec:indexing} are then used in Section~\ref{sec:r_mes} to phrase redundant calibration as a least-squares problem, i.e.
+%In Section~\ref{sec:s_mes}, we present the scalar measurement equation. Next, we define some indexing functions in Section~\ref{sec:indexing}. The measurement 
+%equation and the indexing functions from Section~\ref{sec:indexing} are then used in Section~\ref{sec:r_mes} to phrase redundant calibration as a least-squares problem, i.e.
 Eq.~\eqref{eq:least_squares_red}.
-We present the signal-to-noise ratio (SNR) definition we used in this paper in Section~\ref{sec:snr}. The limitations of this paper are discussed in Section~\ref{sec:scope}.
+%We present the signal-to-noise ratio (SNR) definition we used in this paper in Section~\ref{sec:snr}. The limitations of this paper are discussed in Section~\ref{sec:scope}.
 
@@
 
-\subsection{Indexing Functions}
-\label{sec:indexing}
-In general there are two ways of indexing the true visibilities associated with different baselines.
-We can use the composite index $pq$, which is the strategy that is followed in Eq.~\eqref{eq:vis_definition},
-or we can use a single unique baseline index which we assign in numerical order to each composite index, assuming the composite indexes are arranged in standard baseline order.
-The composite indexes $pq$ are in standard baseline order when they are arranged as follow: $12, 13, 1N, 23\cdots$. Let 
+Eq.~\eqref{eq:vis_definition} can also be expressed in the following vector form 
+\begin{equation}
+\label{eq:vis_definition}
+\bd = \bv + \bn, 
+\end{equation}
+where 
+\begin{align}
+ \left [ \bd \right]_{\alpha_{pq}} &= d_{pq}, & \left [ \bv \right ]_{\alpha_{pq}} &= v_{pq}=g_p y_{pq} \conj{g_q},\nonumber\\
+ \left [ \bn \right ]_{\alpha_{pq}} &= n_{pq}. &  &\label{eq:vec_definitions}
+\end{align}
+The function $\alpha_{pq}$ maps composite antenna indexes to unique single indexes and is defined as
 \begin{equation}
@@
 \end{equation}
-We can use $\alpha_{pq}$ to map composite indexes to unique baseline indexes, i.e. 
-\begin{equation}
-12,13,\cdots,1N,23,\cdots \xrightarrow{\alpha_{pq}} 1,2,3,\cdots,B.
-\end{equation}
-We denote the number of baselines in the array with $B$, i.e. $B = \frac{N^2-N}{2}$.
-
-Recall that redundant baselines sample the exact same visibilities in the $uv$-plane, i.e. if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$. 
-Let us now group all the baselines that measure the same $uv$-modes in a redundant array into the same redundant baseline group.
+For example,
+\begin{equation}
+\{\alpha_{12},\alpha_{13},\alpha_{1N},\alpha_{23},\cdots,\alpha_{N-1N}\} = \{1,2,N-1,N,\cdots,B\},
+\end{equation}
+where $B$ denotes the number of baselines in the array, i.e. $B = \frac{N^2-N}{2}$. The vectors in Eq.~\eqref{eq:vec_definitions} are column vectors of size $B$ (i.e. $p<q$).
+
+We can now formulate calibration as a least-squares problem:
+\begin{equation}
+\label{eq:least_squares}
+\min_{\bg} \Lambda(\bg) = \min_{\bg} \|\br\|_F^2 = \min_{\bg} \|\bd - \bv(\bg)\|_F^2, 
+\end{equation}
+where $\bg =[g_1,\cdots,g_N]^T$.
+
+least squares problems are normally solved by using gradient based methods, like Gauss-Newton and Levenberg-Marquardt .
+
+\begin{align}
+ \bg &=[g_1,\cdots,g_N]^T, & \by &= [y_1,\cdots,y_L]^T,\nonumber\\
+ \bz &= [\bg^T,\by^T]^T. &  &\label{eq:parm_definitions}
+ \end{align}
+
+% \subsection{Indexing Functions}
+% \label{sec:indexing}
+% In general there are two ways of indexing the true visibilities associated with different baselines.
+We can use the composite index $pq$, which is the strategy that is followed in Eq.~\eqref{eq:vis_definition},
+% or we can use a single unique baseline index which we assign in numerical order to each composite index, assuming the composite indexes are arranged in standard baseline order.
+% The composite indexes $pq$ are in standard baseline order when they are arranged as follow: $12, 13, 1N, 23\cdots$. Let 
+% \begin{equation}
+% \alpha_{pq} =
+% \begin{cases}
+% (q-p) + (p-1)\left (N-\frac{1}{2}p \right ) & \textrm{if}~p<q\\
+% %B + (p-q) + (q-1)\left (N-\frac{1}{2}q \right )) & \textrm{if}~p>q\\
+% %B+q + \frac{1}{2}(p-1)(p-2) & \textrm{if}~p>q\\
+% 0 & \textrm{otherwise}
+% \end{cases}.
+% \end{equation}
+% We can use $\alpha_{pq}$ to map composite indexes to unique baseline indexes, i.e. 
+% \begin{equation}
+% 12,13,\cdots,1N,23,\cdots \xrightarrow{\alpha_{pq}} 1,2,3,\cdots,B.
+% \end{equation}
+% We denote the number of baselines in the array with $B$, i.e. $B = \frac{N^2-N}{2}$.
+% 
+% Recall that redundant baselines sample the exact same visibilities in the $uv$-plane, i.e. if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$. 
+% Let us now group all the baselines that measure the same $uv$-modes in a redundant array into the same redundant baseline group.
 Assigning the baselines of a redundant array to different redundant baseline groups has the advantage that it reduces the number of single index values we require.
 We now only require single index values for each redundant baseline group instead of a unique single index label for
-each baseline. Let $\phi_{pq}$ be the function that maps composite antenna pairs to the appropriate single redundant baseline group index. Note that $\phi_{pq}$ is not symmetric as 
-$\phi_{pq} = 0$ if $p\geq q$. We can also define the following symmetric variant of $\phi_{pq}$:
-\begin{equation}
-\zeta_{pq} = 
-\begin{cases}
-\phi_{pq}~\textrm{if}~p \leq q\\
-\phi_{qp}~\textrm{if}~p>q
-\end{cases}.
-\end{equation} 
-It is possible to construct a simple analytic expression for $\zeta_{pq}$ when our array is in an east-west regular configuration, namely $\zeta_{pq} = |q-p|$. 
-It becomes increasingly difficult to construct analytic expressions of $\zeta_{pq}$ for more complicated array layouts. 
-We will refer to $\zeta_{pq}$ as the symmetric geometric function. The empirically constructed symmetric geometry
-functions of three different redundant layouts are depicted in Fig.~\ref{fig:geometry_function}. We denote the range of $\zeta_{pq}$ with $\mathcal{R}(\zeta_{pq})$. The maximal element 
-that $\zeta_{pq}$ can ascertain is denoted by $L$ and can be interpreted as the maximal number of unique redundant baseline groups which can be formed for a given 
-array layout.
+% each baseline. Let $\phi_{pq}$ be the function that maps composite antenna pairs to the appropriate single redundant baseline group index. Note that $\phi_{pq}$ is not symmetric as 
+% $\phi_{pq} = 0$ if $p\geq q$. We can also define the following symmetric variant of $\phi_{pq}$:
+% \begin{equation}
+% \zeta_{pq} = 
+% \begin{cases}
+% \phi_{pq}~\textrm{if}~p \leq q\\
+% \phi_{qp}~\textrm{if}~p>q
+% \end{cases}.
+% \end{equation} 
+% It is possible to construct a simple analytic expression for $\zeta_{pq}$ when our array is in an east-west regular configuration, namely $\zeta_{pq} = |q-p|$. 
+% It becomes increasingly difficult to construct analytic expressions of $\zeta_{pq}$ for more complicated array layouts. 
+% We will refer to $\zeta_{pq}$ as the symmetric geometric function. The empirically constructed symmetric geometry
+% functions of three different redundant layouts are depicted in Fig.~\ref{fig:geometry_function}. We denote the range of $\zeta_{pq}$ with $\mathcal{R}(\zeta_{pq})$. The maximal element 
+% that $\zeta_{pq}$ can ascertain is denoted by $L$ and can be interpreted as the maximal number of unique redundant baseline groups which can be formed for a given 
+% array layout.
 Alternatively, we can interpret the symmetric geometry functions in Fig.~\ref{fig:geometry_function} as geometry matrices.
@@
 The dimension of these geometry
-matrices are $N\times N$.
-
-\begin{figure*}
-\centering
-\subfigure[Hexagonal layout]
-{\includegraphics[width=0.32\textwidth]{./HEX_lay.pdf}\label{fig:HEX_lay}}
-\subfigure[Square layout]
-{\includegraphics[width=0.32\textwidth]{./SQR_lay.pdf}\label{fig:SQR_lay}}
-\subfigure[Regular east-west layout]
-{\includegraphics[width=0.32\textwidth]{./REG_lay.pdf}\label{fig:REG_lay}}
-
-\subfigure[Hexagonal: $\zeta_{pq}$]
-{\includegraphics[width=0.33\textwidth]{./HEX_phi.pdf}\label{fig:HEX_phi}}
-\subfigure[Square: $\zeta_{pq}$]
-{\includegraphics[width=0.33\textwidth]{./SQR_phi.pdf}\label{fig:SQR_phi}}
-\subfigure[Regular east-west: $\zeta_{pq}$]
-{\includegraphics[width=0.33\textwidth]{./REG_phi.pdf}\label{fig:REG_phi}}
-\caption{In the top panel we have three different redundant antenna layouts, namely a hexagonal (left), square (middle) and a regular east-west (right) layout. 
-In the bottom panel we have the respective symmetric redundancy geometry functions for the layouts on the top panel. We used 91 antennas
-to construct the hexagonal layout, while 100 antennas where used in the square and east-west layouts.
+% matrices are $N\times N$. 
+% 
+% \begin{figure*}
+% \centering
+% \subfigure[Hexagonal layout]
+% {\includegraphics[width=0.32\textwidth]{./HEX_lay.pdf}\label{fig:HEX_lay}}
+% \subfigure[Square layout]
+% {\includegraphics[width=0.32\textwidth]{./SQR_lay.pdf}\label{fig:SQR_lay}}
+% \subfigure[Regular east-west layout]
+% {\includegraphics[width=0.32\textwidth]{./REG_lay.pdf}\label{fig:REG_lay}}
+% 
+% \subfigure[Hexagonal: $\zeta_{pq}$]
+% {\includegraphics[width=0.33\textwidth]{./HEX_phi.pdf}\label{fig:HEX_phi}}
+% \subfigure[Square: $\zeta_{pq}$]
+% {\includegraphics[width=0.33\textwidth]{./SQR_phi.pdf}\label{fig:SQR_phi}}
+% \subfigure[Regular east-west: $\zeta_{pq}$]
+% {\includegraphics[width=0.33\textwidth]{./REG_phi.pdf}\label{fig:REG_phi}}
+% \caption{In the top panel we have three different redundant antenna layouts, namely a hexagonal (left), square (middle) and a regular east-west (right) layout. 
+% In the bottom panel we have the respective symmetric redundancy geometry functions for the layouts on the top panel. We used 91 antennas
+% to construct the hexagonal layout, while 100 antennas where used in the square and east-west layouts.
 In the case of the east-west layout we only plot the positions of the first ten antennas.
 The maximal amount of redundant baseline groups that can be formed for
-the hexagonal, square and east-west layouts are 165, 180 and 99 respectively.\label{fig:geometry_function}}
-\end{figure*}
-
-If the function $\phi_{pq}$ is known and one is given one of the two antenna indexes that together form a redundant baseline as well as the redundant baseline group index itself then it is possible 
-to calculate the unknown antenna index. This can be computed via the following two expressions:
-\begin{equation}
-\xi_{ij} = 
-\begin{cases}
-p~\textrm{if}~\exists! ~ p \in \mathbb{N} ~ s.t. ~(\phi_{pi} = j)\\
-0~\textrm{otherwise}
-\end{cases},
-\end{equation}
-and
-\begin{equation}
-\psi_{ij} = 
-\begin{cases}
-q~\textrm{if}~\exists! ~ q \in \mathbb{N} ~ s.t. ~(\phi_{iq} = j)\\
-0~\textrm{otherwise}
-\end{cases}.
-\end{equation}
-We use $\xi_{ij}$ to retrieve the first antenna index of the composite antenna pair associated with a particular baseline, if the index of the first antenna in the composite antenna pair and the unique redundant baseline group index are known, while we use $\psi_{ij}$ to obtain
-a similar result: the second antenna index of the composite antenna pair. %\textbf{NB::} I still need to plot $\zeta_{ij}$ for different antenna layouts.
+% the hexagonal, square and east-west layouts are 165, 180 and 99 respectively.\label{fig:geometry_function}}
+% \end{figure*}
+% 
+% If the function $\phi_{pq}$ is known and one is given one of the two antenna indexes that together form a redundant baseline as well as the redundant baseline group index itself then it is possible 
+% to calculate the unknown antenna index. This can be computed via the following two expressions:
+% \begin{equation}
+% \xi_{ij} = 
+% \begin{cases}
+% p~\textrm{if}~\exists! ~ p \in \mathbb{N} ~ s.t. ~(\phi_{pi} = j)\\
+% 0~\textrm{otherwise}
+% \end{cases},
+% \end{equation}
+% and
+% \begin{equation}
+% \psi_{ij} = 
+% \begin{cases}
+% q~\textrm{if}~\exists! ~ q \in \mathbb{N} ~ s.t. ~(\phi_{iq} = j)\\
+% 0~\textrm{otherwise}
+% \end{cases}.
+% \end{equation}
+% We use $\xi_{ij}$ to retrieve the first antenna index of the composite antenna pair associated with a particular baseline, if the index of the first antenna in the composite antenna pair and the unique redundant baseline group index are known, while we use $\psi_{ij}$ to obtain
+% a similar result: the second antenna index of the composite antenna pair. %\textbf{NB::} I still need to plot $\zeta_{ij}$ for different antenna layouts.
+
+\section{Redundant Wirtinger Calibration}
 
@@
 
+\subsection{Conjugate Gradient Method}
+\label{sec:conj_grad}
+The CG method was independently discovered by \citet{Hestenes1973} and \citet{Stiefel1952}. They later published a joint paper, which is now considered as the seminal
+reference on CG \citep{Hestenes1952}. The iterative CG method is used to solve a particular class of linear system. The CG method is generally used to solve
+\begin{equation}
+\label{eq:linear_system}
+\bb = \bA\bx,
+\end{equation}
+where $\bb\in\mathbb{C}^P$ is a known complex column vector of size $P$, $\bx\in\mathbb{C}^P$ is an unknown complex column vector with the same length as $\bb$, and $\bA\in\mathbb{C}^{P\times P}$ is a square positive-definite Hermitian matrix with the appropriate dimensions.
+Moreover, using the CG method to solve Eq.~\eqref{eq:linear_system} will only be computationally advantageous if $\bA$ is also sparse (contain many zero entries).
+Using the iterative CG method to solve large sparse linear sytems was popularized by \citet{Reid1971}. A good tutorial on the CG method can be found in \citep{Shewchuk1994}.
+The iterative CG method is presented in Algorithm \ref{algo:CG}.
+
+%If $\bA$ is a square positive semi-definite Hermitian matrix, then the conjugate gradient method only converges if $\bb$ is in the column range of $\bA$ (cite Lu2016).
+
+%\begin{algorithm}
+%\caption{Conjugate Gradient Method.
+Inputs: $\bA$, $\bb$, a starting value for $\bx$, maximum number of iterations $k_{\textrm{max}}$ and an error tolerance $\epsilon<1$ Output: $\bx$ the solution to Eq.~\eqref{eq:linear_system}.
+\citep{Shewchuk1994}.}\label{algo:CG}
+%\begin{algorithmic}[1]
+%\State $k \gets 0$
+%\State $\br \gets \bb - \bA\bx$
+%\State $\bp \gets \br$
+%\State $\delta_{\textrm{new}} \gets \br^H\br$
+%\State $\delta_0 \gets \delta_{\textrm{new}}$
+%\While{$k\leq k_{\textrm{max}}$ and $\delta_{\textrm{new}} > \epsilon^2\delta_0$}
+%\State $\alpha \gets \frac{\delta_{\textrm{new}}}{\bp^H\bA\bp}$
+%\State $\bx \gets \bx + \alpha\bp$
+%\State $\br \gets \br - \alpha\bA\bp$
+%\State $\delta_{\textrm{old}} \gets \delta_{\textrm{new}}$ 
+%\State $\delta_{\textrm{new}} \gets \br^H\br$
+%\State $\beta \gets \frac{\delta_{\textrm{new}}}{\delta_{\textrm{old}}}$
+%\State $\bp \gets \br + \beta\bp$
+%\State $k \gets k + 1$
+%\EndWhile
+%\end{algorithmic}
+%\end{algorithm}
+
+\subsubsection{Computational Complexity}
+\citet{Kaniel1966} was able to determine convergence bounds for CG using Chebyshev polynomials. A much more in depth analysis of CG convergence can be found in
+\citep{Sluis1986}. As can be seen from Algorithm \ref{algo:CG}, the CG method makes use of a major outer loop (line 6 of Algorithm \ref{algo:CG}). It turns out that at worst 
+the number of major iterations that CG requires to converge\footnote{If the $\bA$-norm is used to calculate the error.} is proportional to $\sqrt{\kappa}$, where $\kappa$ is the spectral condition number of the matrix $\bA$.
+The mathematical definition of $\kappa(\bA)$ is
+\begin{equation}
+\label{eq:kappa}
+\kappa(\bA) = \frac{\lambda_{\textrm{max}}}{\lambda_{\textrm{min}}}, 
+\end{equation}
+where $\lambda_{\textrm{max}}$ and $\lambda_{\textrm{min}}$ respectively denote the largest and the smallest eigenvalue of $\bA$.
+Moreover, the most expensive operation that is conducted within each major loop of Algorithm~\ref{algo:CG} is the vector-matrix product $\bA\bp$, which has a complexity of $\mathbb{O}(m)$, where $m$ is the number of non-zero entries in $\bA$.
+The computational complexity of CG is therefore 
+\begin{equation}
+\label{eq:cg_bound}
+\mathbb{O}(\sqrt{\kappa}m). 
+\end{equation}
+It should now also be clear to the reader why the CG method is especially applicable to sparse linear systems. In sparse linear systems $m < P^2$.
+
+\subsubsection{Preconditioning}
+\label{sec:precon}
+Preconditioning is a technique which allows us to improve the condition number of a matrix. Let $\bM$ be a positive-definite Hermitian matrix.
+We can now indirectly solve for $\bx$ by using 
+\begin{equation}
+\label{eq:prec}
+\bM^{-1}\bA\bx = \bM^{-1}\bb.
+\end{equation}
+If $\bM$ is a good preconditioner then $\kappa(\bM^{-1}\bA) \ll \kappa(\bA)$, i.e. a good preconditioner lowers the condition number of a matrix which in turn increases the convergence speed 
+of CG (see Eq.~\eqref{eq:cg_bound}).
+In this paper, we denote matrix inversion with $(*)^{-1}$.
+The best possible choice of $\bM$ would therefore be $\bA$, which would lead to $\kappa$ becoming unity (as the largest and smallest eigenvalue of the identity matrix are equal to one).
+If $\kappa$ is one then Eq.~\eqref{eq:cg_bound} simplifies and becomes $\mathbb{O}(m)$.
+However, if we knew the inverse of $\bA$ then obtaining $\bx$ would be trivial and there would be no need for the CG method in the first place.
+In short, a good preconditioner $\bM$ must approximate $\bA$ fairly accurately. However, if it is not much less expensive to 
+compute the inverse of $\bM$ than that of $\bA$ we actually gain nothing (see Eq.~\eqref{eq:prec}).
+
+If the diagonal entries of $\bA$ are much more significant than its off-diagonal entries then the 
+Jacobian preconditioner is a good choice of $\bM$. The Jacobian preconditioner is computed as follows:
+\begin{equation}
+\bM = \bA\odot\bI, 
+\end{equation}
+where $\bI$ is the identity matrix and $\odot$ denotes the Hadamard product. Since the Jacobian preconditioner is a diagonal matrix its inverse can be obtained trivially.
+
+\subsubsection{Normal Equation}
+\label{sec:normal}
+Lets assume that instead of Eq.~\eqref{eq:linear_system} we have a more general linear system, i.e.
+\begin{equation}
+ \bb = \bB\bx,
+\end{equation}
+where $\bb\in\mathbb{C}^K$, $\bx\in\mathbb{C}^P$  and $\bB\in\mathbb{C}^{K \times P}$, with $K > P$. We can use the normal equation 
+\begin{equation}
+\label{eq:normal_equation}
+\bB^H\bB\bx = \bB^H\bb, 
+\end{equation}
+to estimate the $\bx$ that minimizes
+\begin{equation}
+\label{eq:norm_ls}
+\|\bb-\bB\bx\|_F^2. 
+\end{equation}
+In this paper, we denote the Hermitian transpose with $(*)^H$ and the Frobenius norm with $\|*\|_F$.
+
+There are a few important observations that we can make while inspecting Eq.~\eqref{eq:normal_equation}:
+\begin{enumerate}
+\item $\bB^H\bB$ is a square positive semi-definite Hermitian matrix by construction.
+\item $\bB^H\bb$ is in the column range of $\bB^H\bB$; this follows trivially from 
+\begin{equation}
+\bB^H\bb \in \mathcal{R}(\bB^H) \implies \bB^H\bb \in \mathcal{R}(\bB^H\bB).   
+\end{equation}
+\end{enumerate}
+In this paper, we denote the column range of a matrix $\bA$ with $\mathcal{R}(\bA)$.
+
+Recall that we mentioned that in general we can only solve Eq.~\eqref{eq:linear_system} with CG if $\bA$ is a square positive-definite Hermitian matrix.
+The CG method actually has an even broader applicability: it can be used to solve Eq.~\eqref{eq:linear_system} even if $\bA$ is a Hermitian positive semi-definite
+square matrix as long as $\bb$ is in the column range of $\bA$ \citep{Lu2015}.
+In light of \citet{Lu2015}, the two observations in the numbered list we presented in the previous paragraph imply that Eq.~\eqref{eq:normal_equation}
+can be solved by employing the CG method. The solution returned by CG is, however, only unique if $\bB^H\bB$ is positive-definite (invertible). The solution
+obtained by CG, whether it is unique or not, does minimize Eq.~\eqref{eq:norm_ls} in a least squares sense.
+
+Note that if $\bA$ is positive semi-definite, its smallest eigenvalue is zero. This implies that Eq.~\eqref{eq:kappa} is not defined. \citet{Lu2015} show that when $\bA$
+is a square positive semi-definite Hermitian matrix we do not use the spectral condition number to bound the convergence complexity of CG, but the general spectral condition
+number instead. The general spectral condition number is very similar to the spectral condition number; the only difference being that in Eq.~\eqref{eq:kappa} we replace  
+$\lambda_{\textrm{min}}$ with $\lambda_{\textrm{min}}^+$, which denotes the smallest positive eigenvalue.
+
 \section{Redundant \textsc{StEfCal} and the ADI method}

### 2017-02-24 paper.tex
started implementing Gianni's suggestions

@@
 \end{equation}
-For example,
-\begin{equation}
-\{\alpha_{12},\alpha_{13},\alpha_{1N},\alpha_{23},\cdots,\alpha_{N-1N}\} = \{1,2,N-1,N,\cdots,B\},
-\end{equation}
-where $B$ denotes the number of baselines in the array, i.e. $B = \frac{N^2-N}{2}$. The vectors in Eq.~\eqref{eq:vec_definitions} are column vectors of size $B$ (i.e. $p<q$).
+For example, the set
+\begin{equation}
+\{\alpha_{12},\alpha_{13},\cdots,\alpha_{1N},\alpha_{23},\cdots,\alpha_{N-1N}\}
+\end{equation}
+and the set
+\begin{equation}
+\{1,2,\cdots,N-1,N,\cdots,B\},
+\end{equation}
+are equal. We denote the number of baselines in an array with $B$, i.e. $B = \frac{N^2-N}{2}$. The vectors in Eq.~\eqref{eq:vec_definitions} are column vectors of size $B$ (i.e. $p<q$).
 
@@
 \end{equation}
-where $\bg =[g_1,\cdots,g_N]^T$.
-
-least squares problems are normally solved by using gradient based methods, like Gauss-Newton and Levenberg-Marquardt .
+where $\bd$ is the observed visibility vector, $\bv$ is the model predicted visibility vector and $\bg =[g_1,\cdots,g_N]^T$.
+In
+
+Least-square type problems are generally solved by using gradient based minimization algorithms.
+In this paper we will be focusing on the following gradient minimization algorithms: GN and LM \citep{Levenberg1944,Marquardt1963}.
+To employ gradient based algorithms the objective function of a least-squares problem has to be differentiable with respect to the model parameters, which in the case of 
+Eq.~\ref{eq:least_squares} is
 

### 2017-02-27 paper.tex
more corrections

@@
 \begin{equation}
-\{1,2,\cdots,N-1,N,\cdots,B\},
+\{1,2,\cdots,N-1,N,\cdots,B\}
 \end{equation}
@@
 
-We can now formulate calibration as a least-squares problem:
+We can now formulate calibration as an optimization problem:
 \begin{equation}
@@
 \end{equation}
-where $\bd$ is the observed visibility vector, $\bv$ is the model predicted visibility vector and $\bg =[g_1,\cdots,g_N]^T$.
-In
-
-Least-square type problems are generally solved by using gradient based minimization algorithms.
+where $\bd$ is the observed visibility vector, $\bv$ is the model predicted visibility vector and $\bg =[g_1,\cdots,g_N]^T$. Closely inspecting Eq.~\eqref{eq:least_squares},
+we realize that it is in actual fact a least-squares problem, i.e. $\Lambda$ is the objective function, $\br$ is the residual vector, $\bd$ is the observed data vector, $\bv$ is the 
+model-predicted data vector and $\bg$ is the model-parameter vector that minimizes the objective function.
+
+The gradient-based minimization algorithms that are generally used to solve least squares problems generally require the model (in Eq.~\eqref{eq:least_squares} the model is denoted by $\bv$) to be differentiable
+towards each model parameter. In this paper we will be focusing on the following gradient-based minimization algorithms: GN and LM \citep{Levenberg1944,Marquardt1963}.
+When the least squares problem is complex it becomes less straightforward to apply these gradient-based minimization methods, as
+many complex functions are not analytic (no Taylor expansion around a fiducial point exists) and therefore not differentiable if the classic notion of differentiation is used. For instance, if the 
+classical definition of differentiation is used then $\frac{\partial \conj{z}}{\partial z}$, where $z \in \mathbb{C}$, does not exist.
+
+The only way to circumvent the differentiability conundrum associated with complex least square problems is to recast our complex optimization problem in such a way that $\bv$ becomes analytic in its argument.
+The standard way of achieving this is to divide the complex optimization problem into its real and imaginary parts and then to solve for the real and imaginary parts of the parameters separately.
+
+Recently, \citet{Sorber2012} proposed an alternative strategy, which involves the use of Wirtinger calculus \citep{Wirtinger1927}. The Wirtinger derivatives 
+are defined to be 
+\begin{align}
+\label{eq:wir}
+\frac{\partial}{\partial z} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} -  i \frac{\partial}{\partial y} \right ),&\frac{\partial}{\partial \conj{z}} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} +  i \frac{\partial}{\partial y} \right ).
+\end{align}
+Using the above definitions we can now easily compute the following:
+\begin{align}
+\frac{\partial z}{\partial z} & = 1, & \frac{\partial \conj{z}}{\partial z}&=0, & \frac{\partial z}{\partial \conj{z}} & = 0, & \frac{\partial \conj{z}}{\partial \conj{z}}&=1.
+\end{align}
+In this alternative approach we treat the complex parameters and their conjugates as separate variables.
+
+\citet{Smirnov2015} recently proposed to use the Wirtinger derivative to perform normal skymodel-based calibration. To achieve this we first recast Eq.~\eqref{eq:least_squares} to
+\begin{equation}
+\label{eq:least_squares_augmented}
+\min_{\breve{\bg}} \Lambda(\breve{\bg}) = \min_{\breve{\bg}} \|\breve{\br}\|_F^2 = \min_{\breve{\bg}} \|\breve{\bd} - \breve{\bv}(\breve{\bg})\|_F^2, 
+\end{equation}
+where $\breve{\br} = [\br^T,\conj{\br}^T]^T$, $\breve{\bd} = [\bd^T,\conj{\bd}^T]^T$, $\breve{\bv} = [\bv^T,\conj{\bv}^T]^T$ and $\breve{\bg} = [\bg^T,\conj{\bg}^T]^T$.
+
+Least-squares problems are generally solved by using gradient based minimization algorithms.
 In this paper we will be focusing on the following gradient minimization algorithms: GN and LM \citep{Levenberg1944,Marquardt1963}.
-To employ gradient based algorithms the objective function of a least-squares problem has to be differentiable with respect to the model parameters, which in the case of 
-Eq.~\ref{eq:least_squares} is
+
+To employ gradient based algorithms the objective function of a least-squares problem has to be differentiable with respect to the model parameters, which is not the case in 
+Eq.~\ref{eq:least_squares}. The most general way to deal with this difficulty is to split the least-squares problem into its real and imaginary parts and then to solve 
+for the real and imaginary parts of the model-parameter vector separately.
 

### 2017-02-27 paper.tex
more fixes

@@
 
-\subsection{Scalar Measurement Equation}
-\label{sec:s_mes}
+%\subsection{Scalar Measurement Equation}
+%\label{sec:s_mes}
 The observed visibility $d_{pq}$ measured by the baseline formed by antenna $p$ and $q$ is described by
@@
 The gradient-based minimization algorithms that are generally used to solve least squares problems generally require the model (in Eq.~\eqref{eq:least_squares} the model is denoted by $\bv$) to be differentiable
-towards each model parameter. In this paper we will be focusing on the following gradient-based minimization algorithms: GN and LM \citep{Levenberg1944,Marquardt1963}.
+towards each model parameter, i.e.
+these methods make use of a Jacobian matrix $\bJ$.
+In this paper we will be focusing on the following gradient-based minimization algorithms: GN and LM \citep{Levenberg1944,Marquardt1963}.
 When the least squares problem is complex it becomes less straightforward to apply these gradient-based minimization methods, as
@@
 
-\citet{Smirnov2015} recently proposed to use the Wirtinger derivative to perform normal skymodel-based calibration. To achieve this we first recast Eq.~\eqref{eq:least_squares} to
+\citet{Smirnov2015} recently proposed to use the Wirtinger derivatives to perform normal skymodel-based calibration. Eq.~\eqref{eq:least_squares} can be recast to
 \begin{equation}
@@
 where $\breve{\br} = [\br^T,\conj{\br}^T]^T$, $\breve{\bd} = [\bd^T,\conj{\bd}^T]^T$, $\breve{\bv} = [\bv^T,\conj{\bv}^T]^T$ and $\breve{\bg} = [\bg^T,\conj{\bg}^T]^T$. 
-
-Least-squares problems are generally solved by using gradient based minimization algorithms.
-In this paper we will be focusing on the following gradient minimization algorithms: GN and LM \citep{Levenberg1944,Marquardt1963}.
-
-To employ gradient based algorithms the objective function of a least-squares problem has to be differentiable with respect to the model parameters, which is not the case in 
-Eq.~\ref{eq:least_squares}. The most general way to deal with this difficulty is to split the least-squares problem into its real and imaginary parts and then to solve 
-for the real and imaginary parts of the model-parameter vector separately.
-
+The complex GN update step is:
+\begin{equation}
+\label{eq:GN_update}
+ \Delta \breve{\bg} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\br},
+\end{equation}
+with 
+\begin{equation}
+\label{eq:Jacobian}
+\bJ = \begin{bmatrix}
+       \bmJ & \bmJ^*\\
+       \conj{\bmJ}^* & \conj{\bmJ} 
+      \end{bmatrix},
+\end{equation}
+and
 \begin{align}
- \bg &=[g_1,\cdots,g_N]^T, & \by &= [y_1,\cdots,y_L]^T,\nonumber\\
- \bz &= [\bg^T,\by^T]^T. &  &\label{eq:parm_definitions}
+[\bmJ]_{\alpha_{pq},i} &= \frac{\partial v_{pq}}{\partial g_i}, & [\bmJ^*]_{\alpha_{pq},i} &= \frac{\partial v_{pq}}{\partial \conj{g}_i}. 
 \end{align}
+We refer to $\bJ$ as the Jacobian matrix. Note that the Wirtinger derivatives (see Eq.~\eqref{eq:wir}) are employed in Eq.~\eqref{eq:Jacobian}.
+The complex LM update rule is very similar, the major difference being a damping factor $\lambda$ is introduced:
+\begin{equation}
+\label{eq:LM_update}
+\Delta \breve{\bg} = (\bJ^H\bJ + \lambda\bD)^{-1}\bJ^H\breve{\br},
+\end{equation}
+where $\bD=\bI\odot\bJ^H\bJ$. We refer to $\bJ^H\bJ$ as the Hessian matrix $\bH$ and to $\bJ^H\bJ + \lambda\bD$ as the modified Hessian matrix $\bmH$.
+
+Eq.~\eqref{eq:GN_update} or Eq.~\eqref{eq:LM_update} is used to iteratively update the 
+parameter vector $\breve{\bg}$:
+\begin{equation}
+\label{eq:update}
+\breve{\bg}_{k+1} = \breve{\bg}_{k} + \Delta \breve{\bg}_{k}. 
+\end{equation}
+The parameter vector is updated until convergence is reached. To improve readability we do not explicitly denote the iteration subscript $k$ in 
+Eq.~\eqref{eq:GN_update} and Eq.~\eqref{eq:LM_update}.
 
@@
 \section{Redundant Wirtinger Calibration}
-
-\subsection{Redundant Measurement Equation}
-\label{sec:r_mes}
+\label{sec:red_wirtinger}
+
+Recall that redundant baselines sample the exact same visibilities in the $uv$-plane, i.e. if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$. 
+Let us now group all the baselines that measure the same $uv$-modes in a redundant array into the same redundant baseline group.
+Assigning the baselines of a redundant array to different redundant baseline groups has the advantage that it reduces the number of single index values we require.
+We now only require single index values for each redundant baseline group instead of a unique single index label for
+each baseline. Let $\phi_{pq}$ be the function that maps composite antenna pairs to the appropriate single redundant baseline group index. Note that $\phi_{pq}$ is not symmetric as 
+$\phi_{pq} = 0$ if $p\geq q$. We can also define the following symmetric variant of $\phi_{pq}$:
+\begin{equation}
+\zeta_{pq} = 
+\begin{cases}
+\phi_{pq}~\textrm{if}~p \leq q\\
+\phi_{qp}~\textrm{if}~p>q
+\end{cases}.
+\end{equation} 
+It is possible to construct a simple analytic expression for $\zeta_{pq}$ when our array is in an east-west regular configuration, namely $\zeta_{pq} = |q-p|$. 
+It becomes increasingly difficult to construct analytic expressions of $\zeta_{pq}$ for more complicated array layouts. 
+We will refer to $\zeta_{pq}$ as the symmetric geometric function. The empirically constructed symmetric geometry
+functions of three different redundant layouts are depicted in Fig.~\ref{fig:geometry_function}. We denote the range of $\zeta_{pq}$ with $\mathcal{R}(\zeta_{pq})$. The maximal element 
+that $\zeta_{pq}$ can ascertain is denoted by $L$ and can be interpreted as the maximal number of unique redundant baseline groups which can be formed for a given 
+array layout.
+Alternatively, we can interpret the symmetric geometry functions in Fig.~\ref{fig:geometry_function} as geometry matrices.
+In this alternative paradigm, $\zeta_{pq}$ denotes a matrix entry.
+The dimension of these geometry
+matrices are $N\times N$.
 If the array is redundant and the function $\phi_{pq}$ is known we can rewrite Eq.~\eqref{eq:vis_definition} as

### 2017-02-27 paper.tex
more Gianni fixes

@@
 \label{sec:red_wirtinger}
-
-Recall that redundant baselines sample the exact same visibilities in the $uv$-plane, i.e. if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$. 
-Let us now group all the baselines that measure the same $uv$-modes in a redundant array into the same redundant baseline group.
-Assigning the baselines of a redundant array to different redundant baseline groups has the advantage that it reduces the number of single index values we require.
-We now only require single index values for each redundant baseline group instead of a unique single index label for
-each baseline. Let $\phi_{pq}$ be the function that maps composite antenna pairs to the appropriate single redundant baseline group index. Note that $\phi_{pq}$ is not symmetric as 
-$\phi_{pq} = 0$ if $p\geq q$. We can also define the following symmetric variant of $\phi_{pq}$:
+Redundant baselines sample the exact same visibilities in the $uv$-plane, i.e. if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$. 
+We can therefore group all the baselines that measure the exact same visibilities into the same redundant baseline group. We index redundant baseline 
+groups with single group indices instead of composite antenna pairs. Let $\phi_{pq}$ be the function that maps the composite antenna pair associated with a specific baseline 
+to the single group index of the redundant baseline group to which the baseline in question belongs, i.e. if baseline $pq$ and $rs$ are redundant then $\phi_{pq} = \phi_{rs}$.  
+The function $\phi_{pq}$ is not symmetric as 
+$\phi_{pq} = 0$ if $p\geq q$. We can define the following symmetric variant of $\phi_{pq}$:
 \begin{equation}
@@
 \end{equation} 
+If the array is redundant we can rewrite Eq.~\eqref{eq:vis_definition} as
+\begin{equation}
+\label{eq:vis_red}
+d_{pq} = g_{p}\conj{g_q}y_{\phi_{pq}} + n_{pq},
+\end{equation}
+Eq.~\eqref{eq:vis_red} can also be expressed in the following vector form 
+\begin{equation}
+\label{eq:vis_red_vec}
+\bd = \bv + \bn, 
+\end{equation}
+where 
+\begin{align}
+ \left [ \bd \right]_{\alpha_{pq}} &= d_{pq}, & \left [ \bv \right ]_{\alpha_{pq}} &= v_{pq}=g_p y_{\phi_{pq}} \conj{g_q},\nonumber\\
+ \left [ \bn \right ]_{\alpha_{pq}} &= n_{pq}. &  &\label{eq:vec_definitions}
+\end{align}
+The vectors in Eq.~\eqref{eq:vec_definitions} are column vectors of size $B$ (i.e. $p<q$). 
 It is possible to construct a simple analytic expression for $\zeta_{pq}$ when our array is in an east-west regular configuration, namely $\zeta_{pq} = |q-p|$. 
@@
 matrices are $N\times N$.
-If the array is redundant and the function $\phi_{pq}$ is known we can rewrite Eq.~\eqref{eq:vis_definition} as
-\begin{equation}
-\label{eq:vis_red}
-d_{pq} = g_{p}\conj{g_q}y_{\phi_{pq}} + n_{pq}.
-\end{equation}
-Eq.~\eqref{eq:vis_red} can also be expressed in the following vector form 
-\begin{equation}
-\label{eq:vis_red_vec}
-\bd = \bv + \bn, 
-\end{equation}
-where 
-\begin{align}
- \left [ \bd \right]_{\alpha_{pq}} &= d_{pq}, & \left [ \bv \right ]_{\alpha_{pq}} &= v_{pq}=g_p y_{\phi_{pq}} \conj{g_q},\nonumber\\
- \left [ \bn \right ]_{\alpha_{pq}} &= n_{pq}. &  &\label{eq:vec_definitions}
-\end{align}
-The vectors in Eq.~\eqref{eq:vec_definitions} are column vectors of size $B$ (i.e. $p<q$).
 
@@
 
-\section{Results}
+\section{Comparing ADI and PCG}
 \label{sec:results}

### 2017-05-04 paper.tex
started implementing Gianni's comments

@@
 
-\section{Wirtinger Calibration}
-\label{sec:p_statement}
+\section{Skymodel-based Wirtinger Calibration}
+\label{sec:sky_wirtinger}
 %In Section~\ref{sec:s_mes}, we present the scalar measurement equation. Next, we define some indexing functions in Section~\ref{sec:indexing}. The measurement 
@@
 where $g_{p}$ and $g_{q}$ are the direction independent gains associated with antenna $p$ and $q$, $y_{pq}$ denotes the true visibility that baseline $pq$ measured
-and $n_{pq}$ is the thermal noise component.
-Eq.~\eqref{eq:vis_definition} is known as the scalar measurement equation.
-Conjugation is denoted by $\conj{(*)}$.
-The true value of $g_p$, $g_q$ and $y_{pq}$ are unknown.
-We normally try to estimate the true value of
+and $n_{pq}$ is the thermal noise component. Conjugation is denoted by $\conj{(*)}$. The true value of $g_p$, $g_q$ and $y_{pq}$ are unknown. We estimate the true value of
 $g_p$, $g_q$ and $y_{pq}$ by calibrating our data and then imaging the result.
@@
  \left [ \bd \right]_{\alpha_{pq}} &= d_{pq}, & \left [ \bv \right ]_{\alpha_{pq}} &= v_{pq}=g_p y_{pq} \conj{g_q},\nonumber\\
- \left [ \bn \right ]_{\alpha_{pq}} &= n_{pq}. &  &\label{eq:vec_definitions}
+ \left [ \bn \right ]_{\alpha_{pq}} &= n_{pq}, &  &\label{eq:vec_definitions}
 \end{align}
-The function $\alpha_{pq}$ maps composite antenna indexes to unique single indexes and is defined as
+and 
 \begin{equation}
@@
 \end{equation}
-For example, the set
+The function $\alpha_{pq}$ maps composite antenna indexes to unique single indexes, i.e. 
 \begin{equation}
@@
 \end{equation}
-and the set
+and 
 \begin{equation}
@@
 
-We can now formulate calibration as an optimization problem:
+We can now formulate calibration as a least-squares optimization problem:
 \begin{equation}
@@
 \end{equation}
-where $\bd$ is the observed visibility vector, $\bv$ is the model predicted visibility vector and $\bg =[g_1,\cdots,g_N]^T$. Closely inspecting Eq.~\eqref{eq:least_squares},
-we realize that it is in actual fact a least-squares problem, i.e. $\Lambda$ is the objective function, $\br$ is the residual vector, $\bd$ is the observed data vector, $\bv$ is the 
-model-predicted data vector and $\bg$ is the model-parameter vector that minimizes the objective function.
-
-The gradient-based minimization algorithms that are generally used to solve least squares problems generally require the model (in Eq.~\eqref{eq:least_squares} the model is denoted by $\bv$) to be differentiable
-towards each model parameter, i.e.
+where $\Lambda$ is the objective function, $\br$ is the residual vector and $\bg =[g_1,\cdots,g_N]^T$.
+
+The gradient-based minimization algorithms that are generally used to solve least squares problems generally require the model (in Eq.~\eqref{eq:least_squares} the model-predicted vector is denoted by $\bv$) to be differentiable
+towards each model parameter (in Eq.~\eqref{eq:least_squares} the model-parameter vector is denoted by $\bg$), i.e.
 these methods make use of a Jacobian matrix $\bJ$.
@@
 
-The only way to circumvent the differentiability conundrum associated with complex least square problems is to recast our complex optimization problem in such a way that $\bv$ becomes analytic in its argument.
-The standard way of achieving this is to divide the complex optimization problem into its real and imaginary parts and then to solve for the real and imaginary parts of the parameters separately.
+The only way to circumvent the differentiability conundrum associated with complex least squares problems is to recast our complex optimization problem in such a way that $\bv$ becomes analytic in its argument.
+The standard way of achieving this is to divide the complex optimization problem into its real and imaginary parts and then to solve for the real and imaginary parts of the unknown model-parameters separately.
 
@@
 \end{align}
-In this alternative approach we treat the complex parameters and their conjugates as separate variables.
-
-\citet{Smirnov2015} recently proposed to use the Wirtinger derivatives to perform normal skymodel-based calibration. Eq.~\eqref{eq:least_squares} can be recast to
+If we use the Wirtinger derivative, Eq.~\eqref{eq:wir}, instead of the classic notion of a derivative to define the gradient operator then 
+$\bv$ becomes analytic in $\bz$ and $\conj{\bz}$ as a whole.
+The Wirtinger derivative can therefore be used to derive complex variants of the real-valued GN and LM algorithms.
+In these alternative algorithms we treat the complex parameters and their conjugates as separate variables.
+
+\citet{Smirnov2015} recently demonstrated that the Wirtinger derivative can be used to perform normal skymodel-based calibration. We briefly summarize their approach below:
+\begin{enumerate}
+
+\item Eq.~\eqref{eq:least_squares} is first recast to
 \begin{equation}
@@
 where $\breve{\br} = [\br^T,\conj{\br}^T]^T$, $\breve{\bd} = [\bd^T,\conj{\bd}^T]^T$, $\breve{\bv} = [\bv^T,\conj{\bv}^T]^T$ and $\breve{\bg} = [\bg^T,\conj{\bg}^T]^T$.
-The complex GN update step is:
-\begin{equation}
-\label{eq:GN_update}
+\item The complex GN update step is then defined as
+\begin{equation}
+\label{eq:GN_update_skymodel}
  \Delta \breve{\bg} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\br},
@@
 \begin{equation}
-\label{eq:Jacobian}
+\label{eq:Jacobian_skymodel}
 \bJ = \begin{bmatrix}
@@
 \end{align}
-We refer to $\bJ$ as the Jacobian matrix. Note that the Wirtinger derivatives (see Eq.~\eqref{eq:wir}) are employed in Eq.~\eqref{eq:Jacobian}.
-The complex LM update rule is very similar, the major difference being a damping factor $\lambda$ is introduced:
-\begin{equation}
-\label{eq:LM_update}
+Note that the Wirtinger derivative is used in Eq.~\eqref{eq:Jacobian_skymodel} and as such the entries of $\bJ$ are well defined.
+\item The complex LM update rule is very similar, the major difference being a damping factor $\lambda$ is introduced:
+\begin{equation}
+\label{eq:LM_update_skymodel}
 \Delta \breve{\bg} = (\bJ^H\bJ + \lambda\bD)^{-1}\bJ^H\breve{\br},
 \end{equation}
-where $\bD=\bI\odot\bJ^H\bJ$. We refer to $\bJ^H\bJ$ as the Hessian matrix $\bH$ and to $\bJ^H\bJ + \lambda\bD$ as the modified Hessian matrix $\bmH$.
-
-Eq.~\eqref{eq:GN_update} or Eq.~\eqref{eq:LM_update} is used to iteratively update the 
+where $\bD=\bI\odot\bJ^H\bJ$.
+\item Eq.~\eqref{eq:GN_update_skymodel} or Eq.~\eqref{eq:LM_update_skymodel} can now be used iteratively to update the 
 parameter vector $\breve{\bg}$:
@@
 \end{equation}
-The parameter vector is updated until convergence is reached. To improve readability we do not explicitly denote the iteration subscript $k$ in 
-Eq.~\eqref{eq:GN_update} and Eq.~\eqref{eq:LM_update}.
-
+The parameter vector needs to be initialized (an all one vector is one possible choice) and is updated until convergence is reached.
+To improve readability we do not explicitly denote the iteration subscript $k$ in
+Eq.~\eqref{eq:GN_update_skymodel} and Eq.~\eqref{eq:LM_update_skymodel}.
+\end{enumerate}
+We refer to $\bJ^H\bJ$ as the Hessian matrix $\bH$ and to $\bJ^H\bJ + \lambda\bD$ as the modified Hessian matrix $\bmH$ in this paper. 
 % \subsection{Indexing Functions}
@@
 
+\begin{figure*}
+\centering
+\subfigure[Hexagonal layout]
+{\includegraphics[width=0.32\textwidth]{./HEX_lay.pdf}\label{fig:HEX_lay}}
+\subfigure[Square layout]
+{\includegraphics[width=0.32\textwidth]{./SQR_lay.pdf}\label{fig:SQR_lay}}
+\subfigure[Regular east-west layout]
+{\includegraphics[width=0.32\textwidth]{./REG_lay.pdf}\label{fig:REG_lay}}
+
+\subfigure[Hexagonal: $\zeta_{pq}$]
+{\includegraphics[width=0.33\textwidth]{./HEX_phi.pdf}\label{fig:HEX_phi}}
+\subfigure[Square: $\zeta_{pq}$]
+{\includegraphics[width=0.33\textwidth]{./SQR_phi.pdf}\label{fig:SQR_phi}}
+\subfigure[Regular east-west: $\zeta_{pq}$]
+{\includegraphics[width=0.33\textwidth]{./REG_phi.pdf}\label{fig:REG_phi}}
+\caption{In the top panel we have three different redundant antenna layouts, namely a hexagonal (left), square (middle) and a regular east-west (right) layout. 
+In the bottom panel we have the respective symmetric redundancy geometry functions for the layouts on the top panel. We used 91 antennas
+to construct the hexagonal layout, while 100 antennas where used in the square and east-west layouts.
+In the case of the east-west layout we only plot the positions of the first ten antennas.
+The maximal amount of redundant baseline groups that can be formed for
+the hexagonal, square and east-west layouts are 165, 180 and 99 respectively.\label{fig:geometry_function}}
+\end{figure*}
+
 We have now defined enough notation to enable us to formulate redundant calibration as a least-squares problem:
@@
 
-\subsection{Signal-to-noise Ratio}
-\label{sec:snr}
-The following SNR definition is used in this paper  
-\begin{equation}
-\label{}
-\textrm{SNR} = \frac{<\bv\odot\conj{\bv}>_{\nu,t,pq}}{<\bn\odot\conj{\bn}>_{\nu,t,pq}}, 
-\end{equation}
-where $<*>_{\nu,t,pq}$ denotes averaging over frequency, time and baseline. The definition we use here is based on the SNR definitions used in \citet{Liu2010} and \citet{Marthi2014}.
-
-\subsection{Limitations}
-\label{sec:scope}
-We only deal with per-timeslot and frequency channel calibration. No solution smoothing  
-is applied between the different timeslots and channels. The reader interested in this is referred to \citep{Zheng2014}. 
-Moreover, we do not explicitly deal with any of the degeneracies normally associated with redundant calibration \citep{Zheng2014,Kurien2016}.
+% \subsection{Limitations}
+% \label{sec:scope}
+% We only deal with per-timeslot and frequency channel calibration. No solution smoothing  
+% is applied between the different timeslots and channels. The reader interested in this is referred to \citep{Zheng2014}. 
+% Moreover, we do not explicitly deal with any of the degeneracies normally associated with redundant calibration \citep{Zheng2014,Kurien2016}.
 Similar to the approach taken in \citep{Marthi2014}, we assume that the degeneracies present in the redundant calibration solutions are removed with a final round of skymodel-based calibration.
 
-\section{Redundant Calibration}
-\label{sec:main_sec}
-As discussed in Section~\ref{sec:ls}, we can solve Eq.~\eqref{eq:least_squares_red} in one of two ways. We can split the problem into its real and
-imaginary parts and then solve for each part of the parameters seperately or we can employ the notion of Wirtinger derivatives. The former 
+As already alluded to in Section~\ref{sec:sky_wirtinger}, we can solve Eq.~\eqref{eq:least_squares_red} in one of two ways. We can split the problem into its real and
+imaginary parts and then solve for the real and imaginary parts of the unknown parameters separately or we can employ the notion of Wirtinger derivatives. The former 
 approach is discussed in Section~\ref{sec:ri}, while the latter approach is described in Section~\ref{sec:w}.
 
-\subsection{Real and imaginary Framework}
+\subsection{Real and imaginary approach}
 \label{sec:ri}
-As we mentioned in Section~\ref{sec:ls} using least-squares to solve complex problems can be problematic as $\bv$ needs to be analytic in its argument. In the case 
+Using least-squares to solve complex problems can be problematic as $\bv$ needs to be analytic in its argument. In the case 
 of Eq.~\eqref{eq:least_squares_red}, \citet{Liu2010} circumvents this difficulty by 
@@
 
-\subsection{Complex Framework}
+\subsection{Complex Approach}
 \label{sec:w}
 
-\citet{Smirnov2015} recently proposed to use the Wirtinger derivative instead of the classic real and imaginary approach to perform normal skymodel-based calibration. We will now show
-that the Wirtinger derivative can also be applied to redundant calibration.
-
+%\citet{Smirnov2015} recently proposed to use the Wirtinger derivative instead of the classic real and imaginary approach to perform normal skymodel-based calibration. We will now show
+%that the Wirtinger derivative can also be applied to redundant calibration.
 %We can perform redundant calibration by solving the following optimization problem: 
@@
 
-Analogous to \cite{Smirnov2015}, we now recast Eq.~\eqref{eq:least_squares_red} to 
+Analogous to \cite{Smirnov2015}, we recast Eq.~\eqref{eq:least_squares_red} to 
 \begin{equation}
@@
 where $\breve{\bz} = [\bz^T,\conj{\bz}^T]^T$. Eq.~\eqref{eq:least_squares_complex} is the complex conjugate augmented counterpart of Eq.~\eqref{eq:least_squares_red}.
-If we use the Wirtinger derivatives, Eq.~\eqref{eq:wir}, instead of the classic notion of a derivative to define the gradient operator then 
-$\bv$ becomes analytic in $\bz$ and $\conj{\bz}$ as a whole \citep{Smirnov2015}.
 The fact that $\bv$ is now analytic in its argument has the immediate consequence: it becomes possible to apply complex GN
@@
 \label{sec:complex_GN_LM}
-The complex Jacobian is defined to be
+The complex Jacobian associated with Eq.~\eqref{eq:least_squares_complex} is equal to
 \begin{equation}
@@
 where 
-\begin{align}
-\bmJ &= \frac{\partial \bv}{\partial \bz}, & \bmJ^* &= \frac{\partial \bv}{\partial \conj{\bz}}. 
-\end{align}
-In Eq.~\eqref{eq:Jacobian}, $\frac{\partial}{\partial \bz}$ and $\frac{\partial}{\partial \conj{\bz}}$ respectively denote the cogradient operator and the conjugate cogradient operator \citep{Sorber2012}.
+\begin{equation}
+[\bmJ]_{\alpha_{pq},i} = \begin{cases} 
+     \frac{\partial v_{pq}}{\partial g_i} & \textrm{if}~i\leq N \\
+     \frac{\partial v_{pq}}{\partial y_{i-N}} & \textrm{otherwise}  
+\end{cases}, %, & \bmJ^* &= \frac{\partial \bv}{\partial \conj{\bz}}. 
+\end{equation}
+and
+\begin{equation}
+[\bmJ^*]_{\alpha_{pq},i} = \begin{cases} 
+     \frac{\partial v_{pq}}{\partial \conj{g}_i} & \textrm{if}~i\leq N \\
+     \frac{\partial v_{pq}}{\partial \conj{y}_{i-N}} & \textrm{otherwise}  
+\end{cases}. %, & \bmJ^* &= \frac{\partial \bv}{\partial \conj{\bz}}. 
+\end{equation}
 The GN update step is now defined as:
@@
 \end{equation}
-The LM update rule is very similar, the major difference being a damping factor $\lambda$ is introduced:
+The LM update rule is defined in a very similar way:
 \begin{equation}
@@
 \end{equation}
-where $\bD=\bI\odot\bJ^H\bJ$. We refer to $\bJ^H\bJ$ as the Hessian matrix $\bH$ and to $\bJ^H\bJ + \lambda\bD$ as the modified Hessian matrix $\bmH$.
-
-We use Eq.~\eqref{eq:GN_update} or Eq.~\eqref{eq:LM_update} to iteratively update our 
+
+As mentioned in Section We use Eq.~\eqref{eq:GN_update} or Eq.~\eqref{eq:LM_update} to iteratively update our 
 parameter vector as follows:
@@
 
-\subsubsection{Analytic Expressions}
+\subsubsection{Simplifying GN}
+
+The following identities,
+\begin{align}
+\label{eq:identities}
+\frac{1}{3}\bJ\breve{\bz} &= \breve{\bv}, & \bJ^H\breve{\bv} &= \widetilde{\bH}\breve{\bz} 
+\end{align}
+are trivially established by mechanically showing that the left hand side of each identity in Eq.~\eqref{eq:identities} is equal to its right hand side.
+In Eq.~\eqref{eq:identities}, $\widetilde{\bH}$ denotes $\bI\odot\bH$.
+
+Replacing $\breve{\br}$ with $\breve{\bd}-\breve{\bv}$ in Eq.~\eqref{eq:GN_update} results in
+\begin{equation}
+\label{eq:temp_GN_eq}
+\Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H(\breve{\bd}-\breve{\bv}), 
+\end{equation}
+Substituting the first identity of Eq.~\eqref{eq:identities} into Eq.~\eqref{eq:temp_GN_eq} and simplifying the result 
+\begin{equation}
+\label{eq:one_thrid}
+\Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd}-\frac{1}{3}\breve{\bz}.
+\end{equation}
+It now trivially follows that:
+\begin{equation}
+\label{eq:two_thrids}
+\breve{\bz}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{2}{3}\breve{\bz}_{k}. 
+\end{equation}
+Eq.~\eqref{eq:two_thrids} shows us that in the case of redundant calibration we can calculate the GN update step without calculating the residual. 
+% The 
+% normal skymodel-based equivalent of Eq.~\eqref{eq:two_thrids} is equal to \citep{Smirnov2015}:
+% \begin{equation}
+% \label{eq:one_half}
+% \breve{\bg}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{1}{2}\breve{\bg}_{k}, 
+% \end{equation}
+% where $\breve{\bg} = [\bg^T,\conj{\bg}^T]^T$
+% 
+% FIX TRANSPOSE PROBLEM
+
+\section{Comparing ADI and PCG}
+\label{sec:results}
+In Section~\ref{sec:main_sec} we discussed the complex GN and LM algorithms and how they relate to redundant calibration. If we inspect Eq.~\eqref{eq:GN_update} and 
+Eq.~\eqref{eq:LM_update} we see that the most expensive operation used by the GN and LM algorithms is the inversion of the Hessian, $\bH$. The most logical 
+way of speeding up the algorithm is to investigate whether we can do this inversion in some efficient way.
+
+Two recent algorithms have been proposed that are able to invert $\bH$ in an efficient manner, namely the ADI \citep{Marthi2014} and PCG \citep{Liu2010} methods. 
+Note that here we focus solely on comparing the computational complexity of inverting $\bH$ using either one of these two methods. Rating the LM or GN algorithms
+using one of these approaches relative to each other once the computational complexity of inverting $\bH$ is known is trivial. The simulations we performed to compare these two 
+methods are summarized in Section~\ref{sec:simulations}.
+
+%We restrict ourselves to comparing the most 
+%expensive operation of the two methods namely, computing the inverse of the Hessian $\bH$ (or modified Hessian $\bmH$ in the case of LM). It becomes trivial to rate these two 
+%algorithms relative to each other once the computational complexity of the inverse operation is known.
+
+The basic difference between these two speedup algorithms can be understood by inspecting Fig.~\ref{fig:hessian}. The first observation we can make by inspecting 
+Fig.~\ref{fig:hessian_reg} and Fig.~\ref{fig:hessian_hex} is that the diagonal entries of $\bH$ are more significant than the off-diagonal entries in both cases.
+It therefore makes sense to approximate the $\bH$ with its diagonal, which greatly simplifies the computational complexity of inverting 
+$\bH$, i.e. it becomes $\mathbb{O}(P)$\footnote{The cost of constructing 
+the diagonal is not included in this estimate.} \citep{Smirnov2015}. This diagonal-approximation approach is effectively ADI \citep{Marthi2014}. The exact details of why this 
+is the case is discussed in Section~\ref{sec:adi}.
+We will also show how \citet{Marthi2014}, \citet{Smirnov2015} and \citet{Salvini2014} relate to each other in that same section.
+It should be noted here that strictly speaking the ADI method actually refers to the complete LM update step using the approximate $\bH$ and not just
+to the inversion of the approximate $\bH$.
+To improve the readability of the paper, we use the acronym ADI to refer to both the faster approximate inverse approach and the complete LM algorithm that uses it.
+Which one of these two concepts we are actually referring to, however, is always perfectly clear from the context.
+%Furhtermore, the name ADI also implies that with this algorithm we alternate between computing the 
+%gains, their conjugates and the true observed visibilities. %In Appendix~\ref{sec:sbc} we briefly summarize the ADI method that is used for skymodel-based calibration.
+
+The second observation we can make from Fig.~\ref{fig:hessian_reg} and Fig.~\ref{fig:hessian_hex} is that both Hessians contain many zero entries, i.e.
+they are sparse matrices.
+This, in conjunction with the more significant diagonal entries, implies that the PCG method (presented in Section~\ref{sec:conj_grad})
+can be applied to invert them in a computationally efficient manner.
+\citet{Liu2010} proposed the PCG method to compute the inverse of $\bH$, however the computational complexity of this approach was not analyzed therein.
+In Section~\ref{sec:pcg}, we do a proper computational complexity analysis of PCG. We find that the computational complexity of PCG is asymptotically bounded by $\mathbb{O}(P^2)$, which is 
+a lot better than straightforward inversion which takes $\mathbb{O}(P^3)$.
+
+\begin{figure*}
+\centering
+\subfigure[Regular layout]{\includegraphics[width=0.47\textwidth]{./reg_hessian.pdf}\label{fig:hessian_reg}}
+% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
+\subfigure[Hexagonal layout]{\includegraphics[width=0.47\textwidth]{./hex_hessian.pdf}\label{fig:hessian_hex}}
+\caption{The number of analytic terms out of which the entries of the Hessian $\bH$ consist for two different geometric layouts, namely a regular east-west grid with $N = 5$ (left panel) and 
+a hexagonal grid with $N = 7$. The diagonal entries of these two Hessians are clearly more significant than their off-diagonal entries. Moreover, these two Hessians also 
+contain many zero-entries. Note that the locations of the zero-entries are dependent on the geometry of the array layout.
+\label{fig:hessian}} 
+\end{figure*}
+
+%Since we approximate the Hessian in the ADI method with its diagonal the computational complexity of its inverse becomes $\mathbb{O}(P)$\footnote{The cost of constructing 
+%the diagonal is not included in this estimate.}.
+
+Even though the PCG method is computationally more expensive than the ADI method we cannot arrive at a final conclusion at this point. With the ADI method 
+we only approximate the inverse of the Hessian, while in the case of the PCG method we compute the exact inverse. So the question arises whether the LM algorithm using the PCG method 
+will converge much faster than the ADI method even though each iteration (which includes an inversion) is more expensive. We investigate this question in Section~\ref{sec:comparison}.
+
+\subsection{Simulations}
+\label{sec:simulations}
+%The main purpose of the simulations conducted in this paper is to genrate experimental results with which we can compare the computational complexity of ADI and PCG.
+We used an hexagonal geometric layout with an adjustable number of rings during the
+%simulations. 
+For the skymodel we generated 100 flat spectrum sources.
+Each source had a random flux value and position.
+The flux of the sources followed a Pareto distribution (with a shape parameter of 2), while their positions were uniformly
+distributed in a 3$^{\circ}$ by 3$^{\circ}$ square grid around the field center. The remaining fixed simulation parameters can all be found in Table~\ref{tab:fixed_parm}. 
+We also employed three different gain error models which we summarize in Table~\ref{tab:gain_parm}. We found that it did not matter which gain error model we used during
+our simulations, as the model we used did not have an impact on the results.
+Moreover, we had two major experimental setups, we either made use of simulated observations consisting of one frequency channel and multiple time-slots or one time-slot and multiple channels.
+We describe these two experimental setups in more detail in Table~\ref{tab:ch_parm}. As was the case for the gain error models, the experimental setup 
+that was used did not alter the simulation results.
+
+The main flow of a single simulation is as follows.
+We created corrupted visibilities using the fixed simulation parameters presented in Table~\ref{tab:fixed_parm} and either one of the experimental setups summarized in Table~\ref{tab:ch_parm}.
+We then used per time-slot and channel redundant calibration, i.e.
+we use the LM algorithm, which in turn makes use of either the ADI or PCG method, to
+estimate the gains and the true observed visibilities.
+Our initial parameter guess $\bz_0$ was chosen to be an all one vector.
+To obtain statistically significant results we performed, for both the ADI and PCG methods, multiple statistical independent simulations.
+
+\subsubsection{Signal-to-noise Ratio}
+\label{sec:snr}
+The following SNR definition is used in this paper  
+\begin{equation}
+\label{}
+\textrm{SNR} = \frac{<\bv\odot\conj{\bv}>_{\nu,t,pq}}{<\bn\odot\conj{\bn}>_{\nu,t,pq}}, 
+\end{equation}
+where $<*>_{\nu,t,pq}$ denotes averaging over frequency, time and baseline. The definition we use here is based on the SNR definitions used in \citet{Liu2010} and \citet{Marthi2014}.
+
+%We compare ADI with PCG in Section \textbf{??}, by using the results generated from the simulations we performed using 
+%these two methods.
+
+%The results reported in this paper were generated using two major experimental setups, namely a single time-slot and multiple frequency channels setup or 
+%a single frquency channel and multiple time-slots setup. The simulation parameters which where kept fixed in both these setups can be found in 
+%Table~\ref{tab:fixed_parm}. The sky-model we used to simulate the visibilities were the same for both setups.   We used a hexagonal redundant layout for our antenna positions. We produced 
+%results for different
+
+\begin{table*}
+\centering
+\caption{The following fixed simulation parameters were used to generate the results of this paper. We use $U$ to represent 
+a uniform distribution.}
+\begin{tabular}{|c c c c c c c|} 
+\hline
+Declination & Latitude & Minimum baseline & Geometry& Sources & Flux & Spatial \\
+\hline 
+\hline
+ $-74^{\circ}39'37.481''$ & $-30^{\circ}43'17.34''$ & 20 & Hexagonal &100 & Pareto (shape param. of 2) & $U[-3,3]$ (deg.) 
+\end{tabular}
+\label{tab:fixed_parm}
+\end{table*}
+
+\begin{table*}
+\centering
+\caption{The gain error models used in this paper. We have used the symbol $x$ here as a proxy as it can either refer to time-slots or channels. We either
+performed our simulations over multiple time-slots and one frequency channel or one timeslot and multiple frequencey channels (see Table~\ref{tab:ch_parm}).
+Moreover, $c$ in the left most column denotes the speed of light.}
+\begin{tabular}{|c c c c|} 
+\hline
+Number tag & 1 & 2 & 3\\
+Model & Sinusoidal: amplitude and phase & Sinusoidal: real and imaginary parts & Linear phase slope \\ [0.5ex] 
+\hline\hline
+Function & $(A+1)e^{jP}$ & $A\cos(2\pi fx+B)+1.5A+jC\sin(2\pi fx+D)$ & $e^{jP}$ \\ 
+\hline
+Parameters & $A=a\cos(2\pi fx +b)$  & $f=5$ & $P=\tau x$ \\
+ & $P =c \cos(2\pi fx +d)$ & $A,C\sim U[0.5,10]$ & $\tau = \frac{l}{c}$ \\
+ & $f=5$ & $B,D\sim U[0,2\pi]$ &  $l\sim U[5,50]$ (m)\\
+ & $a\sim U[0.8,0.9]$ &  & \\ 
+ & $c\sim U[0.5,5]$ &  &  \\ 
+ & $b,d\sim U[0,2\pi]$ &  &  \\ 
+\hline
+\end{tabular}
+\label{tab:gain_parm}
+\end{table*}
+
+\begin{table}
+\centering
+\caption{We generated results using two major experimental setups. We either used one frequency channel and multiple time-slots, or one time-slot and multiple frequency 
+channels. The most important parameters used in realizing these two major setups are presented here. Note that we refer to the gain error models in Table~\ref{tab:gain_parm} in the 
+last row of this table.}
+\begin{tabular}{|c c c|} 
+\hline
+ & Setup 1 & Setup 2\\
+\hline
+\hline
+ Num. channels & 1024 & 1\\
+$\nu$-range & 100-200 MHz & 1.4 GHz\\
+Num. timeslots & 1 & 50\\
+%Hour angle &-1^h & [-1^h,1^h]\\
+Gain error model (Table~\ref{tab:gain_parm}) & 2 & 1\\
+%Initial guess: $\bz_0$ & $\bone$ & $\bone$\\
+\hline
+\end{tabular}
+\label{tab:ch_parm}
+\end{table}
+
+\subsection{Alternating Direction Implicit Method}
+\label{sec:adi}
+%In Section~\ref{sec:adi} we show how the ADI method used for skymodel-based calibration relates to the redundant ADI method presented in \citet{Marthi2014}. 
+We start this section by briefly reviewing the ADI method as it pertains to skymodel-based calibration in Section~\ref{sec:sbc} \citep{Salvini2014,Smirnov2015}.
+
+In Section~\ref{sec:red_c}, we will show that the fast redundant calibration algorithm, that is presented in \citet{Marthi2014},
+is in fact nothing more than an extension of the skymodel-based ADI method, presented in \citet{Salvini2014}, to the redundant calibration use case (i.e. redundant \textsc{StEfCal}).
+In doing so we will justify the naming convention we have been using for it.
+
+In Section~\ref{sec:adi_results}, we present some ADI simulation results. 
+%We will mainly accomplish this by using the complex framework proposed by \citet{Smirnov2015}.
+
+\subsubsection{Skymodel-based Calibration}
+\label{sec:sbc}
+\textsc{StEfCal} is an ADI method in which the measurement equation, Eq.~\eqref{eq:vis_definition}, is linearized by assuming that the gains are known, but
+that their conjugates are unknown \citep{Mitchell:MWA-cal,Salvini2014}.
+Under this assumption, the conjugates of the gains can be obtained trivially.
+Once the conjugates of the gains have been computed, it is possible to solve for the gains.
+One can then alternate between computing the gains and their conjugates until convergence
+is reached.
+By correctly exploiting the symmetry inherent in the calibration problem, this alternating approach can be replaced by a single iterative new parameter estimation step.
+\cite{Salvini2014} also show that
+the convergence of the algorithm can be improved, if for every even iteration, the basic new parameter estimation step is modified as follows: the current gain estimate is replaced by the average of the current gain estimate and the gain estimate of the previous odd iteration.
+\textsc{StEFCal} lowers the computational complexity of normal skymodel-based calibration from $\mathbb{O}(N^3)$ to $\mathbb{O}(N^2)$.
+
+\citet{Smirnov2015} recently showed that one can also derive \textsc{StEFCal} by viewing normal skymodel-based calibration as a complex optimization problem.
+They realized while inspecting the skymodel-based $\bH$ that its diagonal entries are much more significant than its off-diagonal entries.
+As we mention in Section~\ref{sec:results}, the exact same conclusion can be reached if we inspect the Hessian $\bH$ that is associated with redundant calibration (see Fig.~\ref{fig:hessian}). 
+By approximating the skymodel-based $\bH$ by its diagonal and substituting this approximate Hessian $\widetilde{\bH}$ back into the fundamental skymodel-based LM update step we obtain \citep{Smirnov2015}
+\begin{align}
+\breve{\bg}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bg}_k,\label{eq:stef_lambda}\\
+ &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bg}_k, \label{eq:stef_alpha}  
+\end{align}
+where $\alpha = \frac{1}{1+\lambda}$. Moreover, $\lambda = \frac{1-\alpha}{\alpha}$.
+
+If we choose $\lambda$ to be zero, which corresponds to an $\alpha$ equal to one, we obtain the new parameter estimation step of \textsc{StEfCal} used during odd iterations.
+On the other hand, if we choose $\lambda$ equal to one (which corresponds
+to $\alpha=\frac{1}{2}$) we obtain the new parameter estimation step of \textsc{StEfCal} used during even iterations.
+
+Furthermore, the simplified skymodel-based GN equivalent of Eq.~\eqref{eq:two_thrids} is equal to
+\citep{Smirnov2015}
+\begin{equation}
+\label{eq:one_half}
+\breve{\bg}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{1}{2}\breve{\bg}_{k}. 
+\end{equation}
+Note that in Eq.~\eqref{eq:one_half} the influence that the estimate of the parameters from the previous iteration has on the current iteration is scaled by a half, while
+in Eq.~\eqref{eq:two_thrids} it is scaled by two thirds. Note the similarity between the last terms of Eq.~\eqref{eq:stef_alpha} and Eq.~\eqref{eq:one_half} when $\alpha$ is a half.
+
+\subsubsection{Redundant Calibration}
+\label{sec:red_c}
+In this section, we re-derive the fast redundant calibration algorithm presented in \citet{Marthi2014} using two different approaches.
+Both these approaches differ from the one presented in \citet{Marthi2014}.
+In \citet{Marthi2014}, the algorithm is derived by taking the derivative of the
+objective function $\Lambda$ relative to $\bg$ and $\by$, setting the intermediate results to naught and then solving for the unknown parameters.
+
+It is in fact trivial to extend the ADI method presented in \citet{Salvini2014} to the redundant use case, a.k.a.
+we now alternate between solving for the antenna gains their conjugates and the true observed visibilities.
+See Appendix~\ref{sec:red_stef_ADI} for the details.
+The result of this extension is that we obtain the redundant calibration algorithm presented in \citet{Marthi2014}.
+This implies that the fast redundant calibration algorithm presented in \citet{Marthi2014} can for all intents and purposes be referred to as redundant \textsc{StEfCal} (i.e.
+some kind of ADI method).
+
+Recall, that \citet{Smirnov2015} showed that \textsc{StEfCal} can also be derived by approximating the $\bH$, associated with skymodel-based calibration, with 
+its diagonal.
+To double check if our statement that the algorithm in \citet{Marthi2014} is in actual fact redundant \textsc{StEfCal} (i.e.
+some kind of ADI method) we will now use the diagonal approximation approach presented in \citet{Smirnov2015} to yet again re-derive the redundant calibration algorithm presented in \citet{Marthi2014}.
+The benefit of doing this second additional derivation is that it also enables us to show that the ADI method is in actual fact just a special case of the LM algorithm.
+
+Since the diagonal entries of the $\bH$, associated with redundant calibration, are more significant than its off-diagonal entries it makes 
+sense to try and approximate it with its diagonal (see Fig.~\ref{fig:hessian} and Eq.~\eqref{eq:red_H}).
+If we substitute $\bJ^H\bJ$ with $\widetilde{\bH}$ and replace $\breve{\br}$ with $\breve{\bd} - \breve{\bv}$ in Eq.~\eqref{eq:GN_update} we obtain
+\begin{equation}
+\label{eq:appr_GN}
+ \Delta \breve{\bz} \approx \widetilde{\bH}^{-1}\bJ^H(\breve{\bd}-\breve{\bv})
+\end{equation}
+Utilizing the second identity in Eq.~\eqref{eq:identities} allows us to simplify Eq.~\eqref{eq:appr_GN} to
+\begin{equation}
+  \Delta \breve{\bz} \approx \widetilde{\bH}^{-1}\bJ^H\breve{\bd}-\breve{\bz}.
+\end{equation}
+It now trivially follows that
+\begin{equation}
+ \breve{\bz}_{k+1} \approx \widetilde{\bH}^{-1}\bJ^H\breve{\bd}.
+\end{equation}
+We obtain a similar result if we repeat the above procedure, only in this case we use Eq.~\eqref{eq:LM_update} instead. Thus,
+\begin{align}
+\breve{\bz}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bz}_k,\label{eq:lambda}\\
+ &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bz}_k. \label{eq:alpha}  
+\end{align}
+The analytic expression of $\bJ^H\breve{\bd}$ will be very similar to the analytic 
+expression of $\bJ^H\breve{\br}$, the only difference being that in Eq.~\eqref{eq:ab} the letter $r$ would be replaced by a $d$. If we substitute the analytic expression
+of $\bJ^H\breve{\bd}$ and $\widetilde{\bH}^{-1}$ (which is easily computed as the matrix is diagonal) into Eq.~\eqref{eq:alpha} we obtain the following two update rules:
+\begin{equation}
+\label{eq:g_update}
+g_{i}^{k+1} = \alpha \frac{\sum_{j\neq i} g_j^k \widetilde{y}_{ij}^{~\!\!k} d_{ij}}{\sum_{j\neq i} |g_j^k|^2|y_{\zeta_{ij}}^k|^2} + (1-\alpha) g_i^k, 
+\end{equation}
+and
+\begin{equation}
+\label{eq:y_update}
+y_{i}^{k+1} = \alpha \frac{\sum_{pq \in \mathcal{PQ}_i} \conj{g}_p^k g_q^k d_{pq}}{\sum_{pq \in \mathcal{PQ}_i}|g_p^k|^2|g_q^k|^2} + (1-\alpha) y_i^k. 
+\end{equation}
+The computational complexity of inverting $\widetilde{\bH}$ is $\mathbb{O}(P)$.
+
+Eq.~\eqref{eq:g_update} and Eq.~\eqref{eq:y_update} are the exact same iterative new parameter estimators derived by \citet{Marthi2014}. Also note that Eq.~\eqref{eq:g_update} is the fundamental
+gain estimator associated with skymodel-based \textsc{StEfCal}.
+We have, therefore, successfully used the diagonal approximation approach proposed by \citet{Smirnov2015} to re-derive the fast redundant calibration method presented in
+\citet{Marthi2014}.
+%and in doing so we have confirmed shown that this algorithm is in actual fact nothing more than redundant \textsc{StEfCal}, i.e.
+some form of ADI method, justifying the naming convention we have been using for it.
+%Re-deriving the algorithm presented in \citet{Marthi2014} in this way serves as a further confirmation that it is indeed nothing more than redundant \textsc{StEfCal} (i.e. some kind of ADI method).
+
+The question now arises what value should we use for $\alpha$? In \citet{Marthi2014}, they propose 
+$\alpha = 0.3$. Although an exact mathematical proof for the optimal choice of $\alpha$ still eludes us, it is still possible to
+venture a clever guess by studying what the choice of $\alpha$ is in the case of skymodel-based \textsc{StEFCal}. The value of $\alpha$ is equal to a half 
+in the case of \textsc{StEFCal} (for even iterations); see Section~\ref{sec:sbc}.
+If we wish to mimic this behaviour for the redundant calibration use case we need to scale $\breve{\bz}$ by two thirds in Eq.~\eqref{eq:alpha} (see Eq.~\eqref{eq:two_thrids}) which implies that we should choose $\alpha$ equal to a third, which corresponds to a $\lambda$ equal to two.
+We used this $\lambda$ value in this paper.
+
+\subsubsection{ADI simulation results}
+\label{sec:adi_results}
+Optimality results of ADI were presented in \citet{Marthi2014}. We do not repeat their analysis here. We do, however, present our own ADI results, in Fig.~\ref{fig:prec_error}, that confirm that the
+ADI method can accurately estimate the antenna gains and the observed visibilities at different SNR values. To quantify the exact error in our estimates we used the percentage error $\beta$ defined 
+as 
+\begin{equation}
+\label{eq:beta}
+\beta = \frac{\|\bv - \widehat{\bv}\|_F^2}{\|\bv\|_F^2}.
+\end{equation}
+We have already defined $\bv$ in Eq.~\eqref{eq:vec_definitions}. Moreover, in Eq.~\eqref{eq:beta}, $\widehat{\bv}$ denotes the ADI estimate of $\bv$.
+
+\begin{figure}
+\includegraphics[width=0.47\textwidth]{./prec_error.pdf} 
+\caption{We plot the percentage error $\beta$ between the estimated 
+predicted visibilities and the true predicted visibilities for different SNR values as a function of the number of antennas in the array $N$.
+The graph presented here was generated using the first experimental setup in Table~\ref{tab:ch_parm}.}
+\label{fig:prec_error}
+\end{figure}
+
+% At this point there is no way of knowing if this is mere coincidence. DEF REWRITE THIS.
+% 
+% REWRTIE THIS IN A DIFFERENT WAY. START WITH CHENGULAR AND JUST MENTION THE LINK TO STEFCAL'S CHOICE OF ALPHA.
+% 
+% The update rules in Eq.~\eqref{eq:lambda} and Eq.~\eqref{eq:alpha} have been derived before, by Chengular... The approach they followed is discussed in the Appendix.
+% 
+% Interestingly enough, if we substitute $\alpha = 1$ into Eq.~\ref{eq:g_update} then it reduces to the skymodel-based StEFCal update rule. In Chengular they propose 
+% an $\alpha = 0.3$.
+% 
+% In an attempt to merge the terminology that has independently develop in the general calibration and redundant calibration literature and to emphasize the close
+% relationship between Stefcal and the approach derived by I will use the label Redunandat SteFCal (R-Stefcal) to refer to the NLS method proposed by ... 
+% 
+% ****\\
+% 
+% with $\alpha = \frac{1}{1+\lambda}$. Moreover $\lambda = \frac{1-\alpha}{\alpha}$. So in Marthi and Chengular (2014) they use an $\alpha$ of 0.3, which corresponds to an LM damping factor $\lambda$
+% of $2\frac{1}{3}$.
+% 
+% NB:: NEED TO ADD SNR PLOTS HERE
+% PLUS IMAGE OF HESSIAN FOR THE DISCUSSION
+
+\subsection{Preconditioned Conjugate Gradient Method}
+\label{sec:pcg}
+In this section we investigate the computational complexity of the PCG method when applied to the redundant calibration problem. 
+As we mentioned in Section~\ref{sec:conj_grad} there are two factors which influence the computational complexity of the PCG method, namely the spectral condition number and the sparsity of the matrix we wish to invert (see Eq.~\eqref{eq:cg_bound}).
+We used a $\lambda$ value of two, implying we used Eq.~\eqref{eq:LM_update} instead of Eq.~\eqref{eq:GN_update} throughout this section.
+We discuss the spectral condition number and the sparsity of the modified Hessian $\bmH$, which is a non-singular matrix, in Section~\ref{sec:scn} and Section~\ref{sec:sparsity} respectively.
+It is interesting to note, that the Hessian $\bH$ itself is singular which means that if we are to use Eq.~\eqref{eq:GN_update} instead we would need to be able to invert a singular matrix.
+In Section~\ref{sec:normal},
+we showed that the PCG method would in fact be able to handle such a scenario.
+
+\subsubsection{Spectral Condition Number}
+\label{sec:scn}
+In Section~\ref{sec:precon} we mentioned that the convergence properties of CG can be dramatically improved through preconditioning.
+Inspecting Fig.~\ref{fig:hessian} reveals that the Jacobian preconditioner
+is a good candidate for the redundant use case. In Fig.~\ref{fig:kappa} we see that this choice of preconditioner performs very well, the value of the spectral condition number $\kappa$
+becomes constant and independent of problem size, i.e. the number of antennas in the array. Preconditioning therefore effectively eliminates $\kappa$ from Eq.~\eqref{eq:cg_bound}. This result
+is confirmed by Fig.~\ref{fig:cg_itr}.
+Fig.~\ref{fig:cg_itr} shows us that the number of major iterations needed by the preconditioned conjugate gradient method to invert $\bmH$ is independent of the number of antennas in the array.
+Moreover, the number of major iterations required is also much smaller than the dimension of $\bmH$.
+
+%kappa + itr plots
+\begin{figure*}
+\centering
+\subfigure[$\kappa$]{\includegraphics[width=0.47\textwidth]{./kappa.pdf}\label{fig:kappa}}
+% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
+\subfigure[Iterations required before and after Preconditioning]{\includegraphics[width=0.47\textwidth]{./cg_itr.pdf}\label{fig:cg_itr}}
+\caption{The graphs presented here were generated using the second experimental setup in Table~\ref{tab:ch_parm}.
+In the left panel we have plotted the spectral condition number $\kappa$ of the modified Hessian $\bmH$, before (magenta and red curves) and after preconditioning (blue curve), at different SNR values as a function of the
+number of antennas $N$ in the array.
+The right hand panel confirms this result.
+In the right hand panel we have plotted the number of major iterations (line 6 of Algorithm \ref{algo:CG}) required by the conjugate
+gradient method to invert $\bmH$, before (magenta and red curves) and after preconditioning (blue and green curves), at different SNR values as a function of the number of antennas $N$ in the array.
+\label{fig:kappa_itr}} 
+\end{figure*}
+
+\subsubsection{Sparsity}
+\label{sec:sparsity}
+% \begin{enumerate}
+% \item $\boldsymbol{H}$: is a $(4N-2)\times(4N-2)$ or a $P \times P$ matrix. $N$ denotes the number of antennas and $P$ denotes the number of parameters. This matrix
+% contains $6N^2 - 2N - 2$ non-zero entries. It contains $16N^2 - 16N + 4$ entries. We can construct this matrix with $4N^2 -  4N$ or $\frac{1}{4} P^2 -1$ elementary operations.
+% \item $\boldsymbol{A}$: is a $(2N-1)\times(2N-1)$ matrix. This matrix has $N^2 + N - 1$ non-zero entries. We can construct this matrix with $2\frac{1}{2} (N^2 -  N)$ elementary operations.  
+% \item $\boldsymbol{B}$: is a $(2N-1)\times(2N-1)$ matrix. This matrix has $2N^2 - 2N$ non-zero entries. We can construct this matrix with $1\frac{1}{2} (N^2 -  N)$ elementary operations.
+% \item $\boldsymbol{C}$: is a $N\times N$ matrix. This matrix has $N$ non-zero entries. We can construct this matrix with $N^2-N$ elementary operations.
+% \item $\boldsymbol{D}$: is a $N \times (N-1)$ matrix. This matrix has $\frac{1}{2} (N^2 -  N)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations. 
+% \item $\boldsymbol{E}$: is a $(N-1) \times (N-1)$ matrix. This matrix has $(N-1)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.  
+% \item $\boldsymbol{F}$: is a $N \times N$ matrix. This matrix has $N^2 - N$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.
+% \item $\boldsymbol{G}$: is a $N \times (N-1)$ matrix. This matrix has $\frac{1}{2} (N^2 -  N)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.
+% \item $\boldsymbol{0}$: is a $(N-1) \times (N-1)$ all zero matrix. 
+% \end{enumerate}
+
+%We can construct $\boldsymbol{C}-\boldsymbol{G}$ in $O(N^2)$
+
+The second factor that determines the computational complexity of the conjugate gradient method when used to perform redundant calibration is the number of non-zero
+entries in the modified Hessian matrix $\bmH$. 
+%We have also shown in section that the first factor, the spectral condition number of $\bH$, can be reduced 
+%to a constant by employing preconditioning and can therefore be completely eliminated from the computational complexity equation. 
+A useful matrix quantity which we will use in this section is the sparsity factor $\gamma$ of a matrix, which is the ratio between the number of zero entries and the total number of entries in a matrix.
+We
+can therefore define $\gamma$ as follows:
+\begin{equation}
+ \gamma = \left (1 - \frac{m}{P^2} \right ) 
+\end{equation}
+It is trivial to determine an analytic expression for the sparsity ratio of $\bmH$ when our redundant array is in a regular east-west geometric configuration
+The sparsity and the asymptotic sparsity ratio of $\bmH$ are equal to
+\begin{align}
+\gamma &= \frac{5N^2-7N+3}{8N^2-8N+2} & \gamma_{\infty} &= \lim_{N\rightarrow \infty}\gamma = \frac{5}{8} \label{eq:gamma}. 
+\end{align}
+However when our redundant array makes use of a more complicated geometric layout it becomes much more difficult to construct such expressions, as $\bmH$ itself 
+becomes much harder to express analytically.
+
+In Fig.~\ref{fig:gamma}, we plot the empirically determined sparsity ratios for three different geometric layouts as a function of the number antennas in the array.
+Although, we cannot explicitly calculate
+the limit of the hexagonal and square layout curves, Fig.~\ref{fig:gamma} seems to indicate that they do exist. Moreover, we also plot the left most equation
+of Eq.~\eqref{eq:gamma} (red dashed line) and the right most equation of Eq.~\eqref{eq:gamma} (black dotted curve) in Fig.~\ref{fig:gamma}.
+
+We can now use the sparsity factor to compute the order of the computational complexity of inverting $\bmH$ (assuming the effect of $\kappa$ is negliable) with
+\begin{equation}
+P^{c} = (1 - \gamma)P^2.
+\end{equation}
+Solving for $c$ we obtain
+\begin{align}
+c &= \log_{P}(1 - \gamma) + 2 & c_{\infty} &= \lim_{N\rightarrow \infty} c = 2. \label{eq:c}
+\end{align}
+
+The evaluation of the limit in the rightmost equation of Eq.~\eqref{eq:c} is evaluated as follows: 
+note that
+\begin{equation}
+\label{eq:int_limit}
+\lim_{N\rightarrow\infty} P^{\log_{P}(1-\gamma)} = 1-\gamma_{\infty},
+\end{equation}
+which is only possible if 
+\begin{equation}
+\lim_{N\rightarrow\infty} \log_{P}(1-\gamma) = 0,
+\end{equation}
+implying that $c_{\infty}$ is equal to two.
+
+Eq.~\eqref{eq:int_limit}, implies that we can only compute $c_{\infty}$ in the rightmost equation if $\lim_{N\rightarrow \infty} \gamma$ exists. 
+In Fig.~\ref{fig:c}, we plot the empirically determined value of $c$ for three different geometric layouts as a function of the number antennas in the array.
+Furthermore, in Fig.~\ref{fig:c}, the analytic value of $c$ for the regular east-west grid is plotted as a dashed red curve, while the limit of this analytic curve is plotted with a dotted black curve.
+
+Eq.~\eqref{eq:gamma}, Eq.~\eqref{eq:c} and Fig.~\ref{fig:sparsity} indicate that the computational complexity, under the assumption that $\lim_{N\rightarrow \infty} \gamma$ exists, is asymptotically bounded by $O(P^2)$.
+However, the computational complexity converges very slowly to its asymptotic value, and in general is equal to $O(P^{c})$, where $c \sim 1.7$ if a Hexagonal geometric layout containing less than 200 antennas is employed.
+
+\begin{figure*}
+\centering
+\subfigure[$\gamma$]{\includegraphics[width=0.47\textwidth]{./sparsity.pdf}\label{fig:gamma}}
+% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
+\subfigure[$c$]{\includegraphics[width=0.46\textwidth]{./comp_order.pdf}\label{fig:c}}
+\caption{In the left panel we plot the sparsity ratio $\gamma$ (the number of zero entries relative to the total number of entries in a matrix) of the modified Hessian $\bmH$ as a function of the number
+of antennas $N$ in the array for different geometries, namely an hexagonal grid (blue circles), a square grid (green crosses) and a regular east-west grid (red circles).
+For the regular east-west grid we are able to construct an analytical expression for $\gamma$ (the dashed red line).
+Taking the limit
+of this analytic expression produces the black dotted line. In the right panel we plot the order of the computational cost $c$ for inverting $\bmH$ as a function of the number 
+of antennas in the array for different geometries. We use the same color scheme in the right and left panels. The red dashed line 
+in the right panel is associated with the theoretical curve of $c$ we computed for the east-west regular grid. The limit of the red-dashed curve is two, which we 
+have plotted with a black dotted line.
+\label{fig:sparsity}} 
+\end{figure*}
+
+\subsection{Comparison}
+\label{sec:comparison}
+We are now in a position to make a final comparison between the ADI and PCG methods.
+Inverting $\bH$ by assuming it is a diagonal matrix (i.e.
+using the ADI method) is very cheap, while inverting $\bH$ using PCG is more expensive.
+On the other side of the coin,  one inversion with PCG is much more
+accurate than using the approximate inverse which is associated with the ADI method.
+The question is now, if the inverse using the PCG method is much more accurate will the LM implementation using it require much less LM-iterations than its ADI counterpart.
+Moreover, will the number
+of LM iterations PCG requires be so few that it manages to beat the ADI method after all.
+To answer this we first need to compute the number of iterations, $k$, the LM implementation associated with the ADI method can make in the
+same time it takes the LM implementation associated with the PCG method to perform a single iteration. We can compute this with
+\begin{align}
+\label{eq:k}
+ kP &\approx P^{1.7}, & k &\approx P^{0.7}.
+\end{align}
+Next we need to measure the number of LM iterations we save by using the PCG method and compare it to Eq.\eqref{eq:k}. We make this comparison 
+in Fig.~\ref{fig:out_diff}. Fig.~\ref{fig:out_diff} tells us that the ADI method will outperform the PCG method as the number of LM iterations we save by taking the 
+full inverse is not enough to exceed $k$. In this comparison we have only taken into account the cost of inverting $\bH$. The PCG method might, however, still 
+turn out to be quite useful in practice as it would be more straightforward to apply it to other calibration use cases.
+
+\begin{figure*}
+\centering
+\subfigure[Number of LM iterations required by ADI and PCG.]{\includegraphics[width=0.47\textwidth]{./outerloop.pdf}\label{fig:outerloop}}
+% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
+\subfigure[Difference between ADI and PCG.]{\includegraphics[width=0.47\textwidth]{./diff.pdf}\label{fig:diff}}
+\caption{The graphs presented here were generated using the second experimental setup in Table~\ref{tab:ch_parm}. In the left panel we have the number of LM iterations 
+that is required for the LM algorithm using the ADI (green and blue curve) and the PCG (magenta and red curve) methods to converge to a parameter error tolerance of 1e-6 whilst using different 
+SNR values as a function of the number of antennas $N$ in the array.
+In the right panel we have the average amount of LM iterations (difference between the ADI and PCG curves in the left panel) we save (green and blue curve) by computing the full-inverse with the PCG method whilst using different SNR values.
+The black curve is associated with $k$, the number of iterations the LM implementation using the ADI method can perform in the same time it takes the LM implementation using 
+the PCG method to perform one iteration.
+\label{fig:out_diff}} 
+\end{figure*}
+
+\section{Conclusion}
+In Section~\ref{sec:main_sec} we showed that instead of using the traditional phase-and-amplitude approach to perform redundant calibration we can actually use a purely complex 
+framework. In this framework, the unknown variables and their conjugates are treated as independent variables that have to be estimated. 
+We then derived the GN and LM update steps using this complex framework (see Section~\ref{sec:complex_GN_LM}).
+In this paper we focused more on the LM algorithm than on the GN algorithm.
+The LM algorithm exhibits better convergence properties than the
+GN algorithm. This has the immediate consequence that the amount of algorithmic layers that are needed to perform redundant calibration are reduced.
+
+The most expensive operation associated with the LM algorithm (as well as the GN algorithm) is the inversion of $\bH$.
+We compared the computational complexity of two previously proposed algorithms to efficiently invert $\bH$ for the redundant calibration use case in Section~\ref{sec:results}, namely 
+ADI \citep{Marthi2014} and PCG \citep{Liu2010}.
+We showed that the computational complexity of the ADI method is $\mathbb{O}(P)$, while the computational complexity of the PCG method is asymptotically bounded
+by $\mathbb{O}(P^2)$. Both of these algorithms perform much better than stock standard inversion, which has a computational complexity of $\mathbb{O}(P^3)$.
+Although the ADI method does execute faster than the PCG method it is far less robust as it would be less straightforward to apply to other calibration uses cases. This is due to the fact that,
+the diagonal approximation we make with ADI might not be accurate enough in general.
+
+Lastly, we showed how \citet{Marthi2014}, \citet{Salvini2014} and \citet{Smirnov2015} relate to each other (see Section~\ref{sec:red_c}).
+The ADI method associated with skymodel-based calibration (i.e.
+\textsc{StEfCal}) is presented
+in \citet{Salvini2014}. \citet{Smirnov2015} showed that it is possible to re-derive this skymodel-based ADI method through complex optimization. In particular, they showed that if 
+we approximate the $\bH$ associated with the LM algorithm with its diagonal then we can rederive the ADI update rules presented in \citet{Salvini2014}. 
+\citet{Marthi2014} actually proposes to use the ADI method to perform redundant calibration.
+This fact is not immediately clear when superficially inspecting \citet{Marthi2014}.
+The redundant calibration algorithm presented in \citet{Marthi2014} was derived by taking the partial derivatives of
+the objective function towards the unknown parameters, setting the intermediate results to zero and then solving for the unknowns. In Section~\ref{sec:red_c}, we showed that 
+this algorithm is in actual fact an extension of the skymodel-based ADI method presented in \citet{Salvini2014} to the redundant calibration use case (i.e.
+we showed that this algorithm is in actual fact redundant \textsc{StEfCal}).
+We also re-derived the aforementioned redundant calibration algorithm by applying the diagonal approximation approach presented in \citet{Smirnov2015}, which serves as yet another additional confirmation
+that this algorithm is in fact redundant \textsc{StEfCal} (i.e. that it is some kind of ADI method).
+
+% fiducial initial guess 
+% 
+% The following partial derivatives are now easily computed
+% 
+% \begin{eqnarray}
+% \frac{c_{ij}}{\partial \eta_p} &=& y_{i-j} e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}\\ 
+% \frac{c_{ij}}{\partial \varphi_p} &=&  -i y_{i-j} e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}\\
+% \frac{c_{ij}}{\partial \eta_q} &=& y_{i-j} e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}\\ 
+% \frac{c_{ij}}{\partial \varphi_q} &=&  i y_{i-j} e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}\\
+% \frac{y_{i-j}}{\partial \varphi_q} &=&  e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}
+% \end{eqnarray}
+% 
+% 
+% The Wirtinger derivative is used in the last equation.
+% 
+% \begin{eqnarray}
+% c_{ij} &\approx& c_{ij}^0 + \Delta \eta_p(y_{i-j}^0 e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0}) + \Delta \eta_q(y_{i-j}^0 e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0})\\
+% && -i\Delta\varphi_p (y_{i-j}^0e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0}) +i\Delta\varphi_q (y_{i-j}^0e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0})\\
+% && y_{i-j}^1(y_{i-j}^0e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0})\\
+% &\approx&  c_{ij}^0 + e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_p+ \Delta \eta_q - i\Delta\varphi_p + i\Delta\varphi_p))
+% \end{eqnarray}
+% 
+% \begin{equation}
+% \label{eq:residual}
+% r_{ij} = \delta_{ij} = c_{ij}-c_{ij}^0 = e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_p+ \Delta \eta_q - i\Delta\varphi_p + i\Delta\varphi_p)) 
+% \end{equation}
+% 
+% Eq.~\ref{eq:residual} allows us to construct the following linear system:
+% 
+% \begin{equation}
+% \boldsymbol{J}[\boldsymbol{\Delta \eta},\boldsymbol{\Delta \varphi},\boldsymbol{\Delta y}]^T = \boldsymbol{r}, 
+% \end{equation}
+% 
+% where $\boldsymbol{J}$ is equal to 
+% \begin{equation}
+% \boldsymbol{J} = \bigg [\overbrace{\frac{c_{pq}}{\partial \eta_p}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial \varphi_p}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial y_{t}}}^{t=1\cdots r_s} \bigg ] \bigg \} [pq] = 1\cdots N_{\textrm{b}} (p<q)
+% \end{equation}
+% or
+% \begin{equation}
+% \boldsymbol{J} = \bigg [\frac{c_{pq}}{\partial \boldsymbol{\eta}},~\frac{c_{pq}}{\partial \boldsymbol{\varphi}},~\frac{c_{pq}}{\partial \boldsymbol{y}} \bigg ] \bigg \} [pq] = 1\cdots N_{\textrm{b}} (p<q).
+% \end{equation}
+% 
+% The last column is again a Wirtinger derivative.
+% 
+% Which means we can obtain the least-squares estimate as follows:
+% 
+% \begin{equation}
+% [\boldsymbol{\Delta \eta},\boldsymbol{\Delta \varphi},\boldsymbol{\Delta y}]^T = [\boldsymbol{J}^H\boldsymbol{J}]^{-1}\boldsymbol{J}^H\boldsymbol{r}.
+% \end{equation}
+
+\bibliographystyle{mn2e}
+\bibliography{paper}
+
+\appendix
+
+\section{\textsc{LOGCAL}}
+\label{sec:logcal}
+We can rewrite Eq.~\eqref{eq:vis_red} to \citep{Liu2010}
+\begin{eqnarray}
+\ln |d_{pq}| &=& \ln |g_p| + \ln |g_q| + \ln |y_{\phi_{pq}}| + s_{pq} \label{eq:logcal_amp}\\
+\angle d_{pq} &=& \angle g_p - \angle g_q + \angle \phi_{pq} + t_{pq} \label{eq:logcal_phase}
+\end{eqnarray}
+where
+\begin{align}
+s_{pq} &= \mathscr{R} \left \{1 + \frac{n_{pq}}{g_p\conj{g_q}y_{\phi_{pq}}} \right \}, & t_{pq} &= \mathscr{I}\left \{1 + \frac{n_{pq}}{g_p\conj{g_q}y_{\phi_{pq}}} \right \}.
+\end{align}
+Moreover, $\mathscr{R}\{*\}$ and $\mathscr{I}\{*\}$ denote the real and imaginary operators.
+
+With the aid of Eq.~\eqref{eq:logcal_amp} we can construct the following linear system:
+\begin{equation}
+\label{eq:log_system}
+\bdelta = \bJ\bzeta + \bs, 
+\end{equation}
+where
+\begin{align}
+\left [ \bdelta \right ]_{\alpha_{pq}} &= \ln |d_{pq}|, & \left [ \bs \right ]_{\alpha_{pq}} = s_{pq} 
+\end{align}
+and
+\begin{equation}
+\left [ \bzeta \right ]_{j} = \begin{cases} \ln |g_j| & \textrm{if}~j\leq N \\ \ln |y_{j-N}| & \textrm{otherwise} \end{cases}. 
+\end{equation}
+Furthermore,
+\begin{equation}
+\bJ = 
+\begin{bmatrix}
+\bN & \bM
+\end{bmatrix}
+\end{equation}
+with 
+\begin{equation}
+[\bN]_{\alpha_{pq},j} = \begin{cases}
+       1 & \textrm{if}~(p=j)~\textrm{or}~(q=j)\\
+       0 & \textrm{otherwise}
+      \end{cases}
+\end{equation}
+and
+\begin{equation}
+[\bM]_{\alpha_{pq},j} = \begin{cases}
+       1 & \textrm{if}~\phi_{pq}=j\\
+       0 & \textrm{otherwise}
+      \end{cases}
+\end{equation}
+
+The least squares solution to Eq.\eqref{eq:log_system}, under the incorrect assumption that $s_{pq}$ is normally distributed, is given by
+\begin{equation}
+\bzeta = (\bA^T\bA)^{-1} \bA^T\bdelta. 
+\end{equation}
+How to correctly deal with the fact that $s_{pq}$ is not normally distributed is discussed in \citep{Liu2010}. A similar linear system can be constructed for Eq.~\eqref{eq:logcal_phase}. 
+From the least squares solutions to the linear systems associated with Eq.~\eqref{eq:logcal_amp} and Eq.~\eqref{eq:logcal_phase} we can compute the antenna gains as well as the true visibilities.
+
+\section{\textsc{lincal}}
+\label{sec:lincal}
+%Assume that $\eta_p,\eta_q,\widetilde{\eta}_{\phi_{pq}},\varphi_p,\varphi_q$ and $\widetilde{\varphi}_{\phi_{pq}}$ are real valued variables.
+We can rewrite Eq.~\eqref{eq:vis_red_vec} to \citep{Liu2010}
+\begin{equation}
+\label{eq:lincal_vec}
+\bd = \bv(\bvarrho) + \bn, 
+\end{equation}
+where 
+\begin{align}
+\left [ \bv(\bvarrho) \right ]_{\alpha_{pq}} &= e^{\eta_p - i \varphi_p} e^{\widetilde{\eta}_{\phi_{pq}}- i \widetilde{\varphi}_{\phi_{pq}}} e^{\eta_q + i \varphi_q}\\
+&= v_{pq},
+\end{align}
+and
+\begin{equation}
+\bvarrho = 
+\begin{bmatrix}
+\bmath{\eta} \\
+\bvarphi \\
+\widetilde{\bmath{\eta}} \\
+\widetilde{\bvarphi}
+\end{bmatrix},\qquad
+\begin{aligned}
+\bmath{\eta} &= [\eta_1, \eta_2,\cdots \eta_N]^T\\
+ \bmath{\varphi} &= [\varphi_1, \varphi_2,\cdots \varphi_N]^T\\
+ \widetilde{\bmath{\eta}} &= [\widetilde{\eta}_1, \widetilde{\eta}_2,\cdots \widetilde{\eta}_L]^T\\
+ \widetilde{\bmath{\varphi}} &= [\widetilde{\varphi}_1, \widetilde{\varphi}_2,\cdots \widetilde{\varphi}_L]^T
+\end{aligned}.
+\label{eq:help}
+\end{equation}
+Note that the vectors $\bmath{\eta},~\bvarphi,~\widetilde{\bmath{\eta}}$ and $\widetilde{\bvarphi}$ are real valued.
+
+The first order Taylor expansion of Eq.~\eqref{eq:lincal_vec} around the fiducial point $\bvarrho_0$ is equal to
+\begin{equation}
+\label{eq:d_lin}
+\bd = \bv_0 + \bmJ_0 \Delta \bvarrho  + \bn.
+\end{equation}
+If we calculate the residual, dropping the reference to a specific fiducial point, we obtain: 
+\begin{equation}
+\label{eq:r_lin}
+\br = \bd - \bv = \bmJ \Delta \bvarrho + \bn.
+\end{equation}
+Moreover,
+\begin{equation}
+\label{eq:jac_upper_lincal}
+\bmJ = \begin{bmatrix}
+       \bN & \bM 
+       \end{bmatrix},
+\end{equation}
+where
+\begin{align}
+\bN &= \begin{bmatrix} \bO & \bP \end{bmatrix}, & \bM &= \begin{bmatrix} \bQ & \bR \end{bmatrix}, 
+\end{align}
+\begin{equation}
+ [\bO]_{\alpha_{pq},j} =  \frac{\partial v_{pq}}{\partial \eta_j} = \begin{cases} 
+    v_{pq} &\textrm{if}~(p=j)~\textrm{or}~(q=j)\\
+    0 & \textrm{otherwise}
+   \end{cases},
+\end{equation}
+\begin{equation}
+ [\bP]_{\alpha_{pq},j} =  \frac{\partial v_{pq}}{\partial \varphi_j} = \begin{cases}
+                                                                        i v_{pq} &\textrm{if}~ p=j\\
+                                                                        -i v_{pq} &\textrm{if}~ q=j \\
+                                                                        0 & \textrm{otherwise} 
+                                                                       \end{cases}
+\end{equation}
+\begin{equation}
+ [\bQ]_{\alpha_{pq},j} =  \frac{\partial v_{pq}}{\partial \widetilde{\eta}_j} = \begin{cases} 
+    v_{pq} &\textrm{if}~\phi_{pq}=j\\
+    0 & \textrm{otherwise}
+   \end{cases},
+\end{equation}
+\begin{equation}
+ [\bR]_{\alpha_{pq},j} =  \frac{\partial v_{pq}}{\partial \widetilde{\varphi}_j} = \begin{cases} 
+    i v_{pq} &\textrm{if}~\phi_{pq}=j\\
+    0 & \textrm{otherwise}
+   \end{cases}.
+\end{equation}
+Augmenting the residual by taking into account its conjugate results in 
+\begin{equation}
+\breve{\br} = \bJ \Delta \bvarrho + \bn, 
+\end{equation}
+where $\breve{\br} = \begin{bmatrix}\br^T & \conj{\br}^T \end{bmatrix}^T$ and 
+\begin{equation}
+\label{eq:Jac_lin}
+\bJ = 
+\begin{bmatrix}
+ \bmJ\\  
+ \conj{\bmJ}
+\end{bmatrix}.
+\end{equation}
+The solution to 
+\begin{equation}
+ \min_{\Delta \bvarrho} \| \breve{\br} - \bJ \Delta \bvarrho\|_F^2,
+\end{equation}
+is equal to 
+\begin{equation}
+\label{eq:LINCAL_update}
+\Delta \bvarrho = (\bJ^H\bJ)^{-1} \bJ^H\breve{\br}. 
+\end{equation}
+Eq.~\eqref{eq:LINCAL_update} is used to iteratively determine a new estimate of the parameter vector $\bvarrho$ as follows
+\begin{equation}
+\label{eq:new_rho}
+\bvarrho_{k+1} = \bvarrho_k + \Delta \bvarrho_k. 
+\end{equation}
+Note that we do not explicitly denote the $k$ index in Eq.~\eqref{eq:LINCAL_update}. We update our parameter vector until convergence is reached.
+
+\subsection{Conjugate Gradient Method}
+\label{sec:conj_grad}
+The CG method was independently discovered by \citet{Hestenes1973} and \citet{Stiefel1952}. They later published a joint paper, which is now considered as the seminal
+reference on CG \citep{Hestenes1952}. The iterative CG method is used to solve a particular class of linear system. The CG method is generally used to solve
+\begin{equation}
+\label{eq:linear_system}
+\bb = \bA\bx,
+\end{equation}
+where $\bb\in\mathbb{C}^P$ is a known complex column vector of size $P$, $\bx\in\mathbb{C}^P$ is an unknown complex column vector with the same length as $\bb$, and $\bA\in\mathbb{C}^{P\times P}$ is a square positive-definite Hermitian matrix with the appropriate dimensions.
+Moreover, using the CG method to solve Eq.~\eqref{eq:linear_system} will only be computationally advantageous if $\bA$ is also sparse (contain many zero entries).
+Using the iterative CG method to solve large sparse linear sytems was popularized by \citet{Reid1971}. A good tutorial on the CG method can be found in \citep{Shewchuk1994}.
+The iterative CG method is presented in Algorithm \ref{algo:CG}.
+
+%If $\bA$ is a square positive semi-definite Hermitian matrix, then the conjugate gradient method only converges if $\bb$ is in the column range of $\bA$ (cite Lu2016).
+
+%\begin{algorithm}
+%\caption{Conjugate Gradient Method.
+Inputs: $\bA$, $\bb$, a starting value for $\bx$, maximum number of iterations $k_{\textrm{max}}$ and an error tolerance $\epsilon<1$ Output: $\bx$ the solution to Eq.~\eqref{eq:linear_system}.
+\citep{Shewchuk1994}.}\label{algo:CG}
+%\begin{algorithmic}[1]
+%\State $k \gets 0$
+%\State $\br \gets \bb - \bA\bx$
+%\State $\bp \gets \br$
+%\State $\delta_{\textrm{new}} \gets \br^H\br$
+%\State $\delta_0 \gets \delta_{\textrm{new}}$
+%\While{$k\leq k_{\textrm{max}}$ and $\delta_{\textrm{new}} > \epsilon^2\delta_0$}
+%\State $\alpha \gets \frac{\delta_{\textrm{new}}}{\bp^H\bA\bp}$
+%\State $\bx \gets \bx + \alpha\bp$
+%\State $\br \gets \br - \alpha\bA\bp$
+%\State $\delta_{\textrm{old}} \gets \delta_{\textrm{new}}$ 
+%\State $\delta_{\textrm{new}} \gets \br^H\br$
+%\State $\beta \gets \frac{\delta_{\textrm{new}}}{\delta_{\textrm{old}}}$
+%\State $\bp \gets \br + \beta\bp$
+%\State $k \gets k + 1$
+%\EndWhile
+%\end{algorithmic}
+%\end{algorithm}
+
+\subsubsection{Computational Complexity}
+\citet{Kaniel1966} was able to determine convergence bounds for CG using Chebyshev polynomials. A much more in depth analysis of CG convergence can be found in
+\citep{Sluis1986}. As can be seen from Algorithm \ref{algo:CG}, the CG method makes use of a major outer loop (line 6 of Algorithm \ref{algo:CG}). It turns out that at worst 
+the number of major iterations that CG requires to converge\footnote{If the $\bA$-norm is used to calculate the error.} is proportional to $\sqrt{\kappa}$, where $\kappa$ is the spectral condition number of the matrix $\bA$.
+The mathematical definition of $\kappa(\bA)$ is
+\begin{equation}
+\label{eq:kappa}
+\kappa(\bA) = \frac{\lambda_{\textrm{max}}}{\lambda_{\textrm{min}}}, 
+\end{equation}
+where $\lambda_{\textrm{max}}$ and $\lambda_{\textrm{min}}$ respectively denote the largest and the smallest eigenvalue of $\bA$.
+Moreover, the most expensive operation that is conducted within each major loop of Algorithm~\ref{algo:CG} is the vector-matrix product $\bA\bp$, which has a complexity of $\mathbb{O}(m)$, where $m$ is the number of non-zero entries in $\bA$.
+The computational complexity of CG is therefore 
+\begin{equation}
+\label{eq:cg_bound}
+\mathbb{O}(\sqrt{\kappa}m). 
+\end{equation}
+It should now also be clear to the reader why the CG method is especially applicable to sparse linear systems. In sparse linear systems $m < P^2$.
+
+\subsubsection{Preconditioning}
+\label{sec:precon}
+Preconditioning is a technique which allows us to improve the condition number of a matrix. Let $\bM$ be a positive-definite Hermitian matrix.
+We can now indirectly solve for $\bx$ by using 
+\begin{equation}
+\label{eq:prec}
+\bM^{-1}\bA\bx = \bM^{-1}\bb.
+\end{equation}
+If $\bM$ is a good preconditioner then $\kappa(\bM^{-1}\bA) \ll \kappa(\bA)$, i.e. a good preconditioner lowers the condition number of a matrix which in turn increases the convergence speed 
+of CG (see Eq.~\eqref{eq:cg_bound}).
+In this paper, we denote matrix inversion with $(*)^{-1}$.
+The best possible choice of $\bM$ would therefore be $\bA$, which would lead to $\kappa$ becoming unity (as the largest and smallest eigenvalue of the identity matrix are equal to one).
+If $\kappa$ is one then Eq.~\eqref{eq:cg_bound} simplifies and becomes $\mathbb{O}(m)$.
+However, if we knew the inverse of $\bA$ then obtaining $\bx$ would be trivial and there would be no need for the CG method in the first place.
+In short, a good preconditioner $\bM$ must approximate $\bA$ fairly accurately. However, if it is not much less expensive to 
+compute the inverse of $\bM$ than that of $\bA$ we actually gain nothing (see Eq.~\eqref{eq:prec}).
+
+If the diagonal entries of $\bA$ are much more significant than its off-diagonal entries then the 
+Jacobian preconditioner is a good choice of $\bM$. The Jacobian preconditioner is computed as follows:
+\begin{equation}
+\bM = \bA\odot\bI, 
+\end{equation}
+where $\bI$ is the identity matrix and $\odot$ denotes the Hadamard product. Since the Jacobian preconditioner is a diagonal matrix its inverse can be obtained trivially.
+
+\subsubsection{Normal Equation}
+\label{sec:normal}
+Lets assume that instead of Eq.~\eqref{eq:linear_system} we have a more general linear system, i.e.
+\begin{equation}
+ \bb = \bB\bx,
+\end{equation}
+where $\bb\in\mathbb{C}^K$, $\bx\in\mathbb{C}^P$  and $\bB\in\mathbb{C}^{K \times P}$, with $K > P$. We can use the normal equation 
+\begin{equation}
+\label{eq:normal_equation}
+\bB^H\bB\bx = \bB^H\bb, 
+\end{equation}
+to estimate the $\bx$ that minimizes
+\begin{equation}
+\label{eq:norm_ls}
+\|\bb-\bB\bx\|_F^2. 
+\end{equation}
+In this paper, we denote the Hermitian transpose with $(*)^H$ and the Frobenius norm with $\|*\|_F$.
+
+There are a few important observations that we can make while inspecting Eq.~\eqref{eq:normal_equation}:
+\begin{enumerate}
+\item $\bB^H\bB$ is a square positive semi-definite Hermitian matrix by construction.
+\item $\bB^H\bb$ is in the column range of $\bB^H\bB$; this follows trivially from 
+\begin{equation}
+\bB^H\bb \in \mathcal{R}(\bB^H) \implies \bB^H\bb \in \mathcal{R}(\bB^H\bB).   
+\end{equation}
+\end{enumerate}
+In this paper, we denote the column range of a matrix $\bA$ with $\mathcal{R}(\bA)$.
+
+Recall that we mentioned that in general we can only solve Eq.~\eqref{eq:linear_system} with CG if $\bA$ is a square positive-definite Hermitian matrix.
+The CG method actually has an even broader applicability: it can be used to solve Eq.~\eqref{eq:linear_system} even if $\bA$ is a Hermitian positive semi-definite
+square matrix as long as $\bb$ is in the column range of $\bA$ \citep{Lu2015}.
+In light of \citet{Lu2015}, the two observations in the numbered list we presented in the previous paragraph imply that Eq.~\eqref{eq:normal_equation}
+can be solved by employing the CG method. The solution returned by CG is, however, only unique if $\bB^H\bB$ is positive-definite (invertible). The solution
+obtained by CG, whether it is unique or not, does minimize Eq.~\eqref{eq:norm_ls} in a least squares sense.
+
+Note that if $\bA$ is positive semi-definite, its smallest eigenvalue is zero. This implies that Eq.~\eqref{eq:kappa} is not defined. \citet{Lu2015} show that when $\bA$
+is a square positive semi-definite Hermitian matrix we do not use the spectral condition number to bound the convergence complexity of CG, but the general spectral condition
+number instead. The general spectral condition number is very similar to the spectral condition number; the only difference being that in Eq.~\eqref{eq:kappa} we replace  
+$\lambda_{\textrm{min}}$ with $\lambda_{\textrm{min}}^+$, which denotes the smallest positive eigenvalue.
+
+\section{Redundant Wirtinger Calibration - $\bJ$ and $\bH$}
 \label{sec:analytic}
@@
 \begin{equation}
-\mathcal{RS}_i = \left\{rs\in\mathbb{N}^2|(\phi_{rs} = i) \right\}.
+\mathcal{RS}_i = \left\{rs\in\mathbb{N}^2|(\phi_{rs} = i) \right\},
+\end{equation}
+\begin{equation}
+\xi_{ij} = 
+\begin{cases}
+p~\textrm{if}~\exists! ~ p \in \mathbb{N} ~ s.t. ~(\phi_{pi} = j)\\
+0~\textrm{otherwise}
+\end{cases},
+\end{equation}
+and
+\begin{equation}
+\psi_{ij} = 
+\begin{cases}
+q~\textrm{if}~\exists! ~ q \in \mathbb{N} ~ s.t. ~(\phi_{iq} = j)\\
+0~\textrm{otherwise}
+\end{cases}.
 \end{equation}
@@
 
-\subsubsection{Simplifying GN}
-
-The following identities,
-\begin{align}
-\label{eq:identities}
-\frac{1}{3}\bJ\breve{\bz} &= \breve{\bv}, & \bJ^H\breve{\bv} &= \widetilde{\bH}\breve{\bz} 
-\end{align}
-are trivially established by mechanically showing that the left hand side of each identity in Eq.~\eqref{eq:identities} is equal to its right hand side.
-In Eq.~\eqref{eq:identities}, $\widetilde{\bH}$ denotes $\bI\odot\bH$.
-
-Replacing $\breve{\br}$ with $\breve{\bd}-\breve{\bv}$ in Eq.~\eqref{eq:GN_update} results in
-\begin{equation}
-\label{eq:temp_GN_eq}
-\Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H(\breve{\bd}-\breve{\bv}), 
-\end{equation}
-Substituting the first identity of Eq.~\eqref{eq:identities} into Eq.~\eqref{eq:temp_GN_eq} and simplifying the result 
-\begin{equation}
-\label{eq:one_thrid}
-\Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd}-\frac{1}{3}\breve{\bz}.
-\end{equation}
-It now trivially follows that:
-\begin{equation}
-\label{eq:two_thrids}
-\breve{\bz}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{2}{3}\breve{\bz}_{k}. 
-\end{equation}
-Eq.~\eqref{eq:two_thrids} shows us that in the case of redundant calibration we can calculate the GN update step without calculating the residual. 
-% The 
-% normal skymodel-based equivalent of Eq.~\eqref{eq:two_thrids} is equal to \citep{Smirnov2015}:
-% \begin{equation}
-% \label{eq:one_half}
-% \breve{\bg}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{1}{2}\breve{\bg}_{k}, 
-% \end{equation}
-% where $\breve{\bg} = [\bg^T,\conj{\bg}^T]^T$
-% 
-% FIX TRANSPOSE PROBLEM
-
-\section{Comparing ADI and PCG}
-\label{sec:results}
-In Section~\ref{sec:main_sec} we discussed the complex GN and LM algorithms and how they relate to redundant calibration. If we inspect Eq.~\eqref{eq:GN_update} and 
-Eq.~\eqref{eq:LM_update} we see that the most expensive operation used by the GN and LM algorithms is the inversion of the Hessian, $\bH$. The most logical 
-way of speeding up the algorithm is to investigate whether we can do this inversion in some efficient way.
-
-Two recent algorithms have been proposed that are able to invert $\bH$ in an efficient manner, namely the ADI \citep{Marthi2014} and PCG \citep{Liu2010} methods. 
-Note that here we focus solely on comparing the computational complexity of inverting $\bH$ using either one of these two methods. Rating the LM or GN algorithms
-using one of these approaches relative to each other once the computational complexity of inverting $\bH$ is known is trivial. The simulations we performed to compare these two 
-methods are summarized in Section~\ref{sec:simulations}.
-
-%We restrict ourselves to comparing the most 
-%expensive operation of the two methods namely, computing the inverse of the Hessian $\bH$ (or modified Hessian $\bmH$ in the case of LM). It becomes trivial to rate these two 
-%algorithms relative to each other once the computational complexity of the inverse operation is known.
-
-The basic difference between these two speedup algorithms can be understood by inspecting Fig.~\ref{fig:hessian}. The first observation we can make by inspecting 
-Fig.~\ref{fig:hessian_reg} and Fig.~\ref{fig:hessian_hex} is that the diagonal entries of $\bH$ are more significant than the off-diagonal entries in both cases.
-It therefore makes sense to approximate the $\bH$ with its diagonal, which greatly simplifies the computational complexity of inverting 
-$\bH$, i.e. it becomes $\mathbb{O}(P)$\footnote{The cost of constructing 
-the diagonal is not included in this estimate.} \citep{Smirnov2015}. This diagonal-approximation approach is effectively ADI \citep{Marthi2014}. The exact details of why this 
-is the case is discussed in Section~\ref{sec:adi}.
-We will also show how \citet{Marthi2014}, \citet{Smirnov2015} and \citet{Salvini2014} relate to each other in that same section.
-It should be noted here that strictly speaking the ADI method actually refers to the complete LM update step using the approximate $\bH$ and not just
-to the inversion of the approximate $\bH$.
-To improve the readability of the paper, we use the acronym ADI to refer to both the faster approximate inverse approach and the complete LM algorithm that uses it.
-Which one of these two concepts we are actually referring to, however, is always perfectly clear from the context.
-%Furhtermore, the name ADI also implies that with this algorithm we alternate between computing the 
-%gains, their conjugates and the true observed visibilities. %In Appendix~\ref{sec:sbc} we briefly summarize the ADI method that is used for skymodel-based calibration.
-
-The second observation we can make from Fig.~\ref{fig:hessian_reg} and Fig.~\ref{fig:hessian_hex} is that both Hessians contain many zero entries, i.e.
-they are sparse matrices.
-This, in conjunction with the more significant diagonal entries, implies that the PCG method (presented in Section~\ref{sec:conj_grad})
-can be applied to invert them in a computationally efficient manner.
-\citet{Liu2010} proposed the PCG method to compute the inverse of $\bH$, however the computational complexity of this approach was not analyzed therein.
-In Section~\ref{sec:pcg}, we do a proper computational complexity analysis of PCG. We find that the computational complexity of PCG is asymptotically bounded by $\mathbb{O}(P^2)$, which is 
-a lot better than straightforward inversion which takes $\mathbb{O}(P^3)$.
-
-\begin{figure*}
-\centering
-\subfigure[Regular layout]{\includegraphics[width=0.47\textwidth]{./reg_hessian.pdf}\label{fig:hessian_reg}}
-% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
-\subfigure[Hexagonal layout]{\includegraphics[width=0.47\textwidth]{./hex_hessian.pdf}\label{fig:hessian_hex}}
-\caption{The number of analytic terms out of which the entries of the Hessian $\bH$ consist for two different geometric layouts, namely a regular east-west grid with $N = 5$ (left panel) and 
-a hexagonal grid with $N = 7$. The diagonal entries of these two Hessians are clearly more significant than their off-diagonal entries. Moreover, these two Hessians also 
-contain many zero-entries. Note that the locations of the zero-entries are dependent on the geometry of the array layout.
-\label{fig:hessian}} 
-\end{figure*}
-
-%Since we approximate the Hessian in the ADI method with its diagonal the computational complexity of its inverse becomes $\mathbb{O}(P)$\footnote{The cost of constructing 
-%the diagonal is not included in this estimate.}.
-
-Even though the PCG method is computationally more expensive than the ADI method we cannot arrive at a final conclusion at this point. With the ADI method 
-we only approximate the inverse of the Hessian, while in the case of the PCG method we compute the exact inverse. So the question arises whether the LM algorithm using the PCG method 
-will converge much faster than the ADI method even though each iteration (which includes an inversion) is more expensive. We investigate this question in Section~\ref{sec:comparison}.
-
-\subsection{Simulations}
-\label{sec:simulations}
-%The main purpose of the simulations conducted in this paper is to genrate experimental results with which we can compare the computational complexity of ADI and PCG.
-We used an hexagonal geometric layout with an adjustable number of rings during the
-%simulations. 
-For the skymodel we generated 100 flat spectrum sources.
-Each source had a random flux value and position.
-The flux of the sources followed a Pareto distribution (with a shape parameter of 2), while their positions were uniformly
-distributed in a 3$^{\circ}$ by 3$^{\circ}$ square grid around the field center. The remaining fixed simulation parameters can all be found in Table~\ref{tab:fixed_parm}. 
-We also employed three different gain error models which we summarize in Table~\ref{tab:gain_parm}. We found that it did not matter which gain error model we used during
-our simulations, as the model we used did not have an impact on the results.
-Moreover, we had two major experimental setups, we either made use of simulated observations consisting of one frequency channel and multiple time-slots or one time-slot and multiple channels.
-We describe these two experimental setups in more detail in Table~\ref{tab:ch_parm}. As was the case for the gain error models, the experimental setup 
-that was used did not alter the simulation results.
-
-The main flow of a single simulation is as follows.
-We created corrupted visibilities using the fixed simulation parameters presented in Table~\ref{tab:fixed_parm} and either one of the experimental setups summarized in Table~\ref{tab:ch_parm}.
-We then used per time-slot and channel redundant calibration, i.e.
-we use the LM algorithm, which in turn makes use of either the ADI or PCG method, to
-estimate the gains and the true observed visibilities.
-Our initial parameter guess $\bz_0$ was chosen to be an all one vector.
-To obtain statistically significant results we performed, for both the ADI and PCG methods, multiple statistical independent simulations.
-
-%We compare ADI with PCG in Section \textbf{??}, by using the results generated from the simulations we performed using 
-%these two methods.
-
-%The results reported in this paper were generated using two major experimental setups, namely a single time-slot and multiple frequency channels setup or 
-%a single frquency channel and multiple time-slots setup. The simulation parameters which where kept fixed in both these setups can be found in 
-%Table~\ref{tab:fixed_parm}. The sky-model we used to simulate the visibilities were the same for both setups.   We used a hexagonal redundant layout for our antenna positions. We produced 
-%results for different
-
-\begin{table*}
-\centering
-\caption{The following fixed simulation parameters were used to generate the results of this paper. We use $U$ to represent 
-a uniform distribution.}
-\begin{tabular}{|c c c c c c c|} 
-\hline
-Declination & Latitude & Minimum baseline & Geometry& Sources & Flux & Spatial \\
-\hline 
-\hline
- $-74^{\circ}39'37.481''$ & $-30^{\circ}43'17.34''$ & 20 & Hexagonal &100 & Pareto (shape param. of 2) & $U[-3,3]$ (deg.) 
-\end{tabular}
-\label{tab:fixed_parm}
-\end{table*}
-
-\begin{table*}
-\centering
-\caption{The gain error models used in this paper. We have used the symbol $x$ here as a proxy as it can either refer to time-slots or channels. We either
-performed our simulations over multiple time-slots and one frequency channel or one timeslot and multiple frequencey channels (see Table~\ref{tab:ch_parm}).
-Moreover, $c$ in the left most column denotes the speed of light.}
-\begin{tabular}{|c c c c|} 
-\hline
-Number tag & 1 & 2 & 3\\
-Model & Sinusoidal: amplitude and phase & Sinusoidal: real and imaginary parts & Linear phase slope \\ [0.5ex] 
-\hline\hline
-Function & $(A+1)e^{jP}$ & $A\cos(2\pi fx+B)+1.5A+jC\sin(2\pi fx+D)$ & $e^{jP}$ \\ 
-\hline
-Parameters & $A=a\cos(2\pi fx +b)$  & $f=5$ & $P=\tau x$ \\
- & $P =c \cos(2\pi fx +d)$ & $A,C\sim U[0.5,10]$ & $\tau = \frac{l}{c}$ \\
- & $f=5$ & $B,D\sim U[0,2\pi]$ &  $l\sim U[5,50]$ (m)\\
- & $a\sim U[0.8,0.9]$ &  & \\ 
- & $c\sim U[0.5,5]$ &  &  \\ 
- & $b,d\sim U[0,2\pi]$ &  &  \\ 
-\hline
-\end{tabular}
-\label{tab:gain_parm}
-\end{table*}
-
-\begin{table}
-\centering
-\caption{We generated results using two major experimental setups. We either used one frequency channel and multiple time-slots, or one time-slot and multiple frequency 
-channels. The most important parameters used in realizing these two major setups are presented here. Note that we refer to the gain error models in Table~\ref{tab:gain_parm} in the 
-last row of this table.}
-\begin{tabular}{|c c c|} 
-\hline
- & Setup 1 & Setup 2\\
-\hline
-\hline
- Num. channels & 1024 & 1\\
-$\nu$-range & 100-200 MHz & 1.4 GHz\\
-Num. timeslots & 1 & 50\\
-%Hour angle &-1^h & [-1^h,1^h]\\
-Gain error model (Table~\ref{tab:gain_parm}) & 2 & 1\\
-%Initial guess: $\bz_0$ & $\bone$ & $\bone$\\
-\hline
-\end{tabular}
-\label{tab:ch_parm}
-\end{table}
-
-\subsection{Alternating Direction Implicit Method}
-\label{sec:adi}
-%In Section~\ref{sec:adi} we show how the ADI method used for skymodel-based calibration relates to the redundant ADI method presented in \citet{Marthi2014}. 
-We start this section by briefly reviewing the ADI method as it pertains to skymodel-based calibration in Section~\ref{sec:sbc} \citep{Salvini2014,Smirnov2015}.
-
-In Section~\ref{sec:red_c}, we will show that the fast redundant calibration algorithm, that is presented in \citet{Marthi2014},
-is in fact nothing more than an extension of the skymodel-based ADI method, presented in \citet{Salvini2014}, to the redundant calibration use case (i.e. redundant \textsc{StEfCal}).
-In doing so we will justify the naming convention we have been using for it.
-
-In Section~\ref{sec:adi_results}, we present some ADI simulation results. 
-%We will mainly accomplish this by using the complex framework proposed by \citet{Smirnov2015}.
-
-\subsubsection{Skymodel-based Calibration}
-\label{sec:sbc}
-\textsc{StEfCal} is an ADI method in which the measurement equation, Eq.~\eqref{eq:vis_definition}, is linearized by assuming that the gains are known, but
-that their conjugates are unknown \citep{Mitchell:MWA-cal,Salvini2014}.
-Under this assumption, the conjugates of the gains can be obtained trivially.
-Once the conjugates of the gains have been computed, it is possible to solve for the gains.
-One can then alternate between computing the gains and their conjugates until convergence
-is reached.
-By correctly exploiting the symmetry inherent in the calibration problem, this alternating approach can be replaced by a single iterative new parameter estimation step.
-\cite{Salvini2014} also show that
-the convergence of the algorithm can be improved, if for every even iteration, the basic new parameter estimation step is modified as follows: the current gain estimate is replaced by the average of the current gain estimate and the gain estimate of the previous odd iteration.
-\textsc{StEFCal} lowers the computational complexity of normal skymodel-based calibration from $\mathbb{O}(N^3)$ to $\mathbb{O}(N^2)$.
-
-\citet{Smirnov2015} recently showed that one can also derive \textsc{StEFCal} by viewing normal skymodel-based calibration as a complex optimization problem.
-They realized while inspecting the skymodel-based $\bH$ that its diagonal entries are much more significant than its off-diagonal entries.
-As we mention in Section~\ref{sec:results}, the exact same conclusion can be reached if we inspect the Hessian $\bH$ that is associated with redundant calibration (see Fig.~\ref{fig:hessian}). 
-By approximating the skymodel-based $\bH$ by its diagonal and substituting this approximate Hessian $\widetilde{\bH}$ back into the fundamental skymodel-based LM update step we obtain \citep{Smirnov2015}
-\begin{align}
-\breve{\bg}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bg}_k,\label{eq:stef_lambda}\\
- &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bg}_k, \label{eq:stef_alpha}  
-\end{align}
-where $\alpha = \frac{1}{1+\lambda}$. Moreover, $\lambda = \frac{1-\alpha}{\alpha}$.
-
-If we choose $\lambda$ to be zero, which corresponds to an $\alpha$ equal to one, we obtain the new parameter estimation step of \textsc{StEfCal} used during odd iterations.
-On the other hand, if we choose $\lambda$ equal to one (which corresponds
-to $\alpha=\frac{1}{2}$) we obtain the new parameter estimation step of \textsc{StEfCal} used during even iterations.
-
-Furthermore, the simplified skymodel-based GN equivalent of Eq.~\eqref{eq:two_thrids} is equal to
-\citep{Smirnov2015}
-\begin{equation}
-\label{eq:one_half}
-\breve{\bg}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{1}{2}\breve{\bg}_{k}. 
-\end{equation}
-Note that in Eq.~\eqref{eq:one_half} the influence that the estimate of the parameters from the previous iteration has on the current iteration is scaled by a half, while
-in Eq.~\eqref{eq:two_thrids} it is scaled by two thirds. Note the similarity between the last terms of Eq.~\eqref{eq:stef_alpha} and Eq.~\eqref{eq:one_half} when $\alpha$ is a half.
-
-\subsubsection{Redundant Calibration}
-\label{sec:red_c}
-In this section, we re-derive the fast redundant calibration algorithm presented in \citet{Marthi2014} using two different approaches.
-Both these approaches differ from the one presented in \citet{Marthi2014}.
-In \citet{Marthi2014}, the algorithm is derived by taking the derivative of the
-objective function $\Lambda$ relative to $\bg$ and $\by$, setting the intermediate results to naught and then solving for the unknown parameters.
-
-It is in fact trivial to extend the ADI method presented in \citet{Salvini2014} to the redundant use case, a.k.a.
-we now alternate between solving for the antenna gains their conjugates and the true observed visibilities.
-See Appendix~\ref{sec:red_stef_ADI} for the details.
-The result of this extension is that we obtain the redundant calibration algorithm presented in \citet{Marthi2014}.
-This implies that the fast redundant calibration algorithm presented in \citet{Marthi2014} can for all intents and purposes be referred to as redundant \textsc{StEfCal} (i.e.
-some kind of ADI method).
-
-Recall, that \citet{Smirnov2015} showed that \textsc{StEfCal} can also be derived by approximating the $\bH$, associated with skymodel-based calibration, with 
-its diagonal.
-To double check if our statement that the algorithm in \citet{Marthi2014} is in actual fact redundant \textsc{StEfCal} (i.e.
-some kind of ADI method) we will now use the diagonal approximation approach presented in \citet{Smirnov2015} to yet again re-derive the redundant calibration algorithm presented in \citet{Marthi2014}.
-The benefit of doing this second additional derivation is that it also enables us to show that the ADI method is in actual fact just a special case of the LM algorithm.
-
-Since the diagonal entries of the $\bH$, associated with redundant calibration, are more significant than its off-diagonal entries it makes 
-sense to try and approximate it with its diagonal (see Fig.~\ref{fig:hessian} and Eq.~\eqref{eq:red_H}).
-If we substitute $\bJ^H\bJ$ with $\widetilde{\bH}$ and replace $\breve{\br}$ with $\breve{\bd} - \breve{\bv}$ in Eq.~\eqref{eq:GN_update} we obtain
-\begin{equation}
-\label{eq:appr_GN}
- \Delta \breve{\bz} \approx \widetilde{\bH}^{-1}\bJ^H(\breve{\bd}-\breve{\bv})
-\end{equation}
-Utilizing the second identity in Eq.~\eqref{eq:identities} allows us to simplify Eq.~\eqref{eq:appr_GN} to
-\begin{equation}
-  \Delta \breve{\bz} \approx \widetilde{\bH}^{-1}\bJ^H\breve{\bd}-\breve{\bz}.
-\end{equation}
-It now trivially follows that
-\begin{equation}
- \breve{\bz}_{k+1} \approx \widetilde{\bH}^{-1}\bJ^H\breve{\bd}.
-\end{equation}
-We obtain a similar result if we repeat the above procedure, only in this case we use Eq.~\eqref{eq:LM_update} instead. Thus,
-\begin{align}
-\breve{\bz}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bz}_k,\label{eq:lambda}\\
- &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bz}_k. \label{eq:alpha}  
-\end{align}
-The analytic expression of $\bJ^H\breve{\bd}$ will be very similar to the analytic 
-expression of $\bJ^H\breve{\br}$, the only difference being that in Eq.~\eqref{eq:ab} the letter $r$ would be replaced by a $d$. If we substitute the analytic expression
-of $\bJ^H\breve{\bd}$ and $\widetilde{\bH}^{-1}$ (which is easily computed as the matrix is diagonal) into Eq.~\eqref{eq:alpha} we obtain the following two update rules:
-\begin{equation}
-\label{eq:g_update}
-g_{i}^{k+1} = \alpha \frac{\sum_{j\neq i} g_j^k \widetilde{y}_{ij}^{~\!\!k} d_{ij}}{\sum_{j\neq i} |g_j^k|^2|y_{\zeta_{ij}}^k|^2} + (1-\alpha) g_i^k, 
-\end{equation}
-and
-\begin{equation}
-\label{eq:y_update}
-y_{i}^{k+1} = \alpha \frac{\sum_{pq \in \mathcal{PQ}_i} \conj{g}_p^k g_q^k d_{pq}}{\sum_{pq \in \mathcal{PQ}_i}|g_p^k|^2|g_q^k|^2} + (1-\alpha) y_i^k. 
-\end{equation}
-The computational complexity of inverting $\widetilde{\bH}$ is $\mathbb{O}(P)$.
-
-Eq.~\eqref{eq:g_update} and Eq.~\eqref{eq:y_update} are the exact same iterative new parameter estimators derived by \citet{Marthi2014}. Also note that Eq.~\eqref{eq:g_update} is the fundamental
-gain estimator associated with skymodel-based \textsc{StEfCal}.
-We have, therefore, successfully used the diagonal approximation approach proposed by \citet{Smirnov2015} to re-derive the fast redundant calibration method presented in
-\citet{Marthi2014}.
-%and in doing so we have confirmed shown that this algorithm is in actual fact nothing more than redundant \textsc{StEfCal}, i.e.
-some form of ADI method, justifying the naming convention we have been using for it.
-%Re-deriving the algorithm presented in \citet{Marthi2014} in this way serves as a further confirmation that it is indeed nothing more than redundant \textsc{StEfCal} (i.e. some kind of ADI method).
-
-The question now arises what value should we use for $\alpha$? In \citet{Marthi2014}, they propose 
-$\alpha = 0.3$. Although an exact mathematical proof for the optimal choice of $\alpha$ still eludes us, it is still possible to
-venture a clever guess by studying what the choice of $\alpha$ is in the case of skymodel-based \textsc{StEFCal}. The value of $\alpha$ is equal to a half 
-in the case of \textsc{StEFCal} (for even iterations); see Section~\ref{sec:sbc}.
-If we wish to mimic this behaviour for the redundant calibration use case we need to scale $\breve{\bz}$ by two thirds in Eq.~\eqref{eq:alpha} (see Eq.~\eqref{eq:two_thrids}) which implies that we should choose $\alpha$ equal to a third, which corresponds to a $\lambda$ equal to two.
-We used this $\lambda$ value in this paper.
-
-\subsubsection{ADI simulation results}
-\label{sec:adi_results}
-Optimality results of ADI were presented in \citet{Marthi2014}. We do not repeat their analysis here. We do, however, present our own ADI results, in Fig.~\ref{fig:prec_error}, that confirm that the
-ADI method can accurately estimate the antenna gains and the observed visibilities at different SNR values. To quantify the exact error in our estimates we used the percentage error $\beta$ defined 
-as 
-\begin{equation}
-\label{eq:beta}
-\beta = \frac{\|\bv - \widehat{\bv}\|_F^2}{\|\bv\|_F^2}.
-\end{equation}
-We have already defined $\bv$ in Eq.~\eqref{eq:vec_definitions}. Moreover, in Eq.~\eqref{eq:beta}, $\widehat{\bv}$ denotes the ADI estimate of $\bv$.
-
-\begin{figure}
-\includegraphics[width=0.47\textwidth]{./prec_error.pdf} 
-\caption{We plot the percentage error $\beta$ between the estimated 
-predicted visibilities and the true predicted visibilities for different SNR values as a function of the number of antennas in the array $N$.
-The graph presented here was generated using the first experimental setup in Table~\ref{tab:ch_parm}.}
-\label{fig:prec_error}
-\end{figure}
-
-% At this point there is no way of knowing if this is mere coincidence. DEF REWRITE THIS.
-% 
-% REWRTIE THIS IN A DIFFERENT WAY. START WITH CHENGULAR AND JUST MENTION THE LINK TO STEFCAL'S CHOICE OF ALPHA.
-% 
-% The update rules in Eq.~\eqref{eq:lambda} and Eq.~\eqref{eq:alpha} have been derived before, by Chengular... The approach they followed is discussed in the Appendix.
-% 
-% Interestingly enough, if we substitute $\alpha = 1$ into Eq.~\ref{eq:g_update} then it reduces to the skymodel-based StEFCal update rule. In Chengular they propose 
-% an $\alpha = 0.3$.
-% 
-% In an attempt to merge the terminology that has independently develop in the general calibration and redundant calibration literature and to emphasize the close
-% relationship between Stefcal and the approach derived by I will use the label Redunandat SteFCal (R-Stefcal) to refer to the NLS method proposed by ... 
-% 
-% ****\\
-% 
-% with $\alpha = \frac{1}{1+\lambda}$. Moreover $\lambda = \frac{1-\alpha}{\alpha}$. So in Marthi and Chengular (2014) they use an $\alpha$ of 0.3, which corresponds to an LM damping factor $\lambda$
-% of $2\frac{1}{3}$.
-% 
-% NB:: NEED TO ADD SNR PLOTS HERE
-% PLUS IMAGE OF HESSIAN FOR THE DISCUSSION
-
-\subsection{Preconditioned Conjugate Gradient Method}
-\label{sec:pcg}
-In this section we investigate the computational complexity of the PCG method when applied to the redundant calibration problem. 
-As we mentioned in Section~\ref{sec:conj_grad} there are two factors which influence the computational complexity of the PCG method, namely the spectral condition number and the sparsity of the matrix we wish to invert (see Eq.~\eqref{eq:cg_bound}).
-We used a $\lambda$ value of two, implying we used Eq.~\eqref{eq:LM_update} instead of Eq.~\eqref{eq:GN_update} throughout this section.
-We discuss the spectral condition number and the sparsity of the modified Hessian $\bmH$, which is a non-singular matrix, in Section~\ref{sec:scn} and Section~\ref{sec:sparsity} respectively.
-It is interesting to note, that the Hessian $\bH$ itself is singular which means that if we are to use Eq.~\eqref{eq:GN_update} instead we would need to be able to invert a singular matrix.
-In Section~\ref{sec:normal},
-we showed that the PCG method would in fact be able to handle such a scenario.
-
-\subsubsection{Spectral Condition Number}
-\label{sec:scn}
-In Section~\ref{sec:precon} we mentioned that the convergence properties of CG can be dramatically improved through preconditioning.
-Inspecting Fig.~\ref{fig:hessian} reveals that the Jacobian preconditioner
-is a good candidate for the redundant use case. In Fig.~\ref{fig:kappa} we see that this choice of preconditioner performs very well, the value of the spectral condition number $\kappa$
-becomes constant and independent of problem size, i.e. the number of antennas in the array. Preconditioning therefore effectively eliminates $\kappa$ from Eq.~\eqref{eq:cg_bound}. This result
-is confirmed by Fig.~\ref{fig:cg_itr}.
-Fig.~\ref{fig:cg_itr} shows us that the number of major iterations needed by the preconditioned conjugate gradient method to invert $\bmH$ is independent of the number of antennas in the array.
-Moreover, the number of major iterations required is also much smaller than the dimension of $\bmH$.
-
-%kappa + itr plots
-\begin{figure*}
-\centering
-\subfigure[$\kappa$]{\includegraphics[width=0.47\textwidth]{./kappa.pdf}\label{fig:kappa}}
-% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
-\subfigure[Iterations required before and after Preconditioning]{\includegraphics[width=0.47\textwidth]{./cg_itr.pdf}\label{fig:cg_itr}}
-\caption{The graphs presented here were generated using the second experimental setup in Table~\ref{tab:ch_parm}.
-In the left panel we have plotted the spectral condition number $\kappa$ of the modified Hessian $\bmH$, before (magenta and red curves) and after preconditioning (blue curve), at different SNR values as a function of the
-number of antennas $N$ in the array.
-The right hand panel confirms this result.
-In the right hand panel we have plotted the number of major iterations (line 6 of Algorithm \ref{algo:CG}) required by the conjugate
-gradient method to invert $\bmH$, before (magenta and red curves) and after preconditioning (blue and green curves), at different SNR values as a function of the number of antennas $N$ in the array.
-\label{fig:kappa_itr}} 
-\end{figure*}
-
-\subsubsection{Sparsity}
-\label{sec:sparsity}
-% \begin{enumerate}
-% \item $\boldsymbol{H}$: is a $(4N-2)\times(4N-2)$ or a $P \times P$ matrix. $N$ denotes the number of antennas and $P$ denotes the number of parameters. This matrix
-% contains $6N^2 - 2N - 2$ non-zero entries. It contains $16N^2 - 16N + 4$ entries. We can construct this matrix with $4N^2 -  4N$ or $\frac{1}{4} P^2 -1$ elementary operations.
-% \item $\boldsymbol{A}$: is a $(2N-1)\times(2N-1)$ matrix. This matrix has $N^2 + N - 1$ non-zero entries. We can construct this matrix with $2\frac{1}{2} (N^2 -  N)$ elementary operations.  
-% \item $\boldsymbol{B}$: is a $(2N-1)\times(2N-1)$ matrix. This matrix has $2N^2 - 2N$ non-zero entries. We can construct this matrix with $1\frac{1}{2} (N^2 -  N)$ elementary operations.
-% \item $\boldsymbol{C}$: is a $N\times N$ matrix. This matrix has $N$ non-zero entries. We can construct this matrix with $N^2-N$ elementary operations.
-% \item $\boldsymbol{D}$: is a $N \times (N-1)$ matrix. This matrix has $\frac{1}{2} (N^2 -  N)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations. 
-% \item $\boldsymbol{E}$: is a $(N-1) \times (N-1)$ matrix. This matrix has $(N-1)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.  
-% \item $\boldsymbol{F}$: is a $N \times N$ matrix. This matrix has $N^2 - N$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.
-% \item $\boldsymbol{G}$: is a $N \times (N-1)$ matrix. This matrix has $\frac{1}{2} (N^2 -  N)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.
-% \item $\boldsymbol{0}$: is a $(N-1) \times (N-1)$ all zero matrix. 
-% \end{enumerate}
-
-%We can construct $\boldsymbol{C}-\boldsymbol{G}$ in $O(N^2)$
-
-The second factor that determines the computational complexity of the conjugate gradient method when used to perform redundant calibration is the number of non-zero
-entries in the modified Hessian matrix $\bmH$. 
-%We have also shown in section that the first factor, the spectral condition number of $\bH$, can be reduced 
-%to a constant by employing preconditioning and can therefore be completely eliminated from the computational complexity equation. 
-A useful matrix quantity which we will use in this section is the sparsity factor $\gamma$ of a matrix, which is the ratio between the number of zero entries and the total number of entries in a matrix.
-We
-can therefore define $\gamma$ as follows:
-\begin{equation}
- \gamma = \left (1 - \frac{m}{P^2} \right ) 
-\end{equation}
-It is trivial to determine an analytic expression for the sparsity ratio of $\bmH$ when our redundant array is in a regular east-west geometric configuration
-The sparsity and the asymptotic sparsity ratio of $\bmH$ are equal to
-\begin{align}
-\gamma &= \frac{5N^2-7N+3}{8N^2-8N+2} & \gamma_{\infty} &= \lim_{N\rightarrow \infty}\gamma = \frac{5}{8} \label{eq:gamma}. 
-\end{align}
-However when our redundant array makes use of a more complicated geometric layout it becomes much more difficult to construct such expressions, as $\bmH$ itself 
-becomes much harder to express analytically.
-
-In Fig.~\ref{fig:gamma}, we plot the empirically determined sparsity ratios for three different geometric layouts as a function of the number antennas in the array.
-Although, we cannot explicitly calculate
-the limit of the hexagonal and square layout curves, Fig.~\ref{fig:gamma} seems to indicate that they do exist. Moreover, we also plot the left most equation
-of Eq.~\eqref{eq:gamma} (red dashed line) and the right most equation of Eq.~\eqref{eq:gamma} (black dotted curve) in Fig.~\ref{fig:gamma}.
-
-We can now use the sparsity factor to compute the order of the computational complexity of inverting $\bmH$ (assuming the effect of $\kappa$ is negliable) with
-\begin{equation}
-P^{c} = (1 - \gamma)P^2.
-\end{equation}
-Solving for $c$ we obtain
-\begin{align}
-c &= \log_{P}(1 - \gamma) + 2 & c_{\infty} &= \lim_{N\rightarrow \infty} c = 2. \label{eq:c}
-\end{align}
-
-The evaluation of the limit in the rightmost equation of Eq.~\eqref{eq:c} is evaluated as follows: 
-note that
-\begin{equation}
-\label{eq:int_limit}
-\lim_{N\rightarrow\infty} P^{\log_{P}(1-\gamma)} = 1-\gamma_{\infty},
-\end{equation}
-which is only possible if 
-\begin{equation}
-\lim_{N\rightarrow\infty} \log_{P}(1-\gamma) = 0,
-\end{equation}
-implying that $c_{\infty}$ is equal to two.
-
-Eq.~\eqref{eq:int_limit}, implies that we can only compute $c_{\infty}$ in the rightmost equation if $\lim_{N\rightarrow \infty} \gamma$ exists. 
-In Fig.~\ref{fig:c}, we plot the empirically determined value of $c$ for three different geometric layouts as a function of the number antennas in the array.
-Furthermore, in Fig.~\ref{fig:c}, the analytic value of $c$ for the regular east-west grid is plotted as a dashed red curve, while the limit of this analytic curve is plotted with a dotted black curve.
-
-Eq.~\eqref{eq:gamma}, Eq.~\eqref{eq:c} and Fig.~\ref{fig:sparsity} indicate that the computational complexity, under the assumption that $\lim_{N\rightarrow \infty} \gamma$ exists, is asymptotically bounded by $O(P^2)$.
-However, the computational complexity converges very slowly to its asymptotic value, and in general is equal to $O(P^{c})$, where $c \sim 1.7$ if a Hexagonal geometric layout containing less than 200 antennas is employed.
-
-\begin{figure*}
-\centering
-\subfigure[$\gamma$]{\includegraphics[width=0.47\textwidth]{./sparsity.pdf}\label{fig:gamma}}
-% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
-\subfigure[$c$]{\includegraphics[width=0.46\textwidth]{./comp_order.pdf}\label{fig:c}}
-\caption{In the left panel we plot the sparsity ratio $\gamma$ (the number of zero entries relative to the total number of entries in a matrix) of the modified Hessian $\bmH$ as a function of the number
-of antennas $N$ in the array for different geometries, namely an hexagonal grid (blue circles), a square grid (green crosses) and a regular east-west grid (red circles).
-For the regular east-west grid we are able to construct an analytical expression for $\gamma$ (the dashed red line).
-Taking the limit
-of this analytic expression produces the black dotted line. In the right panel we plot the order of the computational cost $c$ for inverting $\bmH$ as a function of the number 
-of antennas in the array for different geometries. We use the same color scheme in the right and left panels. The red dashed line 
-in the right panel is associated with the theoretical curve of $c$ we computed for the east-west regular grid. The limit of the red-dashed curve is two, which we 
-have plotted with a black dotted line.
-\label{fig:sparsity}} 
-\end{figure*}
-
-\subsection{Comparison}
-\label{sec:comparison}
-We are now in a position to make a final comparison between the ADI and PCG methods.
-Inverting $\bH$ by assuming it is a diagonal matrix (i.e.
-using the ADI method) is very cheap, while inverting $\bH$ using PCG is more expensive.
-On the other side of the coin,  one inversion with PCG is much more
-accurate than using the approximate inverse which is associated with the ADI method.
-The question is now, if the inverse using the PCG method is much more accurate will the LM implementation using it require much less LM-iterations than its ADI counterpart.
-Moreover, will the number
-of LM iterations PCG requires be so few that it manages to beat the ADI method after all.
-To answer this we first need to compute the number of iterations, $k$, the LM implementation associated with the ADI method can make in the
-same time it takes the LM implementation associated with the PCG method to perform a single iteration. We can compute this with
-\begin{align}
-\label{eq:k}
- kP &\approx P^{1.7}, & k &\approx P^{0.7}.
-\end{align}
-Next we need to measure the number of LM iterations we save by using the PCG method and compare it to Eq.\eqref{eq:k}. We make this comparison 
-in Fig.~\ref{fig:out_diff}. Fig.~\ref{fig:out_diff} tells us that the ADI method will outperform the PCG method as the number of LM iterations we save by taking the 
-full inverse is not enough to exceed $k$. In this comparison we have only taken into account the cost of inverting $\bH$. The PCG method might, however, still 
-turn out to be quite useful in practice as it would be more straightforward to apply it to other calibration use cases.
-
-\begin{figure*}
-\centering
-\subfigure[Number of LM iterations required by ADI and PCG.]{\includegraphics[width=0.47\textwidth]{./outerloop.pdf}\label{fig:outerloop}}
-% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
-\subfigure[Difference between ADI and PCG.]{\includegraphics[width=0.47\textwidth]{./diff.pdf}\label{fig:diff}}
-\caption{The graphs presented here were generated using the second experimental setup in Table~\ref{tab:ch_parm}. In the left panel we have the number of LM iterations 
-that is required for the LM algorithm using the ADI (green and blue curve) and the PCG (magenta and red curve) methods to converge to a parameter error tolerance of 1e-6 whilst using different 
-SNR values as a function of the number of antennas $N$ in the array.
-In the right panel we have the average amount of LM iterations (difference between the ADI and PCG curves in the left panel) we save (green and blue curve) by computing the full-inverse with the PCG method whilst using different SNR values.
-The black curve is associated with $k$, the number of iterations the LM implementation using the ADI method can perform in the same time it takes the LM implementation using 
-the PCG method to perform one iteration.
-\label{fig:out_diff}} 
-\end{figure*}
-
-\section{Conclusion}
-In Section~\ref{sec:main_sec} we showed that instead of using the traditional phase-and-amplitude approach to perform redundant calibration we can actually use a purely complex 
-framework. In this framework, the unknown variables and their conjugates are treated as independent variables that have to be estimated. 
-We then derived the GN and LM update steps using this complex framework (see Section~\ref{sec:complex_GN_LM}).
-In this paper we focused more on the LM algorithm than on the GN algorithm.
-The LM algorithm exhibits better convergence properties than the
-GN algorithm. This has the immediate consequence that the amount of algorithmic layers that are needed to perform redundant calibration are reduced.
-
-The most expensive operation associated with the LM algorithm (as well as the GN algorithm) is the inversion of $\bH$.
-We compared the computational complexity of two previously proposed algorithms to efficiently invert $\bH$ for the redundant calibration use case in Section~\ref{sec:results}, namely 
-ADI \citep{Marthi2014} and PCG \citep{Liu2010}.
-We showed that the computational complexity of the ADI method is $\mathbb{O}(P)$, while the computational complexity of the PCG method is asymptotically bounded
-by $\mathbb{O}(P^2)$. Both of these algorithms perform much better than stock standard inversion, which has a computational complexity of $\mathbb{O}(P^3)$.
-Although the ADI method does execute faster than the PCG method it is far less robust as it would be less straightforward to apply to other calibration uses cases. This is due to the fact that,
-the diagonal approximation we make with ADI might not be accurate enough in general.
-
-Lastly, we showed how \citet{Marthi2014}, \citet{Salvini2014} and \citet{Smirnov2015} relate to each other (see Section~\ref{sec:red_c}).
-The ADI method associated with skymodel-based calibration (i.e.
-\textsc{StEfCal}) is presented
-in \citet{Salvini2014}. \citet{Smirnov2015} showed that it is possible to re-derive this skymodel-based ADI method through complex optimization. In particular, they showed that if 
-we approximate the $\bH$ associated with the LM algorithm with its diagonal then we can rederive the ADI update rules presented in \citet{Salvini2014}. 
-\citet{Marthi2014} actually proposes to use the ADI method to perform redundant calibration.
-This fact is not immediately clear when superficially inspecting \citet{Marthi2014}.
-The redundant calibration algorithm presented in \citet{Marthi2014} was derived by taking the partial derivatives of
-the objective function towards the unknown parameters, setting the intermediate results to zero and then solving for the unknowns. In Section~\ref{sec:red_c}, we showed that 
-this algorithm is in actual fact an extension of the skymodel-based ADI method presented in \citet{Salvini2014} to the redundant calibration use case (i.e.
-we showed that this algorithm is in actual fact redundant \textsc{StEfCal}).
-We also re-derived the aforementioned redundant calibration algorithm by applying the diagonal approximation approach presented in \citet{Smirnov2015}, which serves as yet another additional confirmation
-that this algorithm is in fact redundant \textsc{StEfCal} (i.e. that it is some kind of ADI method).
-
-% fiducial initial guess 
-% 
-% The following partial derivatives are now easily computed
-% 
-% \begin{eqnarray}
-% \frac{c_{ij}}{\partial \eta_p} &=& y_{i-j} e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}\\ 
-% \frac{c_{ij}}{\partial \varphi_p} &=&  -i y_{i-j} e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}\\
-% \frac{c_{ij}}{\partial \eta_q} &=& y_{i-j} e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}\\ 
-% \frac{c_{ij}}{\partial \varphi_q} &=&  i y_{i-j} e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}\\
-% \frac{y_{i-j}}{\partial \varphi_q} &=&  e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}
-% \end{eqnarray}
-% 
-% 
-% The Wirtinger derivative is used in the last equation.
-% 
-% \begin{eqnarray}
-% c_{ij} &\approx& c_{ij}^0 + \Delta \eta_p(y_{i-j}^0 e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0}) + \Delta \eta_q(y_{i-j}^0 e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0})\\
-% && -i\Delta\varphi_p (y_{i-j}^0e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0}) +i\Delta\varphi_q (y_{i-j}^0e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0})\\
-% && y_{i-j}^1(y_{i-j}^0e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0})\\
-% &\approx&  c_{ij}^0 + e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_p+ \Delta \eta_q - i\Delta\varphi_p + i\Delta\varphi_p))
-% \end{eqnarray}
-% 
-% \begin{equation}
-% \label{eq:residual}
-% r_{ij} = \delta_{ij} = c_{ij}-c_{ij}^0 = e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_p+ \Delta \eta_q - i\Delta\varphi_p + i\Delta\varphi_p)) 
-% \end{equation}
-% 
-% Eq.~\ref{eq:residual} allows us to construct the following linear system:
-% 
-% \begin{equation}
-% \boldsymbol{J}[\boldsymbol{\Delta \eta},\boldsymbol{\Delta \varphi},\boldsymbol{\Delta y}]^T = \boldsymbol{r}, 
-% \end{equation}
-% 
-% where $\boldsymbol{J}$ is equal to 
-% \begin{equation}
-% \boldsymbol{J} = \bigg [\overbrace{\frac{c_{pq}}{\partial \eta_p}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial \varphi_p}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial y_{t}}}^{t=1\cdots r_s} \bigg ] \bigg \} [pq] = 1\cdots N_{\textrm{b}} (p<q)
-% \end{equation}
-% or
-% \begin{equation}
-% \boldsymbol{J} = \bigg [\frac{c_{pq}}{\partial \boldsymbol{\eta}},~\frac{c_{pq}}{\partial \boldsymbol{\varphi}},~\frac{c_{pq}}{\partial \boldsymbol{y}} \bigg ] \bigg \} [pq] = 1\cdots N_{\textrm{b}} (p<q).
-% \end{equation}
-% 
-% The last column is again a Wirtinger derivative.
-% 
-% Which means we can obtain the least-squares estimate as follows:
-% 
-% \begin{equation}
-% [\boldsymbol{\Delta \eta},\boldsymbol{\Delta \varphi},\boldsymbol{\Delta y}]^T = [\boldsymbol{J}^H\boldsymbol{J}]^{-1}\boldsymbol{J}^H\boldsymbol{r}.
-% \end{equation}
-
-\bibliographystyle{mn2e}
-\bibliography{paper}
-
-\appendix
-
-\section{\textsc{LOGCAL}}
-\label{sec:logcal}
-We can rewrite Eq.~\eqref{eq:vis_red} to \citep{Liu2010}
-\begin{eqnarray}
-\ln |d_{pq}| &=& \ln |g_p| + \ln |g_q| + \ln |y_{\phi_{pq}}| + s_{pq} \label{eq:logcal_amp}\\
-\angle d_{pq} &=& \angle g_p - \angle g_q + \angle \phi_{pq} + t_{pq} \label{eq:logcal_phase}
-\end{eqnarray}
-where
-\begin{align}
-s_{pq} &= \mathscr{R} \left \{1 + \frac{n_{pq}}{g_p\conj{g_q}y_{\phi_{pq}}} \right \}, & t_{pq} &= \mathscr{I}\left \{1 + \frac{n_{pq}}{g_p\conj{g_q}y_{\phi_{pq}}} \right \}.
-\end{align}
-Moreover, $\mathscr{R}\{*\}$ and $\mathscr{I}\{*\}$ denote the real and imaginary operators.
-
-With the aid of Eq.~\eqref{eq:logcal_amp} we can construct the following linear system:
-\begin{equation}
-\label{eq:log_system}
-\bdelta = \bJ\bzeta + \bs, 
-\end{equation}
-where
-\begin{align}
-\left [ \bdelta \right ]_{\alpha_{pq}} &= \ln |d_{pq}|, & \left [ \bs \right ]_{\alpha_{pq}} = s_{pq} 
-\end{align}
-and
-\begin{equation}
-\left [ \bzeta \right ]_{j} = \begin{cases} \ln |g_j| & \textrm{if}~j\leq N \\ \ln |y_{j-N}| & \textrm{otherwise} \end{cases}. 
-\end{equation}
-Furthermore,
-\begin{equation}
-\bJ = 
-\begin{bmatrix}
-\bN & \bM
-\end{bmatrix}
-\end{equation}
-with 
-\begin{equation}
-[\bN]_{\alpha_{pq},j} = \begin{cases}
-       1 & \textrm{if}~(p=j)~\textrm{or}~(q=j)\\
-       0 & \textrm{otherwise}
-      \end{cases}
-\end{equation}
-and
-\begin{equation}
-[\bM]_{\alpha_{pq},j} = \begin{cases}
-       1 & \textrm{if}~\phi_{pq}=j\\
-       0 & \textrm{otherwise}
-      \end{cases}
-\end{equation}
-
-The least squares solution to Eq.\eqref{eq:log_system}, under the incorrect assumption that $s_{pq}$ is normally distributed, is given by
-\begin{equation}
-\bzeta = (\bA^T\bA)^{-1} \bA^T\bdelta. 
-\end{equation}
-How to correctly deal with the fact that $s_{pq}$ is not normally distributed is discussed in \citep{Liu2010}. A similar linear system can be constructed for Eq.~\eqref{eq:logcal_phase}. 
-From the least squares solutions to the linear systems associated with Eq.~\eqref{eq:logcal_amp} and Eq.~\eqref{eq:logcal_phase} we can compute the antenna gains as well as the true visibilities.
-
-\section{\textsc{lincal}}
-\label{sec:lincal}
-%Assume that $\eta_p,\eta_q,\widetilde{\eta}_{\phi_{pq}},\varphi_p,\varphi_q$ and $\widetilde{\varphi}_{\phi_{pq}}$ are real valued variables.
-We can rewrite Eq.~\eqref{eq:vis_red_vec} to \citep{Liu2010}
-\begin{equation}
-\label{eq:lincal_vec}
-\bd = \bv(\bvarrho) + \bn, 
-\end{equation}
-where 
-\begin{align}
-\left [ \bv(\bvarrho) \right ]_{\alpha_{pq}} &= e^{\eta_p - i \varphi_p} e^{\widetilde{\eta}_{\phi_{pq}}- i \widetilde{\varphi}_{\phi_{pq}}} e^{\eta_q + i \varphi_q}\\
-&= v_{pq},
-\end{align}
-and
-\begin{equation}
-\bvarrho = 
-\begin{bmatrix}
-\bmath{\eta} \\
-\bvarphi \\
-\widetilde{\bmath{\eta}} \\
-\widetilde{\bvarphi}
-\end{bmatrix},\qquad
-\begin{aligned}
-\bmath{\eta} &= [\eta_1, \eta_2,\cdots \eta_N]^T\\
- \bmath{\varphi} &= [\varphi_1, \varphi_2,\cdots \varphi_N]^T\\
- \widetilde{\bmath{\eta}} &= [\widetilde{\eta}_1, \widetilde{\eta}_2,\cdots \widetilde{\eta}_L]^T\\
- \widetilde{\bmath{\varphi}} &= [\widetilde{\varphi}_1, \widetilde{\varphi}_2,\cdots \widetilde{\varphi}_L]^T
-\end{aligned}.
-\label{eq:help}
-\end{equation}
-Note that the vectors $\bmath{\eta},~\bvarphi,~\widetilde{\bmath{\eta}}$ and $\widetilde{\bvarphi}$ are real valued.
-
-The first order Taylor expansion of Eq.~\eqref{eq:lincal_vec} around the fiducial point $\bvarrho_0$ is equal to
-\begin{equation}
-\label{eq:d_lin}
-\bd = \bv_0 + \bmJ_0 \Delta \bvarrho  + \bn.
-\end{equation}
-If we calculate the residual, dropping the reference to a specific fiducial point, we obtain: 
-\begin{equation}
-\label{eq:r_lin}
-\br = \bd - \bv = \bmJ \Delta \bvarrho + \bn.
-\end{equation}
-Moreover,
-\begin{equation}
-\label{eq:jac_upper_lincal}
-\bmJ = \begin{bmatrix}
-       \bN & \bM 
-       \end{bmatrix},
-\end{equation}
-where
-\begin{align}
-\bN &= \begin{bmatrix} \bO & \bP \end{bmatrix}, & \bM &= \begin{bmatrix} \bQ & \bR \end{bmatrix}, 
-\end{align}
-\begin{equation}
- [\bO]_{\alpha_{pq},j} =  \frac{\partial v_{pq}}{\partial \eta_j} = \begin{cases} 
-    v_{pq} &\textrm{if}~(p=j)~\textrm{or}~(q=j)\\
-    0 & \textrm{otherwise}
-   \end{cases},
-\end{equation}
-\begin{equation}
- [\bP]_{\alpha_{pq},j} =  \frac{\partial v_{pq}}{\partial \varphi_j} = \begin{cases}
-                                                                        i v_{pq} &\textrm{if}~ p=j\\
-                                                                        -i v_{pq} &\textrm{if}~ q=j \\
-                                                                        0 & \textrm{otherwise} 
-                                                                       \end{cases}
-\end{equation}
-\begin{equation}
- [\bQ]_{\alpha_{pq},j} =  \frac{\partial v_{pq}}{\partial \widetilde{\eta}_j} = \begin{cases} 
-    v_{pq} &\textrm{if}~\phi_{pq}=j\\
-    0 & \textrm{otherwise}
-   \end{cases},
-\end{equation}
-\begin{equation}
- [\bR]_{\alpha_{pq},j} =  \frac{\partial v_{pq}}{\partial \widetilde{\varphi}_j} = \begin{cases} 
-    i v_{pq} &\textrm{if}~\phi_{pq}=j\\
-    0 & \textrm{otherwise}
-   \end{cases}.
-\end{equation}
-Augmenting the residual by taking into account its conjugate results in 
-\begin{equation}
-\breve{\br} = \bJ \Delta \bvarrho + \bn, 
-\end{equation}
-where $\breve{\br} = \begin{bmatrix}\br^T & \conj{\br}^T \end{bmatrix}^T$ and 
-\begin{equation}
-\label{eq:Jac_lin}
-\bJ = 
-\begin{bmatrix}
- \bmJ\\  
- \conj{\bmJ}
-\end{bmatrix}.
-\end{equation}
-The solution to 
-\begin{equation}
- \min_{\Delta \bvarrho} \| \breve{\br} - \bJ \Delta \bvarrho\|_F^2,
-\end{equation}
-is equal to 
-\begin{equation}
-\label{eq:LINCAL_update}
-\Delta \bvarrho = (\bJ^H\bJ)^{-1} \bJ^H\breve{\br}. 
-\end{equation}
-Eq.~\eqref{eq:LINCAL_update} is used to iteratively determine a new estimate of the parameter vector $\bvarrho$ as follows
-\begin{equation}
-\label{eq:new_rho}
-\bvarrho_{k+1} = \bvarrho_k + \Delta \bvarrho_k. 
-\end{equation}
-Note that we do not explicitly denote the $k$ index in Eq.~\eqref{eq:LINCAL_update}. We update our parameter vector until convergence is reached.
-
-\subsection{Conjugate Gradient Method}
-\label{sec:conj_grad}
-The CG method was independently discovered by \citet{Hestenes1973} and \citet{Stiefel1952}. They later published a joint paper, which is now considered as the seminal
-reference on CG \citep{Hestenes1952}. The iterative CG method is used to solve a particular class of linear system. The CG method is generally used to solve
-\begin{equation}
-\label{eq:linear_system}
-\bb = \bA\bx,
-\end{equation}
-where $\bb\in\mathbb{C}^P$ is a known complex column vector of size $P$, $\bx\in\mathbb{C}^P$ is an unknown complex column vector with the same length as $\bb$, and $\bA\in\mathbb{C}^{P\times P}$ is a square positive-definite Hermitian matrix with the appropriate dimensions.
-Moreover, using the CG method to solve Eq.~\eqref{eq:linear_system} will only be computationally advantageous if $\bA$ is also sparse (contain many zero entries).
-Using the iterative CG method to solve large sparse linear sytems was popularized by \citet{Reid1971}. A good tutorial on the CG method can be found in \citep{Shewchuk1994}.
-The iterative CG method is presented in Algorithm \ref{algo:CG}.
-
-%If $\bA$ is a square positive semi-definite Hermitian matrix, then the conjugate gradient method only converges if $\bb$ is in the column range of $\bA$ (cite Lu2016).
-
-%\begin{algorithm}
-%\caption{Conjugate Gradient Method.
-Inputs: $\bA$, $\bb$, a starting value for $\bx$, maximum number of iterations $k_{\textrm{max}}$ and an error tolerance $\epsilon<1$ Output: $\bx$ the solution to Eq.~\eqref{eq:linear_system}.
-\citep{Shewchuk1994}.}\label{algo:CG}
-%\begin{algorithmic}[1]
-%\State $k \gets 0$
-%\State $\br \gets \bb - \bA\bx$
-%\State $\bp \gets \br$
-%\State $\delta_{\textrm{new}} \gets \br^H\br$
-%\State $\delta_0 \gets \delta_{\textrm{new}}$
-%\While{$k\leq k_{\textrm{max}}$ and $\delta_{\textrm{new}} > \epsilon^2\delta_0$}
-%\State $\alpha \gets \frac{\delta_{\textrm{new}}}{\bp^H\bA\bp}$
-%\State $\bx \gets \bx + \alpha\bp$
-%\State $\br \gets \br - \alpha\bA\bp$
-%\State $\delta_{\textrm{old}} \gets \delta_{\textrm{new}}$ 
-%\State $\delta_{\textrm{new}} \gets \br^H\br$
-%\State $\beta \gets \frac{\delta_{\textrm{new}}}{\delta_{\textrm{old}}}$
-%\State $\bp \gets \br + \beta\bp$
-%\State $k \gets k + 1$
-%\EndWhile
-%\end{algorithmic}
-%\end{algorithm}
-
-\subsubsection{Computational Complexity}
-\citet{Kaniel1966} was able to determine convergence bounds for CG using Chebyshev polynomials. A much more in depth analysis of CG convergence can be found in
-\citep{Sluis1986}. As can be seen from Algorithm \ref{algo:CG}, the CG method makes use of a major outer loop (line 6 of Algorithm \ref{algo:CG}). It turns out that at worst 
-the number of major iterations that CG requires to converge\footnote{If the $\bA$-norm is used to calculate the error.} is proportional to $\sqrt{\kappa}$, where $\kappa$ is the spectral condition number of the matrix $\bA$.
-The mathematical definition of $\kappa(\bA)$ is
-\begin{equation}
-\label{eq:kappa}
-\kappa(\bA) = \frac{\lambda_{\textrm{max}}}{\lambda_{\textrm{min}}}, 
-\end{equation}
-where $\lambda_{\textrm{max}}$ and $\lambda_{\textrm{min}}$ respectively denote the largest and the smallest eigenvalue of $\bA$.
-Moreover, the most expensive operation that is conducted within each major loop of Algorithm~\ref{algo:CG} is the vector-matrix product $\bA\bp$, which has a complexity of $\mathbb{O}(m)$, where $m$ is the number of non-zero entries in $\bA$.
-The computational complexity of CG is therefore 
-\begin{equation}
-\label{eq:cg_bound}
-\mathbb{O}(\sqrt{\kappa}m). 
-\end{equation}
-It should now also be clear to the reader why the CG method is especially applicable to sparse linear systems. In sparse linear systems $m < P^2$.
-
-\subsubsection{Preconditioning}
-\label{sec:precon}
-Preconditioning is a technique which allows us to improve the condition number of a matrix. Let $\bM$ be a positive-definite Hermitian matrix.
-We can now indirectly solve for $\bx$ by using 
-\begin{equation}
-\label{eq:prec}
-\bM^{-1}\bA\bx = \bM^{-1}\bb.
-\end{equation}
-If $\bM$ is a good preconditioner then $\kappa(\bM^{-1}\bA) \ll \kappa(\bA)$, i.e. a good preconditioner lowers the condition number of a matrix which in turn increases the convergence speed 
-of CG (see Eq.~\eqref{eq:cg_bound}).
-In this paper, we denote matrix inversion with $(*)^{-1}$.
-The best possible choice of $\bM$ would therefore be $\bA$, which would lead to $\kappa$ becoming unity (as the largest and smallest eigenvalue of the identity matrix are equal to one).
-If $\kappa$ is one then Eq.~\eqref{eq:cg_bound} simplifies and becomes $\mathbb{O}(m)$.
-However, if we knew the inverse of $\bA$ then obtaining $\bx$ would be trivial and there would be no need for the CG method in the first place.
-In short, a good preconditioner $\bM$ must approximate $\bA$ fairly accurately. However, if it is not much less expensive to 
-compute the inverse of $\bM$ than that of $\bA$ we actually gain nothing (see Eq.~\eqref{eq:prec}).
-
-If the diagonal entries of $\bA$ are much more significant than its off-diagonal entries then the 
-Jacobian preconditioner is a good choice of $\bM$. The Jacobian preconditioner is computed as follows:
-\begin{equation}
-\bM = \bA\odot\bI, 
-\end{equation}
-where $\bI$ is the identity matrix and $\odot$ denotes the Hadamard product. Since the Jacobian preconditioner is a diagonal matrix its inverse can be obtained trivially.
-
-\subsubsection{Normal Equation}
-\label{sec:normal}
-Lets assume that instead of Eq.~\eqref{eq:linear_system} we have a more general linear system, i.e.
-\begin{equation}
- \bb = \bB\bx,
-\end{equation}
-where $\bb\in\mathbb{C}^K$, $\bx\in\mathbb{C}^P$  and $\bB\in\mathbb{C}^{K \times P}$, with $K > P$. We can use the normal equation 
-\begin{equation}
-\label{eq:normal_equation}
-\bB^H\bB\bx = \bB^H\bb, 
-\end{equation}
-to estimate the $\bx$ that minimizes
-\begin{equation}
-\label{eq:norm_ls}
-\|\bb-\bB\bx\|_F^2. 
-\end{equation}
-In this paper, we denote the Hermitian transpose with $(*)^H$ and the Frobenius norm with $\|*\|_F$.
-
-There are a few important observations that we can make while inspecting Eq.~\eqref{eq:normal_equation}:
-\begin{enumerate}
-\item $\bB^H\bB$ is a square positive semi-definite Hermitian matrix by construction.
-\item $\bB^H\bb$ is in the column range of $\bB^H\bB$; this follows trivially from 
-\begin{equation}
-\bB^H\bb \in \mathcal{R}(\bB^H) \implies \bB^H\bb \in \mathcal{R}(\bB^H\bB).   
-\end{equation}
-\end{enumerate}
-In this paper, we denote the column range of a matrix $\bA$ with $\mathcal{R}(\bA)$.
-
-Recall that we mentioned that in general we can only solve Eq.~\eqref{eq:linear_system} with CG if $\bA$ is a square positive-definite Hermitian matrix.
-The CG method actually has an even broader applicability: it can be used to solve Eq.~\eqref{eq:linear_system} even if $\bA$ is a Hermitian positive semi-definite
-square matrix as long as $\bb$ is in the column range of $\bA$ \citep{Lu2015}.
-In light of \citet{Lu2015}, the two observations in the numbered list we presented in the previous paragraph imply that Eq.~\eqref{eq:normal_equation}
-can be solved by employing the CG method. The solution returned by CG is, however, only unique if $\bB^H\bB$ is positive-definite (invertible). The solution
-obtained by CG, whether it is unique or not, does minimize Eq.~\eqref{eq:norm_ls} in a least squares sense.
-
-Note that if $\bA$ is positive semi-definite, its smallest eigenvalue is zero. This implies that Eq.~\eqref{eq:kappa} is not defined. \citet{Lu2015} show that when $\bA$
-is a square positive semi-definite Hermitian matrix we do not use the spectral condition number to bound the convergence complexity of CG, but the general spectral condition
-number instead. The general spectral condition number is very similar to the spectral condition number; the only difference being that in Eq.~\eqref{eq:kappa} we replace  
-$\lambda_{\textrm{min}}$ with $\lambda_{\textrm{min}}^+$, which denotes the smallest positive eigenvalue.
-
 \section{Redundant \textsc{StEfCal} and the ADI method}

### 2017-05-05 paper.tex
implemented most of the fixes Gianni, suggested

@@
 This framework reduces the
-
+\textbf{[GIANNI TO REWRITE AND FIX ABSTRACT]}
 has driven advancements in the calibration of radio interferometers and 
@@
 We make use of the parameter-and-conjugate framework in this paper, rather than the more traditional amplitude-and-phase approach as it makes it trivial to relate the alternating direction implicit method and the Levenberg-Marquardt algorithm with one another.
+\textbf{[Could we add JS Kenyon and SJ Wijnholds to author list?]}
 \end{abstract}
@@
 
-We start the paper by presenting some useful preliminaries in Section~\ref{sec:preliminaries}. In Section~\ref{sec:p_statement}, we formulate redundant calibration 
-as a least squares problem, i.e. Eq~\eqref{eq:least_squares_red}. In Section~\ref{sec:main_sec} we use the complex framework proposed by \citet{Smirnov2015} to derive the GN and LM
-update rules (see Eq.~\eqref{eq:GN_update} to Eq.~\eqref{eq:update}), which we can use to solve Eq~\eqref{eq:least_squares_red}.
+In Section~\ref{sec:sky_wirtinger}, we review the skymodel-based complex-and-conjugate calibration approach proposed by \citet{Smirnov2015}.
+In Section~\ref{sec:red_wirtinger}, we show that this same approach can also be used to perform redundant calibration.
 In Section~\ref{sec:results}, we compare the computational complexity of two recently proposed redundant calibration speedup algorithms with one another.
@@
 
-\subsection{Real and imaginary approach}
+\subsection{Real and Imaginary Approach}
 \label{sec:ri}
@@
 
-\subsection{Complex Approach}
+\subsection{Wirtinger Approach}
 \label{sec:w}
@@
 where $\breve{\bz} = [\bz^T,\conj{\bz}^T]^T$. Eq.~\eqref{eq:least_squares_complex} is the complex conjugate augmented counterpart of Eq.~\eqref{eq:least_squares_red}.
-The fact that $\bv$ is now analytic in its argument has the immediate consequence: it becomes possible to apply complex GN
-and complex LM to Eq.~\eqref{eq:least_squares_complex}.
+%The fact that $\bv$ is now analytic in its argument has the immediate consequence: it becomes possible to apply complex GN
+%and complex LM to Eq.~\eqref{eq:least_squares_complex}.
 
@@
 \end{equation}
-
-As mentioned in Section We use Eq.~\eqref{eq:GN_update} or Eq.~\eqref{eq:LM_update} to iteratively update our 
+As was the case in Section~\ref{sec:sky_wirtinger}, we can now use Eq.~\eqref{eq:GN_update} or Eq.~\eqref{eq:LM_update} to iteratively update our 
 parameter vector as follows:
@@
 \end{equation}
-We update our parameter vector by using Eq.~\eqref{eq:update} until convergence is reached. To improve readability we do not explicitly denote the iteration subscript $k$ in 
-Eq.~\eqref{eq:GN_update} and Eq.~\eqref{eq:LM_update}.
+The analytic expressions for $\bJ$, $\bH$ and $\bJ^H\breve{\br}$ can be found in Appendix~\ref{sec:analytic}.
 
@@
 \label{sec:results}
-In Section~\ref{sec:main_sec} we discussed the complex GN and LM algorithms and how they relate to redundant calibration. If we inspect Eq.~\eqref{eq:GN_update} and 
+In Section~\ref{sec:red_wirtinger} we discussed the complex GN and LM algorithms and how they relate to redundant calibration. If we inspect Eq.~\eqref{eq:GN_update} and 
 Eq.~\eqref{eq:LM_update} we see that the most expensive operation used by the GN and LM algorithms is the inversion of the Hessian, $\bH$. The most logical 
@@
 
-\subsubsection{Signal-to-noise Ratio}
-\label{sec:snr}
-The following SNR definition is used in this paper  
+Furthermore, the following SNR definition is used in this paper  
 \begin{equation}
@@
 expression of $\bJ^H\breve{\br}$, the only difference being that in Eq.~\eqref{eq:ab} the letter $r$ would be replaced by a $d$. If we substitute the analytic expression
-of $\bJ^H\breve{\bd}$ and $\widetilde{\bH}^{-1}$ (which is easily computed as the matrix is diagonal) into Eq.~\eqref{eq:alpha} we obtain the following two update rules:
+of $\bJ^H\breve{\bd}$ and $\widetilde{\bH}^{-1}$ (which can easily be constructed using Appendix~\ref{sec:analytic}) into Eq.~\eqref{eq:alpha} we obtain the following two update rules:
 \begin{equation}
@@
 In this section we investigate the computational complexity of the PCG method when applied to the redundant calibration problem. 
-As we mentioned in Section~\ref{sec:conj_grad} there are two factors which influence the computational complexity of the PCG method, namely the spectral condition number and the sparsity of the matrix we wish to invert (see Eq.~\eqref{eq:cg_bound}).
+We briefly review the PCG method in Appendix~\ref{sec:conj_grad}. If the reader is not that familiar with the PCG method we encourage 
+the reader to briefly review Appendix~\ref{sec:conj_grad} before continuing with the paper. The PCG method can be used to efficiently invert Hermitian sparse 
+matrices, which is exactly the kind of matrix $\bH$ is (see Fig.~\ref{fig:hessian} and Eq.~\ref{eq:red_H}).
+As we mention in Appendix~\ref{sec:conj_grad} there are two factors which influence the computational complexity of the PCG method, namely the spectral condition number $\kappa$ and the sparsity of the matrix we wish to invert (see Eq.~\eqref{eq:cg_bound}).
+
 We used a $\lambda$ value of two, implying we used Eq.~\eqref{eq:LM_update} instead of Eq.~\eqref{eq:GN_update} throughout this section.
-We discuss the spectral condition number and the sparsity of the modified Hessian $\bmH$, which is a non-singular matrix, in Section~\ref{sec:scn} and Section~\ref{sec:sparsity} respectively.
+We investigate the spectral condition number and the sparsity of the modified Hessian $\bmH$, which is a non-singular matrix, in Section~\ref{sec:scn} and Section~\ref{sec:sparsity} respectively.
 It is interesting to note, that the Hessian $\bH$ itself is singular which means that if we are to use Eq.~\eqref{eq:GN_update} instead we would need to be able to invert a singular matrix.
-In Section~\ref{sec:normal},
-we showed that the PCG method would in fact be able to handle such a scenario.
+In Appendix~\ref{sec:normal},
+we mention that the PCG method would in fact also be able to pseudo-invert $\bH$ directly.
 
@@
 \label{sec:scn}
-In Section~\ref{sec:precon} we mentioned that the convergence properties of CG can be dramatically improved through preconditioning.
-Inspecting Fig.~\ref{fig:hessian} reveals that the Jacobian preconditioner
-is a good candidate for the redundant use case. In Fig.~\ref{fig:kappa} we see that this choice of preconditioner performs very well, the value of the spectral condition number $\kappa$
+In Appendix~\ref{sec:precon} we mentioned that the convergence properties of CG can be dramatically improved through preconditioning.
+Inspecting Fig.~\ref{fig:hessian}, reveals that the Jacobian preconditioner
+is a good candidate for the redundant use case. In Fig.~\ref{fig:kappa}, we see that this choice of preconditioner performs very well, the value of the spectral condition number $\kappa$
 becomes constant and independent of problem size, i.e. the number of antennas in the array. Preconditioning therefore effectively eliminates $\kappa$ from Eq.~\eqref{eq:cg_bound}. This result
@@
 
-We can now use the sparsity factor to compute the order of the computational complexity of inverting $\bmH$ (assuming the effect of $\kappa$ is negliable) with
+We can now use the sparsity factor to compute the order of the computational complexity of inverting $\bmH$ (assuming the effect of $\kappa$ is negligible) with
 \begin{equation}
@@
 
-The evaluation of the limit in the rightmost equation of Eq.~\eqref{eq:c} is evaluated as follows: 
-note that
+The evaluation of the limit in the rightmost equation of Eq.~\eqref{eq:c} is evaluated as follows: note that
 \begin{equation}
@@
 
-\subsection{Conjugate Gradient Method}
+\section{Conjugate Gradient Method}
 \label{sec:conj_grad}
@@
 
-\subsubsection{Computational Complexity}
+\subsection{Computational Complexity}
 \citet{Kaniel1966} was able to determine convergence bounds for CG using Chebyshev polynomials. A much more in depth analysis of CG convergence can be found in
@@
 
-\subsubsection{Preconditioning}
+\subsection{Preconditioning}
 \label{sec:precon}
@@
 
-\subsubsection{Normal Equation}
+\subsection{Normal Equation}
 \label{sec:normal}
@@
 
-\section{Redundant Wirtinger Calibration - $\bJ$ and $\bH$}
+\section{Redundant Wirtinger Calibration - $\bJ$, $\bH$ and $\bJ^H\breve{\br}$}
 \label{sec:analytic}

### 2017-10-11 paperG2.tex
newest draft from Gianni

--- 2017-05-05 paper.tex
@@
 
-\title[Unknown]{Redundant calibration as a complex optimization problem}
-\author[T.L.~Grobler et al.]{T.L.~Grobler$^{1}$\thanks{E-mail: email@address (AVR)}, G.~Bernardi$^{1,2}$, Z.S.~Ali$^{3}$?, C.L.~Carilli$^{4,5}$?, J.S.~Dillon$^{3}$?, A.~Liu$^{3}$?, \newauthor
-A.R.~Parsons$^{3,6}$? and O.M.~Smirnov$^{1,2}$?\\
+%\title[Complex redundant calibration]{Redundant calibration as a complex optimization problem}
+\title[Redundant interferometric calibration as a complex optimization problem]{Redundant calibration as a complex optimization problem}
+%\author[T.L.~Grobler et al.]{T.L.~Grobler$^{1}$\thanks{E-mail: email@address (AVR)}, G.~Bernardi$^{1,2}$, Z.S.~Ali$^{3}$?, C.L.~Carilli$^{4,5}$?, J.S.~Dillon$^{3}$?, A.~Liu$^{3}$?, \newauthor
+%A.R.~Parsons$^{3,6}$? and O.M.~Smirnov$^{1,2}$?\\
+%$^{1}$Department of Physics and Electronics, Rhodes University, PO Box 94, Grahamstown, 6140, South Africa\\
+%$^{2}$SKA SA, 3rd Floor, The Park, Park Road, Pinelands, 7405, South Africa\\
+%$^{3}$Dept. of Astronomy, University of California, Berkeley, CA 94720, USA\\
+%$^{4}$National Radio Astronomy Obs., Socorro NM\\
+%$^{5}$Cavendish Lab., Cambridge UK\\
+%$^{6}$Radio Astronomy Lab., U. California, Berkeley CA 94720, USA}
+\author[T.L.~Grobler et al.]{T.L.~Grobler$^{1}$\thanks{E-mail: email@address (AVR)}, G.~Bernardi$^{1,2}$, C.L.~Carilli$^{3,4}$, J.S.~Kenyon$^{1}$, A.R.~Parsons$^{5,6}$ \newauthor and O.M.~Smirnov$^{1,2}$\\
 $^{1}$Department of Physics and Electronics, Rhodes University, PO Box 94, Grahamstown, 6140, South Africa\\
 $^{2}$SKA SA, 3rd Floor, The Park, Park Road, Pinelands, 7405, South Africa\\
-$^{3}$Dept. of Astronomy, University of California, Berkeley, CA 94720, USA\\
-$^{4}$National Radio Astronomy Obs., Socorro NM\\
-$^{5}$Cavendish Lab., Cambridge UK\\
+$^{3}$National Radio Astronomy Obs., Socorro NM\\
+$^{4}$Cavendish Lab., Cambridge UK\\
+$^{5}$Dept. of Astronomy, University of California, Berkeley, CA 94720, USA\\
 $^{6}$Radio Astronomy Lab., U. California, Berkeley CA 94720, USA}
@@
 We show that redundant calibration can be formulated in the domain of complex functions of a complex variable.
-This framework reduces the
+This framework reduces the uncertaninties
 \textbf{[GIANNI TO REWRITE AND FIX ABSTRACT]}
@@
 We make use of the parameter-and-conjugate framework in this paper, rather than the more traditional amplitude-and-phase approach as it makes it trivial to relate the alternating direction implicit method and the Levenberg-Marquardt algorithm with one another.
-\textbf{[Could we add JS Kenyon and SJ Wijnholds to author list?]}
 \end{abstract}
@@
 \section{Introduction}
-The effects (i.e.
+The quest for the redshifted 21-cm line from the Epoch of Reionization (EoR) is a frontier of modern observational cosmology and has motivated the construction of a series of new interferometric arrays operating at low frequencies over the last decade.
+Measurements of the EoR are difficult because of the intrinsic faintness of the cosmological signal \citep[see, for instance,][for recent reviews]{Furlanetto2015,McQuinn16} and the presence of foreground emission which is a few orders of magnitude brighter than the EoR anywhere in the sky \citep[e.g.,][]{Bernardi2009,Bernardi2010,Ghosh2011,Dillon2012,Parsons2014} and that can be separated by leveraging only on the different spectral properties (citations needed!!!).
+An exquisite calibration is therefore required as  miscalibrated foreground emission can jeopardize the EoR signal \citep[e.g.,][]{barry2016,grobler2014,ewall-wice2016}.
+These reasons motivated the design and deploment of arrays in redundant configurations like, MIT-EoR  (reference!) the Precision Array to Probe the Epoch of Reionization \citep[PAPER,][]{Parsons2014,Jacobs2015,Ali2015} and the Hydrogen Epoch of Reionization Array \citep[HERA,][]{deboer2017}.
+
+Redundant configurations have historically been abandoned because of their poor imaging performances as they measure only a limited number of spatial scales and, therefore, do not adequately sample the sky brightness distribution.
+For this very reason, however, redundant configurations provide a sensitivity boost on a selected sample of spatial scales that can be tuned to be the highest signal--to--noise ratio modes for EoR measurements \citep[][]{Parsons2012,Dillon2015}.
+
+The second advantage of redundant array configurations is that they, by definition, they measure the same spatial scales {\bf (GB: perhaps worth having the Van cittert-Zernike theorem here?)} regardless of what the sky brightness distribution looks like.
+This means that the corruptions that are introduces by the instrument can be calibrated - up to some extent - without knowing anything about the sky brightness distribution \citep[][]{Noordam1982,Wieringa1992}.
+This property is particularly desirable for EoR observations as it mitigates errors introduced by an incomplete knowledge of the sky brightness distribution.
+
+In this paper we extend the stefcal calibration algorithm to the redundant use case using the complex optimization formalism recently introduced by \cite{Smirnov2015}. 
+We also perform a proper computational analysis of the recently proposed (preconditioned conjugate gradient) PCG method when it is applied to the redundant calibration problem.
+We also compare these two approaches.
+
+present a new algorithm to calibrate redundant arrays based on the complex optimization formalism recently introduced by \cite{Smirnov2015}.
+With respect to current algorithms, it is more robust to initial conditions, it is comparatively fast and {\bf ?????...}.
+We also present an analysis of its computational complexity showing that the LM algorithm shows better convergence properties than the GN currently used \cite{Zheng14}.
+
+The paper is organized as follows: we review the complex calibration formalism by \citet{Smirnov2015,Kurien2016} in Section~\ref{sec:sky_wirtinger}, we extend it to the redundant case in Section~\ref{sec:red_wirtinger}, we test the algorithm performances in Section~\ref{sec:results} and we conclude in Section\ref{sec:conclusions}.
+
+%
+%
+%The effects (i.e.
 the ionosphere, antenna gains, etc.) a celestial signal experiences along its propagation path are modelled with the radio interferometric measurement equation (RIME) \citep{ME1,RRIME1}.
-Calibration is the procedure by which we remove the errors that these propagation effects introduce into interferometric data and is normally accomplished using a 
-least-squares solver.
+Calibration is the procedure by which we remove the errors that these propagation effects introduce into interferometric data and is normally accomplished using a least-squares solver.
 In the case where we only model the antenna gains, which is the most basic form of the RIME, we can define calibration as finding the antenna gains that minimize the difference between our observed and predicted visibilities.
-
-There are two recent developments in skymodel-based calibration which should be explicitly mentioned at this point, since they form the foundation this paper builds on.
-The first is the discovery of \textsc{StEfCal} (Statistically Efficient Calibration) \citep{Mitchell:MWA-cal,Salvini2014}, which
-is an alternating direction implicit (ADI) skymodel-based calibration algorithm. It works by assuming the gains are known while solving for their conjugates and then performing the reverse. 
-This procedure results in the linearization of the calibration algorithm and reduces the computational complexity of calibration from $\mathbb{O}(N^3)$ to $\mathbb{O}(N^2)$, where $N$ denotes the number of antennas in the array.
-The second development is the idea: instead of breaking the calibration problem 
-into its real and imaginary parts and then solving for the real and imaginary parts of the unknown parameters separately we can treat the antenna gains and their 
-conjugates as separate parameters to estimate \citep{Smirnov2015}. \citet{Smirnov2015} also showed that the \textsc{StEfCal} algorithm itself is merely a special case of this
-gain-and-conjugate framework.
-
-Two baselines are redundant if their baseline difference vectors have the same length and orientation, which imply that they measure the same $uv$-modes.
-When we use a redundant array layout, we dramatically
-reduce the number of unknowns. If our array is redundant enough we reduce the number of unknowns to such an extent that we are able to solve for the unknown 
-observed visibilities themselves in addition to the antenna gains. Redundant calibration is the procedure by which we estimate both the unknown observed visibilities and the antenna gains.
-
-Redundant calibration was popularized in the 1980's. \citet{Noordam1982} used redundant calibration and the WSRT (Westerbork Synthesis Radio Telescope) to produce high-dynamic range images of 3C84.
-The first implementations of redundant calibration made use of the logarithm function, because when the logarithm is applied to the scalar measurement equation, redundant 
-calibration is transformed into a linear problem \citep{Wieringa1992,Camps2003,Liu2010}. The implementation associated with this approach is commonly referred to as \textsc{logcal}.
-We briefly discuss \textsc{logcal} in Appendix~\ref{sec:logcal}.
-The next improvement in redundant calibration was to perform a first order Taylor expansion, which is equivalent to the Gauss-Newton (GN) algorithm \citep{Kurien2016}, of the measurement equation and then to solve for the
-perturbations of the model-parameters \citep{Liu2010}.
-\citet{Liu2010} perform the first order Taylor expansion of the measurement equation by splitting the unknown parameters into their phase-and-amplitude
-components. The implementation of  this approach is called \textsc{lincal}. \textsc{lincal} is derived from first principals in Appendix~\ref{sec:lincal}. 
-Currently, redundant calibration is usually implemented using a layered algorithmic strategy.
+%
+%There are two recent developments in skymodel-based calibration which should be explicitly mentioned at this point, since they form the foundation this paper builds on.
+The first is the discovery of \textsc{StEfCal} (Statistically Efficient Calibration) \citep{Mitchell:MWA-cal,Salvini2014}, which is an alternating direction implicit (ADI) skymodel-based calibration algorithm.
+It works by assuming the gains are known while solving for their conjugates and then performing the reverse.
+%This procedure results in the linearization of the calibration algorithm and reduces the computational complexity of calibration from $\mathbb{O}(N^3)$ to $\mathbb{O}(N^2)$, where $N$ denotes the number of antennas in the array.
+%The second development is the idea: instead of breaking the calibration problem into its real and imaginary parts and then solving for the real and imaginary parts of the unknown parameters separately we can treat the antenna gains and their conjugates as separate parameters to estimate \citep{Smirnov2015}.
+\citet{Smirnov2015} also showed that the \textsc{StEfCal} algorithm itself is merely a special case of this gain-and-conjugate framework.
+%
+%Two baselines are redundant if their baseline difference vectors have the same length and orientation, which imply that they measure the same $uv$-modes.
+When we use a redundant array layout, we dramatically reduce the number of unknowns.
+If our array is redundant enough we reduce the number of unknowns to such an extent that we are able to solve for the unknown observed visibilities themselves in addition to the antenna gains.
+Redundant calibration is the procedure by which we estimate both the unknown observed visibilities and the antenna gains.
+%
+%Redundant calibration was popularized in the 1980's. \citet{Noordam1982} used redundant calibration and the WSRT (Westerbork Synthesis Radio Telescope) to produce high-dynamic range images of 3C84.
+%The first implementations of redundant calibration made use of the logarithm function, because when the logarithm is applied to the scalar measurement equation, redundant calibration is transformed into a linear problem \citep{Wieringa1992,Camps2003,Liu2010}.
+The implementation associated with this approach is commonly referred to as \textsc{logcal}.
+%We briefly discuss \textsc{logcal} in Appendix~\ref{sec:logcal}.
+The next improvement in redundant calibration was to perform a first order Taylor expansion, which is equivalent to the Gauss-Newton (GN) algorithm \citep{Kurien2016}, of the measurement equation and then to solve for the perturbations of the model-parameters \citep{Liu2010}.
+\citet{Liu2010} perform the first order Taylor expansion of the measurement equation by splitting the unknown parameters into their phase-and-amplitude components.
+The implementation of  this approach is called \textsc{lincal}.
+\textsc{lincal} is derived from first principals in Appendix~\ref{sec:lincal}.
+%Currently, redundant calibration is usually implemented using a layered algorithmic strategy.
 \textsc{logcal} and \textsc{lincal} are both implemented in \textsc{omnical}\footnote{https://github.com/jeffzhen/omnical}, which is a layered redundant calibration pipeline \citep{Zheng2014,Ali2015}.
-The solutions of \textsc{logcal} are used as initial input for \textsc{lincal}, since the GN algorithm requires that its starting parameter guesses be in the vicinity 
-of the true solution.
-
-It is also interesting to mention here that a completely different calibration strategy can also be used to calibrate redundant arrays, \textsc{CorrCal} \citep{Sievers2017}\footnote{https://github.com/sievers/corrcal2}.
-In \textsc{CorrCal} the unknown
-visibilities are no longer solved for. The sky is actually modelled, via two covariance matrices. 
-The first covariance matrix is associated with known point sources. The second covariance matrix models the remaining sky in a statistical sense. The unknown gains, however, are still 
-estimated. We do not elaborate any further on this approach in this paper.
-
-New instruments like LOFAR (Low Frequency Array) \citep{Noorishad2012}, PAPER (Precision Array for Probing the Epoch of Reionization) \citep{Ali2015} and HERA (Hydrogen Epoch of Reionization Array) \citep{deboer2015} have also sparked a renewed interest in redundant calibration.
-Using redundancy is especially advantageous when we are trying to detect the EoR (Epoch of Reionization) 21 cm signal,
-which is inherently quite faint \citep{Parsons2012}. Since redundant baselines make statistically independent measurements of the same $uv$-modes redundancy can greatly 
-improve the SNR (signal-to-noise ratio) of EoR detection experiments. This explains why PAPER and HERA are highly redundant arrays, as they are both primarily used for EoR 21 cm detection experiments.
-
-\noindent
-\textbf{[GIANNI TO REWRITE AND EXTEND EOR SECTION]}
-
-In this paper we propose to use the complex framework introduced by \citet{Smirnov2015} instead of the traditional amplitude-and-phase approach used by \textsc{lincal}.
-We also propose to use the Levenberg-Marquardt (LM) gradient-based algorithm instead of the Gauss-Newton (GN) algorithm as it exhibits better convergence properties.
-Using the LM algorithm reduces the number of algorithmic layers that are needed to perform redundant calibration.
-
-Two ways of speeding up redundant calibration have recently been proposed. The first technique makes use of the ADI method, in which we alternate between solving the gains, their conjugates and 
-the true observed visibilities \citep{Wijnholds2012,Marthi2014}. Note that the fact that \citet{Marthi2014} make use of the ADI method is not immediately clear
-from the paper itself.
+%The solutions of \textsc{logcal} are used as initial input for \textsc{lincal}, since the GN algorithm requires that its starting parameter guesses be in the vicinity of the true solution.
+%
+%It is also interesting to mention here that a completely different calibration strategy can also be used to calibrate redundant arrays, \textsc{CorrCal} \citep{Sievers2017}\footnote{https://github.com/sievers/corrcal2}.
+In \textsc{CorrCal} the unknown visibilities are no longer solved for.
+The sky is actually modelled, via two covariance matrices.
+%The first covariance matrix is associated with known point sources.
+The second covariance matrix models the remaining sky in a statistical sense.
+The unknown gains, however, are still estimated.
+We do not elaborate any further on this approach in this paper.
+%
+%New instruments like LOFAR (Low Frequency Array) \citep{Noorishad2012}, PAPER (Precision Array for Probing the Epoch of Reionization) \citep{Ali2015} and HERA (Hydrogen Epoch of Reionization Array) \citep{deboer2015} have also sparked a renewed interest in redundant calibration.
+Using redundancy is especially advantageous when we are trying to detect the EoR (Epoch of Reionization) 21 cm signal, which is inherently quite faint \citep{Parsons2012}.
+Since redundant baselines make statistically independent measurements of the same $uv$-modes redundancy can greatly improve the SNR (signal-to-noise ratio) of EoR detection experiments.
+This explains why PAPER and HERA are highly redundant arrays, as they are both primarily used for EoR 21 cm detection experiments.
+%
+%\noindent
+%\textbf{[GIANNI TO REWRITE AND EXTEND EOR SECTION]}
+%
+%In this paper we propose to use the complex framework introduced by \citet{Smirnov2015} instead of the traditional amplitude-and-phase approach used by \textsc{lincal}.
+%We also propose to use the Levenberg-Marquardt (LM) gradient-based algorithm instead of the Gauss-Newton (GN) algorithm as it exhibits better convergence properties.
+%Using the LM algorithm reduces the number of algorithmic layers that are needed to perform redundant calibration. 
+%
+%Two ways of speeding up redundant calibration have recently been proposed.
+The first technique makes use of the ADI method, in which we alternate between solving the gains, their conjugates and the true observed visibilities \citep{Wijnholds2012,Marthi2014}.
+Note that the fact that \citet{Marthi2014} make use of the ADI method is not immediately clear from the paper itself.
 We will elaborate on this in Section~\ref{sec:adi}.
 The second is to use the preconditioned conjugate gradient (PCG) method.
-PCG allows us to exploit the sparsity inherent in the redundant calibration problem
-to speed it up \citep{Liu2010}. A proper comparison between these two approaches has not been done before.
+PCG allows us to exploit the sparsity inherent in the redundant calibration problem to speed it up \citep{Liu2010}.
+A proper comparison between these two approaches has not been done before.
 Therefore, we also compare the ADI \citep{Marthi2014} and PCG \citep{Liu2010} methods in this paper.
-We decided to focus on the complex framework proposed by \citet{Smirnov2015} rather than the traditional phase-and-amplitude approach as it
-allows us to trivially prove that the ADI method proposed by \citep{Marthi2014} is merely a special case of the more general LM algorithm.
-
-%Moreover, this framework also 
-%allows us to show that the ADI method proposed by \citep{Marthi2014} can for all intents and purposes be renamed to redundant \textsc{StEfCal} \citep{Mitchell:MWA-cal,Salvini2014}.
-
-In Section~\ref{sec:sky_wirtinger}, we review the skymodel-based complex-and-conjugate calibration approach proposed by \citet{Smirnov2015}.
-In Section~\ref{sec:red_wirtinger}, we show that this same approach can also be used to perform redundant calibration.
-In Section~\ref{sec:results}, we compare the computational complexity of two recently proposed redundant calibration speedup algorithms with one another.
-We end the paper with some final conclusions.
-
+We decided to focus on the complex framework proposed by \citet{Smirnov2015} rather than the traditional phase-and-amplitude approach as it allows us to trivially prove that the ADI method proposed by \citep{Marthi2014} is merely a special case of the more general LM algorithm.
+%
+%In Section~\ref{sec:sky_wirtinger}, we review the skymodel-based complex-and-conjugate calibration approach proposed by \citet{Smirnov2015}.
+%In Section~\ref{sec:red_wirtinger}, we show that this same approach can also be used to perform redundant calibration.
+%In Section~\ref{sec:results}, we compare the computational complexity of two recently proposed redundant calibration speedup algorithms with one another.
+%We end the paper with some final conclusions.
+%
+%
+%
+%
+%
 % Main contributions of the paper:
@@
 
-{\bf GB: ok here are my comments/questions.
-\begin{itemize}
-\item I think we wanna start with Section 2 being a short review of standard calibration but in the complex optimization framework.
+%{\bf GB: ok here are my comments/questions.
+%\begin{itemize}
+%\item I think we wanna start with Section 2 being a short review of standard calibration but in the complex optimization framework.
 Therefore I would begin with section 3.1, following by section 2.2.
@@
 15;
-\item we need more context for section 3.2.
+%\item we need more context for section 3.2.
 Why do we need to introduce a different indexing?
@@
 As it stands, I think it should remain where it is if we can make a connection to standard calibration, otherwise it should go in section 3.3;
-\item section 3.3 should merge with 4 where redundant calibration is introduced;
-\item section 3.5 goes together with the description of the simulations;
-\item section 4.1: I do not understand the definition in eq.
+%\item section 3.3 should merge with 4 where redundant calibration is introduced;
+%\item section 3.5 goes together with the description of the simulations;
+%\item section 4.1: I do not understand the definition in eq.
 32.
@@
 I also suggest to use the symbol $\equiv$ rather than $=$ for definitions throughout the paper;
-\item can section 4.2.2 be moved to the appendix? Or how much of it?
-\item section 5 should be called something like "Implementation"...
+%\item can section 4.2.2 be moved to the appendix? Or how much of it?
+%\item section 5 should be called something like "Implementation"...
 In Oleg's paper he writes "Inverting $J^H J$ and separability"...
@@
 I also suggest to revert the order in the way this is presented: we first show that the Hessian is almost diagonal (is this only true in the complex represntation?), then we derive ADI and show that is the same as Marthi \& Chengalur;
-\item I have not seen the simulations where you inserted a phase wrap and you calibrate it. Is there a reason why it is not included?
-\end{itemize}
-}
+%\item I have not seen the simulations where you inserted a phase wrap and you calibrate it. Is there a reason why it is not included?
+%\end{itemize}
+%}
 
@@
 % \end{equation}
-% where $\Lambda$ is known as the objective function, $\brho$ is the residual vector, $\bdelta$ is the observed data vector, $\bnu$ is the model-predicted data vector and $\bzeta$ is the model-parameter vector that minimizes Eq.~\eqref{eq:least_squares}.
-% In Eq.~\eqref{eq:least_squares}, we assume that the optimization problem is real valued.
+% where $\Lambda$ is known as the objective function, $\brho$ is the residual vector, $\bdelta$ is the observed data vector, $\bnu$ is the model-predicted data vector and $\bzeta$ is the model-parameter vector that minimizes equation~\ref{eq:least_squares}.
+% In equation~\ref{eq:least_squares}, we assume that the optimization problem is real valued.
 % The main objective of least squares is to find the model-parameters which minimize the squared difference between the observed data and the model-predicted data.
 % 
-% The most standard way of solving Eq.~\eqref{eq:least_squares} is to use a gradient-type minimization algorithm.
+% The most standard way of solving equation~\ref{eq:least_squares} is to use a gradient-type minimization algorithm.
 In this paper we will be focusing on the following gradient minimization algorithms: GN and LM \citep{Levenberg1944,Marquardt1963}.
@@
 
-\section{Skymodel-based Wirtinger Calibration}
+%\section{Skymodel-based Wirtinger Calibration}
+%\section{Calibration described through the Wirtinger formalism}
+\section{Wirtinger calibration}
 \label{sec:sky_wirtinger}
-%In Section~\ref{sec:s_mes}, we present the scalar measurement equation. Next, we define some indexing functions in Section~\ref{sec:indexing}. The measurement 
-%equation and the indexing functions from Section~\ref{sec:indexing} are then used in Section~\ref{sec:r_mes} to phrase redundant calibration as a least-squares problem, i.e.
-Eq.~\eqref{eq:least_squares_red}.
-%We present the signal-to-noise ratio (SNR) definition we used in this paper in Section~\ref{sec:snr}. The limitations of this paper are discussed in Section~\ref{sec:scope}.
-
-%\subsection{Scalar Measurement Equation}
-%\label{sec:s_mes}
-The observed visibility $d_{pq}$ measured by the baseline formed by antenna $p$ and $q$ is described by
+
+In a radio interferometer, the true sky visibilities $y_{pq}$ measured by a baseline formed by antenna $p$ and $q$ are always ``corrupted" by the non-ideal response of the receiver, which is often incorporated into a single, receiver-based complex number $g$.
+The observed visibility $d_{pq}$ is given therefore by \citep{ME1,ME2,RRIME1}
 \begin{equation}
 \label{eq:vis_definition}
-d_{pq} = g_{p}\conj{g_q}y_{pq} + n_{pq},
-\end{equation}
-where $g_{p}$ and $g_{q}$ are the direction independent gains associated with antenna $p$ and $q$, $y_{pq}$ denotes the true visibility that baseline $pq$ measured
-and $n_{pq}$ is the thermal noise component. Conjugation is denoted by $\conj{(*)}$. The true value of $g_p$, $g_q$ and $y_{pq}$ are unknown. We estimate the true value of
-$g_p$, $g_q$ and $y_{pq}$ by calibrating our data and then imaging the result.
-
-The real and imaginary components of the thermal noise are normally distributed with a mean of zero and a standard deviation $\sigma$, where   
-\begin{equation}
-\sigma \propto \frac{T_{\textrm{sys}}}{\sqrt{\Delta \nu \tau}}. 
-\end{equation}
-Here, $T_{\textrm{sys}}$ is equal to the system temperature, $\Delta \nu$ is the observational bandwidth and $\tau$ is the integration time per visibility.
-
-Eq.~\eqref{eq:vis_definition} can also be expressed in the following vector form 
-\begin{equation}
-\label{eq:vis_definition}
+d_{pq} = g_{p}\conj{g_q} \, y_{pq} + n_{pq},
+\end{equation}
+where $\conj{(*)}$ indicates conplex conjugation and $n_{pq}$ is the thermal noise component. 
+The real and imaginary components of the thermal noise are normally distributed with a mean of zero and a
+standard deviation
+\begin{equation}
+\sigma \propto \frac{T_{\textrm{sys}}}{\sqrt{\Delta \nu \tau}},
+\end{equation}
+where $T_{\textrm{sys}}$ is equal to the system temperature, $\Delta \nu$ is the observational bandwidth and $\tau$ is the integration time per visibility.
+
+Considering the number of visibilities $B$ (i.e. baselines) measured by an array of $N$ elements $B = \frac{N^2-N}{2}$, equation~\ref{eq:vis_definition} can be expressed in the following vector form:
+\begin{equation}
+\label{eq:vis_linear_definition}
 \bd = \bv + \bn, 
@@
  \left [ \bd \right]_{\alpha_{pq}} &= d_{pq}, & \left [ \bv \right ]_{\alpha_{pq}} &= v_{pq}=g_p y_{pq} \conj{g_q},\nonumber\\
- \left [ \bn \right ]_{\alpha_{pq}} &= n_{pq}, &  &\label{eq:vec_definitions}
+ \left [ \bn \right ]_{\alpha_{pq}} &= n_{pq}, &  &\label{eq:vec_linear_definitions}
 \end{align}
@@
 \end{equation}
-The function $\alpha_{pq}$ maps composite antenna indexes to unique single indexes, i.e. 
-\begin{equation}
-\{\alpha_{12},\alpha_{13},\cdots,\alpha_{1N},\alpha_{23},\cdots,\alpha_{N-1N}\}
-\end{equation}
-and 
-\begin{equation}
-\{1,2,\cdots,N-1,N,\cdots,B\}
-\end{equation}
-are equal. We denote the number of baselines in an array with $B$, i.e. $B = \frac{N^2-N}{2}$. The vectors in Eq.~\eqref{eq:vec_definitions} are column vectors of size $B$ (i.e. $p<q$).
-
-We can now formulate calibration as a least-squares optimization problem:
+The function $\alpha_{pq}$ therefore maps composite antenna indexes to unique single indexes, i.e:
+\begin{equation}
+\{\alpha_{12},\alpha_{13},\cdots,\alpha_{N-1N}\} = \{1,2,\cdots,B\} 
+\end{equation}
+%\begin{eqnarray}
+%\{1,2,\cdots,N-1,N,\cdots,B\}	 =  \nonumber\\
+%				 =  \{\alpha_{12},\alpha_{13},\cdots,\alpha_{1N},\alpha_{23},\cdots,\alpha_{N-1N}\} 
+%\end{eqnarray}
+%and 
+%\begin{equation}
+%\{1,2,\cdots,N-1,N,\cdots,B\}
+%\end{equation}
+%are equal. We denote the number of baselines in an array with $B$, i.e. $B = \frac{N^2-N}{2}$. 
+The vectors in equation~\ref{eq:vec_linear_definitions} are column vectors of size $B$ (i.e. $p<q$).
+
+Radio interferometric calibration aims to determine the best estimate of $\bg = [g_1,g_2,\cdots,g_N]^T$ in order to correct the data and, following equation~\ref{eq:vis_linear_definition}, can be formulated as a non linear least-squares optimization problem:
 \begin{equation}
@@
 \end{equation}
-where $\Lambda$ is the objective function, $\br$ is the residual vector and $\bg =[g_1,\cdots,g_N]^T$.
-
-The gradient-based minimization algorithms that are generally used to solve least squares problems generally require the model (in Eq.~\eqref{eq:least_squares} the model-predicted vector is denoted by $\bv$) to be differentiable
-towards each model parameter (in Eq.~\eqref{eq:least_squares} the model-parameter vector is denoted by $\bg$), i.e.
-these methods make use of a Jacobian matrix $\bJ$.
-In this paper we will be focusing on the following gradient-based minimization algorithms: GN and LM \citep{Levenberg1944,Marquardt1963}.
-When the least squares problem is complex it becomes less straightforward to apply these gradient-based minimization methods, as
-many complex functions are not analytic (no Taylor expansion around a fiducial point exists) and therefore not differentiable if the classic notion of differentiation is used. For instance, if the 
-classical definition of differentiation is used then $\frac{\partial \conj{z}}{\partial z}$, where $z \in \mathbb{C}$, does not exist.
-
-The only way to circumvent the differentiability conundrum associated with complex least squares problems is to recast our complex optimization problem in such a way that $\bv$ becomes analytic in its argument.
-The standard way of achieving this is to divide the complex optimization problem into its real and imaginary parts and then to solve for the real and imaginary parts of the unknown model-parameters separately.
-
-Recently, \citet{Sorber2012} proposed an alternative strategy, which involves the use of Wirtinger calculus \citep{Wirtinger1927}. The Wirtinger derivatives 
-are defined to be 
+where $\Lambda$ is the objective function and $\br$ is the residual vector.
+In standard interferometric calibration, $y_{pq}$ is assumed to be known at some level, for instance through the observation of previously known calibration sources.
+The knowledge of $y_{pq}$ is often named ``sky model".
+
+The gradient-based minimization algorithms that are generally used to solve non-linear least-squares problems are generally solved by using gradient-based minimization algorithms (i.e.
+Gauss--Newton -- GN -- or Levenberg--Marquardt -- LM) that require the model ($\bv$ in equation~\ref{eq:least_squares}) to be differentiable towards each parameter.
+%i.e. these methods make use of a Jacobian matrix $\bJ$. In this paper we will be focusing on the following gradient-based minimization algorithms: GN and LM \citep{Levenberg1944,Marquardt1963}.
+When the least squares problem is complex, it becomes less straightforward to apply these gradient-based minimization methods, as
+many complex functions are not differentiable if the classic notion of differentiation is used, i.e. $\frac{\partial \conj{z}}{\partial z}$ does not exist if $z \in \mathbb{C}$.
+%are not analytic (no Taylor expansion around a fiducial point exists) and therefore not differentiable if the classic notion of differentiation is used.
+For instance, if the classical definition of differentiation is used then $\frac{\partial \conj{z}}{\partial z}$, where $z \in \mathbb{C}$, does not exist.
+
+In order to circumvent the differentiability conundrum associated with complex least squares problems, standard interferometric calibration divides the complex optimization problem into its real and imaginary parts and solves for the real and imaginary parts of the unknown model parameters separately.
+\citet{Smirnov2015} showed, however, that this approach is not needed if complex calculus \citep{Wirtinger1927} is directly adopted.
+The Wirtinger derivatives are defined as:
 \begin{align}
 \label{eq:wir}
-\frac{\partial}{\partial z} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} -  i \frac{\partial}{\partial y} \right ),&\frac{\partial}{\partial \conj{z}} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} +  i \frac{\partial}{\partial y} \right ).
+\frac{\partial}{\partial z} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} -  i \frac{\partial}{\partial y} \right ),&\frac{\partial}{\partial \conj{z}} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} +  i \frac{\partial}{\partial y} \right ),
 \end{align}
-Using the above definitions we can now easily compute the following:
+which lead to the following relations:
 \begin{align}
+\label{eq:wir_z}
 \frac{\partial z}{\partial z} & = 1, & \frac{\partial \conj{z}}{\partial z}&=0, & \frac{\partial z}{\partial \conj{z}} & = 0, & \frac{\partial \conj{z}}{\partial \conj{z}}&=1.
 \end{align}
-If we use the Wirtinger derivative, Eq.~\eqref{eq:wir}, instead of the classic notion of a derivative to define the gradient operator then 
-$\bv$ becomes analytic in $\bz$ and $\conj{\bz}$ as a whole.
-The Wirtinger derivative can therefore be used to derive complex variants of the real-valued GN and LM algorithms.
-In these alternative algorithms we treat the complex parameters and their conjugates as separate variables.
-
-\citet{Smirnov2015} recently demonstrated that the Wirtinger derivative can be used to perform normal skymodel-based calibration. We briefly summarize their approach below:
-\begin{enumerate}
-
-\item Eq.~\eqref{eq:least_squares} is first recast to
+If the gradient operator is defined using equation~\ref{eq:wir}, the model $\bv$ now becomes analytic in both $\bg$ and $\conj{\bg}$ and equation \ref{eq:wir_z} can be used to derive the complex variants of the real-valued GN and LM algorithms.
+In the complex GN and LM algorithms, complex parameters and their conjugates are treated as separate variables.
+
+Assuming that $y_{pq}$ is known, equation~\ref{eq:least_squares} is recast as \citep{Smirnov2015}:
+%
+%The only way to circumvent the differentiability conundrum associated with complex least squares problems is to recast our complex optimization problem in such a way that $\bv$ becomes analytic in its argument.
+The standard way of achieving this is to divide the complex optimization problem into its real and imaginary parts and then to solve for the real and imaginary parts of the unknown model-parameters separately.
+%
+%Recently, \citet{Sorber2012} proposed an alternative strategy, which involves the use of Wirtinger calculus \citep{Wirtinger1927}. The Wirtinger derivatives are defined to be 
+%
+%\citet{Smirnov2015} recently demonstrated that the Wirtinger derivative can be used to perform normal skymodel-based calibration. We briefly summarize their approach below:
+%
 \begin{equation}
@@
 where $\breve{\br} = [\br^T,\conj{\br}^T]^T$, $\breve{\bd} = [\bd^T,\conj{\bd}^T]^T$, $\breve{\bv} = [\bv^T,\conj{\bv}^T]^T$ and $\breve{\bg} = [\bg^T,\conj{\bg}^T]^T$.
-\item The complex GN update step is then defined as
+
+The complex GN update is therefore defined as:
 \begin{equation}
@@
 \begin{align}
+\label{eq:jac_entries}
 [\bmJ]_{\alpha_{pq},i} &= \frac{\partial v_{pq}}{\partial g_i}, & [\bmJ^*]_{\alpha_{pq},i} &= \frac{\partial v_{pq}}{\partial \conj{g}_i}. 
 \end{align}
-Note that the Wirtinger derivative is used in Eq.~\eqref{eq:Jacobian_skymodel} and as such the entries of $\bJ$ are well defined.
-\item The complex LM update rule is very similar, the major difference being a damping factor $\lambda$ is introduced:
+%
+%
+%Note that the Wirtinger derivative is used in equation~\ref{eq:Jacobian_skymodel} and as such the entries of $\bJ$ are well defined.
+
+The complex LM update is very similar, the major difference being a damping factor $\lambda$ is introduced:
 \begin{equation}
@@
 \end{equation}
-where $\bD=\bI\odot\bJ^H\bJ$.
-\item Eq.~\eqref{eq:GN_update_skymodel} or Eq.~\eqref{eq:LM_update_skymodel} can now be used iteratively to update the 
-parameter vector $\breve{\bg}$:
-\begin{equation}
-\label{eq:update}
+where $\bD=\bI\odot\bJ^H\bJ$. Note the use of the Wirtinger derivatives in equation~\ref{eq:jac_entries}.
+We will refer to $\bJ^H\bJ$ as the Hessian matrix $\bH$ and to $\bJ^H\bJ + \lambda\bD$ as the modified Hessian matrix $\bmH$ throughout this paper.
+
+Equation~\ref{eq:GN_update_skymodel} or~\ref{eq:LM_update_skymodel} can now be used iteratively to update the parameter vector $\breve{\bg}$:
+\begin{equation}
+\label{eq:update_skymodel}
 \breve{\bg}_{k+1} = \breve{\bg}_{k} + \Delta \breve{\bg}_{k}. 
 \end{equation}
-The parameter vector needs to be initialized (an all one vector is one possible choice) and is updated until convergence is reached.
-To improve readability we do not explicitly denote the iteration subscript $k$ in
-Eq.~\eqref{eq:GN_update_skymodel} and Eq.~\eqref{eq:LM_update_skymodel}.
-\end{enumerate}
-We refer to $\bJ^H\bJ$ as the Hessian matrix $\bH$ and to $\bJ^H\bJ + \lambda\bD$ as the modified Hessian matrix $\bmH$ in this paper. 
+until convergence is reached.
+
+In the case of the GN algorithm, the parameter update step simplifies and becomes \citep{Smirnov2015}
+\begin{equation}
+\label{eq:one_half}
+\breve{\bg}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{1}{2}\breve{\bg}_{k}. 
+\end{equation}
+
+%Note that in equation~\ref{eq:one_half} the influence that the estimate of the parameters from the previous iteration has on the current iteration is scaled by a half, while
+%in equation~\ref{eq:two_thrids} it is scaled by two thirds. Note the similarity between the last terms of equation~\ref{eq:stef_alpha} and equation~\ref{eq:one_half} when $\alpha$ is a half.
+
+%
+%The parameter vector needs to be initialized (an all one vector is one possible choice) and is updated until convergence is reached.
+To improve readability we do not explicitly denote the iteration subscript $k$ in equation~\ref{eq:GN_update_skymodel} and equation~\ref{eq:LM_update_skymodel}.
 % \subsection{Indexing Functions}
@@
 % In general there are two ways of indexing the true visibilities associated with different baselines.
-We can use the composite index $pq$, which is the strategy that is followed in Eq.~\eqref{eq:vis_definition},
+We can use the composite index $pq$, which is the strategy that is followed in equation~\ref{eq:vis_definition},
 % or we can use a single unique baseline index which we assign in numerical order to each composite index, assuming the composite indexes are arranged in standard baseline order.
@@
 
+\citet{Smirnov2015} realized that the diagonal entries of the Hessian matrix $\bH$ are much more significant than its off-diagonal entries, i.e.
+$\bH$ is nearly diagonal.
+By approximating $\bH$ by its diagonal and substituting the approximate Hessian matrix the LM parameter update step becomes \citep{Smirnov2015}:
+\begin{align}
+\breve{\bg}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bg}_k,\label{eq:stef_lambda} \nonumber \\
+ &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bg}_k, \label{eq:stef_alpha}  
+\end{align}
+%\begin{align}
+%\breve{\bg}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bg}_k,\label{eq:stef_lambda}\\
+% &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bg}_k, \label{eq:stef_alpha}  
+%\end{align}
+where $\alpha = \frac{1}{1+\lambda}$. Note that equation~\ref{eq:one_half} and equation~\ref{eq:stef_alpha} are   
+not dependent on $\breve{\br}$.
+
+Interestingly enough, if $\lambda = 0$ we obtain the odd parameter update step of \textsc{StEfCal}\footnote{$k\in\{0,2,\cdots\}$}, and if $\lambda=1$ (which corresponds
+to $\alpha=\frac{1}{2}$) we obtain the even parameter update step of \textsc{StEfCal}\footnote{$k\in\{1,3,\cdots\}$} \citep[\textsc{StEfCal},][]{Mitchell:MWA-cal,Salvini2014}.
+In the \textsc{StEfCal} algorithm, the measurement equation, equation~\ref{eq:vis_definition}, is linearized by assuming that the gains are known, but that their conjugates are not.
+Under this assumption, the system of equations become linear and the conjugates of the gains can be obtained straightforwardly.
+Starting from the latest value of the gain conjugates, an updated estimate of the gains can be obtained and the succession continues until convergence is reached.
+Alternating between solving and fixing different sets of parameters (which is exactly what \textsc{StEFCal} does) is in general referred to 
+as the alternating direction implicit method (ADI). The \textsc{StEFCal} algorithm reduces the computational complexity of calibration from $\mathbb{O}(N^3)$ to $\mathbb{O}(N^2)$.
+
+%A special case of solutions referred to as the alternating direction implicit \citep[ADI,][]{Salvini2014} algorithm is obtained by choosing $\lambda = 0$ \citep{Smirnov2015}.
+In particular, this choice of $\lambda$ is the parameter estimation step of the \textsc{StEfCal} algorithm \citep{Mitchell:MWA-cal,Salvini2014}, the most used implementation of ADI algorithms.
+%{\bf (GB: Trienko, I am not sure about where to place the paragraph below till the end of the subsection...
+probably because I'm not sure what ``even" and ``odd" \textsc{StEfCal} steps are...
+can we add the GN step after equation~\ref{eq:stef_lambda}???)}
+%On the other hand, if we choose $\lambda$ equal to one (which corresponds
+%to $\alpha=\frac{1}{2}$) we obtain the new parameter estimation step of \textsc{StEfCal} used during even iterations.
+Note the similarity between the last terms of equation~\ref{eq:stef_alpha} and equation~\ref{eq:one_half} when $\alpha$ is a half.
+%Furthermore, the simplified skymodel-based GN equivalent of equation~\ref{eq:two_thrids} is equal to
+%\citep{Smirnov2015}
+%\begin{equation}
+%\label{eq:one_half}
+%\breve{\bg}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{1}{2}\breve{\bg}_{k}. 
+%\end{equation}
+%Note that in equation~\ref{eq:one_half} the influence that the estimate of the parameters from the previous iteration has on the current iteration is scaled by a half, while
+%in equation~\ref{eq:two_thrids} it is scaled by two thirds.
+
+%\subsection{Alternating Direction Implicit Method}
+%\label{sec:adi}
+%We start this section by briefly reviewing the ADI method as it pertains to skymodel-based calibration in Section~\ref{sec:sbc} \citep{Salvini2014,Smirnov2015}.
+%
+%In Section~\ref{sec:red_c}, we will show that the fast redundant calibration algorithm, that is presented in \citet{Marthi2014}, is in fact nothing more than an extension of the skymodel-based ADI method, presented in \citet{Salvini2014}, to the redundant calibration use case (i.e.
+redundant \textsc{StEfCal}).
+In doing so we will justify the naming convention we have been using for it.
+%
+%In Section~\ref{sec:adi_results}, we present some ADI simulation results. 
+%
+%\subsubsection{Skymodel-based Calibration}
+%\label{sec:sbc}
+%\textsc{StEfCal} is an ADI method in which the measurement equation, equation~\ref{eq:vis_definition}, is linearized by assuming that the gains are known, but that their conjugates are unknown \citep{Mitchell:MWA-cal,Salvini2014}.
+Under this assumption, the conjugates of the gains can be obtained trivially.
+Once the conjugates of the gains have been computed, it is possible to solve for the gains.
+One can then alternate between computing the gains and their conjugates until convergence is reached.
+By correctly exploiting the symmetry inherent in the calibration problem, this alternating approach can be replaced by a single iterative new parameter estimation step.
+\cite{Salvini2014} also show that the convergence of the algorithm can be improved, if for every even iteration, the basic new parameter estimation step is modified as follows: the current gain estimate is replaced by the average of the current gain estimate and the gain estimate of the previous odd iteration.
+%\textsc{StEFCal} lowers the computational complexity of normal skymodel-based calibration from $\mathbb{O}(N^3)$ to $\mathbb{O}(N^2)$. 
+%
+%\citet{Smirnov2015} recently showed that one can also derive \textsc{StEFCal} by viewing normal skymodel-based calibration as a complex optimization problem.
+They realized while inspecting the skymodel-based $\bH$ that its diagonal entries are much more significant than its off-diagonal entries.
+As we mention in Section~\ref{sec:results}, the exact same conclusion can be reached if we inspect the Hessian $\bH$ that is associated with redundant calibration (see Fig.~\ref{fig:hessian}).
+By approximating the skymodel-based $\bH$ by its diagonal and substituting this approximate Hessian $\widetilde{\bH}$ back into the fundamental skymodel-based LM update step we obtain \citep{Smirnov2015}
+%\begin{align}
+%\breve{\bg}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bg}_k,\label{eq:stef_lambda}\\
+% &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bg}_k, \label{eq:stef_alpha}  
+%\end{align}
+%where $\alpha = \frac{1}{1+\lambda}$. Moreover, $\lambda = \frac{1-\alpha}{\alpha}$.
+%
+%If we choose $\lambda$ to be zero, which corresponds to an $\alpha$ equal to one, we obtain the new parameter estimation step of \textsc{StEfCal} used during odd iterations.
+On the other hand, if we choose $\lambda$ equal to one (which corresponds to $\alpha=\frac{1}{2}$) we obtain the new parameter estimation step of \textsc{StEfCal} used during even iterations.
+%
+%Furthermore, the simplified skymodel-based GN equivalent of equation~\ref{eq:two_thrids} is equal to
+%\citep{Smirnov2015}
+%\begin{equation}
+%\label{eq:one_half}
+%\breve{\bg}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{1}{2}\breve{\bg}_{k}. 
+%\end{equation}
+%Note that in equation~\ref{eq:one_half} the influence that the estimate of the parameters from the previous iteration has on the current iteration is scaled by a half, while in equation~\ref{eq:two_thrids} it is scaled by two thirds.
+Note the similarity between the last terms of equation~\ref{eq:stef_alpha} and equation~\ref{eq:one_half} when $\alpha$ is a half.
+
 \section{Redundant Wirtinger Calibration}
 \label{sec:red_wirtinger}
-Redundant baselines sample the exact same visibilities in the $uv$-plane, i.e. if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$. 
-We can therefore group all the baselines that measure the exact same visibilities into the same redundant baseline group. We index redundant baseline 
-groups with single group indices instead of composite antenna pairs. Let $\phi_{pq}$ be the function that maps the composite antenna pair associated with a specific baseline 
-to the single group index of the redundant baseline group to which the baseline in question belongs, i.e. if baseline $pq$ and $rs$ are redundant then $\phi_{pq} = \phi_{rs}$.  
-The function $\phi_{pq}$ is not symmetric as 
-$\phi_{pq} = 0$ if $p\geq q$. We can define the following symmetric variant of $\phi_{pq}$:
+Interferometric baselines are named redundant when they sample the exact same visibilities in the $uv$-plane, i.e.
+if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$.
+It is convenient to group redundant visibilities together and label each group using a single index rather than using their antenna pairs as in equation~\ref{eq:vis_definition}.
+We introduce a function $\phi_{pq}$ that maps the antenna pair associated with a specific baseline to the single group index of the redundant baseline group that includes the baseline in question, i.e.
+if baseline $pq$ and $rs$ are redundant then $\phi_{pq} = \phi_{rs}$.
+The function $\phi_{pq}$ is not symmetric as $\phi_{pq} =
+0$ if $p \geq q$. 
+%Redundant baselines sample the exact same visibilities in the $uv$-plane, i.e.
+if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$.
+We can therefore group all the baselines that measure the exact same visibilities into the same redundant baseline group.
+We %index redundant baseline groups with single group indices instead of composite antenna pairs.
+%Let $\phi_{pq}$ be the function that maps the composite antenna pair associated with a specific baseline 
+%
+%to the single group index of the redundant baseline group to which the baseline in question belongs, i.e. if baseline $pq$ and $rs$ are redundant then $\phi_{pq} = \phi_{rs}$.  
+%The function $\phi_{pq}$ is not symmetric as 
+%$\phi_{pq} = 0$ if $p\geq q$. We can define the following symmetric variant of $\phi_{pq}$:
+%\begin{equation}
+%\zeta_{pq} = 
+%\begin{cases}
+%\phi_{pq}~\textrm{if}~p \leq q\\
+%\phi_{qp}~\textrm{if}~p>q
+%\end{cases}.
+%\end{equation} 
+Equation~\ref{eq:vis_definition} can be re-written for a redundant array as:
+%If the array is redundant we can rewrite equation~\ref{eq:vis_definition} as
+\begin{equation}
+\label{eq:vis_red}
+d_{pq} = g_{p}\conj{g_q}y_{\phi_{pq}} + n_{pq},
+\end{equation}
+with the same vector form as equation~\ref{eq:vis_red} if
+%equation~\ref{eq:vis_red} can also be expressed in the following vector form 
+%\begin{equation}
+%\label{eq:vis_red_vec}
+%\bd = \bv + \bn, 
+%\end{equation}
+%where 
+\begin{align}
+ \left [ \bd \right]_{\alpha_{pq}} &= d_{pq}, & \left [ \bv \right ]_{\alpha_{pq}} &= v_{pq}=g_p y_{\phi_{pq}} \conj{g_q},\nonumber\\
+ \left [ \bn \right ]_{\alpha_{pq}} &= n_{pq}, &  &\label{eq:vec_definitions}
+\end{align}
+where the vectors in Equation~\ref{eq:vec_definitions} are column vectors of size $B$ (i.e. $p<q$).
+
+We also introduce the following symmetric variant of $\phi_{pq}$:
 \begin{equation}
@@
 \phi_{qp}~\textrm{if}~p>q
-\end{cases}.
-\end{equation} 
-If the array is redundant we can rewrite Eq.~\eqref{eq:vis_definition} as
-\begin{equation}
-\label{eq:vis_red}
-d_{pq} = g_{p}\conj{g_q}y_{\phi_{pq}} + n_{pq},
-\end{equation}
-Eq.~\eqref{eq:vis_red} can also be expressed in the following vector form 
-\begin{equation}
-\label{eq:vis_red_vec}
-\bd = \bv + \bn, 
-\end{equation}
-where 
-\begin{align}
- \left [ \bd \right]_{\alpha_{pq}} &= d_{pq}, & \left [ \bv \right ]_{\alpha_{pq}} &= v_{pq}=g_p y_{\phi_{pq}} \conj{g_q},\nonumber\\
- \left [ \bn \right ]_{\alpha_{pq}} &= n_{pq}. &  &\label{eq:vec_definitions}
-\end{align}
-The vectors in Eq.~\eqref{eq:vec_definitions} are column vectors of size $B$ (i.e. $p<q$). 
-It is possible to construct a simple analytic expression for $\zeta_{pq}$ when our array is in an east-west regular configuration, namely $\zeta_{pq} = |q-p|$. 
-It becomes increasingly difficult to construct analytic expressions of $\zeta_{pq}$ for more complicated array layouts. 
-We will refer to $\zeta_{pq}$ as the symmetric geometric function. The empirically constructed symmetric geometry
-functions of three different redundant layouts are depicted in Fig.~\ref{fig:geometry_function}. We denote the range of $\zeta_{pq}$ with $\mathcal{R}(\zeta_{pq})$. The maximal element 
-that $\zeta_{pq}$ can ascertain is denoted by $L$ and can be interpreted as the maximal number of unique redundant baseline groups which can be formed for a given 
-array layout.
-Alternatively, we can interpret the symmetric geometry functions in Fig.~\ref{fig:geometry_function} as geometry matrices.
+\end{cases},
+\end{equation} 
+and we will refer to $\zeta_{pq}$ as the symmetric geometric function.
+It is possible to construct a simple analytic expression for $\zeta_{pq}$ for an east-west regular array, i.e.
+$\zeta_{pq} = |q-p|$.
+It becomes, however, increasingly difficult to construct analytic expressions of $\zeta_{pq}$ for more complicated array layouts.
+The empirically constructed symmetric geometry functions of three different redundant layouts are displayed in Figure~\ref{fig:geometry_function}.
+We denote the range of $\zeta_{pq}$ with $\mathcal{R}(\zeta_{pq})$.
+The maximal element that $\zeta_{pq}$ can ascertain is denoted by $L$ and can be interpreted as the maximal number of unique redundant baseline groups which can be formed for a given array layout.
+%Alternatively, we can interpret the symmetric geometry functions in Fig.~\ref{fig:geometry_function} as geometry matrices.
 In this alternative paradigm, $\zeta_{pq}$ denotes a matrix entry.
-The dimension of these geometry
-matrices are $N\times N$.
+The dimension of these geometry matrices are $N\times N$.
 
@@
 {\includegraphics[width=0.33\textwidth]{./REG_phi.pdf}\label{fig:REG_phi}}
-\caption{In the top panel we have three different redundant antenna layouts, namely a hexagonal (left), square (middle) and a regular east-west (right) layout. 
-In the bottom panel we have the respective symmetric redundancy geometry functions for the layouts on the top panel. We used 91 antennas
-to construct the hexagonal layout, while 100 antennas where used in the square and east-west layouts.
+\caption{Three different redundant antenna layouts: hexagonal (top left), square (top middle) and regular east-west (top right) with their respective symmetric redundancy geometry functions $\zeta_{pq}$ (bottom panels).
+We used 91 antennas to construct the hexagonal layout, while 100 antennas were used in the square and east-west layouts.
 In the case of the east-west layout we only plot the positions of the first ten antennas.
-The maximal amount of redundant baseline groups that can be formed for
-the hexagonal, square and east-west layouts are 165, 180 and 99 respectively.\label{fig:geometry_function}}
+The maximal amount of redundant baseline groups $L$ that can be formed for
+the hexagonal, square and east-west layouts are 165, 180 and 99 respectively.
+The analytic expressions of $L$ for an hexagonal, square and east-west layout is $L = 2N-\frac{1}{2}\sqrt{12N-3}-\frac{1}{2}$,
+$L=2N-2\sqrt{N}$ and $L=N-1$.\label{fig:geometry_function}}
 \end{figure*}
 
-We have now defined enough notation to enable us to formulate redundant calibration as a least-squares problem:
+We can now formulate redundant calibration as a least-squares problem:
 \begin{equation}
@@
  \bg &=[g_1,\cdots,g_N]^T, & \by &= [y_1,\cdots,y_L]^T,\nonumber\\
- \bz &= [\bg^T,\by^T]^T. &  &\label{eq:parm_definitions}
+ \bz &= [\bg^T,\by^T]^T, &  &\label{eq:parm_definitions}
  \end{align}
-The number of model parameters that we need to solve for in Eq.~\eqref{eq:least_squares_red} is $P = 2(N+L)$, since redundant calibration is a complex problem.
-We plot $L$ and $P$ for two different geometric layouts in Fig~\ref{fig:pl}.
-The analytic expressions for the red and the blue curve in Fig.~\ref{fig:ln} are \citep{Camps2003}
-\begin{equation}
-L = 2N-\frac{1}{2}\sqrt{12N-3}-\frac{1}{2} 
-\end{equation}
-and
-\begin{equation}
-L = N - 1 
-\end{equation}
-respectively.
-
-\begin{figure*}
-\centering
-\subfigure[$L$]{\includegraphics[width=0.47\textwidth]{./ln.pdf}\label{fig:ln}}
-% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
-\subfigure[$P$]{\includegraphics[width=0.47\textwidth]{./pn.pdf}\label{fig:pn}}
-\caption{In the left panel we plot the number of unique baseline groups $L$ one can form when using a hexagonal (red curve) and regular east-west layout (blue curve) as 
-a function of the number of antennas $N$ in the array.
-In the right panel we plot the total number of parameters $P$ we need to estimate when employing redundant calibration for a hexagonal (red curve) and regular east-west layout (blue curve) as
-a function of the number of antennas $N$ in the array. 
-\label{fig:pl}} 
-\end{figure*}
+where the number of model parameters to be solved of is now $P = 2(N+L)$, since redundant calibration is a complex problem.
+%The number of model parameters that we need to solve for in equation~\ref{eq:least_squares_red} is $P = 2(N+L)$, since redundant calibration is a complex problem.
 
@@
 
-As already alluded to in Section~\ref{sec:sky_wirtinger}, we can solve Eq.~\eqref{eq:least_squares_red} in one of two ways. We can split the problem into its real and
-imaginary parts and then solve for the real and imaginary parts of the unknown parameters separately or we can employ the notion of Wirtinger derivatives. The former 
-approach is discussed in Section~\ref{sec:ri}, while the latter approach is described in Section~\ref{sec:w}.
-
-\subsection{Real and Imaginary Approach}
-\label{sec:ri}
-Using least-squares to solve complex problems can be problematic as $\bv$ needs to be analytic in its argument. In the case 
-of Eq.~\eqref{eq:least_squares_red}, \citet{Liu2010} circumvents this difficulty by 
-splitting the problem into its phase-and-amplitude components.
-
-The aforementioned is achieved by first rewriting Eq.~\eqref{eq:least_squares_red} as
-\begin{equation}
-\label{eq:least_squares_lincal}
-\min_{\bmath{\varrho}} \|\breve{\br}\| = \min_{\bmath{\varrho}} \|\breve{\bd} - \breve{\bv}(\bmath{\varrho})\|, 
-\end{equation}
-where $\breve{\bd} = [\bd^T,\conj{\bd}^T]^T$, $\breve{\bv} = [\bv^T,\conj{\bv}^T]^T$ and $\bmath{\varrho}$ is the phase and amplitude parameter vector which we define in Appendix~\ref{sec:lincal}. 
-With the aid of Eq.~\eqref{eq:least_squares_lincal} we can derive the algorithm implemented in \textsc{lincal} (see Appendix~\ref{sec:lincal}).
-The Jacobian associated with Eq.~\eqref{eq:least_squares_lincal} is equal to Eq.~\eqref{eq:Jac_lin}. 
+As mentioned in Section~\ref{sec:sky_wirtinger}, equation~\ref{eq:least_squares_red} can be solved by splitting the problem into its real and
+imaginary parts and then solve for the real and imaginary parts of the unknown parameters separately \citep{Wieringa1992,Liu2010,Zheng2014}. In Appendix~\ref{sec:lincal} we review this approach.
+%As already alluded to in Section~\ref{sec:sky_wirtinger}, we can solve equation~\ref{eq:least_squares_red} in one of two ways.
+We can split the problem into its real and imaginary parts and then solve for the real and imaginary parts of the unknown parameters separately or we can employ the notion of Wirtinger derivatives.
+The former approach is discussed in Section~\ref{sec:ri}, while the latter approach is described in Section~\ref{sec:w}.
+%
+%\subsection{Real and Imaginary Approach}
+%\label{sec:ri}
+%Using least-squares to solve complex problems can be problematic as $\bv$ needs to be analytic in its argument.
+In the case of equation~\ref{eq:least_squares_red}, \citet{Liu2010} circumvents this difficulty by splitting the problem into its phase-and-amplitude components.
+%
+%The aforementioned is achieved by first rewriting equation~\ref{eq:least_squares_red} as
+%\begin{equation}
+%\label{eq:least_squares_lincal}
+%\min_{\bmath{\varrho}} \|\breve{\br}\| = \min_{\bmath{\varrho}} \|\breve{\bd} - \breve{\bv}(\bmath{\varrho})\|, 
+%\end{equation}
+%where $\breve{\bd} = [\bd^T,\conj{\bd}^T]^T$, $\breve{\bv} = [\bv^T,\conj{\bv}^T]^T$ and $\bmath{\varrho}$ is the phase and amplitude parameter vector which we define in Appendix~\ref{sec:lincal}. 
+%With the aid of equation~\ref{eq:least_squares_lincal} we can derive the algorithm implemented in \textsc{lincal} (see Appendix~\ref{sec:lincal}).
+The Jacobian associated with equation~\ref{eq:least_squares_lincal} is equal to equation~\ref{eq:Jac_lin}.
 Appendix~\ref{sec:lincal} makes it evident that \textsc{lincal} is equivalent to a basic GN update step \citep{Kurien2016}.
-Therefore, a straightforward way of improving the convergence properties of stock standard \textsc{lincal} would be to simply add the LM damping factor to the basic \textsc{lincal} update step.
-In doing so the need to use $\textsc{logcal}$ before $\textsc{lincal}$ becomes superfluous. The use of $\textsc{logcal}$ may, however, still offer 
-a computational advantage, as it is a very fast method to obtain a good starting solution, and would reduce the number of iterations needed by the LM algorithm.
+%Therefore, a straightforward way of improving the convergence properties of stock standard \textsc{lincal} would be to simply add the LM damping factor to the basic \textsc{lincal} update step.
+%In doing so the need to use $\textsc{logcal}$ before $\textsc{lincal}$ becomes superfluous.
+The use of $\textsc{logcal}$ may, however, still offer a computational advantage, as it is a very fast method to obtain a good starting solution, and would reduce the number of iterations needed by the LM algorithm.
 Proper experiments would need to be conducted to determine whether it is worth keeping $\textsc{logcal}$ in the algorithmic hierarchy.
-
-\subsection{Wirtinger Approach}
-\label{sec:w}
-
+%
+%\subsection{Wirtinger Approach}
+%\label{sec:w}
+%
 %\citet{Smirnov2015} recently proposed to use the Wirtinger derivative instead of the classic real and imaginary approach to perform normal skymodel-based calibration. We will now show
@@
 %where $\bz = [\bg,\by]^T$.
-The other vectors in Eq.~\eqref{eq:least_squares_red} were defined in Section~\ref{} \footnote{In this paper we use the same symbol to denote corrupted predicted visibilities and true observed corrupted visibilities as it
+The other vectors in equation~\ref{eq:least_squares_red} were defined in Section~\ref{} \footnote{In this paper we use the same symbol to denote corrupted predicted visibilities and true observed corrupted visibilities as it
 %is trivial to distinguish which one is implied from the context.
 Using the same symbol for these two quantities improves the readability of the paper.}.
-Note the similarity between Eq.~\eqref{eq:least_squares} and Eq.~\eqref{eq:least_squares_red},
+Note the similarity between equation~\ref{eq:least_squares} and equation~\ref{eq:least_squares_red},
 %i.e. redundant calibration can be cast as a non-linear least squares problem.
 
-Analogous to \cite{Smirnov2015}, we recast Eq.~\eqref{eq:least_squares_red} to 
+We, instead, intend to formulate the redundant calibration problem using Wirtinger calculus. We therefore recast equation~\ref{eq:least_squares_red} as
+%Analogous to \cite{Smirnov2015}, we recast equation~\ref{eq:least_squares_red} to 
 \begin{equation}
@@
 \end{equation}
-where $\breve{\bz} = [\bz^T,\conj{\bz}^T]^T$. Eq.~\eqref{eq:least_squares_complex} is the complex conjugate augmented counterpart of Eq.~\eqref{eq:least_squares_red}.
+where $\breve{\bz} = [\bz^T,\conj{\bz}^T]^T$. 
+%Equation~\ref{eq:least_squares_complex} is the complex conjugate augmented counterpart of equation~\ref{eq:least_squares_red}.
 %The fact that $\bv$ is now analytic in its argument has the immediate consequence: it becomes possible to apply complex GN
-%and complex LM to Eq.~\eqref{eq:least_squares_complex}.
-
-\subsubsection{Complex GN and LM}
-\label{sec:complex_GN_LM}
-The complex Jacobian associated with Eq.~\eqref{eq:least_squares_complex} is equal to
+%and complex LM to equation~\ref{eq:least_squares_complex}. 
+We derive the complex Jacobian associated with equation~\ref{eq:least_squares_complex} to be:
+%\subsubsection{Complex GN and LM}
+%\label{sec:complex_GN_LM}
+%The complex Jacobian associated with equation~\ref{eq:least_squares_complex} is equal to
 \begin{equation}
@@
 \end{equation}
-The GN update step is now defined as:
+We can now calculate the GN and LM update steps to be 
+%The GN update step is now defined as:
 \begin{equation}
 \label{eq:GN_update}
- \Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\br}.
-\end{equation}
-The LM update rule is defined in a very similar way:
+\Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\br}
+%\Delta \breve{\bz} &=& (\bJ^H\bJ + \lambda\bD)^{-1}\bJ^H\breve{\br},
+\end{equation}
+and 
+%The LM update rule is defined in a very similar way:
 \begin{equation}
@@
 \end{equation}
-As was the case in Section~\ref{sec:sky_wirtinger}, we can now use Eq.~\eqref{eq:GN_update} or Eq.~\eqref{eq:LM_update} to iteratively update our 
-parameter vector as follows:
+respectively. As in Section~\ref{sec:sky_wirtinger}, equation~\ref{eq:GN_update} can be used to iteratively update our parameter vector:
+%we can now use equation~\ref{eq:GN_update} or equation~\ref{eq:LM_update} to iteratively update our parameter vector as follows:
 \begin{equation}
@@
 
-\subsubsection{Simplifying GN}
-
-The following identities,
-\begin{align}
-\label{eq:identities}
-\frac{1}{3}\bJ\breve{\bz} &= \breve{\bv}, & \bJ^H\breve{\bv} &= \widetilde{\bH}\breve{\bz} 
-\end{align}
-are trivially established by mechanically showing that the left hand side of each identity in Eq.~\eqref{eq:identities} is equal to its right hand side.
-In Eq.~\eqref{eq:identities}, $\widetilde{\bH}$ denotes $\bI\odot\bH$.
-
-Replacing $\breve{\br}$ with $\breve{\bd}-\breve{\bv}$ in Eq.~\eqref{eq:GN_update} results in
-\begin{equation}
-\label{eq:temp_GN_eq}
-\Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H(\breve{\bd}-\breve{\bv}), 
-\end{equation}
-Substituting the first identity of Eq.~\eqref{eq:identities} into Eq.~\eqref{eq:temp_GN_eq} and simplifying the result 
-\begin{equation}
-\label{eq:one_thrid}
-\Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd}-\frac{1}{3}\breve{\bz}.
-\end{equation}
-It now trivially follows that:
+In the case of the GN algorithm we can simplify equation~\ref{eq:update} even further to obtain
 \begin{equation}
@@
 \end{equation}
-Eq.~\eqref{eq:two_thrids} shows us that in the case of redundant calibration we can calculate the GN update step without calculating the residual. 
-% The 
-% normal skymodel-based equivalent of Eq.~\eqref{eq:two_thrids} is equal to \citep{Smirnov2015}:
-% \begin{equation}
-% \label{eq:one_half}
-% \breve{\bg}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{1}{2}\breve{\bg}_{k}, 
-% \end{equation}
-% where $\breve{\bg} = [\bg^T,\conj{\bg}^T]^T$
-% 
-% FIX TRANSPOSE PROBLEM
-
-\section{Comparing ADI and PCG}
-\label{sec:results}
-In Section~\ref{sec:red_wirtinger} we discussed the complex GN and LM algorithms and how they relate to redundant calibration. If we inspect Eq.~\eqref{eq:GN_update} and 
-Eq.~\eqref{eq:LM_update} we see that the most expensive operation used by the GN and LM algorithms is the inversion of the Hessian, $\bH$. The most logical 
-way of speeding up the algorithm is to investigate whether we can do this inversion in some efficient way.
-
-Two recent algorithms have been proposed that are able to invert $\bH$ in an efficient manner, namely the ADI \citep{Marthi2014} and PCG \citep{Liu2010} methods. 
-Note that here we focus solely on comparing the computational complexity of inverting $\bH$ using either one of these two methods. Rating the LM or GN algorithms
-using one of these approaches relative to each other once the computational complexity of inverting $\bH$ is known is trivial. The simulations we performed to compare these two 
-methods are summarized in Section~\ref{sec:simulations}.
-
-%We restrict ourselves to comparing the most 
-%expensive operation of the two methods namely, computing the inverse of the Hessian $\bH$ (or modified Hessian $\bmH$ in the case of LM). It becomes trivial to rate these two 
-%algorithms relative to each other once the computational complexity of the inverse operation is known.
-
-The basic difference between these two speedup algorithms can be understood by inspecting Fig.~\ref{fig:hessian}. The first observation we can make by inspecting 
-Fig.~\ref{fig:hessian_reg} and Fig.~\ref{fig:hessian_hex} is that the diagonal entries of $\bH$ are more significant than the off-diagonal entries in both cases.
-It therefore makes sense to approximate the $\bH$ with its diagonal, which greatly simplifies the computational complexity of inverting 
-$\bH$, i.e. it becomes $\mathbb{O}(P)$\footnote{The cost of constructing 
-the diagonal is not included in this estimate.} \citep{Smirnov2015}. This diagonal-approximation approach is effectively ADI \citep{Marthi2014}. The exact details of why this 
-is the case is discussed in Section~\ref{sec:adi}.
-We will also show how \citet{Marthi2014}, \citet{Smirnov2015} and \citet{Salvini2014} relate to each other in that same section.
-It should be noted here that strictly speaking the ADI method actually refers to the complete LM update step using the approximate $\bH$ and not just
-to the inversion of the approximate $\bH$.
-To improve the readability of the paper, we use the acronym ADI to refer to both the faster approximate inverse approach and the complete LM algorithm that uses it.
-Which one of these two concepts we are actually referring to, however, is always perfectly clear from the context.
-%Furhtermore, the name ADI also implies that with this algorithm we alternate between computing the 
-%gains, their conjugates and the true observed visibilities. %In Appendix~\ref{sec:sbc} we briefly summarize the ADI method that is used for skymodel-based calibration.
-
-The second observation we can make from Fig.~\ref{fig:hessian_reg} and Fig.~\ref{fig:hessian_hex} is that both Hessians contain many zero entries, i.e.
-they are sparse matrices.
-This, in conjunction with the more significant diagonal entries, implies that the PCG method (presented in Section~\ref{sec:conj_grad})
-can be applied to invert them in a computationally efficient manner.
-\citet{Liu2010} proposed the PCG method to compute the inverse of $\bH$, however the computational complexity of this approach was not analyzed therein.
-In Section~\ref{sec:pcg}, we do a proper computational complexity analysis of PCG. We find that the computational complexity of PCG is asymptotically bounded by $\mathbb{O}(P^2)$, which is 
-a lot better than straightforward inversion which takes $\mathbb{O}(P^3)$.
+The above equation is derived in Appendix~\ref{sec:simplify_GN}.
+Equation~\ref{eq:two_thrids} is the redundant equivalent of equation~\ref{eq:one_half} and it shows us that in the case of redundant calibration we can calculate the GN parameter update step without calculating the residual.
 
@@
 
-%Since we approximate the Hessian in the ADI method with its diagonal the computational complexity of its inverse becomes $\mathbb{O}(P)$\footnote{The cost of constructing 
-%the diagonal is not included in this estimate.}.
-
-Even though the PCG method is computationally more expensive than the ADI method we cannot arrive at a final conclusion at this point. With the ADI method 
-we only approximate the inverse of the Hessian, while in the case of the PCG method we compute the exact inverse. So the question arises whether the LM algorithm using the PCG method 
-will converge much faster than the ADI method even though each iteration (which includes an inversion) is more expensive. We investigate this question in Section~\ref{sec:comparison}.
-
-\subsection{Simulations}
+Figure~\ref{fig:hessian} shows that the Hessian matrix $\bH$ is nearly diagonal and sparse for both a regular and hexagonal layout considered.
+We therefore follow the approach of \citet{Smirnov2015} and approximate the Hessian matrix $\bH$ with its diagonal.
+If we substitute $\bJ^H\bJ$ with $\widetilde{\bH}=\bH\odot\bI$ and replace $\breve{\br}$ with $\breve{\bd} - \breve{\bv}$ in equation~\ref{eq:GN_update} we obtain:
+\begin{equation}
+\label{eq:appr_GN}
+ \Delta \breve{\bz} \approx \widetilde{\bH}^{-1}\bJ^H(\breve{\bd}-\breve{\bv})
+\end{equation}
+Utilizing the second identity in equation~\ref{eq:identities} allows us to simplify equation~\ref{eq:appr_GN} to
+\begin{equation}
+  \Delta \breve{\bz} \approx \widetilde{\bH}^{-1}\bJ^H\breve{\bd}-\breve{\bz},
+\end{equation}
+which leads to
+\begin{equation}
+ \breve{\bz}_{k+1} \approx \widetilde{\bH}^{-1}\bJ^H\breve{\bd}.
+\end{equation}
+We obtain a similar result if we repeat the above procedure, only in this case we use equation~\ref{eq:LM_update} instead. Thus,
+\begin{align}
+\breve{\bz}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bz}_k,\label{eq:lambda}\\
+ &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bz}_k. \label{eq:alpha}  
+\end{align}
+The analytic expression of $\bJ^H\breve{\bd}$ will be very similar to the analytic 
+expression of $\bJ^H\breve{\br}$, the only difference being that in equation~\ref{eq:ab} the letter $r$ would be replaced by a $d$. If we substitute the analytic expression
+of $\bJ^H\breve{\bd}$ and $\widetilde{\bH}^{-1}$ (which can easily be constructed using Appendix~\ref{sec:analytic}) into equation~\ref{eq:alpha} we obtain the following two update rules:
+\begin{equation}
+\label{eq:g_update}
+g_{i}^{k+1} = \alpha \frac{\sum_{j\neq i} g_j^k \widetilde{y}_{ij}^{~\!\!k} d_{ij}}{\sum_{j\neq i} |g_j^k|^2|y_{\zeta_{ij}}^k|^2} + (1-\alpha) g_i^k, 
+\end{equation}
+and
+\begin{equation}
+\label{eq:y_update}
+y_{i}^{k+1} = \alpha \frac{\sum_{pq \in \mathcal{PQ}_i} \conj{g}_p^k g_q^k d_{pq}}{\sum_{pq \in \mathcal{PQ}_i}|g_p^k|^2|g_q^k|^2} + (1-\alpha) y_i^k. 
+\end{equation}
+The computational complexity of inverting $\widetilde{\bH}$ is $\mathbb{O}(P)$.
+We note that equation~\ref{eq:g_update} is the gain estimator associated with \textsc{StEfCal} and that equation~\ref{eq:g_update} and~\ref{eq:y_update} correspond to the parameter estimators derived by \citet{Marthi2014}.
+In \citet{Marthi2014}, equation~\ref{eq:g_update} and~\ref{eq:y_update} are derived by taking the
+derivative of the objective function $\Lambda$ relative to the elements of $\bg$ and $\by$, setting the intermediate results to zero and then solving
+for the unknown parameters.
+The above two equations can also be derived using the ADI method.
+Please see Appendix~\ref{sec:red_stef_ADI} for the details.
+We will refer to the above parameter update steps as redundant \textsc{StEFCal} in the remainder of the
+paper.
+
+%{\bf (GB: derived in which way? It would e useful to say in order to differentiate from us)}.
+
+The choice of $\alpha$ is somewhat unconstrained. 
+In this paper we chose $\alpha=\frac{1}{3}$ (i.e. $\lambda = 2$) and we based our decision on the choice of $\alpha$ in the 
+case of classic \textsc{StEFCal}. In the case of \textsc{StEFCal}, $\alpha = \frac{1}{2}$ (for even iterations),
+which results in the last term of equation~\ref{eq:one_half} and equation~\ref{eq:stef_alpha} being equal. 
+When $\alpha=\frac{1}{3}$, the last term of equation~\ref{eq:two_thrids} and equation~\ref{eq:alpha} are equal.
+
+%In this paper we follow the same choice done in \textsc{StEFCal}, leading to $\alpha = 1/3$.
+The following simulations show this choice to be appropriate {\bf (GB: is this true?
+TRIENKO, I MANAGED TO REACH ONLY TILL THIS POINT.
+I THINK IT IS TIME I HAND THIS BACK TO YOU AND YOU MAKE AN EFFORT TO TIE IT UP.
+i THINK WE CAN GO FROM HERE TO A MINOR REVISION NEXT TIME)}
+
+%equation~\ref{eq:g_update} and equation~\ref{eq:y_update} are the exact same iterative new parameter estimators derived by \citet{Marthi2014}.
+Also note that equation~\ref{eq:g_update} is the fundamental gain estimator associated with skymodel-based \textsc{StEfCal}.
+We have, therefore, successfully used the diagonal approximation approach proposed by \citet{Smirnov2015} to re-derive the fast redundant calibration method presented in  \citet{Marthi2014}.
+%The question now arises what value should we use for $\alpha$? In \citet{Marthi2014}, they propose 
+%$\alpha = 0.3$. Although an exact mathematical proof for the optimal choice of $\alpha$ still eludes us, it is still possible to
+%venture a clever guess by studying what the choice of $\alpha$ is in the case of skymodel-based \textsc{StEFCal}. The value of $\alpha$ is equal to a half 
+%in the case of \textsc{StEFCal} (for even iterations); see Section~\ref{sec:sbc}.
+If we wish to mimic this behaviour for the redundant calibration use case we need to scale $\breve{\bz}$ by two thirds in equation~\ref{eq:alpha} (see equation~\ref{eq:two_thrids}) which implies that we should choose $\alpha$ equal to a third, which corresponds to a $\lambda$ equal to two.
+%We used this $\lambda$ value in this paper. The following simulation results show that this choice produces 
+%good results.
+
+Optimality results of redundant \textsc{StEFCal} were presented in \citet{Marthi2014}.
+We do not repeat their analysis here.
+We do, however, present our own redundant \textsc{StEFCal} results, in Fig.~\ref{fig:prec_error}, which confirm that the $\alpha$ value proposed
+in this paper produce accurate parameter estimates. To quantify the exact error in our estimates we used the percentage error $\beta$ defined 
+as 
+\begin{equation}
+\label{eq:beta}
+\beta = \frac{\|\bv - \widehat{\bv}\|_F^2}{\|\bv\|_F^2}.
+\end{equation}
+We have already defined $\bv$ in equation~\ref{eq:vec_definitions}. Moreover, in equation~\ref{eq:beta}, $\widehat{\bv}$ denotes the parameter estimate obtained with redundant \textsc{StEFCal}.
+
+\begin{figure}
+\includegraphics[width=0.47\textwidth]{./prec_error.pdf} 
+\caption{We plot the percentage error $\beta$ between the estimated 
+predicted visibilities and the true predicted visibilities for different SNR values as a function of the number of antennas in the array $N$.
+The graph presented here was generated using the first experimental setup in Table~\ref{tab:ch_parm}.}
+\label{fig:prec_error}
+\end{figure}
+
+\section{Preconditioned Conjugate Gradient Method}
+\label{sec:pcg}
+In this section we investigate the computational complexity of the PCG method when applied to the redundant calibration problem. 
+We briefly review the PCG method in Appendix~\ref{sec:conj_grad}. If the reader is not that familiar with the PCG method we encourage 
+the reader to briefly review Appendix~\ref{sec:conj_grad} before continuing with the paper. The PCG method can be used to efficiently invert Hermitian sparse 
+matrices, which is exactly the kind of matrix $\bH$ is (see Fig.~\ref{fig:hessian} and Eq.~\ref{eq:red_H}).
+There are two factors which influence the computational complexity of the PCG method, namely the spectral condition number $\kappa$ and the sparsity of the matrix we wish to invert (see equation~\ref{eq:cg_bound}).
+
+We investigate the spectral condition number and the sparsity of the modified Hessian $\bmH$ (i.e.
+$\lambda = 2$), which is a non-singular matrix, in Section~\ref{sec:scn} and Section~\ref{sec:sparsity} respectively.
+It is interesting to note, that the Hessian $\bH$ itself is singular (from empirical tests) and that the PCG method 
+can in principal be be used to pseudo-invert it (see Appendix~\ref{sec:normal}).
+
+\subsection{Spectral Condition Number}
+\label{sec:scn}
+The convergence properties of the conjugate gradient (CG) algorithm can be dramatically improved through preconditioning. Inspecting Fig.~\ref{fig:hessian}, reveals that the Jacobian preconditioner
+is a good candidate for the redundant use case. In Fig.~\ref{fig:kappa}, we see that this choice of preconditioner performs very well, the value of the spectral condition number $\kappa$
+becomes constant and independent of problem size, i.e. the number of antennas in the array. Preconditioning therefore effectively eliminates $\kappa$ from equation~\ref{eq:cg_bound}. This result
+is confirmed by Fig.~\ref{fig:cg_itr}.
+Fig.~\ref{fig:cg_itr} shows us that the number of major iterations needed by the preconditioned conjugate gradient method to invert $\bmH$ is independent of the number of antennas in the array.
+Moreover, the number of major iterations required is also much smaller than the dimension of $\bmH$.
+
+%kappa + itr plots
+\begin{figure*}
+\centering
+\subfigure[$\kappa$]{\includegraphics[width=0.47\textwidth]{./kappa.pdf}\label{fig:kappa}}
+% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
+\subfigure[Iterations required before and after Preconditioning]{\includegraphics[width=0.47\textwidth]{./cg_itr.pdf}\label{fig:cg_itr}}
+\caption{The graphs presented here were generated using the second experimental setup in Table~\ref{tab:ch_parm}.
+In the left panel we have plotted the spectral condition number $\kappa$ of the modified Hessian $\bmH$, before (magenta and red curves) and after preconditioning (blue curve), at different SNR values as a function of the
+number of antennas $N$ in the array.
+The right hand panel confirms this result.
+In the right hand panel we have plotted the number of major iterations (line 6 of Algorithm \ref{algo:CG}) required by the conjugate
+gradient method to invert $\bmH$, before (magenta and red curves) and after preconditioning (blue and green curves), at different SNR values as a function of the number of antennas $N$ in the array.
+\label{fig:kappa_itr}} 
+\end{figure*}
+
+\subsection{Sparsity}
+\label{sec:sparsity}
+% \begin{enumerate}
+% \item $\boldsymbol{H}$: is a $(4N-2)\times(4N-2)$ or a $P \times P$ matrix. $N$ denotes the number of antennas and $P$ denotes the number of parameters. This matrix
+% contains $6N^2 - 2N - 2$ non-zero entries. It contains $16N^2 - 16N + 4$ entries. We can construct this matrix with $4N^2 -  4N$ or $\frac{1}{4} P^2 -1$ elementary operations.
+% \item $\boldsymbol{A}$: is a $(2N-1)\times(2N-1)$ matrix. This matrix has $N^2 + N - 1$ non-zero entries. We can construct this matrix with $2\frac{1}{2} (N^2 -  N)$ elementary operations.  
+% \item $\boldsymbol{B}$: is a $(2N-1)\times(2N-1)$ matrix. This matrix has $2N^2 - 2N$ non-zero entries. We can construct this matrix with $1\frac{1}{2} (N^2 -  N)$ elementary operations.
+% \item $\boldsymbol{C}$: is a $N\times N$ matrix. This matrix has $N$ non-zero entries. We can construct this matrix with $N^2-N$ elementary operations.
+% \item $\boldsymbol{D}$: is a $N \times (N-1)$ matrix. This matrix has $\frac{1}{2} (N^2 -  N)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations. 
+% \item $\boldsymbol{E}$: is a $(N-1) \times (N-1)$ matrix. This matrix has $(N-1)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.  
+% \item $\boldsymbol{F}$: is a $N \times N$ matrix. This matrix has $N^2 - N$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.
+% \item $\boldsymbol{G}$: is a $N \times (N-1)$ matrix. This matrix has $\frac{1}{2} (N^2 -  N)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.
+% \item $\boldsymbol{0}$: is a $(N-1) \times (N-1)$ all zero matrix. 
+% \end{enumerate}
+
+%We can construct $\boldsymbol{C}-\boldsymbol{G}$ in $O(N^2)$
+
+%The second factor that determines the computational complexity of the conjugate gradient method is the number of non-zero
+%entries in the modified Hessian matrix $\bmH$. 
+%We have also shown in section that the first factor, the spectral condition number of $\bH$, can be reduced 
+%to a constant by employing preconditioning and can therefore be completely eliminated from the computational complexity equation. 
+A useful matrix quantity which we will use in this section is the sparsity factor $\gamma$ of a matrix, which is the ratio between the number of zero entries and the total number of entries in a matrix.
+We
+can therefore define $\gamma$ as follows:
+\begin{equation}
+ \gamma = \left (1 - \frac{m}{P^2} \right ) 
+\end{equation}
+It is trivial to determine an analytic expression for the sparsity ratio of $\bmH$ when our redundant array is in a regular east-west geometric configuration
+The sparsity and the asymptotic sparsity ratio of $\bmH$ are equal to
+\begin{align}
+\gamma &= \frac{5N^2-7N+3}{8N^2-8N+2} & \gamma_{\infty} &= \lim_{N\rightarrow \infty}\gamma = \frac{5}{8} \label{eq:gamma}. 
+\end{align}
+However when our redundant array makes use of a more complicated geometric layout it becomes much more difficult to construct such expressions, as $\bmH$ itself 
+becomes much harder to express analytically.
+
+In Fig.~\ref{fig:gamma}, we plot the empirically determined sparsity ratios for three different geometric layouts as a function of the number antennas in the array.
+Although, we cannot explicitly calculate
+the limit of the hexagonal and square layout curves, Fig.~\ref{fig:gamma} seems to indicate that they do exist. Moreover, we also plot the left most equation
+of equation~\ref{eq:gamma} (red dashed line) and the right most equation of equation~\ref{eq:gamma} (black dotted curve) in Fig.~\ref{fig:gamma}.
+
+We can now use the sparsity factor to compute the order of the computational complexity of inverting $\bmH$ (assuming the effect of $\kappa$ is negligible) with
+\begin{equation}
+P^{c} = (1 - \gamma)P^2.
+\end{equation}
+Solving for $c$ we obtain
+\begin{align}
+c &= \log_{P}(1 - \gamma) + 2 & c_{\infty} &= \lim_{N\rightarrow \infty} c = 2. \label{eq:c}
+\end{align}
+
+The evaluation of the limit in the rightmost equation of equation~\ref{eq:c} is discussed below.
+Note that
+\begin{equation}
+\label{eq:int_limit}
+\lim_{N\rightarrow\infty} P^{\log_{P}(1-\gamma)} = 1-\gamma_{\infty},
+\end{equation}
+which is only possible if 
+\begin{equation}
+\lim_{N\rightarrow\infty} \log_{P}(1-\gamma) = 0,
+\end{equation}
+implying that $c_{\infty}$ is equal to two.
+
+equation~\ref{eq:int_limit}, implies that we can only compute $c_{\infty}$ in the rightmost equation if $\lim_{N\rightarrow \infty} \gamma$ exists. 
+In Fig.~\ref{fig:c}, we plot the empirically determined value of $c$ for three different geometric layouts as a function of the number antennas in the array.
+Furthermore, in Fig.~\ref{fig:c}, the analytic value of $c$ for the regular east-west grid is plotted as a dashed red curve, while the limit of this analytic curve is plotted with a dotted black curve.
+
+Equation~\ref{eq:gamma}, equation~\ref{eq:c} and Fig.~\ref{fig:sparsity} indicate that the computational complexity, under the assumption that $\lim_{N\rightarrow \infty} \gamma$ exists, is asymptotically bounded by $O(P^2)$.
+However, the computational complexity converges very slowly to its asymptotic value, and in general is equal to $O(P^{c})$, where $c \sim 1.7$ if a Hexagonal geometric layout containing less than 200 antennas is employed.
+
+\begin{figure*}
+\centering
+\subfigure[$\gamma$]{\includegraphics[width=0.47\textwidth]{./sparsity.pdf}\label{fig:gamma}}
+% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
+\subfigure[$c$]{\includegraphics[width=0.46\textwidth]{./comp_order.pdf}\label{fig:c}}
+\caption{In the left panel we plot the sparsity ratio $\gamma$ (the number of zero entries relative to the total number of entries in a matrix) of the modified Hessian $\bmH$ as a function of the number
+of antennas $N$ in the array for different geometries, namely an hexagonal grid (blue circles), a square grid (green crosses) and a regular east-west grid (red circles).
+For the regular east-west grid we are able to construct an analytical expression for $\gamma$ (the dashed red line).
+Taking the limit
+of this analytic expression produces the black dotted line. In the right panel we plot the order of the computational cost $c$ for inverting $\bmH$ as a function of the number 
+of antennas in the array for different geometries. We use the same color scheme in the right and left panels. The red dashed line 
+in the right panel is associated with the theoretical curve of $c$ we computed for the east-west regular grid. The limit of the red-dashed curve is two, which we 
+have plotted with a black dotted line.
+\label{fig:sparsity}} 
+\end{figure*}
+
+\section{Comparison Results}
+\label{sec:comparison}
+We are now in a position to make a final comparison between redundant \textsc{StEFCal} and the PCG method.
+Inverting $\bmH$ by assuming it is a diagonal matrix (i.e.
+using redundant \textsc{StEFCal}) is computationally inexpensive, while inverting $\bmH$ using PCG is more computationally taxing.
+On the other hand, one inversion with PCG is much more
+accurate than using an approximate inverse.
+The question is now, if the inverse using the PCG method is much more accurate will the LM implementation using it require fewer iterations than redundant \textsc{StEFCal} requires.
+Moreover, will the number
+of LM iterations PCG requires be so few that it manages to outperform redundant \textsc{StEFCal} in the end.
+To answer this we need to compute the number of approximate Hessian inversions $k$, redundant \textsc{StEFCal} computes, in the same time it takes PCG to perform a single Hessian inversion.
+We can compute this with
+\begin{align}
+\label{eq:k} 
+ kP &\approx P^{1.7}, & k &\approx P^{0.7}.
+\end{align}
+Next we need to measure the number of LM iterations we save by using the PCG method and compare it to equation~\ref{eq:k}. We make this comparison 
+in Fig.~\ref{fig:out_diff}. Fig.~\ref{fig:out_diff} tells us that redundant \textsc{StEFCal} will outperform the PCG method as the number of LM iterations we save by taking the 
+full inverse is not enough to exceed $k$. In this comparison we have only taken into account the cost of inverting $\bmH$. The PCG method, however, is still 
+useful in practice as it would be more straightforward to apply it to other calibration use cases.
+
+\begin{figure*}
+\centering
+\subfigure[Number of LM iterations required by redundant \textsc{StEFCal} and PCG.]{\includegraphics[width=0.47\textwidth]{./outerloop.pdf}\label{fig:outerloop}}
+% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
+\subfigure[Difference.]{\includegraphics[width=0.47\textwidth]{./diff.pdf}\label{fig:diff}}
+\caption{The graphs presented here were generated using the second experimental setup in Table~\ref{tab:ch_parm}. In the left panel we have the number of LM iterations 
+that is required by redundant \textsc{StEFCal} (green and blue curve) and the PCG (magenta and red curve) method to converge to a parameter error tolerance of 1e-6 whilst using different 
+SNR values as a function of the number of antennas $N$ in the array.
+In the right panel we have the average amount of LM iterations (difference between the redundant \textsc{StEFCal} and PCG curves in the left panel) we save (green and blue curve) by computing the full-inverse with the PCG method whilst using different SNR values.
+The black curve is associated with $k$, the number of LM iterations redundant \textsc{StEFCal} can perform in the same time it takes the LM implementation using 
+the PCG method to perform one iteration.
+\label{fig:out_diff}} 
+\end{figure*}
+
+%\subsubsection{Simplifying GN}
+
+% The 
+% normal skymodel-based equivalent of equation~\ref{eq:two_thrids} is equal to \citep{Smirnov2015}:
+% \begin{equation}
+% \label{eq:one_half}
+% \breve{\bg}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{1}{2}\breve{\bg}_{k}, 
+% \end{equation}
+% where $\breve{\bg} = [\bg^T,\conj{\bg}^T]^T$
+% 
+% FIX TRANSPOSE PROBLEM
+
+%\section{Comparing ADI and PCG}
+%\section{Implementation} 
+% \section{Redundant \textsc{StEfCal}}
+% \label{sec:results}
+% In Section~\ref{sec:red_wirtinger} we discussed the complex GN and LM algorithms and how they relate to redundant calibration. If we inspect equation~\ref{eq:GN_update} and 
+% equation~\ref{eq:LM_update} we see that the most expensive operation used by the GN and LM algorithms is the inversion of the Hessian, $\bH$. The most logical 
+% way of speeding up the algorithm is to investigate whether we can do this inversion in some efficient way.
+% 
+% Two recent algorithms have been proposed that are able to invert $\bH$ in an efficient manner, namely the ADI \citep{Marthi2014} and PCG \citep{Liu2010} methods. 
+% Note that here we focus solely on comparing the computational complexity of inverting $\bH$ using either one of these two methods. Rating the LM or GN algorithms
+% using one of these approaches relative to each other once the computational complexity of inverting $\bH$ is known is trivial. The simulations we performed to compare these two 
+% methods are summarized in Section~\ref{sec:simulations}.
+% 
+% %We restrict ourselves to comparing the most 
+% %expensive operation of the two methods namely, computing the inverse of the Hessian $\bH$ (or modified Hessian $\bmH$ in the case of LM). It becomes trivial to rate these two 
+% %algorithms relative to each other once the computational complexity of the inverse operation is known. 
+% 
+% The basic difference between these two speedup algorithms can be understood by inspecting Fig.~\ref{fig:hessian}. The first observation we can make by inspecting 
+% Fig.~\ref{fig:hessian_reg} and Fig.~\ref{fig:hessian_hex} is that the diagonal entries of $\bH$ are more significant than the off-diagonal entries in both cases.
+% It therefore makes sense to approximate the $\bH$ with its diagonal, which greatly simplifies the computational complexity of inverting 
+% $\bH$, i.e. it becomes $\mathbb{O}(P)$\footnote{The cost of constructing 
+% the diagonal is not included in this estimate.} \citep{Smirnov2015}. This diagonal-approximation approach is effectively ADI \citep{Marthi2014}. The exact details of why this 
+% is the case is discussed in Section~\ref{sec:adi}.
+We will also show how \citet{Marthi2014}, \citet{Smirnov2015} and \citet{Salvini2014} relate to each other in that same section.
+It should be noted here that strictly speaking the ADI method actually refers to the complete LM update step using the approximate $\bH$ and not just
+% to the inversion of the approximate $\bH$.
+To improve the readability of the paper, we use the acronym ADI to refer to both the faster approximate inverse approach and the complete LM algorithm that uses it.
+% Which one of these two concepts we are actually referring to, however, is always perfectly clear from the context.
+% %Furhtermore, the name ADI also implies that with this algorithm we alternate between computing the 
+% %gains, their conjugates and the true observed visibilities. %In Appendix~\ref{sec:sbc} we briefly summarize the ADI method that is used for skymodel-based calibration.
+% 
+% The second observation we can make from Fig.~\ref{fig:hessian_reg} and Fig.~\ref{fig:hessian_hex} is that both Hessians contain many zero entries, i.e.
+they are sparse matrices.
+This, in conjunction with the more significant diagonal entries, implies that the PCG method (presented in Section~\ref{sec:conj_grad})
+% can be applied to invert them in a computationally efficient manner.
+\citet{Liu2010} proposed the PCG method to compute the inverse of $\bH$, however the computational complexity of this approach was not analyzed therein.
+% In Section~\ref{sec:pcg}, we do a proper computational complexity analysis of PCG. We find that the computational complexity of PCG is asymptotically bounded by $\mathbb{O}(P^2)$, which is 
+% a lot better than straightforward inversion which takes $\mathbb{O}(P^3)$. 
+% 
+% 
+% 
+% %Since we approximate the Hessian in the ADI method with its diagonal the computational complexity of its inverse becomes $\mathbb{O}(P)$\footnote{The cost of constructing 
+% %the diagonal is not included in this estimate.}. 
+% 
+% Even though the PCG method is computationally more expensive than the ADI method we cannot arrive at a final conclusion at this point. With the ADI method 
+% we only approximate the inverse of the Hessian, while in the case of the PCG method we compute the exact inverse. So the question arises whether the LM algorithm using the PCG method 
+% will converge much faster than the ADI method even though each iteration (which includes an inversion) is more expensive. We investigate this question in Section~\ref{sec:comparison}.
+%
+
+% At this point there is no way of knowing if this is mere coincidence. DEF REWRITE THIS.
+% 
+% REWRTIE THIS IN A DIFFERENT WAY. START WITH CHENGULAR AND JUST MENTION THE LINK TO STEFCAL'S CHOICE OF ALPHA.
+% 
+% The update rules in equation~\ref{eq:lambda} and equation~\ref{eq:alpha} have been derived before, by Chengular... The approach they followed is discussed in the Appendix.
+% 
+% Interestingly enough, if we substitute $\alpha = 1$ into Eq.~\ref{eq:g_update} then it reduces to the skymodel-based StEFCal update rule. In Chengular they propose 
+% an $\alpha = 0.3$.
+% 
+% In an attempt to merge the terminology that has independently develop in the general calibration and redundant calibration literature and to emphasize the close
+% relationship between Stefcal and the approach derived by I will use the label Redunandat SteFCal (R-Stefcal) to refer to the NLS method proposed by ... 
+% 
+% ****\\
+% 
+% with $\alpha = \frac{1}{1+\lambda}$. Moreover $\lambda = \frac{1-\alpha}{\alpha}$. So in Marthi and Chengular (2014) they use an $\alpha$ of 0.3, which corresponds to an LM damping factor $\lambda$
+% of $2\frac{1}{3}$.
+% 
+% NB:: NEED TO ADD SNR PLOTS HERE
+% PLUS IMAGE OF HESSIAN FOR THE DISCUSSION
+
+\section{Conclusion}
+\label{sec:conclusions}
+In Section~\ref{sec:red_wirtinger} we extended \textsc{StEfCal} to the redundant use case by using the Wirtinger formalism 
+presented in \citet{Smirnov2015} (i.e. we presented redundant \textsc{StEfCal}). We also pointed out that redundant \textsc{StEfCal} was 
+first derived by \citet{Marthi2014} under a different name using a different approach than the one used in this paper. In Appendix~\ref{sec:red_stef_ADI} we showed that redundant \textsc{StEfCal}
+can also be derived using the ADI method.
+
+\citet{Liu2010} suggested to use the PCG method to speed-up redundant calibration.
+In Section~\ref{sec:pcg} we analyze the computational complexity of the PCG method when applied to the redundant calibration problem.
+In Section~\ref{sec:comparison} we do a comparison between redundant \textsc{StEfCal} and the PCG method. We found 
+that although the PCG method does greatly imporove the speed of redundant calibration it does not outperform redundant \textsc{StEfCal}. 
+% fiducial initial guess 
+% 
+% The following partial derivatives are now easily computed
+% 
+% \begin{eqnarray}
+% \frac{c_{ij}}{\partial \eta_p} &=& y_{i-j} e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}\\ 
+% \frac{c_{ij}}{\partial \varphi_p} &=&  -i y_{i-j} e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}\\
+% \frac{c_{ij}}{\partial \eta_q} &=& y_{i-j} e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}\\ 
+% \frac{c_{ij}}{\partial \varphi_q} &=&  i y_{i-j} e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}\\
+% \frac{y_{i-j}}{\partial \varphi_q} &=&  e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}
+% \end{eqnarray}
+% 
+% 
+% The Wirtinger derivative is used in the last equation.
+% 
+% \begin{eqnarray}
+% c_{ij} &\approx& c_{ij}^0 + \Delta \eta_p(y_{i-j}^0 e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0}) + \Delta \eta_q(y_{i-j}^0 e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0})\\
+% && -i\Delta\varphi_p (y_{i-j}^0e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0}) +i\Delta\varphi_q (y_{i-j}^0e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0})\\
+% && y_{i-j}^1(y_{i-j}^0e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0})\\
+% &\approx&  c_{ij}^0 + e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_p+ \Delta \eta_q - i\Delta\varphi_p + i\Delta\varphi_p))
+% \end{eqnarray}
+% 
+% \begin{equation}
+% \label{eq:residual}
+% r_{ij} = \delta_{ij} = c_{ij}-c_{ij}^0 = e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_p+ \Delta \eta_q - i\Delta\varphi_p + i\Delta\varphi_p)) 
+% \end{equation}
+% 
+% Eq.~\ref{eq:residual} allows us to construct the following linear system:
+% 
+% \begin{equation}
+% \boldsymbol{J}[\boldsymbol{\Delta \eta},\boldsymbol{\Delta \varphi},\boldsymbol{\Delta y}]^T = \boldsymbol{r}, 
+% \end{equation}
+% 
+% where $\boldsymbol{J}$ is equal to 
+% \begin{equation}
+% \boldsymbol{J} = \bigg [\overbrace{\frac{c_{pq}}{\partial \eta_p}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial \varphi_p}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial y_{t}}}^{t=1\cdots r_s} \bigg ] \bigg \} [pq] = 1\cdots N_{\textrm{b}} (p<q)
+% \end{equation}
+% or
+% \begin{equation}
+% \boldsymbol{J} = \bigg [\frac{c_{pq}}{\partial \boldsymbol{\eta}},~\frac{c_{pq}}{\partial \boldsymbol{\varphi}},~\frac{c_{pq}}{\partial \boldsymbol{y}} \bigg ] \bigg \} [pq] = 1\cdots N_{\textrm{b}} (p<q).
+% \end{equation}
+% 
+% The last column is again a Wirtinger derivative.
+% 
+% Which means we can obtain the least-squares estimate as follows:
+% 
+% \begin{equation}
+% [\boldsymbol{\Delta \eta},\boldsymbol{\Delta \varphi},\boldsymbol{\Delta y}]^T = [\boldsymbol{J}^H\boldsymbol{J}]^{-1}\boldsymbol{J}^H\boldsymbol{r}.
+% \end{equation}
+
+\bibliographystyle{mn2e}
+\bibliography{paperG2}
+
+\appendix
+
+\section{Simulations}
 \label{sec:simulations}
-%The main purpose of the simulations conducted in this paper is to genrate experimental results with which we can compare the computational complexity of ADI and PCG.
-We used an hexagonal geometric layout with an adjustable number of rings during the
-%simulations. 
-For the skymodel we generated 100 flat spectrum sources.
-Each source had a random flux value and position.
-The flux of the sources followed a Pareto distribution (with a shape parameter of 2), while their positions were uniformly
-distributed in a 3$^{\circ}$ by 3$^{\circ}$ square grid around the field center. The remaining fixed simulation parameters can all be found in Table~\ref{tab:fixed_parm}. 
-We also employed three different gain error models which we summarize in Table~\ref{tab:gain_parm}. We found that it did not matter which gain error model we used during
-our simulations, as the model we used did not have an impact on the results.
-Moreover, we had two major experimental setups, we either made use of simulated observations consisting of one frequency channel and multiple time-slots or one time-slot and multiple channels.
-We describe these two experimental setups in more detail in Table~\ref{tab:ch_parm}. As was the case for the gain error models, the experimental setup 
-that was used did not alter the simulation results.
-
-The main flow of a single simulation is as follows.
-We created corrupted visibilities using the fixed simulation parameters presented in Table~\ref{tab:fixed_parm} and either one of the experimental setups summarized in Table~\ref{tab:ch_parm}.
-We then used per time-slot and channel redundant calibration, i.e.
-we use the LM algorithm, which in turn makes use of either the ADI or PCG method, to
-estimate the gains and the true observed visibilities.
-Our initial parameter guess $\bz_0$ was chosen to be an all one vector.
-To obtain statistically significant results we performed, for both the ADI and PCG methods, multiple statistical independent simulations.
-
-Furthermore, the following SNR definition is used in this paper  
-\begin{equation}
-\label{}
-\textrm{SNR} = \frac{<\bv\odot\conj{\bv}>_{\nu,t,pq}}{<\bn\odot\conj{\bn}>_{\nu,t,pq}}, 
-\end{equation}
-where $<*>_{\nu,t,pq}$ denotes averaging over frequency, time and baseline. The definition we use here is based on the SNR definitions used in \citet{Liu2010} and \citet{Marthi2014}.
-
-%We compare ADI with PCG in Section \textbf{??}, by using the results generated from the simulations we performed using 
-%these two methods.
-
-%The results reported in this paper were generated using two major experimental setups, namely a single time-slot and multiple frequency channels setup or 
-%a single frquency channel and multiple time-slots setup. The simulation parameters which where kept fixed in both these setups can be found in 
-%Table~\ref{tab:fixed_parm}. The sky-model we used to simulate the visibilities were the same for both setups.   We used a hexagonal redundant layout for our antenna positions. We produced 
-%results for different
-
 \begin{table*}
@@
 \end{table}
-
-\subsection{Alternating Direction Implicit Method}
-\label{sec:adi}
-%In Section~\ref{sec:adi} we show how the ADI method used for skymodel-based calibration relates to the redundant ADI method presented in \citet{Marthi2014}. 
-We start this section by briefly reviewing the ADI method as it pertains to skymodel-based calibration in Section~\ref{sec:sbc} \citep{Salvini2014,Smirnov2015}.
-
-In Section~\ref{sec:red_c}, we will show that the fast redundant calibration algorithm, that is presented in \citet{Marthi2014},
-is in fact nothing more than an extension of the skymodel-based ADI method, presented in \citet{Salvini2014}, to the redundant calibration use case (i.e. redundant \textsc{StEfCal}).
-In doing so we will justify the naming convention we have been using for it.
-
-In Section~\ref{sec:adi_results}, we present some ADI simulation results. 
-%We will mainly accomplish this by using the complex framework proposed by \citet{Smirnov2015}.
-
-\subsubsection{Skymodel-based Calibration}
-\label{sec:sbc}
-\textsc{StEfCal} is an ADI method in which the measurement equation, Eq.~\eqref{eq:vis_definition}, is linearized by assuming that the gains are known, but
-that their conjugates are unknown \citep{Mitchell:MWA-cal,Salvini2014}.
-Under this assumption, the conjugates of the gains can be obtained trivially.
-Once the conjugates of the gains have been computed, it is possible to solve for the gains.
-One can then alternate between computing the gains and their conjugates until convergence
-is reached.
-By correctly exploiting the symmetry inherent in the calibration problem, this alternating approach can be replaced by a single iterative new parameter estimation step.
-\cite{Salvini2014} also show that
-the convergence of the algorithm can be improved, if for every even iteration, the basic new parameter estimation step is modified as follows: the current gain estimate is replaced by the average of the current gain estimate and the gain estimate of the previous odd iteration.
-\textsc{StEFCal} lowers the computational complexity of normal skymodel-based calibration from $\mathbb{O}(N^3)$ to $\mathbb{O}(N^2)$.
-
-\citet{Smirnov2015} recently showed that one can also derive \textsc{StEFCal} by viewing normal skymodel-based calibration as a complex optimization problem.
-They realized while inspecting the skymodel-based $\bH$ that its diagonal entries are much more significant than its off-diagonal entries.
-As we mention in Section~\ref{sec:results}, the exact same conclusion can be reached if we inspect the Hessian $\bH$ that is associated with redundant calibration (see Fig.~\ref{fig:hessian}). 
-By approximating the skymodel-based $\bH$ by its diagonal and substituting this approximate Hessian $\widetilde{\bH}$ back into the fundamental skymodel-based LM update step we obtain \citep{Smirnov2015}
-\begin{align}
-\breve{\bg}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bg}_k,\label{eq:stef_lambda}\\
- &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bg}_k, \label{eq:stef_alpha}  
-\end{align}
-where $\alpha = \frac{1}{1+\lambda}$. Moreover, $\lambda = \frac{1-\alpha}{\alpha}$.
-
-If we choose $\lambda$ to be zero, which corresponds to an $\alpha$ equal to one, we obtain the new parameter estimation step of \textsc{StEfCal} used during odd iterations.
-On the other hand, if we choose $\lambda$ equal to one (which corresponds
-to $\alpha=\frac{1}{2}$) we obtain the new parameter estimation step of \textsc{StEfCal} used during even iterations.
-
-Furthermore, the simplified skymodel-based GN equivalent of Eq.~\eqref{eq:two_thrids} is equal to
-\citep{Smirnov2015}
-\begin{equation}
-\label{eq:one_half}
-\breve{\bg}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{1}{2}\breve{\bg}_{k}. 
-\end{equation}
-Note that in Eq.~\eqref{eq:one_half} the influence that the estimate of the parameters from the previous iteration has on the current iteration is scaled by a half, while
-in Eq.~\eqref{eq:two_thrids} it is scaled by two thirds. Note the similarity between the last terms of Eq.~\eqref{eq:stef_alpha} and Eq.~\eqref{eq:one_half} when $\alpha$ is a half.
-
-\subsubsection{Redundant Calibration}
-\label{sec:red_c}
-In this section, we re-derive the fast redundant calibration algorithm presented in \citet{Marthi2014} using two different approaches.
-Both these approaches differ from the one presented in \citet{Marthi2014}.
-In \citet{Marthi2014}, the algorithm is derived by taking the derivative of the
-objective function $\Lambda$ relative to $\bg$ and $\by$, setting the intermediate results to naught and then solving for the unknown parameters.
-
-It is in fact trivial to extend the ADI method presented in \citet{Salvini2014} to the redundant use case, a.k.a.
-we now alternate between solving for the antenna gains their conjugates and the true observed visibilities.
-See Appendix~\ref{sec:red_stef_ADI} for the details.
-The result of this extension is that we obtain the redundant calibration algorithm presented in \citet{Marthi2014}.
-This implies that the fast redundant calibration algorithm presented in \citet{Marthi2014} can for all intents and purposes be referred to as redundant \textsc{StEfCal} (i.e.
-some kind of ADI method).
-
-Recall, that \citet{Smirnov2015} showed that \textsc{StEfCal} can also be derived by approximating the $\bH$, associated with skymodel-based calibration, with 
-its diagonal.
-To double check if our statement that the algorithm in \citet{Marthi2014} is in actual fact redundant \textsc{StEfCal} (i.e.
-some kind of ADI method) we will now use the diagonal approximation approach presented in \citet{Smirnov2015} to yet again re-derive the redundant calibration algorithm presented in \citet{Marthi2014}.
-The benefit of doing this second additional derivation is that it also enables us to show that the ADI method is in actual fact just a special case of the LM algorithm.
-
-Since the diagonal entries of the $\bH$, associated with redundant calibration, are more significant than its off-diagonal entries it makes 
-sense to try and approximate it with its diagonal (see Fig.~\ref{fig:hessian} and Eq.~\eqref{eq:red_H}).
-If we substitute $\bJ^H\bJ$ with $\widetilde{\bH}$ and replace $\breve{\br}$ with $\breve{\bd} - \breve{\bv}$ in Eq.~\eqref{eq:GN_update} we obtain
-\begin{equation}
-\label{eq:appr_GN}
- \Delta \breve{\bz} \approx \widetilde{\bH}^{-1}\bJ^H(\breve{\bd}-\breve{\bv})
-\end{equation}
-Utilizing the second identity in Eq.~\eqref{eq:identities} allows us to simplify Eq.~\eqref{eq:appr_GN} to
-\begin{equation}
-  \Delta \breve{\bz} \approx \widetilde{\bH}^{-1}\bJ^H\breve{\bd}-\breve{\bz}.
-\end{equation}
-It now trivially follows that
-\begin{equation}
- \breve{\bz}_{k+1} \approx \widetilde{\bH}^{-1}\bJ^H\breve{\bd}.
-\end{equation}
-We obtain a similar result if we repeat the above procedure, only in this case we use Eq.~\eqref{eq:LM_update} instead. Thus,
-\begin{align}
-\breve{\bz}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bz}_k,\label{eq:lambda}\\
- &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bz}_k. \label{eq:alpha}  
-\end{align}
-The analytic expression of $\bJ^H\breve{\bd}$ will be very similar to the analytic 
-expression of $\bJ^H\breve{\br}$, the only difference being that in Eq.~\eqref{eq:ab} the letter $r$ would be replaced by a $d$. If we substitute the analytic expression
-of $\bJ^H\breve{\bd}$ and $\widetilde{\bH}^{-1}$ (which can easily be constructed using Appendix~\ref{sec:analytic}) into Eq.~\eqref{eq:alpha} we obtain the following two update rules:
-\begin{equation}
-\label{eq:g_update}
-g_{i}^{k+1} = \alpha \frac{\sum_{j\neq i} g_j^k \widetilde{y}_{ij}^{~\!\!k} d_{ij}}{\sum_{j\neq i} |g_j^k|^2|y_{\zeta_{ij}}^k|^2} + (1-\alpha) g_i^k, 
-\end{equation}
-and
-\begin{equation}
-\label{eq:y_update}
-y_{i}^{k+1} = \alpha \frac{\sum_{pq \in \mathcal{PQ}_i} \conj{g}_p^k g_q^k d_{pq}}{\sum_{pq \in \mathcal{PQ}_i}|g_p^k|^2|g_q^k|^2} + (1-\alpha) y_i^k. 
-\end{equation}
-The computational complexity of inverting $\widetilde{\bH}$ is $\mathbb{O}(P)$.
-
-Eq.~\eqref{eq:g_update} and Eq.~\eqref{eq:y_update} are the exact same iterative new parameter estimators derived by \citet{Marthi2014}. Also note that Eq.~\eqref{eq:g_update} is the fundamental
-gain estimator associated with skymodel-based \textsc{StEfCal}.
-We have, therefore, successfully used the diagonal approximation approach proposed by \citet{Smirnov2015} to re-derive the fast redundant calibration method presented in
-\citet{Marthi2014}.
-%and in doing so we have confirmed shown that this algorithm is in actual fact nothing more than redundant \textsc{StEfCal}, i.e.
-some form of ADI method, justifying the naming convention we have been using for it.
-%Re-deriving the algorithm presented in \citet{Marthi2014} in this way serves as a further confirmation that it is indeed nothing more than redundant \textsc{StEfCal} (i.e. some kind of ADI method).
-
-The question now arises what value should we use for $\alpha$? In \citet{Marthi2014}, they propose 
-$\alpha = 0.3$. Although an exact mathematical proof for the optimal choice of $\alpha$ still eludes us, it is still possible to
-venture a clever guess by studying what the choice of $\alpha$ is in the case of skymodel-based \textsc{StEFCal}. The value of $\alpha$ is equal to a half 
-in the case of \textsc{StEFCal} (for even iterations); see Section~\ref{sec:sbc}.
-If we wish to mimic this behaviour for the redundant calibration use case we need to scale $\breve{\bz}$ by two thirds in Eq.~\eqref{eq:alpha} (see Eq.~\eqref{eq:two_thrids}) which implies that we should choose $\alpha$ equal to a third, which corresponds to a $\lambda$ equal to two.
-We used this $\lambda$ value in this paper.
-
-\subsubsection{ADI simulation results}
-\label{sec:adi_results}
-Optimality results of ADI were presented in \citet{Marthi2014}. We do not repeat their analysis here. We do, however, present our own ADI results, in Fig.~\ref{fig:prec_error}, that confirm that the
-ADI method can accurately estimate the antenna gains and the observed visibilities at different SNR values. To quantify the exact error in our estimates we used the percentage error $\beta$ defined 
-as 
-\begin{equation}
-\label{eq:beta}
-\beta = \frac{\|\bv - \widehat{\bv}\|_F^2}{\|\bv\|_F^2}.
-\end{equation}
-We have already defined $\bv$ in Eq.~\eqref{eq:vec_definitions}. Moreover, in Eq.~\eqref{eq:beta}, $\widehat{\bv}$ denotes the ADI estimate of $\bv$.
-
-\begin{figure}
-\includegraphics[width=0.47\textwidth]{./prec_error.pdf} 
-\caption{We plot the percentage error $\beta$ between the estimated 
-predicted visibilities and the true predicted visibilities for different SNR values as a function of the number of antennas in the array $N$.
-The graph presented here was generated using the first experimental setup in Table~\ref{tab:ch_parm}.}
-\label{fig:prec_error}
-\end{figure}
-
-% At this point there is no way of knowing if this is mere coincidence. DEF REWRITE THIS.
-% 
-% REWRTIE THIS IN A DIFFERENT WAY. START WITH CHENGULAR AND JUST MENTION THE LINK TO STEFCAL'S CHOICE OF ALPHA.
-% 
-% The update rules in Eq.~\eqref{eq:lambda} and Eq.~\eqref{eq:alpha} have been derived before, by Chengular... The approach they followed is discussed in the Appendix.
-% 
-% Interestingly enough, if we substitute $\alpha = 1$ into Eq.~\ref{eq:g_update} then it reduces to the skymodel-based StEFCal update rule. In Chengular they propose 
-% an $\alpha = 0.3$.
-% 
-% In an attempt to merge the terminology that has independently develop in the general calibration and redundant calibration literature and to emphasize the close
-% relationship between Stefcal and the approach derived by I will use the label Redunandat SteFCal (R-Stefcal) to refer to the NLS method proposed by ... 
-% 
-% ****\\
-% 
-% with $\alpha = \frac{1}{1+\lambda}$. Moreover $\lambda = \frac{1-\alpha}{\alpha}$. So in Marthi and Chengular (2014) they use an $\alpha$ of 0.3, which corresponds to an LM damping factor $\lambda$
-% of $2\frac{1}{3}$.
-% 
-% NB:: NEED TO ADD SNR PLOTS HERE
-% PLUS IMAGE OF HESSIAN FOR THE DISCUSSION
-
-\subsection{Preconditioned Conjugate Gradient Method}
-\label{sec:pcg}
-In this section we investigate the computational complexity of the PCG method when applied to the redundant calibration problem. 
-We briefly review the PCG method in Appendix~\ref{sec:conj_grad}. If the reader is not that familiar with the PCG method we encourage 
-the reader to briefly review Appendix~\ref{sec:conj_grad} before continuing with the paper. The PCG method can be used to efficiently invert Hermitian sparse 
-matrices, which is exactly the kind of matrix $\bH$ is (see Fig.~\ref{fig:hessian} and Eq.~\ref{eq:red_H}).
-As we mention in Appendix~\ref{sec:conj_grad} there are two factors which influence the computational complexity of the PCG method, namely the spectral condition number $\kappa$ and the sparsity of the matrix we wish to invert (see Eq.~\eqref{eq:cg_bound}).
-
-We used a $\lambda$ value of two, implying we used Eq.~\eqref{eq:LM_update} instead of Eq.~\eqref{eq:GN_update} throughout this section.
-We investigate the spectral condition number and the sparsity of the modified Hessian $\bmH$, which is a non-singular matrix, in Section~\ref{sec:scn} and Section~\ref{sec:sparsity} respectively.
-It is interesting to note, that the Hessian $\bH$ itself is singular which means that if we are to use Eq.~\eqref{eq:GN_update} instead we would need to be able to invert a singular matrix.
-In Appendix~\ref{sec:normal},
-we mention that the PCG method would in fact also be able to pseudo-invert $\bH$ directly.
-
-\subsubsection{Spectral Condition Number}
-\label{sec:scn}
-In Appendix~\ref{sec:precon} we mentioned that the convergence properties of CG can be dramatically improved through preconditioning.
-Inspecting Fig.~\ref{fig:hessian}, reveals that the Jacobian preconditioner
-is a good candidate for the redundant use case. In Fig.~\ref{fig:kappa}, we see that this choice of preconditioner performs very well, the value of the spectral condition number $\kappa$
-becomes constant and independent of problem size, i.e. the number of antennas in the array. Preconditioning therefore effectively eliminates $\kappa$ from Eq.~\eqref{eq:cg_bound}. This result
-is confirmed by Fig.~\ref{fig:cg_itr}.
-Fig.~\ref{fig:cg_itr} shows us that the number of major iterations needed by the preconditioned conjugate gradient method to invert $\bmH$ is independent of the number of antennas in the array.
-Moreover, the number of major iterations required is also much smaller than the dimension of $\bmH$.
-
-%kappa + itr plots
-\begin{figure*}
-\centering
-\subfigure[$\kappa$]{\includegraphics[width=0.47\textwidth]{./kappa.pdf}\label{fig:kappa}}
-% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
-\subfigure[Iterations required before and after Preconditioning]{\includegraphics[width=0.47\textwidth]{./cg_itr.pdf}\label{fig:cg_itr}}
-\caption{The graphs presented here were generated using the second experimental setup in Table~\ref{tab:ch_parm}.
-In the left panel we have plotted the spectral condition number $\kappa$ of the modified Hessian $\bmH$, before (magenta and red curves) and after preconditioning (blue curve), at different SNR values as a function of the
-number of antennas $N$ in the array.
-The right hand panel confirms this result.
-In the right hand panel we have plotted the number of major iterations (line 6 of Algorithm \ref{algo:CG}) required by the conjugate
-gradient method to invert $\bmH$, before (magenta and red curves) and after preconditioning (blue and green curves), at different SNR values as a function of the number of antennas $N$ in the array.
-\label{fig:kappa_itr}} 
-\end{figure*}
-
-\subsubsection{Sparsity}
-\label{sec:sparsity}
-% \begin{enumerate}
-% \item $\boldsymbol{H}$: is a $(4N-2)\times(4N-2)$ or a $P \times P$ matrix. $N$ denotes the number of antennas and $P$ denotes the number of parameters. This matrix
-% contains $6N^2 - 2N - 2$ non-zero entries. It contains $16N^2 - 16N + 4$ entries. We can construct this matrix with $4N^2 -  4N$ or $\frac{1}{4} P^2 -1$ elementary operations.
-% \item $\boldsymbol{A}$: is a $(2N-1)\times(2N-1)$ matrix. This matrix has $N^2 + N - 1$ non-zero entries. We can construct this matrix with $2\frac{1}{2} (N^2 -  N)$ elementary operations.  
-% \item $\boldsymbol{B}$: is a $(2N-1)\times(2N-1)$ matrix. This matrix has $2N^2 - 2N$ non-zero entries. We can construct this matrix with $1\frac{1}{2} (N^2 -  N)$ elementary operations.
-% \item $\boldsymbol{C}$: is a $N\times N$ matrix. This matrix has $N$ non-zero entries. We can construct this matrix with $N^2-N$ elementary operations.
-% \item $\boldsymbol{D}$: is a $N \times (N-1)$ matrix. This matrix has $\frac{1}{2} (N^2 -  N)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations. 
-% \item $\boldsymbol{E}$: is a $(N-1) \times (N-1)$ matrix. This matrix has $(N-1)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.  
-% \item $\boldsymbol{F}$: is a $N \times N$ matrix. This matrix has $N^2 - N$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.
-% \item $\boldsymbol{G}$: is a $N \times (N-1)$ matrix. This matrix has $\frac{1}{2} (N^2 -  N)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.
-% \item $\boldsymbol{0}$: is a $(N-1) \times (N-1)$ all zero matrix. 
-% \end{enumerate}
-
-%We can construct $\boldsymbol{C}-\boldsymbol{G}$ in $O(N^2)$
-
-The second factor that determines the computational complexity of the conjugate gradient method when used to perform redundant calibration is the number of non-zero
-entries in the modified Hessian matrix $\bmH$. 
-%We have also shown in section that the first factor, the spectral condition number of $\bH$, can be reduced 
-%to a constant by employing preconditioning and can therefore be completely eliminated from the computational complexity equation. 
-A useful matrix quantity which we will use in this section is the sparsity factor $\gamma$ of a matrix, which is the ratio between the number of zero entries and the total number of entries in a matrix.
-We
-can therefore define $\gamma$ as follows:
-\begin{equation}
- \gamma = \left (1 - \frac{m}{P^2} \right ) 
-\end{equation}
-It is trivial to determine an analytic expression for the sparsity ratio of $\bmH$ when our redundant array is in a regular east-west geometric configuration
-The sparsity and the asymptotic sparsity ratio of $\bmH$ are equal to
-\begin{align}
-\gamma &= \frac{5N^2-7N+3}{8N^2-8N+2} & \gamma_{\infty} &= \lim_{N\rightarrow \infty}\gamma = \frac{5}{8} \label{eq:gamma}. 
-\end{align}
-However when our redundant array makes use of a more complicated geometric layout it becomes much more difficult to construct such expressions, as $\bmH$ itself 
-becomes much harder to express analytically.
-
-In Fig.~\ref{fig:gamma}, we plot the empirically determined sparsity ratios for three different geometric layouts as a function of the number antennas in the array.
-Although, we cannot explicitly calculate
-the limit of the hexagonal and square layout curves, Fig.~\ref{fig:gamma} seems to indicate that they do exist. Moreover, we also plot the left most equation
-of Eq.~\eqref{eq:gamma} (red dashed line) and the right most equation of Eq.~\eqref{eq:gamma} (black dotted curve) in Fig.~\ref{fig:gamma}.
-
-We can now use the sparsity factor to compute the order of the computational complexity of inverting $\bmH$ (assuming the effect of $\kappa$ is negligible) with
-\begin{equation}
-P^{c} = (1 - \gamma)P^2.
-\end{equation}
-Solving for $c$ we obtain
-\begin{align}
-c &= \log_{P}(1 - \gamma) + 2 & c_{\infty} &= \lim_{N\rightarrow \infty} c = 2. \label{eq:c}
-\end{align}
-
-The evaluation of the limit in the rightmost equation of Eq.~\eqref{eq:c} is evaluated as follows: note that
-\begin{equation}
-\label{eq:int_limit}
-\lim_{N\rightarrow\infty} P^{\log_{P}(1-\gamma)} = 1-\gamma_{\infty},
-\end{equation}
-which is only possible if 
-\begin{equation}
-\lim_{N\rightarrow\infty} \log_{P}(1-\gamma) = 0,
-\end{equation}
-implying that $c_{\infty}$ is equal to two.
-
-Eq.~\eqref{eq:int_limit}, implies that we can only compute $c_{\infty}$ in the rightmost equation if $\lim_{N\rightarrow \infty} \gamma$ exists. 
-In Fig.~\ref{fig:c}, we plot the empirically determined value of $c$ for three different geometric layouts as a function of the number antennas in the array.
-Furthermore, in Fig.~\ref{fig:c}, the analytic value of $c$ for the regular east-west grid is plotted as a dashed red curve, while the limit of this analytic curve is plotted with a dotted black curve.
-
-Eq.~\eqref{eq:gamma}, Eq.~\eqref{eq:c} and Fig.~\ref{fig:sparsity} indicate that the computational complexity, under the assumption that $\lim_{N\rightarrow \infty} \gamma$ exists, is asymptotically bounded by $O(P^2)$.
-However, the computational complexity converges very slowly to its asymptotic value, and in general is equal to $O(P^{c})$, where $c \sim 1.7$ if a Hexagonal geometric layout containing less than 200 antennas is employed.
-
-\begin{figure*}
-\centering
-\subfigure[$\gamma$]{\includegraphics[width=0.47\textwidth]{./sparsity.pdf}\label{fig:gamma}}
-% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
-\subfigure[$c$]{\includegraphics[width=0.46\textwidth]{./comp_order.pdf}\label{fig:c}}
-\caption{In the left panel we plot the sparsity ratio $\gamma$ (the number of zero entries relative to the total number of entries in a matrix) of the modified Hessian $\bmH$ as a function of the number
-of antennas $N$ in the array for different geometries, namely an hexagonal grid (blue circles), a square grid (green crosses) and a regular east-west grid (red circles).
-For the regular east-west grid we are able to construct an analytical expression for $\gamma$ (the dashed red line).
-Taking the limit
-of this analytic expression produces the black dotted line. In the right panel we plot the order of the computational cost $c$ for inverting $\bmH$ as a function of the number 
-of antennas in the array for different geometries. We use the same color scheme in the right and left panels. The red dashed line 
-in the right panel is associated with the theoretical curve of $c$ we computed for the east-west regular grid. The limit of the red-dashed curve is two, which we 
-have plotted with a black dotted line.
-\label{fig:sparsity}} 
-\end{figure*}
-
-\subsection{Comparison}
-\label{sec:comparison}
-We are now in a position to make a final comparison between the ADI and PCG methods.
-Inverting $\bH$ by assuming it is a diagonal matrix (i.e.
-using the ADI method) is very cheap, while inverting $\bH$ using PCG is more expensive.
-On the other side of the coin,  one inversion with PCG is much more
-accurate than using the approximate inverse which is associated with the ADI method.
-The question is now, if the inverse using the PCG method is much more accurate will the LM implementation using it require much less LM-iterations than its ADI counterpart.
-Moreover, will the number
-of LM iterations PCG requires be so few that it manages to beat the ADI method after all.
-To answer this we first need to compute the number of iterations, $k$, the LM implementation associated with the ADI method can make in the
-same time it takes the LM implementation associated with the PCG method to perform a single iteration. We can compute this with
-\begin{align}
-\label{eq:k}
- kP &\approx P^{1.7}, & k &\approx P^{0.7}.
-\end{align}
-Next we need to measure the number of LM iterations we save by using the PCG method and compare it to Eq.\eqref{eq:k}. We make this comparison 
-in Fig.~\ref{fig:out_diff}. Fig.~\ref{fig:out_diff} tells us that the ADI method will outperform the PCG method as the number of LM iterations we save by taking the 
-full inverse is not enough to exceed $k$. In this comparison we have only taken into account the cost of inverting $\bH$. The PCG method might, however, still 
-turn out to be quite useful in practice as it would be more straightforward to apply it to other calibration use cases.
-
-\begin{figure*}
-\centering
-\subfigure[Number of LM iterations required by ADI and PCG.]{\includegraphics[width=0.47\textwidth]{./outerloop.pdf}\label{fig:outerloop}}
-% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
-\subfigure[Difference between ADI and PCG.]{\includegraphics[width=0.47\textwidth]{./diff.pdf}\label{fig:diff}}
-\caption{The graphs presented here were generated using the second experimental setup in Table~\ref{tab:ch_parm}. In the left panel we have the number of LM iterations 
-that is required for the LM algorithm using the ADI (green and blue curve) and the PCG (magenta and red curve) methods to converge to a parameter error tolerance of 1e-6 whilst using different 
-SNR values as a function of the number of antennas $N$ in the array.
-In the right panel we have the average amount of LM iterations (difference between the ADI and PCG curves in the left panel) we save (green and blue curve) by computing the full-inverse with the PCG method whilst using different SNR values.
-The black curve is associated with $k$, the number of iterations the LM implementation using the ADI method can perform in the same time it takes the LM implementation using 
-the PCG method to perform one iteration.
-\label{fig:out_diff}} 
-\end{figure*}
-
-\section{Conclusion}
-In Section~\ref{sec:main_sec} we showed that instead of using the traditional phase-and-amplitude approach to perform redundant calibration we can actually use a purely complex 
-framework. In this framework, the unknown variables and their conjugates are treated as independent variables that have to be estimated. 
-We then derived the GN and LM update steps using this complex framework (see Section~\ref{sec:complex_GN_LM}).
-In this paper we focused more on the LM algorithm than on the GN algorithm.
-The LM algorithm exhibits better convergence properties than the
-GN algorithm. This has the immediate consequence that the amount of algorithmic layers that are needed to perform redundant calibration are reduced.
-
-The most expensive operation associated with the LM algorithm (as well as the GN algorithm) is the inversion of $\bH$.
-We compared the computational complexity of two previously proposed algorithms to efficiently invert $\bH$ for the redundant calibration use case in Section~\ref{sec:results}, namely 
-ADI \citep{Marthi2014} and PCG \citep{Liu2010}.
-We showed that the computational complexity of the ADI method is $\mathbb{O}(P)$, while the computational complexity of the PCG method is asymptotically bounded
-by $\mathbb{O}(P^2)$. Both of these algorithms perform much better than stock standard inversion, which has a computational complexity of $\mathbb{O}(P^3)$.
-Although the ADI method does execute faster than the PCG method it is far less robust as it would be less straightforward to apply to other calibration uses cases. This is due to the fact that,
-the diagonal approximation we make with ADI might not be accurate enough in general.
-
-Lastly, we showed how \citet{Marthi2014}, \citet{Salvini2014} and \citet{Smirnov2015} relate to each other (see Section~\ref{sec:red_c}).
-The ADI method associated with skymodel-based calibration (i.e.
-\textsc{StEfCal}) is presented
-in \citet{Salvini2014}. \citet{Smirnov2015} showed that it is possible to re-derive this skymodel-based ADI method through complex optimization. In particular, they showed that if 
-we approximate the $\bH$ associated with the LM algorithm with its diagonal then we can rederive the ADI update rules presented in \citet{Salvini2014}. 
-\citet{Marthi2014} actually proposes to use the ADI method to perform redundant calibration.
-This fact is not immediately clear when superficially inspecting \citet{Marthi2014}.
-The redundant calibration algorithm presented in \citet{Marthi2014} was derived by taking the partial derivatives of
-the objective function towards the unknown parameters, setting the intermediate results to zero and then solving for the unknowns. In Section~\ref{sec:red_c}, we showed that 
-this algorithm is in actual fact an extension of the skymodel-based ADI method presented in \citet{Salvini2014} to the redundant calibration use case (i.e.
-we showed that this algorithm is in actual fact redundant \textsc{StEfCal}).
-We also re-derived the aforementioned redundant calibration algorithm by applying the diagonal approximation approach presented in \citet{Smirnov2015}, which serves as yet another additional confirmation
-that this algorithm is in fact redundant \textsc{StEfCal} (i.e. that it is some kind of ADI method).
-
-% fiducial initial guess 
-% 
-% The following partial derivatives are now easily computed
-% 
-% \begin{eqnarray}
-% \frac{c_{ij}}{\partial \eta_p} &=& y_{i-j} e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}\\ 
-% \frac{c_{ij}}{\partial \varphi_p} &=&  -i y_{i-j} e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}\\
-% \frac{c_{ij}}{\partial \eta_q} &=& y_{i-j} e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}\\ 
-% \frac{c_{ij}}{\partial \varphi_q} &=&  i y_{i-j} e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}\\
-% \frac{y_{i-j}}{\partial \varphi_q} &=&  e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}
-% \end{eqnarray}
-% 
-% 
-% The Wirtinger derivative is used in the last equation.
-% 
-% \begin{eqnarray}
-% c_{ij} &\approx& c_{ij}^0 + \Delta \eta_p(y_{i-j}^0 e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0}) + \Delta \eta_q(y_{i-j}^0 e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0})\\
-% && -i\Delta\varphi_p (y_{i-j}^0e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0}) +i\Delta\varphi_q (y_{i-j}^0e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0})\\
-% && y_{i-j}^1(y_{i-j}^0e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0})\\
-% &\approx&  c_{ij}^0 + e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_p+ \Delta \eta_q - i\Delta\varphi_p + i\Delta\varphi_p))
-% \end{eqnarray}
-% 
-% \begin{equation}
-% \label{eq:residual}
-% r_{ij} = \delta_{ij} = c_{ij}-c_{ij}^0 = e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_p+ \Delta \eta_q - i\Delta\varphi_p + i\Delta\varphi_p)) 
-% \end{equation}
-% 
-% Eq.~\ref{eq:residual} allows us to construct the following linear system:
-% 
-% \begin{equation}
-% \boldsymbol{J}[\boldsymbol{\Delta \eta},\boldsymbol{\Delta \varphi},\boldsymbol{\Delta y}]^T = \boldsymbol{r}, 
-% \end{equation}
-% 
-% where $\boldsymbol{J}$ is equal to 
-% \begin{equation}
-% \boldsymbol{J} = \bigg [\overbrace{\frac{c_{pq}}{\partial \eta_p}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial \varphi_p}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial y_{t}}}^{t=1\cdots r_s} \bigg ] \bigg \} [pq] = 1\cdots N_{\textrm{b}} (p<q)
-% \end{equation}
-% or
-% \begin{equation}
-% \boldsymbol{J} = \bigg [\frac{c_{pq}}{\partial \boldsymbol{\eta}},~\frac{c_{pq}}{\partial \boldsymbol{\varphi}},~\frac{c_{pq}}{\partial \boldsymbol{y}} \bigg ] \bigg \} [pq] = 1\cdots N_{\textrm{b}} (p<q).
-% \end{equation}
-% 
-% The last column is again a Wirtinger derivative.
-% 
-% Which means we can obtain the least-squares estimate as follows:
-% 
-% \begin{equation}
-% [\boldsymbol{\Delta \eta},\boldsymbol{\Delta \varphi},\boldsymbol{\Delta y}]^T = [\boldsymbol{J}^H\boldsymbol{J}]^{-1}\boldsymbol{J}^H\boldsymbol{r}.
-% \end{equation}
-
-\bibliographystyle{mn2e}
-\bibliography{paper}
-
-\appendix
+%The main purpose of the simulations conducted in this paper is to genrate experimental results with which we can compare the computational complexity of ADI and PCG.
+We used an hexagonal geometric layout with an adjustable number of rings during the
+%simulations. 
+In this section we describe the simulations we performed to generate the results we report in the paper. We first discuss some 
+of the simulation parameters we used followed by the general flow of the simulations that were conducted.
+
+For the skymodel we generated 100 flat spectrum sources.
+Each source had a random flux value and position.
+The flux of the sources followed a Pareto distribution (with a shape parameter of 2), while their positions were uniformly
+distributed in a 3$^{\circ}$ by 3$^{\circ}$ square grid around the field center. The remaining fixed simulation parameters can all be found in Table~\ref{tab:fixed_parm}. 
+We also employed three different gain error models which we summarize in Table~\ref{tab:gain_parm}. We found that it did not matter which gain error model we used during
+our simulations, as the model we used did not have an impact on the results.
+Moreover, we had two major experimental setups, we either made use of simulated observations consisting of one frequency channel and multiple time-slots or one time-slot and multiple channels.
+We describe these two experimental setups in more detail in Table~\ref{tab:ch_parm}. As was the case for the gain error models, the experimental setup 
+that was used did not alter the simulation results.
+
+The main flow of a single simulation is as follows.
+We created corrupted visibilities using the fixed simulation parameters presented in Table~\ref{tab:fixed_parm} and either one of the experimental setups summarized in Table~\ref{tab:ch_parm}.
+We then used per time-slot and channel redundant calibration, i.e.
+we use the LM algorithm, which in turn makes use of either the ADI or PCG method, to
+estimate the gains and the true observed visibilities.
+Our initial parameter guess $\bz_0$ was chosen to be an all one vector.
+To obtain statistically significant results we performed, for both the ADI and PCG methods, multiple statistical independent simulations.
+
+Furthermore, the following SNR definition is used in this paper  
+\begin{equation}
+\label{}
+\textrm{SNR} = \frac{<\bv\odot\conj{\bv}>_{\nu,t,pq}}{<\bn\odot\conj{\bn}>_{\nu,t,pq}}, 
+\end{equation}
+where $<*>_{\nu,t,pq}$ denotes averaging over frequency, time and baseline. The definition we use here is based on the SNR definitions used in \citet{Liu2010} and \citet{Marthi2014}.
 
@@
 \label{sec:logcal}
-We can rewrite Eq.~\eqref{eq:vis_red} to \citep{Liu2010}
+We can rewrite equation~\ref{eq:vis_red} to \citep{Liu2010}
 \begin{eqnarray}
@@
 
-With the aid of Eq.~\eqref{eq:logcal_amp} we can construct the following linear system:
+With the aid of equation~\ref{eq:logcal_amp} we can construct the following linear system:
 \begin{equation}
@@
 
-The least squares solution to Eq.\eqref{eq:log_system}, under the incorrect assumption that $s_{pq}$ is normally distributed, is given by
+The least squares solution to Eq.\ref{eq:log_system}, under the incorrect assumption that $s_{pq}$ is normally distributed, is given by
 \begin{equation}
@@
 \end{equation}
-How to correctly deal with the fact that $s_{pq}$ is not normally distributed is discussed in \citep{Liu2010}. A similar linear system can be constructed for Eq.~\eqref{eq:logcal_phase}. 
-From the least squares solutions to the linear systems associated with Eq.~\eqref{eq:logcal_amp} and Eq.~\eqref{eq:logcal_phase} we can compute the antenna gains as well as the true visibilities.
+How to correctly deal with the fact that $s_{pq}$ is not normally distributed is discussed in \citep{Liu2010}. A similar linear system can be constructed for equation~\ref{eq:logcal_phase}. 
+From the least squares solutions to the linear systems associated with equation~\ref{eq:logcal_amp} and equation~\ref{eq:logcal_phase} we can compute the antenna gains as well as the true visibilities.
 
@@
 \label{sec:lincal}
-%Assume that $\eta_p,\eta_q,\widetilde{\eta}_{\phi_{pq}},\varphi_p,\varphi_q$ and $\widetilde{\varphi}_{\phi_{pq}}$ are real valued variables.
-We can rewrite Eq.~\eqref{eq:vis_red_vec} to \citep{Liu2010}
+Assume that $\eta_p,\eta_q,\widetilde{\eta}_{\phi_{pq}},\varphi_p,\varphi_q$ and $\widetilde{\varphi}_{\phi_{pq}}$ are real valued variables.
+We can rewrite equation~\ref{eq:vis_red_vec} to \citep{Liu2010}
 \begin{equation}
@@
 
-The first order Taylor expansion of Eq.~\eqref{eq:lincal_vec} around the fiducial point $\bvarrho_0$ is equal to
+The first order Taylor expansion of equation~\ref{eq:lincal_vec} around the fiducial point $\bvarrho_0$ is equal to
 \begin{equation}
@@
 \end{equation}
-Eq.~\eqref{eq:LINCAL_update} is used to iteratively determine a new estimate of the parameter vector $\bvarrho$ as follows
+equation~\ref{eq:LINCAL_update} is used to iteratively determine a new estimate of the parameter vector $\bvarrho$ as follows
 \begin{equation}
@@
 \end{equation}
-Note that we do not explicitly denote the $k$ index in Eq.~\eqref{eq:LINCAL_update}. We update our parameter vector until convergence is reached.
+Note that we do not explicitly denote the $k$ index in equation~\ref{eq:LINCAL_update}.
+Moreover, equation~\ref{eq:LINCAL_update} is the GN update associated with the following minimization problem \citep{Kurien2016}
+\begin{equation}
+\label{eq:least_squares_lincal}
+\min_{\bmath{\varrho}} \|\breve{\br}\| = \min_{\bmath{\varrho}} \|\breve{\bd} - \breve{\bv}(\bmath{\varrho})\|.
+\end{equation}
 
@@
 where $\bb\in\mathbb{C}^P$ is a known complex column vector of size $P$, $\bx\in\mathbb{C}^P$ is an unknown complex column vector with the same length as $\bb$, and $\bA\in\mathbb{C}^{P\times P}$ is a square positive-definite Hermitian matrix with the appropriate dimensions.
-Moreover, using the CG method to solve Eq.~\eqref{eq:linear_system} will only be computationally advantageous if $\bA$ is also sparse (contain many zero entries).
+Moreover, using the CG method to solve equation~\ref{eq:linear_system} will only be computationally advantageous if $\bA$ is also sparse (contain many zero entries).
 Using the iterative CG method to solve large sparse linear sytems was popularized by \citet{Reid1971}. A good tutorial on the CG method can be found in \citep{Shewchuk1994}.
@@
 %\caption{Conjugate Gradient Method.
-Inputs: $\bA$, $\bb$, a starting value for $\bx$, maximum number of iterations $k_{\textrm{max}}$ and an error tolerance $\epsilon<1$ Output: $\bx$ the solution to Eq.~\eqref{eq:linear_system}.
+Inputs: $\bA$, $\bb$, a starting value for $\bx$, maximum number of iterations $k_{\textrm{max}}$ and an error tolerance $\epsilon<1$ Output: $\bx$ the solution to equation~\ref{eq:linear_system}.
 \citep{Shewchuk1994}.}\label{algo:CG}
@@
 If $\bM$ is a good preconditioner then $\kappa(\bM^{-1}\bA) \ll \kappa(\bA)$, i.e. a good preconditioner lowers the condition number of a matrix which in turn increases the convergence speed 
-of CG (see Eq.~\eqref{eq:cg_bound}).
+of CG (see equation~\ref{eq:cg_bound}).
 In this paper, we denote matrix inversion with $(*)^{-1}$.
 The best possible choice of $\bM$ would therefore be $\bA$, which would lead to $\kappa$ becoming unity (as the largest and smallest eigenvalue of the identity matrix are equal to one).
-If $\kappa$ is one then Eq.~\eqref{eq:cg_bound} simplifies and becomes $\mathbb{O}(m)$.
+If $\kappa$ is one then equation~\ref{eq:cg_bound} simplifies and becomes $\mathbb{O}(m)$.
 However, if we knew the inverse of $\bA$ then obtaining $\bx$ would be trivial and there would be no need for the CG method in the first place.
 In short, a good preconditioner $\bM$ must approximate $\bA$ fairly accurately. However, if it is not much less expensive to 
-compute the inverse of $\bM$ than that of $\bA$ we actually gain nothing (see Eq.~\eqref{eq:prec}).
+compute the inverse of $\bM$ than that of $\bA$ we actually gain nothing (see equation~\ref{eq:prec}).
 
@@
 \label{sec:normal}
-Lets assume that instead of Eq.~\eqref{eq:linear_system} we have a more general linear system, i.e.
+Lets assume that instead of equation~\ref{eq:linear_system} we have a more general linear system, i.e.
 \begin{equation}
@@
 
-There are a few important observations that we can make while inspecting Eq.~\eqref{eq:normal_equation}:
+There are a few important observations that we can make while inspecting equation~\ref{eq:normal_equation}:
 \begin{enumerate}
@@
 
-Recall that we mentioned that in general we can only solve Eq.~\eqref{eq:linear_system} with CG if $\bA$ is a square positive-definite Hermitian matrix.
-The CG method actually has an even broader applicability: it can be used to solve Eq.~\eqref{eq:linear_system} even if $\bA$ is a Hermitian positive semi-definite
+Recall that we mentioned that in general we can only solve equation~\ref{eq:linear_system} with CG if $\bA$ is a square positive-definite Hermitian matrix.
+The CG method actually has an even broader applicability: it can be used to solve equation~\ref{eq:linear_system} even if $\bA$ is a Hermitian positive semi-definite
 square matrix as long as $\bb$ is in the column range of $\bA$ \citep{Lu2015}.
-In light of \citet{Lu2015}, the two observations in the numbered list we presented in the previous paragraph imply that Eq.~\eqref{eq:normal_equation}
+In light of \citet{Lu2015}, the two observations in the numbered list we presented in the previous paragraph imply that equation~\ref{eq:normal_equation}
 can be solved by employing the CG method. The solution returned by CG is, however, only unique if $\bB^H\bB$ is positive-definite (invertible). The solution
-obtained by CG, whether it is unique or not, does minimize Eq.~\eqref{eq:norm_ls} in a least squares sense.
-
-Note that if $\bA$ is positive semi-definite, its smallest eigenvalue is zero. This implies that Eq.~\eqref{eq:kappa} is not defined. \citet{Lu2015} show that when $\bA$
+obtained by CG, whether it is unique or not, does minimize equation~\ref{eq:norm_ls} in a least squares sense.
+
+Note that if $\bA$ is positive semi-definite, its smallest eigenvalue is zero. This implies that equation~\ref{eq:kappa} is not defined. \citet{Lu2015} show that when $\bA$
 is a square positive semi-definite Hermitian matrix we do not use the spectral condition number to bound the convergence complexity of CG, but the general spectral condition
-number instead. The general spectral condition number is very similar to the spectral condition number; the only difference being that in Eq.~\eqref{eq:kappa} we replace  
+number instead. The general spectral condition number is very similar to the spectral condition number; the only difference being that in equation~\ref{eq:kappa} we replace  
 $\lambda_{\textrm{min}}$ with $\lambda_{\textrm{min}}^+$, which denotes the smallest positive eigenvalue.
@@
 \label{sec:analytic}
-If we apply the definition in Eq.~\eqref{eq:Jacobian} to Eq.~\eqref{eq:least_squares_complex} we obtain the following analytic result:
+If we apply the definition in equation~\ref{eq:Jacobian} to equation~\ref{eq:least_squares_complex} we obtain the following analytic result:
 \begin{equation}
@@
 %\end{equation}
-We use $\bzero$ to denote an all zero matrix. It is now trivial to compute the Hessian $\bH$ by using Eq.~\eqref{eq:Jacobian_red}. If we substitute Eq.~\eqref{eq:Jacobian_red} into $\bJ^H\bJ$
+We use $\bzero$ to denote an all zero matrix. It is now trivial to compute the Hessian $\bH$ by using equation~\ref{eq:Jacobian_red}. If we substitute equation~\ref{eq:Jacobian_red} into $\bJ^H\bJ$
 we obtain 
@@
 
-Furthermore, substituting Eq.~\eqref{eq:Jacobian_red} into $\bJ^H\breve{\br}$ results in
+Furthermore, substituting equation~\ref{eq:Jacobian_red} into $\bJ^H\breve{\br}$ results in
 \begin{equation}
@@
 
-\section{Redundant \textsc{StEfCal} and the ADI method}
+\section{Simplifying GN}
+\label{sec:simplify_GN}
+The following identities,
+\begin{align}
+\label{eq:identities}
+\frac{1}{3}\bJ\breve{\bz} &= \breve{\bv}, & \bJ^H\breve{\bv} &= \widetilde{\bH}\breve{\bz} 
+\end{align}
+are trivially established by mechanically showing that the left hand side of each identity in equation~\ref{eq:identities} is equal to its right hand side.
+In equation~\ref{eq:identities}, $\widetilde{\bH}$ denotes $\bI\odot\bH$.
+
+Replacing $\breve{\br}$ with $\breve{\bd}-\breve{\bv}$ in equation~\ref{eq:GN_update} results in
+\begin{equation}
+\label{eq:temp_GN_eq}
+\Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H(\breve{\bd}-\breve{\bv}), 
+\end{equation}
+Substituting the first identity of equation~\ref{eq:identities} into equation~\ref{eq:temp_GN_eq} and simplifying the result 
+\begin{equation}
+\label{eq:one_thrid}
+\Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd}-\frac{1}{3}\breve{\bz}.
+\end{equation}
+
+\section{ADI}
 \label{sec:red_stef_ADI}
-The basic skymodel-based \textsc{StEfCal} update step is equal to  the leftmost term in Eq.~\eqref{eq:g_update} (barring $\alpha$) \citep{Salvini2014}.
-Assume without any loss of generality that the array is in an east-west regular grid. Furthermore, assume that $\boldsymbol{d}$ (see Eq.~\eqref{eq:vis_red_vec}) has been re-ordered to
+The basic skymodel-based \textsc{StEfCal} update step is equal to  the leftmost term in equation~\ref{eq:g_update} (barring $\alpha$) \citep{Salvini2014}.
+Assume without any loss of generality that the array is in an east-west regular grid. Furthermore, assume that $\boldsymbol{d}$ (see equation~\ref{eq:vis_red_vec}) has been re-ordered to
 \begin{equation}
@@
 
-Eq.~\eqref{eq:vis_red_vec} can now be rewritten as 
+equation~\ref{eq:vis_red_vec} can now be rewritten as 
 \begin{equation}
@@
 \end{equation}
-if we assume that $\boldsymbol{g}$ and its conjugate are known vectors. In Eq.~\eqref{eq:linear_system},
+if we assume that $\boldsymbol{g}$ and its conjugate are known vectors. In equation~\ref{eq:linear_system},
 \begin{equation}
@@
 \end{equation}
-Substituting Eq.~\eqref{eq:RHR} and Eq.~\eqref{eq:RHd} into Eq.~\eqref{eq:y_final} and then write out the analytic expression of $y_i$ we obtain
-the leftmost term in Eq.~\eqref{eq:y_update} (barring $\alpha$).
+Substituting equation~\ref{eq:RHR} and equation~\ref{eq:RHd} into equation~\ref{eq:y_final} and then write out the analytic expression of $y_i$ we obtain
+the leftmost term in equation~\ref{eq:y_update} (barring $\alpha$).
+
+%We compare ADI with PCG in Section \textbf{??}, by using the results generated from the simulations we performed using 
+%these two methods.
+
+%The results reported in this paper were generated using two major experimental setups, namely a single time-slot and multiple frequency channels setup or 
+%a single frquency channel and multiple time-slots setup. The simulation parameters which where kept fixed in both these setups can be found in 
+%Table~\ref{tab:fixed_parm}. The sky-model we used to simulate the visibilities were the same for both setups.   We used a hexagonal redundant layout for our antenna positions. We produced 
+%results for different
+
 \label{lastpage}

### 2017-11-01 paper_G_final.tex
worked in newes comments from Gianni

--- 2017-10-11 paperG2.tex
@@
 \newcommand{\nat}{Nature}
+\newcommand{\apj}{The Astrophysical Journal}
+\newcommand{\prd}{Physical Review Letters D}
 \newcommand{\physrep}{Phys. Rep.}
+\newcommand{\araa}{Annual Review of Astronomy and Astrophysics}
+\newcommand{\pasp}{Publications of the Astronomical Society of Pacific}
 
@@
 %$^{6}$Radio Astronomy Lab., U. California, Berkeley CA 94720, USA}
-\author[T.L.~Grobler et al.]{T.L.~Grobler$^{1}$\thanks{E-mail: email@address (AVR)}, G.~Bernardi$^{1,2}$, C.L.~Carilli$^{3,4}$, J.S.~Kenyon$^{1}$, A.R.~Parsons$^{5,6}$ \newauthor and O.M.~Smirnov$^{1,2}$\\
-$^{1}$Department of Physics and Electronics, Rhodes University, PO Box 94, Grahamstown, 6140, South Africa\\
+\author[T.L.~Grobler et al.]{T.L.~Grobler$^{1}$\thanks{E-mail: email@address (AVR)}, G.~Bernardi$^{1,2}$, C.L.~Carilli$^{3,4}$, J.S.~Kenyon$^{1}$, A.R.~Parsons$^{5,6}$?
+\newauthor and O.M.~Smirnov$^{1,2}$\\
+$^{1}$Department of Physics and Electronics, Rhodes University, PO Box 94, Grahamstown, 6140, South Africa {\bf (GB: add your Stellenbosch affilitation)}\\
 $^{2}$SKA SA, 3rd Floor, The Park, Park Road, Pinelands, 7405, South Africa\\
@@
 \begin{abstract}
-%In this paper we show that instead of splitting our redundant calibration problem into its phase-and-amplitude components (the more traditional approach) we can treat the complex unknown model parameters and their complex conjugates as separate unknowns.
-We also derive the Levenberg-Maquardt and Gauss-Newton update steps associated with this alternative parameter-and-conjugate framework.
-We focus more on the Levenberg-Maquardt algorithm in this paper as it exhibits better convergence properties when compared to Gauss-Newton.
-The Gauss-Newton algorithm is currently the de facto standard that is being used to perform redundant calibration.
-It is common practice to implement redundant calibration via a hierarchical algorithmic strategy.
-Using the Levenberg-Maquardt algorithm, therefore, reduces the number of algorithmic layers that are needed to perform redundant calibration.
-We also compare two previously proposed methods to accelerate redundant calibration, namely the alternating direction implicit method and the preconditioned conjugate gradient method.
-We found that the alternating direction implicit method outperforms the preconditioned conjugate gradient method.
-However, the preconditioned conjugate gradient method is more robust as it is more straightforward to adapt to other calibration use cases.
-We make use of the parameter-and-conjugate framework in this paper, rather than the more traditional amplitude-and-phase approach as it makes it trivial to relate the alternating direction implicit method and the Levenberg-Marquardt algorithm with one another.
-Observations of the redshifted 21-cm line from the epoch of reionization have recently motivated the construction of low frequency radio arrays with highly redundant configurations that provide an alternative calibration strategy - ''redundant calibration" - and boosts sensitivity on specific scales.
-We show that redundant calibration can be formulated in the domain of complex functions of a complex variable.
-This framework reduces the uncertaninties
-\textbf{[GIANNI TO REWRITE AND FIX ABSTRACT]}
-has driven advancements in the calibration of radio interferometers and 
-Calibration of the new generation of low frequency radio interferometers faces new challenges due to their very wide field of views 
-The construction of new radio interferometers, particularly operating at low frequencies and 
-In this paper we show that instead of splitting our redundant calibration problem into its phase-and-amplitude components (the more traditional approach) we can treat the complex unknown model parameters and their complex conjugates as separate unknowns.
+Observations of the redshifted 21-cm line from the epoch of reionization have recently motivated the construction of low frequency radio arrays with highly redundant configurations that provide an alternative calibration strategy - ''redundant calibration" - and boosts sensitivity on specific spatial scales.
+In this paper we formulate calibration of redundant interferometric arrays as a complex optimization problem \citep[i.e.,][]{Smirnov2015} and derive the corresponding calibration equation which we then solve using the Levenber-Marquardt algorithm.
+This calibration approach is more robust to initial condition than current algorithms and, leveraging on an approximate matrix inversion (``redundant \textsc{StEfCal}"), allows for a very efficient implementation.
+We investigated the use of the preconditioned conjugate gradient method to solve for the calibration equation without approximating the matrix inversion and found that its computational performance is not competitive with respect to ``redundant \textsc{StEfCal}".
+The implementation of the new algorithm is made publicly available.
+%The construction of new radio interferometers, particularly operating at low frequencies and In this paper we show that instead of splitting our redundant calibration problem into its phase-and-amplitude components (the more traditional approach) we can treat the complex unknown model parameters and their complex conjugates as separate unknowns.
 We also derive the Levenberg-Maquardt and Gauss-Newton update steps associated with this alternative parameter-and-conjugate framework.
@@
 
-%that \textsc{lincal} is effectively the non-linear least-squares gradient-based Gauss-Newton algorithm.
-Therefore, an easy way to reduce the number of layers used in performing redundant calibration is to employ the Levenberg-Marquardt algorithm instead (which would improve the convergence properties of stock standard \textsc{lincal})
-
 \begin{keywords}
-circumstellar matter -- infrared: stars.
+instrumentation: interferometers -- methods: data analysis -- methods: numerical --techniques: interferometric -- cosmology: observations
 \end{keywords}
@@
 The quest for the redshifted 21-cm line from the Epoch of Reionization (EoR) is a frontier of modern observational cosmology and has motivated the construction of a series of new interferometric arrays operating at low frequencies over the last decade.
-Measurements of the EoR are difficult because of the intrinsic faintness of the cosmological signal \citep[see, for instance,][for recent reviews]{Furlanetto2015,McQuinn16} and the presence of foreground emission which is a few orders of magnitude brighter than the EoR anywhere in the sky \citep[e.g.,][]{Bernardi2009,Bernardi2010,Ghosh2011,Dillon2012,Parsons2014} and that can be separated by leveraging only on the different spectral properties (citations needed!!!).
-An exquisite calibration is therefore required as  miscalibrated foreground emission can jeopardize the EoR signal \citep[e.g.,][]{barry2016,grobler2014,ewall-wice2016}.
-These reasons motivated the design and deploment of arrays in redundant configurations like, MIT-EoR  (reference!) the Precision Array to Probe the Epoch of Reionization \citep[PAPER,][]{Parsons2014,Jacobs2015,Ali2015} and the Hydrogen Epoch of Reionization Array \citep[HERA,][]{deboer2017}.
-
-Redundant configurations have historically been abandoned because of their poor imaging performances as they measure only a limited number of spatial scales and, therefore, do not adequately sample the sky brightness distribution.
-For this very reason, however, redundant configurations provide a sensitivity boost on a selected sample of spatial scales that can be tuned to be the highest signal--to--noise ratio modes for EoR measurements \citep[][]{Parsons2012,Dillon2015}.
-
-The second advantage of redundant array configurations is that they, by definition, they measure the same spatial scales {\bf (GB: perhaps worth having the Van cittert-Zernike theorem here?)} regardless of what the sky brightness distribution looks like.
-This means that the corruptions that are introduces by the instrument can be calibrated - up to some extent - without knowing anything about the sky brightness distribution \citep[][]{Noordam1982,Wieringa1992}.
-This property is particularly desirable for EoR observations as it mitigates errors introduced by an incomplete knowledge of the sky brightness distribution.
-
-In this paper we extend the stefcal calibration algorithm to the redundant use case using the complex optimization formalism recently introduced by \cite{Smirnov2015}. 
-We also perform a proper computational analysis of the recently proposed (preconditioned conjugate gradient) PCG method when it is applied to the redundant calibration problem.
-We also compare these two approaches.
-
-present a new algorithm to calibrate redundant arrays based on the complex optimization formalism recently introduced by \cite{Smirnov2015}.
-With respect to current algorithms, it is more robust to initial conditions, it is comparatively fast and {\bf ?????...}.
-We also present an analysis of its computational complexity showing that the LM algorithm shows better convergence properties than the GN currently used \cite{Zheng14}.
-
-The paper is organized as follows: we review the complex calibration formalism by \citet{Smirnov2015,Kurien2016} in Section~\ref{sec:sky_wirtinger}, we extend it to the redundant case in Section~\ref{sec:red_wirtinger}, we test the algorithm performances in Section~\ref{sec:results} and we conclude in Section\ref{sec:conclusions}.
-
-%
-%
-%The effects (i.e.
-the ionosphere, antenna gains, etc.) a celestial signal experiences along its propagation path are modelled with the radio interferometric measurement equation (RIME) \citep{ME1,RRIME1}.
-Calibration is the procedure by which we remove the errors that these propagation effects introduce into interferometric data and is normally accomplished using a least-squares solver.
-In the case where we only model the antenna gains, which is the most basic form of the RIME, we can define calibration as finding the antenna gains that minimize the difference between our observed and predicted visibilities.
-%
-%There are two recent developments in skymodel-based calibration which should be explicitly mentioned at this point, since they form the foundation this paper builds on.
-The first is the discovery of \textsc{StEfCal} (Statistically Efficient Calibration) \citep{Mitchell:MWA-cal,Salvini2014}, which is an alternating direction implicit (ADI) skymodel-based calibration algorithm.
-It works by assuming the gains are known while solving for their conjugates and then performing the reverse.
-%This procedure results in the linearization of the calibration algorithm and reduces the computational complexity of calibration from $\mathbb{O}(N^3)$ to $\mathbb{O}(N^2)$, where $N$ denotes the number of antennas in the array.
-%The second development is the idea: instead of breaking the calibration problem into its real and imaginary parts and then solving for the real and imaginary parts of the unknown parameters separately we can treat the antenna gains and their conjugates as separate parameters to estimate \citep{Smirnov2015}.
-\citet{Smirnov2015} also showed that the \textsc{StEfCal} algorithm itself is merely a special case of this gain-and-conjugate framework.
-%
-%Two baselines are redundant if their baseline difference vectors have the same length and orientation, which imply that they measure the same $uv$-modes.
-When we use a redundant array layout, we dramatically reduce the number of unknowns.
-If our array is redundant enough we reduce the number of unknowns to such an extent that we are able to solve for the unknown observed visibilities themselves in addition to the antenna gains.
-Redundant calibration is the procedure by which we estimate both the unknown observed visibilities and the antenna gains.
-%
-%Redundant calibration was popularized in the 1980's. \citet{Noordam1982} used redundant calibration and the WSRT (Westerbork Synthesis Radio Telescope) to produce high-dynamic range images of 3C84.
-%The first implementations of redundant calibration made use of the logarithm function, because when the logarithm is applied to the scalar measurement equation, redundant calibration is transformed into a linear problem \citep{Wieringa1992,Camps2003,Liu2010}.
-The implementation associated with this approach is commonly referred to as \textsc{logcal}.
-%We briefly discuss \textsc{logcal} in Appendix~\ref{sec:logcal}.
-The next improvement in redundant calibration was to perform a first order Taylor expansion, which is equivalent to the Gauss-Newton (GN) algorithm \citep{Kurien2016}, of the measurement equation and then to solve for the perturbations of the model-parameters \citep{Liu2010}.
-\citet{Liu2010} perform the first order Taylor expansion of the measurement equation by splitting the unknown parameters into their phase-and-amplitude components.
-The implementation of  this approach is called \textsc{lincal}.
-\textsc{lincal} is derived from first principals in Appendix~\ref{sec:lincal}.
-%Currently, redundant calibration is usually implemented using a layered algorithmic strategy.
-\textsc{logcal} and \textsc{lincal} are both implemented in \textsc{omnical}\footnote{https://github.com/jeffzhen/omnical}, which is a layered redundant calibration pipeline \citep{Zheng2014,Ali2015}.
-%The solutions of \textsc{logcal} are used as initial input for \textsc{lincal}, since the GN algorithm requires that its starting parameter guesses be in the vicinity of the true solution.
-%
-%It is also interesting to mention here that a completely different calibration strategy can also be used to calibrate redundant arrays, \textsc{CorrCal} \citep{Sievers2017}\footnote{https://github.com/sievers/corrcal2}.
-In \textsc{CorrCal} the unknown visibilities are no longer solved for.
-The sky is actually modelled, via two covariance matrices.
-%The first covariance matrix is associated with known point sources.
-The second covariance matrix models the remaining sky in a statistical sense.
-The unknown gains, however, are still estimated.
-We do not elaborate any further on this approach in this paper.
-%
-%New instruments like LOFAR (Low Frequency Array) \citep{Noorishad2012}, PAPER (Precision Array for Probing the Epoch of Reionization) \citep{Ali2015} and HERA (Hydrogen Epoch of Reionization Array) \citep{deboer2015} have also sparked a renewed interest in redundant calibration.
-Using redundancy is especially advantageous when we are trying to detect the EoR (Epoch of Reionization) 21 cm signal, which is inherently quite faint \citep{Parsons2012}.
-Since redundant baselines make statistically independent measurements of the same $uv$-modes redundancy can greatly improve the SNR (signal-to-noise ratio) of EoR detection experiments.
-This explains why PAPER and HERA are highly redundant arrays, as they are both primarily used for EoR 21 cm detection experiments.
-%
-%\noindent
-%\textbf{[GIANNI TO REWRITE AND EXTEND EOR SECTION]}
-%
-%In this paper we propose to use the complex framework introduced by \citet{Smirnov2015} instead of the traditional amplitude-and-phase approach used by \textsc{lincal}.
-%We also propose to use the Levenberg-Marquardt (LM) gradient-based algorithm instead of the Gauss-Newton (GN) algorithm as it exhibits better convergence properties.
-%Using the LM algorithm reduces the number of algorithmic layers that are needed to perform redundant calibration. 
-%
-%Two ways of speeding up redundant calibration have recently been proposed.
-The first technique makes use of the ADI method, in which we alternate between solving the gains, their conjugates and the true observed visibilities \citep{Wijnholds2012,Marthi2014}.
-Note that the fact that \citet{Marthi2014} make use of the ADI method is not immediately clear from the paper itself.
-We will elaborate on this in Section~\ref{sec:adi}.
-The second is to use the preconditioned conjugate gradient (PCG) method.
-PCG allows us to exploit the sparsity inherent in the redundant calibration problem to speed it up \citep{Liu2010}.
-A proper comparison between these two approaches has not been done before.
-Therefore, we also compare the ADI \citep{Marthi2014} and PCG \citep{Liu2010} methods in this paper.
-We decided to focus on the complex framework proposed by \citet{Smirnov2015} rather than the traditional phase-and-amplitude approach as it allows us to trivially prove that the ADI method proposed by \citep{Marthi2014} is merely a special case of the more general LM algorithm.
-%
-%In Section~\ref{sec:sky_wirtinger}, we review the skymodel-based complex-and-conjugate calibration approach proposed by \citet{Smirnov2015}.
-%In Section~\ref{sec:red_wirtinger}, we show that this same approach can also be used to perform redundant calibration.
-%In Section~\ref{sec:results}, we compare the computational complexity of two recently proposed redundant calibration speedup algorithms with one another.
-%We end the paper with some final conclusions.
-%
-%
-%
-%
-%
-% Main contributions of the paper:
-% 
-% \begin{enumerate}
-%  \item Present a general framework which unifies all the techniques developed so far showing they are all related (\textsc{lincal}, non-linear estimator, etc ...). They are all non-linear
-%  least-squares techniques, i.e. they either employ Gauss-Newton or the Levenberg-Marquardt algorithms. We have to start with \textsc{lincal} showing that it is GN, then we have to motivate 
-%  why we want to use Oleg's complex optimization framework.
-%  \item Use Oleg's complex optimization framework to re-derive the non-linear technique proposed by Marthi and Chengular. The novelty lies, in the fact that by using Oleg's
-%  framework we can find analytic expressions for the Jacobian, the Hessian and the Jacobian-residual product (which is not even the case for \textsc{lincal}). State that the algorithm is 
-%  effectively related to SteFCal and is eff an independent rediscovery of SteFCal and an extension of it into the redundant domain.
-%  \item Also we present the array geometry function to help us make the derivations from Marthi and Chengular easier to read and understand for a general layout.
-%  \item We also mention at this point that in Oleg's paper the question is raised is there a fast way of computing the exact inverse, we then present this new method, which is called
-%  conjugate gradient method. We discuss the algorithm and the two things which bound its execution time. Which is $kappa$ and $m$ (spectral condition number and its sparsity). Then
-%  we explore both of these parameters.
-%  \end{enumerate}
-%  
-%  **************
-%  FLOW OF PAPER
-%  **************
-%  
-%  \begin{enumerate}
-%  \item We need the definition of visibilities as in Liu. DEFINE SNR HERE ALREADY together maybe with the sigma value of the noise.
-%  \item Introduce the array geometery function.
-%  \item Write down logcal and lincal in matrix form... ?
-%  \item Short discussion of least squares and jacobian and hessian's. General GN and LM update rules.
-%  \item Introduce redundant calibration as a least squares problem. We will use this general fact to derive both popular methods.
-%  \item Propose a possible solution witch leads to lincal. Mention that this approach works as in this form the function is differentiable (a taylor expansion in the 
-%  parameters exists). Maybe mention that we can also divide the problem into real and imaginary etc...
-%  \item Show how this relates to lincal for example ---> show lincal is GN.
-%  \item Now introduce complex optimization ---> Main motivation for switching to the alternative framework is that the the differentials are very simplistic. We wish to show that
-%  we can derive the method of Chengular.
-%  \item Do the derivation of Chengular. 
-%  \item Brief discussion abouth Chengular and SteFCal and Complex Optimization. Here I show that in Oleg's paper he re-derives an algorithm called SteFCal. Stefcal works
-%  on the basis of alternating direction implicit method. The first signs of achieving a similar algorithm already appeared in Stefan's conference paper in which the alternating
-%  idea was first proposed. Then I mention Chegular re-derives the expression by using partial derivatives and extends it to redundant. One could also have used the linear alternating
-%  approach. In an attempt to merge the terminology that has independently develop in the general calibration and redundant calibration literature and to emphasize the close
-%  relationship between Stefcal and the approach derived by I will use the label Redunandat SteFCal (R-Stefcal) to refer to the NLS method proposed by ... Lastly we mention that similarly
-%  to how Oleg re-derived stefcal in, we have achieved the same approach.
-%  \item Faster Exact inverse. A question that Oleg poses in his paper is, does there exist a faster way to take the exact inverse of JhJ? One that is almost linear, and
-%  implies that we can therefore implement the full LM algorithm. The aim here is of course to reduce the number of iterations that are required to converge by using the 
-%  full inverse instead which would hopefully provide enough of a speedup to compensate for the more expensive full-inverse. The algorithm we propose is the conjugate 
-%  gradient method.
-%  \item We give the images of the HESSIAN of both lincal and the complex method (number of terms). So we can mention that both are sparse and contain diagonal entries that 
-%  are more significant than the off-diagonal entries. What linear inversion approach can take advantage of both these phenomenon. One such technique is 
-%  cg. 
-%  \item Briefly discuss CG and how its computationally bounded by its condition number and its sparsity (how does the diagonal play a role).
-%  \item Simulation description
-%  \item Will CG improve things?
-%  \item Now first discuss the condition
-%  number of the Hessian before and after pre-conditioning (pre-codnitioning can only be applied if a good inverse of a matrix is known, if it is known then it can improve 
-%  the spectral condition number of a matrix. We present here the kappa and iteration number graphs for the HEXAGONAL layout. Although the
-%  \item Now we discuss the sparsity results. 
-%  \item Provide a table that theoretical compares R-StEFCal and SPARC.
-%  \item Number of outerloop iterations. 
-%  \item Timing results.
-%  \item Accuracy results.
-%  \item Maybe some freq simulations.
-%  \end{enumerate}
-%  
-%  We
-
-%{\bf GB: ok here are my comments/questions.
-%\begin{itemize}
-%\item I think we wanna start with Section 2 being a short review of standard calibration but in the complex optimization framework.
-Therefore I would begin with section 3.1, following by section 2.2.
-In particular I think that the discussion of least squares should be connected to (current) eq.
-15;
-%\item we need more context for section 3.2.
-Why do we need to introduce a different indexing?
-This should be motivated.
-As it stands, I think it should remain where it is if we can make a connection to standard calibration, otherwise it should go in section 3.3;
-%\item section 3.3 should merge with 4 where redundant calibration is introduced;
-%\item section 3.5 goes together with the description of the simulations;
-%\item section 4.1: I do not understand the definition in eq.
-32.
-How is the ratio of two matrices defined?
-I also suggest to use the symbol $\equiv$ rather than $=$ for definitions throughout the paper;
-%\item can section 4.2.2 be moved to the appendix? Or how much of it?
-%\item section 5 should be called something like "Implementation"...
-In Oleg's paper he writes "Inverting $J^H J$ and separability"...
-I find it a little too direct, but we need to find something else as these are not yet "Results".
-I also suggest to revert the order in the way this is presented: we first show that the Hessian is almost diagonal (is this only true in the complex represntation?), then we derive ADI and show that is the same as Marthi \& Chengalur;
-%\item I have not seen the simulations where you inserted a phase wrap and you calibrate it. Is there a reason why it is not included?
-%\end{itemize}
-%}
-
-% \subsection{Least Squares}
-% \label{sec:ls}
-% The method of nonlinear least squares is tantamount to solving the following optimization problem:  
-% \begin{equation}
-% \label{eq:least_squares}
-% \min_{\bzeta} \Lambda(\bzeta) = \min_{\bzeta} \|\brho\|_F^2 = \min_{\bzeta} \|\bdelta - \bnu(\bzeta)\|_F^2, 
-% \end{equation}
-% where $\Lambda$ is known as the objective function, $\brho$ is the residual vector, $\bdelta$ is the observed data vector, $\bnu$ is the model-predicted data vector and $\bzeta$ is the model-parameter vector that minimizes equation~\ref{eq:least_squares}.
-% In equation~\ref{eq:least_squares}, we assume that the optimization problem is real valued.
-% The main objective of least squares is to find the model-parameters which minimize the squared difference between the observed data and the model-predicted data.
-% 
-% The most standard way of solving equation~\ref{eq:least_squares} is to use a gradient-type minimization algorithm.
-In this paper we will be focusing on the following gradient minimization algorithms: GN and LM \citep{Levenberg1944,Marquardt1963}.
-% The GN update step is defined as
-% \begin{equation}
-%  \Delta \bzeta = (\bJ^T\bJ)^{-1} \bJ^T\brho,
-% \end{equation}
-% while the LM update takes the following form
-% \begin{equation}
-%  \Delta \bzeta = (\bJ^T\bJ+\lambda\bD)^{-1} \bJ^T\brho,
-% \end{equation}
-% where $\bJ$ is the Jacobian matrix (i.e. the derivative of the model $\bnu$ with respect to the parameter vector $\bzeta$), $\lambda$ denotes the 
-% damping factor and $\bD$ is a diagonal matrix whose diagonal entries are equal to the diagonal entries of $\bJ^T\bJ$. In this paper, we denote 
-% matrix transposition with $(*)^T$.
-% These update steps can be used in an iterative manner to determine new estimates of $\bzeta$.
-% 
-% The gradient based minimization algorithms that are generally used to solve least squares problems (i.e. GN and LM) generally require the model $\bnu$ to be differentiable
-% towards each model parameter. When the least squares problem is complex it becomes less straightforward to apply these gradient based minimization methods, as
-% many complex functions are not analytic (no Taylor expansion around a fiducial point exists) and therefore not differentiable if the classic notion of differentiation is used. For instance, if the 
-% classical definition of differentiation is used then $\frac{\partial \conj{z}}{\partial z}$, where $z \in \mathbb{C}$, does not exist.
-% 
-% The only way to circumvent the differentiability conundrum associated with complex least square problems is to recast our complex optimization problem in such a way that $\bnu$ becomes analytic in its argument.
-% The standard way of achieving this is to divide the complex optimization problem into its real and imaginary parts and then to solve for the real and imaginary parts of the parameters separately.
-% 
-% Recently, \citet{Sorber2012} proposed an alternative strategy, which involves the use of Wirtinger calculus \citep{Wirtinger1927}. The Wirtinger derivatives 
-% are defined to be 
-% \begin{align}
-% \label{eq:wir}
-% \frac{\partial}{\partial z} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} -  i \frac{\partial}{\partial y} \right ),&\frac{\partial}{\partial \conj{z}} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} +  i \frac{\partial}{\partial y} \right ).
-% \end{align}
-% Using the above definitions we can now easily compute the following:
-% \begin{align}
-% \frac{\partial z}{\partial z} & = 1, & \frac{\partial \conj{z}}{\partial z}&=0, & \frac{\partial z}{\partial \conj{z}} & = 0, & \frac{\partial \conj{z}}{\partial \conj{z}}&=1.
-% \end{align}
-% In this alternative approach we treat the complex parameters and their conjugates as separate variables.
+EoR measurements are challenging because of the intrinsic faintness of the cosmological signal \citep[see, for instance,][for recent reviews]{Furlanetto2016,McQuinn2016} buried underneath foreground emission which is a few orders of magnitude brighter than the EoR anywhere in the sky \citep[e.g.,][]{Bernardi2009,Bernardi2010,Ghosh2012,Dillon2014,Parsons2014}.
+%and that can be separated by leveraging only on the different spectral properties (citations needed!!!). 
+As miscalibrated foreground emission leads to artifacts that can jeopardize the EoR signal \citep[e.g.,][]{grobler2014,barry2016,ewall-wice2016}, an exquisite interferometric calibration is required.
+Such calibration needs an accurate knowledge of both the sky emission and the instrumental response \citep[e.g.,][]{Smirnov2011c}, unless arrays are deployed in a redundant configuration, i.e.
+with receiving elements placed on regular grids so that they measure many times the same sky brightness emission.
+Leveraging on this property, redundancy is a promising path to achieve high accurate interferometric calibration \citep[][]{Noordam1982,Wieringa1992,Pearson1982,Liu2010,Noorishad2012,Marthi2014,Sievers2017}.
+Moreover, redundant configurations provide a sensitivity boost on a selected sample of spatial scales that can be tuned to be the highest signal--to--noise ratio EoR modes \citep[][]{Parsons2012,Dillon2016}.
+These reasons motivated the design and deploment of EoR arrays in redundant configurations like the MIT-EoR \citep{Zheng2014}, the Precision Array to Probe the Epoch of Reionization \citep[PAPER,][]{Ali2015} and the Hydrogen Epoch of Reionization Array \citep[HERA,][]{deboer2017}.
+
+{\bf (TLG: Please add Pearson 1982 to bibtex.)}
+
+In this paper we present a new algorithm to calibrate redundant arrays based on the complex optimization formalism recently introduced by \cite{Smirnov2015}. 
+With respect to current algorithms, it is more robust to initial conditions, while remaining comparatively fast.
+We also show that given certain approximations this new algorithm reduces to the redundant calibration equivalent of the \textsc{StEfCal} algorithm \citep{Salvini2014}.
+%In this paper we will refer to this simplified version of the new algorithm as redundant StEfCal. 
+We investigate the speed-up that the preconditioned conjugate gradient (PCG) method provides if it is employed by the new algorithm \citep{Liu2010}.
+A comparison between the computational complexity of the optimized new algorithm and redundant StEfCal is also performed.
+
+{\bf (TLG: The following papers on redundant calibration has to be cited in the paper \citep{Noorishad2012} and Sievers2017. It was previously given in the 
+intro and was subsequently removed in the previous iteration made by GB. Unsure were they now belong. GB: let's fix this in the last round)}
+{\bf (TLG: Added Noorishad and Sievers to your red cal list.)}
+
+The paper is organized as follows: we review the complex calibration formalism by \citet{Smirnov2015} in Section~\ref{sec:sky_wirtinger}, we extend it to the redundant case in Section~\ref{sec:red_wirtinger}, we present some computational complexity results in Section~\ref{sec:pcg} and we conclude in Section~\ref{sec:conclusions}.
 
@@
 In a radio interferometer, the true sky visibilities $y_{pq}$ measured by a baseline formed by antenna $p$ and $q$ are always ``corrupted" by the non-ideal response of the receiver, which is often incorporated into a single, receiver-based complex number $g$.
-The observed visibility $d_{pq}$ is given therefore by \citep{ME1,ME2,RRIME1}
+The observed visibility $d_{pq}$ is therefore given by \citep{ME1,ME2,RRIME1}
 \begin{equation}
@@
 (q-p) + (p-1)\left (N-\frac{1}{2}p \right ) & \textrm{if}~p<q\\
-%B + (p-q) + (q-1)\left (N-\frac{1}{2}q \right )) & \textrm{if}~p>q\\
-%B+q + \frac{1}{2}(p-1)(p-2) & \textrm{if}~p>q\\
 0 & \textrm{otherwise}
@@
 \end{equation}
-%\begin{eqnarray}
-%\{1,2,\cdots,N-1,N,\cdots,B\}	 =  \nonumber\\
-%				 =  \{\alpha_{12},\alpha_{13},\cdots,\alpha_{1N},\alpha_{23},\cdots,\alpha_{N-1N}\} 
-%\end{eqnarray}
-%and 
-%\begin{equation}
-%\{1,2,\cdots,N-1,N,\cdots,B\}
-%\end{equation}
-%are equal. We denote the number of baselines in an array with $B$, i.e. $B = \frac{N^2-N}{2}$. 
 The vectors in equation~\ref{eq:vec_linear_definitions} are column vectors of size $B$ (i.e. $p<q$).
@@
 \end{equation}
-where $\Lambda$ is the objective function and $\br$ is the residual vector.
+where $\Lambda$ is the objective function, $\br$ is the residual vector and $\|\cdot\|$ denotes the Frobenius norm.
 In standard interferometric calibration, $y_{pq}$ is assumed to be known at some level, for instance through the observation of previously known calibration sources.
@@
 Gauss--Newton -- GN -- or Levenberg--Marquardt -- LM) that require the model ($\bv$ in equation~\ref{eq:least_squares}) to be differentiable towards each parameter.
-%i.e. these methods make use of a Jacobian matrix $\bJ$. In this paper we will be focusing on the following gradient-based minimization algorithms: GN and LM \citep{Levenberg1944,Marquardt1963}.
 When the least squares problem is complex, it becomes less straightforward to apply these gradient-based minimization methods, as
 many complex functions are not differentiable if the classic notion of differentiation is used, i.e. $\frac{\partial \conj{z}}{\partial z}$ does not exist if $z \in \mathbb{C}$.
-%are not analytic (no Taylor expansion around a fiducial point exists) and therefore not differentiable if the classic notion of differentiation is used.
-For instance, if the classical definition of differentiation is used then $\frac{\partial \conj{z}}{\partial z}$, where $z \in \mathbb{C}$, does not exist.
 
@@
 Assuming that $y_{pq}$ is known, equation~\ref{eq:least_squares} is recast as \citep{Smirnov2015}:
-%
-%The only way to circumvent the differentiability conundrum associated with complex least squares problems is to recast our complex optimization problem in such a way that $\bv$ becomes analytic in its argument.
-The standard way of achieving this is to divide the complex optimization problem into its real and imaginary parts and then to solve for the real and imaginary parts of the unknown model-parameters separately.
-%
-%Recently, \citet{Sorber2012} proposed an alternative strategy, which involves the use of Wirtinger calculus \citep{Wirtinger1927}. The Wirtinger derivatives are defined to be 
-%
-%\citet{Smirnov2015} recently demonstrated that the Wirtinger derivative can be used to perform normal skymodel-based calibration. We briefly summarize their approach below:
-%
 \begin{equation}
@@
 \end{align}
-%
-%
-%Note that the Wirtinger derivative is used in equation~\ref{eq:Jacobian_skymodel} and as such the entries of $\bJ$ are well defined.
 
@@
 \end{equation}
-where $\bD=\bI\odot\bJ^H\bJ$. Note the use of the Wirtinger derivatives in equation~\ref{eq:jac_entries}.
+where $\bD=\bI\odot\bJ^H\bJ$.
+In this paper, the identity matrix is denoted by $\bI$.
+Moreover, the Hadamard product and the Hermition transpose are denoted by $\odot$ and $(\cdot)^H$.
+Note the use of the Wirtinger derivatives in equation~\ref{eq:jac_entries}.
 We will refer to $\bJ^H\bJ$ as the Hessian matrix $\bH$ and to $\bJ^H\bJ + \lambda\bD$ as the modified Hessian matrix $\bmH$ throughout this paper.
@@
 \end{equation}
-
-%Note that in equation~\ref{eq:one_half} the influence that the estimate of the parameters from the previous iteration has on the current iteration is scaled by a half, while
-%in equation~\ref{eq:two_thrids} it is scaled by two thirds. Note the similarity between the last terms of equation~\ref{eq:stef_alpha} and equation~\ref{eq:one_half} when $\alpha$ is a half.
-
-%
-%The parameter vector needs to be initialized (an all one vector is one possible choice) and is updated until convergence is reached.
-To improve readability we do not explicitly denote the iteration subscript $k$ in equation~\ref{eq:GN_update_skymodel} and equation~\ref{eq:LM_update_skymodel}.
-% \subsection{Indexing Functions}
-% \label{sec:indexing}
-% In general there are two ways of indexing the true visibilities associated with different baselines.
-We can use the composite index $pq$, which is the strategy that is followed in equation~\ref{eq:vis_definition},
-% or we can use a single unique baseline index which we assign in numerical order to each composite index, assuming the composite indexes are arranged in standard baseline order.
-% The composite indexes $pq$ are in standard baseline order when they are arranged as follow: $12, 13, 1N, 23\cdots$. Let 
-% \begin{equation}
-% \alpha_{pq} =
-% \begin{cases}
-% (q-p) + (p-1)\left (N-\frac{1}{2}p \right ) & \textrm{if}~p<q\\
-% %B + (p-q) + (q-1)\left (N-\frac{1}{2}q \right )) & \textrm{if}~p>q\\
-% %B+q + \frac{1}{2}(p-1)(p-2) & \textrm{if}~p>q\\
-% 0 & \textrm{otherwise}
-% \end{cases}.
-% \end{equation}
-% We can use $\alpha_{pq}$ to map composite indexes to unique baseline indexes, i.e. 
-% \begin{equation}
-% 12,13,\cdots,1N,23,\cdots \xrightarrow{\alpha_{pq}} 1,2,3,\cdots,B.
-% \end{equation}
-% We denote the number of baselines in the array with $B$, i.e. $B = \frac{N^2-N}{2}$.
-% 
-% Recall that redundant baselines sample the exact same visibilities in the $uv$-plane, i.e. if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$. 
-% Let us now group all the baselines that measure the same $uv$-modes in a redundant array into the same redundant baseline group.
-Assigning the baselines of a redundant array to different redundant baseline groups has the advantage that it reduces the number of single index values we require.
-We now only require single index values for each redundant baseline group instead of a unique single index label for
-% each baseline. Let $\phi_{pq}$ be the function that maps composite antenna pairs to the appropriate single redundant baseline group index. Note that $\phi_{pq}$ is not symmetric as 
-% $\phi_{pq} = 0$ if $p\geq q$. We can also define the following symmetric variant of $\phi_{pq}$:
-% \begin{equation}
-% \zeta_{pq} = 
-% \begin{cases}
-% \phi_{pq}~\textrm{if}~p \leq q\\
-% \phi_{qp}~\textrm{if}~p>q
-% \end{cases}.
-% \end{equation} 
-% It is possible to construct a simple analytic expression for $\zeta_{pq}$ when our array is in an east-west regular configuration, namely $\zeta_{pq} = |q-p|$. 
-% It becomes increasingly difficult to construct analytic expressions of $\zeta_{pq}$ for more complicated array layouts. 
-% We will refer to $\zeta_{pq}$ as the symmetric geometric function. The empirically constructed symmetric geometry
-% functions of three different redundant layouts are depicted in Fig.~\ref{fig:geometry_function}. We denote the range of $\zeta_{pq}$ with $\mathcal{R}(\zeta_{pq})$. The maximal element 
-% that $\zeta_{pq}$ can ascertain is denoted by $L$ and can be interpreted as the maximal number of unique redundant baseline groups which can be formed for a given 
-% array layout.
-Alternatively, we can interpret the symmetric geometry functions in Fig.~\ref{fig:geometry_function} as geometry matrices.
-In this alternative paradigm, $\zeta_{pq}$ denotes a matrix entry.
-The dimension of these geometry
-% matrices are $N\times N$. 
-% 
-% \begin{figure*}
-% \centering
-% \subfigure[Hexagonal layout]
-% {\includegraphics[width=0.32\textwidth]{./HEX_lay.pdf}\label{fig:HEX_lay}}
-% \subfigure[Square layout]
-% {\includegraphics[width=0.32\textwidth]{./SQR_lay.pdf}\label{fig:SQR_lay}}
-% \subfigure[Regular east-west layout]
-% {\includegraphics[width=0.32\textwidth]{./REG_lay.pdf}\label{fig:REG_lay}}
-% 
-% \subfigure[Hexagonal: $\zeta_{pq}$]
-% {\includegraphics[width=0.33\textwidth]{./HEX_phi.pdf}\label{fig:HEX_phi}}
-% \subfigure[Square: $\zeta_{pq}$]
-% {\includegraphics[width=0.33\textwidth]{./SQR_phi.pdf}\label{fig:SQR_phi}}
-% \subfigure[Regular east-west: $\zeta_{pq}$]
-% {\includegraphics[width=0.33\textwidth]{./REG_phi.pdf}\label{fig:REG_phi}}
-% \caption{In the top panel we have three different redundant antenna layouts, namely a hexagonal (left), square (middle) and a regular east-west (right) layout. 
-% In the bottom panel we have the respective symmetric redundancy geometry functions for the layouts on the top panel. We used 91 antennas
-% to construct the hexagonal layout, while 100 antennas where used in the square and east-west layouts.
-In the case of the east-west layout we only plot the positions of the first ten antennas.
-The maximal amount of redundant baseline groups that can be formed for
-% the hexagonal, square and east-west layouts are 165, 180 and 99 respectively.\label{fig:geometry_function}}
-% \end{figure*}
-% 
-% If the function $\phi_{pq}$ is known and one is given one of the two antenna indexes that together form a redundant baseline as well as the redundant baseline group index itself then it is possible 
-% to calculate the unknown antenna index. This can be computed via the following two expressions:
-% \begin{equation}
-% \xi_{ij} = 
-% \begin{cases}
-% p~\textrm{if}~\exists! ~ p \in \mathbb{N} ~ s.t. ~(\phi_{pi} = j)\\
-% 0~\textrm{otherwise}
-% \end{cases},
-% \end{equation}
-% and
-% \begin{equation}
-% \psi_{ij} = 
-% \begin{cases}
-% q~\textrm{if}~\exists! ~ q \in \mathbb{N} ~ s.t. ~(\phi_{iq} = j)\\
-% 0~\textrm{otherwise}
-% \end{cases}.
-% \end{equation}
-% We use $\xi_{ij}$ to retrieve the first antenna index of the composite antenna pair associated with a particular baseline, if the index of the first antenna in the composite antenna pair and the unique redundant baseline group index are known, while we use $\psi_{ij}$ to obtain
-% a similar result: the second antenna index of the composite antenna pair. %\textbf{NB::} I still need to plot $\zeta_{ij}$ for different antenna layouts.
 
@@
 \begin{align}
-\breve{\bg}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bg}_k,\label{eq:stef_lambda} \nonumber \\
- &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bg}_k, \label{eq:stef_alpha}  
-\end{align}
-%\begin{align}
-%\breve{\bg}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bg}_k,\label{eq:stef_lambda}\\
-% &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bg}_k, \label{eq:stef_alpha}  
-%\end{align}
+\breve{\bg}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bg}_k, \nonumber \\
+ &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bg}_k, 
+\label{eq:stef_alpha}  
+\end{align}
 where $\alpha = \frac{1}{1+\lambda}$. Note that equation~\ref{eq:one_half} and equation~\ref{eq:stef_alpha} are   
@@
 
-%A special case of solutions referred to as the alternating direction implicit \citep[ADI,][]{Salvini2014} algorithm is obtained by choosing $\lambda = 0$ \citep{Smirnov2015}.
-In particular, this choice of $\lambda$ is the parameter estimation step of the \textsc{StEfCal} algorithm \citep{Mitchell:MWA-cal,Salvini2014}, the most used implementation of ADI algorithms.
-%{\bf (GB: Trienko, I am not sure about where to place the paragraph below till the end of the subsection...
-probably because I'm not sure what ``even" and ``odd" \textsc{StEfCal} steps are...
-can we add the GN step after equation~\ref{eq:stef_lambda}???)}
-%On the other hand, if we choose $\lambda$ equal to one (which corresponds
-%to $\alpha=\frac{1}{2}$) we obtain the new parameter estimation step of \textsc{StEfCal} used during even iterations.
-Note the similarity between the last terms of equation~\ref{eq:stef_alpha} and equation~\ref{eq:one_half} when $\alpha$ is a half.
-%Furthermore, the simplified skymodel-based GN equivalent of equation~\ref{eq:two_thrids} is equal to
-%\citep{Smirnov2015}
-%\begin{equation}
-%\label{eq:one_half}
-%\breve{\bg}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{1}{2}\breve{\bg}_{k}. 
-%\end{equation}
-%Note that in equation~\ref{eq:one_half} the influence that the estimate of the parameters from the previous iteration has on the current iteration is scaled by a half, while
-%in equation~\ref{eq:two_thrids} it is scaled by two thirds.
-
-%\subsection{Alternating Direction Implicit Method}
-%\label{sec:adi}
-%We start this section by briefly reviewing the ADI method as it pertains to skymodel-based calibration in Section~\ref{sec:sbc} \citep{Salvini2014,Smirnov2015}.
-%
-%In Section~\ref{sec:red_c}, we will show that the fast redundant calibration algorithm, that is presented in \citet{Marthi2014}, is in fact nothing more than an extension of the skymodel-based ADI method, presented in \citet{Salvini2014}, to the redundant calibration use case (i.e.
-redundant \textsc{StEfCal}).
-In doing so we will justify the naming convention we have been using for it.
-%
-%In Section~\ref{sec:adi_results}, we present some ADI simulation results. 
-%
-%\subsubsection{Skymodel-based Calibration}
-%\label{sec:sbc}
-%\textsc{StEfCal} is an ADI method in which the measurement equation, equation~\ref{eq:vis_definition}, is linearized by assuming that the gains are known, but that their conjugates are unknown \citep{Mitchell:MWA-cal,Salvini2014}.
-Under this assumption, the conjugates of the gains can be obtained trivially.
-Once the conjugates of the gains have been computed, it is possible to solve for the gains.
-One can then alternate between computing the gains and their conjugates until convergence is reached.
-By correctly exploiting the symmetry inherent in the calibration problem, this alternating approach can be replaced by a single iterative new parameter estimation step.
-\cite{Salvini2014} also show that the convergence of the algorithm can be improved, if for every even iteration, the basic new parameter estimation step is modified as follows: the current gain estimate is replaced by the average of the current gain estimate and the gain estimate of the previous odd iteration.
-%\textsc{StEFCal} lowers the computational complexity of normal skymodel-based calibration from $\mathbb{O}(N^3)$ to $\mathbb{O}(N^2)$. 
-%
-%\citet{Smirnov2015} recently showed that one can also derive \textsc{StEFCal} by viewing normal skymodel-based calibration as a complex optimization problem.
-They realized while inspecting the skymodel-based $\bH$ that its diagonal entries are much more significant than its off-diagonal entries.
-As we mention in Section~\ref{sec:results}, the exact same conclusion can be reached if we inspect the Hessian $\bH$ that is associated with redundant calibration (see Fig.~\ref{fig:hessian}).
-By approximating the skymodel-based $\bH$ by its diagonal and substituting this approximate Hessian $\widetilde{\bH}$ back into the fundamental skymodel-based LM update step we obtain \citep{Smirnov2015}
-%\begin{align}
-%\breve{\bg}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bg}_k,\label{eq:stef_lambda}\\
-% &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bg}_k, \label{eq:stef_alpha}  
-%\end{align}
-%where $\alpha = \frac{1}{1+\lambda}$. Moreover, $\lambda = \frac{1-\alpha}{\alpha}$.
-%
-%If we choose $\lambda$ to be zero, which corresponds to an $\alpha$ equal to one, we obtain the new parameter estimation step of \textsc{StEfCal} used during odd iterations.
-On the other hand, if we choose $\lambda$ equal to one (which corresponds to $\alpha=\frac{1}{2}$) we obtain the new parameter estimation step of \textsc{StEfCal} used during even iterations.
-%
-%Furthermore, the simplified skymodel-based GN equivalent of equation~\ref{eq:two_thrids} is equal to
-%\citep{Smirnov2015}
-%\begin{equation}
-%\label{eq:one_half}
-%\breve{\bg}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{1}{2}\breve{\bg}_{k}. 
-%\end{equation}
-%Note that in equation~\ref{eq:one_half} the influence that the estimate of the parameters from the previous iteration has on the current iteration is scaled by a half, while in equation~\ref{eq:two_thrids} it is scaled by two thirds.
-Note the similarity between the last terms of equation~\ref{eq:stef_alpha} and equation~\ref{eq:one_half} when $\alpha$ is a half.
-
 \section{Redundant Wirtinger Calibration}
@@
 if baseline $pq$ and $rs$ are redundant then $\phi_{pq} = \phi_{rs}$.
-The function $\phi_{pq}$ is not symmetric as $\phi_{pq} =
-0$ if $p \geq q$. 
-%Redundant baselines sample the exact same visibilities in the $uv$-plane, i.e.
-if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$.
-We can therefore group all the baselines that measure the exact same visibilities into the same redundant baseline group.
-We %index redundant baseline groups with single group indices instead of composite antenna pairs.
-%Let $\phi_{pq}$ be the function that maps the composite antenna pair associated with a specific baseline 
-%
-%to the single group index of the redundant baseline group to which the baseline in question belongs, i.e. if baseline $pq$ and $rs$ are redundant then $\phi_{pq} = \phi_{rs}$.  
-%The function $\phi_{pq}$ is not symmetric as 
-%$\phi_{pq} = 0$ if $p\geq q$. We can define the following symmetric variant of $\phi_{pq}$:
-%\begin{equation}
-%\zeta_{pq} = 
-%\begin{cases}
-%\phi_{pq}~\textrm{if}~p \leq q\\
-%\phi_{qp}~\textrm{if}~p>q
-%\end{cases}.
-%\end{equation} 
+The function $\phi_{pq}$ is not symmetric as $\phi_{pq} = 0$ if $p \geq q$.
 Equation~\ref{eq:vis_definition} can be re-written for a redundant array as:
-%If the array is redundant we can rewrite equation~\ref{eq:vis_definition} as
 \begin{equation}
@@
 with the same vector form as equation~\ref{eq:vis_red} if
-%equation~\ref{eq:vis_red} can also be expressed in the following vector form 
-%\begin{equation}
-%\label{eq:vis_red_vec}
-%\bd = \bv + \bn, 
-%\end{equation}
-%where 
 \begin{align}
@@
 The maximal element that $\zeta_{pq}$ can ascertain is denoted by $L$ and can be interpreted as the maximal number of unique redundant baseline groups which can be formed for a given array layout.
-%Alternatively, we can interpret the symmetric geometry functions in Fig.~\ref{fig:geometry_function} as geometry matrices.
-In this alternative paradigm, $\zeta_{pq}$ denotes a matrix entry.
-The dimension of these geometry matrices are $N\times N$.
 
@@
  \end{align}
-where the number of model parameters to be solved of is now $P = 2(N+L)$, since redundant calibration is a complex problem.
-%The number of model parameters that we need to solve for in equation~\ref{eq:least_squares_red} is $P = 2(N+L)$, since redundant calibration is a complex problem.
-
-% \subsection{Limitations}
-% \label{sec:scope}
-% We only deal with per-timeslot and frequency channel calibration. No solution smoothing  
-% is applied between the different timeslots and channels. The reader interested in this is referred to \citep{Zheng2014}. 
-% Moreover, we do not explicitly deal with any of the degeneracies normally associated with redundant calibration \citep{Zheng2014,Kurien2016}.
-Similar to the approach taken in \citep{Marthi2014}, we assume that the degeneracies present in the redundant calibration solutions are removed with a final round of skymodel-based calibration.
-
-As mentioned in Section~\ref{sec:sky_wirtinger}, equation~\ref{eq:least_squares_red} can be solved by splitting the problem into its real and
-imaginary parts and then solve for the real and imaginary parts of the unknown parameters separately \citep{Wieringa1992,Liu2010,Zheng2014}. In Appendix~\ref{sec:lincal} we review this approach.
-%As already alluded to in Section~\ref{sec:sky_wirtinger}, we can solve equation~\ref{eq:least_squares_red} in one of two ways.
-We can split the problem into its real and imaginary parts and then solve for the real and imaginary parts of the unknown parameters separately or we can employ the notion of Wirtinger derivatives.
-The former approach is discussed in Section~\ref{sec:ri}, while the latter approach is described in Section~\ref{sec:w}.
-%
-%\subsection{Real and Imaginary Approach}
-%\label{sec:ri}
-%Using least-squares to solve complex problems can be problematic as $\bv$ needs to be analytic in its argument.
-In the case of equation~\ref{eq:least_squares_red}, \citet{Liu2010} circumvents this difficulty by splitting the problem into its phase-and-amplitude components.
-%
-%The aforementioned is achieved by first rewriting equation~\ref{eq:least_squares_red} as
-%\begin{equation}
-%\label{eq:least_squares_lincal}
-%\min_{\bmath{\varrho}} \|\breve{\br}\| = \min_{\bmath{\varrho}} \|\breve{\bd} - \breve{\bv}(\bmath{\varrho})\|, 
-%\end{equation}
-%where $\breve{\bd} = [\bd^T,\conj{\bd}^T]^T$, $\breve{\bv} = [\bv^T,\conj{\bv}^T]^T$ and $\bmath{\varrho}$ is the phase and amplitude parameter vector which we define in Appendix~\ref{sec:lincal}. 
-%With the aid of equation~\ref{eq:least_squares_lincal} we can derive the algorithm implemented in \textsc{lincal} (see Appendix~\ref{sec:lincal}).
-The Jacobian associated with equation~\ref{eq:least_squares_lincal} is equal to equation~\ref{eq:Jac_lin}.
-Appendix~\ref{sec:lincal} makes it evident that \textsc{lincal} is equivalent to a basic GN update step \citep{Kurien2016}.
-%Therefore, a straightforward way of improving the convergence properties of stock standard \textsc{lincal} would be to simply add the LM damping factor to the basic \textsc{lincal} update step.
-%In doing so the need to use $\textsc{logcal}$ before $\textsc{lincal}$ becomes superfluous.
-The use of $\textsc{logcal}$ may, however, still offer a computational advantage, as it is a very fast method to obtain a good starting solution, and would reduce the number of iterations needed by the LM algorithm.
-Proper experiments would need to be conducted to determine whether it is worth keeping $\textsc{logcal}$ in the algorithmic hierarchy.
-%
-%\subsection{Wirtinger Approach}
-%\label{sec:w}
-%
-%\citet{Smirnov2015} recently proposed to use the Wirtinger derivative instead of the classic real and imaginary approach to perform normal skymodel-based calibration. We will now show
-%that the Wirtinger derivative can also be applied to redundant calibration.
-%We can perform redundant calibration by solving the following optimization problem: 
-%\begin{equation}
-%\label{eq:least_squares_red}
-%\min_{\bz} \Lambda(\bz) = \min_{\bz} \|\br\|_F^2 = \min_{\bz} \|\bd - \bv(\bz)\|_F^2, 
-%\end{equation}
-%where $\bz = [\bg,\by]^T$.
-The other vectors in equation~\ref{eq:least_squares_red} were defined in Section~\ref{} \footnote{In this paper we use the same symbol to denote corrupted predicted visibilities and true observed corrupted visibilities as it
-%is trivial to distinguish which one is implied from the context.
-Using the same symbol for these two quantities improves the readability of the paper.}.
-Note the similarity between equation~\ref{eq:least_squares} and equation~\ref{eq:least_squares_red},
-%i.e. redundant calibration can be cast as a non-linear least squares problem.
-
-We, instead, intend to formulate the redundant calibration problem using Wirtinger calculus. We therefore recast equation~\ref{eq:least_squares_red} as
-%Analogous to \cite{Smirnov2015}, we recast equation~\ref{eq:least_squares_red} to 
+where the number of model parameters to be solved for is now $P = 2(N+L)$, since redundant calibration is a complex problem.
+
+In literature, equation~\ref{eq:least_squares_red} is solved by splitting the problem into its real and imaginary parts.
+The real and imaginary parts of the unknown parameters are then solved separately \citep{Wieringa1992,Liu2010,Zheng2014}.
+Currently, the above is achieved by using the real-valued GN algorithm \citep{Kurien2016}. 
+We, instead, intend to formulate the redundant calibration problem using Wirtinger calculus and recast equation~\ref{eq:least_squares_red} as
 \begin{equation}
@@
 where $\breve{\bz} = [\bz^T,\conj{\bz}^T]^T$. 
-%Equation~\ref{eq:least_squares_complex} is the complex conjugate augmented counterpart of equation~\ref{eq:least_squares_red}.
-%The fact that $\bv$ is now analytic in its argument has the immediate consequence: it becomes possible to apply complex GN
-%and complex LM to equation~\ref{eq:least_squares_complex}. 
 We derive the complex Jacobian associated with equation~\ref{eq:least_squares_complex} to be:
-%\subsubsection{Complex GN and LM}
-%\label{sec:complex_GN_LM}
-%The complex Jacobian associated with equation~\ref{eq:least_squares_complex} is equal to
 \begin{equation}
@@
 \end{equation}
-We can now calculate the GN and LM update steps to be 
-%The GN update step is now defined as:
+We can now calculate the GN and LM updates to be 
 \begin{equation}
@@
 \Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\br}
-%\Delta \breve{\bz} &=& (\bJ^H\bJ + \lambda\bD)^{-1}\bJ^H\breve{\br},
-\end{equation}
-and 
-%The LM update rule is defined in a very similar way:
+\end{equation}
+and 
 \begin{equation}
@@
 respectively. As in Section~\ref{sec:sky_wirtinger}, equation~\ref{eq:GN_update} can be used to iteratively update our parameter vector:
-%we can now use equation~\ref{eq:GN_update} or equation~\ref{eq:LM_update} to iteratively update our parameter vector as follows:
 \begin{equation}
@@
 The analytic expressions for $\bJ$, $\bH$ and $\bJ^H\breve{\br}$ can be found in Appendix~\ref{sec:analytic}.
-
-In the case of the GN algorithm we can simplify equation~\ref{eq:update} even further to obtain
+Appendix~\ref{sec:analytic} also contain two useful identities, involving $\bJ$ and $\bH$.
+
+In the case of the GN algorithm we can simplify equation~\ref{eq:update} even further. Replacing $\breve{\br}$ with $\breve{\bd}-\breve{\bv}$ in equation~\ref{eq:GN_update} results in
+\begin{equation}
+\label{eq:temp_GN_eq}
+\Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H(\breve{\bd}-\breve{\bv}), 
+\end{equation}
+If we substitute the first identity of equation~\ref{eq:identities} into equation~\ref{eq:temp_GN_eq} and we simplify the 
+result we obtain 
+\begin{equation}
+\label{eq:one_thrid}
+\Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd}-\frac{1}{3}\breve{\bz}.
+\end{equation}
+If we use the above simplified update in equation~\ref{eq:update} it reduces to
 \begin{equation}
@@
 \end{equation}
-The above equation is derived in Appendix~\ref{sec:simplify_GN}.
 Equation~\ref{eq:two_thrids} is the redundant equivalent of equation~\ref{eq:one_half} and it shows us that in the case of redundant calibration we can calculate the GN parameter update step without calculating the residual.
@@
 \label{eq:y_update}
-y_{i}^{k+1} = \alpha \frac{\sum_{pq \in \mathcal{PQ}_i} \conj{g}_p^k g_q^k d_{pq}}{\sum_{pq \in \mathcal{PQ}_i}|g_p^k|^2|g_q^k|^2} + (1-\alpha) y_i^k. 
-\end{equation}
-The computational complexity of inverting $\widetilde{\bH}$ is $\mathbb{O}(P)$.
-We note that equation~\ref{eq:g_update} is the gain estimator associated with \textsc{StEfCal} and that equation~\ref{eq:g_update} and~\ref{eq:y_update} correspond to the parameter estimators derived by \citet{Marthi2014}.
-In \citet{Marthi2014}, equation~\ref{eq:g_update} and~\ref{eq:y_update} are derived by taking the
-derivative of the objective function $\Lambda$ relative to the elements of $\bg$ and $\by$, setting the intermediate results to zero and then solving
-for the unknown parameters.
-The above two equations can also be derived using the ADI method.
-Please see Appendix~\ref{sec:red_stef_ADI} for the details.
-We will refer to the above parameter update steps as redundant \textsc{StEFCal} in the remainder of the
-paper.
-
-%{\bf (GB: derived in which way? It would e useful to say in order to differentiate from us)}.
-
-The choice of $\alpha$ is somewhat unconstrained. 
-In this paper we chose $\alpha=\frac{1}{3}$ (i.e. $\lambda = 2$) and we based our decision on the choice of $\alpha$ in the 
-case of classic \textsc{StEFCal}. In the case of \textsc{StEFCal}, $\alpha = \frac{1}{2}$ (for even iterations),
-which results in the last term of equation~\ref{eq:one_half} and equation~\ref{eq:stef_alpha} being equal. 
+y_{i}^{k+1} = \alpha \frac{\sum_{rs \in \mathcal{RS}_i} \conj{g}_r^k g_s^k d_{rs}}{\sum_{rs \in \mathcal{RS}_i}|g_r^k|^2|g_s^k|^2} + (1-\alpha) y_i^k. 
+\end{equation}
+The the index set $\mathcal{RS}_i$ and the quantity $\widetilde{y}_{ij}$ is defined in equation~\ref{eq:RS} and equation~\ref{eq:y_tilde}, respectively. 
+The computational complexity of inverting $\widetilde{\bH}$ is $\mathbb{O}(P)$. We note that equation~\ref{eq:g_update} is the gain estimator associated with \textsc{StEfCal}.
+
+Equation~\ref{eq:g_update} and~\ref{eq:y_update} were obtained by \citet{Marthi2014} by taking the derivative of the objective function $\Lambda$ relative to the elements of $\bg$ and $\by$, setting the intermediate results to zero and then solving for the unknown parameters (i.e.
+using the gradient descent algorithm).
+We note that their derivation is less general. The LM algorithm has better convergence properties than gradient descent and 
+encompasses the gradient descent algorithm as a special case.
+In Appendix~\ref{sec:red_stef_ADI} we show that equation~\ref{eq:g_update} and~\ref{eq:y_update} can also be derived using the ADI method.
+For this reason, we will refer to the pseudo LM calibration scheme, derived above (equation~\ref{eq:g_update} and~\ref{eq:y_update}), as redundant \textsc{StEFCal}.
+{\bf (GB: this needs a better wording)}.
+{\bf (TLG: rewrote hope it helps. If you wish to (significantly) rewrite the above paragraph please let me know before doing so, so that we 
+do not go around in circles.)}.
+
+The choice of the $\alpha$ parameter is somewhat unconstrained (i.e. \citet{Marthi2014} chose it to be equal to $0.3$) {\bf (GB: what is the choice of $\alpha$ in \citet{Marthi2014}?)}
+{\bf (TLG: added.)}. 
+In this paper we chose $\alpha$ by adopting the same strategy that is used in \textsc{StEfCal},
+i.e. we chose $\alpha$ to be equal to a $\frac{1}{3}$ ($\lambda = 2$). We also carried out simulations to validate this choice.
+%In the case of \textsc{StEFCal}, $\alpha = \frac{1}{2}$ (for even iterations), which results in the last term of equation~\ref{eq:one_half} and equation~\ref{eq:stef_alpha} being equal.
 When $\alpha=\frac{1}{3}$, the last term of equation~\ref{eq:two_thrids} and equation~\ref{eq:alpha} are equal.
 
-%In this paper we follow the same choice done in \textsc{StEFCal}, leading to $\alpha = 1/3$.
-The following simulations show this choice to be appropriate {\bf (GB: is this true?
-TRIENKO, I MANAGED TO REACH ONLY TILL THIS POINT.
-I THINK IT IS TIME I HAND THIS BACK TO YOU AND YOU MAKE AN EFFORT TO TIE IT UP.
-i THINK WE CAN GO FROM HERE TO A MINOR REVISION NEXT TIME)}
-
-%equation~\ref{eq:g_update} and equation~\ref{eq:y_update} are the exact same iterative new parameter estimators derived by \citet{Marthi2014}.
-Also note that equation~\ref{eq:g_update} is the fundamental gain estimator associated with skymodel-based \textsc{StEfCal}.
-We have, therefore, successfully used the diagonal approximation approach proposed by \citet{Smirnov2015} to re-derive the fast redundant calibration method presented in  \citet{Marthi2014}.
-%The question now arises what value should we use for $\alpha$? In \citet{Marthi2014}, they propose 
-%$\alpha = 0.3$. Although an exact mathematical proof for the optimal choice of $\alpha$ still eludes us, it is still possible to
-%venture a clever guess by studying what the choice of $\alpha$ is in the case of skymodel-based \textsc{StEFCal}. The value of $\alpha$ is equal to a half 
-%in the case of \textsc{StEFCal} (for even iterations); see Section~\ref{sec:sbc}.
-If we wish to mimic this behaviour for the redundant calibration use case we need to scale $\breve{\bz}$ by two thirds in equation~\ref{eq:alpha} (see equation~\ref{eq:two_thrids}) which implies that we should choose $\alpha$ equal to a third, which corresponds to a $\lambda$ equal to two.
-%We used this $\lambda$ value in this paper. The following simulation results show that this choice produces 
-%good results.
-
-Optimality results of redundant \textsc{StEFCal} were presented in \citet{Marthi2014}.
-We do not repeat their analysis here.
-We do, however, present our own redundant \textsc{StEFCal} results, in Fig.~\ref{fig:prec_error}, which confirm that the $\alpha$ value proposed
-in this paper produce accurate parameter estimates. To quantify the exact error in our estimates we used the percentage error $\beta$ defined 
-as 
-\begin{equation}
-\label{eq:beta}
-\beta = \frac{\|\bv - \widehat{\bv}\|_F^2}{\|\bv\|_F^2}.
-\end{equation}
-We have already defined $\bv$ in equation~\ref{eq:vec_definitions}. Moreover, in equation~\ref{eq:beta}, $\widehat{\bv}$ denotes the parameter estimate obtained with redundant \textsc{StEFCal}.
-
-\begin{figure}
-\includegraphics[width=0.47\textwidth]{./prec_error.pdf} 
-\caption{We plot the percentage error $\beta$ between the estimated 
-predicted visibilities and the true predicted visibilities for different SNR values as a function of the number of antennas in the array $N$.
-The graph presented here was generated using the first experimental setup in Table~\ref{tab:ch_parm}.}
-\label{fig:prec_error}
-\end{figure}
-
-\section{Preconditioned Conjugate Gradient Method}
-\label{sec:pcg}
-In this section we investigate the computational complexity of the PCG method when applied to the redundant calibration problem. 
-We briefly review the PCG method in Appendix~\ref{sec:conj_grad}. If the reader is not that familiar with the PCG method we encourage 
-the reader to briefly review Appendix~\ref{sec:conj_grad} before continuing with the paper. The PCG method can be used to efficiently invert Hermitian sparse 
-matrices, which is exactly the kind of matrix $\bH$ is (see Fig.~\ref{fig:hessian} and Eq.~\ref{eq:red_H}).
-There are two factors which influence the computational complexity of the PCG method, namely the spectral condition number $\kappa$ and the sparsity of the matrix we wish to invert (see equation~\ref{eq:cg_bound}).
-
-We investigate the spectral condition number and the sparsity of the modified Hessian $\bmH$ (i.e.
-$\lambda = 2$), which is a non-singular matrix, in Section~\ref{sec:scn} and Section~\ref{sec:sparsity} respectively.
-It is interesting to note, that the Hessian $\bH$ itself is singular (from empirical tests) and that the PCG method 
-can in principal be be used to pseudo-invert it (see Appendix~\ref{sec:normal}).
-
-\subsection{Spectral Condition Number}
-\label{sec:scn}
-The convergence properties of the conjugate gradient (CG) algorithm can be dramatically improved through preconditioning. Inspecting Fig.~\ref{fig:hessian}, reveals that the Jacobian preconditioner
-is a good candidate for the redundant use case. In Fig.~\ref{fig:kappa}, we see that this choice of preconditioner performs very well, the value of the spectral condition number $\kappa$
-becomes constant and independent of problem size, i.e. the number of antennas in the array. Preconditioning therefore effectively eliminates $\kappa$ from equation~\ref{eq:cg_bound}. This result
-is confirmed by Fig.~\ref{fig:cg_itr}.
-Fig.~\ref{fig:cg_itr} shows us that the number of major iterations needed by the preconditioned conjugate gradient method to invert $\bmH$ is independent of the number of antennas in the array.
-Moreover, the number of major iterations required is also much smaller than the dimension of $\bmH$.
-
-%kappa + itr plots
-\begin{figure*}
-\centering
-\subfigure[$\kappa$]{\includegraphics[width=0.47\textwidth]{./kappa.pdf}\label{fig:kappa}}
-% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
-\subfigure[Iterations required before and after Preconditioning]{\includegraphics[width=0.47\textwidth]{./cg_itr.pdf}\label{fig:cg_itr}}
-\caption{The graphs presented here were generated using the second experimental setup in Table~\ref{tab:ch_parm}.
-In the left panel we have plotted the spectral condition number $\kappa$ of the modified Hessian $\bmH$, before (magenta and red curves) and after preconditioning (blue curve), at different SNR values as a function of the
-number of antennas $N$ in the array.
-The right hand panel confirms this result.
-In the right hand panel we have plotted the number of major iterations (line 6 of Algorithm \ref{algo:CG}) required by the conjugate
-gradient method to invert $\bmH$, before (magenta and red curves) and after preconditioning (blue and green curves), at different SNR values as a function of the number of antennas $N$ in the array.
-\label{fig:kappa_itr}} 
-\end{figure*}
-
-\subsection{Sparsity}
-\label{sec:sparsity}
-% \begin{enumerate}
-% \item $\boldsymbol{H}$: is a $(4N-2)\times(4N-2)$ or a $P \times P$ matrix. $N$ denotes the number of antennas and $P$ denotes the number of parameters. This matrix
-% contains $6N^2 - 2N - 2$ non-zero entries. It contains $16N^2 - 16N + 4$ entries. We can construct this matrix with $4N^2 -  4N$ or $\frac{1}{4} P^2 -1$ elementary operations.
-% \item $\boldsymbol{A}$: is a $(2N-1)\times(2N-1)$ matrix. This matrix has $N^2 + N - 1$ non-zero entries. We can construct this matrix with $2\frac{1}{2} (N^2 -  N)$ elementary operations.  
-% \item $\boldsymbol{B}$: is a $(2N-1)\times(2N-1)$ matrix. This matrix has $2N^2 - 2N$ non-zero entries. We can construct this matrix with $1\frac{1}{2} (N^2 -  N)$ elementary operations.
-% \item $\boldsymbol{C}$: is a $N\times N$ matrix. This matrix has $N$ non-zero entries. We can construct this matrix with $N^2-N$ elementary operations.
-% \item $\boldsymbol{D}$: is a $N \times (N-1)$ matrix. This matrix has $\frac{1}{2} (N^2 -  N)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations. 
-% \item $\boldsymbol{E}$: is a $(N-1) \times (N-1)$ matrix. This matrix has $(N-1)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.  
-% \item $\boldsymbol{F}$: is a $N \times N$ matrix. This matrix has $N^2 - N$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.
-% \item $\boldsymbol{G}$: is a $N \times (N-1)$ matrix. This matrix has $\frac{1}{2} (N^2 -  N)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.
-% \item $\boldsymbol{0}$: is a $(N-1) \times (N-1)$ all zero matrix. 
-% \end{enumerate}
-
-%We can construct $\boldsymbol{C}-\boldsymbol{G}$ in $O(N^2)$
-
-%The second factor that determines the computational complexity of the conjugate gradient method is the number of non-zero
-%entries in the modified Hessian matrix $\bmH$. 
-%We have also shown in section that the first factor, the spectral condition number of $\bH$, can be reduced 
-%to a constant by employing preconditioning and can therefore be completely eliminated from the computational complexity equation. 
-A useful matrix quantity which we will use in this section is the sparsity factor $\gamma$ of a matrix, which is the ratio between the number of zero entries and the total number of entries in a matrix.
-We
-can therefore define $\gamma$ as follows:
-\begin{equation}
- \gamma = \left (1 - \frac{m}{P^2} \right ) 
-\end{equation}
-It is trivial to determine an analytic expression for the sparsity ratio of $\bmH$ when our redundant array is in a regular east-west geometric configuration
-The sparsity and the asymptotic sparsity ratio of $\bmH$ are equal to
-\begin{align}
-\gamma &= \frac{5N^2-7N+3}{8N^2-8N+2} & \gamma_{\infty} &= \lim_{N\rightarrow \infty}\gamma = \frac{5}{8} \label{eq:gamma}. 
-\end{align}
-However when our redundant array makes use of a more complicated geometric layout it becomes much more difficult to construct such expressions, as $\bmH$ itself 
-becomes much harder to express analytically.
-
-In Fig.~\ref{fig:gamma}, we plot the empirically determined sparsity ratios for three different geometric layouts as a function of the number antennas in the array.
-Although, we cannot explicitly calculate
-the limit of the hexagonal and square layout curves, Fig.~\ref{fig:gamma} seems to indicate that they do exist. Moreover, we also plot the left most equation
-of equation~\ref{eq:gamma} (red dashed line) and the right most equation of equation~\ref{eq:gamma} (black dotted curve) in Fig.~\ref{fig:gamma}.
-
-We can now use the sparsity factor to compute the order of the computational complexity of inverting $\bmH$ (assuming the effect of $\kappa$ is negligible) with
-\begin{equation}
-P^{c} = (1 - \gamma)P^2.
-\end{equation}
-Solving for $c$ we obtain
-\begin{align}
-c &= \log_{P}(1 - \gamma) + 2 & c_{\infty} &= \lim_{N\rightarrow \infty} c = 2. \label{eq:c}
-\end{align}
-
-The evaluation of the limit in the rightmost equation of equation~\ref{eq:c} is discussed below.
-Note that
-\begin{equation}
-\label{eq:int_limit}
-\lim_{N\rightarrow\infty} P^{\log_{P}(1-\gamma)} = 1-\gamma_{\infty},
-\end{equation}
-which is only possible if 
-\begin{equation}
-\lim_{N\rightarrow\infty} \log_{P}(1-\gamma) = 0,
-\end{equation}
-implying that $c_{\infty}$ is equal to two.
-
-equation~\ref{eq:int_limit}, implies that we can only compute $c_{\infty}$ in the rightmost equation if $\lim_{N\rightarrow \infty} \gamma$ exists. 
-In Fig.~\ref{fig:c}, we plot the empirically determined value of $c$ for three different geometric layouts as a function of the number antennas in the array.
-Furthermore, in Fig.~\ref{fig:c}, the analytic value of $c$ for the regular east-west grid is plotted as a dashed red curve, while the limit of this analytic curve is plotted with a dotted black curve.
-
-Equation~\ref{eq:gamma}, equation~\ref{eq:c} and Fig.~\ref{fig:sparsity} indicate that the computational complexity, under the assumption that $\lim_{N\rightarrow \infty} \gamma$ exists, is asymptotically bounded by $O(P^2)$.
-However, the computational complexity converges very slowly to its asymptotic value, and in general is equal to $O(P^{c})$, where $c \sim 1.7$ if a Hexagonal geometric layout containing less than 200 antennas is employed.
-
-\begin{figure*}
-\centering
-\subfigure[$\gamma$]{\includegraphics[width=0.47\textwidth]{./sparsity.pdf}\label{fig:gamma}}
-% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
-\subfigure[$c$]{\includegraphics[width=0.46\textwidth]{./comp_order.pdf}\label{fig:c}}
-\caption{In the left panel we plot the sparsity ratio $\gamma$ (the number of zero entries relative to the total number of entries in a matrix) of the modified Hessian $\bmH$ as a function of the number
-of antennas $N$ in the array for different geometries, namely an hexagonal grid (blue circles), a square grid (green crosses) and a regular east-west grid (red circles).
-For the regular east-west grid we are able to construct an analytical expression for $\gamma$ (the dashed red line).
-Taking the limit
-of this analytic expression produces the black dotted line. In the right panel we plot the order of the computational cost $c$ for inverting $\bmH$ as a function of the number 
-of antennas in the array for different geometries. We use the same color scheme in the right and left panels. The red dashed line 
-in the right panel is associated with the theoretical curve of $c$ we computed for the east-west regular grid. The limit of the red-dashed curve is two, which we 
-have plotted with a black dotted line.
-\label{fig:sparsity}} 
-\end{figure*}
-
-\section{Comparison Results}
-\label{sec:comparison}
-We are now in a position to make a final comparison between redundant \textsc{StEFCal} and the PCG method.
-Inverting $\bmH$ by assuming it is a diagonal matrix (i.e.
-using redundant \textsc{StEFCal}) is computationally inexpensive, while inverting $\bmH$ using PCG is more computationally taxing.
-On the other hand, one inversion with PCG is much more
-accurate than using an approximate inverse.
-The question is now, if the inverse using the PCG method is much more accurate will the LM implementation using it require fewer iterations than redundant \textsc{StEFCal} requires.
-Moreover, will the number
-of LM iterations PCG requires be so few that it manages to outperform redundant \textsc{StEFCal} in the end.
-To answer this we need to compute the number of approximate Hessian inversions $k$, redundant \textsc{StEFCal} computes, in the same time it takes PCG to perform a single Hessian inversion.
-We can compute this with
-\begin{align}
-\label{eq:k} 
- kP &\approx P^{1.7}, & k &\approx P^{0.7}.
-\end{align}
-Next we need to measure the number of LM iterations we save by using the PCG method and compare it to equation~\ref{eq:k}. We make this comparison 
-in Fig.~\ref{fig:out_diff}. Fig.~\ref{fig:out_diff} tells us that redundant \textsc{StEFCal} will outperform the PCG method as the number of LM iterations we save by taking the 
-full inverse is not enough to exceed $k$. In this comparison we have only taken into account the cost of inverting $\bmH$. The PCG method, however, is still 
-useful in practice as it would be more straightforward to apply it to other calibration use cases.
-
-\begin{figure*}
-\centering
-\subfigure[Number of LM iterations required by redundant \textsc{StEFCal} and PCG.]{\includegraphics[width=0.47\textwidth]{./outerloop.pdf}\label{fig:outerloop}}
-% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
-\subfigure[Difference.]{\includegraphics[width=0.47\textwidth]{./diff.pdf}\label{fig:diff}}
-\caption{The graphs presented here were generated using the second experimental setup in Table~\ref{tab:ch_parm}. In the left panel we have the number of LM iterations 
-that is required by redundant \textsc{StEFCal} (green and blue curve) and the PCG (magenta and red curve) method to converge to a parameter error tolerance of 1e-6 whilst using different 
-SNR values as a function of the number of antennas $N$ in the array.
-In the right panel we have the average amount of LM iterations (difference between the redundant \textsc{StEFCal} and PCG curves in the left panel) we save (green and blue curve) by computing the full-inverse with the PCG method whilst using different SNR values.
-The black curve is associated with $k$, the number of LM iterations redundant \textsc{StEFCal} can perform in the same time it takes the LM implementation using 
-the PCG method to perform one iteration.
-\label{fig:out_diff}} 
-\end{figure*}
-
-%\subsubsection{Simplifying GN}
-
-% The 
-% normal skymodel-based equivalent of equation~\ref{eq:two_thrids} is equal to \citep{Smirnov2015}:
-% \begin{equation}
-% \label{eq:one_half}
-% \breve{\bg}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{1}{2}\breve{\bg}_{k}, 
-% \end{equation}
-% where $\breve{\bg} = [\bg^T,\conj{\bg}^T]^T$
-% 
-% FIX TRANSPOSE PROBLEM
-
-%\section{Comparing ADI and PCG}
-%\section{Implementation} 
-% \section{Redundant \textsc{StEfCal}}
-% \label{sec:results}
-% In Section~\ref{sec:red_wirtinger} we discussed the complex GN and LM algorithms and how they relate to redundant calibration. If we inspect equation~\ref{eq:GN_update} and 
-% equation~\ref{eq:LM_update} we see that the most expensive operation used by the GN and LM algorithms is the inversion of the Hessian, $\bH$. The most logical 
-% way of speeding up the algorithm is to investigate whether we can do this inversion in some efficient way.
-% 
-% Two recent algorithms have been proposed that are able to invert $\bH$ in an efficient manner, namely the ADI \citep{Marthi2014} and PCG \citep{Liu2010} methods. 
-% Note that here we focus solely on comparing the computational complexity of inverting $\bH$ using either one of these two methods. Rating the LM or GN algorithms
-% using one of these approaches relative to each other once the computational complexity of inverting $\bH$ is known is trivial. The simulations we performed to compare these two 
-% methods are summarized in Section~\ref{sec:simulations}.
-% 
-% %We restrict ourselves to comparing the most 
-% %expensive operation of the two methods namely, computing the inverse of the Hessian $\bH$ (or modified Hessian $\bmH$ in the case of LM). It becomes trivial to rate these two 
-% %algorithms relative to each other once the computational complexity of the inverse operation is known. 
-% 
-% The basic difference between these two speedup algorithms can be understood by inspecting Fig.~\ref{fig:hessian}. The first observation we can make by inspecting 
-% Fig.~\ref{fig:hessian_reg} and Fig.~\ref{fig:hessian_hex} is that the diagonal entries of $\bH$ are more significant than the off-diagonal entries in both cases.
-% It therefore makes sense to approximate the $\bH$ with its diagonal, which greatly simplifies the computational complexity of inverting 
-% $\bH$, i.e. it becomes $\mathbb{O}(P)$\footnote{The cost of constructing 
-% the diagonal is not included in this estimate.} \citep{Smirnov2015}. This diagonal-approximation approach is effectively ADI \citep{Marthi2014}. The exact details of why this 
-% is the case is discussed in Section~\ref{sec:adi}.
-We will also show how \citet{Marthi2014}, \citet{Smirnov2015} and \citet{Salvini2014} relate to each other in that same section.
-It should be noted here that strictly speaking the ADI method actually refers to the complete LM update step using the approximate $\bH$ and not just
-% to the inversion of the approximate $\bH$.
-To improve the readability of the paper, we use the acronym ADI to refer to both the faster approximate inverse approach and the complete LM algorithm that uses it.
-% Which one of these two concepts we are actually referring to, however, is always perfectly clear from the context.
-% %Furhtermore, the name ADI also implies that with this algorithm we alternate between computing the 
-% %gains, their conjugates and the true observed visibilities. %In Appendix~\ref{sec:sbc} we briefly summarize the ADI method that is used for skymodel-based calibration.
-% 
-% The second observation we can make from Fig.~\ref{fig:hessian_reg} and Fig.~\ref{fig:hessian_hex} is that both Hessians contain many zero entries, i.e.
-they are sparse matrices.
-This, in conjunction with the more significant diagonal entries, implies that the PCG method (presented in Section~\ref{sec:conj_grad})
-% can be applied to invert them in a computationally efficient manner.
-\citet{Liu2010} proposed the PCG method to compute the inverse of $\bH$, however the computational complexity of this approach was not analyzed therein.
-% In Section~\ref{sec:pcg}, we do a proper computational complexity analysis of PCG. We find that the computational complexity of PCG is asymptotically bounded by $\mathbb{O}(P^2)$, which is 
-% a lot better than straightforward inversion which takes $\mathbb{O}(P^3)$. 
-% 
-% 
-% 
-% %Since we approximate the Hessian in the ADI method with its diagonal the computational complexity of its inverse becomes $\mathbb{O}(P)$\footnote{The cost of constructing 
-% %the diagonal is not included in this estimate.}. 
-% 
-% Even though the PCG method is computationally more expensive than the ADI method we cannot arrive at a final conclusion at this point. With the ADI method 
-% we only approximate the inverse of the Hessian, while in the case of the PCG method we compute the exact inverse. So the question arises whether the LM algorithm using the PCG method 
-% will converge much faster than the ADI method even though each iteration (which includes an inversion) is more expensive. We investigate this question in Section~\ref{sec:comparison}.
-%
-
-% At this point there is no way of knowing if this is mere coincidence. DEF REWRITE THIS.
-% 
-% REWRTIE THIS IN A DIFFERENT WAY. START WITH CHENGULAR AND JUST MENTION THE LINK TO STEFCAL'S CHOICE OF ALPHA.
-% 
-% The update rules in equation~\ref{eq:lambda} and equation~\ref{eq:alpha} have been derived before, by Chengular... The approach they followed is discussed in the Appendix.
-% 
-% Interestingly enough, if we substitute $\alpha = 1$ into Eq.~\ref{eq:g_update} then it reduces to the skymodel-based StEFCal update rule. In Chengular they propose 
-% an $\alpha = 0.3$.
-% 
-% In an attempt to merge the terminology that has independently develop in the general calibration and redundant calibration literature and to emphasize the close
-% relationship between Stefcal and the approach derived by I will use the label Redunandat SteFCal (R-Stefcal) to refer to the NLS method proposed by ... 
-% 
-% ****\\
-% 
-% with $\alpha = \frac{1}{1+\lambda}$. Moreover $\lambda = \frac{1-\alpha}{\alpha}$. So in Marthi and Chengular (2014) they use an $\alpha$ of 0.3, which corresponds to an LM damping factor $\lambda$
-% of $2\frac{1}{3}$.
-% 
-% NB:: NEED TO ADD SNR PLOTS HERE
-% PLUS IMAGE OF HESSIAN FOR THE DISCUSSION
-
-\section{Conclusion}
-\label{sec:conclusions}
-In Section~\ref{sec:red_wirtinger} we extended \textsc{StEfCal} to the redundant use case by using the Wirtinger formalism 
-presented in \citet{Smirnov2015} (i.e. we presented redundant \textsc{StEfCal}). We also pointed out that redundant \textsc{StEfCal} was 
-first derived by \citet{Marthi2014} under a different name using a different approach than the one used in this paper. In Appendix~\ref{sec:red_stef_ADI} we showed that redundant \textsc{StEfCal}
-can also be derived using the ADI method.
-
-\citet{Liu2010} suggested to use the PCG method to speed-up redundant calibration.
-In Section~\ref{sec:pcg} we analyze the computational complexity of the PCG method when applied to the redundant calibration problem.
-In Section~\ref{sec:comparison} we do a comparison between redundant \textsc{StEfCal} and the PCG method. We found 
-that although the PCG method does greatly imporove the speed of redundant calibration it does not outperform redundant \textsc{StEfCal}. 
-% fiducial initial guess 
-% 
-% The following partial derivatives are now easily computed
-% 
-% \begin{eqnarray}
-% \frac{c_{ij}}{\partial \eta_p} &=& y_{i-j} e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}\\ 
-% \frac{c_{ij}}{\partial \varphi_p} &=&  -i y_{i-j} e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}\\
-% \frac{c_{ij}}{\partial \eta_q} &=& y_{i-j} e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}\\ 
-% \frac{c_{ij}}{\partial \varphi_q} &=&  i y_{i-j} e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}\\
-% \frac{y_{i-j}}{\partial \varphi_q} &=&  e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}
-% \end{eqnarray}
-% 
-% 
-% The Wirtinger derivative is used in the last equation.
-% 
-% \begin{eqnarray}
-% c_{ij} &\approx& c_{ij}^0 + \Delta \eta_p(y_{i-j}^0 e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0}) + \Delta \eta_q(y_{i-j}^0 e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0})\\
-% && -i\Delta\varphi_p (y_{i-j}^0e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0}) +i\Delta\varphi_q (y_{i-j}^0e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0})\\
-% && y_{i-j}^1(y_{i-j}^0e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0})\\
-% &\approx&  c_{ij}^0 + e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_p+ \Delta \eta_q - i\Delta\varphi_p + i\Delta\varphi_p))
-% \end{eqnarray}
-% 
-% \begin{equation}
-% \label{eq:residual}
-% r_{ij} = \delta_{ij} = c_{ij}-c_{ij}^0 = e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_p+ \Delta \eta_q - i\Delta\varphi_p + i\Delta\varphi_p)) 
-% \end{equation}
-% 
-% Eq.~\ref{eq:residual} allows us to construct the following linear system:
-% 
-% \begin{equation}
-% \boldsymbol{J}[\boldsymbol{\Delta \eta},\boldsymbol{\Delta \varphi},\boldsymbol{\Delta y}]^T = \boldsymbol{r}, 
-% \end{equation}
-% 
-% where $\boldsymbol{J}$ is equal to 
-% \begin{equation}
-% \boldsymbol{J} = \bigg [\overbrace{\frac{c_{pq}}{\partial \eta_p}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial \varphi_p}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial y_{t}}}^{t=1\cdots r_s} \bigg ] \bigg \} [pq] = 1\cdots N_{\textrm{b}} (p<q)
-% \end{equation}
-% or
-% \begin{equation}
-% \boldsymbol{J} = \bigg [\frac{c_{pq}}{\partial \boldsymbol{\eta}},~\frac{c_{pq}}{\partial \boldsymbol{\varphi}},~\frac{c_{pq}}{\partial \boldsymbol{y}} \bigg ] \bigg \} [pq] = 1\cdots N_{\textrm{b}} (p<q).
-% \end{equation}
-% 
-% The last column is again a Wirtinger derivative.
-% 
-% Which means we can obtain the least-squares estimate as follows:
-% 
-% \begin{equation}
-% [\boldsymbol{\Delta \eta},\boldsymbol{\Delta \varphi},\boldsymbol{\Delta y}]^T = [\boldsymbol{J}^H\boldsymbol{J}]^{-1}\boldsymbol{J}^H\boldsymbol{r}.
-% \end{equation}
-
-\bibliographystyle{mn2e}
-\bibliography{paperG2}
-
-\appendix
-
-\section{Simulations}
-\label{sec:simulations}
-\begin{table*}
-\centering
-\caption{The following fixed simulation parameters were used to generate the results of this paper. We use $U$ to represent 
-a uniform distribution.}
-\begin{tabular}{|c c c c c c c|} 
-\hline
-Declination & Latitude & Minimum baseline & Geometry& Sources & Flux & Spatial \\
-\hline 
-\hline
- $-74^{\circ}39'37.481''$ & $-30^{\circ}43'17.34''$ & 20 & Hexagonal &100 & Pareto (shape param. of 2) & $U[-3,3]$ (deg.) 
-\end{tabular}
-\label{tab:fixed_parm}
-\end{table*}
+We generated a sky model that comprises of 100 flat spectrum sources distributed over a 3$^{\circ}$ by 3$^{\circ}$ sky 
+patch centred. The flux density of each source was drawn from a Pareto (power law) distribution 
+with a shape parameter of 2. The chosen shape parameter ensures that it is unlikely that a source, that is more than 20 Jy, is added to the simulation sky model. 
+{\bf (GB: we need some form of motivation for the distribution choice and its parameter)}) 
+{\bf (TLG: this is a bit of an ad-hoc choice, I knew that the flux distribution follows some power-law so a settled on a Pareto distribution. I played around 
+with $a$ until it produced flux values I deemed reasonable. This distribution choice stems from the suppression work I did with Ridhima.
+It be noted, however, that it is the SNR value which influence the results of calibration in general (the SNR of 
+your problem determines how well you can calibrate). Whatever power is in your model is irrelevant, it 
+is the ratio of model power versus noise power which matters. The SNR 
+is a better reflection of what you are doing. Which is why I define it explicitly. Maybe you can word it better.)})
+and the source position drawn from a uniform distribution. 
+We also made use of multiple fictitious telescope layouts each one having a hexagonal geometry (see the left upper image of Figure~\ref{fig:geometry_function} for an example layout).
+The largest hexagonal distributed array
+that we used contained 217 antennas. The minimum baseline of this array was 20 m and its maximum baseline was 320 m.
+{\bf GB (Figure~\ref{????}, with a minimum and a maximum baseline of 20 and ???? m respectively)}. 
+{\bf TLG (Added)}
+
+We "corrupted" visibilities by applying gain errors (Table~\ref{tab:gain_parm}) and calibrated the corrupted visibilities using the redundant \textsc{StEfCal} algorithm.
+Solutions were independently derived for time-step and channel.
+We used 5 independent gain realizations. {\bf (GB: How many different realizations of the gains????)}
+{\bf (TLG: Recall, that we only used 5 as CG and SVD were extremly slow. Moreover, in reality we also have many more ``realizations'' as 
+each channel and time-step can be seen to have a unique ``random gain'' associated with it. Added.)}
 
@@
 performed our simulations over multiple time-slots and one frequency channel or one timeslot and multiple frequencey channels (see Table~\ref{tab:ch_parm}).
-Moreover, $c$ in the left most column denotes the speed of light.}
+Moreover, $c$ in the left most column denotes the speed of light.
+We adopted sinusoidal error models in this paper as they are often used in simulation packages like MeqTrees 
+for the functional testing of algorithms. We also made use of a more physical model, a phase ramp, which is the form 
+that the phase of a banpass solution often takes in practice.
+It is worth pointing out the following.
+In the first model, we model a gain error with amplitude around one and an additional phase error.
+In
+the second model we chose $A$ in such a way that we do not unintentionally produce a zero gain value. For the third model 
+the value of $\tau$ was chosen so that the amount of phase wraps that occur across the observing band is restricted to a realistic number.
+{\bf (GB: we need to describe in more detail the choice of models as well as the choice of parameter ranges - if there is any rational behind them.)}
+{\bf (TLG: Added.)}}
 \begin{tabular}{|c c c c|} 
@@
 \centering
-\caption{We generated results using two major experimental setups. We either used one frequency channel and multiple time-slots, or one time-slot and multiple frequency 
-channels. The most important parameters used in realizing these two major setups are presented here. Note that we refer to the gain error models in Table~\ref{tab:gain_parm} in the 
-last row of this table.}
+\caption{We generated results using two main setups.
+We either used one frequency channel and multiple time-slots, or one time-slot and multiple frequency channels.
+The most important parameters used in realizing these two major setups are presented here.
+%Note that we refer to the gain error models in the last row of Table~\ref{tab:gain_parm}.
+We chose the observational frequency band of setup 1 to coincide with the HERA array. To broaden the scope of our 
+analysis we chose the observational frequency of our second setup to be equal to 1.4~GHz, which is a typical observing frequency of the Westerbork Synthesis Radio Telescope (WSRT). 
+{\bf (GB: we need an explanation as to why we chose 1.4~GHz.)}
+{\bf (TLG: This is enough of a motivation I think. I do not explicitly mention that WSRT is highly redundant as I think it
+is general knowledge.)}}
 \begin{tabular}{|c c c|} 
@@
 %Hour angle &-1^h & [-1^h,1^h]\\
-Gain error model (Table~\ref{tab:gain_parm}) & 2 & 1\\
+%Gain error model (Table~\ref{tab:gain_parm}) & 2 & 1\\
 %Initial guess: $\bz_0$ & $\bone$ & $\bone$\\
@@
 \end{table}
-%The main purpose of the simulations conducted in this paper is to genrate experimental results with which we can compare the computational complexity of ADI and PCG.
-We used an hexagonal geometric layout with an adjustable number of rings during the
-%simulations. 
-In this section we describe the simulations we performed to generate the results we report in the paper. We first discuss some 
-of the simulation parameters we used followed by the general flow of the simulations that were conducted.
-
-For the skymodel we generated 100 flat spectrum sources.
-Each source had a random flux value and position.
-The flux of the sources followed a Pareto distribution (with a shape parameter of 2), while their positions were uniformly
-distributed in a 3$^{\circ}$ by 3$^{\circ}$ square grid around the field center. The remaining fixed simulation parameters can all be found in Table~\ref{tab:fixed_parm}. 
-We also employed three different gain error models which we summarize in Table~\ref{tab:gain_parm}. We found that it did not matter which gain error model we used during
-our simulations, as the model we used did not have an impact on the results.
-Moreover, we had two major experimental setups, we either made use of simulated observations consisting of one frequency channel and multiple time-slots or one time-slot and multiple channels.
-We describe these two experimental setups in more detail in Table~\ref{tab:ch_parm}. As was the case for the gain error models, the experimental setup 
-that was used did not alter the simulation results.
-
-The main flow of a single simulation is as follows.
-We created corrupted visibilities using the fixed simulation parameters presented in Table~\ref{tab:fixed_parm} and either one of the experimental setups summarized in Table~\ref{tab:ch_parm}.
-We then used per time-slot and channel redundant calibration, i.e.
-we use the LM algorithm, which in turn makes use of either the ADI or PCG method, to
-estimate the gains and the true observed visibilities.
-Our initial parameter guess $\bz_0$ was chosen to be an all one vector.
-To obtain statistically significant results we performed, for both the ADI and PCG methods, multiple statistical independent simulations.
-
-Furthermore, the following SNR definition is used in this paper  
+
+The simulated visibilities included a noise term. The purpose of which is to produce a certain signal--to-noise ratio (SNR) in the simulated
+visibilities. We made use of the following definition of \citep[SNR, ][]{Liu2010,Marthi2014}:  
 \begin{equation}
@@
 \end{equation}
-where $<*>_{\nu,t,pq}$ denotes averaging over frequency, time and baseline. The definition we use here is based on the SNR definitions used in \citet{Liu2010} and \citet{Marthi2014}.
+where $<*>_{\nu,t,pq}$ denotes averaging over frequency, time and baseline.
+{\bf (GB; if $n$ is the standard noise term, is the denominator just the noise variance? We need to specify a little more here...)}.
+{\bf (TLG; Yes it is the power in the noise. The average power in signal divided by the average noise power. Recall, I confirmed this 
+expression with Adrian Liu. I do not know what you wish to add here. I am just trying to say that my simulations are proper, I do not use a relative quantity to 
+conduct them. This is the same SNR definition as was used in previous redundant cal papers.)}.
+
+Figure~\ref{fig:prec_error} show the simulation results as a function of SNR and number of antennas. The accuracy of our solutions is quantified through the percentage error
+\begin{equation}
+\label{eq:beta}
+\beta = \frac{\|\bv - \widehat{\bv}\|_F^2}{\|\bv\|_F^2},
+\end{equation}
+where $\widehat{\bv}$ is the redundant \textsc{StEFCal} parameter estimate.
+
+The error magnitude follow the expected behaviour, i.e., it decreases as a function of SNR and number of antennas $N$.
+Interestingly, it reduces to a few percent when $N > 120$ for essentially any choice of SNR.
+
+\citet{Marthi2014} conducted an even more in depth analysis than the one presented here. They found that the total computational cost of redundant StEfCal
+is $\mathbb{O}(P^2)$ and that it approaches the Cram{\'e}r-Rao Bound (CRB) at moderate SNRs. 
+{\bf (GB: is there a similar analysis in \cite{Liu2010} a/o \cite{Marthi2014}? If so, what do they find?)}.
+{\bf (TLG: added.)}
+
+\begin{figure}
+\includegraphics[width=0.47\textwidth]{./prec_error.pdf} 
+\caption{We plot the percentage error $\beta$ between the simulated visibilitied and the visibilities solved for by \textsc{StEFCal} for different SNR values as a function of the number of antennas in the array $N$.
+{\bf (GB: I don't undesrtand the curve with SNR = -1...
+The yaxis should read $\beta$).}}
+\label{fig:prec_error}
+\end{figure}
+
+\section{Preconditioned Conjugate Gradient Method}
+\label{sec:pcg}
+\citet{Liu2010} suggested that the speed of convergence of redundant calibration could be optimized using the 
+conjugate gradient method \citep{Hestenes1952}, which would be computationally advantageous, since the Hessian matrix 
+is sparse \citep{Reid1971}. In this section we study the computational complexity
+of the conjugate gradient method when used to invert the Hessian matrix in equation~\ref{eq:LM_update}, 
+in particular through the use of preconditioning (i.e. the preconditioned conjugate gradient (PCG) method). 
+%(see also Appendix~\ref{sec:conj_grad}).
+
+The computational complexity of CG is  
+\begin{equation}
+\label{eq:cg_bound}
+\mathbb{O}(\sqrt{\kappa}m), 
+\end{equation}
+where $m$ denotes the number of non-zero entries and $\kappa$ denotes the spectral 
+condition number of the matrix which is to be inverted. The spectral condition number $\kappa$ of the matrix $\bA$ is defined as:
+\begin{equation}
+\label{eq:kappa}
+\kappa(\bA) = \frac{\iota_{\textrm{max}}}{\iota_{\textrm{min}}}, 
+\end{equation}
+where $\iota_{\textrm{max}}$ and $\iota_{\textrm{min}}$ respectively denote the largest and the smallest eigenvalue of $\bA$.
+
+Preconditioning is a technique used to improve the spectral condition number of a matrix. Let us consider a generic system of linear equations 
+\begin{equation}
+\label{eq:prec}
+\bA\bx = \bb,
+\end{equation}
+and a positive-definite Hermitian matrix $\bM$ so that 
+\begin{equation}
+\label{eq:prec}
+\bM^{-1}\bA\bx = \bM^{-1}\bb.
+\end{equation}
+
+The matrix $\bM$ is a good preconditioner if:
+\begin{equation}
+\label{eq:kappa}
+\kappa(\bM^{-1}\bA) \ll \kappa(\bA),
+\end{equation}
+i.e., if it lowers lowers the condition number of a matrix.
+If $\bA$ is a nearly diagonal matrix, the Jacobian preconditioner is a natural choice of $\bM$ and can be computed as:
+\begin{equation}
+\bM = \bA\odot\bI, 
+\end{equation}
+where $\bI$ is the identity matrix.
+
+%The most expensive operation ni the PCG algorithm is the vector-matrix product $\bA\bx$, which has a complexity of $\mathbb{O}(m)$, where $m$ is the number of non-zero entries in $\bA$.
+
+In order to quantify the effectiveness of the CG method in redundant calibration, we investigate the spectral condition number and the sparsity of the modified Hessian $\bmH$ (i.e. $\lambda = 2$).
+{\bf (GB: does $\lambda$ indicate eigenvalues?)}
+{\bf (TLG: No, it denotes the LM damping factor, replaced eigenvalues here with $\iota$)}
+
+We generated simulated (i.e. corrupted) visibilities as described in Section~\ref{sec:red_wirtinger}.
+
+{\bf (GB: we may have to add some text to explain between the two differential setups which I may be a little confused myself with...)}
+
+{\bf (TLG1: Not sure what you are asking here. If you are asking what is the difference between red stefcal, PCG and CG I tried to provide more clarity in the text, however, if you 
+are referring to the different simulation setups please see comment TLG2.)}
+
+{\bf (TLG2: Setup 2 was the setup I used initially.
+The parameters I used was influenced by arrays I was accustomed to at that point, like WSRT.
+Later on as you became more involved you suggested setup 1 (and even later still gain model 3) as it is more
+realistic for redundant low-frequency arrays (redoing the red stefcal sims with these parameters did not influence the results though if memory serves me right -- not sure if I still have the old results though to confirm).
+We never redid the sims with PCG as the sims took quite a bit of time (much
+longer than the unvectorized red stefcal, which as we know takes quite a long time) and there 
+would ultimately be now difference in result, the calibration algorithm's speed does not depend on the observational frequency etc... Moreover, 
+Fig. 5 is completely independent of the chosen simulated parameters as it depends only on the analytic expression of the Hessian. Not sure 
+how to phrase or explain the different setups in the paper, as I would like to avoid redoing the simulations at this stage, except if the referee actually requests this.)}
+
+To calibrate, we used the complex LM-algorithm described in Section~\ref{sec:red_wirtinger}. To invert the 
+Hessian we used the CG method, with and without a preconditioner. We made use of the Jacobian preconditioner. 
+Figure~\ref{fig:kappa} shows us that preconditioning reduces the condition number of $\bmH$ to a small constant value and therefore effectively eliminates
+it from equation~\ref{eq:cg_bound}, i.e. equation~\ref{eq:cg_bound} reduces to
+\begin{equation}
+\label{eq:cg_bound2}
+\mathbb{O}(m). 
+\end{equation}
+This result is confirmed by Fig.~\ref{fig:cg_itr}. Fig.~\ref{fig:cg_itr} 
+shows us that the number of major iterations needed by the preconditioned conjugate gradient method to invert 
+$\bmH$ is independent of the number of antennas in the array and that it is much less than the 
+dimension of $\bmH$. 
+%We found that if we used a preconditioner then the spectral condition number $\kappa$ of $\bmH$ improves and becomes completely independent of the number of antennas in the array (see Figure~\ref{fig:kappa_itr})
+
+{\bf - (GB: but it is also much smaller, right?)}.
+{\bf - (TLG: The CG method is normally good to use on classes of matrix problems whose spectral condition number 
+is independent of problem size (this is an ideal property to have to be able to use the 
+CG method). In the problem we are studying here, it is the case even without preconditioning. However, 
+the condition number is high without preconditioning which unfortunately leads to a large number of major iterations (albeit fairly independent 
+of problem size). If we require more than $P$ major iterations, then there is no point to using CG as conventional algorithms will 
+outperform it in any case (here we require more than $P$ major iterations). So here preconditioning is not required to ensure independence of problem size (which is 
+generally the case), but to lower the condition number, which in turn reduces the number of major iterations required to be less than $P$ in a non-trivial way. 
+The relationship which exists between $\kappa$ and the number of major iterations needed depends on starting conditions and the eigenvalue distribution among other things. 
+Since both (preconditioning and not preconditioning) are constant we need to plot both $\kappa$ (this is only an indication of the order of the problem and does not 
+indicate the exact number of iterations due to the complicated relationship of the two) and the number of major iterations used. Because $\kappa$ is constant it does not nec tell 
+you everything about the number of major iterations).
+I think all of this becomes to complicated to explain in the paper and almost irrelevant to the idea i am trying to convey (kappa is constant implies number of major
+iterations is independent of problem size. Emprirical sims show it to be less than $P$. Which is why I try to focus on the basics, hope this makes more sense now?)}
+
+%, which is a non-singular matrix, in Section~\ref{sec:scn} and Section~\ref{sec:sparsity} respectively.
+%It is interesting to note, that the Hessian $\bH$ itself is singular (from empirical tests) and that the PCG method can in principal be be used to pseudo-invert it (see Appendix~\ref{sec:normal}).
+
+\begin{figure*}
+\centering
+\subfigure[$\kappa$]{\includegraphics[width=0.47\textwidth]{./kappa.pdf}\label{fig:kappa}}
+% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
+\subfigure[Iterations required before and after Preconditioning]{\includegraphics[width=0.47\textwidth]{./cg_itr.pdf}\label{fig:cg_itr}}
+\caption{The graphs presented here were generated using the second experimental setup in Table~\ref{tab:ch_parm}.
+Left: spectral condition number $\kappa$ of the modified Hessian $\bmH$as a function of $N$, before (magenta and red curves) and after (blue curve) preconditioning, at different SNR values.
+Right: number of major iterations (line 6 of Algorithm \ref{algo:CG}) required by the conjugate
+gradient method to invert $\bmH$ as a function of the number of antennas $N$ in the array, before (magenta and red curves) and after preconditioning (blue and green curves), at different SNR values.
+Both plots show that the Jacobian preconditioner definitely speed up the conjugate gradient method {\bf(GB: the yaxis label of the right panel should just be "number of iterations"...)}
+\label{fig:kappa_itr}} 
+\end{figure*}
+
+{\bf (GB: the following paragraph may need a better polish up...)}
+{\bf (GB: tried cleaning up the paragraph...)}
+
+Let us now shift our attention towards the remaining factor $m$.
+To get a usefull idea of the computational complexity of CG, we must relate $m$ and $P$.
+This can be achieved by utilizing the modified Hessian's measure of sparsity $\gamma$:
+\begin{equation}
+ \gamma = \left (1 - \frac{m}{P^2} \right ),
+\end{equation}
+where $m$ is the number of non-zero entries in $\bmH$  and $P^2$ is total number of matrix elements.
+
+For a regular east-west geometric configuration the sparsity and the asymptotic sparsity of $\bmH$ can be derived analytically:
+\begin{align}
+\gamma &= \frac{5N^2-7N+3}{8N^2-8N+2} & \gamma_{\infty} &= \lim_{N\rightarrow \infty}\gamma = \frac{5}{8} \label{eq:gamma}. 
+\end{align}
+For more complicated array layouts, however, there is no straightforward analytical solution and we empirically determined the sparsity ratios for three different geometric layouts as a function of the number antennas in the array (see Figure~\ref{fig:gamma}).
+
+It now follows that 
+{\bf (GB: can't follow why it is this way...)}
+{\bf (TLG: Rewrote this to emphasize the link between $m$ and $P$, the computational complexity is only useful, if it 
+is expressed in terms of the number of parameters one whises to solve for. Multiplying the $1-$ sparsity measure with $P^2$ gives you $m$. Moreover,
+since it is smaller than one, this tells us that we are doing better than $P^2$. The reduction in complexity stems from sparse vector matrix multiplication.)}
+\begin{equation}
+P^{c} = m = (1 - \gamma)P^2,
+\end{equation}
+which leads to
+\begin{align}
+c &= \log_{P}(1 - \gamma) + 2 & c_{\infty} &= \lim_{N\rightarrow \infty} c = 2. \label{eq:c}
+\end{align}
+The computational complexity is therefore asymptotically bounded by $O(P^2)$, although it converges very slowly to its asymptotic value, and in general is equal to $O(P^{c})$, where $c \sim 1.7$ in the case of an hexagonal geometric layout with $N < 200$ (see Figure~\ref{fig:c}).
+
+\begin{figure*}
+\centering
+\subfigure[$\gamma$]{\includegraphics[width=0.47\textwidth]{./sparsity.pdf}\label{fig:gamma}}
+% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
+\subfigure[$c$]{\includegraphics[width=0.46\textwidth]{./comp_order.pdf}\label{fig:c}}
+\caption{Left: the sparsity ratio $\gamma$ of the modified Hessian $\bmH$ as a function of the number of antennas $N$ for an hexagonal (blue circles), square (green crosses) and regular east-west grid (red circles) array geometry.
+The red--dashed and black--dotted lines show the analytical expression for $\gamma$ and its limit for the regular east-west grid case (see text for details).
+Right: the order of the computational cost $c$ for inverting $\bmH$ as a function of $N$ for different array geometries (same colour scheme as in the left panel).
+The red--dashed and black--dotttedlines are the analytical expression of $c$ and its limit in the east-west regular grid case {\bf (GB: I would just leave $\gamma$ and $c$ as labels for the y axis as they are defined in the text)}.
+\label{fig:sparsity}} 
+\end{figure*}
+
+We are now finally able to compare the computational complexity of redundant \textsc{StEFCal} and the PCG method.
+Redundant \textsc{StEFCal} is computationally inexpensive as it just needs to invert a diagonal matrix, however, the PCG inversion is accurate and, therefore, may require fewer iterations, ultimately leading to a faster convergence.
+We computed the theoretical number of redundant \textsc{StEFCal} iterations $\Delta k$, in excess
+of the number of LM (implementing PCG) iterations required, for 
+LM to outperform redundant \textsc{StEFCal}
+{\bf (GB: I don't follow the following... perhaps I am missing something.)}
+{\bf (TLG: The argument was logically flawed improved it during this iteration.)}
+\begin{align}
+\label{eq:k} 
+ (k+\Delta k)P >& kP^c, & \Delta k >& k(P^{c-1}-1),
+\end{align}
+and compared it to the empirically obtained average excess number of iterations.
+The results are displayed in Figure~\ref{fig:out_diff}.
+Figure~\ref{fig:out_diff} shows that redundant \textsc{StEFCal} outperforms the PCG method.
+In this comparison we have only taken into account the cost of inverting $\bmH$.
+The PCG method is still
+useful in practice as it would be more straightforward to apply it to other calibration use cases
+
+{\bf (GB: what do you mean with this statements? Which cases?)}.
+{\bf (TLG: I was hoping that Jonathan would have published his work by now. In it he shows that the pointing error 
+formulation has off-diagonal significant elements. We should prob just remove this sentence as I can not reference anything 
+as of yet.)}
+
+\begin{figure*}
+\centering
+\subfigure[Number of LM iterations required by redundant \textsc{StEFCal} and PCG.]{\includegraphics[width=0.47\textwidth]{./outerloop.pdf}\label{fig:outerloop}}
+% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
+\subfigure[Difference.]{\includegraphics[width=0.45\textwidth]{./diff.pdf}\label{fig:diff}}
+\caption{The graphs presented here were generated using the second experimental setup in Table~\ref{tab:ch_parm}.
+Left: number of LM iterations required by redundant \textsc{StEFCal} (green and blue curves) and the PCG (magenta and red curves) methods to converge to a parameter error tolerance of $10^{-6}$ whilst using different SNR values as a function of the number of antennas $N$ in the array {\bf (GB: which configuration?)}.
+Right: average amount of LM iterations (difference between the redundant \textsc{StEFCal} and PCG curves in the left panel) saved (green and blue curve) by computing the full-inverse with the PCG method whilst using different SNR values.
+The black curve is $\Delta k$, the theoretical number
+of redundant \textsc{StEFCal} iterations that are needed in excess of the number of LM iterations used for the LM algorithm to 
+outperform redundant \textsc{StEFCal}. For the black curve we assumed $c=1.7$ and that $k$ could be approximated with the 
+the magenta curve plotted in the left panel.
+\label{fig:out_diff}} 
+\end{figure*}
+
+\section{Conclusion}
+\label{sec:conclusions}
+In this paper we have formulated the calibration of redundant interferometric arrays using the complex optimization formalism \citep{Smirnov2015}.
+We have derived the related Jacobian and updated parameter steps and showed that this approach naturally leads to solving the calibration equations in the fast way used by alternating direction implicit methods \citep[][]{Salvini2014,Marthi2014}.
+The code (redundant \textsc{StEfCal}) is publicly available at ?????.
+We note that, in its current implementation, redundant \textsc{StEfCal} does not solve for the degeneracies inherent to redundant calibration  \citep{Zheng2014,Kurien2016} which will be the subject of future work.
+Compared to current redundant calibration algorithms, redundant \textsc{StEfCal} is more robust to inital conditions and allows for an easier, massive parallelization.
+
+We investigated the computational cost of redundant \textsc{StEfCal} and compared it with the performance of the preconditioned conjugated method suggested by \cite{Liu2010} to solve the calibration equations.
+We found that, although the preconditioned conjugated method method greatly improves the speed of redundant calibration, it still significant underperforms redundant \textsc{StEfCal}.
+
+The characteristics of \textsc{StEfCal} makes it an appealing calibration algorithm for large redundant arrays like HERA \citep{deboer2017}, CHIME (cit!!!!), HIRAX (cit!!!!) or even hybrid arrays like the MWA (cit!!!!).
+
+\section*{Acknowledgements}
+{\bf (GB: you may need to add an acknowledgement to the group grant that supported you last year, pllease, ask Oleg)} G.B.
+acknowledges support from the Royal Society and the Newton Fund under grant NA150184.
+This work is based on the research supported in part by the National Research Foundation of South Africa (grant No.
+103424).
+This research is supported by the South African Research Chairs Initiative of the Department of Science and Technology and National Research Foundation
+
+\bibliographystyle{mn2e}
+\bibliography{paper}
+
+\appendix
 
@@
 \section{Conjugate Gradient Method}
-\label{sec:conj_grad}
-The CG method was independently discovered by \citet{Hestenes1973} and \citet{Stiefel1952}. They later published a joint paper, which is now considered as the seminal
-reference on CG \citep{Hestenes1952}. The iterative CG method is used to solve a particular class of linear system. The CG method is generally used to solve
-\begin{equation}
-\label{eq:linear_system}
-\bb = \bA\bx,
-\end{equation}
-where $\bb\in\mathbb{C}^P$ is a known complex column vector of size $P$, $\bx\in\mathbb{C}^P$ is an unknown complex column vector with the same length as $\bb$, and $\bA\in\mathbb{C}^{P\times P}$ is a square positive-definite Hermitian matrix with the appropriate dimensions.
-Moreover, using the CG method to solve equation~\ref{eq:linear_system} will only be computationally advantageous if $\bA$ is also sparse (contain many zero entries).
-Using the iterative CG method to solve large sparse linear sytems was popularized by \citet{Reid1971}. A good tutorial on the CG method can be found in \citep{Shewchuk1994}.
-The iterative CG method is presented in Algorithm \ref{algo:CG}.
-
-%If $\bA$ is a square positive semi-definite Hermitian matrix, then the conjugate gradient method only converges if $\bb$ is in the column range of $\bA$ (cite Lu2016).
-
-%\begin{algorithm}
-%\caption{Conjugate Gradient Method.
-Inputs: $\bA$, $\bb$, a starting value for $\bx$, maximum number of iterations $k_{\textrm{max}}$ and an error tolerance $\epsilon<1$ Output: $\bx$ the solution to equation~\ref{eq:linear_system}.
-\citep{Shewchuk1994}.}\label{algo:CG}
-%\begin{algorithmic}[1]
-%\State $k \gets 0$
-%\State $\br \gets \bb - \bA\bx$
-%\State $\bp \gets \br$
-%\State $\delta_{\textrm{new}} \gets \br^H\br$
-%\State $\delta_0 \gets \delta_{\textrm{new}}$
-%\While{$k\leq k_{\textrm{max}}$ and $\delta_{\textrm{new}} > \epsilon^2\delta_0$}
-%\State $\alpha \gets \frac{\delta_{\textrm{new}}}{\bp^H\bA\bp}$
-%\State $\bx \gets \bx + \alpha\bp$
-%\State $\br \gets \br - \alpha\bA\bp$
-%\State $\delta_{\textrm{old}} \gets \delta_{\textrm{new}}$ 
-%\State $\delta_{\textrm{new}} \gets \br^H\br$
-%\State $\beta \gets \frac{\delta_{\textrm{new}}}{\delta_{\textrm{old}}}$
-%\State $\bp \gets \br + \beta\bp$
-%\State $k \gets k + 1$
-%\EndWhile
-%\end{algorithmic}
-%\end{algorithm}
-
-\subsection{Computational Complexity}
-\citet{Kaniel1966} was able to determine convergence bounds for CG using Chebyshev polynomials. A much more in depth analysis of CG convergence can be found in
-\citep{Sluis1986}. As can be seen from Algorithm \ref{algo:CG}, the CG method makes use of a major outer loop (line 6 of Algorithm \ref{algo:CG}). It turns out that at worst 
-the number of major iterations that CG requires to converge\footnote{If the $\bA$-norm is used to calculate the error.} is proportional to $\sqrt{\kappa}$, where $\kappa$ is the spectral condition number of the matrix $\bA$.
-The mathematical definition of $\kappa(\bA)$ is
-\begin{equation}
-\label{eq:kappa}
-\kappa(\bA) = \frac{\lambda_{\textrm{max}}}{\lambda_{\textrm{min}}}, 
-\end{equation}
-where $\lambda_{\textrm{max}}$ and $\lambda_{\textrm{min}}$ respectively denote the largest and the smallest eigenvalue of $\bA$.
-Moreover, the most expensive operation that is conducted within each major loop of Algorithm~\ref{algo:CG} is the vector-matrix product $\bA\bp$, which has a complexity of $\mathbb{O}(m)$, where $m$ is the number of non-zero entries in $\bA$.
-The computational complexity of CG is therefore 
-\begin{equation}
-\label{eq:cg_bound}
-\mathbb{O}(\sqrt{\kappa}m). 
-\end{equation}
-It should now also be clear to the reader why the CG method is especially applicable to sparse linear systems. In sparse linear systems $m < P^2$.
-
-\subsection{Preconditioning}
-\label{sec:precon}
-Preconditioning is a technique which allows us to improve the condition number of a matrix. Let $\bM$ be a positive-definite Hermitian matrix.
-We can now indirectly solve for $\bx$ by using 
-\begin{equation}
-\label{eq:prec}
-\bM^{-1}\bA\bx = \bM^{-1}\bb.
-\end{equation}
-If $\bM$ is a good preconditioner then $\kappa(\bM^{-1}\bA) \ll \kappa(\bA)$, i.e. a good preconditioner lowers the condition number of a matrix which in turn increases the convergence speed 
-of CG (see equation~\ref{eq:cg_bound}).
-In this paper, we denote matrix inversion with $(*)^{-1}$.
-The best possible choice of $\bM$ would therefore be $\bA$, which would lead to $\kappa$ becoming unity (as the largest and smallest eigenvalue of the identity matrix are equal to one).
-If $\kappa$ is one then equation~\ref{eq:cg_bound} simplifies and becomes $\mathbb{O}(m)$.
-However, if we knew the inverse of $\bA$ then obtaining $\bx$ would be trivial and there would be no need for the CG method in the first place.
-In short, a good preconditioner $\bM$ must approximate $\bA$ fairly accurately. However, if it is not much less expensive to 
-compute the inverse of $\bM$ than that of $\bA$ we actually gain nothing (see equation~\ref{eq:prec}).
-
-If the diagonal entries of $\bA$ are much more significant than its off-diagonal entries then the 
-Jacobian preconditioner is a good choice of $\bM$. The Jacobian preconditioner is computed as follows:
-\begin{equation}
-\bM = \bA\odot\bI, 
-\end{equation}
-where $\bI$ is the identity matrix and $\odot$ denotes the Hadamard product. Since the Jacobian preconditioner is a diagonal matrix its inverse can be obtained trivially.
 
@@
 
-\section{Redundant Wirtinger Calibration - $\bJ$, $\bH$ and $\bJ^H\breve{\br}$}
+\section{Derivation of redundant Wirtinger calibration}
 \label{sec:analytic}
@@
 
-%\begin{align}
-%\left [ \bM_1 \right ]_{\alpha_{pq},j} &= y_{\phi_{pq}}\conj{g_q}\delta_p^j, & \left [ \bM_2 \right ]_{\alpha_{pq},j} &= g_p\conj{g_q}\delta_{\phi_{pq}}^j 
-%\end{align}
-%\begin{equation}
-%\left [ \bN_1 \right ]_{\alpha_{pq},j} = g_py_{\phi_{pq}}\delta^j_q. 
-%\end{equation}
 We use $\bzero$ to denote an all zero matrix. It is now trivial to compute the Hessian $\bH$ by using equation~\ref{eq:Jacobian_red}. If we substitute equation~\ref{eq:Jacobian_red} into $\bJ^H\bJ$
@@
 \begin{equation}
+\label{eq:RS}
 \mathcal{RS}_i = \left\{rs\in\mathbb{N}^2|(\phi_{rs} = i) \right\},
@@
 \begin{equation}
+\label{eq:y_tilde}
 \widetilde{y}_{ik} = 
@@
 
-% \begin{equation}
-% \boldsymbol{J}^H\breve{\boldsymbol{d}} = 
-% \begin{bmatrix}
-% \sum_{k\neq i } g_k x_{ik}d_{ik}\\
-% \sum_{pq \in \mathcal{I}_j} \conj{g}_p g_q d_{pq}\\
-% \downarrow^{*}
-% \end{bmatrix}
-% \begin{matrix}% matrix for right braces 
-% \coolrightbrace{\sum g_k x_{\zeta_{ik}}d_{ik}}{i = 1\cdots N}\\
-% \coolrightbrace{\sum \conj{g}_p g_q d_{pq}}{j = 1\cdots L}\\
-% \vphantom{\downarrow^{*}}
-% \end{matrix}
-% \end{equation}
+The following identities,
+\begin{align}
+\label{eq:identities}
+\frac{1}{3}\bJ\breve{\bz} &= \breve{\bv}, & \bJ^H\breve{\bv} &= (\bI\odot\bH)\breve{\bz} 
+\end{align}
+can now be trivially established by mechanically showing that the left hand side of each identity in equation~\ref{eq:identities} is equal to its right hand side.
+In equation~\ref{eq:identities}.
 
@@
 
-\section{Simplifying GN}
-\label{sec:simplify_GN}
-The following identities,
-\begin{align}
-\label{eq:identities}
-\frac{1}{3}\bJ\breve{\bz} &= \breve{\bv}, & \bJ^H\breve{\bv} &= \widetilde{\bH}\breve{\bz} 
-\end{align}
-are trivially established by mechanically showing that the left hand side of each identity in equation~\ref{eq:identities} is equal to its right hand side.
-In equation~\ref{eq:identities}, $\widetilde{\bH}$ denotes $\bI\odot\bH$.
-
-Replacing $\breve{\br}$ with $\breve{\bd}-\breve{\bv}$ in equation~\ref{eq:GN_update} results in
-\begin{equation}
-\label{eq:temp_GN_eq}
-\Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H(\breve{\bd}-\breve{\bv}), 
-\end{equation}
-Substituting the first identity of equation~\ref{eq:identities} into equation~\ref{eq:temp_GN_eq} and simplifying the result 
-\begin{equation}
-\label{eq:one_thrid}
-\Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd}-\frac{1}{3}\breve{\bz}.
-\end{equation}
-
 \section{ADI}
@@
 
-%We compare ADI with PCG in Section \textbf{??}, by using the results generated from the simulations we performed using 
-%these two methods.
-
-%The results reported in this paper were generated using two major experimental setups, namely a single time-slot and multiple frequency channels setup or 
-%a single frquency channel and multiple time-slots setup. The simulation parameters which where kept fixed in both these setups can be found in 
-%Table~\ref{tab:fixed_parm}. The sky-model we used to simulate the visibilities were the same for both setups.   We used a hexagonal redundant layout for our antenna positions. We produced 
-%results for different
-
 \label{lastpage}

### 2017-11-01 software/plot_func.py

@@ -728,7 +728,7 @@ def plot_kappa_itr_pres():
     plt.grid('on')
     plt.show()
 
-def plot_k():
+def plot_k_old():
     
     #N = np.array([7,19,37,61,91,108,169,217])
     matplotlib.rcParams.update({'font.size': 22})
@@ -773,6 +773,66 @@ def plot_k():
     plt.plot(N,k,lw=2.0,label='Theory',c='k')
     plt.plot(N1,diff1,lw=2.0,label='SNR=1000',c='b')
     plt.plot(N2,diff2,lw=2.0,label='SNR=5',c='g')
+    plt.xlabel(r'$N$ [antennas]')
+    plt.ylabel(r'# Iterations')
+    plt.legend(loc=5,prop={'size': 18})
+    plt.xlim([37,217])
+    
+    plt.grid('on')
+    
+    plt.show() 
+
+def plot_k_new():
+    
+    #N = np.array([7,19,37,61,91,108,169,217])
+    matplotlib.rcParams.update({'font.size': 22})
+
+    #N = np.arange(37,218)
+    
+    #P = 6*N - np.sqrt(12*N - 3) -1
+
+    #factor = P**(0.7) - 1
+
+    output = open("outerloop_1000.p", 'rb')
+    N1 = pickle.load(output)
+    outerloop_pcg_mean1 = pickle.load(output)
+    outerloop_pcg_std1 = pickle.load(output)
+    outerloop_stef_mean1 = pickle.load(output)
+    outerloop_stef_std1 = pickle.load(output)
+    
+    output.close()
+
+    N1 = N1[2:]
+    outerloop_pcg_mean1 = outerloop_pcg_mean1[2:]
+    outerloop_pcg_std1 = outerloop_pcg_std1[2:]
+    outerloop_stef_mean1 = outerloop_stef_mean1[2:]
+    outerloop_stef_std1 = outerloop_stef_std1[2:]
+    diff1 = outerloop_stef_mean1 - outerloop_pcg_mean1
+    
+    output = open("outerloop_5.p", 'rb')
+    N2 = pickle.load(output)
+    outerloop_pcg_mean2 = pickle.load(output)
+    outerloop_pcg_std2 = pickle.load(output)
+    outerloop_stef_mean2 = pickle.load(output)
+    outerloop_stef_std2 = pickle.load(output)
[…]

### 2017-11-01 software/simulator.py

@@ -503,7 +503,7 @@ class sim():
       num_sources - the amount of values to draw from the pareto distribution.
       plot - if true plot the distribution.
       '''
-      def generate_flux(self,a = 1.2,num_sources = 10,plot = False):
+      def generate_flux(self,a = 2,num_sources = 10000,plot = True):
           y = np.random.pareto(a, size=num_sources)
           if plot:
              count, bins, ignored = plt.hist(y, 100, normed=True)
@@ -1131,8 +1131,10 @@ def Gianni_freq_exp():
     plt.show() 
 
 if __name__ == "__main__":
+   s = sim(layout="REG",order=10)
+   s.generate_flux()
    #example_usage()
-   plot_paper_layouts()
+   #plot_paper_layouts()
    #Gianni_freq_exp()
    
 

### 2017-11-03 paper_G_final.tex
newest version of red cal paper

@@
 \usepackage{mathrsfs}
+\usepackage{url}
 
@@
 %$^{6}$Radio Astronomy Lab., U. California, Berkeley CA 94720, USA}
-\author[T.L.~Grobler et al.]{T.L.~Grobler$^{1}$\thanks{E-mail: email@address (AVR)}, G.~Bernardi$^{1,2}$, C.L.~Carilli$^{3,4}$, J.S.~Kenyon$^{1}$, A.R.~Parsons$^{5,6}$?
-\newauthor and O.M.~Smirnov$^{1,2}$\\
-$^{1}$Department of Physics and Electronics, Rhodes University, PO Box 94, Grahamstown, 6140, South Africa {\bf (GB: add your Stellenbosch affilitation)}\\
-$^{2}$SKA SA, 3rd Floor, The Park, Park Road, Pinelands, 7405, South Africa\\
-$^{3}$National Radio Astronomy Obs., Socorro NM\\
-$^{4}$Cavendish Lab., Cambridge UK\\
-$^{5}$Dept. of Astronomy, University of California, Berkeley, CA 94720, USA\\
-$^{6}$Radio Astronomy Lab., U. California, Berkeley CA 94720, USA}
+\author[T.L.~Grobler et al.]{T.L.~Grobler$^{1}$\thanks{E-mail: email@address (AVR)}, G.~Bernardi$^{2,3}$, C.L.~Carilli$^{4,5}$, J.S.~Kenyon$^{2}$, A.R.~Parsons$^{6,7}$?
+\newauthor and O.M.~Smirnov$^{2,3}$\\
+$^{1}$Dept of Mathematical Sciences, Computer Science Division, Stellenbosch University, Private Bag X1, 7602 Matieland, South Africa\\
+$^{2}$Department of Physics and Electronics, Rhodes University, PO Box 94, Grahamstown, 6140, South Africa\\
+$^{3}$SKA SA, 3rd Floor, The Park, Park Road, Pinelands, 7405, South Africa\\
+$^{4}$National Radio Astronomy Obs., Socorro NM\\
+$^{5}$Cavendish Lab., Cambridge UK\\
+$^{6}$Dept. of Astronomy, University of California, Berkeley, CA 94720, USA\\
+$^{7}$Radio Astronomy Lab., U. California, Berkeley CA 94720, USA}
 \begin{document}
@@
 \begin{abstract}
-Observations of the redshifted 21-cm line from the epoch of reionization have recently motivated the construction of low frequency radio arrays with highly redundant configurations that provide an alternative calibration strategy - ''redundant calibration" - and boosts sensitivity on specific spatial scales.
-In this paper we formulate calibration of redundant interferometric arrays as a complex optimization problem \citep[i.e.,][]{Smirnov2015} and derive the corresponding calibration equation which we then solve using the Levenber-Marquardt algorithm.
-This calibration approach is more robust to initial condition than current algorithms and, leveraging on an approximate matrix inversion (``redundant \textsc{StEfCal}"), allows for a very efficient implementation.
-We investigated the use of the preconditioned conjugate gradient method to solve for the calibration equation without approximating the matrix inversion and found that its computational performance is not competitive with respect to ``redundant \textsc{StEfCal}".
-The implementation of the new algorithm is made publicly available.
+Observations of the redshifted 21-cm line from the epoch of reionization have recently motivated the construction of low 
+frequency radio arrays with highly redundant configurations that provide an alternative calibration strategy - 
+''redundant calibration" - and boosts sensitivity on specific spatial scales. In this paper we formulate 
+calibration of redundant interferometric arrays as a complex optimization problem \citep[i.e.,][]{Smirnov2015}. 
+We then solve this optimization problem via the Levenberg-Marquardt algorithm. This calibration approach is more 
+robust to initial conditions than current algorithms and, leveraging on an approximate matrix inversion, allows for further optimization and an efficient implementation (``redundant \textsc{StEfCal}").
+We also investigated the use of the preconditioned conjugate gradient method as an alternative optimization strategy, instead of
+an approximate matrix inverse, but found that its computational performance is not competitive with respect to ``redundant \textsc{StEfCal}".
+The efficient implementation of this new algorithm is made publicly available.
 %The construction of new radio interferometers, particularly operating at low frequencies and In this paper we show that instead of splitting our redundant calibration problem into its phase-and-amplitude components (the more traditional approach) we can treat the complex unknown model parameters and their complex conjugates as separate unknowns.
@@
 Leveraging on this property, redundancy is a promising path to achieve high accurate interferometric calibration \citep[][]{Noordam1982,Wieringa1992,Pearson1982,Liu2010,Noorishad2012,Marthi2014,Sievers2017}.
-Moreover, redundant configurations provide a sensitivity boost on a selected sample of spatial scales that can be tuned to be the highest signal--to--noise ratio EoR modes \citep[][]{Parsons2012,Dillon2016}.
+Moreover, redundant configurations provide a sensitivity boost on a selected sample of spatial scales that can be tuned to be the highest signal--to--noise ratio (SNR) EoR modes \citep[][]{Parsons2012,Dillon2016}.
 These reasons motivated the design and deploment of EoR arrays in redundant configurations like the MIT-EoR \citep{Zheng2014}, the Precision Array to Probe the Epoch of Reionization \citep[PAPER,][]{Ali2015} and the Hydrogen Epoch of Reionization Array \citep[HERA,][]{deboer2017}.
@@
 The real and imaginary components of the thermal noise are normally distributed with a mean of zero and a
-standard deviation
+standard deviation $\sigma$:
 \begin{equation}
@@
 \end{equation}
-where $\Lambda$ is the objective function, $\br$ is the residual vector and $\|\cdot\|$ denotes the Frobenius norm.
+where $\Lambda$ is the objective function, $\br$ is the residual vector and $\|\cdot\|_F$ denotes the Frobenius norm.
 In standard interferometric calibration, $y_{pq}$ is assumed to be known at some level, for instance through the observation of previously known calibration sources.
@@
 \label{eq:update_skymodel}
-\breve{\bg}_{k+1} = \breve{\bg}_{k} + \Delta \breve{\bg}_{k}. 
+\breve{\bg}_{k+1} = \breve{\bg}_{k} + \Delta \breve{\bg}_{k},
 \end{equation}
@@
 Starting from the latest value of the gain conjugates, an updated estimate of the gains can be obtained and the succession continues until convergence is reached.
-Alternating between solving and fixing different sets of parameters (which is exactly what \textsc{StEFCal} does) is in general referred to 
-as the alternating direction implicit method (ADI). The \textsc{StEFCal} algorithm reduces the computational complexity of calibration from $\mathbb{O}(N^3)$ to $\mathbb{O}(N^2)$.
+Alternating between solving and fixing different sets of parameters (which is exactly what \textsc{StEfCal} does) is in general referred to 
+as the alternating direction implicit method (ADI). The \textsc{StEfCal} algorithm reduces the computational complexity of calibration from $O(N^3)$ to $O(N^2)$.
 
@@
 \end{equation}
-with the same vector form as equation~\ref{eq:vis_red} if
+with the same vector form as equation~\ref{eq:vis_linear_definition} if
 \begin{align}
@@
  \bg &=[g_1,\cdots,g_N]^T, & \by &= [y_1,\cdots,y_L]^T,\nonumber\\
- \bz &= [\bg^T,\by^T]^T, &  &\label{eq:parm_definitions}
+ \bz &= [\bg^T,\by^T]^T. &  &\label{eq:parm_definitions}
  \end{align}
-where the number of model parameters to be solved for is now $P = 2(N+L)$, since redundant calibration is a complex problem.
+The number of model parameters to be solved for is now $P = 2(N+L)$, since redundant calibration is a complex problem.
 
 In literature, equation~\ref{eq:least_squares_red} is solved by splitting the problem into its real and imaginary parts.
-The real and imaginary parts of the unknown parameters are then solved separately \citep{Wieringa1992,Liu2010,Zheng2014}.
+The real and imaginary parts of the unknown parameters are then solved for separately \citep{Wieringa1992,Liu2010,Zheng2014}.
 Currently, the above is achieved by using the real-valued GN algorithm \citep{Kurien2016}. 
@@
 The the index set $\mathcal{RS}_i$ and the quantity $\widetilde{y}_{ij}$ is defined in equation~\ref{eq:RS} and equation~\ref{eq:y_tilde}, respectively. 
-The computational complexity of inverting $\widetilde{\bH}$ is $\mathbb{O}(P)$. We note that equation~\ref{eq:g_update} is the gain estimator associated with \textsc{StEfCal}.
+The computational complexity of inverting $\widetilde{\bH}$ is $O(P)$. We note that equation~\ref{eq:g_update} is the gain estimator associated with \textsc{StEfCal}.
 
@@
 In Appendix~\ref{sec:red_stef_ADI} we show that equation~\ref{eq:g_update} and~\ref{eq:y_update} can also be derived using the ADI method.
-For this reason, we will refer to the pseudo LM calibration scheme, derived above (equation~\ref{eq:g_update} and~\ref{eq:y_update}), as redundant \textsc{StEFCal}.
+For this reason, we refer to the code implementation of the pseudo LM calibration scheme, derived above (equation~\ref{eq:g_update} and~\ref{eq:y_update}), as redundant \textsc{StEfCal}, throughout the rest of the paper.
 {\bf (GB: this needs a better wording)}.
@@
 i.e. we chose $\alpha$ to be equal to a $\frac{1}{3}$ ($\lambda = 2$). We also carried out simulations to validate this choice.
-%In the case of \textsc{StEFCal}, $\alpha = \frac{1}{2}$ (for even iterations), which results in the last term of equation~\ref{eq:one_half} and equation~\ref{eq:stef_alpha} being equal.
+%In the case of \textsc{StEfCal}, $\alpha = \frac{1}{2}$ (for even iterations), which results in the last term of equation~\ref{eq:one_half} and equation~\ref{eq:stef_alpha} being equal.
 When $\alpha=\frac{1}{3}$, the last term of equation~\ref{eq:two_thrids} and equation~\ref{eq:alpha} are equal.
@@
 patch centred. The flux density of each source was drawn from a Pareto (power law) distribution 
-with a shape parameter of 2. The chosen shape parameter ensures that it is unlikely that a source, that is more than 20 Jy, is added to the simulation sky model. 
+with a shape parameter of 2. The chosen shape parameter ensures that it is unlikely that a source, that is more than 20 Jy, is added to the aforementioned sky model. 
 {\bf (GB: we need some form of motivation for the distribution choice and its parameter)}) 
-{\bf (TLG: this is a bit of an ad-hoc choice, I knew that the flux distribution follows some power-law so a settled on a Pareto distribution. I played around 
+{\bf (TLG: this is a bit of an ad-hoc choice, I knew that the flux distribution follows some power-law so I settled on a Pareto distribution. I played around 
 with $a$ until it produced flux values I deemed reasonable. This distribution choice stems from the suppression work I did with Ridhima.
-It be noted, however, that it is the SNR value which influence the results of calibration in general (the SNR of 
+It should be noted, however, that it is the SNR value which influence the results of calibration in general (the SNR of 
 your problem determines how well you can calibrate). Whatever power is in your model is irrelevant, it 
 is the ratio of model power versus noise power which matters. The SNR 
-is a better reflection of what you are doing. Which is why I define it explicitly. Maybe you can word it better.)})
+is a true reflection of the solvability of your problem. Which is why I use it in the paper. PLEASE HELP HERE AS I HAVE NO OTHER MOTIVATION
+FOR THE PARETO DITRO???.)})
 and the source position drawn from a uniform distribution. 
@@
 
-We "corrupted" visibilities by applying gain errors (Table~\ref{tab:gain_parm}) and calibrated the corrupted visibilities using the redundant \textsc{StEfCal} algorithm.
-Solutions were independently derived for time-step and channel.
+We "corrupted" visibilities by applying gain errors (Table~\ref{tab:gain_parm}) and calibrated the corrupted visibilities using redundant \textsc{StEfCal}.
+Solutions were independently derived for each time-step and channel.
 We used 5 independent gain realizations. {\bf (GB: How many different realizations of the gains????)}
@@
 
-The simulated visibilities included a noise term. The purpose of which is to produce a certain signal--to-noise ratio (SNR) in the simulated
+The simulated visibilities included a noise term. The purpose of which is to produce a certain SNR in the simulated
 visibilities. We made use of the following definition of \citep[SNR, ][]{Liu2010,Marthi2014}:  
@@
 \label{}
-\textrm{SNR} = \frac{<\bv\odot\conj{\bv}>_{\nu,t,pq}}{<\bn\odot\conj{\bn}>_{\nu,t,pq}}, 
+\textrm{SNR} = 10\log \left (\frac{<\bv\odot\conj{\bv}>_{\nu,t,pq}}{<\bn\odot\conj{\bn}>_{\nu,t,pq}} \right ), 
 \end{equation}
@@
 expression with Adrian Liu. I do not know what you wish to add here. I am just trying to say that my simulations are proper, I do not use a relative quantity to 
-conduct them. This is the same SNR definition as was used in previous redundant cal papers.)}.
+conduct them. This is the same SNR definition as was used in previous redundant cal papers)}.
 
@@
 \end{equation}
-where $\widehat{\bv}$ is the redundant \textsc{StEFCal} parameter estimate.
+where $\widehat{\bv}$ is the redundant \textsc{StEfCal} parameter estimate.
 
@@
 
-\citet{Marthi2014} conducted an even more in depth analysis than the one presented here. They found that the total computational cost of redundant StEfCal
-is $\mathbb{O}(P^2)$ and that it approaches the Cram{\'e}r-Rao Bound (CRB) at moderate SNRs. 
+\citet{Marthi2014} conducted an even more in depth analysis than the one presented here. They found that the total computational cost of the algorithm associated with redundant StEfCal
+is $O(P^2)$ and that it approaches the Cram{\'e}r-Rao Bound (CRB) at moderate SNRs. 
 {\bf (GB: is there a similar analysis in \cite{Liu2010} a/o \cite{Marthi2014}? If so, what do they find?)}.
-{\bf (TLG: added.)}
+{\bf (TLG: Added.)}
 
@@
 \includegraphics[width=0.47\textwidth]{./prec_error.pdf} 
-\caption{We plot the percentage error $\beta$ between the simulated visibilitied and the visibilities solved for by \textsc{StEFCal} for different SNR values as a function of the number of antennas in the array $N$.
+\caption{We plot the percentage error $\beta$ between the simulated visibilitied and the visibilities solved for by redundant \textsc{StEfCal} for different SNR values as a function of the number of antennas in the array $N$.
 {\bf (GB: I don't undesrtand the curve with SNR = -1...
-The yaxis should read $\beta$).}}
+The yaxis should read $\beta$).} {\bf (TLG: Added $\beta$, SNR is expressed in dB.
+Made this more explicit).}}
 \label{fig:prec_error}
@@
 \citet{Liu2010} suggested that the speed of convergence of redundant calibration could be optimized using the 
-conjugate gradient method \citep{Hestenes1952}, which would be computationally advantageous, since the Hessian matrix 
-is sparse \citep{Reid1971}. In this section we study the computational complexity
-of the conjugate gradient method when used to invert the Hessian matrix in equation~\ref{eq:LM_update}, 
-in particular through the use of preconditioning (i.e. the preconditioned conjugate gradient (PCG) method). 
+conjugate gradient method \citep{Hestenes1952}, which would be computationally advantageous, since the Hessian matrix associated 
+with redundant calibration (see Figure~\ref{fig:hessian}) is sparse \citep{Reid1971}. In this section we study the computational complexity
+of the conjugate gradient (CG) method when it is used to invert the modified Hessian matrix (see equation~\ref{eq:LM_update}), 
+in particular when preconditioning is used (i.e.
+the PCG method).
+Interestingly, emperical tests suggests that the unmodified Hessian itself is singular.
+It is therefore important to mention here, that the CG method
+can pseudo invert the unmodified Hessian as well, i.e.
+the CG method can be directly applied to equation~\ref{eq:GN_update}, since the Hessian is a positive semi-definite Hermition matrix and the vector $\bJ^H\breve{\br}$ is an element of its column range \citep{Lu2015}.
+
 %(see also Appendix~\ref{sec:conj_grad}).
@@
 \label{eq:cg_bound}
-\mathbb{O}(\sqrt{\kappa}m), 
+O(\sqrt{\kappa}m), 
 \end{equation}
@@
 \end{equation}
-i.e., if it lowers lowers the condition number of a matrix.
+i.e. if it lowers lowers the condition number of a matrix.
 If $\bA$ is a nearly diagonal matrix, the Jacobian preconditioner is a natural choice of $\bM$ and can be computed as:
 \begin{equation}
-\bM = \bA\odot\bI, 
-\end{equation}
-where $\bI$ is the identity matrix.
-
-%The most expensive operation ni the PCG algorithm is the vector-matrix product $\bA\bx$, which has a complexity of $\mathbb{O}(m)$, where $m$ is the number of non-zero entries in $\bA$.
+\bM = \bA\odot\bI. 
+\end{equation}
+
+%The most expensive operation ni the PCG algorithm is the vector-matrix product $\bA\bx$, which has a complexity of $O(m)$, where $m$ is the number of non-zero entries in $\bA$.
 
@@
 Later on as you became more involved you suggested setup 1 (and even later still gain model 3) as it is more
-realistic for redundant low-frequency arrays (redoing the red stefcal sims with these parameters did not influence the results though if memory serves me right -- not sure if I still have the old results though to confirm).
+realistic for redundant low-frequency arrays.
+Redoing the red stefcal sims with these parameters did not influence the results though if memory serves me right -- not sure if I still have the old results though to confirm.
 We never redid the sims with PCG as the sims took quite a bit of time (much
-longer than the unvectorized red stefcal, which as we know takes quite a long time) and there 
+longer than unvectorized red stefcal, which as we know takes quite a long time) and there 
 would ultimately be now difference in result, the calibration algorithm's speed does not depend on the observational frequency etc... Moreover, 
 Fig. 5 is completely independent of the chosen simulated parameters as it depends only on the analytic expression of the Hessian. Not sure 
-how to phrase or explain the different setups in the paper, as I would like to avoid redoing the simulations at this stage, except if the referee actually requests this.)}
-
-To calibrate, we used the complex LM-algorithm described in Section~\ref{sec:red_wirtinger}. To invert the 
+how to phrase or explain the different setups in the paper, as I would like to avoid redoing the simulations at this stage, except if the referee actually requests this. PLEASE HELP HERE. 
+UNSURE HOW TO MOTIVATE OR EXPLAIN THE SETUPS. NOT SURE WHERE SUCH A DISCUSSION BELONGS.)}
+
+To calibrate, we used the complex LM-algorithm described in Section~\ref{sec:red_wirtinger}. To invert the modified
 Hessian we used the CG method, with and without a preconditioner. We made use of the Jacobian preconditioner. 
@@
 \label{eq:cg_bound2}
-\mathbb{O}(m). 
+O(m). 
 \end{equation}
 This result is confirmed by Fig.~\ref{fig:cg_itr}. Fig.~\ref{fig:cg_itr} 
-shows us that the number of major iterations needed by the preconditioned conjugate gradient method to invert 
+shows us that the number of major iterations needed by the PCG method to invert 
 $\bmH$ is independent of the number of antennas in the array and that it is much less than the 
@@
 you everything about the number of major iterations).
-I think all of this becomes to complicated to explain in the paper and almost irrelevant to the idea i am trying to convey (kappa is constant implies number of major
-iterations is independent of problem size. Emprirical sims show it to be less than $P$. Which is why I try to focus on the basics, hope this makes more sense now?)}
+I think all of this becomes to complicated to explain in the paper and almost irrelevant for the point I am trying to make in the paper: kappa is constant implies number of major
+iterations is independent of problem size. Moreover, empirical sims show it to be less than $P$. Which is why I try to focus on the basics, hope this makes more sense now?)}
 
@@
 \subfigure[Iterations required before and after Preconditioning]{\includegraphics[width=0.47\textwidth]{./cg_itr.pdf}\label{fig:cg_itr}}
-\caption{The graphs presented here were generated using the second experimental setup in Table~\ref{tab:ch_parm}.
-Left: spectral condition number $\kappa$ of the modified Hessian $\bmH$as a function of $N$, before (magenta and red curves) and after (blue curve) preconditioning, at different SNR values.
-Right: number of major iterations (line 6 of Algorithm \ref{algo:CG}) required by the conjugate
+\caption{The graphs presented here were generated using the simulations discussed in Section~\ref{sec:red_wirtinger}.
+Left: spectral condition number $\kappa$ of the modified Hessian $\bmH$ as a function of $N$, before (magenta and red curves) and after (blue curve) preconditioning, at different SNR values.
+Right: number of major iterations required by the conjugate
 gradient method to invert $\bmH$ as a function of the number of antennas $N$ in the array, before (magenta and red curves) and after preconditioning (blue and green curves), at different SNR values.
@@
 \end{figure*}
+
+%The graphs presented here were generated using the second experimental setup in Table~\ref{tab:ch_parm}.
 
@@
 \subfigure[$c$]{\includegraphics[width=0.46\textwidth]{./comp_order.pdf}\label{fig:c}}
-\caption{Left: the sparsity ratio $\gamma$ of the modified Hessian $\bmH$ as a function of the number of antennas $N$ for an hexagonal (blue circles), square (green crosses) and regular east-west grid (red circles) array geometry.
+\caption{Left: the sparsity ratio $\gamma$ of the modified Hessian $\bmH$ as a function of the number of antennas $N$ for an hexagonal (blue circles), square (green crosses) and regular east-west (red circles) array geometry.
 The red--dashed and black--dotted lines show the analytical expression for $\gamma$ and its limit for the regular east-west grid case (see text for details).
@@
 
-We are now finally able to compare the computational complexity of redundant \textsc{StEFCal} and the PCG method.
-Redundant \textsc{StEFCal} is computationally inexpensive as it just needs to invert a diagonal matrix, however, the PCG inversion is accurate and, therefore, may require fewer iterations, ultimately leading to a faster convergence.
-We computed the theoretical number of redundant \textsc{StEFCal} iterations $\Delta k$, in excess
+We are now finally able to compare the computational complexity of redundant \textsc{StEfCal} and the PCG method.
+Redundant \textsc{StEfCal} is computationally inexpensive as it just needs to invert a diagonal matrix, however, the PCG inversion is accurate and, therefore, may require fewer iterations, ultimately leading to a faster convergence.
+We computed the theoretical number of redundant \textsc{StEfCal} iterations $\Delta k$, in excess
 of the number of LM (implementing PCG) iterations required, for 
-LM to outperform redundant \textsc{StEFCal}
+LM to outperform redundant \textsc{StEfCal}
 {\bf (GB: I don't follow the following... perhaps I am missing something.)}
@@
 The results are displayed in Figure~\ref{fig:out_diff}.
-Figure~\ref{fig:out_diff} shows that redundant \textsc{StEFCal} outperforms the PCG method.
+Figure~\ref{fig:out_diff} shows that redundant \textsc{StEfCal} outperforms the PCG method.
 In this comparison we have only taken into account the cost of inverting $\bmH$.
+We also ignored the cost of preconditioning.
 The PCG method is still
@@
 \centering
-\subfigure[Number of LM iterations required by redundant \textsc{StEFCal} and PCG.]{\includegraphics[width=0.47\textwidth]{./outerloop.pdf}\label{fig:outerloop}}
+\subfigure[Number of LM iterations required by redundant \textsc{StEfCal} and PCG.]{\includegraphics[width=0.45\textwidth]{./outerloop.pdf}\label{fig:outerloop}}
 % V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
 \subfigure[Difference.]{\includegraphics[width=0.45\textwidth]{./diff.pdf}\label{fig:diff}}
-\caption{The graphs presented here were generated using the second experimental setup in Table~\ref{tab:ch_parm}.
-Left: number of LM iterations required by redundant \textsc{StEFCal} (green and blue curves) and the PCG (magenta and red curves) methods to converge to a parameter error tolerance of $10^{-6}$ whilst using different SNR values as a function of the number of antennas $N$ in the array {\bf (GB: which configuration?)}.
-Right: average amount of LM iterations (difference between the redundant \textsc{StEFCal} and PCG curves in the left panel) saved (green and blue curve) by computing the full-inverse with the PCG method whilst using different SNR values.
+\caption{The graphs presented here were generated using the simulations discussed in Section~\ref{sec:red_wirtinger}.
+Left: number of LM iterations required by redundant \textsc{StEfCal} (green and blue curves) and the PCG (magenta and red curves) methods to converge to a parameter error tolerance of $10^{-6}$ whilst using different SNR values as a function of the number of antennas $N$ in the array {\bf (GB: which configuration?) - (TLG: hexagonal as mentioned in sims from red wir sec)}.
+Right: average amount of LM iterations (difference between the redundant \textsc{StEfCal} and PCG curves in the left panel) saved (green and blue curve) by computing the full-inverse with the PCG method whilst using different SNR values.
 The black curve is $\Delta k$, the theoretical number
-of redundant \textsc{StEFCal} iterations that are needed in excess of the number of LM iterations used for the LM algorithm to 
-outperform redundant \textsc{StEFCal}. For the black curve we assumed $c=1.7$ and that $k$ could be approximated with the 
+of redundant \textsc{StEfCal} iterations that are needed in excess of the number of LM iterations used for the LM algorithm to 
+outperform redundant \textsc{StEfCal}. For the black curve we assumed $c=1.7$ and that $k$ could be approximated with the 
 the magenta curve plotted in the left panel.
@@
 \label{sec:conclusions}
-In this paper we have formulated the calibration of redundant interferometric arrays using the complex optimization formalism \citep{Smirnov2015}.
-We have derived the related Jacobian and updated parameter steps and showed that this approach naturally leads to solving the calibration equations in the fast way used by alternating direction implicit methods \citep[][]{Salvini2014,Marthi2014}.
-The code (redundant \textsc{StEfCal}) is publicly available at ?????.
+In this paper we have formulated the calibration of redundant interferometric arrays using the complex optimization formalism 
+\citep{Smirnov2015}. We derived the associated complex Jacobian and the GN and LM parameter update steps.
+We also showed that the LM parameter update step can be simplified to obtain an ADI type of algorithm \citep[][]{Salvini2014,Marthi2014}.
+Our code implementation of this algorithm (redundant \textsc{StEfCal}) is publicly available at 
+\url{https://github.com/Trienko/heracommissioning/blob/master/code/stef.py}. 
 We note that, in its current implementation, redundant \textsc{StEfCal} does not solve for the degeneracies inherent to redundant calibration  \citep{Zheng2014,Kurien2016} which will be the subject of future work.
@@
 
-We investigated the computational cost of redundant \textsc{StEfCal} and compared it with the performance of the preconditioned conjugated method suggested by \cite{Liu2010} to solve the calibration equations.
-We found that, although the preconditioned conjugated method method greatly improves the speed of redundant calibration, it still significant underperforms redundant \textsc{StEfCal}.
-
-The characteristics of \textsc{StEfCal} makes it an appealing calibration algorithm for large redundant arrays like HERA \citep{deboer2017}, CHIME (cit!!!!), HIRAX (cit!!!!) or even hybrid arrays like the MWA (cit!!!!).
+We investigated the computational cost of redundant \textsc{StEfCal} and compared it with the performance of the PCG method (approach suggested by \cite{Liu2010}).
+We found that, although the PCG method method greatly improves the speed of redundant calibration, it still significant underperforms in comparison to redundant \textsc{StEfCal}.
+
+The characteristics of redundant \textsc{StEfCal} makes it an appealing calibration algorithm for large redundant arrays like HERA \citep{deboer2017}, CHIME (cit!!!!), HIRAX (cit!!!!) or even hybrid arrays like the MWA (cit!!!!).
+{\bf (TLG: Please add missing references.)}
 
 \section*{Acknowledgements}
-{\bf (GB: you may need to add an acknowledgement to the group grant that supported you last year, pllease, ask Oleg)} G.B.
+{\bf (GB: you may need to add an acknowledgement to the group grant that supported you last year, please, ask Oleg)}. 
+{\bf (TLG: wish to change my affiliation on this paper to Stellenbosch (which I did for now), while acknowledging SKA, Rhodes here. This 
+is to get some backing from my current division to continue with the research. GB, would like to know your thoughts 
+on this. Have not done so as I need to ask Oleg for approval)}.
+G.B.
 acknowledges support from the Royal Society and the Newton Fund under grant NA150184.
@@
 
-\section{\textsc{LOGCAL}}
-\label{sec:logcal}
-We can rewrite equation~\ref{eq:vis_red} to \citep{Liu2010}
-\begin{eqnarray}
-\ln |d_{pq}| &=& \ln |g_p| + \ln |g_q| + \ln |y_{\phi_{pq}}| + s_{pq} \label{eq:logcal_amp}\\
-\angle d_{pq} &=& \angle g_p - \angle g_q + \angle \phi_{pq} + t_{pq} \label{eq:logcal_phase}
-\end{eqnarray}
-where
-\begin{align}
-s_{pq} &= \mathscr{R} \left \{1 + \frac{n_{pq}}{g_p\conj{g_q}y_{\phi_{pq}}} \right \}, & t_{pq} &= \mathscr{I}\left \{1 + \frac{n_{pq}}{g_p\conj{g_q}y_{\phi_{pq}}} \right \}.
-\end{align}
-Moreover, $\mathscr{R}\{*\}$ and $\mathscr{I}\{*\}$ denote the real and imaginary operators.
-
-With the aid of equation~\ref{eq:logcal_amp} we can construct the following linear system:
-\begin{equation}
-\label{eq:log_system}
-\bdelta = \bJ\bzeta + \bs, 
-\end{equation}
-where
-\begin{align}
-\left [ \bdelta \right ]_{\alpha_{pq}} &= \ln |d_{pq}|, & \left [ \bs \right ]_{\alpha_{pq}} = s_{pq} 
-\end{align}
-and
-\begin{equation}
-\left [ \bzeta \right ]_{j} = \begin{cases} \ln |g_j| & \textrm{if}~j\leq N \\ \ln |y_{j-N}| & \textrm{otherwise} \end{cases}. 
-\end{equation}
-Furthermore,
-\begin{equation}
-\bJ = 
-\begin{bmatrix}
-\bN & \bM
-\end{bmatrix}
-\end{equation}
-with 
-\begin{equation}
-[\bN]_{\alpha_{pq},j} = \begin{cases}
-       1 & \textrm{if}~(p=j)~\textrm{or}~(q=j)\\
-       0 & \textrm{otherwise}
-      \end{cases}
-\end{equation}
-and
-\begin{equation}
-[\bM]_{\alpha_{pq},j} = \begin{cases}
-       1 & \textrm{if}~\phi_{pq}=j\\
-       0 & \textrm{otherwise}
-      \end{cases}
-\end{equation}
-
-The least squares solution to Eq.\ref{eq:log_system}, under the incorrect assumption that $s_{pq}$ is normally distributed, is given by
-\begin{equation}
-\bzeta = (\bA^T\bA)^{-1} \bA^T\bdelta. 
-\end{equation}
-How to correctly deal with the fact that $s_{pq}$ is not normally distributed is discussed in \citep{Liu2010}. A similar linear system can be constructed for equation~\ref{eq:logcal_phase}. 
-From the least squares solutions to the linear systems associated with equation~\ref{eq:logcal_amp} and equation~\ref{eq:logcal_phase} we can compute the antenna gains as well as the true visibilities.
-
-\section{\textsc{lincal}}
-\label{sec:lincal}
-Assume that $\eta_p,\eta_q,\widetilde{\eta}_{\phi_{pq}},\varphi_p,\varphi_q$ and $\widetilde{\varphi}_{\phi_{pq}}$ are real valued variables.
-We can rewrite equation~\ref{eq:vis_red_vec} to \citep{Liu2010}
-\begin{equation}
-\label{eq:lincal_vec}
-\bd = \bv(\bvarrho) + \bn, 
-\end{equation}
-where 
-\begin{align}
-\left [ \bv(\bvarrho) \right ]_{\alpha_{pq}} &= e^{\eta_p - i \varphi_p} e^{\widetilde{\eta}_{\phi_{pq}}- i \widetilde{\varphi}_{\phi_{pq}}} e^{\eta_q + i \varphi_q}\\
-&= v_{pq},
-\end{align}
-and
-\begin{equation}
-\bvarrho = 
-\begin{bmatrix}
-\bmath{\eta} \\
-\bvarphi \\
-\widetilde{\bmath{\eta}} \\
-\widetilde{\bvarphi}
-\end{bmatrix},\qquad
-\begin{aligned}
-\bmath{\eta} &= [\eta_1, \eta_2,\cdots \eta_N]^T\\
- \bmath{\varphi} &= [\varphi_1, \varphi_2,\cdots \varphi_N]^T\\
- \widetilde{\bmath{\eta}} &= [\widetilde{\eta}_1, \widetilde{\eta}_2,\cdots \widetilde{\eta}_L]^T\\
- \widetilde{\bmath{\varphi}} &= [\widetilde{\varphi}_1, \widetilde{\varphi}_2,\cdots \widetilde{\varphi}_L]^T
-\end{aligned}.
-\label{eq:help}
-\end{equation}
-Note that the vectors $\bmath{\eta},~\bvarphi,~\widetilde{\bmath{\eta}}$ and $\widetilde{\bvarphi}$ are real valued.
-
-The first order Taylor expansion of equation~\ref{eq:lincal_vec} around the fiducial point $\bvarrho_0$ is equal to
-\begin{equation}
-\label{eq:d_lin}
-\bd = \bv_0 + \bmJ_0 \Delta \bvarrho  + \bn.
-\end{equation}
-If we calculate the residual, dropping the reference to a specific fiducial point, we obtain: 
-\begin{equation}
-\label{eq:r_lin}
-\br = \bd - \bv = \bmJ \Delta \bvarrho + \bn.
-\end{equation}
-Moreover,
-\begin{equation}
-\label{eq:jac_upper_lincal}
-\bmJ = \begin{bmatrix}
-       \bN & \bM 
-       \end{bmatrix},
-\end{equation}
-where
-\begin{align}
-\bN &= \begin{bmatrix} \bO & \bP \end{bmatrix}, & \bM &= \begin{bmatrix} \bQ & \bR \end{bmatrix}, 
-\end{align}
-\begin{equation}
- [\bO]_{\alpha_{pq},j} =  \frac{\partial v_{pq}}{\partial \eta_j} = \begin{cases} 
-    v_{pq} &\textrm{if}~(p=j)~\textrm{or}~(q=j)\\
-    0 & \textrm{otherwise}
-   \end{cases},
-\end{equation}
-\begin{equation}
- [\bP]_{\alpha_{pq},j} =  \frac{\partial v_{pq}}{\partial \varphi_j} = \begin{cases}
-                                                                        i v_{pq} &\textrm{if}~ p=j\\
-                                                                        -i v_{pq} &\textrm{if}~ q=j \\
-                                                                        0 & \textrm{otherwise} 
-                                                                       \end{cases}
-\end{equation}
-\begin{equation}
- [\bQ]_{\alpha_{pq},j} =  \frac{\partial v_{pq}}{\partial \widetilde{\eta}_j} = \begin{cases} 
-    v_{pq} &\textrm{if}~\phi_{pq}=j\\
-    0 & \textrm{otherwise}
-   \end{cases},
-\end{equation}
-\begin{equation}
- [\bR]_{\alpha_{pq},j} =  \frac{\partial v_{pq}}{\partial \widetilde{\varphi}_j} = \begin{cases} 
-    i v_{pq} &\textrm{if}~\phi_{pq}=j\\
-    0 & \textrm{otherwise}
-   \end{cases}.
-\end{equation}
-Augmenting the residual by taking into account its conjugate results in 
-\begin{equation}
-\breve{\br} = \bJ \Delta \bvarrho + \bn, 
-\end{equation}
-where $\breve{\br} = \begin{bmatrix}\br^T & \conj{\br}^T \end{bmatrix}^T$ and 
-\begin{equation}
-\label{eq:Jac_lin}
-\bJ = 
-\begin{bmatrix}
- \bmJ\\  
- \conj{\bmJ}
-\end{bmatrix}.
-\end{equation}
-The solution to 
-\begin{equation}
- \min_{\Delta \bvarrho} \| \breve{\br} - \bJ \Delta \bvarrho\|_F^2,
-\end{equation}
-is equal to 
-\begin{equation}
-\label{eq:LINCAL_update}
-\Delta \bvarrho = (\bJ^H\bJ)^{-1} \bJ^H\breve{\br}. 
-\end{equation}
-equation~\ref{eq:LINCAL_update} is used to iteratively determine a new estimate of the parameter vector $\bvarrho$ as follows
-\begin{equation}
-\label{eq:new_rho}
-\bvarrho_{k+1} = \bvarrho_k + \Delta \bvarrho_k. 
-\end{equation}
-Note that we do not explicitly denote the $k$ index in equation~\ref{eq:LINCAL_update}.
-Moreover, equation~\ref{eq:LINCAL_update} is the GN update associated with the following minimization problem \citep{Kurien2016}
-\begin{equation}
-\label{eq:least_squares_lincal}
-\min_{\bmath{\varrho}} \|\breve{\br}\| = \min_{\bmath{\varrho}} \|\breve{\bd} - \breve{\bv}(\bmath{\varrho})\|.
-\end{equation}
-
-\section{Conjugate Gradient Method}
-
-\subsection{Normal Equation}
-\label{sec:normal}
-Lets assume that instead of equation~\ref{eq:linear_system} we have a more general linear system, i.e.
-\begin{equation}
- \bb = \bB\bx,
-\end{equation}
-where $\bb\in\mathbb{C}^K$, $\bx\in\mathbb{C}^P$  and $\bB\in\mathbb{C}^{K \times P}$, with $K > P$. We can use the normal equation 
-\begin{equation}
-\label{eq:normal_equation}
-\bB^H\bB\bx = \bB^H\bb, 
-\end{equation}
-to estimate the $\bx$ that minimizes
-\begin{equation}
-\label{eq:norm_ls}
-\|\bb-\bB\bx\|_F^2. 
-\end{equation}
-In this paper, we denote the Hermitian transpose with $(*)^H$ and the Frobenius norm with $\|*\|_F$.
-
-There are a few important observations that we can make while inspecting equation~\ref{eq:normal_equation}:
-\begin{enumerate}
-\item $\bB^H\bB$ is a square positive semi-definite Hermitian matrix by construction.
-\item $\bB^H\bb$ is in the column range of $\bB^H\bB$; this follows trivially from 
-\begin{equation}
-\bB^H\bb \in \mathcal{R}(\bB^H) \implies \bB^H\bb \in \mathcal{R}(\bB^H\bB).   
-\end{equation}
-\end{enumerate}
-In this paper, we denote the column range of a matrix $\bA$ with $\mathcal{R}(\bA)$.
-
-Recall that we mentioned that in general we can only solve equation~\ref{eq:linear_system} with CG if $\bA$ is a square positive-definite Hermitian matrix.
+{\bf (TLG: I have removed logcal and lincal as well as the PCG section. My main concern about this is that the aim of these
+sections was to show off the usefulness of the geometry function, it simplifies, at least from 
+a mathematecial perspective, the derivation of the older algorithms. This insight has now been lost. On the other 
+hand keeping them and trying to make them fit, which may cause further delays, at this point simply no longer 
+seems worht the effort. I have incorporated the remaining ideas left in the PCG section into the main text.)}.
+% \section{\textsc{LOGCAL}}
+% \label{sec:logcal}
+% We can rewrite equation~\ref{eq:vis_red} to \citep{Liu2010}
+% \begin{eqnarray}
+% \ln |d_{pq}| &=& \ln |g_p| + \ln |g_q| + \ln |y_{\phi_{pq}}| + s_{pq} \label{eq:logcal_amp}\\
+% \angle d_{pq} &=& \angle g_p - \angle g_q + \angle \phi_{pq} + t_{pq} \label{eq:logcal_phase}
+% \end{eqnarray}
+% where
+% \begin{align}
+% s_{pq} &= \mathscr{R} \left \{1 + \frac{n_{pq}}{g_p\conj{g_q}y_{\phi_{pq}}} \right \}, & t_{pq} &= \mathscr{I}\left \{1 + \frac{n_{pq}}{g_p\conj{g_q}y_{\phi_{pq}}} \right \}.
+% \end{align}
+% Moreover, $\mathscr{R}\{*\}$ and $\mathscr{I}\{*\}$ denote the real and imaginary operators. 
+% 
+% With the aid of equation~\ref{eq:logcal_amp} we can construct the following linear system:
+% \begin{equation}
+% \label{eq:log_system}
+% \bdelta = \bJ\bzeta + \bs, 
+% \end{equation}
+% where
+% \begin{align}
+% \left [ \bdelta \right ]_{\alpha_{pq}} &= \ln |d_{pq}|, & \left [ \bs \right ]_{\alpha_{pq}} = s_{pq} 
+% \end{align}
+% and
+% \begin{equation}
+% \left [ \bzeta \right ]_{j} = \begin{cases} \ln |g_j| & \textrm{if}~j\leq N \\ \ln |y_{j-N}| & \textrm{otherwise} \end{cases}. 
+% \end{equation}
+% Furthermore,
+% \begin{equation}
+% \bJ = 
+% \begin{bmatrix}
+% \bN & \bM
+% \end{bmatrix}
+% \end{equation}
+% with 
+% \begin{equation}
+% [\bN]_{\alpha_{pq},j} = \begin{cases}
+%        1 & \textrm{if}~(p=j)~\textrm{or}~(q=j)\\
+%        0 & \textrm{otherwise}
+%       \end{cases}
+% \end{equation}
+% and
+% \begin{equation}
+% [\bM]_{\alpha_{pq},j} = \begin{cases}
+%        1 & \textrm{if}~\phi_{pq}=j\\
+%        0 & \textrm{otherwise}
+%       \end{cases}
+% \end{equation}
+% 
+% The least squares solution to Eq.\ref{eq:log_system}, under the incorrect assumption that $s_{pq}$ is normally distributed, is given by
+% \begin{equation}
+% \bzeta = (\bA^T\bA)^{-1} \bA^T\bdelta. 
+% \end{equation}
+% How to correctly deal with the fact that $s_{pq}$ is not normally distributed is discussed in \citep{Liu2010}. A similar linear system can be constructed for equation~\ref{eq:logcal_phase}. 
+% From the least squares solutions to the linear systems associated with equation~\ref{eq:logcal_amp} and equation~\ref{eq:logcal_phase} we can compute the antenna gains as well as the true visibilities.
+% 
+% \section{\textsc{lincal}}
+% \label{sec:lincal}
+% Assume that $\eta_p,\eta_q,\widetilde{\eta}_{\phi_{pq}},\varphi_p,\varphi_q$ and $\widetilde{\varphi}_{\phi_{pq}}$ are real valued variables.
+% We can rewrite equation~\ref{eq:vis_red_vec} to \citep{Liu2010}
+% \begin{equation}
+% \label{eq:lincal_vec}
+% \bd = \bv(\bvarrho) + \bn, 
+% \end{equation}
+% where 
+% \begin{align}
+% \left [ \bv(\bvarrho) \right ]_{\alpha_{pq}} &= e^{\eta_p - i \varphi_p} e^{\widetilde{\eta}_{\phi_{pq}}- i \widetilde{\varphi}_{\phi_{pq}}} e^{\eta_q + i \varphi_q}\\
+% &= v_{pq},
+% \end{align}
+% and
+% \begin{equation}
+% \bvarrho = 
+% \begin{bmatrix}
+% \bmath{\eta} \\
+% \bvarphi \\
+% \widetilde{\bmath{\eta}} \\
+% \widetilde{\bvarphi}
+% \end{bmatrix},\qquad
+% \begin{aligned}
+% \bmath{\eta} &= [\eta_1, \eta_2,\cdots \eta_N]^T\\
+%  \bmath{\varphi} &= [\varphi_1, \varphi_2,\cdots \varphi_N]^T\\
+%  \widetilde{\bmath{\eta}} &= [\widetilde{\eta}_1, \widetilde{\eta}_2,\cdots \widetilde{\eta}_L]^T\\
+%  \widetilde{\bmath{\varphi}} &= [\widetilde{\varphi}_1, \widetilde{\varphi}_2,\cdots \widetilde{\varphi}_L]^T
+% \end{aligned}.
+% \label{eq:help}
+% \end{equation}
+% Note that the vectors $\bmath{\eta},~\bvarphi,~\widetilde{\bmath{\eta}}$ and $\widetilde{\bvarphi}$ are real valued.
+% 
+% The first order Taylor expansion of equation~\ref{eq:lincal_vec} around the fiducial point $\bvarrho_0$ is equal to
+% \begin{equation}
+% \label{eq:d_lin}
+% \bd = \bv_0 + \bmJ_0 \Delta \bvarrho  + \bn.
+% \end{equation}
+% If we calculate the residual, dropping the reference to a specific fiducial point, we obtain: 
+% \begin{equation}
+% \label{eq:r_lin}
+% \br = \bd - \bv = \bmJ \Delta \bvarrho + \bn.
+% \end{equation}
+% Moreover,
+% \begin{equation}
+% \label{eq:jac_upper_lincal}
+% \bmJ = \begin{bmatrix}
+%        \bN & \bM 
+%        \end{bmatrix},
+% \end{equation}
+% where
+% \begin{align}
+% \bN &= \begin{bmatrix} \bO & \bP \end{bmatrix}, & \bM &= \begin{bmatrix} \bQ & \bR \end{bmatrix}, 
+% \end{align}
+% \begin{equation}
+%  [\bO]_{\alpha_{pq},j} =  \frac{\partial v_{pq}}{\partial \eta_j} = \begin{cases} 
+%     v_{pq} &\textrm{if}~(p=j)~\textrm{or}~(q=j)\\
+%     0 & \textrm{otherwise}
+%    \end{cases},
+% \end{equation}
+% \begin{equation}
+%  [\bP]_{\alpha_{pq},j} =  \frac{\partial v_{pq}}{\partial \varphi_j} = \begin{cases}
+%                                                                         i v_{pq} &\textrm{if}~ p=j\\
+%                                                                         -i v_{pq} &\textrm{if}~ q=j \\
+%                                                                         0 & \textrm{otherwise} 
+%                                                                        \end{cases}
+% \end{equation}
+% \begin{equation}
+%  [\bQ]_{\alpha_{pq},j} =  \frac{\partial v_{pq}}{\partial \widetilde{\eta}_j} = \begin{cases} 
+%     v_{pq} &\textrm{if}~\phi_{pq}=j\\
+%     0 & \textrm{otherwise}
+%    \end{cases},
+% \end{equation}
+% \begin{equation}
+%  [\bR]_{\alpha_{pq},j} =  \frac{\partial v_{pq}}{\partial \widetilde{\varphi}_j} = \begin{cases} 
+%     i v_{pq} &\textrm{if}~\phi_{pq}=j\\
+%     0 & \textrm{otherwise}
+%    \end{cases}.
+% \end{equation}
+% Augmenting the residual by taking into account its conjugate results in 
+% \begin{equation}
+% \breve{\br} = \bJ \Delta \bvarrho + \bn, 
+% \end{equation}
+% where $\breve{\br} = \begin{bmatrix}\br^T & \conj{\br}^T \end{bmatrix}^T$ and 
+% \begin{equation}
+% \label{eq:Jac_lin}
+% \bJ = 
+% \begin{bmatrix}
+%  \bmJ\\  
+%  \conj{\bmJ}
+% \end{bmatrix}.
+% \end{equation}
+% The solution to 
+% \begin{equation}
+%  \min_{\Delta \bvarrho} \| \breve{\br} - \bJ \Delta \bvarrho\|_F^2,
+% \end{equation}
+% is equal to 
+% \begin{equation}
+% \label{eq:LINCAL_update}
+% \Delta \bvarrho = (\bJ^H\bJ)^{-1} \bJ^H\breve{\br}. 
+% \end{equation}
+% equation~\ref{eq:LINCAL_update} is used to iteratively determine a new estimate of the parameter vector $\bvarrho$ as follows
+% \begin{equation}
+% \label{eq:new_rho}
+% \bvarrho_{k+1} = \bvarrho_k + \Delta \bvarrho_k. 
+% \end{equation}
+% Note that we do not explicitly denote the $k$ index in equation~\ref{eq:LINCAL_update}.
+% Moreover, equation~\ref{eq:LINCAL_update} is the GN update associated with the following minimization problem \citep{Kurien2016}
+% \begin{equation}
+% \label{eq:least_squares_lincal}
+% \min_{\bmath{\varrho}} \|\breve{\br}\| = \min_{\bmath{\varrho}} \|\breve{\bd} - \breve{\bv}(\bmath{\varrho})\|.
+% \end{equation}
+
+% \section{Conjugate Gradient Method}
+% 
+% \subsection{Normal Equation}
+% \label{sec:normal}
+% Lets assume that instead of equation~\ref{eq:linear_system} we have a more general linear system, i.e.
+% \begin{equation}
+%  \bb = \bB\bx,
+% \end{equation}
+% where $\bb\in\mathbb{C}^K$, $\bx\in\mathbb{C}^P$  and $\bB\in\mathbb{C}^{K \times P}$, with $K > P$. We can use the normal equation 
+% \begin{equation}
+% \label{eq:normal_equation}
+% \bB^H\bB\bx = \bB^H\bb, 
+% \end{equation}
+% to estimate the $\bx$ that minimizes
+% \begin{equation}
+% \label{eq:norm_ls}
+% \|\bb-\bB\bx\|_F^2. 
+% \end{equation}
+% In this paper, we denote the Hermitian transpose with $(*)^H$ and the Frobenius norm with $\|*\|_F$.
+% 
+% There are a few important observations that we can make while inspecting equation~\ref{eq:normal_equation}:
+% \begin{enumerate}
+% \item $\bB^H\bB$ is a square positive semi-definite Hermitian matrix by construction.
+% \item $\bB^H\bb$ is in the column range of $\bB^H\bB$; this follows trivially from 
+% \begin{equation}
+% \bB^H\bb \in \mathcal{R}(\bB^H) \implies \bB^H\bb \in \mathcal{R}(\bB^H\bB).   
+% \end{equation}
+% \end{enumerate}
+% In this paper, we denote the column range of a matrix $\bA$ with $\mathcal{R}(\bA)$.
+% 
+% Recall that we mentioned that in general we can only solve equation~\ref{eq:linear_system} with CG if $\bA$ is a square positive-definite Hermitian matrix.
 The CG method actually has an even broader applicability: it can be used to solve equation~\ref{eq:linear_system} even if $\bA$ is a Hermitian positive semi-definite
-square matrix as long as $\bb$ is in the column range of $\bA$ \citep{Lu2015}.
+% square matrix as long as $\bb$ is in the column range of $\bA$ \citep{Lu2015}.
 In light of \citet{Lu2015}, the two observations in the numbered list we presented in the previous paragraph imply that equation~\ref{eq:normal_equation}
-can be solved by employing the CG method. The solution returned by CG is, however, only unique if $\bB^H\bB$ is positive-definite (invertible). The solution
-obtained by CG, whether it is unique or not, does minimize equation~\ref{eq:norm_ls} in a least squares sense.
-
-Note that if $\bA$ is positive semi-definite, its smallest eigenvalue is zero. This implies that equation~\ref{eq:kappa} is not defined. \citet{Lu2015} show that when $\bA$
-is a square positive semi-definite Hermitian matrix we do not use the spectral condition number to bound the convergence complexity of CG, but the general spectral condition
-number instead. The general spectral condition number is very similar to the spectral condition number; the only difference being that in equation~\ref{eq:kappa} we replace  
-$\lambda_{\textrm{min}}$ with $\lambda_{\textrm{min}}^+$, which denotes the smallest positive eigenvalue.
+% can be solved by employing the CG method. The solution returned by CG is, however, only unique if $\bB^H\bB$ is positive-definite (invertible). The solution
+% obtained by CG, whether it is unique or not, does minimize equation~\ref{eq:norm_ls} in a least squares sense.
+% 
+% Note that if $\bA$ is positive semi-definite, its smallest eigenvalue is zero. This implies that equation~\ref{eq:kappa} is not defined. \citet{Lu2015} show that when $\bA$
+% is a square positive semi-definite Hermitian matrix we do not use the spectral condition number to bound the convergence complexity of CG, but the general spectral condition
+% number instead. The general spectral condition number is very similar to the spectral condition number; the only difference being that in equation~\ref{eq:kappa} we replace  
+% $\lambda_{\textrm{min}}$ with $\lambda_{\textrm{min}}^+$, which denotes the smallest positive eigenvalue.
 
@@
 
-The following identities,
+Furthermore,
 \begin{align}
@@
 \end{align}
-can now be trivially established by mechanically showing that the left hand side of each identity in equation~\ref{eq:identities} is equal to its right hand side.
+The above identities can be trivially established by mechanically showing that the left hand side of each expression in equation~\ref{eq:identities} is equal to its right hand side.
 In equation~\ref{eq:identities}.
@@
 \centering
-\caption{The dimensions of the matrices defined in Section~\ref{sec:analytic}.}
+\caption{The dimensions of the matrices defined in Appendix~\ref{sec:analytic}.}
 \begin{tabular}{|c c|} 
@@
 The basic skymodel-based \textsc{StEfCal} update step is equal to  the leftmost term in equation~\ref{eq:g_update} (barring $\alpha$) \citep{Salvini2014}.
-Assume without any loss of generality that the array is in an east-west regular grid. Furthermore, assume that $\boldsymbol{d}$ (see equation~\ref{eq:vis_red_vec}) has been re-ordered to
+Assume without any loss of generality that the array is in an east-west regular grid. Furthermore, assume that $\boldsymbol{d}$ (see equation~\ref{eq:vis_linear_definition}) has been re-ordered 
+and that the result of this re-ordering is the following
 \begin{equation}
@@
 
-equation~\ref{eq:vis_red_vec} can now be rewritten as 
+Equation~\ref{eq:vis_linear_definition} can now be rewritten as 
 \begin{equation}
@@
 \end{equation}
-Substituting equation~\ref{eq:RHR} and equation~\ref{eq:RHd} into equation~\ref{eq:y_final} and then write out the analytic expression of $y_i$ we obtain
-the leftmost term in equation~\ref{eq:y_update} (barring $\alpha$).
-
+Substituting equation~\ref{eq:RHR} and equation~\ref{eq:RHd} into equation~\ref{eq:y_final} and simplifying the result leads to the leftmost term in equation~\ref{eq:y_update} (if we bar $\alpha$ and consider
+only the $i$-th entry of $\by$).
 \label{lastpage}

### 2017-11-03 software/plot_func.py

@@ -574,7 +574,7 @@ def plot_prec_err_paper():
         precentage_stef_std1 = precentage_stef_std1[2:]
         output.close()
           
-        ax.plot(N1,precentage_stef_mean1*100,color[k],lw=2,label="SNR="+str(SNR[k]))
+        ax.plot(N1,precentage_stef_mean1*100,color[k],lw=2,label="SNR="+str(SNR[k])+" dB")
         if SNR[k] <> 1000:
            ax.fill_between(N1, (precentage_stef_mean1-precentage_stef_std1)*100, (precentage_stef_mean1+precentage_stef_std1)*100,alpha=0.2, edgecolor='k', facecolor=color[k])
 
@@ -582,7 +582,7 @@ def plot_prec_err_paper():
     ax.set_ylim([-5,30])
     ax.set_xlim([37,217])
     ax.set_xlabel("$N$ [antennas]")
-    ax.set_ylabel("% Error")
+    ax.set_ylabel(r"$\beta$")
     plt.grid('on')
     plt.show()
 
@@ -651,17 +651,17 @@ def plot_outerloop_pres():
     matplotlib.rcParams.update({'font.size': 22})
     fig = plt.figure()
     ax = fig.add_subplot(1,1,1)
-    ax.plot(N1,outerloop_pcg_mean1,"r",lw=2,label="PCG (SNR=1000)")
+    ax.plot(N1,outerloop_pcg_mean1,"r",lw=2,label="PCG (SNR=1000 dB)")
     ax.fill_between(N1,outerloop_pcg_mean1-outerloop_pcg_std1, outerloop_pcg_mean1+outerloop_pcg_std1,alpha=0.2, edgecolor='k', facecolor='r')
-    ax.plot(N1,outerloop_stef_mean1,"b",lw=2,label="ADI (SNR=1000)")
+    ax.plot(N1,outerloop_stef_mean1,"b",lw=2,label="Red StEfCal (SNR=1000 dB)")
     ax.fill_between(N1, outerloop_stef_mean1-outerloop_stef_std1, outerloop_stef_mean1+outerloop_stef_std1,alpha=0.2, edgecolor='k', facecolor='b')
-    ax.plot(N1,outerloop_pcg_mean2,"m",lw=2,label="PCG (SNR=5)")
[…]

### 2017-11-12 paper.tex
added the almost final version of the paper

@@
 \newcommand{\nat}{Nature}
+\newcommand{\apj}{The Astrophysical Journal}
+\newcommand{\prd}{Physical Review Letters D}
 \newcommand{\physrep}{Phys. Rep.}
+\newcommand{\araa}{Annual Review of Astronomy and Astrophysics}
+\newcommand{\pasp}{Publications of the Astronomical Society of Pacific}
+\newcommand{\procspie}{Proceedings of the SPIE}
+\newcommand{\pasa}{Publications of the Astronomical Society of Australia}
 
@@
 
-\title[Unknown]{Redundant calibration as a complex optimization problem}
-\author[T.L.~Grobler et al.]{T.L.~Grobler$^{1}$\thanks{E-mail: email@address (AVR)}, G.~Bernardi$^{1,2}$, Z.S.~Ali$^{3}$?, C.L.~Carilli$^{4,5}$?, J.S.~Dillon$^{3}$?, A.~Liu$^{3}$?, \newauthor
-A.R.~Parsons$^{3,6}$? and O.M.~Smirnov$^{1,2}$?\\
-$^{1}$Department of Physics and Electronics, Rhodes University, PO Box 94, Grahamstown, 6140, South Africa\\
-$^{2}$SKA SA, 3rd Floor, The Park, Park Road, Pinelands, 7405, South Africa\\
-$^{3}$Dept. of Astronomy, University of California, Berkeley, CA 94720, USA\\
+%\title[Complex redundant calibration]{Redundant calibration as a complex optimization problem}
+\title[Redundant interferometric calibration as a complex optimization problem]{Redundant calibration as a complex optimization problem}
+%\author[T.L.~Grobler et al.]{T.L.~Grobler$^{1}$\thanks{E-mail: email@address (AVR)}, G.~Bernardi$^{1,2}$, Z.S.~Ali$^{3}$?, C.L.~Carilli$^{4,5}$?, J.S.~Dillon$^{3}$?, A.~Liu$^{3}$?, \newauthor
+%A.R.~Parsons$^{3,6}$? and O.M.~Smirnov$^{1,2}$?\\
+%$^{1}$Department of Physics and Electronics, Rhodes University, PO Box 94, Grahamstown, 6140, South Africa\\
+%$^{2}$SKA SA, 3rd Floor, The Park, Park Road, Pinelands, 7405, South Africa\\
+%$^{3}$Dept. of Astronomy, University of California, Berkeley, CA 94720, USA\\
+%$^{4}$National Radio Astronomy Obs., Socorro NM\\
+%$^{5}$Cavendish Lab., Cambridge UK\\
+%$^{6}$Radio Astronomy Lab., U. California, Berkeley CA 94720, USA}
+\author[T.L.~Grobler et al.]{T.L.~Grobler$^{1,2}$\thanks{E-mail: email@address (AVR)}, G.~Bernardi$^{2,3}$, C.L.~Carilli$^{4,5}$, J.S.~Kenyon$^{2}$, A.R.~Parsons$^{6,7}$?
+\newauthor and O.M.~Smirnov$^{2,3}$\\
+$^{1}$Dept of Mathematical Sciences, Computer Science Division, Stellenbosch University, Private Bag X1, 7602 Matieland, South Africa\\
+$^{2}$Department of Physics and Electronics, Rhodes University, PO Box 94, Grahamstown, 6140, South Africa\\
+$^{3}$SKA SA, 3rd Floor, The Park, Park Road, Pinelands, 7405, South Africa\\
 $^{4}$National Radio Astronomy Obs., Socorro NM\\
 $^{5}$Cavendish Lab., Cambridge UK\\
-$^{6}$Radio Astronomy Lab., U. California, Berkeley CA 94720, USA}
+$^{6}$Dept. of Astronomy, University of California, Berkeley, CA 94720, USA\\
+$^{7}$Radio Astronomy Lab., U. California, Berkeley CA 94720, USA}
 \begin{document}
@@
 \begin{abstract}
-%In this paper we show that instead of splitting our redundant calibration problem into its phase-and-amplitude components (the more traditional approach) we can treat the complex unknown model parameters and their complex conjugates as separate unknowns.
+Observations of the redshifted 21-cm line from the epoch of reionization have recently motivated the construction of low 
+frequency radio arrays with highly redundant configurations that provide an alternative calibration strategy - 
+''redundant calibration" - and boosts sensitivity on specific spatial scales. In this paper we formulate 
+calibration of redundant interferometric arrays as a complex optimization problem \citep[i.e.,][]{Smirnov2015}. 
+We then solve this optimization problem via the Levenberg-Marquardt algorithm. This calibration approach is more 
+robust to initial conditions than current algorithms and, leveraging on an approximate matrix inversion, allows for further optimization and an efficient implementation (``redundant \textsc{StEfCal}").
+We also investigated the use of the preconditioned conjugate gradient method as an alternative optimization strategy, instead of
+an approximate matrix inverse, but found that its computational performance is not competitive with respect to ``redundant \textsc{StEfCal}".
+The efficient implementation of this new algorithm is made publicly available.
+%The construction of new radio interferometers, particularly operating at low frequencies and In this paper we show that instead of splitting our redundant calibration problem into its phase-and-amplitude components (the more traditional approach) we can treat the complex unknown model parameters and their complex conjugates as separate unknowns.
 We also derive the Levenberg-Maquardt and Gauss-Newton update steps associated with this alternative parameter-and-conjugate framework.
@@
 We make use of the parameter-and-conjugate framework in this paper, rather than the more traditional amplitude-and-phase approach as it makes it trivial to relate the alternating direction implicit method and the Levenberg-Marquardt algorithm with one another.
-Observations of the redshifted 21-cm line from the epoch of reionization have recently motivated the construction of low frequency radio arrays with highly redundant configurations that provide an alternative calibration strategy - ''redundant calibration" - and boosts sensitivity on specific scales.
-We show that redundant calibration can be formulated in the domain of complex functions of a complex variable.
-This framework reduces the
-\textbf{[GIANNI TO REWRITE AND FIX ABSTRACT]}
-has driven advancements in the calibration of radio interferometers and 
-Calibration of the new generation of low frequency radio interferometers faces new challenges due to their very wide field of views 
-The construction of new radio interferometers, particularly operating at low frequencies and 
-In this paper we show that instead of splitting our redundant calibration problem into its phase-and-amplitude components (the more traditional approach) we can treat the complex unknown model parameters and their complex conjugates as separate unknowns.
-We also derive the Levenberg-Maquardt and Gauss-Newton update steps associated with this alternative parameter-and-conjugate framework.
-We focus more on the Levenberg-Maquardt algorithm in this paper as it exhibits better convergence properties when compared to Gauss-Newton.
-The Gauss-Newton algorithm is currently the de facto standard that is being used to perform redundant calibration.
-It is common practice to implement redundant calibration via a hierarchical algorithmic strategy.
-Using the Levenberg-Maquardt algorithm, therefore, reduces the number of algorithmic layers that are needed to perform redundant calibration.
-We also compare two previously proposed methods to accelerate redundant calibration, namely the alternating direction implicit method and the preconditioned conjugate gradient method.
-We found that the alternating direction implicit method outperforms the preconditioned conjugate gradient method.
-However, the preconditioned conjugate gradient method is more robust as it is more straightforward to adapt to other calibration use cases.
-We make use of the parameter-and-conjugate framework in this paper, rather than the more traditional amplitude-and-phase approach as it makes it trivial to relate the alternating direction implicit method and the Levenberg-Marquardt algorithm with one another.
-\textbf{[Could we add JS Kenyon and SJ Wijnholds to author list?]}
 \end{abstract}
 
-%that \textsc{lincal} is effectively the non-linear least-squares gradient-based Gauss-Newton algorithm.
-Therefore, an easy way to reduce the number of layers used in performing redundant calibration is to employ the Levenberg-Marquardt algorithm instead (which would improve the convergence properties of stock standard \textsc{lincal})
-
 \begin{keywords}
-circumstellar matter -- infrared: stars.
+instrumentation: interferometers -- methods: data analysis -- methods: numerical --techniques: interferometric -- cosmology: observations
 \end{keywords}
@@
 \section{Introduction}
-The effects (i.e.
-the ionosphere, antenna gains, etc.) a celestial signal experiences along its propagation path are modelled with the radio interferometric measurement equation (RIME) \citep{ME1,RRIME1}.
-Calibration is the procedure by which we remove the errors that these propagation effects introduce into interferometric data and is normally accomplished using a 
-least-squares solver.
-In the case where we only model the antenna gains, which is the most basic form of the RIME, we can define calibration as finding the antenna gains that minimize the difference between our observed and predicted visibilities.
-
-There are two recent developments in skymodel-based calibration which should be explicitly mentioned at this point, since they form the foundation this paper builds on.
-The first is the discovery of \textsc{StEfCal} (Statistically Efficient Calibration) \citep{Mitchell:MWA-cal,Salvini2014}, which
-is an alternating direction implicit (ADI) skymodel-based calibration algorithm. It works by assuming the gains are known while solving for their conjugates and then performing the reverse. 
-This procedure results in the linearization of the calibration algorithm and reduces the computational complexity of calibration from $\mathbb{O}(N^3)$ to $\mathbb{O}(N^2)$, where $N$ denotes the number of antennas in the array.
-The second development is the idea: instead of breaking the calibration problem 
-into its real and imaginary parts and then solving for the real and imaginary parts of the unknown parameters separately we can treat the antenna gains and their 
-conjugates as separate parameters to estimate \citep{Smirnov2015}. \citet{Smirnov2015} also showed that the \textsc{StEfCal} algorithm itself is merely a special case of this
-gain-and-conjugate framework.
-
-Two baselines are redundant if their baseline difference vectors have the same length and orientation, which imply that they measure the same $uv$-modes.
-When we use a redundant array layout, we dramatically
-reduce the number of unknowns. If our array is redundant enough we reduce the number of unknowns to such an extent that we are able to solve for the unknown 
-observed visibilities themselves in addition to the antenna gains. Redundant calibration is the procedure by which we estimate both the unknown observed visibilities and the antenna gains.
-
-Redundant calibration was popularized in the 1980's. \citet{Noordam1982} used redundant calibration and the WSRT (Westerbork Synthesis Radio Telescope) to produce high-dynamic range images of 3C84.
-The first implementations of redundant calibration made use of the logarithm function, because when the logarithm is applied to the scalar measurement equation, redundant 
-calibration is transformed into a linear problem \citep{Wieringa1992,Camps2003,Liu2010}. The implementation associated with this approach is commonly referred to as \textsc{logcal}.
-We briefly discuss \textsc{logcal} in Appendix~\ref{sec:logcal}.
-The next improvement in redundant calibration was to perform a first order Taylor expansion, which is equivalent to the Gauss-Newton (GN) algorithm \citep{Kurien2016}, of the measurement equation and then to solve for the
-perturbations of the model-parameters \citep{Liu2010}.
-\citet{Liu2010} perform the first order Taylor expansion of the measurement equation by splitting the unknown parameters into their phase-and-amplitude
-components. The implementation of  this approach is called \textsc{lincal}. \textsc{lincal} is derived from first principals in Appendix~\ref{sec:lincal}. 
-Currently, redundant calibration is usually implemented using a layered algorithmic strategy.
-\textsc{logcal} and \textsc{lincal} are both implemented in \textsc{omnical}\footnote{https://github.com/jeffzhen/omnical}, which is a layered redundant calibration pipeline \citep{Zheng2014,Ali2015}.
-The solutions of \textsc{logcal} are used as initial input for \textsc{lincal}, since the GN algorithm requires that its starting parameter guesses be in the vicinity 
-of the true solution.
-
-It is also interesting to mention here that a completely different calibration strategy can also be used to calibrate redundant arrays, \textsc{CorrCal} \citep{Sievers2017}\footnote{https://github.com/sievers/corrcal2}.
-In \textsc{CorrCal} the unknown
-visibilities are no longer solved for. The sky is actually modelled, via two covariance matrices. 
-The first covariance matrix is associated with known point sources. The second covariance matrix models the remaining sky in a statistical sense. The unknown gains, however, are still 
-estimated. We do not elaborate any further on this approach in this paper.
-
-New instruments like LOFAR (Low Frequency Array) \citep{Noorishad2012}, PAPER (Precision Array for Probing the Epoch of Reionization) \citep{Ali2015} and HERA (Hydrogen Epoch of Reionization Array) \citep{deboer2015} have also sparked a renewed interest in redundant calibration.
-Using redundancy is especially advantageous when we are trying to detect the EoR (Epoch of Reionization) 21 cm signal,
-which is inherently quite faint \citep{Parsons2012}. Since redundant baselines make statistically independent measurements of the same $uv$-modes redundancy can greatly 
-improve the SNR (signal-to-noise ratio) of EoR detection experiments. This explains why PAPER and HERA are highly redundant arrays, as they are both primarily used for EoR 21 cm detection experiments.
-
-\noindent
-\textbf{[GIANNI TO REWRITE AND EXTEND EOR SECTION]}
-
-In this paper we propose to use the complex framework introduced by \citet{Smirnov2015} instead of the traditional amplitude-and-phase approach used by \textsc{lincal}.
-We also propose to use the Levenberg-Marquardt (LM) gradient-based algorithm instead of the Gauss-Newton (GN) algorithm as it exhibits better convergence properties.
-Using the LM algorithm reduces the number of algorithmic layers that are needed to perform redundant calibration.
-
-Two ways of speeding up redundant calibration have recently been proposed. The first technique makes use of the ADI method, in which we alternate between solving the gains, their conjugates and 
-the true observed visibilities \citep{Wijnholds2012,Marthi2014}. Note that the fact that \citet{Marthi2014} make use of the ADI method is not immediately clear
-from the paper itself.
-We will elaborate on this in Section~\ref{sec:adi}.
-The second is to use the preconditioned conjugate gradient (PCG) method.
-PCG allows us to exploit the sparsity inherent in the redundant calibration problem
-to speed it up \citep{Liu2010}. A proper comparison between these two approaches has not been done before.
-Therefore, we also compare the ADI \citep{Marthi2014} and PCG \citep{Liu2010} methods in this paper.
-We decided to focus on the complex framework proposed by \citet{Smirnov2015} rather than the traditional phase-and-amplitude approach as it
-allows us to trivially prove that the ADI method proposed by \citep{Marthi2014} is merely a special case of the more general LM algorithm.
-
-%Moreover, this framework also 
-%allows us to show that the ADI method proposed by \citep{Marthi2014} can for all intents and purposes be renamed to redundant \textsc{StEfCal} \citep{Mitchell:MWA-cal,Salvini2014}.
-
-In Section~\ref{sec:sky_wirtinger}, we review the skymodel-based complex-and-conjugate calibration approach proposed by \citet{Smirnov2015}.
-In Section~\ref{sec:red_wirtinger}, we show that this same approach can also be used to perform redundant calibration.
-In Section~\ref{sec:results}, we compare the computational complexity of two recently proposed redundant calibration speedup algorithms with one another.
-We end the paper with some final conclusions.
-
-% Main contributions of the paper:
-% 
-% \begin{enumerate}
-%  \item Present a general framework which unifies all the techniques developed so far showing they are all related (\textsc{lincal}, non-linear estimator, etc ...). They are all non-linear
-%  least-squares techniques, i.e. they either employ Gauss-Newton or the Levenberg-Marquardt algorithms. We have to start with \textsc{lincal} showing that it is GN, then we have to motivate 
-%  why we want to use Oleg's complex optimization framework.
-%  \item Use Oleg's complex optimization framework to re-derive the non-linear technique proposed by Marthi and Chengular. The novelty lies, in the fact that by using Oleg's
-%  framework we can find analytic expressions for the Jacobian, the Hessian and the Jacobian-residual product (which is not even the case for \textsc{lincal}). State that the algorithm is 
-%  effectively related to SteFCal and is eff an independent rediscovery of SteFCal and an extension of it into the redundant domain.
-%  \item Also we present the array geometry function to help us make the derivations from Marthi and Chengular easier to read and understand for a general layout.
-%  \item We also mention at this point that in Oleg's paper the question is raised is there a fast way of computing the exact inverse, we then present this new method, which is called
-%  conjugate gradient method. We discuss the algorithm and the two things which bound its execution time. Which is $kappa$ and $m$ (spectral condition number and its sparsity). Then
-%  we explore both of these parameters.
-%  \end{enumerate}
-%  
-%  **************
-%  FLOW OF PAPER
-%  **************
-%  
-%  \begin{enumerate}
-%  \item We need the definition of visibilities as in Liu. DEFINE SNR HERE ALREADY together maybe with the sigma value of the noise.
-%  \item Introduce the array geometery function.
-%  \item Write down logcal and lincal in matrix form... ?
-%  \item Short discussion of least squares and jacobian and hessian's. General GN and LM update rules.
-%  \item Introduce redundant calibration as a least squares problem. We will use this general fact to derive both popular methods.
-%  \item Propose a possible solution witch leads to lincal. Mention that this approach works as in this form the function is differentiable (a taylor expansion in the 
-%  parameters exists). Maybe mention that we can also divide the problem into real and imaginary etc...
-%  \item Show how this relates to lincal for example ---> show lincal is GN.
-%  \item Now introduce complex optimization ---> Main motivation for switching to the alternative framework is that the the differentials are very simplistic. We wish to show that
-%  we can derive the method of Chengular.
-%  \item Do the derivation of Chengular. 
-%  \item Brief discussion abouth Chengular and SteFCal and Complex Optimization. Here I show that in Oleg's paper he re-derives an algorithm called SteFCal. Stefcal works
-%  on the basis of alternating direction implicit method. The first signs of achieving a similar algorithm already appeared in Stefan's conference paper in which the alternating
-%  idea was first proposed. Then I mention Chegular re-derives the expression by using partial derivatives and extends it to redundant. One could also have used the linear alternating
-%  approach. In an attempt to merge the terminology that has independently develop in the general calibration and redundant calibration literature and to emphasize the close
-%  relationship between Stefcal and the approach derived by I will use the label Redunandat SteFCal (R-Stefcal) to refer to the NLS method proposed by ... Lastly we mention that similarly
-%  to how Oleg re-derived stefcal in, we have achieved the same approach.
-%  \item Faster Exact inverse. A question that Oleg poses in his paper is, does there exist a faster way to take the exact inverse of JhJ? One that is almost linear, and
-%  implies that we can therefore implement the full LM algorithm. The aim here is of course to reduce the number of iterations that are required to converge by using the 
-%  full inverse instead which would hopefully provide enough of a speedup to compensate for the more expensive full-inverse. The algorithm we propose is the conjugate 
-%  gradient method.
-%  \item We give the images of the HESSIAN of both lincal and the complex method (number of terms). So we can mention that both are sparse and contain diagonal entries that 
-%  are more significant than the off-diagonal entries. What linear inversion approach can take advantage of both these phenomenon. One such technique is 
-%  cg. 
-%  \item Briefly discuss CG and how its computationally bounded by its condition number and its sparsity (how does the diagonal play a role).
-%  \item Simulation description
-%  \item Will CG improve things?
-%  \item Now first discuss the condition
-%  number of the Hessian before and after pre-conditioning (pre-codnitioning can only be applied if a good inverse of a matrix is known, if it is known then it can improve 
-%  the spectral condition number of a matrix. We present here the kappa and iteration number graphs for the HEXAGONAL layout. Although the
-%  \item Now we discuss the sparsity results. 
-%  \item Provide a table that theoretical compares R-StEFCal and SPARC.
-%  \item Number of outerloop iterations. 
-%  \item Timing results.
-%  \item Accuracy results.
-%  \item Maybe some freq simulations.
-%  \end{enumerate}
-%  
-%  We
-
-{\bf GB: ok here are my comments/questions.
-\begin{itemize}
-\item I think we wanna start with Section 2 being a short review of standard calibration but in the complex optimization framework.
-Therefore I would begin with section 3.1, following by section 2.2.
-In particular I think that the discussion of least squares should be connected to (current) eq.
-15;
-\item we need more context for section 3.2.
-Why do we need to introduce a different indexing?
-This should be motivated.
-As it stands, I think it should remain where it is if we can make a connection to standard calibration, otherwise it should go in section 3.3;
-\item section 3.3 should merge with 4 where redundant calibration is introduced;
-\item section 3.5 goes together with the description of the simulations;
-\item section 4.1: I do not understand the definition in eq.
-32.
-How is the ratio of two matrices defined?
-I also suggest to use the symbol $\equiv$ rather than $=$ for definitions throughout the paper;
-\item can section 4.2.2 be moved to the appendix? Or how much of it?
-\item section 5 should be called something like "Implementation"...
-In Oleg's paper he writes "Inverting $J^H J$ and separability"...
-I find it a little too direct, but we need to find something else as these are not yet "Results".
-I also suggest to revert the order in the way this is presented: we first show that the Hessian is almost diagonal (is this only true in the complex represntation?), then we derive ADI and show that is the same as Marthi \& Chengalur;
-\item I have not seen the simulations where you inserted a phase wrap and you calibrate it. Is there a reason why it is not included?
-\end{itemize}
-}
-
-% \subsection{Least Squares}
-% \label{sec:ls}
-% The method of nonlinear least squares is tantamount to solving the following optimization problem:  
-% \begin{equation}
-% \label{eq:least_squares}
-% \min_{\bzeta} \Lambda(\bzeta) = \min_{\bzeta} \|\brho\|_F^2 = \min_{\bzeta} \|\bdelta - \bnu(\bzeta)\|_F^2, 
-% \end{equation}
-% where $\Lambda$ is known as the objective function, $\brho$ is the residual vector, $\bdelta$ is the observed data vector, $\bnu$ is the model-predicted data vector and $\bzeta$ is the model-parameter vector that minimizes Eq.~\eqref{eq:least_squares}.
-% In Eq.~\eqref{eq:least_squares}, we assume that the optimization problem is real valued.
-% The main objective of least squares is to find the model-parameters which minimize the squared difference between the observed data and the model-predicted data.
-% 
-% The most standard way of solving Eq.~\eqref{eq:least_squares} is to use a gradient-type minimization algorithm.
-In this paper we will be focusing on the following gradient minimization algorithms: GN and LM \citep{Levenberg1944,Marquardt1963}.
-% The GN update step is defined as
-% \begin{equation}
-%  \Delta \bzeta = (\bJ^T\bJ)^{-1} \bJ^T\brho,
-% \end{equation}
-% while the LM update takes the following form
-% \begin{equation}
-%  \Delta \bzeta = (\bJ^T\bJ+\lambda\bD)^{-1} \bJ^T\brho,
-% \end{equation}
-% where $\bJ$ is the Jacobian matrix (i.e. the derivative of the model $\bnu$ with respect to the parameter vector $\bzeta$), $\lambda$ denotes the 
-% damping factor and $\bD$ is a diagonal matrix whose diagonal entries are equal to the diagonal entries of $\bJ^T\bJ$. In this paper, we denote 
-% matrix transposition with $(*)^T$.
-% These update steps can be used in an iterative manner to determine new estimates of $\bzeta$.
-% 
-% The gradient based minimization algorithms that are generally used to solve least squares problems (i.e. GN and LM) generally require the model $\bnu$ to be differentiable
-% towards each model parameter. When the least squares problem is complex it becomes less straightforward to apply these gradient based minimization methods, as
-% many complex functions are not analytic (no Taylor expansion around a fiducial point exists) and therefore not differentiable if the classic notion of differentiation is used. For instance, if the 
-% classical definition of differentiation is used then $\frac{\partial \conj{z}}{\partial z}$, where $z \in \mathbb{C}$, does not exist.
-% 
-% The only way to circumvent the differentiability conundrum associated with complex least square problems is to recast our complex optimization problem in such a way that $\bnu$ becomes analytic in its argument.
-% The standard way of achieving this is to divide the complex optimization problem into its real and imaginary parts and then to solve for the real and imaginary parts of the parameters separately.
-% 
-% Recently, \citet{Sorber2012} proposed an alternative strategy, which involves the use of Wirtinger calculus \citep{Wirtinger1927}. The Wirtinger derivatives 
-% are defined to be 
-% \begin{align}
-% \label{eq:wir}
-% \frac{\partial}{\partial z} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} -  i \frac{\partial}{\partial y} \right ),&\frac{\partial}{\partial \conj{z}} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} +  i \frac{\partial}{\partial y} \right ).
-% \end{align}
-% Using the above definitions we can now easily compute the following:
-% \begin{align}
-% \frac{\partial z}{\partial z} & = 1, & \frac{\partial \conj{z}}{\partial z}&=0, & \frac{\partial z}{\partial \conj{z}} & = 0, & \frac{\partial \conj{z}}{\partial \conj{z}}&=1.
-% \end{align}
-% In this alternative approach we treat the complex parameters and their conjugates as separate variables.
-
-\section{Skymodel-based Wirtinger Calibration}
+The quest for the redshifted 21-cm line from the Epoch of Reionization (EoR) is a frontier of modern observational cosmology and has motivated the construction of a series of new interferometric arrays operating at low frequencies over the last decade.
+EoR measurements are challenging because of the intrinsic faintness of the cosmological signal \citep[see, for instance,][for recent reviews]{Furlanetto2016,McQuinn2016} buried underneath foreground emission which is a few orders of magnitude brighter than the EoR anywhere in the sky \citep[e.g.,][]{Bernardi2009,Bernardi2010,Ghosh2012,Dillon2014,Parsons2014}.
+%and that can be separated by leveraging only on the different spectral properties (citations needed!!!). 
+As miscalibrated foreground emission leads to artifacts that can jeopardize the EoR signal \citep[e.g.,][]{grobler2014,barry2016,ewall-wice2016}, an exquisite interferometric calibration is required.
+Such calibration needs an accurate knowledge of both the sky emission and the instrumental response \citep[e.g.,][]{Smirnov2011c}, unless arrays are deployed in a redundant configuration, i.e.
+with receiving elements placed on regular grids so that they measure many times the same sky brightness emission.
+Leveraging on this property, redundancy is a promising path to achieve high accurate interferometric calibration \citep[][]{Noordam1982,Wieringa1992,Pearson1984,Liu2010,Noorishad2012,Marthi2014,Sievers2017}.
+Moreover, redundant configurations provide a sensitivity boost on a selected sample of spatial scales that can be tuned to be the highest signal--to--noise ratio (SNR) EoR modes \citep[][]{Parsons2012,Dillon2016}.
+These reasons motivated the design and deploment of EoR arrays in redundant configurations like the MIT-EoR \citep{Zheng2014}, the Precision Array to Probe the Epoch of Reionization \citep[PAPER,][]{Ali2015} and the Hydrogen Epoch of Reionization Array \citep[HERA,][]{deboer2017}.
+
+In this paper we present a new algorithm to calibrate redundant arrays based on the complex optimization formalism recently introduced by \cite{Smirnov2015}. 
+With respect to current algorithms, it is more robust to initial conditions, while remaining comparatively fast.
+We also show that given certain approximations this new algorithm reduces to the redundant calibration equivalent of the \textsc{StEfCal} algorithm \citep{Salvini2014}.
+%In this paper we will refer to this simplified version of the new algorithm as redundant StEfCal. 
+We investigate the speed-up that the preconditioned conjugate gradient (PCG) method provides if it is employed by the new algorithm \citep{Liu2010}.
+A comparison between the computational complexity of the optimized new algorithm and redundant StEfCal is also performed.
+
+The paper is organized as follows: we review the complex calibration formalism by \citet{Smirnov2015} in Section~\ref{sec:sky_wirtinger}, we extend it to the redundant case in Section~\ref{sec:red_wirtinger}, we present some computational complexity results in Section~\ref{sec:pcg} and we conclude in Section~\ref{sec:conclusions}.
+
+%\section{Skymodel-based Wirtinger Calibration}
+%\section{Calibration described through the Wirtinger formalism}
+\section{Wirtinger calibration}
 \label{sec:sky_wirtinger}
-%In Section~\ref{sec:s_mes}, we present the scalar measurement equation. Next, we define some indexing functions in Section~\ref{sec:indexing}. The measurement 
-%equation and the indexing functions from Section~\ref{sec:indexing} are then used in Section~\ref{sec:r_mes} to phrase redundant calibration as a least-squares problem, i.e.
-Eq.~\eqref{eq:least_squares_red}.
-%We present the signal-to-noise ratio (SNR) definition we used in this paper in Section~\ref{sec:snr}. The limitations of this paper are discussed in Section~\ref{sec:scope}.
-
-%\subsection{Scalar Measurement Equation}
-%\label{sec:s_mes}
-The observed visibility $d_{pq}$ measured by the baseline formed by antenna $p$ and $q$ is described by
+
+In a radio interferometer, the true sky visibilities $y_{pq}$ measured by a baseline formed by antenna $p$ and $q$ are always ``corrupted" by the non-ideal response of the receiver, which is often incorporated into a single, receiver-based complex number $g$.
+The observed visibility $d_{pq}$ is therefore given by \citep{ME1,ME2,RRIME1}
 \begin{equation}
 \label{eq:vis_definition}
-d_{pq} = g_{p}\conj{g_q}y_{pq} + n_{pq},
-\end{equation}
-where $g_{p}$ and $g_{q}$ are the direction independent gains associated with antenna $p$ and $q$, $y_{pq}$ denotes the true visibility that baseline $pq$ measured
-and $n_{pq}$ is the thermal noise component. Conjugation is denoted by $\conj{(*)}$. The true value of $g_p$, $g_q$ and $y_{pq}$ are unknown. We estimate the true value of
-$g_p$, $g_q$ and $y_{pq}$ by calibrating our data and then imaging the result.
-
-The real and imaginary components of the thermal noise are normally distributed with a mean of zero and a standard deviation $\sigma$, where   
-\begin{equation}
-\sigma \propto \frac{T_{\textrm{sys}}}{\sqrt{\Delta \nu \tau}}. 
-\end{equation}
-Here, $T_{\textrm{sys}}$ is equal to the system temperature, $\Delta \nu$ is the observational bandwidth and $\tau$ is the integration time per visibility.
-
-Eq.~\eqref{eq:vis_definition} can also be expressed in the following vector form 
-\begin{equation}
-\label{eq:vis_definition}
+d_{pq} = g_{p}\conj{g_q} \, y_{pq} + n_{pq},
+\end{equation}
+where $\conj{(*)}$ indicates conplex conjugation and $n_{pq}$ is the thermal noise component. 
+The real and imaginary components of the thermal noise are normally distributed with a mean of zero and a
+standard deviation $\sigma$:
+\begin{equation}
+\sigma \propto \frac{T_{\textrm{sys}}}{\sqrt{\Delta \nu \tau}},
+\end{equation}
+where $T_{\textrm{sys}}$ is equal to the system temperature, $\Delta \nu$ is the observational bandwidth and $\tau$ is the integration time per visibility.
+
+Considering the number of visibilities $B$ (i.e. baselines) measured by an array of $N$ elements $B = \frac{N^2-N}{2}$, equation~\ref{eq:vis_definition} can be expressed in the following vector form:
+\begin{equation}
+\label{eq:vis_linear_definition}
 \bd = \bv + \bn, 
@@
 \end{equation}
-The function $\alpha_{pq}$ maps composite antenna indexes to unique single indexes, i.e. 
-\begin{equation}
-\{\alpha_{12},\alpha_{13},\cdots,\alpha_{1N},\alpha_{23},\cdots,\alpha_{N-1N}\}
-\end{equation}
-and 
-\begin{equation}
-\{1,2,\cdots,N-1,N,\cdots,B\}
-\end{equation}
-are equal. We denote the number of baselines in an array with $B$, i.e. $B = \frac{N^2-N}{2}$. The vectors in Eq.~\eqref{eq:vec_definitions} are column vectors of size $B$ (i.e. $p<q$).
-
-We can now formulate calibration as a least-squares optimization problem:
+The function $\alpha_{pq}$ therefore maps composite antenna indexes to unique single indexes, i.e:
+\begin{equation}
+\{\alpha_{12},\alpha_{13},\cdots,\alpha_{N-1N}\} = \{1,2,\cdots,B\} 
+\end{equation}
+The vectors in equation~\ref{eq:vec_linear_definitions} are column vectors of size $B$ (i.e. $p<q$).
+
+Radio interferometric calibration aims to determine the best estimate of $\bg = [g_1,g_2,\cdots,g_N]^T$ in order to correct the data and, following equation~\ref{eq:vis_linear_definition}, can be formulated as a non linear least-squares optimization problem:
 \begin{equation}
@@
 \end{equation}
-where $\Lambda$ is the objective function, $\br$ is the residual vector and $\bg =[g_1,\cdots,g_N]^T$.
-
-The gradient-based minimization algorithms that are generally used to solve least squares problems generally require the model (in Eq.~\eqref{eq:least_squares} the model-predicted vector is denoted by $\bv$) to be differentiable
-towards each model parameter (in Eq.~\eqref{eq:least_squares} the model-parameter vector is denoted by $\bg$), i.e.
-these methods make use of a Jacobian matrix $\bJ$.
-In this paper we will be focusing on the following gradient-based minimization algorithms: GN and LM \citep{Levenberg1944,Marquardt1963}.
-When the least squares problem is complex it becomes less straightforward to apply these gradient-based minimization methods, as
-many complex functions are not analytic (no Taylor expansion around a fiducial point exists) and therefore not differentiable if the classic notion of differentiation is used. For instance, if the 
-classical definition of differentiation is used then $\frac{\partial \conj{z}}{\partial z}$, where $z \in \mathbb{C}$, does not exist.
-
-The only way to circumvent the differentiability conundrum associated with complex least squares problems is to recast our complex optimization problem in such a way that $\bv$ becomes analytic in its argument.
-The standard way of achieving this is to divide the complex optimization problem into its real and imaginary parts and then to solve for the real and imaginary parts of the unknown model-parameters separately.
-
-Recently, \citet{Sorber2012} proposed an alternative strategy, which involves the use of Wirtinger calculus \citep{Wirtinger1927}. The Wirtinger derivatives 
-are defined to be 
+where $\Lambda$ is the objective function, $\br$ is the residual vector and $\|\cdot\|_F$ denotes the Frobenius norm.
+In standard interferometric calibration, $y_{pq}$ is assumed to be known at some level, for instance through the observation of previously known calibration sources.
+The knowledge of $y_{pq}$ is often named ``sky model".
+
+The gradient-based minimization algorithms that are generally used to solve non-linear least-squares problems are generally solved by using gradient-based minimization algorithms (i.e.
+Gauss--Newton -- GN -- or Levenberg--Marquardt -- LM) that require the model ($\bv$ in equation~\ref{eq:least_squares}) to be differentiable towards each parameter.
+When the least squares problem is complex, it becomes less straightforward to apply these gradient-based minimization methods, as
+many complex functions are not differentiable if the classic notion of differentiation is used, i.e. $\frac{\partial \conj{z}}{\partial z}$ does not exist if $z \in \mathbb{C}$.
+
+In order to circumvent the differentiability conundrum associated with complex least squares problems, standard interferometric calibration divides the complex optimization problem into its real and imaginary parts and solves for the real and imaginary parts of the unknown model parameters separately.
+\citet{Smirnov2015} showed, however, that this approach is not needed if complex calculus \citep{Wirtinger1927} is directly adopted.
+The Wirtinger derivatives are defined as:
 \begin{align}
 \label{eq:wir}
-\frac{\partial}{\partial z} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} -  i \frac{\partial}{\partial y} \right ),&\frac{\partial}{\partial \conj{z}} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} +  i \frac{\partial}{\partial y} \right ).
-\end{align}
-Using the above definitions we can now easily compute the following:
-\begin{align}
+\frac{\partial}{\partial z} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} -  i \frac{\partial}{\partial y} \right ),&\frac{\partial}{\partial \conj{z}} &= \frac{1}{2}\left ( \frac{\partial}{\partial x} +  i \frac{\partial}{\partial y} \right ),
+\end{align}
+which lead to the following relations:
+\begin{align}
+\label{eq:wir_z}
 \frac{\partial z}{\partial z} & = 1, & \frac{\partial \conj{z}}{\partial z}&=0, & \frac{\partial z}{\partial \conj{z}} & = 0, & \frac{\partial \conj{z}}{\partial \conj{z}}&=1.
 \end{align}
-If we use the Wirtinger derivative, Eq.~\eqref{eq:wir}, instead of the classic notion of a derivative to define the gradient operator then 
-$\bv$ becomes analytic in $\bz$ and $\conj{\bz}$ as a whole.
-The Wirtinger derivative can therefore be used to derive complex variants of the real-valued GN and LM algorithms.
-In these alternative algorithms we treat the complex parameters and their conjugates as separate variables.
-
-\citet{Smirnov2015} recently demonstrated that the Wirtinger derivative can be used to perform normal skymodel-based calibration. We briefly summarize their approach below:
-\begin{enumerate}
-
-\item Eq.~\eqref{eq:least_squares} is first recast to
+If the gradient operator is defined using equation~\ref{eq:wir}, the model $\bv$ now becomes analytic in both $\bg$ and $\conj{\bg}$ and equation \ref{eq:wir_z} can be used to derive the complex variants of the real-valued GN and LM algorithms.
+In the complex GN and LM algorithms, complex parameters and their conjugates are treated as separate variables.
+
+Assuming that $y_{pq}$ is known, equation~\ref{eq:least_squares} is recast as \citep{Smirnov2015}:
 \begin{equation}
@@
 \begin{align}
+\label{eq:jac_entries}
 [\bmJ]_{\alpha_{pq},i} &= \frac{\partial v_{pq}}{\partial g_i}, & [\bmJ^*]_{\alpha_{pq},i} &= \frac{\partial v_{pq}}{\partial \conj{g}_i}. 
 \end{align}
-Note that the Wirtinger derivative is used in Eq.~\eqref{eq:Jacobian_skymodel} and as such the entries of $\bJ$ are well defined.
-\item The complex LM update rule is very similar, the major difference being a damping factor $\lambda$ is introduced:
+
+The complex LM update is very similar, the major difference being a damping factor $\lambda$ is introduced:
 \begin{equation}
@@
 where $\bD=\bI\odot\bJ^H\bJ$.
-\item Eq.~\eqref{eq:GN_update_skymodel} or Eq.~\eqref{eq:LM_update_skymodel} can now be used iteratively to update the 
-parameter vector $\breve{\bg}$:
-\begin{equation}
-\label{eq:update}
-\breve{\bg}_{k+1} = \breve{\bg}_{k} + \Delta \breve{\bg}_{k}. 
-\end{equation}
-The parameter vector needs to be initialized (an all one vector is one possible choice) and is updated until convergence is reached.
-To improve readability we do not explicitly denote the iteration subscript $k$ in
-Eq.~\eqref{eq:GN_update_skymodel} and Eq.~\eqref{eq:LM_update_skymodel}.
-\end{enumerate}
-We refer to $\bJ^H\bJ$ as the Hessian matrix $\bH$ and to $\bJ^H\bJ + \lambda\bD$ as the modified Hessian matrix $\bmH$ in this paper. 
-% \subsection{Indexing Functions}
-% \label{sec:indexing}
-% In general there are two ways of indexing the true visibilities associated with different baselines.
-We can use the composite index $pq$, which is the strategy that is followed in Eq.~\eqref{eq:vis_definition},
-% or we can use a single unique baseline index which we assign in numerical order to each composite index, assuming the composite indexes are arranged in standard baseline order.
-% The composite indexes $pq$ are in standard baseline order when they are arranged as follow: $12, 13, 1N, 23\cdots$. Let 
-% \begin{equation}
-% \alpha_{pq} =
-% \begin{cases}
-% (q-p) + (p-1)\left (N-\frac{1}{2}p \right ) & \textrm{if}~p<q\\
-% %B + (p-q) + (q-1)\left (N-\frac{1}{2}q \right )) & \textrm{if}~p>q\\
-% %B+q + \frac{1}{2}(p-1)(p-2) & \textrm{if}~p>q\\
-% 0 & \textrm{otherwise}
-% \end{cases}.
-% \end{equation}
-% We can use $\alpha_{pq}$ to map composite indexes to unique baseline indexes, i.e. 
-% \begin{equation}
-% 12,13,\cdots,1N,23,\cdots \xrightarrow{\alpha_{pq}} 1,2,3,\cdots,B.
-% \end{equation}
-% We denote the number of baselines in the array with $B$, i.e. $B = \frac{N^2-N}{2}$.
-% 
-% Recall that redundant baselines sample the exact same visibilities in the $uv$-plane, i.e. if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$. 
-% Let us now group all the baselines that measure the same $uv$-modes in a redundant array into the same redundant baseline group.
-Assigning the baselines of a redundant array to different redundant baseline groups has the advantage that it reduces the number of single index values we require.
-We now only require single index values for each redundant baseline group instead of a unique single index label for
-% each baseline. Let $\phi_{pq}$ be the function that maps composite antenna pairs to the appropriate single redundant baseline group index. Note that $\phi_{pq}$ is not symmetric as 
-% $\phi_{pq} = 0$ if $p\geq q$. We can also define the following symmetric variant of $\phi_{pq}$:
-% \begin{equation}
-% \zeta_{pq} = 
-% \begin{cases}
-% \phi_{pq}~\textrm{if}~p \leq q\\
-% \phi_{qp}~\textrm{if}~p>q
-% \end{cases}.
-% \end{equation} 
-% It is possible to construct a simple analytic expression for $\zeta_{pq}$ when our array is in an east-west regular configuration, namely $\zeta_{pq} = |q-p|$. 
-% It becomes increasingly difficult to construct analytic expressions of $\zeta_{pq}$ for more complicated array layouts. 
-% We will refer to $\zeta_{pq}$ as the symmetric geometric function. The empirically constructed symmetric geometry
-% functions of three different redundant layouts are depicted in Fig.~\ref{fig:geometry_function}. We denote the range of $\zeta_{pq}$ with $\mathcal{R}(\zeta_{pq})$. The maximal element 
-% that $\zeta_{pq}$ can ascertain is denoted by $L$ and can be interpreted as the maximal number of unique redundant baseline groups which can be formed for a given 
-% array layout.
-Alternatively, we can interpret the symmetric geometry functions in Fig.~\ref{fig:geometry_function} as geometry matrices.
-In this alternative paradigm, $\zeta_{pq}$ denotes a matrix entry.
-The dimension of these geometry
-% matrices are $N\times N$. 
-% 
-% \begin{figure*}
-% \centering
-% \subfigure[Hexagonal layout]
-% {\includegraphics[width=0.32\textwidth]{./HEX_lay.pdf}\label{fig:HEX_lay}}
-% \subfigure[Square layout]
-% {\includegraphics[width=0.32\textwidth]{./SQR_lay.pdf}\label{fig:SQR_lay}}
-% \subfigure[Regular east-west layout]
-% {\includegraphics[width=0.32\textwidth]{./REG_lay.pdf}\label{fig:REG_lay}}
-% 
-% \subfigure[Hexagonal: $\zeta_{pq}$]
-% {\includegraphics[width=0.33\textwidth]{./HEX_phi.pdf}\label{fig:HEX_phi}}
-% \subfigure[Square: $\zeta_{pq}$]
-% {\includegraphics[width=0.33\textwidth]{./SQR_phi.pdf}\label{fig:SQR_phi}}
-% \subfigure[Regular east-west: $\zeta_{pq}$]
-% {\includegraphics[width=0.33\textwidth]{./REG_phi.pdf}\label{fig:REG_phi}}
-% \caption{In the top panel we have three different redundant antenna layouts, namely a hexagonal (left), square (middle) and a regular east-west (right) layout. 
-% In the bottom panel we have the respective symmetric redundancy geometry functions for the layouts on the top panel. We used 91 antennas
-% to construct the hexagonal layout, while 100 antennas where used in the square and east-west layouts.
-In the case of the east-west layout we only plot the positions of the first ten antennas.
-The maximal amount of redundant baseline groups that can be formed for
-% the hexagonal, square and east-west layouts are 165, 180 and 99 respectively.\label{fig:geometry_function}}
-% \end{figure*}
-% 
-% If the function $\phi_{pq}$ is known and one is given one of the two antenna indexes that together form a redundant baseline as well as the redundant baseline group index itself then it is possible 
-% to calculate the unknown antenna index. This can be computed via the following two expressions:
-% \begin{equation}
-% \xi_{ij} = 
-% \begin{cases}
-% p~\textrm{if}~\exists! ~ p \in \mathbb{N} ~ s.t. ~(\phi_{pi} = j)\\
-% 0~\textrm{otherwise}
-% \end{cases},
-% \end{equation}
-% and
-% \begin{equation}
-% \psi_{ij} = 
-% \begin{cases}
-% q~\textrm{if}~\exists! ~ q \in \mathbb{N} ~ s.t. ~(\phi_{iq} = j)\\
-% 0~\textrm{otherwise}
-% \end{cases}.
-% \end{equation}
-% We use $\xi_{ij}$ to retrieve the first antenna index of the composite antenna pair associated with a particular baseline, if the index of the first antenna in the composite antenna pair and the unique redundant baseline group index are known, while we use $\psi_{ij}$ to obtain
-% a similar result: the second antenna index of the composite antenna pair. %\textbf{NB::} I still need to plot $\zeta_{ij}$ for different antenna layouts.
+In this paper, the identity matrix is denoted by $\bI$.
+Moreover, the Hadamard product and the Hermition transpose are denoted by $\odot$ and $(\cdot)^H$.
+Note the use of the Wirtinger derivatives in equation~\ref{eq:jac_entries}.
+We will refer to $\bJ^H\bJ$ as the Hessian matrix $\bH$ and to $\bJ^H\bJ + \lambda\bD$ as the modified Hessian matrix $\bmH$ throughout this paper.
+
+Equation~\ref{eq:GN_update_skymodel} or~\ref{eq:LM_update_skymodel} can now be used iteratively to update the parameter vector $\breve{\bg}$:
+\begin{equation}
+\label{eq:update_skymodel}
+\breve{\bg}_{k+1} = \breve{\bg}_{k} + \Delta \breve{\bg}_{k},
+\end{equation}
+until convergence is reached.
+
+In the case of the GN algorithm, the parameter update step simplifies and becomes \citep{Smirnov2015}
+\begin{equation}
+\label{eq:one_half}
+\breve{\bg}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{1}{2}\breve{\bg}_{k}. 
+\end{equation}
+
+\citet{Smirnov2015} realized that the diagonal entries of the Hessian matrix $\bH$ are much more significant than its off-diagonal entries, i.e.
+$\bH$ is nearly diagonal.
+By approximating $\bH$ by its diagonal and substituting the approximate Hessian matrix the LM parameter update step becomes \citep{Smirnov2015}:
+\begin{align}
+\breve{\bg}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bg}_k, \nonumber \\
+ &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bg}_k, 
+\label{eq:stef_alpha}  
+\end{align}
+where $\alpha = \frac{1}{1+\lambda}$. Note that equation~\ref{eq:one_half} and equation~\ref{eq:stef_alpha} are   
+not dependent on $\breve{\br}$.
+
+Interestingly enough, if $\lambda = 0$ we obtain the odd parameter update step of \textsc{StEfCal}\footnote{$k\in\{0,2,\cdots\}$}, and if $\lambda=1$ (which corresponds
+to $\alpha=\frac{1}{2}$) we obtain the even parameter update step of \textsc{StEfCal}\footnote{$k\in\{1,3,\cdots\}$} \citep[\textsc{StEfCal},][]{Mitchell:MWA-cal,Salvini2014}.
+In the \textsc{StEfCal} algorithm, the measurement equation, equation~\ref{eq:vis_definition}, is linearized by assuming that the gains are known, but that their conjugates are not.
+Under this assumption, the system of equations become linear and the conjugates of the gains can be obtained straightforwardly.
+Starting from the latest value of the gain conjugates, an updated estimate of the gains can be obtained and the succession continues until convergence is reached.
+Alternating between solving and fixing different sets of parameters (which is exactly what \textsc{StEfCal} does) is in general referred to 
+as the alternating direction implicit method (ADI). The \textsc{StEfCal} algorithm reduces the computational complexity of calibration from $O(N^3)$ to $O(N^2)$.
 
@@
 \label{sec:red_wirtinger}
-Redundant baselines sample the exact same visibilities in the $uv$-plane, i.e. if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$. 
-We can therefore group all the baselines that measure the exact same visibilities into the same redundant baseline group. We index redundant baseline 
-groups with single group indices instead of composite antenna pairs. Let $\phi_{pq}$ be the function that maps the composite antenna pair associated with a specific baseline 
-to the single group index of the redundant baseline group to which the baseline in question belongs, i.e. if baseline $pq$ and $rs$ are redundant then $\phi_{pq} = \phi_{rs}$.  
-The function $\phi_{pq}$ is not symmetric as 
-$\phi_{pq} = 0$ if $p\geq q$. We can define the following symmetric variant of $\phi_{pq}$:
+Interferometric baselines are named redundant when they sample the exact same visibilities in the $uv$-plane, i.e.
+if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$.
+It is convenient to group redundant visibilities together and label each group using a single index rather than using their antenna pairs as in equation~\ref{eq:vis_definition}.
+We introduce a function $\phi_{pq}$ that maps the antenna pair associated with a specific baseline to the single group index of the redundant baseline group that includes the baseline in question, i.e.
+if baseline $pq$ and $rs$ are redundant then $\phi_{pq} = \phi_{rs}$.
+The function $\phi_{pq}$ is not symmetric as $\phi_{pq} = 0$ if $p \geq q$.
+Equation~\ref{eq:vis_definition} can be re-written for a redundant array as:
+\begin{equation}
+\label{eq:vis_red}
+d_{pq} = g_{p}\conj{g_q}y_{\phi_{pq}} + n_{pq},
+\end{equation}
+with the same vector form as equation~\ref{eq:vis_linear_definition} if
+\begin{align}
+ \left [ \bd \right]_{\alpha_{pq}} &= d_{pq}, & \left [ \bv \right ]_{\alpha_{pq}} &= v_{pq}=g_p y_{\phi_{pq}} \conj{g_q},\nonumber\\
+ \left [ \bn \right ]_{\alpha_{pq}} &= n_{pq}, &  &\label{eq:vec_definitions}
+\end{align}
+where the vectors in Equation~\ref{eq:vec_definitions} are column vectors of size $B$ (i.e. $p<q$).
+
+We also introduce the following symmetric variant of $\phi_{pq}$:
 \begin{equation}
@@
 \phi_{qp}~\textrm{if}~p>q
-\end{cases}.
-\end{equation} 
-If the array is redundant we can rewrite Eq.~\eqref{eq:vis_definition} as
-\begin{equation}
-\label{eq:vis_red}
-d_{pq} = g_{p}\conj{g_q}y_{\phi_{pq}} + n_{pq},
-\end{equation}
-Eq.~\eqref{eq:vis_red} can also be expressed in the following vector form 
-\begin{equation}
-\label{eq:vis_red_vec}
-\bd = \bv + \bn, 
-\end{equation}
-where 
-\begin{align}
- \left [ \bd \right]_{\alpha_{pq}} &= d_{pq}, & \left [ \bv \right ]_{\alpha_{pq}} &= v_{pq}=g_p y_{\phi_{pq}} \conj{g_q},\nonumber\\
- \left [ \bn \right ]_{\alpha_{pq}} &= n_{pq}. &  &\label{eq:vec_definitions}
-\end{align}
-The vectors in Eq.~\eqref{eq:vec_definitions} are column vectors of size $B$ (i.e. $p<q$). 
-It is possible to construct a simple analytic expression for $\zeta_{pq}$ when our array is in an east-west regular configuration, namely $\zeta_{pq} = |q-p|$. 
-It becomes increasingly difficult to construct analytic expressions of $\zeta_{pq}$ for more complicated array layouts. 
-We will refer to $\zeta_{pq}$ as the symmetric geometric function. The empirically constructed symmetric geometry
-functions of three different redundant layouts are depicted in Fig.~\ref{fig:geometry_function}. We denote the range of $\zeta_{pq}$ with $\mathcal{R}(\zeta_{pq})$. The maximal element 
-that $\zeta_{pq}$ can ascertain is denoted by $L$ and can be interpreted as the maximal number of unique redundant baseline groups which can be formed for a given 
-array layout.
-Alternatively, we can interpret the symmetric geometry functions in Fig.~\ref{fig:geometry_function} as geometry matrices.
-In this alternative paradigm, $\zeta_{pq}$ denotes a matrix entry.
-The dimension of these geometry
-matrices are $N\times N$.
+\end{cases},
+\end{equation} 
+and we will refer to $\zeta_{pq}$ as the symmetric geometric function.
+It is possible to construct a simple analytic expression for $\zeta_{pq}$ for an east-west regular array, i.e.
+$\zeta_{pq} = |q-p|$.
+It becomes, however, increasingly difficult to construct analytic expressions of $\zeta_{pq}$ for more complicated array layouts.
+The empirically constructed symmetric geometry functions of three different redundant layouts are displayed in Figure~\ref{fig:geometry_function}.
+We denote the range of $\zeta_{pq}$ with $\mathcal{R}(\zeta_{pq})$.
+The maximal element that $\zeta_{pq}$ can ascertain is denoted by $L$ and can be interpreted as the maximal number of unique redundant baseline groups which can be formed for a given array layout.
 
@@
  \end{align}
-The number of model parameters that we need to solve for in Eq.~\eqref{eq:least_squares_red} is $P = 2(N+L)$, since redundant calibration is a complex problem.
-We plot $L$ and $P$ for two different geometric layouts in Fig~\ref{fig:pl}.
-The analytic expressions for the red and the blue curve in Fig.~\ref{fig:ln} are \citep{Camps2003}
-\begin{equation}
-L = 2N-\frac{1}{2}\sqrt{12N-3}-\frac{1}{2} 
-\end{equation}
-and
-\begin{equation}
-L = N - 1 
-\end{equation}
-respectively.
-
-\begin{figure*}
-\centering
-\subfigure[$L$]{\includegraphics[width=0.47\textwidth]{./ln.pdf}\label{fig:ln}}
-% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
-\subfigure[$P$]{\includegraphics[width=0.47\textwidth]{./pn.pdf}\label{fig:pn}}
-\caption{In the left panel we plot the number of unique baseline groups $L$ one can form when using a hexagonal (red curve) and regular east-west layout (blue curve) as 
-a function of the number of antennas $N$ in the array.
-In the right panel we plot the total number of parameters $P$ we need to estimate when employing redundant calibration for a hexagonal (red curve) and regular east-west layout (blue curve) as
-a function of the number of antennas $N$ in the array. 
-\label{fig:pl}} 
-\end{figure*}
-
-% \subsection{Limitations}
-% \label{sec:scope}
-% We only deal with per-timeslot and frequency channel calibration. No solution smoothing  
-% is applied between the different timeslots and channels. The reader interested in this is referred to \citep{Zheng2014}. 
-% Moreover, we do not explicitly deal with any of the degeneracies normally associated with redundant calibration \citep{Zheng2014,Kurien2016}.
-Similar to the approach taken in \citep{Marthi2014}, we assume that the degeneracies present in the redundant calibration solutions are removed with a final round of skymodel-based calibration.
-
-As already alluded to in Section~\ref{sec:sky_wirtinger}, we can solve Eq.~\eqref{eq:least_squares_red} in one of two ways. We can split the problem into its real and
-imaginary parts and then solve for the real and imaginary parts of the unknown parameters separately or we can employ the notion of Wirtinger derivatives. The former 
-approach is discussed in Section~\ref{sec:ri}, while the latter approach is described in Section~\ref{sec:w}.
-
-\subsection{Real and Imaginary Approach}
-\label{sec:ri}
-Using least-squares to solve complex problems can be problematic as $\bv$ needs to be analytic in its argument. In the case 
-of Eq.~\eqref{eq:least_squares_red}, \citet{Liu2010} circumvents this difficulty by 
-splitting the problem into its phase-and-amplitude components.
-
-The aforementioned is achieved by first rewriting Eq.~\eqref{eq:least_squares_red} as
-\begin{equation}
-\label{eq:least_squares_lincal}
-\min_{\bmath{\varrho}} \|\breve{\br}\| = \min_{\bmath{\varrho}} \|\breve{\bd} - \breve{\bv}(\bmath{\varrho})\|, 
-\end{equation}
-where $\breve{\bd} = [\bd^T,\conj{\bd}^T]^T$, $\breve{\bv} = [\bv^T,\conj{\bv}^T]^T$ and $\bmath{\varrho}$ is the phase and amplitude parameter vector which we define in Appendix~\ref{sec:lincal}. 
-With the aid of Eq.~\eqref{eq:least_squares_lincal} we can derive the algorithm implemented in \textsc{lincal} (see Appendix~\ref{sec:lincal}).
-The Jacobian associated with Eq.~\eqref{eq:least_squares_lincal} is equal to Eq.~\eqref{eq:Jac_lin}. 
-Appendix~\ref{sec:lincal} makes it evident that \textsc{lincal} is equivalent to a basic GN update step \citep{Kurien2016}. 
-Therefore, a straightforward way of improving the convergence properties of stock standard \textsc{lincal} would be to simply add the LM damping factor to the basic \textsc{lincal} update step.
-In doing so the need to use $\textsc{logcal}$ before $\textsc{lincal}$ becomes superfluous. The use of $\textsc{logcal}$ may, however, still offer 
-a computational advantage, as it is a very fast method to obtain a good starting solution, and would reduce the number of iterations needed by the LM algorithm.
-Proper experiments would need to be conducted to determine whether it is worth keeping $\textsc{logcal}$ in the algorithmic hierarchy.
-
-\subsection{Wirtinger Approach}
-\label{sec:w}
-
-%\citet{Smirnov2015} recently proposed to use the Wirtinger derivative instead of the classic real and imaginary approach to perform normal skymodel-based calibration. We will now show
-%that the Wirtinger derivative can also be applied to redundant calibration.
-%We can perform redundant calibration by solving the following optimization problem: 
-%\begin{equation}
-%\label{eq:least_squares_red}
-%\min_{\bz} \Lambda(\bz) = \min_{\bz} \|\br\|_F^2 = \min_{\bz} \|\bd - \bv(\bz)\|_F^2, 
-%\end{equation}
-%where $\bz = [\bg,\by]^T$.
-The other vectors in Eq.~\eqref{eq:least_squares_red} were defined in Section~\ref{} \footnote{In this paper we use the same symbol to denote corrupted predicted visibilities and true observed corrupted visibilities as it
-%is trivial to distinguish which one is implied from the context.
-Using the same symbol for these two quantities improves the readability of the paper.}.
-Note the similarity between Eq.~\eqref{eq:least_squares} and Eq.~\eqref{eq:least_squares_red},
-%i.e. redundant calibration can be cast as a non-linear least squares problem.
-
-Analogous to \cite{Smirnov2015}, we recast Eq.~\eqref{eq:least_squares_red} to 
+The number of model parameters to be solved for is now $P = 2(N+L)$, since redundant calibration is a complex problem.
+
+In literature, equation~\ref{eq:least_squares_red} is solved by splitting the problem into its real and imaginary parts.
+The real and imaginary parts of the unknown parameters are then solved for separately \citep{Wieringa1992,Liu2010,Zheng2014}.
+Currently, the above is achieved by using the real-valued GN algorithm \citep{Kurien2016}. 
+We, instead, intend to formulate the redundant calibration problem using Wirtinger calculus and recast equation~\ref{eq:least_squares_red} as
 \begin{equation}
@@
 \end{equation}
-where $\breve{\bz} = [\bz^T,\conj{\bz}^T]^T$. Eq.~\eqref{eq:least_squares_complex} is the complex conjugate augmented counterpart of Eq.~\eqref{eq:least_squares_red}.
-%The fact that $\bv$ is now analytic in its argument has the immediate consequence: it becomes possible to apply complex GN
-%and complex LM to Eq.~\eqref{eq:least_squares_complex}.
-
-\subsubsection{Complex GN and LM}
-\label{sec:complex_GN_LM}
-The complex Jacobian associated with Eq.~\eqref{eq:least_squares_complex} is equal to
+where $\breve{\bz} = [\bz^T,\conj{\bz}^T]^T$. 
+We derive the complex Jacobian associated with equation~\ref{eq:least_squares_complex} to be:
 \begin{equation}
@@
 \end{equation}
-The GN update step is now defined as:
+We can now calculate the GN and LM updates to be 
 \begin{equation}
 \label{eq:GN_update}
- \Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\br}.
-\end{equation}
-The LM update rule is defined in a very similar way:
+\Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\br}
+\end{equation}
+and 
 \begin{equation}
@@
 \end{equation}
-As was the case in Section~\ref{sec:sky_wirtinger}, we can now use Eq.~\eqref{eq:GN_update} or Eq.~\eqref{eq:LM_update} to iteratively update our 
-parameter vector as follows:
+respectively. As in Section~\ref{sec:sky_wirtinger}, equation~\ref{eq:GN_update} can be used to iteratively update our parameter vector:
 \begin{equation}
@@
 The analytic expressions for $\bJ$, $\bH$ and $\bJ^H\breve{\br}$ can be found in Appendix~\ref{sec:analytic}.
-
-\subsubsection{Simplifying GN}
-
-The following identities,
-\begin{align}
-\label{eq:identities}
-\frac{1}{3}\bJ\breve{\bz} &= \breve{\bv}, & \bJ^H\breve{\bv} &= \widetilde{\bH}\breve{\bz} 
-\end{align}
-are trivially established by mechanically showing that the left hand side of each identity in Eq.~\eqref{eq:identities} is equal to its right hand side.
-In Eq.~\eqref{eq:identities}, $\widetilde{\bH}$ denotes $\bI\odot\bH$.
-
-Replacing $\breve{\br}$ with $\breve{\bd}-\breve{\bv}$ in Eq.~\eqref{eq:GN_update} results in
+Appendix~\ref{sec:analytic} also contain two useful identities, involving $\bJ$ and $\bH$.
+
+In the case of the GN algorithm we can simplify equation~\ref{eq:update} even further. Replacing $\breve{\br}$ with $\breve{\bd}-\breve{\bv}$ in equation~\ref{eq:GN_update} results in
 \begin{equation}
@@
 \end{equation}
-Substituting the first identity of Eq.~\eqref{eq:identities} into Eq.~\eqref{eq:temp_GN_eq} and simplifying the result 
+If we substitute the first identity of equation~\ref{eq:identities} into equation~\ref{eq:temp_GN_eq} and we simplify the 
+result we obtain 
 \begin{equation}
@@
 \end{equation}
-It now trivially follows that:
+If we use the above simplified update in equation~\ref{eq:update} it reduces to
 \begin{equation}
@@
 \end{equation}
-Eq.~\eqref{eq:two_thrids} shows us that in the case of redundant calibration we can calculate the GN update step without calculating the residual. 
-% The 
-% normal skymodel-based equivalent of Eq.~\eqref{eq:two_thrids} is equal to \citep{Smirnov2015}:
-% \begin{equation}
-% \label{eq:one_half}
-% \breve{\bg}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{1}{2}\breve{\bg}_{k}, 
-% \end{equation}
-% where $\breve{\bg} = [\bg^T,\conj{\bg}^T]^T$
-% 
-% FIX TRANSPOSE PROBLEM
-
-\section{Comparing ADI and PCG}
-\label{sec:results}
-In Section~\ref{sec:red_wirtinger} we discussed the complex GN and LM algorithms and how they relate to redundant calibration. If we inspect Eq.~\eqref{eq:GN_update} and 
-Eq.~\eqref{eq:LM_update} we see that the most expensive operation used by the GN and LM algorithms is the inversion of the Hessian, $\bH$. The most logical 
-way of speeding up the algorithm is to investigate whether we can do this inversion in some efficient way.
-
-Two recent algorithms have been proposed that are able to invert $\bH$ in an efficient manner, namely the ADI \citep{Marthi2014} and PCG \citep{Liu2010} methods. 
-Note that here we focus solely on comparing the computational complexity of inverting $\bH$ using either one of these two methods. Rating the LM or GN algorithms
-using one of these approaches relative to each other once the computational complexity of inverting $\bH$ is known is trivial. The simulations we performed to compare these two 
-methods are summarized in Section~\ref{sec:simulations}.
-
-%We restrict ourselves to comparing the most 
-%expensive operation of the two methods namely, computing the inverse of the Hessian $\bH$ (or modified Hessian $\bmH$ in the case of LM). It becomes trivial to rate these two 
-%algorithms relative to each other once the computational complexity of the inverse operation is known.
-
-The basic difference between these two speedup algorithms can be understood by inspecting Fig.~\ref{fig:hessian}. The first observation we can make by inspecting 
-Fig.~\ref{fig:hessian_reg} and Fig.~\ref{fig:hessian_hex} is that the diagonal entries of $\bH$ are more significant than the off-diagonal entries in both cases.
-It therefore makes sense to approximate the $\bH$ with its diagonal, which greatly simplifies the computational complexity of inverting 
-$\bH$, i.e. it becomes $\mathbb{O}(P)$\footnote{The cost of constructing 
-the diagonal is not included in this estimate.} \citep{Smirnov2015}. This diagonal-approximation approach is effectively ADI \citep{Marthi2014}. The exact details of why this 
-is the case is discussed in Section~\ref{sec:adi}.
-We will also show how \citet{Marthi2014}, \citet{Smirnov2015} and \citet{Salvini2014} relate to each other in that same section.
-It should be noted here that strictly speaking the ADI method actually refers to the complete LM update step using the approximate $\bH$ and not just
-to the inversion of the approximate $\bH$.
-To improve the readability of the paper, we use the acronym ADI to refer to both the faster approximate inverse approach and the complete LM algorithm that uses it.
-Which one of these two concepts we are actually referring to, however, is always perfectly clear from the context.
-%Furhtermore, the name ADI also implies that with this algorithm we alternate between computing the 
-%gains, their conjugates and the true observed visibilities. %In Appendix~\ref{sec:sbc} we briefly summarize the ADI method that is used for skymodel-based calibration.
-
-The second observation we can make from Fig.~\ref{fig:hessian_reg} and Fig.~\ref{fig:hessian_hex} is that both Hessians contain many zero entries, i.e.
-they are sparse matrices.
-This, in conjunction with the more significant diagonal entries, implies that the PCG method (presented in Section~\ref{sec:conj_grad})
-can be applied to invert them in a computationally efficient manner.
-\citet{Liu2010} proposed the PCG method to compute the inverse of $\bH$, however the computational complexity of this approach was not analyzed therein.
-In Section~\ref{sec:pcg}, we do a proper computational complexity analysis of PCG. We find that the computational complexity of PCG is asymptotically bounded by $\mathbb{O}(P^2)$, which is 
-a lot better than straightforward inversion which takes $\mathbb{O}(P^3)$.
+Equation~\ref{eq:two_thrids} is the redundant equivalent of equation~\ref{eq:one_half} and it shows us that in the case of redundant calibration we can calculate the GN parameter update step without calculating the residual.
 
@@
 
-%Since we approximate the Hessian in the ADI method with its diagonal the computational complexity of its inverse becomes $\mathbb{O}(P)$\footnote{The cost of constructing 
-%the diagonal is not included in this estimate.}.
-
-Even though the PCG method is computationally more expensive than the ADI method we cannot arrive at a final conclusion at this point. With the ADI method 
-we only approximate the inverse of the Hessian, while in the case of the PCG method we compute the exact inverse. So the question arises whether the LM algorithm using the PCG method 
-will converge much faster than the ADI method even though each iteration (which includes an inversion) is more expensive. We investigate this question in Section~\ref{sec:comparison}.
-
-\subsection{Simulations}
-\label{sec:simulations}
-%The main purpose of the simulations conducted in this paper is to genrate experimental results with which we can compare the computational complexity of ADI and PCG.
-We used an hexagonal geometric layout with an adjustable number of rings during the
-%simulations. 
-For the skymodel we generated 100 flat spectrum sources.
-Each source had a random flux value and position.
-The flux of the sources followed a Pareto distribution (with a shape parameter of 2), while their positions were uniformly
-distributed in a 3$^{\circ}$ by 3$^{\circ}$ square grid around the field center. The remaining fixed simulation parameters can all be found in Table~\ref{tab:fixed_parm}. 
-We also employed three different gain error models which we summarize in Table~\ref{tab:gain_parm}. We found that it did not matter which gain error model we used during
-our simulations, as the model we used did not have an impact on the results.
-Moreover, we had two major experimental setups, we either made use of simulated observations consisting of one frequency channel and multiple time-slots or one time-slot and multiple channels.
-We describe these two experimental setups in more detail in Table~\ref{tab:ch_parm}. As was the case for the gain error models, the experimental setup 
-that was used did not alter the simulation results.
-
-The main flow of a single simulation is as follows.
-We created corrupted visibilities using the fixed simulation parameters presented in Table~\ref{tab:fixed_parm} and either one of the experimental setups summarized in Table~\ref{tab:ch_parm}.
-We then used per time-slot and channel redundant calibration, i.e.
-we use the LM algorithm, which in turn makes use of either the ADI or PCG method, to
-estimate the gains and the true observed visibilities.
-Our initial parameter guess $\bz_0$ was chosen to be an all one vector.
-To obtain statistically significant results we performed, for both the ADI and PCG methods, multiple statistical independent simulations.
-
-Furthermore, the following SNR definition is used in this paper  
-\begin{equation}
-\label{}
-\textrm{SNR} = \frac{<\bv\odot\conj{\bv}>_{\nu,t,pq}}{<\bn\odot\conj{\bn}>_{\nu,t,pq}}, 
-\end{equation}
-where $<*>_{\nu,t,pq}$ denotes averaging over frequency, time and baseline. The definition we use here is based on the SNR definitions used in \citet{Liu2010} and \citet{Marthi2014}.
-
-%We compare ADI with PCG in Section \textbf{??}, by using the results generated from the simulations we performed using 
-%these two methods.
-
-%The results reported in this paper were generated using two major experimental setups, namely a single time-slot and multiple frequency channels setup or 
-%a single frquency channel and multiple time-slots setup. The simulation parameters which where kept fixed in both these setups can be found in 
-%Table~\ref{tab:fixed_parm}. The sky-model we used to simulate the visibilities were the same for both setups.   We used a hexagonal redundant layout for our antenna positions. We produced 
-%results for different
-
-\begin{table*}
-\centering
-\caption{The following fixed simulation parameters were used to generate the results of this paper. We use $U$ to represent 
-a uniform distribution.}
-\begin{tabular}{|c c c c c c c|} 
-\hline
-Declination & Latitude & Minimum baseline & Geometry& Sources & Flux & Spatial \\
-\hline 
-\hline
- $-74^{\circ}39'37.481''$ & $-30^{\circ}43'17.34''$ & 20 & Hexagonal &100 & Pareto (shape param. of 2) & $U[-3,3]$ (deg.) 
-\end{tabular}
-\label{tab:fixed_parm}
-\end{table*}
+Figure~\ref{fig:hessian} shows that the Hessian matrix $\bH$ is nearly diagonal and sparse for both a regular and hexagonal layout considered.
+We therefore follow the approach of \citet{Smirnov2015} and approximate the Hessian matrix $\bH$ with its diagonal.
+If we substitute $\bJ^H\bJ$ with $\widetilde{\bH}=\bH\odot\bI$ and replace $\breve{\br}$ with $\breve{\bd} - \breve{\bv}$ in equation~\ref{eq:GN_update} we obtain:
+\begin{equation}
+\label{eq:appr_GN}
+ \Delta \breve{\bz} \approx \widetilde{\bH}^{-1}\bJ^H(\breve{\bd}-\breve{\bv})
+\end{equation}
+Utilizing the second identity in equation~\ref{eq:identities} allows us to simplify equation~\ref{eq:appr_GN} to
+\begin{equation}
+  \Delta \breve{\bz} \approx \widetilde{\bH}^{-1}\bJ^H\breve{\bd}-\breve{\bz},
+\end{equation}
+which leads to
+\begin{equation}
+ \breve{\bz}_{k+1} \approx \widetilde{\bH}^{-1}\bJ^H\breve{\bd}.
+\end{equation}
+We obtain a similar result if we repeat the above procedure, only in this case we use equation~\ref{eq:LM_update} instead. Thus,
+\begin{align}
+\breve{\bz}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bz}_k,\label{eq:lambda}\\
+ &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bz}_k. \label{eq:alpha}  
+\end{align}
+The analytic expression of $\bJ^H\breve{\bd}$ will be very similar to the analytic 
+expression of $\bJ^H\breve{\br}$, the only difference being that in equation~\ref{eq:ab} the letter $r$ would be replaced by a $d$. If we substitute the analytic expression
+of $\bJ^H\breve{\bd}$ and $\widetilde{\bH}^{-1}$ (which can easily be constructed using Appendix~\ref{sec:analytic}) into equation~\ref{eq:alpha} we obtain the following two update rules:
+\begin{equation}
+\label{eq:g_update}
+g_{i}^{k+1} = \alpha \frac{\sum_{j\neq i} g_j^k \widetilde{y}_{ij}^{~\!\!k} d_{ij}}{\sum_{j\neq i} |g_j^k|^2|y_{\zeta_{ij}}^k|^2} + (1-\alpha) g_i^k, 
+\end{equation}
+and
+\begin{equation}
+\label{eq:y_update}
+y_{i}^{k+1} = \alpha \frac{\sum_{rs \in \mathcal{RS}_i} \conj{g}_r^k g_s^k d_{rs}}{\sum_{rs \in \mathcal{RS}_i}|g_r^k|^2|g_s^k|^2} + (1-\alpha) y_i^k. 
+\end{equation}
+The the index set $\mathcal{RS}_i$ and the quantity $\widetilde{y}_{ij}$ is defined in equation~\ref{eq:RS} and equation~\ref{eq:y_tilde}, respectively. 
+The computational complexity of inverting $\widetilde{\bH}$ is $O(P)$. We note that equation~\ref{eq:g_update} is the gain estimator associated with \textsc{StEfCal}.
+
+Equation~\ref{eq:g_update} and~\ref{eq:y_update} were obtained by \citet{Marthi2014} by taking the derivative of the objective function $\Lambda$ relative to the elements of $\bg$ and $\by$, setting the intermediate results to zero and then solving for the unknown parameters (i.e.
+using the gradient descent algorithm).
+We note that their derivation is less general. The LM algorithm has better convergence properties than gradient descent and 
+encompasses the gradient descent algorithm as a special case.
+In Appendix~\ref{sec:red_stef_ADI} we show that equation~\ref{eq:g_update} and~\ref{eq:y_update} can also be derived using the ADI method.
+For this reason, we refer to the code implementation of the pseudo LM calibration scheme, derived above (equation~\ref{eq:g_update} and~\ref{eq:y_update}), as redundant \textsc{StEfCal}, throughout the rest of the paper.
+
+The choice of the $\alpha$ parameter is somewhat unconstrained. In this paper we chose $\alpha$ by adopting the same strategy that is used by \textsc{StEfCal} and \citet{Marthi2014},
+i.e. we chose $\alpha$ to be equal to a $\frac{1}{3}$ ($\lambda = 2$). We also carried out simulations to validate this choice.
+%In the case of \textsc{StEfCal}, $\alpha = \frac{1}{2}$ (for even iterations), which results in the last term of equation~\ref{eq:one_half} and equation~\ref{eq:stef_alpha} being equal.
+When $\alpha=\frac{1}{3}$, the last term of equation~\ref{eq:two_thrids} and equation~\ref{eq:alpha} are equal.
+
+We generated a sky model that comprises of 100 flat spectrum sources distributed over a 3$^{\circ}$ by 3$^{\circ}$ sky 
+patch centred. The flux density of each source was drawn from a power law distribution with a slope of 2 and the source position drawn from a uniform distribution. 
+We also made use of multiple fictitious telescope layouts each one having a hexagonal geometry (see the left upper image of Figure~\ref{fig:geometry_function} for an example layout).
+The largest hexagonal array that we used has 217 antennas, with a minimum and maximum baseline of of 20~m and 320~m respectively.
+
+We "corrupted" visibilities by applying gain errors (Table~\ref{tab:gain_parm}) and calibrated the corrupted visibilities using redundant \textsc{StEfCal}.
+Solutions were independently derived for each time-step and channel for five realizations.
 
@@
 performed our simulations over multiple time-slots and one frequency channel or one timeslot and multiple frequencey channels (see Table~\ref{tab:ch_parm}).
-Moreover, $c$ in the left most column denotes the speed of light.}
+Moreover, $c$ in the left most column denotes the speed of light.
+We adopted a sinusoidal error model as well as phase slope across frequency that mimics a real case of physical delays between different antennas.
+It is worth pointing out the following. In the first model, we model a gain error with amplitude around one and an additional phase error. In 
+the second model we chose $A$ in such a way that we do not unintentionally produce a zero gain value. For the third model 
+the value of $\tau$ was chosen so that the amount of phase wraps that occur across the observing band is restricted to a realistic number.
+{\bf (GB: we need a final better description of the models, but we can work this out later).}}
 \begin{tabular}{|c c c c|} 
@@
 \centering
-\caption{We generated results using two major experimental setups. We either used one frequency channel and multiple time-slots, or one time-slot and multiple frequency 
-channels. The most important parameters used in realizing these two major setups are presented here. Note that we refer to the gain error models in Table~\ref{tab:gain_parm} in the 
-last row of this table.}
+\caption{We generated results using two main setups.
+We either used one frequency channel and multiple time-slots, or one time-slot and multiple frequency channels.
+The most important parameters used in realizing these two major setups are presented here.
+%Note that we refer to the gain error models in the last row of Table~\ref{tab:gain_parm}.
+We chose the observational frequency band of setup 1 to coincide with the HERA array. To broaden the scope of our 
+analysis we chose the observational frequency of our second setup to be equal to 1.4~GHz, which is a typical observing frequency of the Westerbork Synthesis Radio Telescope.}
 \begin{tabular}{|c c c|} 
@@
 
-\subsection{Alternating Direction Implicit Method}
-\label{sec:adi}
-%In Section~\ref{sec:adi} we show how the ADI method used for skymodel-based calibration relates to the redundant ADI method presented in \citet{Marthi2014}. 
-We start this section by briefly reviewing the ADI method as it pertains to skymodel-based calibration in Section~\ref{sec:sbc} \citep{Salvini2014,Smirnov2015}.
-
-In Section~\ref{sec:red_c}, we will show that the fast redundant calibration algorithm, that is presented in \citet{Marthi2014},
-is in fact nothing more than an extension of the skymodel-based ADI method, presented in \citet{Salvini2014}, to the redundant calibration use case (i.e. redundant \textsc{StEfCal}).
-In doing so we will justify the naming convention we have been using for it.
-
-In Section~\ref{sec:adi_results}, we present some ADI simulation results. 
-%We will mainly accomplish this by using the complex framework proposed by \citet{Smirnov2015}.
-
-\subsubsection{Skymodel-based Calibration}
-\label{sec:sbc}
-\textsc{StEfCal} is an ADI method in which the measurement equation, Eq.~\eqref{eq:vis_definition}, is linearized by assuming that the gains are known, but
-that their conjugates are unknown \citep{Mitchell:MWA-cal,Salvini2014}.
-Under this assumption, the conjugates of the gains can be obtained trivially.
-Once the conjugates of the gains have been computed, it is possible to solve for the gains.
-One can then alternate between computing the gains and their conjugates until convergence
-is reached.
-By correctly exploiting the symmetry inherent in the calibration problem, this alternating approach can be replaced by a single iterative new parameter estimation step.
-\cite{Salvini2014} also show that
-the convergence of the algorithm can be improved, if for every even iteration, the basic new parameter estimation step is modified as follows: the current gain estimate is replaced by the average of the current gain estimate and the gain estimate of the previous odd iteration.
-\textsc{StEFCal} lowers the computational complexity of normal skymodel-based calibration from $\mathbb{O}(N^3)$ to $\mathbb{O}(N^2)$.
-
-\citet{Smirnov2015} recently showed that one can also derive \textsc{StEFCal} by viewing normal skymodel-based calibration as a complex optimization problem.
-They realized while inspecting the skymodel-based $\bH$ that its diagonal entries are much more significant than its off-diagonal entries.
-As we mention in Section~\ref{sec:results}, the exact same conclusion can be reached if we inspect the Hessian $\bH$ that is associated with redundant calibration (see Fig.~\ref{fig:hessian}). 
-By approximating the skymodel-based $\bH$ by its diagonal and substituting this approximate Hessian $\widetilde{\bH}$ back into the fundamental skymodel-based LM update step we obtain \citep{Smirnov2015}
-\begin{align}
-\breve{\bg}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bg}_k,\label{eq:stef_lambda}\\
- &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bg}_k, \label{eq:stef_alpha}  
-\end{align}
-where $\alpha = \frac{1}{1+\lambda}$. Moreover, $\lambda = \frac{1-\alpha}{\alpha}$.
-
-If we choose $\lambda$ to be zero, which corresponds to an $\alpha$ equal to one, we obtain the new parameter estimation step of \textsc{StEfCal} used during odd iterations.
-On the other hand, if we choose $\lambda$ equal to one (which corresponds
-to $\alpha=\frac{1}{2}$) we obtain the new parameter estimation step of \textsc{StEfCal} used during even iterations.
-
-Furthermore, the simplified skymodel-based GN equivalent of Eq.~\eqref{eq:two_thrids} is equal to
-\citep{Smirnov2015}
-\begin{equation}
-\label{eq:one_half}
-\breve{\bg}_{k+1} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\bd} + \frac{1}{2}\breve{\bg}_{k}. 
-\end{equation}
-Note that in Eq.~\eqref{eq:one_half} the influence that the estimate of the parameters from the previous iteration has on the current iteration is scaled by a half, while
-in Eq.~\eqref{eq:two_thrids} it is scaled by two thirds. Note the similarity between the last terms of Eq.~\eqref{eq:stef_alpha} and Eq.~\eqref{eq:one_half} when $\alpha$ is a half.
-
-\subsubsection{Redundant Calibration}
-\label{sec:red_c}
-In this section, we re-derive the fast redundant calibration algorithm presented in \citet{Marthi2014} using two different approaches.
-Both these approaches differ from the one presented in \citet{Marthi2014}.
-In \citet{Marthi2014}, the algorithm is derived by taking the derivative of the
-objective function $\Lambda$ relative to $\bg$ and $\by$, setting the intermediate results to naught and then solving for the unknown parameters.
-
-It is in fact trivial to extend the ADI method presented in \citet{Salvini2014} to the redundant use case, a.k.a.
-we now alternate between solving for the antenna gains their conjugates and the true observed visibilities.
-See Appendix~\ref{sec:red_stef_ADI} for the details.
-The result of this extension is that we obtain the redundant calibration algorithm presented in \citet{Marthi2014}.
-This implies that the fast redundant calibration algorithm presented in \citet{Marthi2014} can for all intents and purposes be referred to as redundant \textsc{StEfCal} (i.e.
-some kind of ADI method).
-
-Recall, that \citet{Smirnov2015} showed that \textsc{StEfCal} can also be derived by approximating the $\bH$, associated with skymodel-based calibration, with 
-its diagonal.
-To double check if our statement that the algorithm in \citet{Marthi2014} is in actual fact redundant \textsc{StEfCal} (i.e.
-some kind of ADI method) we will now use the diagonal approximation approach presented in \citet{Smirnov2015} to yet again re-derive the redundant calibration algorithm presented in \citet{Marthi2014}.
-The benefit of doing this second additional derivation is that it also enables us to show that the ADI method is in actual fact just a special case of the LM algorithm.
-
-Since the diagonal entries of the $\bH$, associated with redundant calibration, are more significant than its off-diagonal entries it makes 
-sense to try and approximate it with its diagonal (see Fig.~\ref{fig:hessian} and Eq.~\eqref{eq:red_H}).
-If we substitute $\bJ^H\bJ$ with $\widetilde{\bH}$ and replace $\breve{\br}$ with $\breve{\bd} - \breve{\bv}$ in Eq.~\eqref{eq:GN_update} we obtain
-\begin{equation}
-\label{eq:appr_GN}
- \Delta \breve{\bz} \approx \widetilde{\bH}^{-1}\bJ^H(\breve{\bd}-\breve{\bv})
-\end{equation}
-Utilizing the second identity in Eq.~\eqref{eq:identities} allows us to simplify Eq.~\eqref{eq:appr_GN} to
-\begin{equation}
-  \Delta \breve{\bz} \approx \widetilde{\bH}^{-1}\bJ^H\breve{\bd}-\breve{\bz}.
-\end{equation}
-It now trivially follows that
-\begin{equation}
- \breve{\bz}_{k+1} \approx \widetilde{\bH}^{-1}\bJ^H\breve{\bd}.
-\end{equation}
-We obtain a similar result if we repeat the above procedure, only in this case we use Eq.~\eqref{eq:LM_update} instead. Thus,
-\begin{align}
-\breve{\bz}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bz}_k,\label{eq:lambda}\\
- &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bz}_k. \label{eq:alpha}  
-\end{align}
-The analytic expression of $\bJ^H\breve{\bd}$ will be very similar to the analytic 
-expression of $\bJ^H\breve{\br}$, the only difference being that in Eq.~\eqref{eq:ab} the letter $r$ would be replaced by a $d$. If we substitute the analytic expression
-of $\bJ^H\breve{\bd}$ and $\widetilde{\bH}^{-1}$ (which can easily be constructed using Appendix~\ref{sec:analytic}) into Eq.~\eqref{eq:alpha} we obtain the following two update rules:
-\begin{equation}
-\label{eq:g_update}
-g_{i}^{k+1} = \alpha \frac{\sum_{j\neq i} g_j^k \widetilde{y}_{ij}^{~\!\!k} d_{ij}}{\sum_{j\neq i} |g_j^k|^2|y_{\zeta_{ij}}^k|^2} + (1-\alpha) g_i^k, 
-\end{equation}
-and
-\begin{equation}
-\label{eq:y_update}
-y_{i}^{k+1} = \alpha \frac{\sum_{pq \in \mathcal{PQ}_i} \conj{g}_p^k g_q^k d_{pq}}{\sum_{pq \in \mathcal{PQ}_i}|g_p^k|^2|g_q^k|^2} + (1-\alpha) y_i^k. 
-\end{equation}
-The computational complexity of inverting $\widetilde{\bH}$ is $\mathbb{O}(P)$.
-
-Eq.~\eqref{eq:g_update} and Eq.~\eqref{eq:y_update} are the exact same iterative new parameter estimators derived by \citet{Marthi2014}. Also note that Eq.~\eqref{eq:g_update} is the fundamental
-gain estimator associated with skymodel-based \textsc{StEfCal}.
-We have, therefore, successfully used the diagonal approximation approach proposed by \citet{Smirnov2015} to re-derive the fast redundant calibration method presented in
-\citet{Marthi2014}.
-%and in doing so we have confirmed shown that this algorithm is in actual fact nothing more than redundant \textsc{StEfCal}, i.e.
-some form of ADI method, justifying the naming convention we have been using for it.
-%Re-deriving the algorithm presented in \citet{Marthi2014} in this way serves as a further confirmation that it is indeed nothing more than redundant \textsc{StEfCal} (i.e. some kind of ADI method).
-
-The question now arises what value should we use for $\alpha$? In \citet{Marthi2014}, they propose 
-$\alpha = 0.3$. Although an exact mathematical proof for the optimal choice of $\alpha$ still eludes us, it is still possible to
-venture a clever guess by studying what the choice of $\alpha$ is in the case of skymodel-based \textsc{StEFCal}. The value of $\alpha$ is equal to a half 
-in the case of \textsc{StEFCal} (for even iterations); see Section~\ref{sec:sbc}.
-If we wish to mimic this behaviour for the redundant calibration use case we need to scale $\breve{\bz}$ by two thirds in Eq.~\eqref{eq:alpha} (see Eq.~\eqref{eq:two_thrids}) which implies that we should choose $\alpha$ equal to a third, which corresponds to a $\lambda$ equal to two.
-We used this $\lambda$ value in this paper.
-
-\subsubsection{ADI simulation results}
-\label{sec:adi_results}
-Optimality results of ADI were presented in \citet{Marthi2014}. We do not repeat their analysis here. We do, however, present our own ADI results, in Fig.~\ref{fig:prec_error}, that confirm that the
-ADI method can accurately estimate the antenna gains and the observed visibilities at different SNR values. To quantify the exact error in our estimates we used the percentage error $\beta$ defined 
-as 
+The simulated visibilities included a noise term. The purpose of which is to produce a certain SNR in the simulated
+visibilities. We made use of the following definition of \citep[SNR, ][]{Liu2010,Marthi2014}:  
+\begin{equation}
+\label{}
+\textrm{SNR} = 10\log \left (\frac{<\bv\odot\conj{\bv}>_{\nu,t,pq}}{<\bn\odot\conj{\bn}>_{\nu,t,pq}} \right ), 
+\end{equation}
+where $<*>_{\nu,t,pq}$ denotes averaging over frequency, time and baseline.
+
+Figure~\ref{fig:prec_error} show the simulation results as a function of SNR and number of antennas. The accuracy of our solutions is quantified through the percentage error
 \begin{equation}
 \label{eq:beta}
-\beta = \frac{\|\bv - \widehat{\bv}\|_F^2}{\|\bv\|_F^2}.
-\end{equation}
-We have already defined $\bv$ in Eq.~\eqref{eq:vec_definitions}. Moreover, in Eq.~\eqref{eq:beta}, $\widehat{\bv}$ denotes the ADI estimate of $\bv$.
+\beta = \frac{\|\bv - \widehat{\bv}\|_F^2}{\|\bv\|_F^2},
+\end{equation}
+where $\widehat{\bv}$ is the redundant \textsc{StEfCal} parameter estimate.
+
+The error magnitude follows the expected behaviour, i.e., it decreases as a function of SNR and number of antennas $N$.
+Interestingly, it reduces to a few percent when $N > 120$ for essentially any choice of SNR.
+%\citet{Marthi2014} conducted an even more in depth analysis than the one presented here.
+They found that the total computational cost of the algorithm associated with redundant StEfCal is $O(P^2)$ and that it approaches the Cram{\'e}r-Rao Bound (CRB) at moderate SNRs.
 
@@
 \includegraphics[width=0.47\textwidth]{./prec_error.pdf} 
-\caption{We plot the percentage error $\beta$ between the estimated 
-predicted visibilities and the true predicted visibilities for different SNR values as a function of the number of antennas in the array $N$.
-The graph presented here was generated using the first experimental setup in Table~\ref{tab:ch_parm}.}
+\caption{We plot the percentage error $\beta$ between the simulated visibilitied and the visibilities solved for by redundant \textsc{StEfCal} for different SNR values as a function of the number of antennas in the array $N$.}
 \label{fig:prec_error}
@@
 
-% At this point there is no way of knowing if this is mere coincidence. DEF REWRITE THIS.
-% 
-% REWRTIE THIS IN A DIFFERENT WAY. START WITH CHENGULAR AND JUST MENTION THE LINK TO STEFCAL'S CHOICE OF ALPHA.
-% 
-% The update rules in Eq.~\eqref{eq:lambda} and Eq.~\eqref{eq:alpha} have been derived before, by Chengular... The approach they followed is discussed in the Appendix.
-% 
-% Interestingly enough, if we substitute $\alpha = 1$ into Eq.~\ref{eq:g_update} then it reduces to the skymodel-based StEFCal update rule. In Chengular they propose 
-% an $\alpha = 0.3$.
-% 
-% In an attempt to merge the terminology that has independently develop in the general calibration and redundant calibration literature and to emphasize the close
-% relationship between Stefcal and the approach derived by I will use the label Redunandat SteFCal (R-Stefcal) to refer to the NLS method proposed by ... 
-% 
-% ****\\
-% 
-% with $\alpha = \frac{1}{1+\lambda}$. Moreover $\lambda = \frac{1-\alpha}{\alpha}$. So in Marthi and Chengular (2014) they use an $\alpha$ of 0.3, which corresponds to an LM damping factor $\lambda$
-% of $2\frac{1}{3}$.
-% 
-% NB:: NEED TO ADD SNR PLOTS HERE
-% PLUS IMAGE OF HESSIAN FOR THE DISCUSSION
-
-\subsection{Preconditioned Conjugate Gradient Method}
+\section{Preconditioned Conjugate Gradient Method}
 \label{sec:pcg}
-In this section we investigate the computational complexity of the PCG method when applied to the redundant calibration problem. 
-We briefly review the PCG method in Appendix~\ref{sec:conj_grad}. If the reader is not that familiar with the PCG method we encourage 
-the reader to briefly review Appendix~\ref{sec:conj_grad} before continuing with the paper. The PCG method can be used to efficiently invert Hermitian sparse 
-matrices, which is exactly the kind of matrix $\bH$ is (see Fig.~\ref{fig:hessian} and Eq.~\ref{eq:red_H}).
-As we mention in Appendix~\ref{sec:conj_grad} there are two factors which influence the computational complexity of the PCG method, namely the spectral condition number $\kappa$ and the sparsity of the matrix we wish to invert (see Eq.~\eqref{eq:cg_bound}).
-
-We used a $\lambda$ value of two, implying we used Eq.~\eqref{eq:LM_update} instead of Eq.~\eqref{eq:GN_update} throughout this section.
-We investigate the spectral condition number and the sparsity of the modified Hessian $\bmH$, which is a non-singular matrix, in Section~\ref{sec:scn} and Section~\ref{sec:sparsity} respectively.
-It is interesting to note, that the Hessian $\bH$ itself is singular which means that if we are to use Eq.~\eqref{eq:GN_update} instead we would need to be able to invert a singular matrix.
-In Appendix~\ref{sec:normal},
-we mention that the PCG method would in fact also be able to pseudo-invert $\bH$ directly.
-
-\subsubsection{Spectral Condition Number}
-\label{sec:scn}
-In Appendix~\ref{sec:precon} we mentioned that the convergence properties of CG can be dramatically improved through preconditioning.
-Inspecting Fig.~\ref{fig:hessian}, reveals that the Jacobian preconditioner
-is a good candidate for the redundant use case. In Fig.~\ref{fig:kappa}, we see that this choice of preconditioner performs very well, the value of the spectral condition number $\kappa$
-becomes constant and independent of problem size, i.e. the number of antennas in the array. Preconditioning therefore effectively eliminates $\kappa$ from Eq.~\eqref{eq:cg_bound}. This result
-is confirmed by Fig.~\ref{fig:cg_itr}.
-Fig.~\ref{fig:cg_itr} shows us that the number of major iterations needed by the preconditioned conjugate gradient method to invert $\bmH$ is independent of the number of antennas in the array.
-Moreover, the number of major iterations required is also much smaller than the dimension of $\bmH$.
-
-%kappa + itr plots
+\citet{Liu2010} suggested that the speed of convergence of redundant calibration could be optimized using the 
+conjugate gradient method \citep{Hestenes1952}, which would be computationally advantageous, since the Hessian matrix associated 
+with redundant calibration (see Figure~\ref{fig:hessian}) is sparse \citep{Reid1971}. In this section we study the computational complexity
+of the conjugate gradient (CG) method when it is used to invert the modified Hessian matrix (see equation~\ref{eq:LM_update}), 
+in particular when preconditioning is used (i.e.
+the PCG method).
+Interestingly, emperical tests suggests that the unmodified Hessian itself is singular.
+It is therefore important to mention here, that the CG method
+can pseudo invert the unmodified Hessian as well, i.e.
+the CG method can be directly applied to equation~\ref{eq:GN_update}, since the Hessian is a positive semi-definite Hermition matrix and the vector $\bJ^H\breve{\br}$ is an element of its column range \citep{Lu2015}.
+
+%(see also Appendix~\ref{sec:conj_grad}).
+
+The computational complexity of CG is  
+\begin{equation}
+\label{eq:cg_bound}
+O(\sqrt{\kappa}m), 
+\end{equation}
+where $m$ denotes the number of non-zero entries and $\kappa$ denotes the spectral 
+condition number of the matrix which is to be inverted. The spectral condition number $\kappa$ of the matrix $\bA$ is defined as:
+\begin{equation}
+\label{eq:kappa}
+\kappa(\bA) = \frac{\iota_{\textrm{max}}}{\iota_{\textrm{min}}}, 
+\end{equation}
+where $\iota_{\textrm{max}}$ and $\iota_{\textrm{min}}$ respectively denote the largest and the smallest eigenvalue of $\bA$.
+
+Preconditioning is a technique used to improve the spectral condition number of a matrix. Let us consider a generic system of linear equations 
+\begin{equation}
+\label{eq:prec}
+\bA\bx = \bb,
+\end{equation}
+and a positive-definite Hermitian matrix $\bM$ so that 
+\begin{equation}
+\label{eq:prec}
+\bM^{-1}\bA\bx = \bM^{-1}\bb.
+\end{equation}
+
+The matrix $\bM$ is a good preconditioner if:
+\begin{equation}
+\label{eq:kappa}
+\kappa(\bM^{-1}\bA) \ll \kappa(\bA),
+\end{equation}
+i.e. if it lowers lowers the condition number of a matrix.
+If $\bA$ is a nearly diagonal matrix, the Jacobian preconditioner is a natural choice of $\bM$ and can be computed as:
+\begin{equation}
+\bM = \bA\odot\bI. 
+\end{equation}
+
+%The most expensive operation ni the PCG algorithm is the vector-matrix product $\bA\bx$, which has a complexity of $O(m)$, where $m$ is the number of non-zero entries in $\bA$.
+
+In order to quantify the effectiveness of the CG method in redundant calibration, we investigate the spectral condition number and the sparsity of the modified Hessian $\bmH$ (i.e. $\lambda = 2$).
+
+We generated simulated (i.e. corrupted) visibilities according to models described in Table~\ref{tab:ch_parm} {\bf (GB: we need some more explanation of the simulations... next round)}. 
+%
+%{\bf (TLG2: Setup 2 was the setup I used initially.
+The parameters I used was influenced by arrays I was accustomed to at that point, like WSRT.
+Later on as you became more involved you suggested setup 1 (and even later still gain model 3) as it is more realistic for redundant low-frequency arrays.
+Redoing the red stefcal sims with these parameters did not influence the results though if memory serves me right -- not sure if I still have the old results though to confirm.
+We never redid the sims with PCG as the sims took quite a bit of time (much longer than unvectorized red stefcal, which as we know takes quite a long time) and there would ultimately be now difference in result, the calibration algorithm's speed does not depend on the observational frequency etc...
+Moreover, Fig.
+5 is completely independent of the chosen simulated parameters as it depends only on the analytic expression of the Hessian.
+Not sure how to phrase or explain the different setups in the paper, as I would like to avoid redoing the simulations at this stage, except if the referee actually requests this.
+PLEASE HELP HERE.
+UNSURE HOW TO MOTIVATE OR EXPLAIN THE SETUPS.
+NOT SURE WHERE SUCH A DISCUSSION BELONGS.)}
+We used the complex LM-algorithm described in Section~\ref{sec:red_wirtinger} to calibrate the corrupted visibilities. To invert the modified
+Hessian we used the CG method, with and without a preconditioner. We made use of the Jacobian preconditioner. 
+Figure~\ref{fig:kappa} shows us that preconditioning reduces the condition number of $\bmH$ to a small constant value and therefore effectively eliminates
+it from equation~\ref{eq:cg_bound}, i.e. equation~\ref{eq:cg_bound} reduces to
+\begin{equation}
+\label{eq:cg_bound2}
+O(m). 
+\end{equation}
+This result is confirmed by Fig.~\ref{fig:cg_itr}. Fig.~\ref{fig:cg_itr} 
+shows us that the number of major iterations needed by the PCG method to invert 
+$\bmH$ is independent of the number of antennas in the array and that it is much less than the 
+dimension of $\bmH$. 
+%We found that if we used a preconditioner then the spectral condition number $\kappa$ of $\bmH$ improves and becomes completely independent of the number of antennas in the array (see Figure~\ref{fig:kappa_itr})
+
+%{\bf - (GB: but it is also much smaller, right?)}.
+%{\bf - (TLG: The CG method is normally good to use on classes of matrix problems whose spectral condition number is independent of problem size (this is an ideal property to have to be able to use the CG method).
+In the problem we are studying here, it is the case even without preconditioning.
+However, the condition number is high without preconditioning which unfortunately leads to a large number of major iterations (albeit fairly independent of problem size).
+If we require more than $P$ major iterations, then there is no point to using CG as conventional algorithms will outperform it in any case (here we require more than $P$ major iterations).
+So here preconditioning is not required to ensure independence of problem size (which is generally the case), but to lower the condition number, which in turn reduces the number of major iterations required to be less than $P$ in a non-trivial way.
+The relationship which exists between $\kappa$ and the number of major iterations needed depends on starting conditions and the eigenvalue distribution among other things.
+Since both (preconditioning and not preconditioning) are constant we need to plot both $\kappa$ (this is only an indication of the order of the problem and does not indicate the exact number of iterations due to the complicated relationship of the two) and the number of major iterations used.
+Because $\kappa$ is constant it does not nec tell you everything about the number of major iterations).
+I think all of this becomes to complicated to explain in the paper and almost irrelevant for the point I am trying to make in the paper: kappa is constant implies number of major iterations is independent of problem size.
+Moreover, empirical sims show it to be less than $P$.
+Which is why I try to focus on the basics, hope this makes more sense now?)}
+
+%, which is a non-singular matrix, in Section~\ref{sec:scn} and Section~\ref{sec:sparsity} respectively.
+%It is interesting to note, that the Hessian $\bH$ itself is singular (from empirical tests) and that the PCG method can in principal be be used to pseudo-invert it (see Appendix~\ref{sec:normal}).
+
 \begin{figure*}
@@
 \subfigure[Iterations required before and after Preconditioning]{\includegraphics[width=0.47\textwidth]{./cg_itr.pdf}\label{fig:cg_itr}}
-\caption{The graphs presented here were generated using the second experimental setup in Table~\ref{tab:ch_parm}.
-In the left panel we have plotted the spectral condition number $\kappa$ of the modified Hessian $\bmH$, before (magenta and red curves) and after preconditioning (blue curve), at different SNR values as a function of the
-number of antennas $N$ in the array.
-The right hand panel confirms this result.
-In the right hand panel we have plotted the number of major iterations (line 6 of Algorithm \ref{algo:CG}) required by the conjugate
-gradient method to invert $\bmH$, before (magenta and red curves) and after preconditioning (blue and green curves), at different SNR values as a function of the number of antennas $N$ in the array.
+\caption{The graphs presented here were generated using the simulations discussed in Section~\ref{sec:red_wirtinger}.
+Left: spectral condition number $\kappa$ of the modified Hessian $\bmH$ as a function of $N$, before (magenta and red curves) and after (blue curve) preconditioning, at different SNR values.
+Right: number of major iterations required by the conjugate
+gradient method to invert $\bmH$ as a function of the number of antennas $N$ in the array, before (magenta and red curves) and after preconditioning (blue and green curves), at different SNR values.
+Both plots show that the Jacobian preconditioner definitely speed up the conjugate gradient method.
 \label{fig:kappa_itr}}
@@
 
-\subsubsection{Sparsity}
-\label{sec:sparsity}
-% \begin{enumerate}
-% \item $\boldsymbol{H}$: is a $(4N-2)\times(4N-2)$ or a $P \times P$ matrix. $N$ denotes the number of antennas and $P$ denotes the number of parameters. This matrix
-% contains $6N^2 - 2N - 2$ non-zero entries. It contains $16N^2 - 16N + 4$ entries. We can construct this matrix with $4N^2 -  4N$ or $\frac{1}{4} P^2 -1$ elementary operations.
-% \item $\boldsymbol{A}$: is a $(2N-1)\times(2N-1)$ matrix. This matrix has $N^2 + N - 1$ non-zero entries. We can construct this matrix with $2\frac{1}{2} (N^2 -  N)$ elementary operations.  
-% \item $\boldsymbol{B}$: is a $(2N-1)\times(2N-1)$ matrix. This matrix has $2N^2 - 2N$ non-zero entries. We can construct this matrix with $1\frac{1}{2} (N^2 -  N)$ elementary operations.
-% \item $\boldsymbol{C}$: is a $N\times N$ matrix. This matrix has $N$ non-zero entries. We can construct this matrix with $N^2-N$ elementary operations.
-% \item $\boldsymbol{D}$: is a $N \times (N-1)$ matrix. This matrix has $\frac{1}{2} (N^2 -  N)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations. 
-% \item $\boldsymbol{E}$: is a $(N-1) \times (N-1)$ matrix. This matrix has $(N-1)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.  
-% \item $\boldsymbol{F}$: is a $N \times N$ matrix. This matrix has $N^2 - N$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.
-% \item $\boldsymbol{G}$: is a $N \times (N-1)$ matrix. This matrix has $\frac{1}{2} (N^2 -  N)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.
-% \item $\boldsymbol{0}$: is a $(N-1) \times (N-1)$ all zero matrix. 
-% \end{enumerate}
-
-%We can construct $\boldsymbol{C}-\boldsymbol{G}$ in $O(N^2)$
-
-The second factor that determines the computational complexity of the conjugate gradient method when used to perform redundant calibration is the number of non-zero
-entries in the modified Hessian matrix $\bmH$. 
-%We have also shown in section that the first factor, the spectral condition number of $\bH$, can be reduced 
-%to a constant by employing preconditioning and can therefore be completely eliminated from the computational complexity equation. 
-A useful matrix quantity which we will use in this section is the sparsity factor $\gamma$ of a matrix, which is the ratio between the number of zero entries and the total number of entries in a matrix.
-We
-can therefore define $\gamma$ as follows:
-\begin{equation}
- \gamma = \left (1 - \frac{m}{P^2} \right ) 
-\end{equation}
-It is trivial to determine an analytic expression for the sparsity ratio of $\bmH$ when our redundant array is in a regular east-west geometric configuration
-The sparsity and the asymptotic sparsity ratio of $\bmH$ are equal to
+%The graphs presented here were generated using the second experimental setup in Table~\ref{tab:ch_parm}.
+
+Let us now shift our attention towards the remaining factor $m$.
+To get a usefull idea of the computational complexity of CG, we must relate $m$ and $P$.
+This can be achieved by utilizing the modified Hessian's measure of sparsity $\gamma$:
+\begin{equation}
+ \gamma = \left (1 - \frac{m}{P^2} \right ),
+\end{equation}
+where $m$ is the number of non-zero entries in $\bmH$  and $P^2$ is total number of matrix elements.
+
+For a regular east-west geometric configuration the sparsity and the asymptotic sparsity of $\bmH$ can be derived analytically:
 \begin{align}
@@
 \end{align}
-However when our redundant array makes use of a more complicated geometric layout it becomes much more difficult to construct such expressions, as $\bmH$ itself 
-becomes much harder to express analytically.
-
-In Fig.~\ref{fig:gamma}, we plot the empirically determined sparsity ratios for three different geometric layouts as a function of the number antennas in the array.
-Although, we cannot explicitly calculate
-the limit of the hexagonal and square layout curves, Fig.~\ref{fig:gamma} seems to indicate that they do exist. Moreover, we also plot the left most equation
-of Eq.~\eqref{eq:gamma} (red dashed line) and the right most equation of Eq.~\eqref{eq:gamma} (black dotted curve) in Fig.~\ref{fig:gamma}.
-
-We can now use the sparsity factor to compute the order of the computational complexity of inverting $\bmH$ (assuming the effect of $\kappa$ is negligible) with
-\begin{equation}
-P^{c} = (1 - \gamma)P^2.
-\end{equation}
-Solving for $c$ we obtain
+For more complicated array layouts, however, there is no straightforward analytical solution and we empirically determined the sparsity ratios for three different geometric layouts as a function of the number antennas in the array (see Figure~\ref{fig:gamma}).
+
+It now follows that 
+%{\bf (GB: can't follow why it is this way...)}
+%{\bf (TLG: Rewrote this to emphasize the link between $m$ and $P$, the computational complexity is only useful, if it is expressed in terms of the number of parameters one whises to solve for.
+Multiplying the $1-$ sparsity measure with $P^2$ gives you $m$.
+Moreover,since it is smaller than one, this tells us that we are doing better than $P^2$.
+The reduction in complexity stems from sparse vector matrix multiplication.)}
+\begin{equation}
+P^{c} = m = (1 - \gamma)P^2,
+\end{equation}
+which leads to
 \begin{align}
@@
 \end{align}
-
-The evaluation of the limit in the rightmost equation of Eq.~\eqref{eq:c} is evaluated as follows: note that
-\begin{equation}
-\label{eq:int_limit}
-\lim_{N\rightarrow\infty} P^{\log_{P}(1-\gamma)} = 1-\gamma_{\infty},
-\end{equation}
-which is only possible if 
-\begin{equation}
-\lim_{N\rightarrow\infty} \log_{P}(1-\gamma) = 0,
-\end{equation}
-implying that $c_{\infty}$ is equal to two.
-
-Eq.~\eqref{eq:int_limit}, implies that we can only compute $c_{\infty}$ in the rightmost equation if $\lim_{N\rightarrow \infty} \gamma$ exists. 
-In Fig.~\ref{fig:c}, we plot the empirically determined value of $c$ for three different geometric layouts as a function of the number antennas in the array.
-Furthermore, in Fig.~\ref{fig:c}, the analytic value of $c$ for the regular east-west grid is plotted as a dashed red curve, while the limit of this analytic curve is plotted with a dotted black curve.
-
-Eq.~\eqref{eq:gamma}, Eq.~\eqref{eq:c} and Fig.~\ref{fig:sparsity} indicate that the computational complexity, under the assumption that $\lim_{N\rightarrow \infty} \gamma$ exists, is asymptotically bounded by $O(P^2)$.
-However, the computational complexity converges very slowly to its asymptotic value, and in general is equal to $O(P^{c})$, where $c \sim 1.7$ if a Hexagonal geometric layout containing less than 200 antennas is employed.
+The computational complexity is therefore asymptotically bounded by $O(P^2)$, although it converges very slowly to its asymptotic value, and in general is equal to $O(P^{c})$, where $c \sim 1.7$ in the case of an hexagonal geometric layout with $N < 200$ (see Figure~\ref{fig:c}).
 
@@
 \subfigure[$c$]{\includegraphics[width=0.46\textwidth]{./comp_order.pdf}\label{fig:c}}
-\caption{In the left panel we plot the sparsity ratio $\gamma$ (the number of zero entries relative to the total number of entries in a matrix) of the modified Hessian $\bmH$ as a function of the number
-of antennas $N$ in the array for different geometries, namely an hexagonal grid (blue circles), a square grid (green crosses) and a regular east-west grid (red circles).
-For the regular east-west grid we are able to construct an analytical expression for $\gamma$ (the dashed red line).
-Taking the limit
-of this analytic expression produces the black dotted line. In the right panel we plot the order of the computational cost $c$ for inverting $\bmH$ as a function of the number 
-of antennas in the array for different geometries. We use the same color scheme in the right and left panels. The red dashed line 
-in the right panel is associated with the theoretical curve of $c$ we computed for the east-west regular grid. The limit of the red-dashed curve is two, which we 
-have plotted with a black dotted line.
+\caption{Left: the sparsity ratio $\gamma$ of the modified Hessian $\bmH$ as a function of the number of antennas $N$ for an hexagonal (blue circles), square (green crosses) and regular east-west (red circles) array geometry.
+The red--dashed and black--dotted lines show the analytical expression for $\gamma$ and its limit for the regular east-west grid case (see text for details).
+Right: the order of the computational cost $c$ for inverting $\bmH$ as a function of $N$ for different array geometries (same colour scheme as in the left panel).
+The red--dashed and black--dotttedlines are the analytical expression of $c$ and its limit in the east-west regular grid case.
 \label{fig:sparsity}} 
@@
 
-\subsection{Comparison}
-\label{sec:comparison}
-We are now in a position to make a final comparison between the ADI and PCG methods.
-Inverting $\bH$ by assuming it is a diagonal matrix (i.e.
-using the ADI method) is very cheap, while inverting $\bH$ using PCG is more expensive.
-On the other side of the coin,  one inversion with PCG is much more
-accurate than using the approximate inverse which is associated with the ADI method.
-The question is now, if the inverse using the PCG method is much more accurate will the LM implementation using it require much less LM-iterations than its ADI counterpart.
-Moreover, will the number
-of LM iterations PCG requires be so few that it manages to beat the ADI method after all.
-To answer this we first need to compute the number of iterations, $k$, the LM implementation associated with the ADI method can make in the
-same time it takes the LM implementation associated with the PCG method to perform a single iteration. We can compute this with
+We are now finally able to compare the computational complexity of redundant \textsc{StEfCal} and the PCG method.
+Redundant \textsc{StEfCal} is computationally inexpensive as it just needs to invert a diagonal matrix, however, the PCG inversion is accurate and, therefore, may require fewer iterations, ultimately leading to a faster convergence.
+We computed the theoretical number of redundant \textsc{StEfCal} iterations $\Delta k$, in excess of the number of LM (implementing PCG) iterations required, for LM to outperform redundant \textsc{StEfCal}
 \begin{align}
 \label{eq:k} 
- kP &\approx P^{1.7}, & k &\approx P^{0.7}.
-\end{align}
-Next we need to measure the number of LM iterations we save by using the PCG method and compare it to Eq.\eqref{eq:k}. We make this comparison 
-in Fig.~\ref{fig:out_diff}. Fig.~\ref{fig:out_diff} tells us that the ADI method will outperform the PCG method as the number of LM iterations we save by taking the 
-full inverse is not enough to exceed $k$. In this comparison we have only taken into account the cost of inverting $\bH$. The PCG method might, however, still 
-turn out to be quite useful in practice as it would be more straightforward to apply it to other calibration use cases.
+ (k+\Delta k)P >& \, kP^c, & \Delta k >& \, k(P^{c-1}-1),
+\end{align}
+and compared it to the empirically obtained average excess number of iterations.
+The results are displayed in Figure~\ref{fig:out_diff} which shows that redundant \textsc{StEfCal} outperforms the PCG method.
+We note that, in this comparison, we have only taken into account the cost of inverting $\bmH$ and ignored the cost of preconditioning.
+%The PCG method is still useful in practice as it would be more straightforward to apply it to other calibration use cases 
+%
+%{\bf (GB: what do you mean with this statements? Which cases?)}.
+%{\bf (TLG: I was hoping that Jonathan would have published his work by now.
+In it he shows that the pointing error formulation has off-diagonal significant elements.
+We should prob just remove this sentence as I can not reference anything as of yet.)}
 
@@
 \centering
-\subfigure[Number of LM iterations required by ADI and PCG.]{\includegraphics[width=0.47\textwidth]{./outerloop.pdf}\label{fig:outerloop}}
+\subfigure[Number of LM iterations required by redundant \textsc{StEfCal} and PCG.]{\includegraphics[width=0.45\textwidth]{./outerloop.pdf}\label{fig:outerloop}}
 % V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
-\subfigure[Difference between ADI and PCG.]{\includegraphics[width=0.47\textwidth]{./diff.pdf}\label{fig:diff}}
-\caption{The graphs presented here were generated using the second experimental setup in Table~\ref{tab:ch_parm}. In the left panel we have the number of LM iterations 
-that is required for the LM algorithm using the ADI (green and blue curve) and the PCG (magenta and red curve) methods to converge to a parameter error tolerance of 1e-6 whilst using different 
-SNR values as a function of the number of antennas $N$ in the array.
-In the right panel we have the average amount of LM iterations (difference between the ADI and PCG curves in the left panel) we save (green and blue curve) by computing the full-inverse with the PCG method whilst using different SNR values.
-The black curve is associated with $k$, the number of iterations the LM implementation using the ADI method can perform in the same time it takes the LM implementation using 
-the PCG method to perform one iteration.
+\subfigure[Difference.]{\includegraphics[width=0.45\textwidth]{./diff.pdf}\label{fig:diff}}
+\caption{The graphs presented here were generated using the simulations discussed in Section~\ref{sec:red_wirtinger}.
+Left: number of LM iterations required by redundant \textsc{StEfCal} (green and blue curves) and the PCG (magenta and red curves) methods to converge to a parameter error tolerance of $10^{-6}$ whilst using different SNR values as a function of the number of antennas $N$ in the array.
+Right: average amount of LM iterations (difference between the redundant \textsc{StEfCal} and PCG curves in the left panel) saved (green and blue curve) by computing the full-inverse with the PCG method whilst using different SNR values.
+The black curve is $\Delta k$, the theoretical number
+of redundant \textsc{StEfCal} iterations that are needed in excess of the number of LM iterations used for the LM algorithm to 
+outperform redundant \textsc{StEfCal}. For the black curve we assumed $c=1.7$ and that $k$ could be approximated with the 
+the magenta curve plotted in the left panel.
 \label{fig:out_diff}} 
@@
 \section{Conclusion}
-In Section~\ref{sec:main_sec} we showed that instead of using the traditional phase-and-amplitude approach to perform redundant calibration we can actually use a purely complex 
-framework. In this framework, the unknown variables and their conjugates are treated as independent variables that have to be estimated. 
-We then derived the GN and LM update steps using this complex framework (see Section~\ref{sec:complex_GN_LM}).
-In this paper we focused more on the LM algorithm than on the GN algorithm.
-The LM algorithm exhibits better convergence properties than the
-GN algorithm. This has the immediate consequence that the amount of algorithmic layers that are needed to perform redundant calibration are reduced.
-
-The most expensive operation associated with the LM algorithm (as well as the GN algorithm) is the inversion of $\bH$.
-We compared the computational complexity of two previously proposed algorithms to efficiently invert $\bH$ for the redundant calibration use case in Section~\ref{sec:results}, namely 
-ADI \citep{Marthi2014} and PCG \citep{Liu2010}.
-We showed that the computational complexity of the ADI method is $\mathbb{O}(P)$, while the computational complexity of the PCG method is asymptotically bounded
-by $\mathbb{O}(P^2)$. Both of these algorithms perform much better than stock standard inversion, which has a computational complexity of $\mathbb{O}(P^3)$.
-Although the ADI method does execute faster than the PCG method it is far less robust as it would be less straightforward to apply to other calibration uses cases. This is due to the fact that,
-the diagonal approximation we make with ADI might not be accurate enough in general.
-
-Lastly, we showed how \citet{Marthi2014}, \citet{Salvini2014} and \citet{Smirnov2015} relate to each other (see Section~\ref{sec:red_c}).
-The ADI method associated with skymodel-based calibration (i.e.
-\textsc{StEfCal}) is presented
-in \citet{Salvini2014}. \citet{Smirnov2015} showed that it is possible to re-derive this skymodel-based ADI method through complex optimization. In particular, they showed that if 
-we approximate the $\bH$ associated with the LM algorithm with its diagonal then we can rederive the ADI update rules presented in \citet{Salvini2014}. 
-\citet{Marthi2014} actually proposes to use the ADI method to perform redundant calibration.
-This fact is not immediately clear when superficially inspecting \citet{Marthi2014}.
-The redundant calibration algorithm presented in \citet{Marthi2014} was derived by taking the partial derivatives of
-the objective function towards the unknown parameters, setting the intermediate results to zero and then solving for the unknowns. In Section~\ref{sec:red_c}, we showed that 
-this algorithm is in actual fact an extension of the skymodel-based ADI method presented in \citet{Salvini2014} to the redundant calibration use case (i.e.
-we showed that this algorithm is in actual fact redundant \textsc{StEfCal}).
-We also re-derived the aforementioned redundant calibration algorithm by applying the diagonal approximation approach presented in \citet{Smirnov2015}, which serves as yet another additional confirmation
-that this algorithm is in fact redundant \textsc{StEfCal} (i.e. that it is some kind of ADI method).
-
-% fiducial initial guess 
-% 
-% The following partial derivatives are now easily computed
-% 
-% \begin{eqnarray}
-% \frac{c_{ij}}{\partial \eta_p} &=& y_{i-j} e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}\\ 
-% \frac{c_{ij}}{\partial \varphi_p} &=&  -i y_{i-j} e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}\\
-% \frac{c_{ij}}{\partial \eta_q} &=& y_{i-j} e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}\\ 
-% \frac{c_{ij}}{\partial \varphi_q} &=&  i y_{i-j} e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}\\
-% \frac{y_{i-j}}{\partial \varphi_q} &=&  e^{\eta_p - i \varphi_p} e^{\eta_q - i \varphi_q}
-% \end{eqnarray}
-% 
-% 
-% The Wirtinger derivative is used in the last equation.
-% 
-% \begin{eqnarray}
-% c_{ij} &\approx& c_{ij}^0 + \Delta \eta_p(y_{i-j}^0 e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0}) + \Delta \eta_q(y_{i-j}^0 e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0})\\
-% && -i\Delta\varphi_p (y_{i-j}^0e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0}) +i\Delta\varphi_q (y_{i-j}^0e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0})\\
-% && y_{i-j}^1(y_{i-j}^0e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0})\\
-% &\approx&  c_{ij}^0 + e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_p+ \Delta \eta_q - i\Delta\varphi_p + i\Delta\varphi_p))
-% \end{eqnarray}
-% 
-% \begin{equation}
-% \label{eq:residual}
-% r_{ij} = \delta_{ij} = c_{ij}-c_{ij}^0 = e^{\eta_p^0 - i \varphi_p^0} e^{\eta_q^0 - i \varphi_q^0}(y_{i-j}^1 + y_{i-j}^0( \Delta \eta_p+ \Delta \eta_q - i\Delta\varphi_p + i\Delta\varphi_p)) 
-% \end{equation}
-% 
-% Eq.~\ref{eq:residual} allows us to construct the following linear system:
-% 
-% \begin{equation}
-% \boldsymbol{J}[\boldsymbol{\Delta \eta},\boldsymbol{\Delta \varphi},\boldsymbol{\Delta y}]^T = \boldsymbol{r}, 
-% \end{equation}
-% 
-% where $\boldsymbol{J}$ is equal to 
-% \begin{equation}
-% \boldsymbol{J} = \bigg [\overbrace{\frac{c_{pq}}{\partial \eta_p}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial \varphi_p}}^{i= 1\cdots N},~\overbrace{\frac{c_{pq}}{\partial y_{t}}}^{t=1\cdots r_s} \bigg ] \bigg \} [pq] = 1\cdots N_{\textrm{b}} (p<q)
-% \end{equation}
-% or
-% \begin{equation}
-% \boldsymbol{J} = \bigg [\frac{c_{pq}}{\partial \boldsymbol{\eta}},~\frac{c_{pq}}{\partial \boldsymbol{\varphi}},~\frac{c_{pq}}{\partial \boldsymbol{y}} \bigg ] \bigg \} [pq] = 1\cdots N_{\textrm{b}} (p<q).
-% \end{equation}
-% 
-% The last column is again a Wirtinger derivative.
-% 
-% Which means we can obtain the least-squares estimate as follows:
-% 
-% \begin{equation}
-% [\boldsymbol{\Delta \eta},\boldsymbol{\Delta \varphi},\boldsymbol{\Delta y}]^T = [\boldsymbol{J}^H\boldsymbol{J}]^{-1}\boldsymbol{J}^H\boldsymbol{r}.
-% \end{equation}
+\label{sec:conclusions}
+In this paper we have formulated the calibration of redundant interferometric arrays using the complex optimization formalism 
+\citep{Smirnov2015}. We derived the associated complex Jacobian and the GN and LM parameter update steps.
+We also showed that the LM parameter update step can be simplified to obtain an ADI type of algorithm \citep[][]{Salvini2014,Marthi2014}.
+Our code implementation of this algorithm (redundant \textsc{StEfCal}) is publicly available at 
+\url{https://github.com/Trienko/heracommissioning/blob/master/code/stef.py}. 
+We note that, in its current implementation, redundant \textsc{StEfCal} does not solve for the degeneracies inherent to redundant calibration  \citep{Zheng2014,Kurien2016} which will be the subject of future work.
+Compared to current redundant calibration algorithms, redundant \textsc{StEfCal} is more robust to inital conditions and allows for an easier, massive parallelization.
+
+We investigated the computational cost of redundant \textsc{StEfCal} and compared it with the performance of the PCG method (approach suggested by \cite{Liu2010}).
+We found that, although the PCG method method greatly improves the speed of redundant calibration, it still significant underperforms in comparison to redundant \textsc{StEfCal}.
+
+The characteristics of redundant \textsc{StEfCal} make it an appealing calibration algorithm for large redundant arrays like HERA \citep{deboer2017}, CHIME \citep{Bandura2014}, HIRAX \citep{Newburgh2016} or even hybrid arrays like the MWA \citep{Tingay2013} in its updated phase.
+
+\section*{Acknowledgements}
+G.B.
+acknowledges support from the Royal Society and the Newton Fund under grant NA150184.
+This work is based on the research supported in part by the National Research Foundation of South Africa (grant No.
+103424).
+This research is supported by the South African Research Chairs Initiative of the Department of Science and Technology and National Research Foundation,
 
@@
 
-\section{\textsc{LOGCAL}}
-\label{sec:logcal}
-We can rewrite Eq.~\eqref{eq:vis_red} to \citep{Liu2010}
-\begin{eqnarray}
-\ln |d_{pq}| &=& \ln |g_p| + \ln |g_q| + \ln |y_{\phi_{pq}}| + s_{pq} \label{eq:logcal_amp}\\
-\angle d_{pq} &=& \angle g_p - \angle g_q + \angle \phi_{pq} + t_{pq} \label{eq:logcal_phase}
-\end{eqnarray}
-where
-\begin{align}
-s_{pq} &= \mathscr{R} \left \{1 + \frac{n_{pq}}{g_p\conj{g_q}y_{\phi_{pq}}} \right \}, & t_{pq} &= \mathscr{I}\left \{1 + \frac{n_{pq}}{g_p\conj{g_q}y_{\phi_{pq}}} \right \}.
-\end{align}
-Moreover, $\mathscr{R}\{*\}$ and $\mathscr{I}\{*\}$ denote the real and imaginary operators.
-
-With the aid of Eq.~\eqref{eq:logcal_amp} we can construct the following linear system:
-\begin{equation}
-\label{eq:log_system}
-\bdelta = \bJ\bzeta + \bs, 
-\end{equation}
-where
-\begin{align}
-\left [ \bdelta \right ]_{\alpha_{pq}} &= \ln |d_{pq}|, & \left [ \bs \right ]_{\alpha_{pq}} = s_{pq} 
-\end{align}
-and
-\begin{equation}
-\left [ \bzeta \right ]_{j} = \begin{cases} \ln |g_j| & \textrm{if}~j\leq N \\ \ln |y_{j-N}| & \textrm{otherwise} \end{cases}. 
-\end{equation}
-Furthermore,
-\begin{equation}
-\bJ = 
-\begin{bmatrix}
-\bN & \bM
-\end{bmatrix}
-\end{equation}
-with 
-\begin{equation}
-[\bN]_{\alpha_{pq},j} = \begin{cases}
-       1 & \textrm{if}~(p=j)~\textrm{or}~(q=j)\\
-       0 & \textrm{otherwise}
-      \end{cases}
-\end{equation}
-and
-\begin{equation}
-[\bM]_{\alpha_{pq},j} = \begin{cases}
-       1 & \textrm{if}~\phi_{pq}=j\\
-       0 & \textrm{otherwise}
-      \end{cases}
-\end{equation}
-
-The least squares solution to Eq.\eqref{eq:log_system}, under the incorrect assumption that $s_{pq}$ is normally distributed, is given by
-\begin{equation}
-\bzeta = (\bA^T\bA)^{-1} \bA^T\bdelta. 
-\end{equation}
-How to correctly deal with the fact that $s_{pq}$ is not normally distributed is discussed in \citep{Liu2010}. A similar linear system can be constructed for Eq.~\eqref{eq:logcal_phase}. 
-From the least squares solutions to the linear systems associated with Eq.~\eqref{eq:logcal_amp} and Eq.~\eqref{eq:logcal_phase} we can compute the antenna gains as well as the true visibilities.
-
-\section{\textsc{lincal}}
-\label{sec:lincal}
-%Assume that $\eta_p,\eta_q,\widetilde{\eta}_{\phi_{pq}},\varphi_p,\varphi_q$ and $\widetilde{\varphi}_{\phi_{pq}}$ are real valued variables.
-We can rewrite Eq.~\eqref{eq:vis_red_vec} to \citep{Liu2010}
-\begin{equation}
-\label{eq:lincal_vec}
-\bd = \bv(\bvarrho) + \bn, 
-\end{equation}
-where 
-\begin{align}
-\left [ \bv(\bvarrho) \right ]_{\alpha_{pq}} &= e^{\eta_p - i \varphi_p} e^{\widetilde{\eta}_{\phi_{pq}}- i \widetilde{\varphi}_{\phi_{pq}}} e^{\eta_q + i \varphi_q}\\
-&= v_{pq},
-\end{align}
-and
-\begin{equation}
-\bvarrho = 
-\begin{bmatrix}
-\bmath{\eta} \\
-\bvarphi \\
-\widetilde{\bmath{\eta}} \\
-\widetilde{\bvarphi}
-\end{bmatrix},\qquad
-\begin{aligned}
-\bmath{\eta} &= [\eta_1, \eta_2,\cdots \eta_N]^T\\
- \bmath{\varphi} &= [\varphi_1, \varphi_2,\cdots \varphi_N]^T\\
- \widetilde{\bmath{\eta}} &= [\widetilde{\eta}_1, \widetilde{\eta}_2,\cdots \widetilde{\eta}_L]^T\\
- \widetilde{\bmath{\varphi}} &= [\widetilde{\varphi}_1, \widetilde{\varphi}_2,\cdots \widetilde{\varphi}_L]^T
-\end{aligned}.
-\label{eq:help}
-\end{equation}
-Note that the vectors $\bmath{\eta},~\bvarphi,~\widetilde{\bmath{\eta}}$ and $\widetilde{\bvarphi}$ are real valued.
-
-The first order Taylor expansion of Eq.~\eqref{eq:lincal_vec} around the fiducial point $\bvarrho_0$ is equal to
-\begin{equation}
-\label{eq:d_lin}
-\bd = \bv_0 + \bmJ_0 \Delta \bvarrho  + \bn.
-\end{equation}
-If we calculate the residual, dropping the reference to a specific fiducial point, we obtain: 
-\begin{equation}
-\label{eq:r_lin}
-\br = \bd - \bv = \bmJ \Delta \bvarrho + \bn.
-\end{equation}
-Moreover,
-\begin{equation}
-\label{eq:jac_upper_lincal}
-\bmJ = \begin{bmatrix}
-       \bN & \bM 
-       \end{bmatrix},
-\end{equation}
-where
-\begin{align}
-\bN &= \begin{bmatrix} \bO & \bP \end{bmatrix}, & \bM &= \begin{bmatrix} \bQ & \bR \end{bmatrix}, 
-\end{align}
-\begin{equation}
- [\bO]_{\alpha_{pq},j} =  \frac{\partial v_{pq}}{\partial \eta_j} = \begin{cases} 
-    v_{pq} &\textrm{if}~(p=j)~\textrm{or}~(q=j)\\
-    0 & \textrm{otherwise}
-   \end{cases},
-\end{equation}
-\begin{equation}
- [\bP]_{\alpha_{pq},j} =  \frac{\partial v_{pq}}{\partial \varphi_j} = \begin{cases}
-                                                                        i v_{pq} &\textrm{if}~ p=j\\
-                                                                        -i v_{pq} &\textrm{if}~ q=j \\
-                                                                        0 & \textrm{otherwise} 
-                                                                       \end{cases}
-\end{equation}
-\begin{equation}
- [\bQ]_{\alpha_{pq},j} =  \frac{\partial v_{pq}}{\partial \widetilde{\eta}_j} = \begin{cases} 
-    v_{pq} &\textrm{if}~\phi_{pq}=j\\
-    0 & \textrm{otherwise}
-   \end{cases},
-\end{equation}
-\begin{equation}
- [\bR]_{\alpha_{pq},j} =  \frac{\partial v_{pq}}{\partial \widetilde{\varphi}_j} = \begin{cases} 
-    i v_{pq} &\textrm{if}~\phi_{pq}=j\\
-    0 & \textrm{otherwise}
-   \end{cases}.
-\end{equation}
-Augmenting the residual by taking into account its conjugate results in 
-\begin{equation}
-\breve{\br} = \bJ \Delta \bvarrho + \bn, 
-\end{equation}
-where $\breve{\br} = \begin{bmatrix}\br^T & \conj{\br}^T \end{bmatrix}^T$ and 
-\begin{equation}
-\label{eq:Jac_lin}
-\bJ = 
-\begin{bmatrix}
- \bmJ\\  
- \conj{\bmJ}
-\end{bmatrix}.
-\end{equation}
-The solution to 
-\begin{equation}
- \min_{\Delta \bvarrho} \| \breve{\br} - \bJ \Delta \bvarrho\|_F^2,
-\end{equation}
-is equal to 
-\begin{equation}
-\label{eq:LINCAL_update}
-\Delta \bvarrho = (\bJ^H\bJ)^{-1} \bJ^H\breve{\br}. 
-\end{equation}
-Eq.~\eqref{eq:LINCAL_update} is used to iteratively determine a new estimate of the parameter vector $\bvarrho$ as follows
-\begin{equation}
-\label{eq:new_rho}
-\bvarrho_{k+1} = \bvarrho_k + \Delta \bvarrho_k. 
-\end{equation}
-Note that we do not explicitly denote the $k$ index in Eq.~\eqref{eq:LINCAL_update}. We update our parameter vector until convergence is reached.
-
-\section{Conjugate Gradient Method}
-\label{sec:conj_grad}
-The CG method was independently discovered by \citet{Hestenes1973} and \citet{Stiefel1952}. They later published a joint paper, which is now considered as the seminal
-reference on CG \citep{Hestenes1952}. The iterative CG method is used to solve a particular class of linear system. The CG method is generally used to solve
-\begin{equation}
-\label{eq:linear_system}
-\bb = \bA\bx,
-\end{equation}
-where $\bb\in\mathbb{C}^P$ is a known complex column vector of size $P$, $\bx\in\mathbb{C}^P$ is an unknown complex column vector with the same length as $\bb$, and $\bA\in\mathbb{C}^{P\times P}$ is a square positive-definite Hermitian matrix with the appropriate dimensions.
-Moreover, using the CG method to solve Eq.~\eqref{eq:linear_system} will only be computationally advantageous if $\bA$ is also sparse (contain many zero entries).
-Using the iterative CG method to solve large sparse linear sytems was popularized by \citet{Reid1971}. A good tutorial on the CG method can be found in \citep{Shewchuk1994}.
-The iterative CG method is presented in Algorithm \ref{algo:CG}.
-
-%If $\bA$ is a square positive semi-definite Hermitian matrix, then the conjugate gradient method only converges if $\bb$ is in the column range of $\bA$ (cite Lu2016).
-
-%\begin{algorithm}
-%\caption{Conjugate Gradient Method.
-Inputs: $\bA$, $\bb$, a starting value for $\bx$, maximum number of iterations $k_{\textrm{max}}$ and an error tolerance $\epsilon<1$ Output: $\bx$ the solution to Eq.~\eqref{eq:linear_system}.
-\citep{Shewchuk1994}.}\label{algo:CG}
-%\begin{algorithmic}[1]
-%\State $k \gets 0$
-%\State $\br \gets \bb - \bA\bx$
-%\State $\bp \gets \br$
-%\State $\delta_{\textrm{new}} \gets \br^H\br$
-%\State $\delta_0 \gets \delta_{\textrm{new}}$
-%\While{$k\leq k_{\textrm{max}}$ and $\delta_{\textrm{new}} > \epsilon^2\delta_0$}
-%\State $\alpha \gets \frac{\delta_{\textrm{new}}}{\bp^H\bA\bp}$
-%\State $\bx \gets \bx + \alpha\bp$
-%\State $\br \gets \br - \alpha\bA\bp$
-%\State $\delta_{\textrm{old}} \gets \delta_{\textrm{new}}$ 
-%\State $\delta_{\textrm{new}} \gets \br^H\br$
-%\State $\beta \gets \frac{\delta_{\textrm{new}}}{\delta_{\textrm{old}}}$
-%\State $\bp \gets \br + \beta\bp$
-%\State $k \gets k + 1$
-%\EndWhile
-%\end{algorithmic}
-%\end{algorithm}
-
-\subsection{Computational Complexity}
-\citet{Kaniel1966} was able to determine convergence bounds for CG using Chebyshev polynomials. A much more in depth analysis of CG convergence can be found in
-\citep{Sluis1986}. As can be seen from Algorithm \ref{algo:CG}, the CG method makes use of a major outer loop (line 6 of Algorithm \ref{algo:CG}). It turns out that at worst 
-the number of major iterations that CG requires to converge\footnote{If the $\bA$-norm is used to calculate the error.} is proportional to $\sqrt{\kappa}$, where $\kappa$ is the spectral condition number of the matrix $\bA$.
-The mathematical definition of $\kappa(\bA)$ is
-\begin{equation}
-\label{eq:kappa}
-\kappa(\bA) = \frac{\lambda_{\textrm{max}}}{\lambda_{\textrm{min}}}, 
-\end{equation}
-where $\lambda_{\textrm{max}}$ and $\lambda_{\textrm{min}}$ respectively denote the largest and the smallest eigenvalue of $\bA$.
-Moreover, the most expensive operation that is conducted within each major loop of Algorithm~\ref{algo:CG} is the vector-matrix product $\bA\bp$, which has a complexity of $\mathbb{O}(m)$, where $m$ is the number of non-zero entries in $\bA$.
-The computational complexity of CG is therefore 
-\begin{equation}
-\label{eq:cg_bound}
-\mathbb{O}(\sqrt{\kappa}m). 
-\end{equation}
-It should now also be clear to the reader why the CG method is especially applicable to sparse linear systems. In sparse linear systems $m < P^2$.
-
-\subsection{Preconditioning}
-\label{sec:precon}
-Preconditioning is a technique which allows us to improve the condition number of a matrix. Let $\bM$ be a positive-definite Hermitian matrix.
-We can now indirectly solve for $\bx$ by using 
-\begin{equation}
-\label{eq:prec}
-\bM^{-1}\bA\bx = \bM^{-1}\bb.
-\end{equation}
-If $\bM$ is a good preconditioner then $\kappa(\bM^{-1}\bA) \ll \kappa(\bA)$, i.e. a good preconditioner lowers the condition number of a matrix which in turn increases the convergence speed 
-of CG (see Eq.~\eqref{eq:cg_bound}).
-In this paper, we denote matrix inversion with $(*)^{-1}$.
-The best possible choice of $\bM$ would therefore be $\bA$, which would lead to $\kappa$ becoming unity (as the largest and smallest eigenvalue of the identity matrix are equal to one).
-If $\kappa$ is one then Eq.~\eqref{eq:cg_bound} simplifies and becomes $\mathbb{O}(m)$.
-However, if we knew the inverse of $\bA$ then obtaining $\bx$ would be trivial and there would be no need for the CG method in the first place.
-In short, a good preconditioner $\bM$ must approximate $\bA$ fairly accurately. However, if it is not much less expensive to 
-compute the inverse of $\bM$ than that of $\bA$ we actually gain nothing (see Eq.~\eqref{eq:prec}).
-
-If the diagonal entries of $\bA$ are much more significant than its off-diagonal entries then the 
-Jacobian preconditioner is a good choice of $\bM$. The Jacobian preconditioner is computed as follows:
-\begin{equation}
-\bM = \bA\odot\bI, 
-\end{equation}
-where $\bI$ is the identity matrix and $\odot$ denotes the Hadamard product. Since the Jacobian preconditioner is a diagonal matrix its inverse can be obtained trivially.
-
-\subsection{Normal Equation}
-\label{sec:normal}
-Lets assume that instead of Eq.~\eqref{eq:linear_system} we have a more general linear system, i.e.
-\begin{equation}
- \bb = \bB\bx,
-\end{equation}
-where $\bb\in\mathbb{C}^K$, $\bx\in\mathbb{C}^P$  and $\bB\in\mathbb{C}^{K \times P}$, with $K > P$. We can use the normal equation 
-\begin{equation}
-\label{eq:normal_equation}
-\bB^H\bB\bx = \bB^H\bb, 
-\end{equation}
-to estimate the $\bx$ that minimizes
-\begin{equation}
-\label{eq:norm_ls}
-\|\bb-\bB\bx\|_F^2. 
-\end{equation}
-In this paper, we denote the Hermitian transpose with $(*)^H$ and the Frobenius norm with $\|*\|_F$.
-
-There are a few important observations that we can make while inspecting Eq.~\eqref{eq:normal_equation}:
-\begin{enumerate}
-\item $\bB^H\bB$ is a square positive semi-definite Hermitian matrix by construction.
-\item $\bB^H\bb$ is in the column range of $\bB^H\bB$; this follows trivially from 
-\begin{equation}
-\bB^H\bb \in \mathcal{R}(\bB^H) \implies \bB^H\bb \in \mathcal{R}(\bB^H\bB).   
-\end{equation}
-\end{enumerate}
-In this paper, we denote the column range of a matrix $\bA$ with $\mathcal{R}(\bA)$.
-
-Recall that we mentioned that in general we can only solve Eq.~\eqref{eq:linear_system} with CG if $\bA$ is a square positive-definite Hermitian matrix.
-The CG method actually has an even broader applicability: it can be used to solve Eq.~\eqref{eq:linear_system} even if $\bA$ is a Hermitian positive semi-definite
-square matrix as long as $\bb$ is in the column range of $\bA$ \citep{Lu2015}.
-In light of \citet{Lu2015}, the two observations in the numbered list we presented in the previous paragraph imply that Eq.~\eqref{eq:normal_equation}
-can be solved by employing the CG method. The solution returned by CG is, however, only unique if $\bB^H\bB$ is positive-definite (invertible). The solution
-obtained by CG, whether it is unique or not, does minimize Eq.~\eqref{eq:norm_ls} in a least squares sense.
-
-Note that if $\bA$ is positive semi-definite, its smallest eigenvalue is zero. This implies that Eq.~\eqref{eq:kappa} is not defined. \citet{Lu2015} show that when $\bA$
-is a square positive semi-definite Hermitian matrix we do not use the spectral condition number to bound the convergence complexity of CG, but the general spectral condition
-number instead. The general spectral condition number is very similar to the spectral condition number; the only difference being that in Eq.~\eqref{eq:kappa} we replace  
-$\lambda_{\textrm{min}}$ with $\lambda_{\textrm{min}}^+$, which denotes the smallest positive eigenvalue.
-
-\section{Redundant Wirtinger Calibration - $\bJ$, $\bH$ and $\bJ^H\breve{\br}$}
+{\bf (GB: We need to bring back the logcal and lincal derivation using the geometric function.)}
+%{\bf (TLG: I have removed logcal and lincal as well as the PCG section.
+My main concern about this is that the aim of these sections was to show off the usefulness of the geometry function, it simplifies, at least from a mathematecial perspective, the derivation of the older algorithms.
+This insight has now been lost.
+On the other hand keeping them and trying to make them fit, which may cause further delays, at this point simply no longer seems worht the effort.
+I have incorporated the remaining ideas left in the PCG section into the main text.)}.
+%
+% \section{\textsc{LOGCAL}}
+% \label{sec:logcal}
+% We can rewrite equation~\ref{eq:vis_red} to \citep{Liu2010}
+% \begin{eqnarray}
+% \ln |d_{pq}| &=& \ln |g_p| + \ln |g_q| + \ln |y_{\phi_{pq}}| + s_{pq} \label{eq:logcal_amp}\\
+% \angle d_{pq} &=& \angle g_p - \angle g_q + \angle \phi_{pq} + t_{pq} \label{eq:logcal_phase}
+% \end{eqnarray}
+% where
+% \begin{align}
+% s_{pq} &= \mathscr{R} \left \{1 + \frac{n_{pq}}{g_p\conj{g_q}y_{\phi_{pq}}} \right \}, & t_{pq} &= \mathscr{I}\left \{1 + \frac{n_{pq}}{g_p\conj{g_q}y_{\phi_{pq}}} \right \}.
+% \end{align}
+% Moreover, $\mathscr{R}\{*\}$ and $\mathscr{I}\{*\}$ denote the real and imaginary operators. 
+% 
+% With the aid of equation~\ref{eq:logcal_amp} we can construct the following linear system:
+% \begin{equation}
+% \label{eq:log_system}
+% \bdelta = \bJ\bzeta + \bs, 
+% \end{equation}
+% where
+% \begin{align}
+% \left [ \bdelta \right ]_{\alpha_{pq}} &= \ln |d_{pq}|, & \left [ \bs \right ]_{\alpha_{pq}} = s_{pq} 
+% \end{align}
+% and
+% \begin{equation}
+% \left [ \bzeta \right ]_{j} = \begin{cases} \ln |g_j| & \textrm{if}~j\leq N \\ \ln |y_{j-N}| & \textrm{otherwise} \end{cases}. 
+% \end{equation}
+% Furthermore,
+% \begin{equation}
+% \bJ = 
+% \begin{bmatrix}
+% \bN & \bM
+% \end{bmatrix}
+% \end{equation}
+% with 
+% \begin{equation}
+% [\bN]_{\alpha_{pq},j} = \begin{cases}
+%        1 & \textrm{if}~(p=j)~\textrm{or}~(q=j)\\
+%        0 & \textrm{otherwise}
+%       \end{cases}
+% \end{equation}
+% and
+% \begin{equation}
+% [\bM]_{\alpha_{pq},j} = \begin{cases}
+%        1 & \textrm{if}~\phi_{pq}=j\\
+%        0 & \textrm{otherwise}
+%       \end{cases}
+% \end{equation}
+% 
+% The least squares solution to Eq.\ref{eq:log_system}, under the incorrect assumption that $s_{pq}$ is normally distributed, is given by
+% \begin{equation}
+% \bzeta = (\bA^T\bA)^{-1} \bA^T\bdelta. 
+% \end{equation}
+% How to correctly deal with the fact that $s_{pq}$ is not normally distributed is discussed in \citep{Liu2010}. A similar linear system can be constructed for equation~\ref{eq:logcal_phase}. 
+% From the least squares solutions to the linear systems associated with equation~\ref{eq:logcal_amp} and equation~\ref{eq:logcal_phase} we can compute the antenna gains as well as the true visibilities.
+% 
+% \section{\textsc{lincal}}
+% \label{sec:lincal}
+% Assume that $\eta_p,\eta_q,\widetilde{\eta}_{\phi_{pq}},\varphi_p,\varphi_q$ and $\widetilde{\varphi}_{\phi_{pq}}$ are real valued variables.
+% We can rewrite equation~\ref{eq:vis_red_vec} to \citep{Liu2010}
+% \begin{equation}
+% \label{eq:lincal_vec}
+% \bd = \bv(\bvarrho) + \bn, 
+% \end{equation}
+% where 
+% \begin{align}
+% \left [ \bv(\bvarrho) \right ]_{\alpha_{pq}} &= e^{\eta_p - i \varphi_p} e^{\widetilde{\eta}_{\phi_{pq}}- i \widetilde{\varphi}_{\phi_{pq}}} e^{\eta_q + i \varphi_q}\\
+% &= v_{pq},
+% \end{align}
+% and
+% \begin{equation}
+% \bvarrho = 
+% \begin{bmatrix}
+% \bmath{\eta} \\
+% \bvarphi \\
+% \widetilde{\bmath{\eta}} \\
+% \widetilde{\bvarphi}
+% \end{bmatrix},\qquad
+% \begin{aligned}
+% \bmath{\eta} &= [\eta_1, \eta_2,\cdots \eta_N]^T\\
+%  \bmath{\varphi} &= [\varphi_1, \varphi_2,\cdots \varphi_N]^T\\
+%  \widetilde{\bmath{\eta}} &= [\widetilde{\eta}_1, \widetilde{\eta}_2,\cdots \widetilde{\eta}_L]^T\\
+%  \widetilde{\bmath{\varphi}} &= [\widetilde{\varphi}_1, \widetilde{\varphi}_2,\cdots \widetilde{\varphi}_L]^T
+% \end{aligned}.
+% \label{eq:help}
+% \end{equation}
+% Note that the vectors $\bmath{\eta},~\bvarphi,~\widetilde{\bmath{\eta}}$ and $\widetilde{\bvarphi}$ are real valued.
+% 
+% The first order Taylor expansion of equation~\ref{eq:lincal_vec} around the fiducial point $\bvarrho_0$ is equal to
+% \begin{equation}
+% \label{eq:d_lin}
+% \bd = \bv_0 + \bmJ_0 \Delta \bvarrho  + \bn.
+% \end{equation}
+% If we calculate the residual, dropping the reference to a specific fiducial point, we obtain: 
+% \begin{equation}
+% \label{eq:r_lin}
+% \br = \bd - \bv = \bmJ \Delta \bvarrho + \bn.
+% \end{equation}
+% Moreover,
+% \begin{equation}
+% \label{eq:jac_upper_lincal}
+% \bmJ = \begin{bmatrix}
+%        \bN & \bM 
+%        \end{bmatrix},
+% \end{equation}
+% where
+% \begin{align}
+% \bN &= \begin{bmatrix} \bO & \bP \end{bmatrix}, & \bM &= \begin{bmatrix} \bQ & \bR \end{bmatrix}, 
+% \end{align}
+% \begin{equation}
+%  [\bO]_{\alpha_{pq},j} =  \frac{\partial v_{pq}}{\partial \eta_j} = \begin{cases} 
+%     v_{pq} &\textrm{if}~(p=j)~\textrm{or}~(q=j)\\
+%     0 & \textrm{otherwise}
+%    \end{cases},
+% \end{equation}
+% \begin{equation}
+%  [\bP]_{\alpha_{pq},j} =  \frac{\partial v_{pq}}{\partial \varphi_j} = \begin{cases}
+%                                                                         i v_{pq} &\textrm{if}~ p=j\\
+%                                                                         -i v_{pq} &\textrm{if}~ q=j \\
+%                                                                         0 & \textrm{otherwise} 
+%                                                                        \end{cases}
+% \end{equation}
+% \begin{equation}
+%  [\bQ]_{\alpha_{pq},j} =  \frac{\partial v_{pq}}{\partial \widetilde{\eta}_j} = \begin{cases} 
+%     v_{pq} &\textrm{if}~\phi_{pq}=j\\
+%     0 & \textrm{otherwise}
+%    \end{cases},
+% \end{equation}
+% \begin{equation}
+%  [\bR]_{\alpha_{pq},j} =  \frac{\partial v_{pq}}{\partial \widetilde{\varphi}_j} = \begin{cases} 
+%     i v_{pq} &\textrm{if}~\phi_{pq}=j\\
+%     0 & \textrm{otherwise}
+%    \end{cases}.
+% \end{equation}
+% Augmenting the residual by taking into account its conjugate results in 
+% \begin{equation}
+% \breve{\br} = \bJ \Delta \bvarrho + \bn, 
+% \end{equation}
+% where $\breve{\br} = \begin{bmatrix}\br^T & \conj{\br}^T \end{bmatrix}^T$ and 
+% \begin{equation}
+% \label{eq:Jac_lin}
+% \bJ = 
+% \begin{bmatrix}
+%  \bmJ\\  
+%  \conj{\bmJ}
+% \end{bmatrix}.
+% \end{equation}
+% The solution to 
+% \begin{equation}
+%  \min_{\Delta \bvarrho} \| \breve{\br} - \bJ \Delta \bvarrho\|_F^2,
+% \end{equation}
+% is equal to 
+% \begin{equation}
+% \label{eq:LINCAL_update}
+% \Delta \bvarrho = (\bJ^H\bJ)^{-1} \bJ^H\breve{\br}. 
+% \end{equation}
+% equation~\ref{eq:LINCAL_update} is used to iteratively determine a new estimate of the parameter vector $\bvarrho$ as follows
+% \begin{equation}
+% \label{eq:new_rho}
+% \bvarrho_{k+1} = \bvarrho_k + \Delta \bvarrho_k. 
+% \end{equation}
+% Note that we do not explicitly denote the $k$ index in equation~\ref{eq:LINCAL_update}.
+% Moreover, equation~\ref{eq:LINCAL_update} is the GN update associated with the following minimization problem \citep{Kurien2016}
+% \begin{equation}
+% \label{eq:least_squares_lincal}
+% \min_{\bmath{\varrho}} \|\breve{\br}\| = \min_{\bmath{\varrho}} \|\breve{\bd} - \breve{\bv}(\bmath{\varrho})\|.
+% \end{equation}
+
+% \section{Conjugate Gradient Method}
+% 
+% \subsection{Normal Equation}
+% \label{sec:normal}
+% Lets assume that instead of equation~\ref{eq:linear_system} we have a more general linear system, i.e.
+% \begin{equation}
+%  \bb = \bB\bx,
+% \end{equation}
+% where $\bb\in\mathbb{C}^K$, $\bx\in\mathbb{C}^P$  and $\bB\in\mathbb{C}^{K \times P}$, with $K > P$. We can use the normal equation 
+% \begin{equation}
+% \label{eq:normal_equation}
+% \bB^H\bB\bx = \bB^H\bb, 
+% \end{equation}
+% to estimate the $\bx$ that minimizes
+% \begin{equation}
+% \label{eq:norm_ls}
+% \|\bb-\bB\bx\|_F^2. 
+% \end{equation}
+% In this paper, we denote the Hermitian transpose with $(*)^H$ and the Frobenius norm with $\|*\|_F$.
+% 
+% There are a few important observations that we can make while inspecting equation~\ref{eq:normal_equation}:
+% \begin{enumerate}
+% \item $\bB^H\bB$ is a square positive semi-definite Hermitian matrix by construction.
+% \item $\bB^H\bb$ is in the column range of $\bB^H\bB$; this follows trivially from 
+% \begin{equation}
+% \bB^H\bb \in \mathcal{R}(\bB^H) \implies \bB^H\bb \in \mathcal{R}(\bB^H\bB).   
+% \end{equation}
+% \end{enumerate}
+% In this paper, we denote the column range of a matrix $\bA$ with $\mathcal{R}(\bA)$.
+% 
+% Recall that we mentioned that in general we can only solve equation~\ref{eq:linear_system} with CG if $\bA$ is a square positive-definite Hermitian matrix.
+The CG method actually has an even broader applicability: it can be used to solve equation~\ref{eq:linear_system} even if $\bA$ is a Hermitian positive semi-definite
+% square matrix as long as $\bb$ is in the column range of $\bA$ \citep{Lu2015}.
+In light of \citet{Lu2015}, the two observations in the numbered list we presented in the previous paragraph imply that equation~\ref{eq:normal_equation}
+% can be solved by employing the CG method. The solution returned by CG is, however, only unique if $\bB^H\bB$ is positive-definite (invertible). The solution
+% obtained by CG, whether it is unique or not, does minimize equation~\ref{eq:norm_ls} in a least squares sense.
+% 
+% Note that if $\bA$ is positive semi-definite, its smallest eigenvalue is zero. This implies that equation~\ref{eq:kappa} is not defined. \citet{Lu2015} show that when $\bA$
+% is a square positive semi-definite Hermitian matrix we do not use the spectral condition number to bound the convergence complexity of CG, but the general spectral condition
+% number instead. The general spectral condition number is very similar to the spectral condition number; the only difference being that in equation~\ref{eq:kappa} we replace  
+% $\lambda_{\textrm{min}}$ with $\lambda_{\textrm{min}}^+$, which denotes the smallest positive eigenvalue.
+
+\section{Derivation of redundant Wirtinger calibration}
 \label{sec:analytic}
-If we apply the definition in Eq.~\eqref{eq:Jacobian} to Eq.~\eqref{eq:least_squares_complex} we obtain the following analytic result:
+If we apply the definition in equation~\ref{eq:Jacobian} to equation~\ref{eq:least_squares_complex} we obtain the following analytic result:
 \begin{equation}
@@
 
-%\begin{align}
-%\left [ \bM_1 \right ]_{\alpha_{pq},j} &= y_{\phi_{pq}}\conj{g_q}\delta_p^j, & \left [ \bM_2 \right ]_{\alpha_{pq},j} &= g_p\conj{g_q}\delta_{\phi_{pq}}^j 
-%\end{align}
-%\begin{equation}
-%\left [ \bN_1 \right ]_{\alpha_{pq},j} = g_py_{\phi_{pq}}\delta^j_q. 
-%\end{equation}
-We use $\bzero$ to denote an all zero matrix. It is now trivial to compute the Hessian $\bH$ by using Eq.~\eqref{eq:Jacobian_red}. If we substitute Eq.~\eqref{eq:Jacobian_red} into $\bJ^H\bJ$
+We use $\bzero$ to denote an all zero matrix. It is now trivial to compute the Hessian $\bH$ by using equation~\ref{eq:Jacobian_red}. If we substitute equation~\ref{eq:Jacobian_red} into $\bJ^H\bJ$
 we obtain 
@@
 
-% \begin{equation}
-% \boldsymbol{J}^H\breve{\boldsymbol{d}} = 
-% \begin{bmatrix}
-% \sum_{k\neq i } g_k x_{ik}d_{ik}\\
-% \sum_{pq \in \mathcal{I}_j} \conj{g}_p g_q d_{pq}\\
-% \downarrow^{*}
-% \end{bmatrix}
-% \begin{matrix}% matrix for right braces 
-% \coolrightbrace{\sum g_k x_{\zeta_{ik}}d_{ik}}{i = 1\cdots N}\\
-% \coolrightbrace{\sum \conj{g}_p g_q d_{pq}}{j = 1\cdots L}\\
-% \vphantom{\downarrow^{*}}
-% \end{matrix}
-% \end{equation}
+Furthermore,
+\begin{align}
+\label{eq:identities}
+\frac{1}{3}\bJ\breve{\bz} &= \breve{\bv}, & \bJ^H\breve{\bv} &= (\bI\odot\bH)\breve{\bz} 
+\end{align}
+The above identities can be trivially established by mechanically showing that the left hand side of each expression in equation~\ref{eq:identities} is equal to its right hand side.
+In equation~\ref{eq:identities}.
 
@@
 
-\section{Redundant \textsc{StEfCal} and the ADI method}
+\section{ADI}
 \label{sec:red_stef_ADI}
-The basic skymodel-based \textsc{StEfCal} update step is equal to  the leftmost term in Eq.~\eqref{eq:g_update} (barring $\alpha$) \citep{Salvini2014}.
-Assume without any loss of generality that the array is in an east-west regular grid. Furthermore, assume that $\boldsymbol{d}$ (see Eq.~\eqref{eq:vis_red_vec}) has been re-ordered to
+The basic skymodel-based \textsc{StEfCal} update step is equal to  the leftmost term in equation~\ref{eq:g_update} (barring $\alpha$) \citep{Salvini2014}.
+Assume without any loss of generality that the array is in an east-west regular grid. Furthermore, assume that $\boldsymbol{d}$ (see equation~\ref{eq:vis_linear_definition}) has been re-ordered 
+and that the result of this re-ordering is the following
 \begin{equation}
@@
 
-Eq.~\eqref{eq:vis_red_vec} can now be rewritten as 
+Equation~\ref{eq:vis_linear_definition} can now be rewritten as 
 \begin{equation}
@@
 \end{equation}
-Substituting Eq.~\eqref{eq:RHR} and Eq.~\eqref{eq:RHd} into Eq.~\eqref{eq:y_final} and then write out the analytic expression of $y_i$ we obtain
-the leftmost term in Eq.~\eqref{eq:y_update} (barring $\alpha$).
+Substituting equation~\ref{eq:RHR} and equation~\ref{eq:RHd} into equation~\ref{eq:y_final} and simplifying the result leads to the leftmost term in equation~\ref{eq:y_update} (if we bar $\alpha$ and consider
+only the $i$-th entry of $\by$).
 \label{lastpage}

### 2017-11-12 paperG2.tex
added the almost final version of the paper

@@
 
-The second advantage of redundant array configurations is that they, by definition, they measure the same spatial scales {\bf (GB: perhaps worth having the Van cittert-Zernike theorem here?)} regardless of what the sky brightness distribution looks like.
+The second advantage of redundant array configurations is that they, by definition, they measure the same spatial scales
+
+{\bf (GB: perhaps worth having the Van cittert-Zernike theorem here?)}
+
+{\bf (TLG: Leave the decision to Gianni.)}
+
+regardless of what the sky brightness distribution looks like.
 This means that the corruptions that are introduces by the instrument can be calibrated - up to some extent - without knowing anything about the sky brightness distribution \citep[][]{Noordam1982,Wieringa1992}.
@@
 
-In this paper we extend the stefcal calibration algorithm to the redundant use case using the complex optimization formalism recently introduced by \cite{Smirnov2015}. 
-We also perform a proper computational analysis of the recently proposed (preconditioned conjugate gradient) PCG method when it is applied to the redundant calibration problem.
-We also compare these two approaches.
-
-present a new algorithm to calibrate redundant arrays based on the complex optimization formalism recently introduced by \cite{Smirnov2015}.
-With respect to current algorithms, it is more robust to initial conditions, it is comparatively fast and {\bf ?????...}.
-We also present an analysis of its computational complexity showing that the LM algorithm shows better convergence properties than the GN currently used \cite{Zheng14}.
-
-The paper is organized as follows: we review the complex calibration formalism by \citet{Smirnov2015,Kurien2016} in Section~\ref{sec:sky_wirtinger}, we extend it to the redundant case in Section~\ref{sec:red_wirtinger}, we test the algorithm performances in Section~\ref{sec:results} and we conclude in Section\ref{sec:conclusions}.
+%In this paper we extend the stefcal calibration algorithm to the redundant use case using the complex optimization formalism recently introduced by \cite{Smirnov2015}. 
+%We also perform a proper computational analysis of the recently proposed (preconditioned conjugate gradient) PCG method when it is applied to the redundant calibration problem.
+%We also compare these two approaches.
+
+{\bf (TLG: I have only modified the paragraph below in the introduction to try and point out the point of the paper. I hope this makes 
+more sense now given the the restructuring GB suggested.)}
+
+In this paper we present a new algorithm to calibrate redundant arrays based on the complex optimization formalism recently introduced by \cite{Smirnov2015}. 
+With respect to current algorithms, it is more robust to initial conditions, while remaining comparatively fast. We also show 
+that given certain approximations this new algorithm reduces to the redundant calibration equivalent of the StEfCal algorithm \citep{Salvini2014}.
+In this paper we will refer to this simplified version of the new algorithm as redundant StEfCal. We also 
+investigate the speed-up that the preconditioned conjugate gradient (PCG) method provides if it is employed by the new algorithm \citep{Liu2010}.
+A comparison between the computational complexity of the optimized new algorithm and redundant StEfCal is also performed.
+
+{\bf (TLG: The following papers on redundant calibration has to be cited in the paper \citep{Noorishad2012} and Sievers2017. It was previously given in the 
+intro and was subsequently removed in the previous iteration made by GB. Unsure were they now belong.)}
+
+%Redundant StEfCal is much faster 
+%than WRC. 
+%We also investigate the speed-up that is achievable by employing
+%Moreover, a proper computational complexity comparison is made between redundant StEfCal and the recently proposed  when applied to the redundant calibration problem is conducted.
+%With respect to current algorithms, it is more robust to initial conditions, it is comparatively fast and {\bf ?????...}. 
+%We also compare these two approaches.
+%We also present an analysis of its computational complexity showing that the LM algorithm shows better convergence properties than the GN currently used \cite{Zheng14}.
+
+The paper is organized as follows: we review the complex calibration formalism by \citet{Smirnov2015} in Section~\ref{sec:sky_wirtinger}, we extend it to the redundant case in Section~\ref{sec:red_wirtinger}, we test the algorithm performances in Section~\ref{sec:results} and we conclude in Section\ref{sec:conclusions}.
 
@@
 
+{\bf (GB: Trienko, I am not sure about where to place the paragraph below till the end of the subsection...
+probably because I'm not sure what ``even" and ``odd" \textsc{StEfCal} steps are...
+can we add the GN step after equation~\ref{eq:stef_lambda}???)}
+
+{\bf (TLG: I have rewritten the above to make the odd and even stefcal steps more clear. I have also move the GN simplified equation before the 
+stefcal explanation. GB to rewrite.)}
+
 %A special case of solutions referred to as the alternating direction implicit \citep[ADI,][]{Salvini2014} algorithm is obtained by choosing $\lambda = 0$ \citep{Smirnov2015}.
@@
 where the number of model parameters to be solved of is now $P = 2(N+L)$, since redundant calibration is a complex problem.
+{\bf (GB: is the following important? I thought it could be removed...) We plot $L$ and $P$ for two different geometric layouts in Fig~\ref{fig:pl}.
+The analytic expressions for the red and the blue curve in Fig.~\ref{fig:ln} are \citep{Camps2003}
+\begin{equation}
+L = 2N-\frac{1}{2}\sqrt{12N-3}-\frac{1}{2} 
+\end{equation}
+and
+\begin{equation}
+L = N - 1 
+\end{equation}
+respectively.}
+
+{\bf (TLG: Done.)}
+
 %The number of model parameters that we need to solve for in equation~\ref{eq:least_squares_red} is $P = 2(N+L)$, since redundant calibration is a complex problem.
 
-% \subsection{Limitations}
-% \label{sec:scope}
-% We only deal with per-timeslot and frequency channel calibration. No solution smoothing  
-% is applied between the different timeslots and channels. The reader interested in this is referred to \citep{Zheng2014}. 
-% Moreover, we do not explicitly deal with any of the degeneracies normally associated with redundant calibration \citep{Zheng2014,Kurien2016}.
+{\bf (TLG: This section got removed during a previous edit. It needs to be put back somewhere????)}
+
+\subsection{Limitations}
+ \label{sec:scope}
+ We only deal with per-timeslot and frequency channel calibration. No solution smoothing  
+ is applied between the different timeslots and channels. The reader interested in this is referred to \citep{Zheng2014}. 
+ Moreover, we do not explicitly deal with any of the degeneracies normally associated with redundant calibration \citep{Zheng2014,Kurien2016}.
 Similar to the approach taken in \citep{Marthi2014}, we assume that the degeneracies present in the redundant calibration solutions are removed with a final round of skymodel-based calibration.
@@
 As mentioned in Section~\ref{sec:sky_wirtinger}, equation~\ref{eq:least_squares_red} can be solved by splitting the problem into its real and
-imaginary parts and then solve for the real and imaginary parts of the unknown parameters separately \citep{Wieringa1992,Liu2010,Zheng2014}. In Appendix~\ref{sec:lincal} we review this approach.
+imaginary parts and then solve for the real and imaginary parts of the unknown parameters separately \citep{Wieringa1992,Liu2010,Zheng2014}.
+In Appendix~\ref{sec:logcal} and~\ref{sec:lincal} we review this approach.
 %As already alluded to in Section~\ref{sec:sky_wirtinger}, we can solve equation~\ref{eq:least_squares_red} in one of two ways.
@@
 
-%{\bf (GB: derived in which way? It would e useful to say in order to differentiate from us)}.
+{\bf (GB: derived in which way? It would be useful to say in order to differentiate from us)}.
+{\bf (TLG: Added)}.
 
@@
 When $\alpha=\frac{1}{3}$, the last term of equation~\ref{eq:two_thrids} and equation~\ref{eq:alpha} are equal.
+
+{\bf (GB: is this true?
+TRIENKO, I MANAGED TO REACH ONLY TILL THIS POINT.
+I THINK IT IS TIME I HAND THIS BACK TO YOU AND YOU MAKE AN EFFORT TO TIE IT UP.
+i THINK WE CAN GO FROM HERE TO A MINOR REVISION NEXT TIME)}
+{\bf (TLG: tried to rephrase is it clear now?}
 

### 2017-11-17 paper.tex
added most of Jonathan's comments

@@
 Observations of the redshifted 21-cm line from the epoch of reionization have recently motivated the construction of low 
-frequency radio arrays with highly redundant configurations that provide an alternative calibration strategy - 
-''redundant calibration" - and boosts sensitivity on specific spatial scales. In this paper we formulate 
-calibration of redundant interferometric arrays as a complex optimization problem \citep[i.e.,][]{Smirnov2015}. 
-We then solve this optimization problem via the Levenberg-Marquardt algorithm. This calibration approach is more 
-robust to initial conditions than current algorithms and, leveraging on an approximate matrix inversion, allows for further optimization and an efficient implementation (``redundant \textsc{StEfCal}").
-We also investigated the use of the preconditioned conjugate gradient method as an alternative optimization strategy, instead of
-an approximate matrix inverse, but found that its computational performance is not competitive with respect to ``redundant \textsc{StEfCal}".
+frequency radio arrays with highly redundant configurations. These configurations provide an alternative calibration strategy - 
+''redundant calibration" - and boosts sensitivity on specific spatial scales. In this paper, we formulate 
+calibration of redundant interferometric arrays as a complex optimization problem {\bf SJK: suggests removing citation} \citep[i.e.,][]{Smirnov2015}. 
+We solve this optimization problem via the Levenberg-Marquardt algorithm. This calibration approach is more 
+robust to initial conditions than current algorithms and, by leveraging an approximate matrix inversion, allows for further optimization and an efficient implementation (``redundant \textsc{StEfCal}").
+{\bf (SJK: ...boasts competitive [big O notation here] scaling.
+We coin the term "redundant StEfCal" for our algorithm and implementation)}.
+We also investigated using the preconditioned conjugate gradient method as an alternative to the approximate matrix inverse, but found that its computational performance is not competitive with respect to ``redundant \textsc{StEfCal}".
 The efficient implementation of this new algorithm is made publicly available.
@@
 %and that can be separated by leveraging only on the different spectral properties (citations needed!!!). 
-As miscalibrated foreground emission leads to artifacts that can jeopardize the EoR signal \citep[e.g.,][]{grobler2014,barry2016,ewall-wice2016}, an exquisite interferometric calibration is required.
+As miscalibrated foreground emission leads to artifacts that can jeopardize the EoR signal \citep[e.g.,][]{grobler2014,barry2016,ewall-wice2016}, an exceptional interferometric calibration is required.
 Such calibration needs an accurate knowledge of both the sky emission and the instrumental response \citep[e.g.,][]{Smirnov2011c}, unless arrays are deployed in a redundant configuration, i.e.
-with receiving elements placed on regular grids so that they measure many times the same sky brightness emission.
-Leveraging on this property, redundancy is a promising path to achieve high accurate interferometric calibration \citep[][]{Noordam1982,Wieringa1992,Pearson1984,Liu2010,Noorishad2012,Marthi2014,Sievers2017}.
+with receiving elements placed on regular grids so that multiple baselines measure the same sky brightness emission.
+Leveraging this property, redundancy is a promising path to achieve highly accurate interferometric calibration \citep[][]{Noordam1982,Wieringa1992,Pearson1984,Liu2010,Noorishad2012,Marthi2014,Sievers2017}.
 Moreover, redundant configurations provide a sensitivity boost on a selected sample of spatial scales that can be tuned to be the highest signal--to--noise ratio (SNR) EoR modes \citep[][]{Parsons2012,Dillon2016}.
@@
 
-In this paper we present a new algorithm to calibrate redundant arrays based on the complex optimization formalism recently introduced by \cite{Smirnov2015}. 
+In this paper we, present a new algorithm to calibrate redundant arrays based on the complex optimization formalism recently introduced by \cite{Smirnov2015}. 
 With respect to current algorithms, it is more robust to initial conditions, while remaining comparatively fast.
@@
 
-The paper is organized as follows: we review the complex calibration formalism by \citet{Smirnov2015} in Section~\ref{sec:sky_wirtinger}, we extend it to the redundant case in Section~\ref{sec:red_wirtinger}, we present some computational complexity results in Section~\ref{sec:pcg} and we conclude in Section~\ref{sec:conclusions}.
+The paper is organized as follows: we review the complex calibration formalism by \citet{Smirnov2015} in Section~\ref{sec:sky_wirtinger}, we extend it to the redundant case in Section~\ref{sec:red_wirtinger}, we present some computational complexity results in Section~\ref{sec:pcg} and we presnt our conclusions in Section~\ref{sec:conclusions}.
 
@@
 
-In a radio interferometer, the true sky visibilities $y_{pq}$ measured by a baseline formed by antenna $p$ and $q$ are always ``corrupted" by the non-ideal response of the receiver, which is often incorporated into a single, receiver-based complex number $g$.
+In a radio interferometer, the true sky visibilities $y_{pq}$ measured by a baseline formed by antenna $p$ and $q$ are always ``corrupted" by the non-ideal response of the receiver, which is often incorporated into a single, receiver-based complex number $g$ (i.e.
+an antenna gain).
 The observed visibility $d_{pq}$ is therefore given by \citep{ME1,ME2,RRIME1}
@@
 
+{\bf JSK: I would introduce the alpha quantity on its own before attaching it as an index in the definition of d.
+The reason for this is that eq.
+4 and 5 seem very complicated without first understanding the meaning of alpha.
+Could have a paragraph starting ``In order to express equation 1 in vector form, we first introduce $\alpha$ which maps...''}
+
+{\bf TLG: Was like this earlier, got changed during the edits. GB let me know if it should be changed back.}
+
 Radio interferometric calibration aims to determine the best estimate of $\bg = [g_1,g_2,\cdots,g_N]^T$ in order to correct the data and, following equation~\ref{eq:vis_linear_definition}, can be formulated as a non linear least-squares optimization problem:
@@
 In standard interferometric calibration, $y_{pq}$ is assumed to be known at some level, for instance through the observation of previously known calibration sources.
-The knowledge of $y_{pq}$ is often named ``sky model".
-
-The gradient-based minimization algorithms that are generally used to solve non-linear least-squares problems are generally solved by using gradient-based minimization algorithms (i.e.
-Gauss--Newton -- GN -- or Levenberg--Marquardt -- LM) that require the model ($\bv$ in equation~\ref{eq:least_squares}) to be differentiable towards each parameter.
+
+Non-linear least-squares problems are generally solved by using gradient-based minimization algorithms (i.e.
+Gauss--Newton (GN) -- or Levenberg--Marquardt (LM)) that require the model ($\bv$ in equation~\ref{eq:least_squares}) to be differentiable towards each parameter.
 When the least squares problem is complex, it becomes less straightforward to apply these gradient-based minimization methods, as
@@
 In order to circumvent the differentiability conundrum associated with complex least squares problems, standard interferometric calibration divides the complex optimization problem into its real and imaginary parts and solves for the real and imaginary parts of the unknown model parameters separately.
-\citet{Smirnov2015} showed, however, that this approach is not needed if complex calculus \citep{Wirtinger1927} is directly adopted.
+\citet{Smirnov2015} showed, however, that this approach is not needed if complex calculus \citep{Wirtinger1927} is directly adopted {\bf JSK: employed}.
 The Wirtinger derivatives are defined as:
@@
 
-The complex LM update is very similar, the major difference being a damping factor $\lambda$ is introduced:
+The complex LM update is very similar, the major difference being the introduction of a damping parameter, $\lambda$:
 \begin{equation}
@@
 In this paper, the identity matrix is denoted by $\bI$.
-Moreover, the Hadamard product and the Hermition transpose are denoted by $\odot$ and $(\cdot)^H$.
+Moreover, the Hadamard product and the Hermition transpose are denoted by $\odot$ and $(\cdot)^H$ respectively.
 Note the use of the Wirtinger derivatives in equation~\ref{eq:jac_entries}.
@@
 \breve{\bg}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bg}_k, \nonumber \\
- &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bg}_k, 
+ &= \rho \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\rho)\breve{\bg}_k, 
 \label{eq:stef_alpha}  
 \end{align}
-where $\alpha = \frac{1}{1+\lambda}$. Note that equation~\ref{eq:one_half} and equation~\ref{eq:stef_alpha} are   
+where $\rho = \frac{1}{1+\lambda}$. Note that equation~\ref{eq:one_half} and equation~\ref{eq:stef_alpha} are   
 not dependent on $\breve{\br}$.
@@
 Interestingly enough, if $\lambda = 0$ we obtain the odd parameter update step of \textsc{StEfCal}\footnote{$k\in\{0,2,\cdots\}$}, and if $\lambda=1$ (which corresponds
-to $\alpha=\frac{1}{2}$) we obtain the even parameter update step of \textsc{StEfCal}\footnote{$k\in\{1,3,\cdots\}$} \citep[\textsc{StEfCal},][]{Mitchell:MWA-cal,Salvini2014}.
-In the \textsc{StEfCal} algorithm, the measurement equation, equation~\ref{eq:vis_definition}, is linearized by assuming that the gains are known, but that their conjugates are not.
-Under this assumption, the system of equations become linear and the conjugates of the gains can be obtained straightforwardly.
-Starting from the latest value of the gain conjugates, an updated estimate of the gains can be obtained and the succession continues until convergence is reached.
-Alternating between solving and fixing different sets of parameters (which is exactly what \textsc{StEfCal} does) is in general referred to 
-as the alternating direction implicit method (ADI). The \textsc{StEfCal} algorithm reduces the computational complexity of calibration from $O(N^3)$ to $O(N^2)$.
+to $\rho=\frac{1}{2}$) we obtain the even parameter update step of \textsc{StEfCal}\footnote{$k\in\{1,3,\cdots\}$} \citep[\textsc{StEfCal},][]{Mitchell:MWA-cal,Salvini2014}.
+In the \textsc{StEfCal} algorithm, the measurement equation (equation~\ref{eq:vis_definition}) is linearized by assuming that the gains are known, but that their conjugates are not.
+Under this assumption, the system of equations become linear and the conjugates of the gains can be obtained in a straightforward manner.
+Starting from the latest value of the gain conjugates, an updated estimate of the gains can be obtained iteratively until convergence is reached.
+Alternating between solving and fixing different sets of parameters (which is exactly what \textsc{StEfCal} does) is referred to 
+as the alternating direction implicit (ADI) method. The \textsc{StEfCal} algorithm reduces the computational complexity of calibration from $O(N^3)$ to $O(N^2)$.
 
@@
 \label{sec:red_wirtinger}
-Interferometric baselines are named redundant when they sample the exact same visibilities in the $uv$-plane, i.e.
-if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$.
+Interferometric baselines are redundant when they sample the exact same visibilities in the $uv$-plane, i.e. if baseline $pq$ and $rs$ are redundant then $y_{pq} = y_{rs}$. 
+A redundant array layout allows us to solve for the unknown observed visibilities themselves in
+addition to the antenna gains (see equation \ref{eq:vec_linear_definitions}).
+This is true, since in the case of a redundant layout equation \ref{eq:vec_linear_definitions} is already an overdetermined
+system (there is no need to first predict visibilities from a pre-existing skymodel).
+
+{\bf SJK: I feel there needs to be some additional emphasis placed on the fact that the model visibilities are a solvable parameter.
+I found a great deal of emphasis is placed on the redundancy, but not why that redundancy is so crucial to the solution procedure.}
+{\bf TLG: Added paragraph, above, GB want to phrase it better??}
+
+%When we employ a redundant array, we dramatically reduce the number 
+%of unknowns (assuming we have no prior knowledge of the sky). If our array is redundant enough we reduce
+%the number of unknowns to such an extent that we are able
+%to solve for  Redundant calibration is the
+%procedure by which we estimate both the unknown observed
+%visibilities and the antenna gains.
+
 It is convenient to group redundant visibilities together and label each group using a single index rather than using their antenna pairs as in equation~\ref{eq:vis_definition}.
-We introduce a function $\phi_{pq}$ that maps the antenna pair associated with a specific baseline to the single group index of the redundant baseline group that includes the baseline in question, i.e.
-if baseline $pq$ and $rs$ are redundant then $\phi_{pq} = \phi_{rs}$.
+We introduce a function $\phi$ that maps the antenna pair associated with a specific baseline to its corresponding redundant baseline group, i.e.
+if baseline $pq$ and $rs$ are redundant then $\phi_{pq} = \phi_{rs}$ (implying they belong to the same group).
 The function $\phi_{pq}$ is not symmetric as $\phi_{pq} = 0$ if $p \geq q$.
@@
 We denote the range of $\zeta_{pq}$ with $\mathcal{R}(\zeta_{pq})$.
-The maximal element that $\zeta_{pq}$ can ascertain is denoted by $L$ and can be interpreted as the maximal number of unique redundant baseline groups which can be formed for a given array layout.
+The maximal element that $\zeta_{pq}$ can ascertain is denoted by $L$ and can be interpreted as the greatest number of unique redundant baseline groups which can be formed for a given array layout.
 
@@
 {\includegraphics[width=0.33\textwidth]{./REG_phi.pdf}\label{fig:REG_phi}}
-\caption{Three different redundant antenna layouts: hexagonal (top left), square (top middle) and regular east-west (top right) with their respective symmetric redundancy geometry functions $\zeta_{pq}$ (bottom panels).
+\caption{Three different redundant antenna layouts: hexagonal (top left), square (top middle) and regular east-west (top right) with their associated symmetric redundancy geometry functions $\zeta_{pq}$ (bottom panels).
 We used 91 antennas to construct the hexagonal layout, while 100 antennas were used in the square and east-west layouts.
@@
 the hexagonal, square and east-west layouts are 165, 180 and 99 respectively.
-The analytic expressions of $L$ for an hexagonal, square and east-west layout is $L = 2N-\frac{1}{2}\sqrt{12N-3}-\frac{1}{2}$,
-$L=2N-2\sqrt{N}$ and $L=N-1$.\label{fig:geometry_function}}
+The analytic expressions of $L$ for an hexagonal, square and east-west layout are $L = 2N-\frac{1}{2}\sqrt{12N-3}-\frac{1}{2}$,
+$L=2N-2\sqrt{N}$ and $L=N-1$ respectively.\label{fig:geometry_function}}
 \end{figure*}
@@
 The analytic expressions for $\bJ$, $\bH$ and $\bJ^H\breve{\br}$ can be found in Appendix~\ref{sec:analytic}.
-Appendix~\ref{sec:analytic} also contain two useful identities, involving $\bJ$ and $\bH$.
+Appendix~\ref{sec:analytic} also contains two useful identities involving $\bJ$ and $\bH$.
 
@@
 
-Figure~\ref{fig:hessian} shows that the Hessian matrix $\bH$ is nearly diagonal and sparse for both a regular and hexagonal layout considered.
+Figure~\ref{fig:hessian} shows that the Hessian matrix $\bH$ is nearly diagonal and sparse for both the regular east-west and hexagonal layouts we considered.
 We therefore follow the approach of \citet{Smirnov2015} and approximate the Hessian matrix $\bH$ with its diagonal.
@@
 \end{equation}
-We obtain a similar result if we repeat the above procedure, only in this case we use equation~\ref{eq:LM_update} instead. Thus,
+Using equation equation~\ref{eq:LM_update}, we follow the same procedure and obtain a similar result for LM
 \begin{align}
 \breve{\bz}_{k+1} &\approx \frac{1}{1+\lambda}\widetilde{\bH}^{-1}\bJ^H\breve{\bd} + \frac{\lambda}{1+\lambda} \breve{\bz}_k,\label{eq:lambda}\\
- &= \alpha \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\alpha)\breve{\bz}_k. \label{eq:alpha}  
+ &= \rho \widetilde{\bH}^{-1}\bJ^H\breve{\bd} + (1-\rho)\breve{\bz}_k. \label{eq:alpha}  
 \end{align}
@@
 \label{eq:g_update}
-g_{i}^{k+1} = \alpha \frac{\sum_{j\neq i} g_j^k \widetilde{y}_{ij}^{~\!\!k} d_{ij}}{\sum_{j\neq i} |g_j^k|^2|y_{\zeta_{ij}}^k|^2} + (1-\alpha) g_i^k, 
+g_{i}^{k+1} = \rho \frac{\sum_{j\neq i} g_j^k \widetilde{y}_{ij}^{~\!\!k} d_{ij}}{\sum_{j\neq i} |g_j^k|^2|y_{\zeta_{ij}}^k|^2} + (1-\rho) g_i^k, 
 \end{equation}
@@
 \label{eq:y_update}
-y_{i}^{k+1} = \alpha \frac{\sum_{rs \in \mathcal{RS}_i} \conj{g}_r^k g_s^k d_{rs}}{\sum_{rs \in \mathcal{RS}_i}|g_r^k|^2|g_s^k|^2} + (1-\alpha) y_i^k. 
-\end{equation}
-The the index set $\mathcal{RS}_i$ and the quantity $\widetilde{y}_{ij}$ is defined in equation~\ref{eq:RS} and equation~\ref{eq:y_tilde}, respectively. 
+y_{i}^{k+1} = \rho \frac{\sum_{rs \in \mathcal{RS}_i} \conj{g}_r^k g_s^k d_{rs}}{\sum_{rs \in \mathcal{RS}_i}|g_r^k|^2|g_s^k|^2} + (1-\rho) y_i^k. 
+\end{equation}
+The the index set $\mathcal{RS}_i$ and the quantity $\widetilde{y}_{ij}$ are defined in equation~\ref{eq:RS} and equation~\ref{eq:y_tilde}, respectively. 
 The computational complexity of inverting $\widetilde{\bH}$ is $O(P)$. We note that equation~\ref{eq:g_update} is the gain estimator associated with \textsc{StEfCal}.
@@
 In Appendix~\ref{sec:red_stef_ADI} we show that equation~\ref{eq:g_update} and~\ref{eq:y_update} can also be derived using the ADI method.
-For this reason, we refer to the code implementation of the pseudo LM calibration scheme, derived above (equation~\ref{eq:g_update} and~\ref{eq:y_update}), as redundant \textsc{StEfCal}, throughout the rest of the paper.
-
-The choice of the $\alpha$ parameter is somewhat unconstrained. In this paper we chose $\alpha$ by adopting the same strategy that is used by \textsc{StEfCal} and \citet{Marthi2014},
-i.e. we chose $\alpha$ to be equal to a $\frac{1}{3}$ ($\lambda = 2$). We also carried out simulations to validate this choice.
-%In the case of \textsc{StEfCal}, $\alpha = \frac{1}{2}$ (for even iterations), which results in the last term of equation~\ref{eq:one_half} and equation~\ref{eq:stef_alpha} being equal.
-When $\alpha=\frac{1}{3}$, the last term of equation~\ref{eq:two_thrids} and equation~\ref{eq:alpha} are equal.
-
-We generated a sky model that comprises of 100 flat spectrum sources distributed over a 3$^{\circ}$ by 3$^{\circ}$ sky 
-patch centred. The flux density of each source was drawn from a power law distribution with a slope of 2 and the source position drawn from a uniform distribution. 
+For this reason, we refer to the implementation of the pseudo-LM calibration scheme derived above equation~\ref{eq:g_update} and~\ref{eq:y_update} as redundant \textsc{StEfCal} throughout the rest of the paper.
+
+The choice of the $\rho$ parameter is somewhat unconstrained. In this paper we chose $\rho$ by adopting the same strategy that is used by \textsc{StEfCal} and \citet{Marthi2014},
+i.e. we chose $\rho$ to be equal to a $\frac{1}{3}$ ($\lambda = 2$). We also carried out simulations to validate this choice.
+%In the case of \textsc{StEfCal}, $\rho = \frac{1}{2}$ (for even iterations), which results in the last term of equation~\ref{eq:one_half} and equation~\ref{eq:stef_alpha} being equal.
+When $\rho=\frac{1}{3}$, the last term of equation~\ref{eq:two_thrids} and equation~\ref{eq:alpha} are equal.
+
+We generated a sky model that comprised of 100 flat spectrum sources distributed over a 3$^{\circ}$ by 3$^{\circ}$ sky 
+patch. The flux density of each source was drawn from a power law distribution with a slope of 2 and the source position were drawn from a uniform distribution. 
 We also made use of multiple fictitious telescope layouts each one having a hexagonal geometry (see the left upper image of Figure~\ref{fig:geometry_function} for an example layout).
@@
 
-We "corrupted" visibilities by applying gain errors (Table~\ref{tab:gain_parm}) and calibrated the corrupted visibilities using redundant \textsc{StEfCal}.
+We corrupted visibilities by applying gain errors (Table~\ref{tab:gain_parm}) and calibrated the corrupted visibilities using redundant \textsc{StEfCal}.
 Solutions were independently derived for each time-step and channel for five realizations.
@@
 \caption{The gain error models used in this paper. We have used the symbol $x$ here as a proxy as it can either refer to time-slots or channels. We either
-performed our simulations over multiple time-slots and one frequency channel or one timeslot and multiple frequencey channels (see Table~\ref{tab:ch_parm}).
+performed our simulations over multiple time-slots and one frequency channel or one timeslot and multiple frequency channels (see Table~\ref{tab:ch_parm}).
 Moreover, $c$ in the left most column denotes the speed of light.
 We adopted a sinusoidal error model as well as phase slope across frequency that mimics a real case of physical delays between different antennas.
-It is worth pointing out the following. In the first model, we model a gain error with amplitude around one and an additional phase error. In 
+In the first model, we model a gain error with amplitude around one and an additional phase error. In 
 the second model we chose $A$ in such a way that we do not unintentionally produce a zero gain value. For the third model 
 the value of $\tau$ was chosen so that the amount of phase wraps that occur across the observing band is restricted to a realistic number.
-{\bf (GB: we need a final better description of the models, but we can work this out later).}}
+{\bf (GB: we need a final better description of the models, but we can work this out later).} {\bf SJK: Agree with Gianni here - this caption needs a rework.
+The descriptions of the models are a bit clumsy.}}
 \begin{tabular}{|c c c c|} 
@@
 
-The simulated visibilities included a noise term. The purpose of which is to produce a certain SNR in the simulated
+The simulated visibilities included a noise term, the purpose of which was to produce a certain SNR in the simulated
 visibilities. We made use of the following definition of \citep[SNR, ][]{Liu2010,Marthi2014}:  
@@
 
-Figure~\ref{fig:prec_error} show the simulation results as a function of SNR and number of antennas. The accuracy of our solutions is quantified through the percentage error
+Figure~\ref{fig:prec_error} shows the simulation results as a function of SNR and number of antennas. The accuracy of our solutions is quantified through the percentage error
 \begin{equation}
@@
 \includegraphics[width=0.47\textwidth]{./prec_error.pdf} 
-\caption{We plot the percentage error $\beta$ between the simulated visibilitied and the visibilities solved for by redundant \textsc{StEfCal} for different SNR values as a function of the number of antennas in the array $N$.}
+\caption{We plot the percentage error $\beta$ between the simulated visibilities and the visibilities solved for by redundant \textsc{StEfCal} for different SNR values as a function of the number of antennas ($N$) in the array.}
 \label{fig:prec_error}
@@
 the PCG method).
-Interestingly, emperical tests suggests that the unmodified Hessian itself is singular.
-It is therefore important to mention here, that the CG method
+Interestingly, empirical tests suggest that the unmodified Hessian itself is singular.
+It is therefore important to mention, that the CG method
 can pseudo invert the unmodified Hessian as well, i.e.
-the CG method can be directly applied to equation~\ref{eq:GN_update}, since the Hessian is a positive semi-definite Hermition matrix and the vector $\bJ^H\breve{\br}$ is an element of its column range \citep{Lu2015}.
+the CG method can be directly applied to equation~\ref{eq:GN_update}, because the Hessian is a positive semi-definite Hermitian matrix and the vector $\bJ^H\breve{\br}$ is an element of its column range \citep{Lu2015}.
 
@@
 
-The computational complexity of CG is  
+The computational complexity of the CG method is  
 \begin{equation}
@@
 \end{equation}
-where $m$ denotes the number of non-zero entries and $\kappa$ denotes the spectral 
+where $m$ denotes the number of non-zero entries in and $\kappa$ denotes the spectral 
 condition number of the matrix which is to be inverted. The spectral condition number $\kappa$ of the matrix $\bA$ is defined as:
@@
 \end{equation}
-where $\iota_{\textrm{max}}$ and $\iota_{\textrm{min}}$ respectively denote the largest and the smallest eigenvalue of $\bA$.
+where $\iota_{\textrm{max}}$ and $\iota_{\textrm{min}}$ denote the largest and the smallest eigenvalue of $\bA$ respectively.
 
@@
 We used the complex LM-algorithm described in Section~\ref{sec:red_wirtinger} to calibrate the corrupted visibilities. To invert the modified
-Hessian we used the CG method, with and without a preconditioner. We made use of the Jacobian preconditioner. 
+Hessian we used the CG method, with and without a Jacobian preconditioner. 
 Figure~\ref{fig:kappa} shows us that preconditioning reduces the condition number of $\bmH$ to a small constant value and therefore effectively eliminates
@@
 gradient method to invert $\bmH$ as a function of the number of antennas $N$ in the array, before (magenta and red curves) and after preconditioning (blue and green curves), at different SNR values.
-Both plots show that the Jacobian preconditioner definitely speed up the conjugate gradient method.
+Both plots show that the Jacobian preconditioner definitely speeds up the conjugate gradient method.
 \label{fig:kappa_itr}}
@@
 Let us now shift our attention towards the remaining factor $m$.
-To get a usefull idea of the computational complexity of CG, we must relate $m$ and $P$.
+To get a useful idea of the computational complexity of CG, we must relate $m$ and $P$.
 This can be achieved by utilizing the modified Hessian's measure of sparsity $\gamma$:
@@
 \end{align}
+
+{\bf SJK: What does this mean?
+I feel the notation of this aymptotic quantity needs a little more explanation.
+I can be inferred from the maths, but I think a brief qualitative description would be good.}
+{\bf TLG: Started to add an in depth description. I find the description actually adding confusion and thus removed it. See Latex for added paragraph, just below comment.}
+%The sparsity of a matrix is defined as the ratio
+%obtained by dividing the zero-entries of a matrix by the total amount of entries in the matrix. The asymptotic sparsity 
+%of a class of matrix is obtained by computing the sparsity of this class of matrices as their dimension tends to infinity.
+
 For more complicated array layouts, however, there is no straightforward analytical solution and we empirically determined the sparsity ratios for three different geometric layouts as a function of the number antennas in the array (see Figure~\ref{fig:gamma}).
@@
 \end{align}
-The computational complexity is therefore asymptotically bounded by $O(P^2)$, although it converges very slowly to its asymptotic value, and in general is equal to $O(P^{c})$, where $c \sim 1.7$ in the case of an hexagonal geometric layout with $N < 200$ (see Figure~\ref{fig:c}).
+The computational complexity is therefore asymptotically bounded by $O(P^2)$, although it converges very slowly to its asymptotic value, and in general is equal to $O(P^{c})$ (with $c<2$).
+In the case of an hexagonal geometric layout with $N < 200$ we have that $c \sim 1.7$ (see Figure~\ref{fig:c}).
 
@@
 Redundant \textsc{StEfCal} is computationally inexpensive as it just needs to invert a diagonal matrix, however, the PCG inversion is accurate and, therefore, may require fewer iterations, ultimately leading to a faster convergence.
-We computed the theoretical number of redundant \textsc{StEfCal} iterations $\Delta k$, in excess of the number of LM (implementing PCG) iterations required, for LM to outperform redundant \textsc{StEfCal}
+We computed the theoretical number of redundant \textsc{StEfCal} iterations $\Delta k$ in excess of the number of LM (implementing PCG) iterations required for LM to outperform redundant \textsc{StEfCal}
 \begin{align}
 \label{eq:k} 
- (k+\Delta k)P >& \, kP^c, & \Delta k >& \, k(P^{c-1}-1),
-\end{align}
-and compared it to the empirically obtained average excess number of iterations.
+ (k+\Delta k)P >& \, kP^c, & \Delta k >& \, k(P^{c-1}-1).
+\end{align}
+We then compared $\Delta k$ with the empirically obtained average excess number of iterations.
 The results are displayed in Figure~\ref{fig:out_diff} which shows that redundant \textsc{StEfCal} outperforms the PCG method.
@@
 We note that, in its current implementation, redundant \textsc{StEfCal} does not solve for the degeneracies inherent to redundant calibration  \citep{Zheng2014,Kurien2016} which will be the subject of future work.
-Compared to current redundant calibration algorithms, redundant \textsc{StEfCal} is more robust to inital conditions and allows for an easier, massive parallelization.
-
-We investigated the computational cost of redundant \textsc{StEfCal} and compared it with the performance of the PCG method (approach suggested by \cite{Liu2010}).
-We found that, although the PCG method method greatly improves the speed of redundant calibration, it still significant underperforms in comparison to redundant \textsc{StEfCal}.
+Compared to current redundant calibration algorithms, redundant \textsc{StEfCal} is more robust to inital conditions and allows for easier parallelization.
+
+We investigated the computational cost of redundant \textsc{StEfCal} and compared it with the performance of the PCG method (as suggested by \cite{Liu2010}).
+We found that, although the PCG method method greatly improves the speed of redundant calibration, it still significantly underperforms when compared to redundant \textsc{StEfCal}.
 
@@
 \end{equation}
-Substituting equation~\ref{eq:RHR} and equation~\ref{eq:RHd} into equation~\ref{eq:y_final} and simplifying the result leads to the leftmost term in equation~\ref{eq:y_update} (if we bar $\alpha$ and consider
+Substituting equation~\ref{eq:RHR} and equation~\ref{eq:RHd} into equation~\ref{eq:y_final} and simplifying the result leads to the leftmost term in equation~\ref{eq:y_update} (if we bar $\rho$ and consider
 only the $i$-th entry of $\by$).

### 2017-12-14 paper.tex
final edits from Trienko

@@
 ''redundant calibration" - and boosts sensitivity on specific spatial scales. In this paper, we formulate 
-calibration of redundant interferometric arrays as a complex optimization problem {\bf SJK: suggests removing citation} \citep[i.e.,][]{Smirnov2015}. 
+calibration of redundant interferometric arrays as a complex optimization problem. 
 We solve this optimization problem via the Levenberg-Marquardt algorithm. This calibration approach is more 
 robust to initial conditions than current algorithms and, by leveraging an approximate matrix inversion, allows for further optimization and an efficient implementation (``redundant \textsc{StEfCal}").
-{\bf (SJK: ...boasts competitive [big O notation here] scaling.
-We coin the term "redundant StEfCal" for our algorithm and implementation)}.
 We also investigated using the preconditioned conjugate gradient method as an alternative to the approximate matrix inverse, but found that its computational performance is not competitive with respect to ``redundant \textsc{StEfCal}".
@@
 Moreover, redundant configurations provide a sensitivity boost on a selected sample of spatial scales that can be tuned to be the highest signal--to--noise ratio (SNR) EoR modes \citep[][]{Parsons2012,Dillon2016}.
-These reasons motivated the design and deploment of EoR arrays in redundant configurations like the MIT-EoR \citep{Zheng2014}, the Precision Array to Probe the Epoch of Reionization \citep[PAPER,][]{Ali2015} and the Hydrogen Epoch of Reionization Array \citep[HERA,][]{deboer2017}.
+These reasons motivated the design and deployment of EoR arrays in redundant configurations like the MIT-EoR \citep{Zheng2014}, the Precision Array to Probe the Epoch of Reionization \citep[PAPER,][]{Ali2015} and the Hydrogen Epoch of Reionization Array \citep[HERA,][]{deboer2017}.
 
@@
 We investigate the speed-up that the preconditioned conjugate gradient (PCG) method provides if it is employed by the new algorithm \citep{Liu2010}.
-A comparison between the computational complexity of the optimized new algorithm and redundant StEfCal is also performed.
+A comparison between the computational complexity of the optimized new algorithm and redundant \textsc{StEfCal} is also performed.
 
@@
 
-{\bf JSK: I would introduce the alpha quantity on its own before attaching it as an index in the definition of d.
-The reason for this is that eq.
-4 and 5 seem very complicated without first understanding the meaning of alpha.
-Could have a paragraph starting ``In order to express equation 1 in vector form, we first introduce $\alpha$ which maps...''}
-
-{\bf TLG: Was like this earlier, got changed during the edits. GB let me know if it should be changed back.}
-
 Radio interferometric calibration aims to determine the best estimate of $\bg = [g_1,g_2,\cdots,g_N]^T$ in order to correct the data and, following equation~\ref{eq:vis_linear_definition}, can be formulated as a non linear least-squares optimization problem:
@@
 In order to circumvent the differentiability conundrum associated with complex least squares problems, standard interferometric calibration divides the complex optimization problem into its real and imaginary parts and solves for the real and imaginary parts of the unknown model parameters separately.
-\citet{Smirnov2015} showed, however, that this approach is not needed if complex calculus \citep{Wirtinger1927} is directly adopted {\bf JSK: employed}.
+\citet{Smirnov2015} showed, however, that this approach is not needed if complex calculus \citep{Wirtinger1927} is adopted.
 The Wirtinger derivatives are defined as:
@@
 This is true, since in the case of a redundant layout equation \ref{eq:vec_linear_definitions} is already an overdetermined
-system (there is no need to first predict visibilities from a pre-existing skymodel).
-
-{\bf SJK: I feel there needs to be some additional emphasis placed on the fact that the model visibilities are a solvable parameter.
-I found a great deal of emphasis is placed on the redundancy, but not why that redundancy is so crucial to the solution procedure.}
-{\bf TLG: Added paragraph, above, GB want to phrase it better??}
+system even before having predicted visibilities from a pre-existing skymodel.
 
@@
 \end{equation}
-Using equation equation~\ref{eq:LM_update}, we follow the same procedure and obtain a similar result for LM
+Using equation~\ref{eq:LM_update}, we follow the same procedure and obtain a similar result for LM
 \begin{align}
@@
 We note that their derivation is less general. The LM algorithm has better convergence properties than gradient descent and 
-encompasses the gradient descent algorithm as a special case.
-In Appendix~\ref{sec:red_stef_ADI} we show that equation~\ref{eq:g_update} and~\ref{eq:y_update} can also be derived using the ADI method.
-For this reason, we refer to the implementation of the pseudo-LM calibration scheme derived above equation~\ref{eq:g_update} and~\ref{eq:y_update} as redundant \textsc{StEfCal} throughout the rest of the paper.
+encompasses the gradient descent algorithm as a special case. In Appendix~\ref{sec:red_stef_ADI} we show that equation~\ref{eq:g_update} and~\ref{eq:y_update} can also be derived using the ADI method.
+For this reason, we refer to the implementation of the pseudo-LM calibration scheme derived above, i.e.
+equation~\ref{eq:g_update} and~\ref{eq:y_update}, as redundant \textsc{StEfCal} throughout the rest of the paper.
+Interestingly, \citet{Marthi2014} were not the first to make use of the ADI method to perform redundant calibration, a slower alternative ADI based calibration algorithm is presented in \citet{Wijnholds2012}.
 
@@
 We generated a sky model that comprised of 100 flat spectrum sources distributed over a 3$^{\circ}$ by 3$^{\circ}$ sky 
-patch. The flux density of each source was drawn from a power law distribution with a slope of 2 and the source position were drawn from a uniform distribution. 
+patch. The flux density of each source was drawn from a power law distribution with a slope of 2 and the source position was drawn from a uniform distribution. 
 We also made use of multiple fictitious telescope layouts each one having a hexagonal geometry (see the left upper image of Figure~\ref{fig:geometry_function} for an example layout).
-The largest hexagonal array that we used has 217 antennas, with a minimum and maximum baseline of of 20~m and 320~m respectively.
+The largest hexagonal array that we used has 217 antennas, with a minimum and maximum baseline of 20~m and 320~m respectively.
 
@@
 Moreover, $c$ in the left most column denotes the speed of light.
-We adopted a sinusoidal error model as well as phase slope across frequency that mimics a real case of physical delays between different antennas.
+We adopted a sinusoidal error model, similar to the sinusoidal error models used within \textsc{MeqTrees} \citep{Noordam2010}, as well as a phase slope across frequency which mimics a real case of physical delays between different antennas.
 In the first model, we model a gain error with amplitude around one and an additional phase error. In 
-the second model we chose $A$ in such a way that we do not unintentionally produce a zero gain value. For the third model 
-the value of $\tau$ was chosen so that the amount of phase wraps that occur across the observing band is restricted to a realistic number.
-{\bf (GB: we need a final better description of the models, but we can work this out later).} {\bf SJK: Agree with Gianni here - this caption needs a rework.
-The descriptions of the models are a bit clumsy.}}
+the second model we chose $A$ in such a way that we do not unintentionally produce nonsensical gain values, i.e. a zero gain value. For the third model 
+the value of $\tau$ was chosen so that the amount of phase wraps that occur across the observing band is restricted to a realistic number.}
 \begin{tabular}{|c c c c|} 
@@
 
-We generated simulated (i.e. corrupted) visibilities according to models described in Table~\ref{tab:ch_parm} {\bf (GB: we need some more explanation of the simulations... next round)}. 
+We generated simulated (i.e. corrupted) visibilities according to models described in Table~\ref{tab:gain_parm}. 
 %
@@
 \end{align}
-
-{\bf SJK: What does this mean?
-I feel the notation of this aymptotic quantity needs a little more explanation.
-I can be inferred from the maths, but I think a brief qualitative description would be good.}
-{\bf TLG: Started to add an in depth description. I find the description actually adding confusion and thus removed it. See Latex for added paragraph, just below comment.}
-%The sparsity of a matrix is defined as the ratio
 %obtained by dividing the zero-entries of a matrix by the total amount of entries in the matrix. The asymptotic sparsity 
@@
 We investigated the computational cost of redundant \textsc{StEfCal} and compared it with the performance of the PCG method (as suggested by \cite{Liu2010}).
-We found that, although the PCG method method greatly improves the speed of redundant calibration, it still significantly underperforms when compared to redundant \textsc{StEfCal}.
+We found that, although the PCG method greatly improves the speed of redundant calibration, it still significantly underperforms when compared to redundant \textsc{StEfCal}.
 
@@
 \section*{Acknowledgements}
+This work is based upon research supported by the South
+African Research Chairs Initiative of the Department of Sci-
+ence and Technology and National Research Foundation.
+
 G.B.
@@
 
-{\bf (GB: We need to bring back the logcal and lincal derivation using the geometric function.)}
+%{\bf (GB: We need to bring back the logcal and lincal derivation using the geometric function.)}
 %{\bf (TLG: I have removed logcal and lincal as well as the PCG section.

### 2017-12-15 paper.tex
Gianni's last comments

@@
 %$^{6}$Radio Astronomy Lab., U. California, Berkeley CA 94720, USA}
-\author[T.L.~Grobler et al.]{T.L.~Grobler$^{1,2}$\thanks{E-mail: email@address (AVR)}, G.~Bernardi$^{2,3}$, C.L.~Carilli$^{4,5}$, J.S.~Kenyon$^{2}$, A.R.~Parsons$^{6,7}$?
-\newauthor and O.M.~Smirnov$^{2,3}$\\
+\author[T.L.~Grobler et al.]{T.L.~Grobler$^{1,2}$\thanks{E-mail:tlgrobler@sun.ac.za}, G.~Bernardi$^{2,3,4}$, J.S.~Kenyon$^{2}$, A.R.~Parsons$^{5,6}$ and O.M.~Smirnov$^{2,3}$\\
 $^{1}$Dept of Mathematical Sciences, Computer Science Division, Stellenbosch University, Private Bag X1, 7602 Matieland, South Africa\\
@@
 $^{3}$SKA SA, 3rd Floor, The Park, Park Road, Pinelands, 7405, South Africa\\
-$^{4}$National Radio Astronomy Obs., Socorro NM\\
-$^{5}$Cavendish Lab., Cambridge UK\\
-$^{6}$Dept. of Astronomy, University of California, Berkeley, CA 94720, USA\\
-$^{7}$Radio Astronomy Lab., U. California, Berkeley CA 94720, USA}
+$^{4}$INAF-Istituto di Radioastronomia, via Gobetti 101, 40129, Bologna, Italy\\
+$^{5}$Dept. of Astronomy, University of California, Berkeley, CA 94720, USA\\
+$^{6}$Radio Astronomy Lab., U. California, Berkeley CA 94720, USA}
 \begin{document}

### 2017-12-18 paper.tex
final edit from Gianni

@@
 \section*{Acknowledgements}
-This work is based upon research supported by the South
-African Research Chairs Initiative of the Department of Sci-
-ence and Technology and National Research Foundation.
-
-G.B.
-acknowledges support from the Royal Society and the Newton Fund under grant NA150184.
+This work is based upon research supported by the South African Research Chairs Initiative of the Department of Science and Technology and National Research Foundation.
+GB acknowledges support from the Royal Society and the Newton Fund under grant NA150184.
 This work is based on the research supported in part by the National Research Foundation of South Africa (grant No.
 103424).
-This research is supported by the South African Research Chairs Initiative of the Department of Science and Technology and National Research Foundation,
 
@@
 The above identities can be trivially established by mechanically showing that the left hand side of each expression in equation~\ref{eq:identities} is equal to its right hand side.
-In equation~\ref{eq:identities}.
 

### 2018-02-01 paper.tex
newest version of redcal paper

@@
 %\title[Complex redundant calibration]{Redundant calibration as a complex optimization problem}
-\title[Redundant interferometric calibration as a complex optimization problem]{Redundant calibration as a complex optimization problem}
+\title[Redundant calibration and Complex Optimization]{Redundant interferometric calibration as a complex optimization problem}
 %\author[T.L.~Grobler et al.]{T.L.~Grobler$^{1}$\thanks{E-mail: email@address (AVR)}, G.~Bernardi$^{1,2}$, Z.S.~Ali$^{3}$?, C.L.~Carilli$^{4,5}$?, J.S.~Dillon$^{3}$?, A.~Liu$^{3}$?, \newauthor
@@
 As miscalibrated foreground emission leads to artifacts that can jeopardize the EoR signal \citep[e.g.,][]{grobler2014,barry2016,ewall-wice2016}, an exceptional interferometric calibration is required.
-Such calibration needs an accurate knowledge of both the sky emission and the instrumental response \citep[e.g.,][]{Smirnov2011c}, unless arrays are deployed in a redundant configuration, i.e.
-with receiving elements placed on regular grids so that multiple baselines measure the same sky brightness emission.
-Leveraging this property, redundancy is a promising path to achieve highly accurate interferometric calibration \citep[][]{Noordam1982,Wieringa1992,Pearson1984,Liu2010,Noorishad2012,Marthi2014,Sievers2017}.
-Moreover, redundant configurations provide a sensitivity boost on a selected sample of spatial scales that can be tuned to be the highest signal--to--noise ratio (SNR) EoR modes \citep[][]{Parsons2012,Dillon2016}.
+Such calibration needs an accurate knowledge of both the sky emission and the instrumental response \citep[e.g.,][]{Smirnov2011c}.
+A skymodel is required for standard calibration as it would be an ill-posed problem if one tries to directly solve for all of the unknowns, i.e. the antenna gains and the uncorrupted 
+visibilities, without first attempting to simplify the calibration problem. When the uncorrupted visibilities, however, are predicted from a predefined skymodel the problem becomes solvable.
+In contrast, if an array is deployed in a redundant configuration, i.e. with receiving elements placed on regular grids so that multiple baselines measure the same sky brightness emission,
+we circumvent the need for a skymodel altogether. This is true, since redundancy leads to fewer unknowns (i.e. number of unique 
+$uv$-modes) and if the array is redundant enough it reduces the number of unknowns to such an extend that the calibration problem becomes solvable without having to predict the uncorrupted visibilities from a skymodel.
+Redundancy, therefore, is a promising path to achieve highly accurate interferometric calibration \citep[][]{Noordam1982,Wieringa1992,Pearson1984,Liu2010,Noorishad2012,Marthi2014,Sievers2017}.
+Moreover, redundant configurations provide a sensitivity boost on particular spatial scales that can be tuned to be the highest signal--to--noise ratio (SNR) EoR modes \citep[][]{Parsons2012,Dillon2016}.
 These reasons motivated the design and deployment of EoR arrays in redundant configurations like the MIT-EoR \citep{Zheng2014}, the Precision Array to Probe the Epoch of Reionization \citep[PAPER,][]{Ali2015} and the Hydrogen Epoch of Reionization Array \citep[HERA,][]{deboer2017}.
@@
 
+A summary of the notation that is used within the paper is presented in 
+Table~\ref{tab:glossary}. We also discuss the overset notation used within the paper in a bit more detail below:
+\begin{enumerate}
+ \item $\widetilde{x}$ -- This notation is used to denote a new scalar value which was derived from the scalar $x$ using a proper mathematical definition.  
+ \item $\conj{x}$ -- This notation denotes the conjugation of its operand (the conjugation of $x$). For the readers convenience we will redefine this operator when it is
+ first used witin the paper.
+ \item $\widehat{x}$ -- This notation denotes that the quantity is an estimated value.
+ \item $\breve{\mathbf{x}}$ -- This notation denotes an augmented vector, i.e. $\breve{\mathbf{x}} = [\mathbf{x}^T,\conj{\mathbf{x}}^T]^T$. 
+\end{enumerate}
+
 The paper is organized as follows: we review the complex calibration formalism by \citet{Smirnov2015} in Section~\ref{sec:sky_wirtinger}, we extend it to the redundant case in Section~\ref{sec:red_wirtinger}, we present some computational complexity results in Section~\ref{sec:pcg} and we presnt our conclusions in Section~\ref{sec:conclusions}.
@@
 
+\begin{table}
+\centering
+\caption{Notation and frequently used symbols.}
+\begin{tabular}{|l l|} 
+\hline
+$x$, $\mathbf{x}$ and $\mathbf{X}$ & scalar, vector and matrix\\
+$\mathbb{N}$ & set of natural numbers\\
+$\mathbb{C}$ & set of complex numbers\\
+$\bI$ & Identity matrix\\
+$\alpha_{pq}$, $\phi_{pq}$, $\zeta_{pq}$, $\xi_{pq}$ and & indexing functions\\
+$\psi_{pq}:\mathbb{N}^2\rightarrow\mathbb{N}$ &\\
+$\conj{(*)}$ & conjugation of operand\\
+$(*)^{2}$ & square operand\\
+$\|*\|_F$ & Frobenius norm\\
+$(*)^{-1}$ & Matrix inversion\\
+$\odot$ & Hadamard product\\
+&(element-wise product)\\
+$(*)^H$ & Hermitian transpose\\
+$(*)^T$ & Matrix transpose\\
+$<*>_x$ & averaging in x\\
+$\frac{\partial}{\partial z}$ & Wirtinger derivative\\
+$\widetilde{x}$ & a new quantity derived from $x$\\
+$\widehat{x}$ & estimated quantity\\
+$\breve{\mathbf{x}}$ & augmented vector $[\mathbf{x}^T,\conj{\mathbf{x}}^T]^T$\\
+$i$ & $\sqrt{-1}$\\
+$N$,$B$,$L$ and $P$ & number of antennas, baselines,\\
+&redundant groups and parameters\\
+$\bd$, $\bv$, $\bg$, $\by$, $\br$ and $\bn$& data, predicted, antenna,\\
+&true visibility, residual\\
+&and noise vector.\\ 
+$\bz$ & $\bz = [\bg^T,\by^T]^T$\\
+$\Lambda$ & objective function\\
+$\lambda$ and $\rho$ & damping factor and $\rho = \frac{1}{1+\lambda}$\\
+$\bJ$ & Jacobian matrix\\
+$\bH$ & $\bJ^H\bJ$ (Hessian matrix)\\
+$\bmH$ & modified Hessian matrix\\
+$[\mathbf{X}]_{ij}$& element $ij$ of matrix $\mathbf{X}$\\
+$x_{ij}$ & composite antenna index\\
+$x_i$ & antenna or redundant group index\\
+$x^k$ & iteration number\\
+$\bM$ & preconditioner matrix\\
+$m$ & number of non-zero entries\\
+&in a matrix\\
+$\kappa$ & spectral condition number\\
+&of a matrix\\
+$\gamma$ & sparsity factor of a matrix\\
+$\Delta x$ & parameter update\\
+\hline
+\end{tabular}
+\label{tab:glossary}
+\end{table}
+
+\begin{table}
+\centering
+\caption{The dimensions of the Jacobian matrices, and their respective sub-matrices, defined in Section~\ref{sec:sky_wirtinger} and Section~\ref{sec:red_wirtinger}.}
+\begin{tabular}{|c c|} 
+\hline
+Matrix & Dimension\\
+\hline
+\hline
+\multicolumn{2}{c}{Wirtinger Calibration}\\
+\hline
+$\bJ$ & $2B \times 2N$ \\
+$\bmJ_1$ & $B \times N$\\
+$\bmJ_2$ & $B \times N$\\
+\hline
+\hline
+\multicolumn{2}{c}{Redundant Wirtinger Calibration}\\
+\hline
+$\bJ$ & $2B \times P$ \\
+$\bmJ_1$ & $B \times (N+L)$\\
+$\bmJ_2$ & $B \times (N+L)$\\
+\hline
+\end{tabular}
+\label{tab:matrix_dimensions_main}
+\end{table}
+
 In a radio interferometer, the true sky visibilities $y_{pq}$ measured by a baseline formed by antenna $p$ and $q$ are always ``corrupted" by the non-ideal response of the receiver, which is often incorporated into a single, receiver-based complex number $g$ (i.e.
@@
 \end{equation}
-where $\conj{(*)}$ indicates conplex conjugation and $n_{pq}$ is the thermal noise component. 
+where $\conj{(*)}$ indicates complex conjugation and $n_{pq}$ is the thermal noise component. 
 The real and imaginary components of the thermal noise are normally distributed with a mean of zero and a
@@
 \end{equation}
-The vectors in equation~\ref{eq:vec_linear_definitions} are column vectors of size $B$ (i.e. $p<q$).
+The vectors in equation~\ref{eq:vec_linear_definitions} are column vectors of size $B$ (i.e. $p<q$). It is important to point out here that 
+all of the mathematical definitions in this paper assumes that we are only considering composite indices belonging to the set $\{rs|r<s\}$. Fore the sake of mathematical 
+completeness, however, $\alpha_{pq}$ is defined for composite indices which do not belong to this set.
 
@@
 \bJ = \begin{bmatrix}
-       \bmJ & \bmJ^*\\
-       \conj{\bmJ}^* & \conj{\bmJ} 
+       \bmJ_1 & \bmJ_2\\
+       \conj{\bmJ}_2 & \conj{\bmJ}_1 
       \end{bmatrix},
@@
 \label{eq:jac_entries}
-[\bmJ]_{\alpha_{pq},i} &= \frac{\partial v_{pq}}{\partial g_i}, & [\bmJ^*]_{\alpha_{pq},i} &= \frac{\partial v_{pq}}{\partial \conj{g}_i}. 
+[\bmJ_1]_{\alpha_{pq},i} &= \frac{\partial v_{pq}}{\partial g_i}, & [\bmJ_2]_{\alpha_{pq},i} &= \frac{\partial v_{pq}}{\partial \conj{g}_i}. 
 \end{align}
+The matrix $\bJ$ is generally referred to as the Jacobian\footnote{a matrix composed of first-order partial derivatives} matrix.
 
@@
 where $\bD=\bI\odot\bJ^H\bJ$.
-In this paper, the identity matrix is denoted by $\bI$.
-Moreover, the Hadamard product and the Hermition transpose are denoted by $\odot$ and $(\cdot)^H$ respectively.
+
+In this paper, we use single subscript indices (e.g.
+$i$) to refer to a specific antenna (or as will become apparent later a specific redundant group) and composite subscript indices (e.g.
+$pq$) to refer to
+a specific baseline. If these indices appear in the definition of matrix elements then their allowed ranges are determined by the dimension of the matrix that 
+is being defined (see Table~\ref{tab:matrix_dimensions_main}).
+Furthermore, the identity matrix is denoted by $\bI$.
+Moreover, the Hadamard\footnote{element-wise product} product and the Hermition transpose are denoted by $\odot$ and $(\cdot)^H$ respectively \citep{Liu2008}.
 Note the use of the Wirtinger derivatives in equation~\ref{eq:jac_entries}.
-We will refer to $\bJ^H\bJ$ as the Hessian matrix $\bH$ and to $\bJ^H\bJ + \lambda\bD$ as the modified Hessian matrix $\bmH$ throughout this paper.
+We will refer to $\bJ^H\bJ$ as the Hessian\footnote{a square matrix composed of second-order partial derivatives} matrix $\bH$ and to $\bJ^H\bJ + \lambda\bD$ as the modified Hessian matrix $\bmH$ throughout this paper \citep{Madsen:LM}.
 
@@
 if baseline $pq$ and $rs$ are redundant then $\phi_{pq} = \phi_{rs}$ (implying they belong to the same group).
-The function $\phi_{pq}$ is not symmetric as $\phi_{pq} = 0$ if $p \geq q$.
-Equation~\ref{eq:vis_definition} can be re-written for a redundant array as:
+To be exact, $\phi$ maps the composite index $pq$ to its group index only if $pq\in\{rs|r\leq s\}$. If $pq\notin\{rs|r\leq s\}$ then the 
+composite index $pq$ is mapped to zero. The function $\phi_{pq}$ is, therefore, not symmetric. Equation~\ref{eq:vis_definition} can be re-written for a redundant array as:
 \begin{equation}
@@
  \end{align}
-The number of model parameters to be solved for is now $P = 2(N+L)$, since redundant calibration is a complex problem.
+The number of model parameters to be solved for is now $P = 2(N+L)$, since redundant calibration is a complex problem. Note that equation~\ref{eq:least_squares_red} is only solvable 
+(i.e. the array is redundant enough) if $L+N \leq B$.
 
@@
 where $\breve{\bz} = [\bz^T,\conj{\bz}^T]^T$.
+
+\begin{figure}
+\includegraphics[width=0.47\textwidth]{./flow_diagram.pdf}
+\caption{A flow chart representing the procedure one would follow to partially construct the Jacobian matrix (i.e. $\bmJ$) associated with Wirtinger redundant calibration. \label{fig:flow_chart}}
+\end{figure}
+%\cite{Liu2008}\cite{Madsen:LM}
+
 We derive the complex Jacobian associated with equation~\ref{eq:least_squares_complex} to be:
@@
 \begin{equation}
-[\bmJ]_{\alpha_{pq},i} = \begin{cases} 
+\label{eq:j1}
+[\bmJ_1]_{\alpha_{pq},i} = \begin{cases} 
      \frac{\partial v_{pq}}{\partial g_i} & \textrm{if}~i\leq N \\
@@
 \begin{equation}
-[\bmJ^*]_{\alpha_{pq},i} = \begin{cases} 
+\label{eq:j2}
+[\bmJ_2]_{\alpha_{pq},i} = \begin{cases} 
      \frac{\partial v_{pq}}{\partial \conj{g}_i} & \textrm{if}~i\leq N \\
@@
 \end{equation}
+
+Note that we employ the same subscript indexing notation that we used in Section~\ref{sec:sky_wirtinger} in equation~\ref{eq:j1} and equation~\ref{eq:j2}.
+To further aid the reader in understanding this indexing notation we refer him/her to Figure~\ref{fig:flow_chart} which depicts a flow diagram of the matrix construction procedure 
+with which $\bmJ_1$ can be constructed. The range of values the indexes in equation~\ref{eq:j1} and equation~\ref{eq:j2} can attain should also 
+be clear to the reader after having inspected Figure~\ref{fig:flow_chart} and Table~\ref{tab:matrix_dimensions_main}.
+
 We can now calculate the GN and LM updates to be 
@@
 
-The simulated visibilities included a noise term, the purpose of which was to produce a certain SNR in the simulated
-visibilities. We made use of the following definition of \citep[SNR, ][]{Liu2010,Marthi2014}:  
-\begin{equation}
-\label{}
+Note that our simulations are almost ideal as we did not include a primary beam response, nor did we incorporate time and frequency smearing into our simulation.
+
+We also did not explicitly define our noise in terms of the integration time, channel bandwidth and $T_{\textrm{sys}}$, we instead
+follow the same approach described in \citep{Liu2010,Marthi2014} and make use of the definition of SNR to introduce noise into our visibilities. 
+We use the following definition of 
+\citep[SNR, ][]{Liu2010,Marthi2014}:  
+\begin{equation}
+\label{eq:SNR}
 \textrm{SNR} = 10\log \left (\frac{<\bv\odot\conj{\bv}>_{\nu,t,pq}}{<\bn\odot\conj{\bn}>_{\nu,t,pq}} \right ), 
 \end{equation}
-where $<*>_{\nu,t,pq}$ denotes averaging over frequency, time and baseline.
+where $<*>_{\nu,t,pq}$ denotes averaging over frequency, time and baseline. It should be pointed out, however, that by producing visibilities with different SNR values you are 
+effectively either changing your integration time or your channel bandwidth or both (assuming a single $T_{\textrm{sys}}$ value for your instrument). The reason we decided to also adopt
+equation~\ref{eq:SNR} is that it has become standard practice to present the accuracy of parameter estimation algorithms as a function of SNR.
 
@@
 \label{sec:pcg}
-\citet{Liu2010} suggested that the speed of convergence of redundant calibration could be optimized using the 
+\citet{Liu2010} suggested that the execution speed of redundant calibration could be reduced using the 
 conjugate gradient method \citep{Hestenes1952}, which would be computationally advantageous, since the Hessian matrix associated 
@@
 
+\section{Future Outlook}
+We plan to apply redundant \textsc{StEfCal} to HERA data in the near future.
+The reason for applying redundant \textsc{StEfCal} to real data is two-fold.
+First, we wish to validate the theoretical computational
+complexity of redundant \textsc{StEfCal} (derived earlier in the paper). Second, we would like to test whether it could be used to calibrate HERA in near-realtime. 
+We are also interested in paralyzing our current implementation and to see how much can be gained by doing so. We would also like to 
+conduct a perturbation analyses, similar to the one conducted by \cite{Liu2010}, to estimate the error which is introduced into our 
+estimated gains and visibilities when the array contains baselines which are not perfectly redundant. We are also interested in quantifying 
+the error which is introduced into the estimated gains and visibilities because of differing primary beam patterns between array elements \citep{Noorishad2012}.
+
 \section*{Acknowledgements}
@@
 \label{sec:red_stef_ADI}
-The basic skymodel-based \textsc{StEfCal} update step is equal to  the leftmost term in equation~\ref{eq:g_update} (barring $\alpha$) \citep{Salvini2014}.
+The basic skymodel-based \textsc{StEfCal} update step is equal to  the leftmost term in equation~\ref{eq:g_update} (barring $\rho$) \citep{Salvini2014}.
 Assume without any loss of generality that the array is in an east-west regular grid. Furthermore, assume that $\boldsymbol{d}$ (see equation~\ref{eq:vis_linear_definition}) has been re-ordered 
@@
 \begin{equation}
-\widetilde{\boldsymbol{d}} = \left[d{12},\cdots,d_{N-1,N},d_{13},\cdots,d_{N-2,N},\cdots,d_{1N}\right]^T .
+\widetilde{\boldsymbol{d}} = \left[d_{12},\cdots,d_{N-1,N},d_{13},\cdots,d_{N-2,N},\cdots,d_{1N}\right]^T .
 \end{equation}

### 2018-02-08 paper.tex
second round of corrections

@@
 \usepackage{graphicx}
-%\usepackage{algorithm}
-%\usepackage[noend]{algpseudocode}
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
 \usepackage{mathrsfs}
@@
 
-The paper is organized as follows: we review the complex calibration formalism by \citet{Smirnov2015} in Section~\ref{sec:sky_wirtinger}, we extend it to the redundant case in Section~\ref{sec:red_wirtinger}, we present some computational complexity results in Section~\ref{sec:pcg} and we presnt our conclusions in Section~\ref{sec:conclusions}.
+The paper is organized as follows: we review the complex calibration formalism by \citet{Smirnov2015} in Section~\ref{sec:sky_wirtinger}, we extend it to the redundant case in Section~\ref{sec:red_wirtinger}, we present some computational complexity results in Section~\ref{sec:pcg} and we present our conclusions in Section~\ref{sec:conclusions}.
 
@@
 $\psi_{pq}:\mathbb{N}^2\rightarrow\mathbb{N}$ &\\
-$\conj{(*)}$ & conjugation of operand\\
-$(*)^{2}$ & square operand\\
-$\|*\|_F$ & Frobenius norm\\
-$(*)^{-1}$ & Matrix inversion\\
+$\conj{x}$ & conjugation\\
+$\conj{\mathbf{x}},\conj{\mathbf{X}}$ & element-wise conjugation\\
+$x^{2}$ & square\\
+$\|\mathbf{x}\|_F$ & Frobenius norm\\
+$\mathbf{X}^{-1}$ & Matrix inversion\\
 $\odot$ & Hadamard product\\
 &(element-wise product)\\
-$(*)^H$ & Hermitian transpose\\
-$(*)^T$ & Matrix transpose\\
-$<*>_x$ & averaging in x\\
+$\mathbf{X}^H$ & Hermitian transpose\\
+$\mathbf{x}^T,\mathbf{X}^T$ & Transpose\\
+$<\mathbf{x}>_x$ & averaging over $x$\\
 $\frac{\partial}{\partial z}$ & Wirtinger derivative\\
@@
 \end{equation}
-where $\conj{(*)}$ indicates complex conjugation and $n_{pq}$ is the thermal noise component. 
+where $\conj{\conj{x}}$ indicates complex conjugation and $n_{pq}$ is the thermal noise component. 
 The real and imaginary components of the thermal noise are normally distributed with a mean of zero and a
@@
 The vectors in equation~\ref{eq:vec_linear_definitions} are column vectors of size $B$ (i.e. $p<q$). It is important to point out here that 
-all of the mathematical definitions in this paper assumes that we are only considering composite indices belonging to the set $\{rs|r<s\}$. Fore the sake of mathematical 
-completeness, however, $\alpha_{pq}$ is defined for composite indices which do not belong to this set.
+all of the mathematical definitions in this paper assumes that we are only considering composite indices belonging to the set $\{rs|r<s\}$. For the sake of mathematical 
+completeness, however, $\alpha_{pq}$ is defined for composite indices which do not belong to this set. Also note that, while the indexing function $\alpha_{pq}$ does not attain 
+the value zero in practice the zero indexing value does play a very important role in some of the definitions used within this paper (see for example Eq.~\ref{eq:D_mat}).
 
@@
 \end{equation}
-where $\Lambda$ is the objective function, $\br$ is the residual vector and $\|\cdot\|_F$ denotes the Frobenius norm.
+where $\Lambda$ is the objective function, $\br$ is the residual vector and $\|\mathbf{x}\|_F$ denotes the Frobenius norm.
 In standard interferometric calibration, $y_{pq}$ is assumed to be known at some level, for instance through the observation of previously known calibration sources.
@@
 Furthermore, the identity matrix is denoted by $\bI$.
-Moreover, the Hadamard\footnote{element-wise product} product and the Hermition transpose are denoted by $\odot$ and $(\cdot)^H$ respectively \citep{Liu2008}.
+Moreover, the Hadamard\footnote{element-wise product} product and the Hermition transpose are denoted by $\odot$ and $\mathbf{X}^H$ respectively \citep{Liu2008}.
 Note the use of the Wirtinger derivatives in equation~\ref{eq:jac_entries}.
@@
 \includegraphics[width=0.47\textwidth]{./flow_diagram.pdf}
-\caption{A flow chart representing the procedure one would follow to partially construct the Jacobian matrix (i.e. $\bmJ$) associated with Wirtinger redundant calibration. \label{fig:flow_chart}}
+\caption{A flow chart representing the procedure one would follow to partially construct the Jacobian matrix (i.e. $\bmJ_1$) associated with Wirtinger redundant 
+calibration. The red circle represents the start of the flow diagram. The blue circle represents the end of the diagram. Blue diamonds denote loop 
+conditionals, while green diamonds denote simple conditionals. The diagram elements following the red arrows just below a loop conditional element all form part 
+of the main body of the loop which has the loop conditional of the aforementioned loop conditional element in its definition.
+The pseudocode associated with this flow diagram is given in Algorithm 1.
+\label{fig:flow_chart}}
 \end{figure}
-%\cite{Liu2008}\cite{Madsen:LM}
 
@@
 
+\begin{algorithm}
+\caption{Constructing $\bmJ_1$}\label{euclid}
+\begin{algorithmic}[1]
+\State $p \gets 1$
+\While {$p\leq N$}
+\State $q \gets p + 1$
+\While {$q\leq N$}
+\State $\alpha  \gets (q-p) + (p-1)\left (N-\frac{1}{2} \right )$
+\State $i \gets 1$
+\While {$i\leq (N + L)$}
+\If {$i \leq N$}
+\State $\bmJ_1[\alpha,i] \gets \frac{\partial v_{pq}}{\partial g_i}$
+\Else
+\State $\bmJ_1[\alpha,i] \gets \frac{\partial v_{pq}}{\partial y_{i-N}}$
+\EndIf
+\State $i\gets i + 1$
+\EndWhile
+\State $q \gets q + 1$
+\EndWhile
+\State $p \gets p + 1$
+\EndWhile
+\end{algorithmic}
+\end{algorithm}
+
 Note that we employ the same subscript indexing notation that we used in Section~\ref{sec:sky_wirtinger} in equation~\ref{eq:j1} and equation~\ref{eq:j2}.
-To further aid the reader in understanding this indexing notation we refer him/her to Figure~\ref{fig:flow_chart} which depicts a flow diagram of the matrix construction procedure 
+To further aid the reader in understanding this indexing notation we refer him/her to Figure~\ref{fig:flow_chart} (also see Algorithm 1) which depicts a flow diagram of the matrix construction procedure
 with which $\bmJ_1$ can be constructed. The range of values the indexes in equation~\ref{eq:j1} and equation~\ref{eq:j2} can attain should also 
@@
 \end{equation}
-where $<*>_{\nu,t,pq}$ denotes averaging over frequency, time and baseline. It should be pointed out, however, that by producing visibilities with different SNR values you are 
+where $<\mathbf{x}>_{\nu,t,pq}$ denotes averaging over frequency, time and baseline. It should be pointed out, however, that by producing visibilities with different SNR values you are 
 effectively either changing your integration time or your channel bandwidth or both (assuming a single $T_{\textrm{sys}}$ value for your instrument). The reason we decided to also adopt
@@
 \url{https://github.com/Trienko/heracommissioning/blob/master/code/stef.py}. 
-We note that, in its current implementation, redundant \textsc{StEfCal} does not solve for the degeneracies inherent to redundant calibration  \citep{Zheng2014,Kurien2016} which will be the subject of future work.
+We note that, in its current implementation, redundant \textsc{StEfCal} does not solve for the degeneracies inherent to redundant calibration  \citep{Zheng2014,Kurien2016,dillon2017} which will be the subject of future work.
 Compared to current redundant calibration algorithms, redundant \textsc{StEfCal} is more robust to inital conditions and allows for easier parallelization.
@@
 complexity of redundant \textsc{StEfCal} (derived earlier in the paper). Second, we would like to test whether it could be used to calibrate HERA in near-realtime. 
-We are also interested in paralyzing our current implementation and to see how much can be gained by doing so. We would also like to 
+We are also interested in parallelizing our current implementation and to see how much can be gained by doing so. We would also like to 
 conduct a perturbation analyses, similar to the one conducted by \cite{Liu2010}, to estimate the error which is introduced into our 
@@
 % \end{equation}
-% In this paper, we denote the Hermitian transpose with $(*)^H$ and the Frobenius norm with $\|*\|_F$.
+% In this paper, we denote the Hermitian transpose with $(\cdot)^H$ and the Frobenius norm with $\|*\|_F$.
 % 
@@
 \begin{equation}
+\label{eq:D_mat}
 [\bD]_{ij} = 
